### 2014-08-28 D1-3.tex
Renamed file. Inserted rough TOC.

\documentclass[external]{20120615_deliverable_template_ukob}

\usepackage{xspace}
\usepackage{amsmath,amsthm,amssymb,url,color}
\theoremstyle{definition}
\newtheorem{example}{Example}
\usepackage{xcolor}
\usepackage{color, colortbl}
%\usepackage{showkeys}

\LGtitle{\LiveGovThirtyTitleFonts \textbf{D1.3 - Privacy Aware Sensor Data Storage and Miner} }
%\LGnumber{WP.NR} % e.g 6.7 no letters D or ID

\LGnumber{1.2}

\LGwp{WP1 - Reality Sensing and Mining}

\LGdissemination{PU - Public}

\LGcontractdate{Month 30, October 2014}

\LGactualdate{Month 30, October 2014}

\LGtask{T1.3, T1.4}

\LGtype{Prototype}

%\LGnature{$<$ Report $|$ Prototype $|$ Demonstrator $|$ Other $>$}

%\LGapproved % activate only if approved
\LGfinal

\LGversion{12}
%\LGstaffmonths{Resources spent on the deliverable}

%\LGdistribution{WP leaders, PMB members, European Commission}

%\LGkeywords{}

%\LGabstract{Put abstract here (separate paragraphs using
%{\tt\char92 par}.)}
\LGabstract{

% \TODO{Add abstract}

%  This deliverable describes the mobile sensor mining component.
%  Along with this document we supply the source code of the
%  components, documentation (javadoc), and pre-compiled packages for
%  direct installation and testing on mobile devices.

}

%\LGhistory{01}{2011-10-01}{First draft}{Yiannis Kompatsiaris}
%\LGhistory{02}{2011-10-02}{Modifications}{Sotiris Diplaris}
%\LGhistory{03}{2011-10-03}{Further Modifications \& Corrections}{Sotiris Diplaris}
%\LGhistory{04}{2011-10-05}{Further Modifications \& Corrections}{Sotiris Diplaris}

%------------------------ macro for change portrait to landscape
\usepackage{calc,graphicx,pdflscape,eso-pic} % needed for printing
                                             % headers and footers
\newlength\landscapewidth
\newlength\landscapeheight
\newcommand\landscapepagestyle{ % command which prepare page
                                % for landscape printing,
                                % change to landscape is
                                % achieved by pdflscape
\clearpage \thispagestyle{empty}
\setlength\landscapewidth{247mm}
\setlength\landscapeheight{161mm}
\AddToShipoutPicture*{\AtPageCenter{ % this make "layer" for
                                     % printing the header and footer
\rotatebox[origin=c]{90}{
\hspace*{-2em} % for adjusting position of
               % header and footer
\parbox{\landscapewidth}{\vskip-.57\landscapeheight
%\centerline{\WEKNOWIT \hfill \textbf{\Large \bf{IDx.x - V0} \normalsize }} %header text for Internal Deliverables
                                                                         % notation: IDx.x V{LGversion}
\centerline{\LiveGovLogo \hfill \textbf{\Large \bf{Dx.x - V0} \normalsize }} %header text for External Deliverables
                                                                        % notation: Dx.x V{LGversion}
\vspace{-5pt} \rule{\landscapewidth}{0.4pt} \\
\rule{0pt}{\landscapeheight} \\
%\rule{\landscapewidth}{0.4pt} \\
\centerline{Page \thepage\ }  % footer text
} } } } }

%--------------------------------------------------------------------------------------------------------------%
\usepackage{appendix}
\pretolerance=10000 % prevent overflow lines for $math$ elements

\begin{document}

% add a \LGaddhistory{version}{date}{reason}{revised by} for each new
% version
\begin{LGhistory}
\LGaddhistory{01}{2014-07-17}{Outline}{Heinrich Hartmann}

\end{LGhistory}

\newcommand{\LGaddauthorNoPhone}[3]{\hline  #1 &  #2 & %
   \parbox{3em}{E-mail:} \small #3 \\
}

% add a
% \LGaddauthor{Partner}{Name}{Telephone}{Fax}{Email} for each author
\begin{LGauthors}

\LGaddauthor{UKob}{Heinrich Hartmann}{+49 261 287 2759}%
{+49 261 287 100 2759}{\small hartmann@uni-koblenz.de}

\LGaddauthor{UKob}{Christoph Schaefer}{+49 261 287 2786 }%
{+49 261 287 100 2786}{\small chrisschaefer@uni-koblenz.de}

%% \LGaddauthor{UKob}{Christoph Kling}{+49 261 287 2702 }%
%% {+49 261 287 100 2702}{\small ckling@uni-koblenz.de}

%% \LGaddauthorNoPhone{MTS}{Laura Niittyl\"a}%
%% {\small Laura.Niittyla@mattersoft.fi}

\end{LGauthors}

\begin{LGExecutiveSummary}
  \vspace{10pt}

  The deliverable is accompanied with source code of the components in
  Java, API documentation (javadoc), and pre-compiled packages for
  direct installation and testing on mobile devices.

\end{LGExecutiveSummary}

% add a \LGaddabbreviation{ABBR}{Explanation} for each abbreviation
\begin{LGAbbreviations}
%\LGaddabbreviation{LG}{Live+Gov}

\LGaddabbreviation{\textbf{AIDL}}
{Android Interface Description Language}

\LGaddabbreviation{\textbf{AJAX}}
{Asynchronous JavaScript and XML}

\LGaddabbreviation{\textbf{ALC}}
{Attribute Language with Complement}

\LGaddabbreviation{\textbf{API}}
{Application Programming Interface}

\LGaddabbreviation{\textbf{GeSA}}
{Geographical Semantic Analysis}

\LGaddabbreviation{\textbf{GPS}}
{Global Positioning System}

\LGaddabbreviation{\textbf{GSM}}
{Global System for Mobile Communications}

\LGaddabbreviation{\textbf{HTML}}
{HyperText Markup Language}

\LGaddabbreviation{\textbf{HTTP}}
{Hypertext Transfer Protocol}

\LGaddabbreviation{\textbf{ID}}
{Identifier}

\LGaddabbreviation{\textbf{IEEE 802.11}}
{see WIFI, WLAN}

\LGaddabbreviation{\textbf{JSON}}
{JavaScript Object Notation}

\LGaddabbreviation{\textbf{LAN}}
{Local Area Network}

\LGaddabbreviation{\textbf{REST}}
{Representational State Transfer}

\LGaddabbreviation{\textbf{RF-ID}}
{Radio-Frequency Identification}

\LGaddabbreviation{\textbf{SDCF}}
{Sensor Data Collection Framework}

\LGaddabbreviation{\textbf{SQL}}
{Structured Query Language}

\LGaddabbreviation{\textbf{SVM}}
{Support Vector Machine}

\LGaddabbreviation{\textbf{TBox}}
{Terminological Box}

\LGaddabbreviation{\textbf{UI}}
{User Interface}

\LGaddabbreviation{\textbf{URL}}
{Uniform Resource Locator}

\LGaddabbreviation{\textbf{UUID}}
{Universal Unique Device Identifier}

\LGaddabbreviation{\textbf{WP}}
{Work Package}

\LGaddabbreviation{\textbf{WIFI}}
{Wireless Fidelity (IEEE 802.11), WLAN}

\LGaddabbreviation{\textbf{WLAN}}
{Wireless Local Area Network}

\LGaddabbreviation{\textbf{XML}}
{Extensible Markup Language}

\LGaddabbreviation{~\\}
{~~~}

\end{LGAbbreviations}

% add a \LGaddterm{Term}{Explanation} for each entry of the glossary
%\begin{LGGlossary}
%\LGaddterm{Term}{Definition text here}
%\LGaddterm{Term2}{Definition text here}
%\end{LGGlossary}

\setcounter{tocdepth}{1}

% add this for a table of contents
\LGTOC

\newpage
\newcommand{\todo}[2]{\textcolor{magenta}{#1: #2}}
%\newcommand{\todo}[2]{}

\chapter{Introduction}
\label{chap:Introduction}

\clearpage

\chapter{Privacy Protection}
Intro
\section{Historical Definitions}
Aristotle
Warren and Brandeis
Fried
\section{Privacy Definition and Taxonomy}
\section{Privacy and Law}
European Convention on Human Rights - Article 8
EU Data Protection Direvtive
\section{IT Security Analysis}
Live+Gov Privacy Protection Analysis
Step 1. World Analysis
Assets: Privacy
IT-Systems and Humans
Conflict of Interests
Vulnerabilities
Step 2. Potential Analysis
Step 3. Plan Development
\section{Implementation}
* Sensor Data Privacy Control
* Rising Awareness

\chapter{Improved Sensor Data Mining Methods}
\section{Battery Awareness of Sensor Collector}
Results from Projektprakikum (Wifi/Zip)

\section{Service Line Detection (new method)}
MA thesis results from Sven Milker

\section{Issue Analysis}
Christoph Schäfer with Niko Beck

\bibliography{D1-3}

\end{document}

### 2014-09-11 D1-3.tex
Added Privacy Protection tex from Max.
Signed-off-by: Heinrich Hartmann <hartmann@uni-koblenz.de>

@@
 
+\input{PrivacyProtection-2014-09-09}
+
 \clearpage
 
-\chapter{Privacy Protection}
-Intro
-\section{Historical Definitions}
-Aristotle
-Warren and Brandeis
-Fried
-\section{Privacy Definition and Taxonomy}
-\section{Privacy and Law}
-European Convention on Human Rights - Article 8
-EU Data Protection Direvtive
-\section{IT Security Analysis}
-Live+Gov Privacy Protection Analysis
-Step 1. World Analysis
-Assets: Privacy
-IT-Systems and Humans
-Conflict of Interests
-Vulnerabilities
-Step 2. Potential Analysis
-Step 3. Plan Development
 \section{Implementation}
-* Sensor Data Privacy Control
-* Rising Awareness
+\subsection{Authentification and Access Control}
+\subsection{Anonymization and Pseudonoymization}
+\subsection{Randomization}
 

### 2014-09-11 PrivacyProtection-2014-09-09.tex
Added Privacy Protection tex from Max.
Signed-off-by: Heinrich Hartmann <hartmann@uni-koblenz.de>

\section{Privacy}

The goal of this document to analyze and identify the threads to personal privacy that are posed by collecting, storing and processing sensor data from mobile phones.
We derive concrete privacy protection measures that address the main risks involved with handling such data.
In the last part we describe how these measures are implemented in the Live+Gov toolkit.

As privacy is a very general and hard to grasp term, we need to fix a definition of privacy that is suitable for our needs.
As background information we include an overview about historical treatments of privacy as well as legal regulation of privacy in the European Union.
Based on this we propose a definition of privacy as {\em control over personal data}, and introduce a taxonomy of privacy attributes that give specify the term {\em personal data} in the context of mobile
sensor data collection.

The complexity of our systems and the variety of threads make a great number of counter measures plausible.
We approach this complexity with the aid of a general security analysis model developed in [Grimm].
We give a brief introduction to this model and perform a IT Security Analysis with respect to the privacy asset for our system.

\subsection{Histrocial Approaches to Privacy}

\subsubsection{Aristotle}

Public Sphere vs. Private Sphere of politics. This distinction is concerned with the politic (public) life of politicians in early democracies opposed by their domestic (private) life.

\subsubsection{Warren and Brandeis}

The Right to Privacy, 1890: Establish the term Informational Privacy. They argue innovations like photography and a growing press create the need for a new right to privacy, "the right to be left alone".

The term informational privacy as conceptual foundation of privacy originates from the juristic discussion The Right to Privacy (1890) by Warren and Brandeis.
The arise of new technologies like photography combined with a growing press created an increased publicity.
Warren and Brandeis were worried that existing law like castle doctrines or libel and slander could not suffice.
They felt a need for a new right to "protect the extent to wich one's thoughts, sentiments, and emotions could be shared with others" [1].
This right would be "the right to be left alone".

The right to "protect the extent to wich one's thoughts, sentiments, and emotions could be shared with others" by Warren and Brandeis was later extend to cover arbitrary information and is now kown as the privacy concept "... the control we have over information about ourselves ..." [2].

\subsubsection{Fried}

C. Freid, An Anatomy of Values, Harvard Univ. Press, 1970

Defines privacy as the ability to control information about oneself. He argues that limited knowledge does not automatically create privacy.

\subsection{Legal Aspects of Privacy}

* Present legal view on privacy from European angle: ``Data Protection Directive`` and Implementation.

* Some US Law

\subsubsection{European Convention on Human Rights - Article 8}

``(1) Everyone has the right to respect for his private and family life, his home and his correspondence. (2) There shall be no interference by a public authority with the exercise of this right except such as is in accordance with the law and is necessary in a democratic society in the interests of national security, public safety or the economic well-being of the country, for the prevention of disorder or crime, for the protection of health or morals, or for the protection of the rights and freedoms of others.'' {[}\hyperref[references]{3}{]}

``The European Court of Human has given this article a very broad
interpretation in its juresprudence.'' {[}\hyperref[references]{2}{]}

\subsubsection{EU Data Protection Direvtive}

\emph{Directive 95/46/EC}

EU directives in general do not hold any direct legal binding for EU
citzens. Member states have to create their own legal implementation.

``The directive regulates the processing of personal data regardless of
whether such processing is automated or not.''
{[}\hyperref[references]{1}{]} where:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{personal data} is ``any information relating to an identified
  or identifiable natural person {[}\ldots{}{]} (art. 2a)''
  {[}\hyperref[references]{1}{]}
\item
  \textbf{processing} is ``any operation or set of operations which is
  performed upon personal data {[}\ldots{}{]} (art. 2b)''
  {[}\hyperref[references]{1}{]}
\end{itemize}

The ``identified or identifiable natural person'' is called \textbf{data
subject}.

The sole responsibility for compliance is held by so called
\textbf{controllers}. A controller is an actor ``which alone or jointly
with others dertimines the purposes and means of the processing of
personal data; (art 2d)'' {[}\hyperref[references]{1}{]}

(\textbf{Note:} this rule is applicable ``whenever the controller uses
equipment situated within the EU''{[}\hyperref[references]{1}{]}. This
will hold for every e-commerce provider \textbf{inside or outside} of
the EU because the customer's computer is \emph{situated within the EU}
anyways.)

``Personal data should not be processed at all, except when certain
conditions are met'' {[}\hyperref[references]{1}{]}. These conditions
incorporate the seven OECD princepals and are futher categorized in the
categories:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Transparency}: ``The data subject has the right to be informed
  when his personal data is being processed {[}\ldots{}{]} (art. 10 and
  11)'' {[}\hyperref[references]{1}{]} and he has the right how and by
  whom it is processed. Additionally other speficied requirements have
  to be met, i.e.~giving consent.
\item
  \textbf{Legitimate prupose}: ``Personal data can only be processed for
  specified explicit and legitimate pruposes and may not be processes
  futher in a way incompatible with those pruposes (art. 6b)''
  {[}\hyperref[references]{1}{]}.
\item
  \textbf{Proportionality}: ``Personal data may be processed only
  insofar as it is adequate {[}\ldots{}{]} in relation to {[}its{]}
  pruposes {[}\ldots{}{]}; every reasonable step must be taken to ensure
  that data wich are inaccurate or incomplete {[}\ldots{}{]} are erased
  or rectified (art. 6)'' {[}\hyperref[references]{1}{]}
\end{itemize}

(\textbf{Note:} The principles incorporated by the EU directive seem to
closely map to \emph{the seven principles for user privacy control}.)

Member states have to create independent \textbf{supervisory
authorities} for personal data processing where controllers have to
register before they start to process any data. The record has to be
stored in public register {[}\hyperref[references]{1}{]}.

Data transfer to thrid countries outside of the EU requires those
countries to have a similar data protection level. Exceptions exists:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Save Harbor
\item
  Passenger Name Record Aggreement
\end{itemize}

all between the US and the EU, although the US has no comparable law of
data protection. The US Supreme Court regards privacy as
\textbf{implicit} right granted with the First Amendment. However,
european law states privacy as \textbf{explicit} right (ECHR Art. 8).
Additionally the US government follows a doctrine which favors the
economy to implement self-regulation {[}\hyperref[references]{1}{]}.

\subsubsection{Implementation of the Data Protection Direvtive}

\paragraph{Germany}

Germany implements Directive 95/46 with the \emph{Federal Data
Protection Act} from 2001 known as \emph{Bundesdatenschutzgesetz (BDSG)}
{[}\hyperref[references]{1}{]}{[}\hyperref[references]{2}{]}{[}\hyperref[references]{6}{]}.
However, Germany has violated the directive in two points:

\begin{enumerate}
\itemsep1pt\parskip0pt\parsep0pt
\item
  The BDSG has become effective three years too late, thus the EC filed
  a treaty violation proceeding against Germany
  {[}\hyperref[references]{6}{]}.
\item
  The BDSG does not implement \textbf{independet} supervisory
  authorities. The \emph{Bundesdatenschutzbeauftragter} is subordinate
  to the Ministiry of Interior. Although he is not subject to technical
  oversight (\emph{Fachaufsicht}), he is subject to staff supervision by
  the government (\emph{Rechtsaufsicht} durch die Bundesregierung und
  \emph{Dienstaufsicht} durch das Innenministerium) and budget oversight
  (including approval of employees) by the ministry
  {[}\hyperref[references]{7}{]}. Thus the EC filed a treaty violation
  proceeding against Germany, again {[}\hyperref[references]{6}{]}. In
  March 2010 Germany was found guilty of violation of Directive 95/46 by
  the ECJ {[}\hyperref[references]{6}{]}.
\end{enumerate}

(\textbf{Note:} Germany still violates point 2!)

The states of Germany have their own implementation of Directive 95/46
(\emph{Landesdatenschutzgesetze}). Federal public authorites are only
bound to their federal law. {[}\hyperref[references]{2}{]}

Churches are not subject the BDSG. {[}\hyperref[references]{2}{]}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Roman Catholic Church:}
  \href{http://de.wikipedia.org/wiki/Anordnung_\%C3\%BCber_den_kirchlichen_Datenschutz}{Anordnung
  über den kirchlichen Datenschutz (KDO)}
\item
  \textbf{German Protestant Churches (Synode der Evangelischen Kirche in
  Deutschland):}
  \href{http://de.wikipedia.org/wiki/Datenschutzgesetz_der_Evangelischen_Kirche_in_Deutschland}{Kirchengesetz
  über den Datenschutz der Evangelischen Kirche in Deutschland}
\end{itemize}

Data subjects have the right to\ldots{}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \ldots{} \textbf{disclosure} of
\item
  wether data about them is stored and processed
\item
  which data is stored and processed
\item
  the data sources
\item
  \ldots{} \textbf{correction} of false data
\item
  \ldots{} \textbf{file complaints} to the supervisory authority
\item
  \ldots{} have data \textbf{deleted} or \textbf{blocked}, but
  controllers can prohibit deletion in favour of blocking
\item
  \ldots{} \textbf{Decline} third party access to the data
\end{itemize}

\paragraph{United Kingdom}

The UK implements Directive 95/46 with the \emph{Data Protection Act
1998 (DPA)}
{[}\hyperref[references]{1}{]}{[}\hyperref[references]{3}{]}.

The act is knwon for its high complexity: a manual record of phone
numbers for business purposes could be hold subject to the DPA
{[}\hyperref[references]{3}{]}. Although the act seems to fully cover
the directive. Even higher restriction apply for \emph{``sensitive
personal data''} (race, ethnicity, politics, religion, trade union
status, health, sex life or criminal record), i.e. \textbf{consent} must
be given freely and has to be explicit. {[}\hyperref[references]{3}{]}

\emph{``The Act's definition of''personal data" covers any data that can
be used to identify a living individual. Anonymised or aggregated data
is not regulated by the Act, providing the anonymisation or aggregation
has not been done in a reversible way. Individuals can be identified by
various means including their name and address, telephone number or
Email address. The Act applies only to data which is held, or intended
to be held, on computers (`equipment operating automatically in response
to instructions given for that purpose'), or held in a `relevant filing
system'.}

\emph{In some cases even a paper address book can be classified as a
`relevant filing system', for example diaries used to support commercial
activities such as a salesperson's diary.}

\emph{The Freedom of Information Act 2000 modified the act for public
bodies and authorities, and the Durant case modified the interpretation
of the act by providing case law and precedent.}

\emph{The Data Protection Act creates rights for those who have their
data stored, and responsibilities for those who store, process or
transmit such data. The person who has their data processed has the
right to:}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \emph{View the data an organisation holds on them. A `subject access
  request' can be obtained for a nominal fee. As of January 2014, the
  maximum fee is £2 for requests to credit reference agencies, £50 for
  health and educational request, and £10 per individual otherwise,}
\item
  \emph{Request that incorrect information be corrected. If the company
  ignores the request, a court can order the data to be corrected or
  destroyed, and in some cases compensation can be awarded.}
\item
  \emph{Require that data is not used in any way that may potentially
  cause damage or distress.}
\item
  *Require that their data is not used for direct marketing.``*
  {[}\hyperref[references]{3}{]}
\end{itemize}

\paragraph{France}

France implements Directive 95/46 with Law 2004-801 modifying law 78-17
of 6.1.1978 (\emph{Loi n° 2004-801 du 6 août 2004} modifying \emph{loi
n°78-17 relative à l'informatique, aux fichiers et aux libertés})
{[}\hyperref[references]{1}{]}{[}\hyperref[references]{10}{]}{[}\hyperref[references]{11}{]}.

\paragraph{United States of America}

The USA do not implement the directive, nor is there any obligation for
them to do so. However, companies subject to US jurisdiction can be
certified to comply with the seven principles enforced by Directive
95/46 (the seven OECD recommendations) {[}\hyperref[references]{5}{]}.
Thus, those companies will act as \emph{safe harbors}. Without
certification foreign companies are not allowed to send customer data
back home {[}\hyperref[references]{9}{]}.

(\textbf{Note:} \emph{Safe harbor} in general is the legal concept to
regulate that a certain conduct will be deemed
{[}\hyperref[references]{4}{]}, but in germany the term is commonly used
as synonym for the agreement between the EU and the US regarding
Directive 95/46.)

\subsection{Privacy Definition and Taxonomy}

\subsubsection{Defining Privacy}

Defining privacy is a challenge which seems impossible. This is well put
to words by Serge Gutwirth, who notes: \emph{``The notion of privacy
remains out of the grasp of every academic chasing it. Even when it is
cornered by such additional modifiers as `our' privacy, it still finds a
way to remain elusive.''} {[}\hyperref[references]{2}{]}.

Many privacy researchers seem to ``forcus on the ways in which privacy
can be infrenged'' {[}\hyperref[references]{1}{]}. Thus they try to
create taxonomies of \emph{privacy harms} instead of taxonomies of
\emph{privacy types}. Those two differ in the respect that the former
focuses on threats to prohibit whereas the latter focuses on values to
protect. So one should rather evaluate what aspects are precious about
privacy and develop measures to ensure their security than only forbid
single actions against privacy {[}\hyperref[references]{1}{]}.

\begin{center}
We follow Fried in defining privacy as the ability to control data about personal information.
\end{center}

* What is the relation to legal aspects?

* What should be personal information?

* Use Seven Privacy Types be Friedewald, Finn and Wrigtht.

\paragraph{The Seven Types of Privacy}

Seven Types of Privacy based on four types of privacy by Clarke.
Clarke's four types are outdated by contemporary technologies and no longer adequate. 
In order to fix this Friedewald, Finn and Wright extend the former four to the now introduced seven types privacy as follows:

\begin{enumerate}
\item \textbf{Privacy of the Person.}

This type is the right keep body functions and body characteristics
private {[}\hyperref[references]{1}{]}. It maps the one by Clarke.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Body Characteristics
\item
  Weight
\item
  Height
\item
  Other body measures (shoulder width)
\item
  Fingerprints
\item
  DNA Sequence
\item
  Medical Conitions
\item
  Orthopedic conditions (e.g.~limping)
\item
  Having a cold
\item
  \ldots{}
\end{itemize}
\end{enumerate}

\subparagraph{Privacy of Behaviour and Action}

This type is also concerned with the ``protection against disclosure of
personal matters'' {[}\hyperref[references]{1}{]} through behaviour,
however Clarke's distinction between ``casual observation {[}\ldots{}{]}
systematic recording and storage of information about those activities''
{[}\hyperref[references]{1}{]}is lifted.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  regular visit at church, backery, doctors
\item
  sexual habits
\item
  political activities
\end{itemize}

\subparagraph{Privacy of Communication}

It ``aims to avoid the interception of communications''
{[}\hyperref[references]{1}{]} either electronic or face-to-face.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Briefgeheimnis (in german Legislation)
\item
  Email contents
\item
  Personal direct communication
\item
  Right to free discussion, i.e.~without third parties listening
\end{itemize}

\subparagraph{Privacy of Data and Image}

This type is concerned with ``making sure that individuals's data is not
automatically available to other individuals and organisations''
{[}\hyperref[references]{1}{]}. This is Informational Privacy in an
intuitive sense.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Telefonnummer
\item
  IP Addresse
\item
  Public-administrative Data (Date of Birth, Melderegister)
\item
  Data held by organizations, like Banks or Insurance Companies
\item
  All data that is stored in online services (facebook)
\end{itemize}

\subparagraph{Privacy of Thoughts and Feelings}

This type is the right ``not to share their thoughts or feelings or to
have those thoughts or feeling (sic!) revealed''
{[}\hyperref[references]{1}{]}. It is concerned with (automatic) emotion
detection. This type is the counterpart the \emph{Privacy of the Person}
like body and mind are counterparts of one another (dualism).

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Current feeling: Depression, tiredness, stressed, awake
\item
  Thoughts in general
\end{itemize}

\subparagraph{6 Privacy of Location and Space}

This type is the right ``to move about in public or semi-public space
withoug being identified, tracked or monitored''
{[}\hyperref[references]{1}{]}. Additionally this type is concerned with
the protection of one's home and private places (``right to solitude''
{[}\hyperref[references]{1}{]}).

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  GPS position tracking
\item
  Location of home address
\end{itemize}

\subparagraph{Privacy of Association}

This type is the right ``to associate with whomever {[}one{]} wish,
withoug being monitored'' {[}\hyperref[references]{1}{]}. It is
concerned with the protection against the automatic record of one's
contacts. It does not imply that one is monitored because of the
associations.

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Friends
\item
  Joining of orgnaizations (e.g.~political parties)
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{IT Security}

\subsection{Basic Terminology}

%\includegraphics{diagrams/png/itsecterminology.png}
\input{./figures/Basic-IT-Security-Terminology}

\subsubsection{threat}

\emph{threat} is a class of potential events whose manifestation can
cause damage or harm.

\subsubsection{vulnerability}

A weakness which can cause the loss of security.

\subsubsection{hazard}

Concrete risk due to one or more concrete vulnerabilities \textbf{AND}
corresponding threats.

\texttt{threat + vulnerability -\textgreater{} hazard}

\subsubsection{incident}

An event caused by a threat manifestation against one or more
corresponding vulnerabilities.

\texttt{threat manifestation + vulnerability -\textgreater{} hazard}

\subsubsection{accident}

An incident which did cause damage or harm. (Translates to
\emph{Schandensfall/Zwischenfall/Störfall} in german)

\subsubsection{non-accident}

An incident which \textbf{did not} cause any damage or harm.

\subsubsection{measure}

A well defined set of one or more activities which reduce or eliminate
vulnerabilities.

\subsection{IT Security Analysis according to Grimm et. al}

Grimm et al. create a reference model for conducting IT security
analyses consisting of:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  an \textbf{ontology}, which aims to organize common security
  terminology in a reasonable and practical way
\item
  and a systematic analysis \textbf{procedure} based on that ontology
\end{itemize}

\subsubsection{Ontology}

%\includegraphics{../diagrams/png/itsec-ref-model-grimm.png}
\input{./figures/The-IT-Security-Reference-Model-Grimm}

\paragraph{Views}

The ontology is organized in four views:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{World:} Contains all components describing the current state.
\item
  \textbf{Potential:} Contains all components describing both the
  desired and dreaded state.
\item
  \textbf{Plan:} Contains all conceptual components required to develop
  sufficient measures to achieve and secure the desired state
  \textbf{or} to create the dreaded state. The \textbf{Plan} view
  \emph{realizes} the \textbf{Potential} view of the system.
  (\textbf{Note:} this should rather be translated with \emph{Concept})
\item
  \textbf{Event:} Contains all actual operations and events through out
  the production phase. The \textbf{Event} view \emph{instantiates} the
  \textbf{Plan} view of the system.
\end{itemize}

\paragraph{Components}

\subparagraph{World}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Assets:} Things of value to one or more stakeholders. The
  value can be \emph{``hard''} (money, data, etc.) or \emph{``soft''}
  (trust,privacy,etc.).
\item
  \textbf{IT-Systems:} The IT-Systems under study.
\item
  \textbf{Humans:} All identifiable actors of the system under study.
  Many problems arise due to misunderstandings during man-machine
  interaction.
\item
  \textbf{Conflicts of Interests:} Different actors have different
  interests. Those interest can be in conflict. A trival conflict is the
  \emph{Attacker-Attackee}-Conflict: a service provider offers private
  data storage, therefore the provider is interested in having the
  access restricted. An attacker is naturally interested in easy access
  in order captialize the stolen data. However, heavy security
  restriction are also a burden for maintenance staff, as they are
  interested in having an easy life. This kind of conflict can create
  vulnerabilities.
\item
  \textbf{Vulnerabilities:} All indentifiable weaknesses in the current
  system layout.
\item
  \textbf{Interactions:} Assets, IT-Systems, Humans and Vulnerabilites
  are in continous interaction with each other in order to
  \emph{support} the \textbf{Business Objectives}. Those interactionscan
  also \emph{lead to} \textbf{Threats}, i.e.~having unencrypted
  communication with a server. Therefore all interactions which occur in
  the system's outline have to be documented.
\end{itemize}

\subparagraph{Potential}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Chances/Risks:} Chances and Risks \emph{affect}
  \textbf{Business Objectives} (and \textbf{Threats}).
\item
  \textbf{Business Objectives:} Chances the system under study. This
  describes the desired state.
\item
  \textbf{Threats:} Risks for business objectives, this can simply be
  the integrety of the IT-System. Threats \emph{target} one or more
  \textbf{Assets} and violate \textbf{Security Requirements}. This
  describes the dreaded state.
\item
  \textbf{Security Requirements:} A set of distinct requirements for a
  safe and secure system. Security requirements \emph{hold for} one or
  more \textbf{Assets}. This describes the potential state after this
  procedure.
\end{itemize}

\subparagraph{Plan}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Business Model:} The plan to achieve \textbf{Business
  Objectives}
\item
  \textbf{Accident Scenario:} This is either the plan of a certain
  attack or the concrete outline of random disaster. Although the latter
  is most likely hard to anticipate.
\item
  \textbf{Attack Technique:} A specific technique or technology to
  attack IT-Systems (Man in the Middel, Phishing, etc.). (\textbf{Note:}
  This component is not conform to the glossary as the term
  \emph{accident} there covers both disaster and attack!)
\item
  \textbf{Security Measures:} The plan to achieve a safe and secure
  system. Each \textbf{Security Mesaure} \emph{opposes} an
  \textbf{Accident Scenario} or an \textbf{Attack Technique}.
\end{itemize}

\subparagraph{Event}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Business Process:} The actual, running instance of the
  \textbf{Busieness Modell}
\item
  \textbf{Accidents:} All actually happened accidents.
\item
  \textbf{Attacks:} All actually happened attacks. (\textbf{Note:} This
  component is not conform to the glossary as the term \emph{accident}
  there covers both disaster and attack!)
\item
  \textbf{Security Operations:} Instances of \textbf{Security Measures}.
  Security Measures \emph{secure} the system against \textbf{Accidents}
  and \textbf{Attacks}.
\end{itemize}

\subsubsection{Procedure}

The analysis procedure is an incremental and iterative process following
the four views of the previously described ontology.

%\includegraphics{../diagrams/png/itsec-ref-model-grimm-procedure.png}
\input{./figures/The-IT-Security-Reference-Analysis-Grimm}

\paragraph{Step 1. World Analysis}

At first, one has to outline the current state of the system under
study. This includes description of:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  all \textbf{Assets} which must be protected
\item
  all relevant \textbf{IT-Systems}
\item
  all involved \textbf{Humans} and their \textbf{Conflicts of Interests}
\item
  all known \textbf{Vulnerabilities}
\item
  and all important \textbf{Interactions} between the former components.
\end{itemize}

\paragraph{Step 2. Potential Analysis}

Secondly, one needs to outline the potential state of the system under
study. This includes both the dreaded (\textbf{Threats} and
\textbf{Risks}) and the desired (\textbf{Business Objectives},
\textbf{Chances}, \textbf{Security Requirements} state. This step
produces four artefacts:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  a \emph{threat specifiation}, which identifies the \textbf{Threat} and
  its targeted \textbf{Assets}
\item
  a \emph{threat risk evaluation}, which detrmines the likelyhood of a
  threat manifestation
\item
  a \emph{hazard matrix}, which maps threats to known
  \textbf{Vulnerabilities} and identifies potential hazards
\item
  a \emph{security requirement specifiation}, which specifies
  requirments in order to deal with identified hazards
\end{itemize}

\paragraph{Step 3. Plan Development}

Based on \textbf{Step 2.}, the identified hazards are used alongside
realistic \textbf{Accident Scenarios} (\textbf{Attack Techniques}) to
create a \emph{risk matrix}. With this matrix it is possible to decide
if the risk is acceptable or not. Together with the previously specified
\textbf{Security Requirements}, the matrix is used to define adequate
\textbf{Security Measures}. Like a \textbf{Business Model} is an
abstract concept to achieve \textbf{Business Objectives}, this step
creates an abstract concept to improve the system's security.

\paragraph{Step 4. Plan Deployment}

Finally, the \textbf{Security Measures} have to be implemented.
Additionally, all \textbf{Business Operations}, \textbf{Accidents}
(\textbf{Attacks}) and executed \textbf{Security Operations} will be
recorded through out the production phase.

The implementation of \textbf{Security Measures} evantually changes the
\textbf{World} and renders the conducted analysis outdated. So this
analysis procedure needs to be conducted again.

\subsubsection{Abstraction Levels of the Reference Model}

The Reference Model can be used on different levels of abstraction. This
means each component can be used within a wide range of granularity, for
instance the security measure \emph{Encryption} can be explored in
genral or on the level of different concrete encryption tools; or on the
even finer level of concrete algorithms.

The utilized abstraction level is not important for the analysis
procedure, it depends on the intended audience for the analysis.
However, it is important to use one abstraction level consistently
through out the analysis.

\section{Live+Gov Privacy Protection Analysis}

\subsection{Step 1. World Analysis}

\subsubsection{Assets: Privacy}

We focus on the Asset privacy as described in Section \ref{sec:taxonomy}

\subsubsection{IT-Systems \& Humans}

\paragraph{Server Side Mining Scenario (Figure \ref{figure:Live+Gov Operation Scenario 01 - Server Side Mining})}

%\includegraphics{../diagrams/png/scenario01-ServerSideMining.png}
\input{./figures/LG-Operation-Scenario-01-Server-Side-Mining}

A citizen (a) carries a mobile device (b) running the L+G client. The
mobile device collects sensor data and sends it to the L+G Data Cetner
(c). The L+G Data Center also sends beneficial information for citizens
to the mobile device.

The L+G Maintenance Crew (d) maintaince the L+G Data Center.

The Aggregation Monitor (e) queries the L+G Data Center for aggregated
data. Local Authorities (f) use the Aggregation Monitor to get
information in order to improve public services. \textbf{In this
scenario the mobile the server conducts data mining on aggregated user
data and additionally trys to combine it with publicly accessible data
to enhance the results.}

When no connection to the L+G servers is available, the mining
end-products are stored on the mobile device.

\paragraph{Mobile Mining Scenario (Figure \ref{figure:Live+Gov Operation Scenario 02 - Mobile Mining})}

%\includegraphics{../diagrams/png/scenario02-MobileMining.png}
\input{./figures/LG-Operation-Scenario-02-Mobile-Mining}

A citizen (a) carries a mobile device (b) running the L+G client. The
mobile device collects sensor data and sends it to the L+G Data Cetner
(c). The L+G Data Center also sends beneficial information for citizens
to the mobile device. \textbf{In this scenario the mobile device also
conducts data mining previously to sending the results to the L+G Data
Center.}

The L+G Maintenance Crew (d) maintaince the L+G Data Center.

The Aggregation Monitor (e) queries the L+G Data Center for aggregated
data. Local Authorities (f) use the Aggregation Monitor to get
information in order to improve public services.

When no connection to the L+G servers is available, the mining
end-products are stored on the mobile device.

\subsubsection{Conflict of Interests (Figure \ref{figure:Live+Gov Conflicts of Interests})}

%\includegraphics{../diagrams/png/conflict-of-interests.png}
\input{./figures/LG-Conflicts-Of-Interests}

\textbf{Citizen vs L+G Maintenance Crew}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  citizens want to be unidentifiable
\item
  citizens want certain informations about them to be secret
\item
  technical staff needs more or less unrestricted access to data in
  order to maintain the system
\end{itemize}

\textbf{Citizen vs Local Authority}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  citizens want to be unidentifiable
\item
  citizens want certain informations about them to be secret
\end{itemize}

\textbf{L+G Maintenance Crew vs Local Authority}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  staff wants proper payment
\item
  local authorities want to reduce cost
\end{itemize}

\textbf{Criminal vs Citizen}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  citizens want to be unidentifiable
\item
  citizens want certain informations about them to be secret
\item
  criminals want to posses certain information about a citizen
\end{itemize}

\textbf{Criminal vs L+G Maintenance Crew}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  criminals want to harm the system
\item
  technical staff has to keep the system safe and sound
\end{itemize}

\textbf{Criminal vs Local Authority}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  criminals want to harm the system
\item
  criminals want access to collected data
\item
  local authorities have to keep collected data restricted
\item
  local authorities want a stable system
\end{itemize}

\subsubsection{Vulnerabilities}

%\includegraphics{../diagrams/png/scenario-vulnerabilities.png}
\input{./figures/LG-Vulnerabilities}

\paragraph{Combined Scenario (Figure \ref{figure:Live+Gov Vulnerabilities})}

A citizen (a) carries a mobile device (b) running the L+G client. The
mobile device collects sensor data and sends it to the L+G Data Cetner
(c). The L+G Data Center also sends beneficial information for citizens
to the mobile device.

The L+G Maintenance Crew (d) maintaince the L+G Data Center.

The Aggregation Monitor (e) queries the L+G Data Center for aggregated
data. Local Authorities (f) use the Aggregation Monitor to get
information in order to improve public services.

A criminal (g) threatens the L+G system by corrupting the L+G
Maintenance Crew or Local Authority, or by forcing access to L+G
system's hardware, software or communication. (It is also possible for
criminals to corrupt the citizen, although technically this would be no
threat to privacy but rather an exercise of privacy
{[}\hyperref[references]{1}{]}.)

\paragraph{List of Vulnerabilities}

The vulnerabilities of Live+Gov system outline are:

\begin{itemize}
\item
  unencrypted data transmission (Man In The Middel)
\item
  insecure mobile devices (Exploit)
\item
  insecure servers (Exploit)
\item
  inadequate access rules for
  \begin{itemize}
  \item
    mobile devices (citizens)
  \item
    servers (staff)
  \item
    applications (local authority)
  \end{itemize}
\item
  corrupt/unhappy authorities or staff (\textbf{Could be actually a
  threat not a vulnerability})
\end{itemize}

(\textbf{Note:} The vulnerabilities seem to mostly open possibilities to
violate the \textbf{Privacy of Data and Image}, as this type is concerd
with making data \emph{automatically} available to others.)

\subsection{Step 2. Potential Analysis}

\subsubsection{Threats}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Unauthorized access to privacy sensitive data}, caused by
  \begin{itemize}
  \item
    \textbf{Excessive Data Mining} Linking sensor data provided by
    citizens with additional sources can produce more privacy sensitive
    data.
  \item
    \textbf{Corrupt Local Authorities} Local Authorities have certain data
    access, they could hand over this access for monetary reasons.
  \item
    \textbf{Corrupt/Unhappy Staff} Staff members also have a certain data
    access and they also could hand over this access for monetary reasons
    or as a form of payback for unfair treatment.
  \item
    \textbf{Competent Attackers} Competent Attackers are always a threat
    for IT-systems. Normaly they have monetary reasons to exploit a
    system, but there is also a possibility for proof-of-concept like
    attacks.
  \end{itemize}
\end{itemize}

\paragraph{Implicit Privacy Type Violation (Figure \ref{figure:Implicit Privacy Violation Matrix})}

%\includegraphics{../diagrams/png/implicit-privacy-violation-matrix.png}
\input{./figures/Implicit-Privacy-Violation-Matrix}

\subparagraph{Privacy of The Person}

The Privacy of The Person is concerned with one's biometric privacy. If
this type is violated, following implicit violations are possible:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Privacy of The Person:} reflexive violation
\item
  \textbf{Privacy of Behaviour and Action:} none
\item
  \textbf{Privacy of Communication:} none
\item
  \textbf{Privacy of Data and Image:} none
\item
  \textbf{Privacy of Thoughts and Feelings:} Some psychological diseases
  (e.g.~depression) have physiological impact. Such physiological
  patterns could be detected.
\item
  \textbf{Privacy of Location and Space:} none
\item
  \textbf{Privacy of Association:} none
\end{itemize}

\subparagraph{Privacy of Behaviour and Action}

The Privacy of Behaviour and Action is concerned with one's privacy
regarding social activities (religious, political, sexual, \ldots{}). If
this type is violated, following implicit violations are possible:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Privacy of The Person:} Religious practices may include body
  modifications (e.g.~circumcision).
\item
  \textbf{Privacy of Behaviour and Action:} reflexive violation
\item
  \textbf{Privacy of Communication:} none
\item
  \textbf{Privacy of Data and Image:} none
\item
  \textbf{Privacy of Thoughts and Feelings:} Social activities in
  general depend on a certain intellectual attitude. Such an actvity is
  the expressions of such an attitude.
\item
  \textbf{Privacy of Location and Space:} none
\item
  \textbf{Privacy of Association:} Recording religious, politcal or
  sexual activities can reveal association with churches, political
  parties or sexual partners.
\end{itemize}

\subparagraph{Privacy of Communication}

The Privacy of Communication is concerned with not havin such
communication (corresponence or vis-a-vis) intercepted. This is very
broad type of privacy. Depending on the contens of the intercepted
communication every other type can be violated:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Privacy of The Person:} Communication about body
  characteristics.
\item
  \textbf{Privacy of Behaviour and Action:} Communicatio about social
  activities.
\item
  \textbf{Privacy of Communication:} reflexive violation
\item
  \textbf{Privacy of Data and Image:} Communication containing one's
  passwords or other sensitive data.
\item
  \textbf{Privacy of Thoughts and Feelings:} Communication of thoughts
  and feelings, e.g.~wiretapping a flirt or a catholic confession
  ritual.
\item
  \textbf{Privacy of Location and Space:} Interception of face-to-face
  communication is only possible if one's location and space is violated
  (wiretapping).
\item
  \textbf{Privacy of Association:} Communication about one's
  associations (family members, churches, etc.).
\end{itemize}

\subparagraph{Privacy of Data and Image}

The Privacy of Data and Image is concered with one's data not beeing
automatically availabel to others. This also is a very broad type of
privacy. Depending of the data or image contents erery other type can be
violated:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Privacy of The Person:} Images or stored biometric information
  reveal one's physical characteristics.
\item
  \textbf{Privacy of Behaviour and Action:} Images or diaries can reveal
  one's social activities.
\item
  \textbf{Privacy of Communication:} Modern communication systems
  usually contain some sort of archive function, e.g.~E-mail clients do
  not automatically delete messages. Such messages are data and reveal
  one's communication.
\item
  \textbf{Privacy of Data and Image:} reflexive violation
\item
  \textbf{Privacy of Thoughts and Feelings:} Images can show one's
  emotional state.
\item
  \textbf{Privacy of Location and Space:} Images can reveal one's
  location, e.g.~making a picture in front of the Eifel Tower.
\item
  \textbf{Privacy of Association:} E-mail data can also reveal
  association.
\end{itemize}

\subparagraph{Privacy of Thoughts and Feelings}

The Privacy of Thoughts and Feelings is concerned with keeping such
thoughts and feelings secret. If this type is violated, following
implicit violations are possible:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Privacy of The Person:} Thoughts and feelings can reveal
  medical conditions.
\item
  \textbf{Privacy of Behaviour and Action:} Thoughts and feelings can
  reveal a certain attitudes which create a foundation for certain
  social activities.
\item
  \textbf{Privacy of Communication:} none
\item
  \textbf{Privacy of Data and Image:} none
\item
  \textbf{Privacy of Thoughts and Feelings:} reflexive violation
\item
  \textbf{Privacy of Location and Space:} none
\item
  \textbf{Privacy of Association:} Thoughts and feelings can reveal
  individual association, e.g amorous feelings for a certain person.
\end{itemize}

\subparagraph{Privacy of Location and Space}

The Privacy of Location and Space is concerned with one's right to move
freely without being tracked and one's right to private places. If this
type is violated, following implicit violations are possible:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Privacy of The Person:} Frequently visited doctors can reveal
  certain medical conditions, if such doctors are known specialsts. In
  general it could imply ill-being.
\item
  \textbf{Privacy of Behaviour and Action:} Frequently visited places in
  general can reveal association and hence implies social activities.
\item
  \textbf{Privacy of Communication:} If one's location is knwon, it is
  possible to intercept (wiretap) one's communication. This also may
  violate the right to private spaces.
\item
  \textbf{Privacy of Data and Image:} If one's location is known, it is
  possible shoot pictures. This violates the right to one's image
  (\emph{``Recht am eigenen Bild''}).
\item
  \textbf{Privacy of Thoughts and Feelings:} Frequently visited persons
  may imply certain thoughts and feelings, e.g.~having a mistress.
\item
  \textbf{Privacy of Location and Space:} reflexive violation
\item
  \textbf{Privacy of Association:} Frequently visted places can reveal
  associations simply by searching in maps or yellow-pages.
\end{itemize}

\subparagraph{Privacy of Association}

The Privacy of Association is concerned with one's right to associate
with whomever one wants, without that association having recorded.If
this type is violated, following implicit violations are possible:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Privacy of The Person:} Association with tatoo artists could
  imply having tatoos or other body modifications
\item
  \textbf{Privacy of Behaviour and Action:} Association with churches or
  poltical organizations could imply certain activities.
\item
  \textbf{Privacy of Communication:} none
\item
  \textbf{Privacy of Data and Image:} none
\item
  \textbf{Privacy of Thoughts and Feelings:} Association with churchs of
  political organizations could imply a certain intellectual attitude.
\item
  \textbf{Privacy of Location and Space:} none
\item
  \textbf{Privacy of Association:} reflexive violation
\end{itemize}

\paragraph{Sensor Privacy Matrix (Figure \ref{figure:Live+Gov Sensor-Privacy Matrix})}

%\includegraphics{../diagrams/png/sensor-privacy-matrix.png}
\input{./figures/LG-Sensor-Privacy-Matrix}

\subparagraph{Privacy of The Person}

The \textbf{Privacy of The Person} is generally concerned with one could
best understand as \emph{Biometric Privacy}. Friedewald et al. paraphase
it as \emph{``{[}\ldots{}{]} the right to keep body functions and body
characteristics {[}\ldots{}{]} private''}. \textbf{Accelerometer},
\textbf{Rotation Vector} and \textbf{Gyroscope} measure the physical
movement of the mobile device on all three axes. If the moblile device
is carried \emph{``normally''} its safe to say that those sensors also
measure the movments of its carrier. So his privacy is infringed
regarding biometric behaviour, as it is captured automatically.
(\textbf{Note:} This is not to be confused with the \textbf{Privacy of
Behaviour and Action} which is used for the social aspects of behaviour
action, e.g.~praying, sexual habits or poltical activities.)

\subparagraph{Privacy of Data and Image}

The \textbf{Privacy of Data and Image} demands, that
\emph{``individual's data is not automatically available to other
individuals''}. This type of privacy is trivially threatened because
here sensor data is indvidual data, a priori. So every sensor violates
the privacy of data and image, as data is transported into a foreign
system where operators have access to it.

\subparagraph{Privacy of Location and Space}

According to Friedewald et al., the \textbf{Privacy of Location and
Space} is concerned with one's \emph{``right to move about in public or
semi-public space withoug being identified, tracked or monitored.''}.
This is the location aspect of this type. The space aspect is concerned
with one's \emph{``right to solitude''}, which generally includes one's
right to an inviolate home and other private spaces. Obviously the
\textbf{GPS} and \textbf{GSM} sensors violate such right about not being
tracked, because they reveal the position of the mobile device and its
carrier. The \textbf{GSM} sensor gives the exact cell, the mobile device
has registered with at the current moment. The \textbf{GPS} sensor gives
the current longitude and latitude, the current global position of the
mobile device and its carrier, although there is some artificial
inaccuracy within civil use.

The \textbf{WLAN} and \textbf{Bluetooth} sensors record lists of the
currently available local wireless networks and bluetooth clients. If
such are known stationary entities, those sensors are considered as
dangerous as the \textbf{GSM} sensor for the carriers locational
privacy.

The \textbf{Magnetic field} sensor is not regarded very dangerous to the
carrier's privacy, because it does not allow very precise localization.
But it can limit the possibilities for the global position of the mobile
device. Here, it is just named for completeness sake.

\subparagraph{Privacy of Associtation}

The \textbf{Privacy of Association} sates that everyone has the
\emph{``right to associate with whomever they whish, without being
monitored {[}automatically without reasonable suspicion{]}''}. This
includes indivisuals and organizations. \textbf{WLAN} and
\textbf{Bluetooth} sensors provide the ability to monitor such
associations if their lists contain kown entities. If one frequently
connects with an organizational wirelss network, e.g.~an university
network, an association can be deduced (student or staff). The same goes
for the \textbf{Bluetooth} sensor, if it is stationary. Additionally, if
the recorded bluetooth clients are mobile, it is more or less possible
to deduce association with the technical identity of (yet) anonymous
individuals.

Additionally, the \textbf{GSM} sensor could provide the association with
the GSM operator (\textbf{NEEDS TO BE VERIFIED!}).

\paragraph{Excessive Data Mining}

\subparagraph{1 Privacy of the Person}

\emph{``This type is the right keep body functions and body
characteristics private''} {[}\hyperref[references]{1}{]}. Here we are
concerned with the personal physiological and psychological privacy of
citizens. This privacy is threatened if \textbf{body characteristics}
(weight, height, body measures, fingerprints, dna,\ldots{}) or
\textbf{medical conditions} (limping, having a cold, suffering from
depression, visiting therapists or other doctors) \textbf{can be
revealed or inferred}.

The Live+Gov project collects sensor data from mobile devices
(accellerometer, location, \ldots{}). By applying \emph{human activity
recognition (HAR)} techniques we could detect certain movement patterns
and in fact detect limping or other pathologic movements. Additionally
we know the position with of indiviual citizens with a sufficient
accurancy, and the Live+Gov project is only applied to a destinct urban
area. So we could link HAR data with the yellow pages, filter for
medical specialists and limit the possibilities of pathologic
conditions. But even without HAR data we could determine a likelihood
for certain conplaints. Frequent vistits to the dentist does not imply
healthy teeth.

\subparagraph{2 Privacy of Behaviour and Action}

\emph{``This type is also concerned with the `protection against
disclosure of personal matters' through behaviour''}
{[}\hyperref[references]{1}{]}. This type is concered with religous
practices, sexaul habits, political activities, etc. revealed through
observation. For the Live+Gove project this type is related to the
Privacy of Location and Space and the Privacy of Association. Locational
data is recorded by default and by linking it to maps and public
registers like phone books those *``personal matters*" could be
disclosed by frequently visited places (churches, brothels, party head
quaters, \ldots{}).

\subparagraph{3 Privacy of Communication}

\emph{``It `aims to avoid the interception of communications' either
electronic or face-to-face''} {[}\hyperref[references]{1}{]}. This type
is threatened if we intrude the secrecy of correspondence, posts and
telecommunications, personal direct communication or right to free
discussion without third parties listening. The Live+Giv project only
could threaten this type by using mobile client application as trojan
horse, collecting any communication data (chat, sms, microphone).

\subparagraph{4 Privacy of Data and Image}

\emph{``This type is concerned with `making sure that individuals's data
is not automatically available to other individuals and
organisations'\,''} {[}\hyperref[references]{1}{]}. This is
Informational Privacy in an intuitive sense regarding data like:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Phone Number
\item
  IP Address
\item
  Public-administrative Data (Date of Birth, population register)
\item
  Data held by organizations, like Banks or Insurance Companies
\item
  All data that is stored in online services (Facebook)
\end{itemize}

The Live+Gov project does not threaten this type directly if we assume a
secure and closed system. However, this privacy type could be threatened
by carelessly ignoring known (technical) vulnerabilities regarding data
security. Additionally we could threaten the Privacy of Data and Image
indirectly by linking collected data with other sources.

\subparagraph{5 Privacy of Thoughts and Feelings}

\emph{``This type is the right `not to share their thoughts or feelings
or to have those thoughts or feeling (sic!) revealed'\,''}
{[}\hyperref[references]{1}{]}. This means thoughts and emotions must
not be detected autmatically. Intuitively the Live+Gov system seems
unable to threaten this type. However, considering the issue component
of the Urban Maintaince use case, this might reveal information about
one's thoughts regarding the community in a positive manner. Solely by
taking part we could assume a caring personality. This threatens one's
privacy in a rather technical sense, but does not necessarily impose any
harm.

An other way of threatening this type could be constructed by recording
one's voice with the phone's microphone and run emotion detecting
algorithms against this data.

\subparagraph{6 Privacy of Location and Space}

\emph{``This type is the right `to move about in public or semi-public
space withoug being identified, tracked or monitored'. Additionally this
type is concerned with the protection of one's home and private places
(`right to solitude')''} {[}\hyperref[references]{1}{]}. The location
dimension of this type is relatively easy to understand: The
geo-position of citizens cannot be monitored by default. However, by
actively taking part in the Live+Gov project, the locational privacy of
citizens is threatened by default because data of the location sensor
will be collected. The space dimension is more complex, but can be
simplified with a \emph{``right to solitude''}. This dimension could be
threatend with the invasion of personal space in any means, i.e.~by
disrespecting one's right to an inviolate home or by undercutting one's
comfort zone in an conversation. Live+Gov only utilizes the mobile
devices of citizens, so we could disrespect the former by activating the
phone's micropone and start recording.

\subparagraph{7 Privacy of Association}

\emph{``This type is the right `to associate with whomever {[}one{]}
wish, withoug being monitored'.''} {[}\hyperref[references]{1}{]}. This
means that one's associations must not be recoreded by default
independet from any suspicion. Anyhow, it does not mean that this right
cannot be forfeit given a reasonable suspicion. The Live+Gov project
could easily threaten this type just by linking locational data with
yellow pages and a map. Even if an association graph cannot be deducted
for individual citizens, we could aggregate the data to create an
association graph for a whole population as the Live+Gov system is
applied to a restricted urban area.

\subsection{Step 3. Plan Development}

\subsubsection{Security Measures}

\paragraph{The 7 C's of user privacy control (Figure \ref{figure:The Seven Cs of User Privacy Control})}

This is a note on an excerpt from the article \emph{Sociotechnical
Architecture for Online Privacy} {[}\hyperref[references]{1}{]} called
\textbf{The 7 C's of user privacy control}. Those 7 C's are aspects
which should be covered by measures for implementing user privacy. They
derive from an interpretation of privacy which could be summarized as
\emph{``One's ability to control/seclude information about themselve''}.

%\includegraphics{../diagrams/png/The7Cs.png}
\input{./figures/The-7-Cs-Of-User-Privacy-Control}

\subparagraph{Comprehension}

\emph{``Users should \textbf{understand} ho personal identifiable
information (PII) is handled, who's collecting it and for what purpose,
and who will proccess the PII and for what purpose. Users are entitled
to know all parties that can access their PII, the limits to processing
transparency, why the PII data is being requested, when the data will
expire (Either from a collection or database), and what happens to it
after that. This category also include legal rights around PII, and the
imolications of a contract when one is formed.''}

This C implements transparancy regarding user data and user privacy.
Comprehension should answer the following questions:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \emph{WHO} collects data?
\item
  \emph{WHAT} data will be collected?
\item
  \emph{WHY} will data be collected and processed?
\item
  \emph{HOW} will data be collected and processed?
\item
  \emph{WHEN} will data expire?
\item
  What is allowed?
\item
  What choices ar possible?
\end{itemize}

All in all information of what's happening and why has to made
accessible for users.

\subparagraph{Consciousness \textbf{(critical!)}}

\emph{``Users should \textbf{be aware} of when data collections occurs,
when a contract is being formed between a user and data collector when
their PII is set to expire, who's collecting the data, with whom the
data will be shared, how to subsequently access the PII, and the
purposes for which the data is being collected.''}

This C seems to be critical for privacy protection. Consciousness
complements Comprehension in respect that the latter just states that
hard facts need to be delivered. However, those facts might get hidden
in a terms and conditions section which nobody reads but still accepts
anyway. In order to prevent that Consciousness states that a certain
level of \textbf{Awareness} of those facts needs to be establisched.

\subparagraph{Choice}

\emph{``Users should \textbf{have choices} regarding data collection
activies in terms of opting in or out, whether or not to provide data,
andhow to cerrect their data.''}

Self explaining. This is the actual control enabled by the 7 C's.

\subparagraph{Consnet}

\emph{``Users must first \textbf{consent} (meaning informed, explicit,
unambiguous agreement) to data collection, use, and storage proposals
for any PII. Privacy consent mechanisms should explicitly incorporate
the mechanisms of comprehension, consciousness, limitations, and
choice.''}

This C might be special case of Choice. Before taking part a user should
have the choice wether to join or not (Opt-In).

\subparagraph{Context}

\emph{``Users should \textbf{be able to change pricacy preferences}
according to context. Situational or physical context - such as crowded
situations (for example, when at a service desk where severeal people
can listen in on your exchange when you provide a phone number, or when
you're in an online community chat room) - is different from wehn you
perform a buy transaction with Amazon.com or in rooms with cameras
(where digitization makes the information permanent and unmistakably
you) and data context (such as the sensitivity of data, for example
health data could dictate different actions on the same PII in different
contexts.''}

Self explaining. Refines Choice in context sensitive manner.

\subparagraph{Confinement}

\emph{``Users should \textbf{be able to set limits} on who may access
their PII, for what purposes, and where and possibly when it may be
stored. Setting limits could provide some good opportiunities for future
negotiation between vendors and users.''}

Self explaining. Refines Choice regarding data collection and
processing.

\subparagraph{Consistency}

\emph{``Users should \textbf{anticipate} with reasonable certainty what
will occur if any action involving their PII is taken. That is, certain
actions should be predictable on user acces of PII or giving out of
PII.''}

Information given by Comprehension needs to be reliable to found
choices.

\paragraph{The 2 Steps of the 7 C's (Figure \ref{figure:The 2 Steps of the 7 Cs of User Privacy Control})}

If we look closer at the 7 C's and how they try to enable control, we
see that a 2 step approach is taken:

%\includegraphics{../diagrams/png/7Cs2Steps.png}
\input{./figures/The-2-Stepos-Of-The-7-Cs-Of-User-Privacy-Control}

\subparagraph{(I) Enable \emph{Adequate} Control}

The 7 C's try to enable control through choices. A user should be able
to choose which data can be collected and processed depending on context
and who will have access to the data. But cannot be random. In order to
make substantiated choices and enable \emph{adequate} control a user
needs have

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Comprehension:} access to hard facts
\item
  \textbf{Consistency:} trust that those facts are reliable
\item
  \textbf{Conciousness:} awareness of those facts to found choices
\end{itemize}

\subparagraph{(II) Enable \emph{Actual} Control}

Naturally after creating a certain amount of knowledge, a user needs
also access to opportunities to make use of it. Therefore a user needs
possibilites to actually make choices. Additionally to having choices at
all, the 7 C's have 3 special choice categories:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Choice}
  \begin{itemize}
  \item
    \textbf{Consent:} the choice to opt-in
  \item
    \textbf{Confinement:} the choice to set limits regarding user data
  \item
    \textbf{Context:} the choice to set limits regarding user data
    depending on certain context
  \end{itemize}

\end{itemize}

\section{References}

### 2014-09-11 figures/Implicit-Privacy-Violation-Matrix.tex
Added Privacy Protection tex from Max.
Signed-off-by: Heinrich Hartmann <hartmann@uni-koblenz.de>

\begin{figure}
\centering
\includegraphics{diagrams/png/implicit-privacy-violation-matrix.png}

\begin{flushleft}
\scriptsize
\textbf{Legend:}
The matrix above shows the following relation: \emph{Type X can be
implicitly violated by the violation of type Y.}
\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  The X-Axis shows the Seven Types of Privacy according to Friedewald et
  al., which could be violated implicitly
\item
  The Y-Axis shows the Seven Types of Privacy according to Friedewald et
  al., which have been violated explicitly
\item
  Big black bullet points denote, that an implicit violation is possible
\item
  Big grey bullet points only denote, that the relation is reflexive
  (\emph{a R a}). They are only shown for completeness sake and not
  discussed further, because they denote a trivial fact.
\end{itemize}
\end{flushleft}

\caption{Implicit Privacy Violation Matrix}
\label{figure:Implicit Privacy Violation Matrix}
\end{figure}

### 2014-09-11 figures/LG-Operation-Scenario-01-Server-Side-Mining.tex
Added Privacy Protection tex from Max.
Signed-off-by: Heinrich Hartmann <hartmann@uni-koblenz.de>

\begin{figure}[h]
\centering
\includegraphics{diagrams/png/scenario01-ServerSideMining.png}

\begin{flushleft}
\scriptsize
\textbf{Legend:}
\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{(a) Citizen:} User of the L+G client application whose privacy
  is at stake.
\item
  \textbf{(b) Mobile Device:} Runs the L+G client application, produces
  and stores data sensitive to the users privacy.
\item
  \textbf{(c) L+G Data Center:} Runs the L+G services, proccesses and
  stores user data.
\item
  \textbf{(d) L+G Maintenance Crew:} Technical staff with access to
  critical infrastructure.
\item
  \textbf{(e) Aggregation Monitor:} Interface to aggregated user data.
\item
  \textbf{(f) Local Authority:} Provider of the L+G system.
\end{itemize}
\end{flushleft}

\caption{Live+Gov Operation Scenario 01 - Server Side Mining}
\label{figure:Live+Gov Operation Scenario 01 - Server Side Mining}
\end{figure}

### 2014-09-11 figures/LG-Sensor-Privacy-Matrix.tex
Added Privacy Protection tex from Max.
Signed-off-by: Heinrich Hartmann <hartmann@uni-koblenz.de>

\begin{figure}
\centering
\includegraphics{diagrams/png/sensor-privacy-matrix.png}

\begin{flushleft}
\scriptsize
\textbf{Legend:}
\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  The x-axis lists the 7 Types of Privacy according to Friedewald et al.
\item
  The y-axis lists all mobile sensors currently used by the Live+Gov
  procject.
\item
  The big black bullet point dentotes that there is a \textbf{direct}
  violoation of a privacy type from a sensor
\end{itemize}
\end{flushleft}

\caption{Live+Gov Sensor-Privacy Matrix}
\label{figure:Live+Gov Sensor-Privacy Matrix}
\end{figure}

### 2014-09-14 figures/ThreatTable.tex
started ThreatTable with mapping to CofI and Vulnerabilities

% =================================================
% Threat Table Macros

% Const Column Width
\newcommand{\ThreatTableColWidth}{5cm}

% Const Header Row Height
\newcommand{\ThreatTableHeaderRowHeight}{0.7cm}

% Const Content Row Height
\newcommand{\ThreatTableContentRowHeight}{1cm}

% Define Header Cell Makro
\newcommand{\ThreatTableHeaderCell}[1]{
\begin{minipage}[t][\ThreatTableHeaderRowHeight][c]{\ThreatTableColWidth}
\centering
\textbf{#1}
\end{minipage}
}

% Define Content Cell Makro
\newcommand{\ThreatTableContentCell}[1]{
\begin{minipage}[t][][c]{\ThreatTableColWidth}
\begin{flushleft}
\scriptsize
#1
\newline
\end{flushleft}
\end{minipage}
}

% Define Header Row Makro
\newcommand{\ThreatTableHeaderRow}[4]{
\ThreatTableHeaderCell{#1}
&\ThreatTableHeaderCell{#2}
&\ThreatTableHeaderCell{#3}
&\ThreatTableHeaderCell{#4}
\\ \hline
}

% Define Content Row Makro
\newcommand{\ThreatTableContentRow}[4]{
\ThreatTableContentCell{#1}
&\ThreatTableContentCell{#2}
&\ThreatTableContentCell{#3}
&\ThreatTableContentCell{#4}
\\ \hline
}

%  Define Threat Table Environment
\newenvironment{ThreatTable}
{
\begin{tabular}{|c|c|c|c|}
\hline
\ThreatTableHeaderRow
{Description}
{Conflict of Interest}
{Vulnerabilites}
{Assets (Privacy Types)}
}
{\end{tabular}}

% =================================================
% Threat Table Content

\begin{landscape}
\begin{figure}
\centering
\begin{ThreatTable}

\ThreatTableContentRow
{\textbf{(Zero Day) Exploits}
\\A criminal uses (Zero Day) Exploits to obtain access to hardware or software which stores or processes
privacy sensitive data in order to get that data.}
{Criminals want to obtain access or data for personal profit, i.e. by (re-)selling the data or  by processing it themselves. 
However, criminals may have no financial interest, they could 
\\\textbf{Criminal vs Citizen}
\\
\\ \textbf{Criminal vs Local Authorities}
\\ \textbf{Criminal vs Maintenance Staff}}
{Unsecure Hardware- or Software-Interfaces}
{\textbf{Explicit:} 1,4,6,7}

\ThreatTableContentRow
{\textbf{Man in the Middle} 
\\A criminal intercepts communication between mobile device and server or between server and application.}
{arg2}
{Unencrypted hardware or software communication}
{\textbf{Explicit:} 1,4,6,7}

\ThreatTableContentRow
{\textbf{Corrupt Employees} 
\\ Employee sells private data to advertisers}
{\textbf{Employee vs Citizen} 
\\ Employees want to make personal profit by selling citizen data. 
\\ Citizens provide data for public improvement, they don't want their data to be used for other purposes.}
{Full DB access of Employees}
{\textbf{Explicit:} 1,4,6,7}

\ThreatTableContentRow
{\textbf{Corrupt Local Authorities}}
{arg2}
{arg3}
{\textbf{Explicit:} 1,4,6,7}

\ThreatTableContentRow
{\textbf{Careless Citizen}}
{arg2}
{arg3}
{\textbf{Explicit:} 1,4,6,7}

\ThreatTableContentRow
{\textbf{Intransparent Data Mining}}
{arg2}
{arg3}
{arg4}

\end{ThreatTable}

\caption{Threat Table}
\end{figure}
\end{landscape}

### 2014-09-15 figures/ThreatTable.tex
completed threat table

@@
 % Const Header Row Height
-\newcommand{\ThreatTableHeaderRowHeight}{0.7cm}
+\newcommand{\ThreatTableHeaderRowHeight}{0.5cm}
 
@@
 \centering
+\scriptsize
 \textbf{#1}
@@
 \begin{flushleft}
-\scriptsize
+\tiny
 #1
@@
 {Criminals want to obtain access or data for personal profit, i.e. by (re-)selling the data or  by processing it themselves. 
-However, criminals may have no financial interest, they could 
+However, criminals may have no financial interest, they could also gain personal (ego) profit by testing proof-of-concept attacks.
 \\\textbf{Criminal vs Citizen}
-\\
-\\ \textbf{Criminal vs Local Authorities}
-\\ \textbf{Criminal vs Maintenance Staff}}
-{Unsecure Hardware- or Software-Interfaces}
+\\Citizens only allowed Local Authorities to use their data. They want their data to be secret to others. (Additionally, citizens are
+also interested in a working system, which they payed for via taxes.)
+\\\textbf{Criminal vs Local Authorities}
+\\Local Authorities have capital and reputation invested in a working system. Successful attacks undermine both.
+\\\textbf{Criminal vs Maintenance Staff}
+\\Staff members have a professional ethos and a duty to provide working systems. Successful attacks offend the former and obstruct the latter.}
+{Unsecured Hardware- or Software-Interfaces}
 {\textbf{Explicit:} 1,4,6,7}
@@
 \\A criminal intercepts communication between mobile device and server or between server and application.}
-{arg2}
+{\textit{Like Exploits}}
 {Unencrypted hardware or software communication}
@@
 {\textbf{Corrupt Employees} 
-\\ Employee sells private data to advertisers}
-{\textbf{Employee vs Citizen} 
-\\ Employees want to make personal profit by selling citizen data. 
-\\ Citizens provide data for public improvement, they don't want their data to be used for other purposes.}
-{Full DB access of Employees}
+\\ An employee abuses his database access to obtain private citizen data in order to sell it to advertisers.}
+{\textbf{(Corrupt) Employee vs Citizen} 
+\\ Corrupt Employees want to make personal profit by selling citizen data. 
+Citizens provide data for public improvement, they don't want their data to be used for other purposes, which
+may lead to negative effects for themselves.
+Employees in general need easy database access to do their job. But this also means easy access to privacy
+sensitive data of citizens. This diametrically opposes the interest of citizens to have such information unknown
+to other individuals.}
+{Full database access of Employees}
 {\textbf{Explicit:} 1,4,6,7}
@@
 \ThreatTableContentRow
-{\textbf{Corrupt Local Authorities}}
-{arg2}
-{arg3}
+{\textbf{Corrupt Local Authorities}
+\\A member of the Local Authority abuses his access to applications 
+to obtain aggregated citizen data in order to use it for illegitimate purposes, e.g. selling it.}
+{\textbf{(Corrupt) Local Authority Member vs Citizen}
+\\\textit{Like Corrupt Employees}}
+{Full application access of Local Authorities}
 {\textbf{Explicit:} 1,4,6,7}
@@
 \ThreatTableContentRow
-{\textbf{Careless Citizen}}
-{arg2}
-{arg3}
+{\textbf{Careless Citizen}
+\\A careless citizen allows others (Criminals) to have unrestricted access to his mobile device. 
+Hence, he creates to possibility to install spy-ware or have the device destroyed.}
+{\textbf{Criminal vs Citizen}
+\\Criminals want to have access to mobile devices to obtain private data of citizens in order to 
+gain personal profit - or to simply render the device useless. On the other hand, citizens have a 
+natural interest in keeping personal data secret in order to prevent financial loss or because having
+sensitive information accessible to others violates their privacy.}
+{Insufficient access rules for mobile devices}
 {\textbf{Explicit:} 1,4,6,7}
@@
 \ThreatTableContentRow
-{\textbf{Intransparent Data Mining}}
-{arg2}
-{arg3}
-{arg4}
+{\textbf{Intransparent Data Mining}
+\\Local Authorities or System Providers use their technical knowledge, data mining capabilities and 
+additional data sources to obtain/create more information about citizens.}
+{\textbf{Citizen vs Local Authority or System Provider}
+\\Citizens only agreed to share certain sensitive data with Local Authorities and System Providers to help society.
+They are not interested in negative effects as a result of such a good willing act.
+However, Local Authorities and System Providers have an interest to maximize the profit of their investments.
+Local Authorities could (secretly) use mined data for security or health care issues. System Providers could
+(secretly) sell mined data to illegitimate customers, e.g. the SCHUFA. This could lead to repressive behaviour of law
+enforcement or negative scores.}
+{Unaware Citizens}
+{1,2,3,4,5,6,7}
 
 \end{ThreatTable}
+
+{\scriptsize (\textbf{Note for Meeting:} I think we need discrete (numbered) lists of interests for each actor in the \textit{Humans} section.)}
 

### 2014-09-15 PrivacyProtection-2014-09-09.tex
started ThreatTable with mapping to CofI and Vulnerabilities
added combined diagram for it systems
Partial cleanup up formatting and arrangement of sections.

@@
-\section{Privacy}
+\chapter{Privacy Protection}
 
@@
 
-\subsection{Histrocial Approaches to Privacy}
-
-\subsubsection{Aristotle}
+\section{Histrocial Approaches to Privacy}
+
+\subsection{Aristotle}
 
@@
 
-\subsubsection{Warren and Brandeis}
+\subsection{Warren and Brandeis}
 
@@
 
-\subsubsection{Fried}
+\subsection{Fried}
 
@@
 
-\subsection{Legal Aspects of Privacy}
+\section{Legal Aspects of Privacy}
 
@@
 
-\subsubsection{European Convention on Human Rights - Article 8}
+\subsection{European Convention on Human Rights - Article 8}
 
@@
 
-\subsubsection{EU Data Protection Direvtive}
+\subsection{EU Data Protection Direvtive}
 
@@
 
-EU directives in general do not hold any direct legal binding for EU
-citzens. Member states have to create their own legal implementation.
-
-``The directive regulates the processing of personal data regardless of
-whether such processing is automated or not.''
+EU directives in general do not hold any direct legal binding for EU citzens.
+Member states have to create their own legal implementation.
+
+``The directive regulates the processing of personal data regardless of whether such processing is automated or not.''
 {[}\hyperref[references]{1}{]} where:
@@
 \begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \textbf{personal data} is ``any information relating to an identified
-  or identifiable natural person {[}\ldots{}{]} (art. 2a)''
+\item
+  \textbf{personal data} is ``any information relating to an identified or identifiable natural person {[}\ldots{}{]} (art. 2a)''
   {[}\hyperref[references]{1}{]}
 \item
-  \textbf{processing} is ``any operation or set of operations which is
-  performed upon personal data {[}\ldots{}{]} (art. 2b)''
+  \textbf{processing} is ``any operation or set of operations which is performed upon personal data {[}\ldots{}{]} (art. 2b)''
   {[}\hyperref[references]{1}{]}
@@
 
-The ``identified or identifiable natural person'' is called \textbf{data
-subject}.
-
-The sole responsibility for compliance is held by so called
-\textbf{controllers}. A controller is an actor ``which alone or jointly
-with others dertimines the purposes and means of the processing of
-personal data; (art 2d)'' {[}\hyperref[references]{1}{]}
-
-(\textbf{Note:} this rule is applicable ``whenever the controller uses
-equipment situated within the EU''{[}\hyperref[references]{1}{]}. This
-will hold for every e-commerce provider \textbf{inside or outside} of
-the EU because the customer's computer is \emph{situated within the EU}
-anyways.)
-
-``Personal data should not be processed at all, except when certain
-conditions are met'' {[}\hyperref[references]{1}{]}. These conditions
-incorporate the seven OECD princepals and are futher categorized in the
-categories:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \textbf{Transparency}: ``The data subject has the right to be informed
-  when his personal data is being processed {[}\ldots{}{]} (art. 10 and
-  11)'' {[}\hyperref[references]{1}{]} and he has the right how and by
-  whom it is processed. Additionally other speficied requirements have
-  to be met, i.e.~giving consent.
-\item
-  \textbf{Legitimate prupose}: ``Personal data can only be processed for
-  specified explicit and legitimate pruposes and may not be processes
-  futher in a way incompatible with those pruposes (art. 6b)''
+The ``identified or identifiable natural person'' is called \textbf{data subject}.
+
+The sole responsibility for compliance is held by so called \textbf{controllers}.
+A controller is an actor ``which alone or jointly with others dertimines the purposes and means of the processing of personal data; (art 2d)''
+{[}\hyperref[references]{1}{]}
+
+This rule is applicable ``whenever the controller uses equipment situated within the EU''{[}\hyperref[references]{1}{]}.
+This will hold for every e-commerce provider \textbf{inside or outside} of the EU because the customer's computer is \emph{situated within the EU} anyways.
+
+``Personal data should not be processed at all, except when certain conditions are met'' {[}\hyperref[references]{1}{]}.
+These conditions incorporate the seven OECD princepals and are futher categorized in the categories:
+
+\begin{itemize}
+\item
+\textbf{Transparency}: ``The data subject has the right to be informed when his personal data is being processed {[}\ldots{}{]} (art.
+10 and 11)'' {[}\hyperref[references]{1}{]} and he has the right how and by whom it is processed.
+Additionally other speficied requirements have to be met, i.e.~giving consent.
+\item
+\textbf{Legitimate prupose}: ``Personal data can only be processed for specified explicit and legitimate pruposes and may not be processes futher in a way incompatible with those pruposes (art. 6b)''
 {[}\hyperref[references]{1}{]}.
 \item
-  \textbf{Proportionality}: ``Personal data may be processed only
-  insofar as it is adequate {[}\ldots{}{]} in relation to {[}its{]}
-  pruposes {[}\ldots{}{]}; every reasonable step must be taken to ensure
-  that data wich are inaccurate or incomplete {[}\ldots{}{]} are erased
-  or rectified (art. 6)'' {[}\hyperref[references]{1}{]}
-\end{itemize}
-
-(\textbf{Note:} The principles incorporated by the EU directive seem to
-closely map to \emph{the seven principles for user privacy control}.)
-
-Member states have to create independent \textbf{supervisory
-authorities} for personal data processing where controllers have to
-register before they start to process any data. The record has to be
-stored in public register {[}\hyperref[references]{1}{]}.
-
-Data transfer to thrid countries outside of the EU requires those
-countries to have a similar data protection level. Exceptions exists:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
+\textbf{Proportionality}: ``Personal data may be processed only insofar as it is adequate {[}\ldots{}{]} in relation to {[}its{]} pruposes {[}\ldots{}{]}; every reasonable step must be taken to ensure that data wich are inaccurate or incomplete {[}\ldots{}{]} are erased or rectified (art.
+6)'' {[}\hyperref[references]{1}{]}
+\end{itemize}
+
+(\textbf{Note:} The principles incorporated by the EU directive seem to closely map to \emph{the seven principles for user privacy control}.)
+
+Member states have to create independent \textbf{supervisory authorities} for personal data processing where controllers have to register before they start to process any data.
+The record has to be stored in public register {[}\hyperref[references]{1}{]}.
+
+Data transfer to thrid countries outside of the EU requires those countries to have a similar data protection level. Exceptions exists:
+
+\begin{itemize}
 \item
@@
 
-all between the US and the EU, although the US has no comparable law of
-data protection. The US Supreme Court regards privacy as
-\textbf{implicit} right granted with the First Amendment. However,
-european law states privacy as \textbf{explicit} right (ECHR Art. 8).
-Additionally the US government follows a doctrine which favors the
-economy to implement self-regulation {[}\hyperref[references]{1}{]}.
-
-\subsubsection{Implementation of the Data Protection Direvtive}
-
-\paragraph{Germany}
-
-Germany implements Directive 95/46 with the \emph{Federal Data
-Protection Act} from 2001 known as \emph{Bundesdatenschutzgesetz (BDSG)}
-{[}\hyperref[references]{1}{]}{[}\hyperref[references]{2}{]}{[}\hyperref[references]{6}{]}.
+all between the US and the EU, although the US has no comparable law of data protection. 
+The US Supreme Court regards privacy as \textbf{implicit} right granted with the First Amendment. However, european law states privacy as \textbf{explicit} right (ECHR Art. 8).
+Additionally the US government follows a doctrine which favors the economy to implement self-regulation {[}\hyperref[references]{1}{]}.
+
+\subsection{Implementation of the Data Protection Direvtive}
+
+\subsubsection*{Germany}
+
+Germany implements Directive 95/46 with the \emph{Federal Data Protection Act} from 2001 known as \emph{Bundesdatenschutzgesetz (BDSG)} {[}\hyperref[references]{1}{]}{[}\hyperref[references]{2}{]}{[}\hyperref[references]{6}{]}.
 However, Germany has violated the directive in two points:
@@
 \begin{enumerate}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  The BDSG has become effective three years too late, thus the EC filed
-  a treaty violation proceeding against Germany
-  {[}\hyperref[references]{6}{]}.
-\item
-  The BDSG does not implement \textbf{independet} supervisory
-  authorities. The \emph{Bundesdatenschutzbeauftragter} is subordinate
-  to the Ministiry of Interior. Although he is not subject to technical
-  oversight (\emph{Fachaufsicht}), he is subject to staff supervision by
-  the government (\emph{Rechtsaufsicht} durch die Bundesregierung und
-  \emph{Dienstaufsicht} durch das Innenministerium) and budget oversight
-  (including approval of employees) by the ministry
-  {[}\hyperref[references]{7}{]}. Thus the EC filed a treaty violation
-  proceeding against Germany, again {[}\hyperref[references]{6}{]}. In
-  March 2010 Germany was found guilty of violation of Directive 95/46 by
-  the ECJ {[}\hyperref[references]{6}{]}.
+\item
+The BDSG has become effective three years too late, thus the EC filed a treaty violation proceeding against Germany {[}\hyperref[references]{6}{]}.
+
+\item
+The BDSG does not implement \textbf{independet} supervisory authorities. 
+The \emph{Bundesdatenschutzbeauftragter} is subordinate to the Ministiry of Interior. 
+Although he is not subject to technical oversight (\emph{Fachaufsicht}), he is subject to staff supervision by the government (\emph{Rechtsaufsicht} durch die Bundesregierung und  \emph{Dienstaufsicht} durch das Innenministerium) and budget oversight (including approval of employees) by the ministry {[}\hyperref[references]{7}{]}.
+Thus the EC filed a treaty violation proceeding against Germany, again {[}\hyperref[references]{6}{]}. 
+In March 2010 Germany was found guilty of violation of Directive 95/46 by the ECJ {[}\hyperref[references]{6}{]}.
 \end{enumerate}
 
-(\textbf{Note:} Germany still violates point 2!)
-
-The states of Germany have their own implementation of Directive 95/46
-(\emph{Landesdatenschutzgesetze}). Federal public authorites are only
-bound to their federal law. {[}\hyperref[references]{2}{]}
+Germany still violates point 2!
+
+The states of Germany have their own implementation of Directive 95/46 (\emph{Landesdatenschutzgesetze}).
+Federal public authorites are only bound to their federal law. {[}\hyperref[references]{2}{]}
 
@@
 \begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
+
 \item
 \textbf{Roman Catholic Church:}
-  \href{http://de.wikipedia.org/wiki/Anordnung_\%C3\%BCber_den_kirchlichen_Datenschutz}{Anordnung
-  über den kirchlichen Datenschutz (KDO)}
-\item
-  \textbf{German Protestant Churches (Synode der Evangelischen Kirche in
-  Deutschland):}
-  \href{http://de.wikipedia.org/wiki/Datenschutzgesetz_der_Evangelischen_Kirche_in_Deutschland}{Kirchengesetz
-  über den Datenschutz der Evangelischen Kirche in Deutschland}
-\end{itemize}
-
-Data subjects have the right to\ldots{}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \ldots{} \textbf{disclosure} of
-\item
-  wether data about them is stored and processed
-\item
-  which data is stored and processed
-\item
-  the data sources
-\item
-  \ldots{} \textbf{correction} of false data
-\item
-  \ldots{} \textbf{file complaints} to the supervisory authority
-\item
-  \ldots{} have data \textbf{deleted} or \textbf{blocked}, but
-  controllers can prohibit deletion in favour of blocking
-\item
-  \ldots{} \textbf{Decline} third party access to the data
-\end{itemize}
-
-\paragraph{United Kingdom}
-
-The UK implements Directive 95/46 with the \emph{Data Protection Act
-1998 (DPA)}
-{[}\hyperref[references]{1}{]}{[}\hyperref[references]{3}{]}.
-
-The act is knwon for its high complexity: a manual record of phone
-numbers for business purposes could be hold subject to the DPA
-{[}\hyperref[references]{3}{]}. Although the act seems to fully cover
-the directive. Even higher restriction apply for \emph{``sensitive
-personal data''} (race, ethnicity, politics, religion, trade union
-status, health, sex life or criminal record), i.e. \textbf{consent} must
-be given freely and has to be explicit. {[}\hyperref[references]{3}{]}
-
-\emph{``The Act's definition of''personal data" covers any data that can
-be used to identify a living individual. Anonymised or aggregated data
-is not regulated by the Act, providing the anonymisation or aggregation
-has not been done in a reversible way. Individuals can be identified by
-various means including their name and address, telephone number or
-Email address. The Act applies only to data which is held, or intended
-to be held, on computers (`equipment operating automatically in response
-to instructions given for that purpose'), or held in a `relevant filing
-system'.}
-
-\emph{In some cases even a paper address book can be classified as a
-`relevant filing system', for example diaries used to support commercial
-activities such as a salesperson's diary.}
-
-\emph{The Freedom of Information Act 2000 modified the act for public
-bodies and authorities, and the Durant case modified the interpretation
-of the act by providing case law and precedent.}
-
-\emph{The Data Protection Act creates rights for those who have their
-data stored, and responsibilities for those who store, process or
-transmit such data. The person who has their data processed has the
-right to:}
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \emph{View the data an organisation holds on them. A `subject access
-  request' can be obtained for a nominal fee. As of January 2014, the
-  maximum fee is £2 for requests to credit reference agencies, £50 for
-  health and educational request, and £10 per individual otherwise,}
-\item
-  \emph{Request that incorrect information be corrected. If the company
-  ignores the request, a court can order the data to be corrected or
-  destroyed, and in some cases compensation can be awarded.}
-\item
-  \emph{Require that data is not used in any way that may potentially
-  cause damage or distress.}
-\item
-  *Require that their data is not used for direct marketing.``*
+\href{http://de.wikipedia.org/wiki/Anordnung_\%C3\%BCber_den_kirchlichen_Datenschutz}{Anordnung  über den kirchlichen Datenschutz (KDO)}
+
+\item
+\textbf{German Protestant Churches (Synode der Evangelischen Kirche in Deutschland):}
+\href{http://de.wikipedia.org/wiki/Datenschutzgesetz_der_Evangelischen_Kirche_in_Deutschland}{Kirchengesetz  über den Datenschutz der Evangelischen Kirche in Deutschland}
+\end{itemize}
+
+Data subjects have the following rights.
+
+\begin{itemize}
+\item
+Right to disclosure of wether data about them is stored and processed, which data is stored and processed and the data sources.
+\item
+Right to correction of false data.
+
+\item
+Right to file complaints to the supervisory authority.
+
+\item
+Right to have data deleted or blocked, but controllers can prohibit deletion in favour of blocking.
+
+\item
+Right to decline third party access to the data.
+\end{itemize}
+
+\subsubsection*{United Kingdom}
+
+The UK implements Directive 95/46 with the \emph{Data Protection Act 1998 (DPA)} {[}\hyperref[references]{1}{]}{[}\hyperref[references]{3}{]}.
+
+The act is knwon for its high complexity: a manual record of phone numbers for business purposes could be hold subject to the DPA {[}\hyperref[references]{3}{]}.
+Although the act seems to fully cover the directive.
+Even higher restriction apply for \emph{``sensitive personal data''} (race, ethnicity, politics, religion, trade union status, health, sex life or criminal record), i.e.
+\textbf{consent} must be given freely and has to be explicit.
 {[}\hyperref[references]{3}{]}
-\end{itemize}
-
-\paragraph{France}
-
-France implements Directive 95/46 with Law 2004-801 modifying law 78-17
-of 6.1.1978 (\emph{Loi n° 2004-801 du 6 août 2004} modifying \emph{loi
-n°78-17 relative à l'informatique, aux fichiers et aux libertés})
-{[}\hyperref[references]{1}{]}{[}\hyperref[references]{10}{]}{[}\hyperref[references]{11}{]}.
-
-\paragraph{United States of America}
-
-The USA do not implement the directive, nor is there any obligation for
-them to do so. However, companies subject to US jurisdiction can be
-certified to comply with the seven principles enforced by Directive
-95/46 (the seven OECD recommendations) {[}\hyperref[references]{5}{]}.
-Thus, those companies will act as \emph{safe harbors}. Without
-certification foreign companies are not allowed to send customer data
-back home {[}\hyperref[references]{9}{]}.
-
-(\textbf{Note:} \emph{Safe harbor} in general is the legal concept to
-regulate that a certain conduct will be deemed
-{[}\hyperref[references]{4}{]}, but in germany the term is commonly used
-as synonym for the agreement between the EU and the US regarding
-Directive 95/46.)
-
-\subsection{Privacy Definition and Taxonomy}
-
-\subsubsection{Defining Privacy}
-
-Defining privacy is a challenge which seems impossible. This is well put
-to words by Serge Gutwirth, who notes: \emph{``The notion of privacy
-remains out of the grasp of every academic chasing it. Even when it is
-cornered by such additional modifiers as `our' privacy, it still finds a
-way to remain elusive.''} {[}\hyperref[references]{2}{]}.
-
-Many privacy researchers seem to ``forcus on the ways in which privacy
-can be infrenged'' {[}\hyperref[references]{1}{]}. Thus they try to
-create taxonomies of \emph{privacy harms} instead of taxonomies of
-\emph{privacy types}. Those two differ in the respect that the former
-focuses on threats to prohibit whereas the latter focuses on values to
-protect. So one should rather evaluate what aspects are precious about
-privacy and develop measures to ensure their security than only forbid
-single actions against privacy {[}\hyperref[references]{1}{]}.
-
-\begin{center}
-We follow Fried in defining privacy as the ability to control data about personal information.
-\end{center}
+
+\emph{``The Act's definition of''personal data" covers any data that can be used to identify a living individual.
+Anonymised or aggregated data is not regulated by the Act, providing the anonymisation or aggregation has not been done in a reversible way.
+Individuals can be identified by various means including their name and address, telephone number or
+Email address.
+The Act applies only to data which is held, or intended to be held, on computers (`equipment operating automatically in response to instructions given for that purpose'), or held in a `relevant filing system'.}
+
+\emph{In some cases even a paper address book can be classified as a `relevant filing system', for example diaries used to support commercial activities such as a salesperson's diary.}
+
+\emph{The Freedom of Information Act 2000 modified the act for public bodies and authorities, and the Durant case modified the interpretation of the act by providing case law and precedent.}
+
+\emph{The Data Protection Act creates rights for those who have their data stored, and responsibilities for those who store, process or transmit such data.
+The person who has their data processed has the right to:}
+
+\begin{itemize}
+
+\item
+View the data an organisation holds on them.
+A `subject access request' can be obtained for a nominal fee.
+As of January 2014, the maximum fee is £2 for requests to credit reference agencies, £50 for health and educational request, and £10 per individual otherwise.
+
+\item
+Request that incorrect information be corrected.
+If the company ignores the request, a court can order the data to be corrected or destroyed, and in some cases compensation can be awarded.
+
+\item
+Require that data is not used in any way that may potentially cause damage or distress.
+
+\item
+Require that their data is not used for direct marketing. {[}\hyperref[references]{3}{]}
+\end{itemize}
+
+\subsubsection*{France}
+
+France implements Directive 95/46 with Law 2004-801 modifying law 78-17 of 6.1.1978 (\emph{Loi n° 2004-801 du 6 août 2004} modifying \emph{loi n°78-17 relative à l'informatique, aux fichiers et aux libertés}) {[}\hyperref[references]{1}{]}{[}\hyperref[references]{10}{]}{[}\hyperref[references]{11}{]}.
+
+\subsubsection*{United States of America}
+
+The USA do not implement the directive, nor is there any obligation for them to do so.
+However, companies subject to US jurisdiction can be certified to comply with the seven principles enforced by Directive 95/46 (the seven OECD recommendations) {[}\hyperref[references]{5}{]}.
+Thus, those companies will act as \emph{safe harbors}. Without certification foreign companies are not allowed to send customer data back home {[}\hyperref[references]{9}{]}.
+
+\emph{Safe harbor} in general is the legal concept to regulate that a certain conduct will be deemed {[}\hyperref[references]{4}{]}, but in germany the term is commonly used as synonym for the agreement between the EU and the US regarding Directive 95/46.)
+
+\section{Privacy Definition and Taxonomy}
+
+\subsection{Defining Privacy}
+
+Defining privacy is a challenge which seems impossible.
+This is well put to words by Serge Gutwirth, who notes: \emph{``The notion of privacy remains out of the grasp of every academic chasing it.
+Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.''} {[}\hyperref[references]{2}{]}.
+
+Many privacy researchers seem to ``forcus on the ways in which privacy can be infrenged'' {[}\hyperref[references]{1}{]}.
+Thus they try to create taxonomies of \emph{privacy harms} instead of taxonomies of \emph{privacy types}.
+Those two differ in the respect that the former focuses on threats to prohibit whereas the latter focuses on values to protect.
+So one should rather evaluate what aspects are precious about privacy and develop measures to ensure their security than only forbid single actions against privacy {[}\hyperref[references]{1}{]}.
+
+%% TODO %%
+
+* We follow Fried in defining privacy as the ability to control data about personal information.
 
@@
 
-\paragraph{The Seven Types of Privacy}
+\subsection{The Seven Types of Privacy}
 
@@
 \begin{enumerate}
+
 \item \textbf{Privacy of the Person.}
@@
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Body Characteristics
-\item
-  Weight
-\item
-  Height
-\item
-  Other body measures (shoulder width)
-\item
-  Fingerprints
-\item
-  DNA Sequence
-\item
-  Medical Conitions
-\item
-  Orthopedic conditions (e.g.~limping)
-\item
-  Having a cold
-\item
-  \ldots{}
-\end{itemize}
-\end{enumerate}
-
-\subparagraph{Privacy of Behaviour and Action}
-
-This type is also concerned with the ``protection against disclosure of
-personal matters'' {[}\hyperref[references]{1}{]} through behaviour,
-however Clarke's distinction between ``casual observation {[}\ldots{}{]}
-systematic recording and storage of information about those activities''
-{[}\hyperref[references]{1}{]}is lifted.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  regular visit at church, backery, doctors
-\item
-  sexual habits
-\item
-  political activities
-\end{itemize}
-
-\subparagraph{Privacy of Communication}
-
-It ``aims to avoid the interception of communications''
-{[}\hyperref[references]{1}{]} either electronic or face-to-face.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Briefgeheimnis (in german Legislation)
-\item
-  Email contents
-\item
-  Personal direct communication
-\item
-  Right to free discussion, i.e.~without third parties listening
-\end{itemize}
-
-\subparagraph{Privacy of Data and Image}
+Examples include:
+\begin{itemize}
+\item Body Characteristics: weight, height, sholder width, \ldots
+\item Biometric Properties: Fingerprints, DNA Sequence, \ldots
+\item Medical Conitions: Orthopedic conditions (e.g.~limping), Having a cold, \ldots
+\end{itemize}
+
+\item \textbf{Privacy of Behaviour and Action}
+
+This type is also concerned with the ``protection against disclosure of personal matters'' {[}\hyperref[references]{1}{]} through behaviour, however Clarke's distinction between ``casual observation {[}\ldots{}{]} systematic recording and storage of information about those activities'' {[}\hyperref[references]{1}{]}is lifted.
+
+Examples include:
+\begin{itemize}
+\item regular visit at church, backery, doctors
+\item sexual habits
+\item political activities
+\end{itemize}
+
+\item \textbf{Privacy of Communication}
+
+It ``aims to avoid the interception of communications'' {[}\hyperref[references]{1}{]} either electronic or face-to-face.
+
+Examples include:
+\begin{itemize}
+\item Secrecy of letters ('Briefgeheimnis' in german Legislation).
+\item Email contents
+\item Personal direct communication
+\item Right to free discussion, i.e.~without third parties listening
+\end{itemize}
+
+\item \textbf{Privacy of Data and Image}
 
@@
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Telefonnummer
-\item
-  IP Addresse
-\item
-  Public-administrative Data (Date of Birth, Melderegister)
-\item
-  Data held by organizations, like Banks or Insurance Companies
-\item
-  All data that is stored in online services (facebook)
-\end{itemize}
-
-\subparagraph{Privacy of Thoughts and Feelings}
+Examples include:
+\begin{itemize}
+\item Phone number
+\item IP address
+\item Public-administrative Data (Date of Birth, Melderegister)
+\item Data held by organizations, like Banks or Insurance Companies
+\item All data that is stored in online services (facebook)
+\end{itemize}
+
+\item \textbf{Privacy of Thoughts and Feelings}
 
@@
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Current feeling: Depression, tiredness, stressed, awake
-\item
-  Thoughts in general
-\end{itemize}
-
-\subparagraph{6 Privacy of Location and Space}
+Examples include:
+\begin{itemize}
+\item Current feeling: Depression, tiredness, stressed, awake
+\item Thoughts in general.
+\end{itemize}
+
+\item \textbf{Privacy of Location and Space}
 
 This type is the right ``to move about in public or semi-public space
-withoug being identified, tracked or monitored''
-{[}\hyperref[references]{1}{]}. Additionally this type is concerned with
-the protection of one's home and private places (``right to solitude''
-{[}\hyperref[references]{1}{]}).
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  GPS position tracking
-\item
-  Location of home address
+withoug being identified, tracked or monitored'' {[}\hyperref[references]{1}{]}.
+Additionally this type is concerned with the protection of one's home and private places (``right to solitude'' {[}\hyperref[references]{1}{]}).
+
+Examples include:
+\begin{itemize}
+\item GPS position tracking
+\item Location of home address
+\end{itemize}
+
+\item \textbf{Privacy of Association}
+
+This type is the right ``to associate with whomever {[}one{]} wish, withoug being monitored'' {[}\hyperref[references]{1}{]}.
+It is concerned with the protection against the automatic record of one's contacts.
+It does not imply that one is monitored because of the associations.
+
+Examples include:
+\begin{itemize}
+\item Friends
+\item Joining of orgnaizations (e.g.~political parties)
+\end{itemize}
+\end{enumerate}
+
+\subsection{Implicit Privacy Type Violation}
+
+Cf. Figure \ref{figure:Implicit Privacy Violation Matrix}
+
+%\includegraphics{../diagrams/png/implicit-privacy-violation-matrix.png}
+\input{./figures/Implicit-Privacy-Violation-Matrix}
+\input{./figures/ThreatTable}
+
+\subparagraph{Privacy of The Person}
+
+The Privacy of The Person is concerned with one's biometric privacy. If
+this type is violated, following implicit violations are possible:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \textbf{Privacy of The Person:} reflexive violation
+\item
+  \textbf{Privacy of Behaviour and Action:} none
+\item
+  \textbf{Privacy of Communication:} none
+\item
+  \textbf{Privacy of Data and Image:} none
+\item
+  \textbf{Privacy of Thoughts and Feelings:} Some psychological diseases
+  (e.g.~depression) have physiological impact. Such physiological
+  patterns could be detected.
+\item
+  \textbf{Privacy of Location and Space:} none
+\item
+  \textbf{Privacy of Association:} none
+\end{itemize}
+
+\subparagraph{Privacy of Behaviour and Action}
+
+The Privacy of Behaviour and Action is concerned with one's privacy
+regarding social activities (religious, political, sexual, \ldots{}). If
+this type is violated, following implicit violations are possible:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \textbf{Privacy of The Person:} Religious practices may include body
+  modifications (e.g.~circumcision).
+\item
+  \textbf{Privacy of Behaviour and Action:} reflexive violation
+\item
+  \textbf{Privacy of Communication:} none
+\item
+  \textbf{Privacy of Data and Image:} none
+\item
+  \textbf{Privacy of Thoughts and Feelings:} Social activities in
+  general depend on a certain intellectual attitude. Such an actvity is
+  the expressions of such an attitude.
+\item
+  \textbf{Privacy of Location and Space:} none
+\item
+  \textbf{Privacy of Association:} Recording religious, politcal or
+  sexual activities can reveal association with churches, political
+  parties or sexual partners.
+\end{itemize}
+
+\subparagraph{Privacy of Communication}
+
+The Privacy of Communication is concerned with not havin such
+communication (corresponence or vis-a-vis) intercepted. This is very
+broad type of privacy. Depending on the contens of the intercepted
+communication every other type can be violated:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \textbf{Privacy of The Person:} Communication about body
+  characteristics.
+\item
+  \textbf{Privacy of Behaviour and Action:} Communicatio about social
+  activities.
+\item
+  \textbf{Privacy of Communication:} reflexive violation
+\item
+  \textbf{Privacy of Data and Image:} Communication containing one's
+  passwords or other sensitive data.
+\item
+  \textbf{Privacy of Thoughts and Feelings:} Communication of thoughts
+  and feelings, e.g.~wiretapping a flirt or a catholic confession
+  ritual.
+\item
+  \textbf{Privacy of Location and Space:} Interception of face-to-face
+  communication is only possible if one's location and space is violated
+  (wiretapping).
+\item
+  \textbf{Privacy of Association:} Communication about one's
+  associations (family members, churches, etc.).
+\end{itemize}
+
+\subparagraph{Privacy of Data and Image}
+
+The Privacy of Data and Image is concered with one's data not beeing
+automatically availabel to others. This also is a very broad type of
+privacy. Depending of the data or image contents erery other type can be
+violated:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \textbf{Privacy of The Person:} Images or stored biometric information
+  reveal one's physical characteristics.
+\item
+  \textbf{Privacy of Behaviour and Action:} Images or diaries can reveal
+  one's social activities.
+\item
+  \textbf{Privacy of Communication:} Modern communication systems
+  usually contain some sort of archive function, e.g.~E-mail clients do
+  not automatically delete messages. Such messages are data and reveal
+  one's communication.
+\item
+  \textbf{Privacy of Data and Image:} reflexive violation
+\item
+  \textbf{Privacy of Thoughts and Feelings:} Images can show one's
+  emotional state.
+\item
+  \textbf{Privacy of Location and Space:} Images can reveal one's
+  location, e.g.~making a picture in front of the Eifel Tower.
+\item
+  \textbf{Privacy of Association:} E-mail data can also reveal
+  association.
+\end{itemize}
+
+\subparagraph{Privacy of Thoughts and Feelings}
+
+The Privacy of Thoughts and Feelings is concerned with keeping such
+thoughts and feelings secret. If this type is violated, following
+implicit violations are possible:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \textbf{Privacy of The Person:} Thoughts and feelings can reveal
+  medical conditions.
+\item
+  \textbf{Privacy of Behaviour and Action:} Thoughts and feelings can
+  reveal a certain attitudes which create a foundation for certain
+  social activities.
+\item
+  \textbf{Privacy of Communication:} none
+\item
+  \textbf{Privacy of Data and Image:} none
+\item
+  \textbf{Privacy of Thoughts and Feelings:} reflexive violation
+\item
+  \textbf{Privacy of Location and Space:} none
+\item
+  \textbf{Privacy of Association:} Thoughts and feelings can reveal
+  individual association, e.g amorous feelings for a certain person.
+\end{itemize}
+
+\subparagraph{Privacy of Location and Space}
+
+The Privacy of Location and Space is concerned with one's right to move
+freely without being tracked and one's right to private places. If this
+type is violated, following implicit violations are possible:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \textbf{Privacy of The Person:} Frequently visited doctors can reveal
+  certain medical conditions, if such doctors are known specialsts. In
+  general it could imply ill-being.
+\item
+  \textbf{Privacy of Behaviour and Action:} Frequently visited places in
+  general can reveal association and hence implies social activities.
+\item
+  \textbf{Privacy of Communication:} If one's location is knwon, it is
+  possible to intercept (wiretap) one's communication. This also may
+  violate the right to private spaces.
+\item
+  \textbf{Privacy of Data and Image:} If one's location is known, it is
+  possible shoot pictures. This violates the right to one's image
+  (\emph{``Recht am eigenen Bild''}).
+\item
+  \textbf{Privacy of Thoughts and Feelings:} Frequently visited persons
+  may imply certain thoughts and feelings, e.g.~having a mistress.
+\item
+  \textbf{Privacy of Location and Space:} reflexive violation
+\item
+  \textbf{Privacy of Association:} Frequently visted places can reveal
+  associations simply by searching in maps or yellow-pages.
 \end{itemize}
@@
 
-This type is the right ``to associate with whomever {[}one{]} wish,
-withoug being monitored'' {[}\hyperref[references]{1}{]}. It is
-concerned with the protection against the automatic record of one's
-contacts. It does not imply that one is monitored because of the
-associations.
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  Friends
-\item
-  Joining of orgnaizations (e.g.~political parties)
+The Privacy of Association is concerned with one's right to associate
+with whomever one wants, without that association having recorded.If
+this type is violated, following implicit violations are possible:
+
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+\item
+  \textbf{Privacy of The Person:} Association with tatoo artists could
+  imply having tatoos or other body modifications
+\item
+  \textbf{Privacy of Behaviour and Action:} Association with churches or
+  poltical organizations could imply certain activities.
+\item
+  \textbf{Privacy of Communication:} none
+\item
+  \textbf{Privacy of Data and Image:} none
+\item
+  \textbf{Privacy of Thoughts and Feelings:} Association with churchs of
+  political organizations could imply a certain intellectual attitude.
+\item
+  \textbf{Privacy of Location and Space:} none
+\item
+  \textbf{Privacy of Association:} reflexive violation
 \end{itemize}
@@
 \itemsep1pt\parskip0pt\parsep0pt
-\item
-  citizens want to be unidentifiable
-\item
-  citizens want certain informations about them to be secret
+
+\item Citizen want to exercise their legal right to privacy. I.e. their right to control information about themselves.
+- citizens want to be unidentifiable
+- citizens want certain informations about them to be secret
+\item Citizen wants to prevent abuse of their private data, and the related negative consequences.
 \item
@@
   \end{itemize}
-\end{itemize}
-
-\paragraph{Implicit Privacy Type Violation (Figure \ref{figure:Implicit Privacy Violation Matrix})}
-
-%\includegraphics{../diagrams/png/implicit-privacy-violation-matrix.png}
-\input{./figures/Implicit-Privacy-Violation-Matrix}
-
-\subparagraph{Privacy of The Person}
-
-The Privacy of The Person is concerned with one's biometric privacy. If
-this type is violated, following implicit violations are possible:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \textbf{Privacy of The Person:} reflexive violation
-\item
-  \textbf{Privacy of Behaviour and Action:} none
-\item
-  \textbf{Privacy of Communication:} none
-\item
-  \textbf{Privacy of Data and Image:} none
-\item
-  \textbf{Privacy of Thoughts and Feelings:} Some psychological diseases
-  (e.g.~depression) have physiological impact. Such physiological
-  patterns could be detected.
-\item
-  \textbf{Privacy of Location and Space:} none
-\item
-  \textbf{Privacy of Association:} none
-\end{itemize}
-
-\subparagraph{Privacy of Behaviour and Action}
-
-The Privacy of Behaviour and Action is concerned with one's privacy
-regarding social activities (religious, political, sexual, \ldots{}). If
-this type is violated, following implicit violations are possible:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \textbf{Privacy of The Person:} Religious practices may include body
-  modifications (e.g.~circumcision).
-\item
-  \textbf{Privacy of Behaviour and Action:} reflexive violation
-\item
-  \textbf{Privacy of Communication:} none
-\item
-  \textbf{Privacy of Data and Image:} none
-\item
-  \textbf{Privacy of Thoughts and Feelings:} Social activities in
-  general depend on a certain intellectual attitude. Such an actvity is
-  the expressions of such an attitude.
-\item
-  \textbf{Privacy of Location and Space:} none
-\item
-  \textbf{Privacy of Association:} Recording religious, politcal or
-  sexual activities can reveal association with churches, political
-  parties or sexual partners.
-\end{itemize}
-
-\subparagraph{Privacy of Communication}
-
-The Privacy of Communication is concerned with not havin such
-communication (corresponence or vis-a-vis) intercepted. This is very
-broad type of privacy. Depending on the contens of the intercepted
-communication every other type can be violated:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \textbf{Privacy of The Person:} Communication about body
-  characteristics.
-\item
-  \textbf{Privacy of Behaviour and Action:} Communicatio about social
-  activities.
-\item
-  \textbf{Privacy of Communication:} reflexive violation
-\item
-  \textbf{Privacy of Data and Image:} Communication containing one's
-  passwords or other sensitive data.
-\item
-  \textbf{Privacy of Thoughts and Feelings:} Communication of thoughts
-  and feelings, e.g.~wiretapping a flirt or a catholic confession
-  ritual.
-\item
-  \textbf{Privacy of Location and Space:} Interception of face-to-face
-  communication is only possible if one's location and space is violated
-  (wiretapping).
-\item
-  \textbf{Privacy of Association:} Communication about one's
-  associations (family members, churches, etc.).
-\end{itemize}
-
-\subparagraph{Privacy of Data and Image}
-
-The Privacy of Data and Image is concered with one's data not beeing
-automatically availabel to others. This also is a very broad type of
-privacy. Depending of the data or image contents erery other type can be
-violated:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \textbf{Privacy of The Person:} Images or stored biometric information
-  reveal one's physical characteristics.
-\item
-  \textbf{Privacy of Behaviour and Action:} Images or diaries can reveal
-  one's social activities.
-\item
-  \textbf{Privacy of Communication:} Modern communication systems
-  usually contain some sort of archive function, e.g.~E-mail clients do
-  not automatically delete messages. Such messages are data and reveal
-  one's communication.
-\item
-  \textbf{Privacy of Data and Image:} reflexive violation
-\item
-  \textbf{Privacy of Thoughts and Feelings:} Images can show one's
-  emotional state.
-\item
-  \textbf{Privacy of Location and Space:} Images can reveal one's
-  location, e.g.~making a picture in front of the Eifel Tower.
-\item
-  \textbf{Privacy of Association:} E-mail data can also reveal
-  association.
-\end{itemize}
-
-\subparagraph{Privacy of Thoughts and Feelings}
-
-The Privacy of Thoughts and Feelings is concerned with keeping such
-thoughts and feelings secret. If this type is violated, following
-implicit violations are possible:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \textbf{Privacy of The Person:} Thoughts and feelings can reveal
-  medical conditions.
-\item
-  \textbf{Privacy of Behaviour and Action:} Thoughts and feelings can
-  reveal a certain attitudes which create a foundation for certain
-  social activities.
-\item
-  \textbf{Privacy of Communication:} none
-\item
-  \textbf{Privacy of Data and Image:} none
-\item
-  \textbf{Privacy of Thoughts and Feelings:} reflexive violation
-\item
-  \textbf{Privacy of Location and Space:} none
-\item
-  \textbf{Privacy of Association:} Thoughts and feelings can reveal
-  individual association, e.g amorous feelings for a certain person.
-\end{itemize}
-
-\subparagraph{Privacy of Location and Space}
-
-The Privacy of Location and Space is concerned with one's right to move
-freely without being tracked and one's right to private places. If this
-type is violated, following implicit violations are possible:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \textbf{Privacy of The Person:} Frequently visited doctors can reveal
-  certain medical conditions, if such doctors are known specialsts. In
-  general it could imply ill-being.
-\item
-  \textbf{Privacy of Behaviour and Action:} Frequently visited places in
-  general can reveal association and hence implies social activities.
-\item
-  \textbf{Privacy of Communication:} If one's location is knwon, it is
-  possible to intercept (wiretap) one's communication. This also may
-  violate the right to private spaces.
-\item
-  \textbf{Privacy of Data and Image:} If one's location is known, it is
-  possible shoot pictures. This violates the right to one's image
-  (\emph{``Recht am eigenen Bild''}).
-\item
-  \textbf{Privacy of Thoughts and Feelings:} Frequently visited persons
-  may imply certain thoughts and feelings, e.g.~having a mistress.
-\item
-  \textbf{Privacy of Location and Space:} reflexive violation
-\item
-  \textbf{Privacy of Association:} Frequently visted places can reveal
-  associations simply by searching in maps or yellow-pages.
-\end{itemize}
-
-\subparagraph{Privacy of Association}
-
-The Privacy of Association is concerned with one's right to associate
-with whomever one wants, without that association having recorded.If
-this type is violated, following implicit violations are possible:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  \textbf{Privacy of The Person:} Association with tatoo artists could
-  imply having tatoos or other body modifications
-\item
-  \textbf{Privacy of Behaviour and Action:} Association with churches or
-  poltical organizations could imply certain activities.
-\item
-  \textbf{Privacy of Communication:} none
-\item
-  \textbf{Privacy of Data and Image:} none
-\item
-  \textbf{Privacy of Thoughts and Feelings:} Association with churchs of
-  political organizations could imply a certain intellectual attitude.
-\item
-  \textbf{Privacy of Location and Space:} none
-\item
-  \textbf{Privacy of Association:} reflexive violation
 \end{itemize}

### 2014-09-15 figures/ThreatTable.tex
Partial cleanup up formatting and arrangement of sections.

@@
 \\\textbf{Criminal vs Maintenance Staff}
-\\Staff members have a professional ethos and a duty to provide working systems. Successful attacks offend the former and obstruct the latter.}
+\\Staff members have a professional ethos and a duty to provide working systems. Successful attacks offend the former and obstruct the latter.
+\\Staff member employed by L+G or contractor. The Business operation is thretened by loss of confidential data.
+}
 {Unsecured Hardware- or Software-Interfaces}

### 2014-09-15 figures/IT-Systems.tex

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{diagrams/png/it-systems.png}

\begin{flushleft}
\scriptsize
\textbf{Legend:}
\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt

\item
\textbf{(a) Citizen:} 
User of the L+G client application whose privacy is at stake.

\item
\textbf{(b) System Provider:} Provides technical infrastructure.

\item
\textbf{(c) Local Authority:} Provider of the L+G system.

\item
\textbf{(d) Mobile Device:}
Runs the L+G client application, produces and stores data sensitive to the users privacy.

\item
\textbf{(e) Data Center:} Runs the L+G services, processes and stores user data.

\item
\textbf{(f) Report Tool:} Interface to aggregated user data.

\end{itemize}
\end{flushleft}

\caption{IT Systems}
\label{figure:IT Systems}
\end{figure}

### 2014-09-15 PrivacyProtection-2014-09-09.tex
Further restructuring and cleanup.

@@
-\chapter{Privacy Protection}
+\chapter{Legal and Ethical aspects of Privacy}
 
@@
 \input{./figures/Implicit-Privacy-Violation-Matrix}
-\input{./figures/ThreatTable}
-
-\subparagraph{Privacy of The Person}
+
+\paragraph*{Privacy of The Person}
 
@@
 \begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
+
 \item
@@
 
-\subparagraph{Privacy of Behaviour and Action}
+\paragraph*{Privacy of Behaviour and Action}
 
@@
 
-\subparagraph{Privacy of Communication}
+\paragraph*{Privacy of Communication}
 
@@
 
-\subparagraph{Privacy of Data and Image}
+\paragraph*{Privacy of Data and Image}
 
@@
 
-\subparagraph{Privacy of Thoughts and Feelings}
+\paragraph*{Privacy of Thoughts and Feelings}
 
@@
 
-\subparagraph{Privacy of Location and Space}
+\paragraph*{Privacy of Location and Space}
 
@@
 
-\subparagraph{Privacy of Association}
+\paragraph*{Privacy of Association}
 
@@
 
-\section{IT Security}
+\pagebreak
+
+\chapter{Privacy Analysis of Live+Gov Systems}
+
+\section{IT Security Analysis according to Grimm et. al}
+
+Grimm et al. create a reference model for conducting IT security
+analyses consisting of:
+
+\begin{itemize}
+
+\item
+  a \textbf{model}, which aims to organize common security
+  terminology in a reasonable and practical way
+\item
+  and a systematic analysis \textbf{procedure} based on that ontology
+\end{itemize}
 
@@
 
-\subsubsection{threat}
-
-\emph{threat} is a class of potential events whose manifestation can
+\begin{itemize}
+
+\item A \emph{threat} is a class of potential events whose manifestation can
 cause damage or harm.
 
-\subsubsection{vulnerability}
-
-A weakness which can cause the loss of security.
-
-\subsubsection{hazard}
-
-Concrete risk due to one or more concrete vulnerabilities \textbf{AND}
-corresponding threats.
+\item A \emph{vulnerability} is a weakness which can cause the loss of security.
+
+\item a \emph{hazard} is a concrete risk due to one or more concrete vulnerabilities and corresponding threats.
 
@@
 
-\subsubsection{incident}
-
-An event caused by a threat manifestation against one or more
+\item An \emph{incident} is an event caused by a threat manifestation against one or more
 corresponding vulnerabilities.
@@
 
-\subsubsection{accident}
-
-An incident which did cause damage or harm. (Translates to
-\emph{Schandensfall/Zwischenfall/Störfall} in german)
-
-\subsubsection{non-accident}
-
-An incident which \textbf{did not} cause any damage or harm.
-
-\subsubsection{measure}
-
-A well defined set of one or more activities which reduce or eliminate
+\item An \emph{accident} is an incident which did cause damage or harm. (Translates to 'Schandensfall/Zwischenfall/Störfall' in the German language).
+
+\item A \emph {non-accident} is an incident which did not cause any damage or harm.
+
+\item A security \emph{measure} is a set of one or more activities which reduce or eliminate
 vulnerabilities.
 
-\subsection{IT Security Analysis according to Grimm et. al}
-
-Grimm et al. create a reference model for conducting IT security
-analyses consisting of:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-\item
-  an \textbf{ontology}, which aims to organize common security
-  terminology in a reasonable and practical way
-\item
-  and a systematic analysis \textbf{procedure} based on that ontology
-\end{itemize}
-
-\subsubsection{Ontology}
+\end{itemize}
+
+\subsection{Model}
 
@@
 
-The ontology is organized in four views:
-
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
+The model is organized in four views:
+
+\begin{itemize}
+
 \item
@@
 \input{./figures/LG-Operation-Scenario-01-Server-Side-Mining}
+\input{./figures/IT-Systems}
 
@@
 \begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
 
@@
 
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
+\input{./figures/ThreatTable}
+
+\begin{itemize}
+
 \item
@@
 
-\section{References}
+% \section{References}

### 2014-09-15 PrivacyProtection-2014-09-09.tex
refactored the it-systems section

--- 2014-09-15 PrivacyProtection-2014-09-09.tex
@@
 
-\subsubsection{IT-Systems \& Humans}
-
-\paragraph{Server Side Mining Scenario (Figure \ref{figure:Live+Gov Operation Scenario 01 - Server Side Mining})}
-
-%\includegraphics{../diagrams/png/scenario01-ServerSideMining.png}
-\input{./figures/LG-Operation-Scenario-01-Server-Side-Mining}
-
-A citizen (a) carries a mobile device (b) running the L+G client. The
-mobile device collects sensor data and sends it to the L+G Data Cetner
-(c). The L+G Data Center also sends beneficial information for citizens
-to the mobile device.
-
-The L+G Maintenance Crew (d) maintaince the L+G Data Center.
-
-The Aggregation Monitor (e) queries the L+G Data Center for aggregated
-data. Local Authorities (f) use the Aggregation Monitor to get
-information in order to improve public services. \textbf{In this
+\subsubsection{IT-Systems}
+
+This section outlines the general architecture (Figure \ref{figure:IT Systems}) of IT systems for public monitoring comparable to the Live+Gov project.
+
+\input{./figures/IT-Systems}
+
+\textbf{Description:}
+A \textit{Citizen (a)} carries a mobile device running the \textit{Sensor Collector (d)} application.
+The \textit{Sensor Collector} application collects sensor data (accelerometer, GPS, GSM, ...) and sends it to the \textit{Data Cetner (c)}. 
+The \textit{Data Center} on the other side sends beneficial information (traffic jam reports, bus schedule) for \textit{Citizens} to the mobile device.
+The \textit{System Provider (b)} provides and operates technical infrastructure like the Data Center and the \textit{Report Tool (f)}.
+The \textit{Report Tool} queries the \textit{Data Center} for aggregated data to visualize in form of charts and other means suitable to help understanding of monitored \textit{Citizens}. 
+\textit{Local Authorities (c)} use the \textit{Report Tool} to get information in order to understand citizen movement and improve public services.
+
+In such systems, the most valuable information is not the raw sensor data, the beneficial information for \textit{Local Authorities} is produced via \textit{Data Mining}.
+
+\textbf{In this
 scenario the mobile the server conducts data mining on aggregated user
@@
 
-\paragraph{Mobile Mining Scenario (Figure \ref{figure:Live+Gov Operation Scenario 02 - Mobile Mining})}
-
-%\includegraphics{../diagrams/png/scenario02-MobileMining.png}
-\input{./figures/LG-Operation-Scenario-02-Mobile-Mining}
-
-A citizen (a) carries a mobile device (b) running the L+G client. The
-mobile device collects sensor data and sends it to the L+G Data Cetner
-(c). The L+G Data Center also sends beneficial information for citizens
-to the mobile device. \textbf{In this scenario the mobile device also
-conducts data mining previously to sending the results to the L+G Data
-Center.}
-
-The L+G Maintenance Crew (d) maintaince the L+G Data Center.
-
-The Aggregation Monitor (e) queries the L+G Data Center for aggregated
-data. Local Authorities (f) use the Aggregation Monitor to get
-information in order to improve public services.
-
-When no connection to the L+G servers is available, the mining
-end-products are stored on the mobile device.
+\subsubsection{Humans}
 

### 2014-09-15 PrivacyProtection-2014-09-09.tex

--- 2014-09-15 PrivacyProtection-2014-09-09.tex
@@
 
-\subsubsection{IT-Systems \& Humans}
-
-\paragraph{Server Side Mining Scenario (Figure \ref{figure:Live+Gov Operation Scenario 01 - Server Side Mining})}
-
-%\includegraphics{../diagrams/png/scenario01-ServerSideMining.png}
-\input{./figures/LG-Operation-Scenario-01-Server-Side-Mining}
+\subsubsection{IT-Systems}
+
+This section outlines the general architecture (Figure \ref{figure:IT Systems}) of IT systems for public monitoring comparable to the Live+Gov project.
+
 \input{./figures/IT-Systems}
 
-A citizen (a) carries a mobile device (b) running the L+G client. The
-mobile device collects sensor data and sends it to the L+G Data Cetner
-(c). The L+G Data Center also sends beneficial information for citizens
-to the mobile device.
-
-The L+G Maintenance Crew (d) maintaince the L+G Data Center.
-
-The Aggregation Monitor (e) queries the L+G Data Center for aggregated
-data. Local Authorities (f) use the Aggregation Monitor to get
-information in order to improve public services. \textbf{In this
+\textbf{Description:}
+A \textit{Citizen (a)} carries a mobile device running the \textit{Sensor Collector (d)} application.
+The \textit{Sensor Collector} application collects sensor data (accelerometer, GPS, GSM, ...) and sends it to the \textit{Data Cetner (c)}. 
+The \textit{Data Center} on the other side sends beneficial information (traffic jam reports, bus schedule) for \textit{Citizens} to the mobile device.
+The \textit{System Provider (b)} provides and operates technical infrastructure like the Data Center and the \textit{Report Tool (f)}.
+The \textit{Report Tool} queries the \textit{Data Center} for aggregated data to visualize in form of charts and other means suitable to help understanding of monitored \textit{Citizens}. 
+\textit{Local Authorities (c)} use the \textit{Report Tool} to get information in order to understand citizen movement and improve public services.
+
+In such systems, the most valuable information is not the raw sensor data, the beneficial information for \textit{Local Authorities} is produced via \textit{Data Mining}.
+
+\textbf{In this
 scenario the mobile the server conducts data mining on aggregated user

### 2014-09-15 PrivacyProtection-2014-09-09.tex
Adding structure
finished description of it-systems

@@
 
-\textbf{Description:}
 A \textit{Citizen (a)} carries a mobile device running the \textit{Sensor Collector (d)} application.
@@
 
-In such systems, the most valuable information is not the raw sensor data, the beneficial information for \textit{Local Authorities} is produced via \textit{Data Mining}.
-
-\textbf{In this
-scenario the mobile the server conducts data mining on aggregated user
-data and additionally trys to combine it with publicly accessible data
-to enhance the results.}
-
-When no connection to the L+G servers is available, the mining
-end-products are stored on the mobile device.
+In such systems, the most valuable information is not the raw sensor data, the beneficial information for \textit{Local Authorities} is produced via \textit{Data Mining} (Human Activity Recognition: detect running, walking, standing, ...).
+Usually, \textit{Data Mining} is heavily conduct within the \textit{Data Center}, but simpler mining processes can also be conducted on mobile devices.
 

### 2014-09-15 PrivacyProtection-2014-09-09.tex
finished interest description of citizens

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\pagebreak
+
 \section{Live+Gov Privacy Protection Analysis}
@@
 \subsubsection{IT-Systems}
+\label{subsubsection:it-systems}
 
@@
 \subsubsection{Humans}
+\label{subsubsection:humans}
+
+This section outlines the human actors previously introduced within the IT system architecture (\ref{subsubsection:it-systems}).
+It is described which individual interests they have invested in monitoring systems comparable to the Live+Gov system. 
+Also the additional actor \textit{Criminal} is introduced as an possible obstacle.
+
+\textbf{Citizen:}
+Citizens who take part in such urban monitoring systems share privacy sensitive data about themselves. 
+They do it out of freewill and with consent to certain agreements on secrecy and proportionality.
+So naturally they are interested in not having private information about them revealed to unknown third parties.
+This would interfere with their ability to control such information, thus they cannot exercise privacy (according to Fried) anymore.
+
+Beyond that, they share information with trust in the legitimate and proportional use of that information. 
+Citizens expect their data is only used for public improvement within the agreed bounds.
+They do not expect that their data is used for illegitimate purposes like scoring procedures, which could have a possible negative effect for them.
+
+In general, Citizens are not interested in negative results for themselves.
+This for financial loss. 
+Local Authorities pay System Providers for their services.
+Typically Local Authorities capitalize themselves via taxes.
+So Citizens also have financial interests invested in the monitoring system.
+
+In short, Citizens are interested in ...
+\begin{itemize}
+\item ... confidential use of privacy sensitive data
+\item ... legitimate use of privacy sensitive data
+\item ... proportional use of privacy sensitive data
+\item ... successful use of privacy sensitive data
+\end{itemize}
+
+\textbf{System Provider:} Lorem Ipsum
+
+\textbf{Local Authority:} Lorem Ipsum
+
+\textbf{Criminal:} Lorem Ipsum
 

### 2014-09-15 PrivacyProtection-2014-09-09.tex
Refactoring Reference Model

--- 2014-09-15 PrivacyProtection-2014-09-09.tex
@@
 \chapter{Legal and Ethical aspects of Privacy}
-
-The goal of this document to analyze and identify the threads to personal privacy that are posed by collecting, storing and processing sensor data from mobile phones.
-We derive concrete privacy protection measures that address the main risks involved with handling such data.
-In the last part we describe how these measures are implemented in the Live+Gov toolkit.
 
@@
 As background information we include an overview about historical treatments of privacy as well as legal regulation of privacy in the European Union.
-Based on this we propose a definition of privacy as {\em control over personal data}, and introduce a taxonomy of privacy attributes that give specify the term {\em personal data} in the context of mobile
-sensor data collection.
-
-The complexity of our systems and the variety of threads make a great number of counter measures plausible.
-We approach this complexity with the aid of a general security analysis model developed in [Grimm].
-We give a brief introduction to this model and perform a IT Security Analysis with respect to the privacy asset for our system.
+Based on this we propose a definition of privacy as {\em control over personal data}, and introduce a taxonomy of privacy attributes that give specify the term {\em personal data} in the context of mobile sensor data collection.
 
@@
 
+The goal of this chapter to analyze and identify the threads to personal privacy that are posed by collecting, storing and processing sensor data from mobile phones.
+We derive concrete privacy protection measures that address the main risks involved with handling such data.
+
+The complexity of our systems and the variety of threads make a great number of counter measures plausible.
+We approach this complexity with the aid of a general security analysis model developed in \cite{Grimm}.
+We give a brief introduction to this model and perform a IT Security Analysis with respect to the privacy asset for our system.
+
 \section{IT Security Analysis according to Grimm et. al}
 
-Grimm et al. create a reference model for conducting IT security
-analyses consisting of:
-
-\begin{itemize}
-
-\item
-  a \textbf{model}, which aims to organize common security
-  terminology in a reasonable and practical way
-\item
-  and a systematic analysis \textbf{procedure} based on that ontology
-\end{itemize}
-
-\subsection{Basic Terminology}
+We follow the Reference Model for IT Security Analyis as described in \cite{Grimm}.
+It supersedes earlier efforts by \cite{Avizienis}.
+
+The \emph{reference model} consists of a \emph{model}, which aims to organize common security terminology in a reasonable and practical way and a systematic analysis \emph{procedure} based on that model.
+
+In this section we give a brief overview over the reference model.
+Before we can start with the description of model and procedure we introduce the following basic terminology (cf. Figure \ref{fig:terminology}).
 
@@
 
-\paragraph{Views}
-
-The model is organized in four views:
-
-\begin{itemize}
-
-\item
-  \textbf{World:} Contains all components describing the current state.
-\item
-  \textbf{Potential:} Contains all components describing both the
-  desired and dreaded state.
-\item
-  \textbf{Plan:} Contains all conceptual components required to develop
-  sufficient measures to achieve and secure the desired state
-  \textbf{or} to create the dreaded state. The \textbf{Plan} view
-  \emph{realizes} the \textbf{Potential} view of the system.
-  (\textbf{Note:} this should rather be translated with \emph{Concept})
-\item
-  \textbf{Event:} Contains all actual operations and events through out
-  the production phase. The \textbf{Event} view \emph{instantiates} the
-  \textbf{Plan} view of the system.
-\end{itemize}
-
-\paragraph{Components}
-
-\subparagraph{World}
-
-\begin{itemize}
-
-\item
-  \textbf{Assets:} Things of value to one or more stakeholders. The
-  value can be \emph{``hard''} (money, data, etc.) or \emph{``soft''}
-  (trust,privacy,etc.).
-\item
-  \textbf{IT-Systems:} The IT-Systems under study.
-\item
-  \textbf{Humans:} All identifiable actors of the system under study.
-  Many problems arise due to misunderstandings during man-machine
-  interaction.
-\item
-  \textbf{Conflicts of Interests:} Different actors have different
-  interests. Those interest can be in conflict. A trival conflict is the
-  \emph{Attacker-Attackee}-Conflict: a service provider offers private
-  data storage, therefore the provider is interested in having the
-  access restricted. An attacker is naturally interested in easy access
-  in order captialize the stolen data. However, heavy security
-  restriction are also a burden for maintenance staff, as they are
-  interested in having an easy life. This kind of conflict can create
-  vulnerabilities.
-\item
-  \textbf{Vulnerabilities:} All indentifiable weaknesses in the current
-  system layout.
-\item
-  \textbf{Interactions:} Assets, IT-Systems, Humans and Vulnerabilites
-  are in continous interaction with each other in order to
-  \emph{support} the \textbf{Business Objectives}. Those interactionscan
-  also \emph{lead to} \textbf{Threats}, i.e.~having unencrypted
-  communication with a server. Therefore all interactions which occur in
-  the system's outline have to be documented.
-\end{itemize}
-
-\subparagraph{Potential}
-
-\begin{itemize}
-
-\item
-  \textbf{Chances/Risks:} Chances and Risks \emph{affect}
-  \textbf{Business Objectives} (and \textbf{Threats}).
-\item
-  \textbf{Business Objectives:} Chances the system under study. This
-  describes the desired state.
-\item
-  \textbf{Threats:} Risks for business objectives, this can simply be
-  the integrety of the IT-System. Threats \emph{target} one or more
-  \textbf{Assets} and violate \textbf{Security Requirements}. This
-  describes the dreaded state.
-\item
-  \textbf{Security Requirements:} A set of distinct requirements for a
-  safe and secure system. Security requirements \emph{hold for} one or
-  more \textbf{Assets}. This describes the potential state after this
-  procedure.
-\end{itemize}
-
-\subparagraph{Plan}
-
-\begin{itemize}
-
-\item
-  \textbf{Business Model:} The plan to achieve \textbf{Business
-  Objectives}
-\item
-  \textbf{Accident Scenario:} This is either the plan of a certain
-  attack or the concrete outline of random disaster. Although the latter
-  is most likely hard to anticipate.
-\item
-  \textbf{Attack Technique:} A specific technique or technology to
-  attack IT-Systems (Man in the Middel, Phishing, etc.). (\textbf{Note:}
-  This component is not conform to the glossary as the term
-  \emph{accident} there covers both disaster and attack!)
-\item
-  \textbf{Security Measures:} The plan to achieve a safe and secure
-  system. Each \textbf{Security Mesaure} \emph{opposes} an
-  \textbf{Accident Scenario} or an \textbf{Attack Technique}.
-\end{itemize}
-
-\subparagraph{Event}
-
-\begin{itemize}
-
+The model is depictured in Figure \ref{fig:refmodel}.
+It is organized in four views (round boexes) that contain a number of components (rectangular boxes).
+
+The \textbf{world view} contains all components describing the current state.
+It consists of the following components:
+\begin{itemize}
+\item
+\textbf{Assets.}
+Things of value to one or more stakeholders.
+The value can be ``hard'' (money, data, etc.) or ``soft'' (trust,privacy,etc.).
+
+\item \textbf{IT-Systems.}
+The IT-Systems under study.
+
+\item \textbf{Humans.}
+All identifiable actors of the system under study.
+Many problems arise due to misunderstandings during man-machine interaction.
+
+\item \textbf{Conflicts of Interests.}
+Different actors have different interests.
+Those interest can be in conflict.
+A trival conflict is the \emph{Attacker-Attackee-Conflict}:
+A service provider offers private data storage, therefore the provider is interested in having the access restricted.
+An attacker is naturally interested in easy access in order captialize the stolen data.
+However, heavy security restriction are also a burden for the service operator, as they are interested in efficient work flows.
+This kind of conflict can create vulnerabilities.
+
+\item \textbf{Vulnerabilities.}
+All indentifiable weaknesses in the current system layout.
+
+\item \textbf{Interactions.}
+Assets, IT-Systems, Humans and Vulnerabilites are in continous interaction with each other in order to support the business objectives.
+Those interactionscan also lead to threats, e.g. having unencrypted communication with a server.
+Therefore all interactions which occur in the system's outline have to be documented.
+\end{itemize}
+
+The \textbf{potential view} describes both the desired and dreaded state.
+It consists of the following components.
+\begin{itemize}
+\item \textbf{Chances/Risk.}
+Chances and Risks affect business objectives and threats.
+
+\item \textbf{Business Objectives.}
+Chances the system under study.
+This describes the desired state.
+
+\item \textbf{Threats.}
+Risks for business objectives, this can simply be the integrety of the IT-System.
+Threats target one or more assets and violate security requirements}.
+This describes the dreaded state.
+
+\item
+\textbf{Security Requirements:} A set of distinct requirements for a safe and secure system.
+Security requirements hold for one or more assets.
+This describes the potential state after this procedure.
+\end{itemize}
+
+The \textbf{concept view} is a realization of the the potential view of the system.
+It describes a plan of sufficient measures to secure the desired state and to aviod the dreaded state of the system.
+It contians the following components:
+\begin{itemize}
+\item \textbf{Business Model.}
+The plan to achieve business objectives.
+
+\item \textbf{Accident Scenario.}
+This is either the plan of a certain attack or the concrete outline of random disaster.
+Although the latter is most likely hard to anticipate.
+
+\item \textbf{Attack Technique.}
+A specific technique or technology to attack IT-Systems (Man in the Middel, Phishing, etc.).
+(\textbf{Note:} This component is not conform to the glossary as the term \emph{accident} there covers both disaster and attack!)
+
+\item \textbf{Security Measures.}
+The plan to achieve a safe and secure system.
+Each security mesaure opposes an accident scenario or an attack technique.
+
+\end{itemize}
+
+The \textbf{event view} contains all actual operations and events through out the production phase.
+The event view instantiates the concept view of the system.
+It contains the following components:
+\begin{itemize}
 \item
@@
 
-\subsubsection{Procedure}
+\subsection{Procedure}
 
@@
 through out the analysis.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\pagebreak
 

### 2014-09-15 PrivacyProtection-2014-09-09.tex
modified actors description

--- 2014-09-15 PrivacyProtection-2014-09-09.tex
@@
 
-\subsubsection{Humans}
+\subsubsection{Actors}
 \label{subsubsection:humans}
@@
 \textbf{Citizen:}
-Citizens who take part in such urban monitoring systems share privacy sensitive data about themselves. 
-They do it out of freewill and with consent to certain agreements on secrecy and proportionality.
-So naturally they are interested in not having private information about them revealed to unknown third parties.
-This would interfere with their ability to control such information, thus they cannot exercise privacy (according to Fried) anymore.
-
-Beyond that, they share information with trust in the legitimate and proportional use of that information. 
-Citizens expect their data is only used for public improvement within the agreed bounds.
-They do not expect that their data is used for illegitimate purposes like scoring procedures, which could have a possible negative effect for them.
-
-In general, Citizens are not interested in negative results for themselves.
-This for financial loss. 
-Local Authorities pay System Providers for their services.
-Typically Local Authorities capitalize themselves via taxes.
-So Citizens also have financial interests invested in the monitoring system.
+\begin{itemize}
+\item legal right to privacy
+\item no spam
+\item no stalking
+\item no loss / gain of money
+\item physical wellbeing / personal health (attack because someone knows your gps location)
+\item transparency: who has my data
+\item good pricing of insurance policies
+\end{itemize}
 
@@
 
-\textbf{System Provider:} Lorem Ipsum
-
-\textbf{Local Authority:} Lorem Ipsum
-
-\textbf{Criminal:} Lorem Ipsum
+\textbf{Criminal:}
+Criminals are persons who do not have legal access to the IT systems of a proposed monitoring system.
+Usually, monetary interests lead their activities.
+For example they want to obtain access to critical systems to steal sensitive data or to get the system under their control.
+Controlled systems could be leased as part of a bot net. 
+Stolen data could simply be sold as is or used for illegitimate purposes, e.g. spam or phishing attacks - or excessive data mining.
+
+Because Local Authorities are involved in the general outline of a public monitoring system, the possibility for politically motivated attacks cannot be ignored. 
+Criminals could want to harm or destroy the systems in order to damage the reputation of Local Authorities (politicians or other officials) or to make a political statement of their own, e.g.
+"Stop Surveillance!".
+
+Another possible motivation for criminal activities could be social appreciation.
+A hacker could attack critical infrastructure just to show his skills.
+
+In short, Criminals are interested in ...
+\begin{itemize}
+\item ... financial profit
+\item ... high social standing
+\item ... political power
+\end{itemize}
+
+\textbf{System Provider:}
+System Providers provide and operate the technical infrastructure (hardware and software) of the proposed monitoring system.
+They are normal companies which offer to provide a service, so different roles of employees are involved:
+an administrator, who maintains and operates the running system;
+a programmer/developer, who develops the system;
+a support manager, who handles customer relations;
+...
+
+As companies, they follow typical business objectives. 
+They are interested in gaining financial profit, which heavily depends on the customer's satisfaction.
+But it also depends on the happiness of their employees. 
+Employees have a claim to professional excellence.
+They want to deliver a good job within their means.
+This effects product quality and thus effects customer satisfaction.
+If employees cannot satisfy their demand for professional excellence, they might get discontent, deliver poor work or produce higher costs through sick days.
+The worst case scenario could be, that a discontent employee gets angry and steals data or harms the running systems.
+
+In short, System Providers are interested in ...
+\begin{itemize}
+\item ... financial profit
+\item ... good working conditions
+\item ... professional excellence
+\end{itemize}
+
+\textbf{Local Authority:}
+
+In short, Local Authorities are interested in ...
+\begin{itemize}
+\item ... business intelligence (profit, RoI)
+\item ... political reputation
+\end{itemize}
 

### 2014-09-15 PrivacyProtection-2014-09-09.tex

--- 2014-09-15 PrivacyProtection-2014-09-09.tex
@@
 
-\subsubsection{Humans}
+\subsubsection{Actors}
+\label{subsubsection:humans}
+
+This section outlines the human actors previously introduced within the IT system architecture (\ref{subsubsection:it-systems}).
+It is described which individual interests they have invested in monitoring systems comparable to the Live+Gov system. 
+Also the additional actor \textit{Criminal} is introduced as an possible obstacle.
+
+\textbf{Citizen:}
+\begin{itemize}
+\item legal right to privacy
+\item no spam
+\item no stalking
+\item no loss / gain of money
+\item physical wellbeing / personal health (attack because someone knows your gps location)
+\item transparency: who has my data
+\item good pricing of insurance policies
+\end{itemize}
+
+In short, Citizens are interested in ...
+\begin{itemize}
+\item ... confidential use of privacy sensitive data
+\item ... legitimate use of privacy sensitive data
+\item ... proportional use of privacy sensitive data
+\item ... successful use of privacy sensitive data
+\end{itemize}
+
+\textbf{Criminal:}
+Criminals are persons who do not have legal access to the IT systems of a proposed monitoring system.
+Usually, monetary interests lead their activities.
+For example they want to obtain access to critical systems to steal sensitive data or to get the system under their control.
+Controlled systems could be leased as part of a bot net. 
+Stolen data could simply be sold as is or used for illegitimate purposes, e.g. spam or phishing attacks - or excessive data mining.
+
+Because Local Authorities are involved in the general outline of a public monitoring system, the possibility for politically motivated attacks cannot be ignored. 
+Criminals could want to harm or destroy the systems in order to damage the reputation of Local Authorities (politicians or other officials) or to make a political statement of their own, e.g.
+"Stop Surveillance!".
+
+Another possible motivation for criminal activities could be social appreciation.
+A hacker could attack critical infrastructure just to show his skills.
+
+In short, Criminals are interested in ...
+\begin{itemize}
+\item ... financial profit
+\item ... high social standing
+\item ... political power
+\end{itemize}
+
+\textbf{System Provider:}
+System Providers provide and operate the technical infrastructure (hardware and software) of the proposed monitoring system.
+They are normal companies which offer to provide a service, so different roles of employees are involved:
+an administrator, who maintains and operates the running system;
+a programmer/developer, who develops the system;
+a support manager, who handles customer relations;
+...
+
+As companies, they follow typical business objectives. 
+They are interested in gaining financial profit, which heavily depends on the customer's satisfaction.
+But it also depends on the happiness of their employees. 
+Employees have a claim to professional excellence.
+They want to deliver a good job within their means.
+This effects product quality and thus effects customer satisfaction.
+If employees cannot satisfy their demand for professional excellence, they might get discontent, deliver poor work or produce higher costs through sick days.
+The worst case scenario could be, that a discontent employee gets angry and steals data or harms the running systems.
+
+In short, System Providers are interested in ...
+\begin{itemize}
+\item ... financial profit
+\item ... good working conditions
+\item ... professional excellence
+\end{itemize}
+
+\textbf{Local Authority:}
+
+In short, Local Authorities are interested in ...
+\begin{itemize}
+\item ... business intelligence (profit, RoI)
+\item ... political reputation
+\end{itemize}
 

### 2014-09-15 PrivacyProtection-2014-09-09.tex
WIP reviews model

--- 2014-09-15 PrivacyProtection-2014-09-09.tex
@@
 
-The \emph{reference model} consists of a \emph{model}, which aims to organize common security terminology in a reasonable and practical way and a systematic analysis \emph{procedure} based on that model.
-
+The reference model consists of a \emph{model} and a \emph{procedure}.
+The model aims to organize common security terminology in a reasonable and practical way.
+The procedure describes a systematic way to analysis the IT system based on that model.
 In this section we give a brief overview over the reference model.
-Before we can start with the description of model and procedure we introduce the following basic terminology (cf. Figure \ref{fig:terminology}).
-
-%\includegraphics{diagrams/png/itsecterminology.png}
-\input{./figures/Basic-IT-Security-Terminology}
-
-\begin{itemize}
-
-\item A \emph{threat} is a class of potential events whose manifestation can
-cause damage or harm.
-
-\item A \emph{vulnerability} is a weakness which can cause the loss of security.
-
-\item a \emph{hazard} is a concrete risk due to one or more concrete vulnerabilities and corresponding threats.
-
-\texttt{threat + vulnerability -\textgreater{} hazard}
-
-\item An \emph{incident} is an event caused by a threat manifestation against one or more
-corresponding vulnerabilities.
-
-\texttt{threat manifestation + vulnerability -\textgreater{} hazard}
-
-\item An \emph{accident} is an incident which did cause damage or harm. (Translates to 'Schandensfall/Zwischenfall/Störfall' in the German language).
-
-\item A \emph {non-accident} is an incident which did not cause any damage or harm.
-
-\item A security \emph{measure} is a set of one or more activities which reduce or eliminate
-vulnerabilities.
-
-\end{itemize}
 
@@
 \item
-\textbf{Assets.}
-Things of value to one or more stakeholders.
-The value can be ``hard'' (money, data, etc.) or ``soft'' (trust,privacy,etc.).
+\textbf{Assets.} Things of value to one or more stakeholders.
+The value can be ``hard'' (money, data, etc.) or ``soft'' (trust, privacy,etc.).
+In our case the only asset we are concerned with is the privacy of the citizen.
 
 \item \textbf{IT-Systems.}
-The IT-Systems under study.
-
-\item \textbf{Humans.}
-All identifiable actors of the system under study.
-Many problems arise due to misunderstandings during man-machine interaction.
+A description of the IT-Systems under study.
+The level of garnularity has to be detailed enough to express all possible threats to the assets at stake.
+
+\item \textbf{Actors.}
+All identifiable stakeholders of the system under study.
+Typical actors include, users, developers, clients and criminals.
 
 \item \textbf{Conflicts of Interests.}
-Different actors have different interests.
-Those interest can be in conflict.
-A trival conflict is the \emph{Attacker-Attackee-Conflict}:
-A service provider offers private data storage, therefore the provider is interested in having the access restricted.
-An attacker is naturally interested in easy access in order captialize the stolen data.
-However, heavy security restriction are also a burden for the service operator, as they are interested in efficient work flows.
-This kind of conflict can create vulnerabilities.
+Different actors have different interests which can be in conflict which each other.
+These conflicts of interest are the origin of all attacks to the system.
+
+A typical conflict is the \emph{Criminal-User-Conflict}:
+A user wants to keep control over their private data.
+A criminal wants to gain money.
+The possibility of selling private data (user profiles) to advertizers, renders both interests conflicting.
 
 \item \textbf{Vulnerabilities.}
-All indentifiable weaknesses in the current system layout.
+All indentifiable weaknesses in the IT-System.
+
+In the example of the cirminal-user-conflict, the criminal has to exploit a vulnerability, e.g. a weak password, to gain access to the private data abou the user.
 
 \item \textbf{Interactions.}
-Assets, IT-Systems, Humans and Vulnerabilites are in continous interaction with each other in order to support the business objectives.
-Those interactionscan also lead to threats, e.g. having unencrypted communication with a server.
-Therefore all interactions which occur in the system's outline have to be documented.
-\end{itemize}
-
-The \textbf{potential view} describes both the desired and dreaded state.
-It consists of the following components.
-\begin{itemize}
+This point captures all possible interactions between assets, IT-Systems, humans and vulnerabilites.
+It is described in more detail in the next view.
+\end{itemize}
+
+The \textbf{potential view} describes the intended and unintended interactions of the components of the world view.
+The intended interactions support the underlying business objectives.
+Unintended interactions lead to threats.
+The potential view consists of the following components.
+
+\begin{itemize}
+\item \textbf{Business Objectives.}
+Describes how IT-system and actors should interact in order to reach the goals of operation.
+
+\item \textbf{Threats.}
+A threat is a potential interaction that destroys or harms assets of the system.
+Concrete realizations of threats can be \emph{attacks} or \emph{accidents}.
+Attacks are executed by an actor in response to a conflict of interest.
+Accidents are harmful interactions that are not willfully caused by an actor.
+
 \item \textbf{Chances/Risk.}
-Chances and Risks affect business objectives and threats.
-
-\item \textbf{Business Objectives.}
-Chances the system under study.
-This describes the desired state.
-
-\item \textbf{Threats.}
-Risks for business objectives, this can simply be the integrety of the IT-System.
-Threats target one or more assets and violate security requirements}.
-This describes the dreaded state.
-
-\item
-\textbf{Security Requirements:} A set of distinct requirements for a safe and secure system.
-Security requirements hold for one or more assets.
-This describes the potential state after this procedure.
+Evaluation of chances and risks associated to the business objectives and threats.
+The risk associated to a threat is its expected loss.
+A chance associated to a intended interaction is its expected gain.
+
+In the case that, the loss can be quantified monetary, and the likelihood of occurence of a threat can
+be modeled probabilistically, the risk is given by the procuct
+\[ \text{risk} = \text{loss} \cdot P[\text{threat}]. \]
+In practice such a quantitative risk evaluation is often not possible, and a qualitative, heursitic, analysis is performed instead.
+
+\item \textbf{Security Requirements.}
+Targeted change in specify interactions that may not occure within the system in order to acieve its business objectives.
+Security requirements are targeted to protect one or more assets.
+
 \end{itemize}
@@
 The \textbf{concept view} is a realization of the the potential view of the system.
-It describes a plan of sufficient measures to secure the desired state and to aviod the dreaded state of the system.
+It describes a plan of sufficient measures to secure the intended interactions and to aviod the unintended interactions.
 It contians the following components:
@@
 A specific technique or technology to attack IT-Systems (Man in the Middel, Phishing, etc.).
-(\textbf{Note:} This component is not conform to the glossary as the term \emph{accident} there covers both disaster and attack!)
+
+%TODO: (\textbf{Note:} This component is not conform to the glossary as the term \emph{accident} there covers both disaster and attack!)
 
@@
 Each security mesaure opposes an accident scenario or an attack technique.
-
-\end{itemize}
-
-The \textbf{event view} contains all actual operations and events through out the production phase.
+\end{itemize}
+
+The \textbf{event view} contains all actual events through out the lifetime of the system.
 The event view instantiates the concept view of the system.
@@
 \item
-  \textbf{Business Process:} The actual, running instance of the
-  \textbf{Busieness Modell}
+  \textbf{Business Process:} The actual, running instance of the business model.
 \item
 \textbf{Accidents:} All actually happened accidents.
+
 \item
@@
 there covers both disaster and attack!)
+
 \item

### 2014-09-15 PrivacyProtection-2014-09-09.tex
Cleand up Grimm Model.

@@
 The model aims to organize common security terminology in a reasonable and practical way.
-The procedure describes a systematic way to analysis the IT system based on that model.
+The procedure describes a method to analysis the IT system based on that model.
 In this section we give a brief overview over the reference model.
@@
 
-The \textbf{world view} contains all components describing the current state.
+The \emph{world view} contains all components describing the current state.
 It consists of the following components:
 \begin{itemize}
-\item
-\textbf{Assets.} Things of value to one or more stakeholders.
-The value can be ``hard'' (money, data, etc.) or ``soft'' (trust, privacy,etc.).
-In our case the only asset we are concerned with is the privacy of the citizen.
-
-\item \textbf{IT-Systems.}
-A description of the IT-Systems under study.
-The level of garnularity has to be detailed enough to express all possible threats to the assets at stake.
-
 \item \textbf{Actors.}
@@
 Typical actors include, users, developers, clients and criminals.
+
+\item
+\textbf{Assets.} Things of value to one or more stakeholders.
+The value can be hard (money, data, etc.) or soft (trust, privacy,etc.).
+In our case the only asset we are concerned with is the privacy of the citizen.
+
+\item \textbf{IT-Systems.}
+The releveant IT-Systems under study. This encompasses hardware (e.g. servers, network infrastructure), as well as software and third party services.
+The level of garnularity has to be detailed enough to express all possible threats to the assets at stake.
 
@@
 
-The \textbf{potential view} describes the intended and unintended interactions of the components of the world view.
+The \emph{potential view} displays the intended and unintended interactions of the components in the world view.
 The intended interactions support the underlying business objectives.
@@
 The potential view consists of the following components.
-
 \begin{itemize}
 \item \textbf{Business Objectives.}
-Describes how IT-system and actors should interact in order to reach the goals of operation.
+Interaction of IT-system and actors that realize a business goals of the system owner.
 
@@
 \item \textbf{Security Requirements.}
-Targeted change in specify interactions that may not occure within the system in order to acieve its business objectives.
+A set of interactions (e.g. threats) that shall not occure within the system in order to acieve its business objectives.
 Security requirements are targeted to protect one or more assets.
 
-\end{itemize}
-
-The \textbf{concept view} is a realization of the the potential view of the system.
-It describes a plan of sufficient measures to secure the intended interactions and to aviod the unintended interactions.
+An example of a security requirement is that a given communication channel shall not be infringed by criminals.
+\end{itemize}
+
+The \emph{concept view} is a realization of the the potential view of the system.
+It specifies important interactions that require further planning.
 It contians the following components:
@@
 \item \textbf{Accident Scenario.}
-This is either the plan of a certain attack or the concrete outline of random disaster.
-Although the latter is most likely hard to anticipate.
+A concrete outline of an interaction that leads to an accident.
+In particular the asset under threatened and expolited vulnerability need to be described.
 
@@
 A specific technique or technology to attack IT-Systems (Man in the Middel, Phishing, etc.).
-
-%TODO: (\textbf{Note:} This component is not conform to the glossary as the term \emph{accident} there covers both disaster and attack!)
+In particular the attacking actor, the conflict of interest and the expolited vulnerability need to be described.
 
 \item \textbf{Security Measures.}
-The plan to achieve a safe and secure system.
-Each security mesaure opposes an accident scenario or an attack technique.
-\end{itemize}
-
-The \textbf{event view} contains all actual events through out the lifetime of the system.
+It describes a plan of sufficient measures to secure the intended interactions and to aviod the unintended interactions.
+Each security mesaure targets a vulnearbility of the system in order to reduce a risk for a certain thread.
+\end{itemize}
+
+The \emph{event view} contains all actual events through out the lifetime of the system.
 The event view instantiates the concept view of the system.
@@
 \begin{itemize}
-\item
-  \textbf{Business Process:} The actual, running instance of the business model.
-\item
-\textbf{Accidents:} All actually happened accidents.
-
-\item
-\textbf{Attacks:} All actually happened attacks. (\textbf{Note:} This
-component is not conform to the glossary as the term \emph{accident}
-there covers both disaster and attack!)
-
-\item
-\textbf{Security Operations:} Instances of \textbf{Security Measures}.
-Security Measures \emph{secure} the system against \textbf{Accidents}
-and \textbf{Attacks}.
+\item \textbf{Business Process.}
+The actual, running instance of the business model.
+
+\item \textbf{Accidents.}
+All actually happened accidents.
+
+\item \textbf{Attacks.}
+All actually happened attacks.
+
+\item \textbf{Security Operations.}
+Instances of security measures.
 \end{itemize}
@@
 
-The analysis procedure is an incremental and iterative process following
-the four views of the previously described ontology.
+The analysis procedure is an incremental and iterative process following the four views of the previously described model.
 
@@
 
-\paragraph{Step 1. World Analysis}
-
-At first, one has to outline the current state of the system under
-study. This includes description of:
-
-\begin{itemize}
-
-\item
-  all \textbf{Assets} which must be protected
-\item
-  all relevant \textbf{IT-Systems}
-\item
-  all involved \textbf{Humans} and their \textbf{Conflicts of Interests}
-\item
-  all known \textbf{Vulnerabilities}
-\item
-  and all important \textbf{Interactions} between the former components.
-\end{itemize}
-
-\paragraph{Step 2. Potential Analysis}
-
-Secondly, one needs to outline the potential state of the system under
-study. This includes both the dreaded (\textbf{Threats} and
-\textbf{Risks}) and the desired (\textbf{Business Objectives},
-\textbf{Chances}, \textbf{Security Requirements} state. This step
-produces four artefacts:
-
-\begin{itemize}
-
-\item
-  a \emph{threat specifiation}, which identifies the \textbf{Threat} and
-  its targeted \textbf{Assets}
-\item
-  a \emph{threat risk evaluation}, which detrmines the likelyhood of a
-  threat manifestation
-\item
-  a \emph{hazard matrix}, which maps threats to known
-  \textbf{Vulnerabilities} and identifies potential hazards
-\item
-  a \emph{security requirement specifiation}, which specifies
-  requirments in order to deal with identified hazards
-\end{itemize}
-
-\paragraph{Step 3. Plan Development}
-
-Based on \textbf{Step 2.}, the identified hazards are used alongside
-realistic \textbf{Accident Scenarios} (\textbf{Attack Techniques}) to
-create a \emph{risk matrix}. With this matrix it is possible to decide
-if the risk is acceptable or not. Together with the previously specified
-\textbf{Security Requirements}, the matrix is used to define adequate
-\textbf{Security Measures}. Like a \textbf{Business Model} is an
-abstract concept to achieve \textbf{Business Objectives}, this step
-creates an abstract concept to improve the system's security.
-
-\paragraph{Step 4. Plan Deployment}
-
-Finally, the \textbf{Security Measures} have to be implemented.
-Additionally, all \textbf{Business Operations}, \textbf{Accidents}
-(\textbf{Attacks}) and executed \textbf{Security Operations} will be
-recorded through out the production phase.
-
-The implementation of \textbf{Security Measures} evantually changes the
-\textbf{World} and renders the conducted analysis outdated. So this
-analysis procedure needs to be conducted again.
-
-\subsubsection{Abstraction Levels of the Reference Model}
-
-The Reference Model can be used on different levels of abstraction. This
-means each component can be used within a wide range of granularity, for
-instance the security measure \emph{Encryption} can be explored in
-genral or on the level of different concrete encryption tools; or on the
-even finer level of concrete algorithms.
-
-The utilized abstraction level is not important for the analysis
-procedure, it depends on the intended audience for the analysis.
-However, it is important to use one abstraction level consistently
-through out the analysis.
+\paragraph*{Step 1. World Analysis}
+
+At first, one has to outline the current state of the system under study. This includes description of:
+
+\begin{itemize}
+\item all \textbf{Assets} which must be protected
+\item all relevant \textbf{IT-Systems}
+\item all involved \textbf{Humans} and their \textbf{Conflicts of Interests}
+\item all known \textbf{Vulnerabilities}
+\item and all important \textbf{Interactions} between the former components.
+\end{itemize}
+
+\paragraph*{Step 2. Potential Analysis}
+
+Secondly, one needs to outline the potential insteractions of the system under study.
+This includes both the unintended interactions (threads) and the intended interactions (business objectives).
+This step produces four artefacts:
+\begin{itemize}
+\item a \emph{threat specifiation}, which identifies the threat, its targeted assets, the involved actors and thier conflicts of interest.
+\item a \emph{threat risk evaluation}, which quantifies the likelyhood of a threat manifestation in relation to its associated loss.
+\item a \emph{security requirement specifiation}, which specifies requirments in order to deal with identified hazards
+\end{itemize}
+
+\paragraph*{Step 3. Plan Development}
+
+Based on \textbf{Step 2.}, the identified hazards are used alongside realistic accident ccenarios and attack techniques to create a \emph{risk matrix}.
+With this matrix it is possible to decide if the risk is acceptable or not.
+Together with the previously specified security requirements, the matrix is used to define adequate security measures.
+Like a business model is an abstract concept to achieve business objectives, this step creates an concept to improve the system's security.
+
+\paragraph*{Step 4. Plan Deployment}
+
+Finally, the security measures have to be implemented.
+Additionally, all business operations, accidents, attacks and executed security operations will be recorded in the following time.
+
+The implementation of security measures changes the world view (e.g. IT-systems, actors) and renders the conducted analysis outdated.
+So this analysis procedure needs to be conducted again.
+
+\subsection{Abstraction Levels of the Reference Model}
+
+The Reference Model can be used on different levels of abstraction.
+This means each component can be used within a wide range of granularity, for instance the security measure \emph{Encryption} can be explored in genral or on the level of different concrete encryption tools; or on the even finer level of concrete algorithms.
+
+The utilized abstraction level is not important for the analysis procedure, it depends on the intended audience for the analysis.
+However, it is important to use one abstraction level consistently through out the analysis.
 

### 2014-09-17 PrivacyProtection-2014-09-09.tex
local changes

--- 2014-09-15 PrivacyProtection-2014-09-09.tex
@@
 
-The reference model consists of a \emph{model} and a \emph{procedure}.
-The model aims to organize common security terminology in a reasonable and practical way.
-The procedure describes a systematic way to analysis the IT system based on that model.
+The \emph{reference model} consists of a \emph{model}, which aims to organize common security terminology in a reasonable and practical way and a systematic analysis \emph{procedure} based on that model.
+
 In this section we give a brief overview over the reference model.
+Before we can start with the description of model and procedure we introduce the following basic terminology (cf. Figure \ref{fig:terminology}).
+
+%\includegraphics{diagrams/png/itsecterminology.png}
+\input{./figures/Basic-IT-Security-Terminology}
+
+\begin{itemize}
+
+\item A \emph{threat} is a class of potential events whose manifestation can
+cause damage or harm.
+
+\item A \emph{vulnerability} is a weakness which can cause the loss of security.
+
+\item a \emph{hazard} is a concrete risk due to one or more concrete vulnerabilities and corresponding threats.
+
+\texttt{threat + vulnerability -\textgreater{} hazard}
+
+\item An \emph{incident} is an event caused by a threat manifestation against one or more
+corresponding vulnerabilities.
+
+\texttt{threat manifestation + vulnerability -\textgreater{} hazard}
+
+\item An \emph{accident} is an incident which did cause damage or harm. (Translates to 'Schandensfall/Zwischenfall/Störfall' in the German language).
+
+\item A \emph {non-accident} is an incident which did not cause any damage or harm.
+
+\item A security \emph{measure} is a set of one or more activities which reduce or eliminate
+vulnerabilities.
+
+\end{itemize}
 
@@
 \item
-\textbf{Assets.} Things of value to one or more stakeholders.
-The value can be ``hard'' (money, data, etc.) or ``soft'' (trust, privacy,etc.).
-In our case the only asset we are concerned with is the privacy of the citizen.
+\textbf{Assets.}
+Things of value to one or more stakeholders.
+The value can be ``hard'' (money, data, etc.) or ``soft'' (trust,privacy,etc.).
 
 \item \textbf{IT-Systems.}
-A description of the IT-Systems under study.
-The level of garnularity has to be detailed enough to express all possible threats to the assets at stake.
-
-\item \textbf{Actors.}
-All identifiable stakeholders of the system under study.
-Typical actors include, users, developers, clients and criminals.
+The IT-Systems under study.
+
+\item \textbf{Humans.}
+All identifiable actors of the system under study.
+Many problems arise due to misunderstandings during man-machine interaction.
 
 \item \textbf{Conflicts of Interests.}
-Different actors have different interests which can be in conflict which each other.
-These conflicts of interest are the origin of all attacks to the system.
-
-A typical conflict is the \emph{Criminal-User-Conflict}:
-A user wants to keep control over their private data.
-A criminal wants to gain money.
-The possibility of selling private data (user profiles) to advertizers, renders both interests conflicting.
+Different actors have different interests.
+Those interest can be in conflict.
+A trival conflict is the \emph{Attacker-Attackee-Conflict}:
+A service provider offers private data storage, therefore the provider is interested in having the access restricted.
+An attacker is naturally interested in easy access in order captialize the stolen data.
+However, heavy security restriction are also a burden for the service operator, as they are interested in efficient work flows.
+This kind of conflict can create vulnerabilities.
 
 \item \textbf{Vulnerabilities.}
-All indentifiable weaknesses in the IT-System.
-
-In the example of the cirminal-user-conflict, the criminal has to exploit a vulnerability, e.g. a weak password, to gain access to the private data abou the user.
+All indentifiable weaknesses in the current system layout.
 
 \item \textbf{Interactions.}
-This point captures all possible interactions between assets, IT-Systems, humans and vulnerabilites.
-It is described in more detail in the next view.
-\end{itemize}
-
-The \textbf{potential view} describes the intended and unintended interactions of the components of the world view.
-The intended interactions support the underlying business objectives.
-Unintended interactions lead to threats.
-The potential view consists of the following components.
-
-\begin{itemize}
+Assets, IT-Systems, Humans and Vulnerabilites are in continous interaction with each other in order to support the business objectives.
+Those interactionscan also lead to threats, e.g. having unencrypted communication with a server.
+Therefore all interactions which occur in the system's outline have to be documented.
+\end{itemize}
+
+The \textbf{potential view} describes both the desired and dreaded state.
+It consists of the following components.
+\begin{itemize}
+\item \textbf{Chances/Risk.}
+Chances and Risks affect business objectives and threats.
+
 \item \textbf{Business Objectives.}
-Describes how IT-system and actors should interact in order to reach the goals of operation.
+Chances the system under study.
+This describes the desired state.
 
 \item \textbf{Threats.}
-A threat is a potential interaction that destroys or harms assets of the system.
-Concrete realizations of threats can be \emph{attacks} or \emph{accidents}.
-Attacks are executed by an actor in response to a conflict of interest.
-Accidents are harmful interactions that are not willfully caused by an actor.
-
-\item \textbf{Chances/Risk.}
-Evaluation of chances and risks associated to the business objectives and threats.
-The risk associated to a threat is its expected loss.
-A chance associated to a intended interaction is its expected gain.
-
-In the case that, the loss can be quantified monetary, and the likelihood of occurence of a threat can
-be modeled probabilistically, the risk is given by the procuct
-\[ \text{risk} = \text{loss} \cdot P[\text{threat}]. \]
-In practice such a quantitative risk evaluation is often not possible, and a qualitative, heursitic, analysis is performed instead.
-
-\item \textbf{Security Requirements.}
-Targeted change in specify interactions that may not occure within the system in order to acieve its business objectives.
-Security requirements are targeted to protect one or more assets.
-
+Risks for business objectives, this can simply be the integrety of the IT-System.
+Threats target one or more assets and violate security requirements.
+This describes the dreaded state.
+
+\item
+\textbf{Security Requirements:} A set of distinct requirements for a safe and secure system.
+Security requirements hold for one or more assets.
+This describes the potential state after this procedure.
 \end{itemize}
@@
 The \textbf{concept view} is a realization of the the potential view of the system.
-It describes a plan of sufficient measures to secure the intended interactions and to aviod the unintended interactions.
+It describes a plan of sufficient measures to secure the desired state and to aviod the dreaded state of the system.
 It contians the following components:
@@
 A specific technique or technology to attack IT-Systems (Man in the Middel, Phishing, etc.).
-
-%TODO: (\textbf{Note:} This component is not conform to the glossary as the term \emph{accident} there covers both disaster and attack!)
+(\textbf{Note:} This component is not conform to the glossary as the term \emph{accident} there covers both disaster and attack!)
 
@@
 Each security mesaure opposes an accident scenario or an attack technique.
-\end{itemize}
-
-The \textbf{event view} contains all actual events through out the lifetime of the system.
+
+\end{itemize}
+
+The \textbf{event view} contains all actual operations and events through out the production phase.
 The event view instantiates the concept view of the system.
@@
 \item
-  \textbf{Business Process:} The actual, running instance of the business model.
+  \textbf{Business Process:} The actual, running instance of the
+  \textbf{Busieness Modell}
 \item
   \textbf{Accidents:} All actually happened accidents.
-
 \item
@@
   there covers both disaster and attack!)
-
 \item

### 2014-09-29 PrivacyProtection-2014-09-09.tex
Resolved conflict

--- 2014-09-15 PrivacyProtection-2014-09-09.tex
@@
 \chapter{Legal and Ethical aspects of Privacy}
+\label{chap:privacy}
 
@@
 \section{Privacy Definition and Taxonomy}
+\label{sec:taxonomy}
 
@@
 At first, one has to outline the current state of the system under study. This includes description of:
-
 \begin{itemize}
@@
 \item all known \textbf{Vulnerabilities}
-\item and all important \textbf{Interactions} between the former components.
+% \item and all important \textbf{Interactions} between the former components.
 \end{itemize}
@@
 
-We focus on the Asset privacy as described in Section \ref{sec:taxonomy}
+In this document we focus our attention to only one asset: The privacy of the citizen.
+
+%TODO: UPDATE
+Our definition of privacy is descibed in detail in Chapter \ref{chap:privacy}.
+In Section \ref{sec:taxonomy} we define privacy as the ``control over private data'' and introduce the following seven different types of privacy:
+\begin{enumerate}
+\item Privacy of the Person
+\item Privacy of Behaviour and Action
+\item Privacy of Communication
+\item Privacy of Data and Image
+\item Privacy of Thoughts and Feelings
+\item Privacy of Location and Space
+\item Privacy os Association
+\end{enumerate}
+
+Please refer to Section \ref{sec:taxonomy} for more details.
 
@@
 
+The IT infrastructure of the Live+Gov system, i.e.
+the Live+Gov toolkit and the customization of the software components are described in detail in various project deliverables: D4.1, D4.3, D1.1, D5.1.
+In this section we give an abstraction of those systems from the perspective of WP1.
+
 \input{./figures/IT-Systems}
 
-A \textit{Citizen (a)} carries a mobile device running the \textit{Sensor Collector (d)} application.
-The \textit{Sensor Collector} application collects sensor data (accelerometer, GPS, GSM, ...) and sends it to the \textit{Data Cetner (c)}. 
-The \textit{Data Center} on the other side sends beneficial information (traffic jam reports, bus schedule) for \textit{Citizens} to the mobile device.
-The \textit{System Provider (b)} provides and operates technical infrastructure like the Data Center and the \textit{Report Tool (f)}.
+A citizen (a) carries a mobile device running the \emph{sensor collector} (d) application.
+The sensor collector application collects sensor data (accelerometer, GPS, GSM, ...) and sends it to the \emph{data center} (c).
+The sensor collector can also be able to perform certain data mining operations.
+Examples of such data mining operations include human activity recognition,the detection of service lines, detection of characters and faces from images.
+
+The \emph{data center} stores and processes sensor data collected with the sensor collector application.
+Again the amount
+
+ sends beneficial information (traffic jam reports, bus schedule) for \textit{Citizens} to the mobile device.
+The \textit{System Provider (b)} provides and operates technical infrastructure like the \textit{Data Center} and the \textit{Report Tool (f)}.
 The \textit{Report Tool} queries the \textit{Data Center} for aggregated data to visualize in form of charts and other means suitable to help understanding of monitored \textit{Citizens}. 
@@
 In such systems, the most valuable information is not the raw sensor data, the beneficial information for \textit{Local Authorities} is produced via \textit{Data Mining} (Human Activity Recognition: detect running, walking, standing, ...).
-Usually, \textit{Data Mining} is heavily conduct within the \textit{Data Center}, but simpler mining processes can also be conducted on mobile devices.
+Usually, \textit{Data Mining} is heavily conducted within the \textit{Data Center}, but simpler mining processes can also be conducted on mobile devices.
 
@@
 \textbf{Citizen:} 
-\begin{itemize}
-\item legal right to privacy
-\item no spam
-\item no stalking
-\item no loss / gain of money
-\item physical wellbeing / personal health (attack because someone knows your gps location)
-\item transparency: who has my data
-\item good pricing of insurance policies
-\end{itemize}
+Citizens are persons who take part in the monitoring system as the monitored subjects.
+They offer their privacy sensitive data, because by participating they expect a higher level of comfort in their lives through the offered services in return.
+E.g. they will have access to real time bus schedules and traffic jam reports, localized social networking or they generally benefit from public improvement by Local Authorities.
+
+However, they have a legal right to privacy and are interested in having that privacy respected by others.
+For example, all humans have a natural interest in physical wellbeing and personal health. 
+If someone monitors their immediate location (GPS) or frequently used routes (stalking), people with access to that information could attack and harm citizens.
+
+In general, citizens are not interested in having information about them exploited by others. 
+This includes just annoying spam, where their technical identity is used to send them unwanted commercials.
+But it also includes more severe phishing attacks, which try to rob citizens of their TAN numbers and disrespect their financial interests.
+
+Such financial interests of monitored citizens, the interest in not losing / gaining money, could be disrespected on an even larger scale.
+Collected data could be used for scoring and analysis, which may affect the pricing of policies or credits, if sold to insurances or banks.
+
+So citizens have an inherent interest in transparency and disclosure of who uses their data for which purposes, as it may affect their physical and financial wellbeing.
 
@@
 \textbf{Local Authority:} 
+Local Authorities are public offices (ministry, agency, department, ...) or other external public service providers which act as direct customers of Service Providers.
+They order a proposed monitoring and mining system specialized for their needs. 
+For example a low level department, like departments for public mobility, orders a monitoring and mining system to better understand and improve urban traffic flow.
+
+Such systems are heavy investments, so naturally Local Authorities are interested in a profitable return.
+However, the return of investment is not directly of financial nature.
+Like Service Providers their financial gain depends on customer satisfaction. 
+Customers for Local Authorities are either citizens, who use their services, or politicians, who order their services.
+The happiness of both sides is interdependent. 
+Citizens are happy customers if the services, e.g. public mobility, work well. 
+If citizens are happy, it is more likely that politicians gain reputation, as they organize the public services through Local Authorities.
+
+One way of achieving these business objectives is by utilizing business intelligence. 
+A proposed monitoring and mining system can provide such intelligence.
+For public mobility services it can provide traffic jam detection to maximize traffic flow.
+Or it can provide user profiling to maximize (targeted) publicity.
+
+Additionally, since Local Authorities act like corporations comparable to Service Providers, they are also interested in good working conditions.
+As discontent employees may harm the system by disclosure of business intelligence.
 
@@
 \begin{itemize}
-\item ... business intelligence (profit, RoI)
+\item ... business intelligence
 \item ... political reputation
+\item ... good working conditions
 \end{itemize}

### 2014-09-29 PrivacyProtection-2014-09-09.tex

@@
 
-\subsubsection{Conflict of Interests (Figure \ref{figure:Live+Gov Conflicts of Interests})}
-
-%\includegraphics{../diagrams/png/conflict-of-interests.png}
+\subsubsection{Conflicts of Interests}
+\label{subsubsection:Conflicts of Interests}
+This section outlines the Conflicts of Interests (Figure \ref{figure:Live+Gov Conflicts of Interests}) between the actors of the proposed IT system architecture. 
+The individual interests of all actors is already described in the previous section and are not elaborated any further.
+The emphasis here is put on the existing conflicts.
+
 \input{./figures/LG-Conflicts-Of-Interests}

### 2014-10-08 PrivacyProtection-2014-09-09.tex

@@
 
-\paragraph*{Step 3. Plan Development}
+\paragraph*{Step 3. Concept Development}
 
@@
 
-\paragraph*{Step 4. Plan Deployment}
+\paragraph*{Step 4. Concept Deployment}
 
@@
 
-\subsubsection{IT-Systems}
+\subsubsection{IT-Systems \& Interactions}
 \label{subsubsection:it-systems}
@@
 This section outlines the general architecture (Figure \ref{figure:IT Systems}) of IT systems for public monitoring comparable to the Live+Gov project.
+This includes a description of it technical infrastructure and the interactions between its components.
 
@@
 
- sends beneficial information (traffic jam reports, bus schedule) for \textit{Citizens} to the mobile device.
+Vice versa, it sends beneficial information (traffic jam reports, bus schedule) for \textit{Citizens} to the mobile device.
 The \textit{System Provider (b)} provides and operates technical infrastructure like the \textit{Data Center} and the \textit{Report Tool (f)}.
@@
 In general, citizens are not interested in having information about them exploited by others. 
-This includes just annoying spam, where their technical identity is used to send them unwanted commercials.
+This includes just annoying spam, where their technical identity is used to send unwanted commercials.
 But it also includes more severe phishing attacks, which try to rob citizens of their TAN numbers and disrespect their financial interests.
@@
 Another possible motivation for criminal activities could be social appreciation.
-A hacker could attack critical infrastructure just to show his skills.
+A hacker could attack critical infrastructure just to prove his skills.
 
@@
  a support manager, who handles customer relations;
-...
-
-As companies, they follow typical business objectives. 
-They are interested in gaining financial profit, which heavily depends on the customer's satisfaction.
-But it also depends on the happiness of their employees. 
+ etc.
+
+As companies, they are interested in gaining financial profit.
+Among other things, this depends on customer satisfaction, employee happiness and task complexity.
+Unsatisfied customers may not want to pay for the service or do not continue the business relation.
+Moreover, unsatisfied customers can create a bad reputation, which affects the market for future customers.
+
+Customer satisfaction is connected with the quality of the offered product or service.
+This quality depends on the happiness of employees. 
 Employees have a claim to professional excellence.
 They want to deliver a good job within their means.
-This effects product quality and thus effects customer satisfaction.
-If employees cannot satisfy their demand for professional excellence, they might get discontent, deliver poor work or produce higher costs through sick days.
+If employees cannot satisfy their demand for professional excellence, they might get discontent and deliver poor work.
+Moreover, unhappy employees can produce higher costs through sick days.
+Ill employees stay on payroll, but stress the project schedule by splitting the workload on fewer heads.
 The worst case scenario could be, that a discontent employee gets angry and steals data or harms the running systems.
+
+At last, the financial success of System Providers depends on the task complexity.
+The complexity of a task has to be in reasonable bounds, so that System Providers can complete it within time, with a satisfying quality.
+If a task has a higher complexity than expected, financial loss is almost determined.
+Either System Providers need to hire additional competence to meet schedule and requirements.
+Or System Providers they stress the time-line, which also results in a higher man-hour salary ratio and additionally endangers customer satisfaction.
+Ultimately, high task complexities can affect employee happiness, if employees cannot complete it within their claim to professional excellence.
 
@@
 \item ... professional excellence
+\item ... reasonable complexity
 \end{itemize}
@@
 The individual interests of all actors is already described in the previous section and are not elaborated any further.
-The emphasis here is put on the existing conflicts.
+The emphasis here is put on prominent existing conflicts, because they provide a foundation for vulnerabilities and subsequent threats.
 
@@
 
-\textbf{Citizen vs L+G Maintenance Crew}
-
-\begin{itemize}
-
-\item Citizen want to exercise their legal right to privacy. I.e. their right to control information about themselves.
-- citizens want to be unidentifiable
-- citizens want certain informations about them to be secret
-\item Citizen wants to prevent abuse of their private data, and the related negative consequences.
-\item
-  technical staff needs more or less unrestricted access to data in
-  order to maintain the system
-\end{itemize}
-
-\textbf{Citizen vs Local Authority}
-
-\begin{itemize}
-
-\item
-  citizens want to be unidentifiable
-\item
-  citizens want certain informations about them to be secret
-\end{itemize}
-
-\textbf{L+G Maintenance Crew vs Local Authority}
-
-\begin{itemize}
-
-\item
-  staff wants proper payment
-\item
-  local authorities want to reduce cost
-\end{itemize}
-
-\textbf{Criminal vs Citizen}
-
-\begin{itemize}
-
-\item
-  citizens want to be unidentifiable
-\item
-  citizens want certain informations about them to be secret
-\item
-  criminals want to posses certain information about a citizen
-\end{itemize}
-
-\textbf{Criminal vs L+G Maintenance Crew}
-
-\begin{itemize}
-
-\item
-  criminals want to harm the system
-\item
-  technical staff has to keep the system safe and sound
-\end{itemize}
-
-\textbf{Criminal vs Local Authority}
-
-\begin{itemize}
-
-\item
-  criminals want to harm the system
-\item
-  criminals want access to collected data
-\item
-  local authorities have to keep collected data restricted
-\item
-  local authorities want a stable system
-\end{itemize}
+\textbf{System Complexity vs Privacy:}
+System Providers offer a service to Local Authorities, which is to provide and maintain a monitoring and mining system, e.g. for public mobility.
+This system shall produce business intelligence, so that Local Authorities can improve their public services.
+This task in it self has a high technical complexity and is the sole asset with financial return for System Providers.
+However, this task operates on privacy sensitive data provided by monitored Citizens.
+In order to ensure their privacy, System Provider would have to implement additional mechanisms, which allow Citizens to exercise control of their data.
+But this will eventually raise the complexity of the monitoring and mining system.
+
+\textbf{Business Intelligence vs Privacy:}
+Local Authorities order a monitoring and mining system from System Providers, which allows them to produce business intelligence for public services.
+The system is a heavy investment for Local Authorities, so they are interested in as much intelligence as possible to achieve a profitable return.
+But the intelligence is the result of data mining conducted on privacy sensitive data of monitored Citizens.
+They are interested in the successful usage of their data, in a sense that they are also benefactors, e.g. improvement of public mobility.
+However, the main interest of Citizens lies in maintaining control over their data.
+In order do that, they need disclosure of the exact purposes their data is used for, so they can select which data to share.
+These purposes cannot be (secretly) exceeded.
+
+\textbf{Money vs Privacy:}
+Criminals can gain financial profit from stealing privacy sensitive data.
+For example by selling raw contact information to advertisers or by selling mined data to scoring services.
+In such cases, Citizens lose complete control over their data.
+
+\textbf{Money vs Professional Reputation:}
+Criminals have various business models as optional foundation for attacks on System Providers.
+For instance, they can try to invade the infrastructure for e-espionage reasons, to get control over servers to create a bot-net or to steal user data.
+All these approaches are motivated by financial interests.
+Gathered information can be sold, zombie servers can be leased.
+A successful attack proves the technical competence of System Providers wrong and subsequently harms their professional reputation.
+This can lead to a loss of future customers or a decrease of stock price for registered companies.
+Eventually the financial interests of System Providers are endangered.
+
+\textbf{Political Power vs Political Reputation:}
+Besides monetary reasons, Criminals can be motivated by political reasons to attack the monitoring and mining system.
+Criminals can break the system to make a political statement of their own, e.g. "Stop Surveillance!".
+Or they can steal user data to prove the system insecure.
+Both would harm the political reputation of Local Authorities, who endangered the privacy of monitored Citizens.
+Criminals would gain political power, because their point of view would seem to be valid.
 
 \subsubsection{Vulnerabilities}
+This section outlines the vulnerabilities (Figure \ref{figure:Live+Gov Vulnerabilities}) of the proposed monitoring and mining system.
+Note that vulnerabilities are not necessarily of technical nature.
+The weaknesses of IT systems are often created due to misuse or misconfiguration of the various components by one or more actors.
+
+\input{./figures/LG-Vulnerabilities}
+
+\textbf{Unencrypted Data Transmission:}
+The proposed monitoring and mining system uses HTTP to exchange data between the Sensor Collector, the Data Center and the Report Tool.
+Per default, HTTP is a clear text protocol.
+This means, one can intercept the connection and read all sensitive information, which is send between the components.
+That is: passwords, raw sensor data and data mining results
+
+\textbf{Inadequate Access Rules:}
+The proposed IT system infrastructure has various accesses to privacy sensitive data.
+System Provider staff has access to Data Center hardware and software like databases, web-servers and other inspection tools.
+Local Authority staff has access to the Report Tool.
+This all enables staff members to have potential access to privacy sensitive information.
+Those accesses have to be secured against unauthorized third parties, e.g. Criminals.
+Moreover, we need to ensure that no single person has to many access rights.
+For example, a system administrator should not be able to secretly download the whole database on a flash-drive.
+
+\textbf{Insufficiently Secured Infrastructure:}
+The proposed monitoring system consists of many hardware and software components, each with its own concrete weaknesses.
+For instance, operating systems can be outdated or not subject to frequent updates or virus scans.
+Web-applications can be carelessly implemented and not protected against SQL-Injections or Cross-Site-Scripting attacks.
+Databases can be ill-configured, so that access from outside the system is possible.
+All those weak points can be subject to various known exploit techniques.
+
+\textbf{Unaware Monitoring Subjects:}
+We define privacy as one's ability to control information about oneself. 
+In order to do that, monitored subjects need to know, that they are monitored, who monitors them, what information is recorded and for what purposes.
+Subjects who are not aware of these things cannot effectively preserve control and thus lose their privacy.
 
 %\includegraphics{../diagrams/png/scenario-vulnerabilities.png}
-\input{./figures/LG-Vulnerabilities}
-
-\paragraph{Combined Scenario (Figure \ref{figure:Live+Gov Vulnerabilities})}
-
-A citizen (a) carries a mobile device (b) running the L+G client. The
-mobile device collects sensor data and sends it to the L+G Data Cetner
-(c). The L+G Data Center also sends beneficial information for citizens
-to the mobile device.
-
-The L+G Maintenance Crew (d) maintaince the L+G Data Center.
-
-The Aggregation Monitor (e) queries the L+G Data Center for aggregated
-data. Local Authorities (f) use the Aggregation Monitor to get
-information in order to improve public services.
-
-A criminal (g) threatens the L+G system by corrupting the L+G
-Maintenance Crew or Local Authority, or by forcing access to L+G
-system's hardware, software or communication. (It is also possible for
-criminals to corrupt the citizen, although technically this would be no
-threat to privacy but rather an exercise of privacy
-{[}\hyperref[references]{1}{]}.)
-
-\paragraph{List of Vulnerabilities}
-
-The vulnerabilities of Live+Gov system outline are:
-
-\begin{itemize}
-\item
-  unencrypted data transmission (Man In The Middel)
-\item
-  insecure mobile devices (Exploit)
-\item
-  insecure servers (Exploit)
-\item
-  inadequate access rules for
-  \begin{itemize}
-  \item
-    mobile devices (citizens)
-  \item
-    servers (staff)
-  \item
-    applications (local authority)
-  \end{itemize}
-\item
-  corrupt/unhappy authorities or staff (\textbf{Could be actually a
-  threat not a vulnerability})
-\end{itemize}
-
-(\textbf{Note:} The vulnerabilities seem to mostly open possibilities to
-violate the \textbf{Privacy of Data and Image}, as this type is concerd
-with making data \emph{automatically} available to others.)
+
+%\paragraph{List of Vulnerabilities}
+%
+%The vulnerabilities of Live+Gov system outline are:
+%
+%\begin{itemize}
+%\item
+%  unencrypted data transmission (Man In The Middel)
+%\item
+%  insecure mobile devices (Exploit)
+%\item
+%  insecure servers (Exploit)
+%\item
+%  inadequate access rules for
+%  \begin{itemize}
+%  \item
+%    mobile devices (citizens)
+%  \item
+%    servers (staff)
+%  \item
+%    applications (local authority)
+%  \end{itemize}
+%\item
+%  corrupt/unhappy authorities or staff (\textbf{Could be actually a
+%  threat not a vulnerability})
+%\end{itemize}
+%
+%(\textbf{Note:} The vulnerabilities seem to mostly open possibilities to
+%violate the \textbf{Privacy of Data and Image}, as this type is concerd
+%with making data \emph{automatically} available to others.)
 
@@
 \subsubsection{Threats}
-
+\input{./figures/LG-Threats}
 \input{./figures/ThreatTable}

### 2014-10-08 D1-3.tex
Partial cleanup up formatting and arrangement of sections.
Added authors. Upated TOC.  Cleanup Formatting mainly.

@@
 
+\LGaddauthorNoPhone{UKob}{Maximilan Meffert}%
+{}
+
+\LGaddauthorNoPhone{BIZ}{Maite de Arana Agiretxe}%
+{\small komunikazioa.tekniko1@biscaytik.eu}
+
 %% \LGaddauthor{UKob}{Christoph Kling}{+49 261 287 2702 }%
@@
 
-\setcounter{tocdepth}{1}
+\setcounter{tocdepth}{2}
 
@@
 
-\section{Implementation}
-\subsection{Authentification and Access Control}
-\subsection{Anonymization and Pseudonoymization}
-\subsection{Randomization}
-
-\chapter{Improved Sensor Data Mining Methods}
-\section{Battery Awareness of Sensor Collector}
-Results from Projektprakikum (Wifi/Zip)
-
-\section{Service Line Detection (new method)}
-MA thesis results from Sven Milker
-
-\section{Issue Analysis}
-Christoph Schäfer with Niko Beck
+\chapter{Implementation of Privacy Protection Measures}
+\subsection{Privacy Policy}
+\subsection{Privacy Dashboard}
+\subsection{Transfer Encryption (HTTPS/SSL)}
+\subsection{Database Access Policies}
+\subsection{Anonymization of recorded trips}
+\subsection{Randomization of GPS data}
 
 \bibliography{D1-3}
-
 \end{document}

### 2014-10-08 PrivacyProtection-2014-09-09.tex
Added authors. Upated TOC.  Cleanup Formatting mainly.

@@
-\chapter{Legal and Ethical aspects of Privacy}
+>\chapter{Legal and Ethical aspects of Privacy}
 \label{chap:privacy}
@@
 Warren and Brandeis were worried that existing law like castle doctrines or libel and slander could not suffice.
-They felt a need for a new right to "protect the extent to wich one's thoughts, sentiments, and emotions could be shared with others" [1].
+They felt a need for a new right to "protect the extent to wich one's thoughts, sentiments, and emotions could be shared with others" \cite{1}.
 This right would be "the right to be left alone".
@@
 The right to "protect the extent to wich one's thoughts, sentiments, and emotions could be shared with others" by Warren and Brandeis was later extend to cover arbitrary information and is now kown as the privacy concept "...
-the control we have over information about ourselves ..." [2].
+the control we have over information about ourselves ..." \cite{2}.
 
@@
 ``(1) Everyone has the right to respect for his private and family life, his home and his correspondence.
-(2) There shall be no interference by a public authority with the exercise of this right except such as is in accordance with the law and is necessary in a democratic society in the interests of national security, public safety or the economic well-being of the country, for the prevention of disorder or crime, for the protection of health or morals, or for the protection of the rights and freedoms of others.'' {[}\hyperref[references]{3}{]}
+(2) There shall be no interference by a public authority with the exercise of this right except such as is in accordance with the law and is necessary in a democratic society in the interests of national security, public safety or the economic well-being of the country, for the prevention of disorder or crime, for the protection of health or morals, or for the protection of the rights and freedoms of others.'' \cite{2}
 
 ``The European Court of Human has given this article a very broad
-interpretation in its juresprudence.'' {[}\hyperref[references]{2}{]}
+interpretation in its juresprudence.'' \cite{2}
 
@@
 
-``The directive regulates the processing of personal data regardless of whether such processing is automated or not.''
-{[}\hyperref[references]{1}{]} where:
-
-\begin{itemize}
-\item
-  \textbf{personal data} is ``any information relating to an identified or identifiable natural person {[}\ldots{}{]} (art. 2a)''
-  {[}\hyperref[references]{1}{]}
-\item
-  \textbf{processing} is ``any operation or set of operations which is performed upon personal data {[}\ldots{}{]} (art. 2b)''
-  {[}\hyperref[references]{1}{]}
-\end{itemize}
-
-The ``identified or identifiable natural person'' is called \textbf{data subject}.
+``The directive regulates the processing of personal data regardless of whether such processing is automated or not.'' \cite{1} where \emph{personal data} is ``any information relating to an identified or identifiable natural person [...] (art.
+2a)'' and \emph{processing} is ``any operation or set of operations which is performed upon personal data [...] (art.
+2b)'' \cite{1}
+
+The ``identified or identifiable natural person'' is called \emph{data subject}.
 
@@
 A controller is an actor ``which alone or jointly with others dertimines the purposes and means of the processing of personal data; (art 2d)''
-{[}\hyperref[references]{1}{]}
-
-This rule is applicable ``whenever the controller uses equipment situated within the EU''{[}\hyperref[references]{1}{]}.
+\cite{1}
+
+This rule is applicable ``whenever the controller uses equipment situated within the EU''\cite{1}.
 This will hold for every e-commerce provider \textbf{inside or outside} of the EU because the customer's computer is \emph{situated within the EU} anyways.
 
-``Personal data should not be processed at all, except when certain conditions are met'' {[}\hyperref[references]{1}{]}.
+``Personal data should not be processed at all, except when certain conditions are met'' \cite{1}.
 These conditions incorporate the seven OECD princepals and are futher categorized in the categories:
@@
 \item
-\textbf{Transparency}: ``The data subject has the right to be informed when his personal data is being processed {[}\ldots{}{]} (art.
-10 and 11)'' {[}\hyperref[references]{1}{]} and he has the right how and by whom it is processed.
+\textbf{Transparency}: ``The data subject has the right to be informed when his personal data is being processed [...] (art. 10 and 11)'' \cite{1} and he has the right how and by whom it is processed.
 Additionally other speficied requirements have to be met, i.e.~giving consent.
@@
 \textbf{Legitimate prupose}: ``Personal data can only be processed for specified explicit and legitimate pruposes and may not be processes futher in a way incompatible with those pruposes (art. 6b)''
-{[}\hyperref[references]{1}{]}.
-\item
-\textbf{Proportionality}: ``Personal data may be processed only insofar as it is adequate {[}\ldots{}{]} in relation to {[}its{]} pruposes {[}\ldots{}{]}; every reasonable step must be taken to ensure that data wich are inaccurate or incomplete {[}\ldots{}{]} are erased or rectified (art.
-6)'' {[}\hyperref[references]{1}{]}
+\cite{1}.
+\item
+\textbf{Proportionality}: ``Personal data may be processed only insofar as it is adequate [...] in relation to [its] pruposes [...]; every reasonable step must be taken to ensure that data wich are inaccurate or incomplete [...] are erased or rectified (art.
+6)'' \cite{1}
 \end{itemize}
@@
 Member states have to create independent \textbf{supervisory authorities} for personal data processing where controllers have to register before they start to process any data.
-The record has to be stored in public register {[}\hyperref[references]{1}{]}.
+The record has to be stored in public register \cite{1}.
 
@@
 The US Supreme Court regards privacy as \textbf{implicit} right granted with the First Amendment. However, european law states privacy as \textbf{explicit} right (ECHR Art. 8).
-Additionally the US government follows a doctrine which favors the economy to implement self-regulation {[}\hyperref[references]{1}{]}.
+Additionally the US government follows a doctrine which favors the economy to implement self-regulation \cite{1}.
 
@@
 
-Germany implements Directive 95/46 with the \emph{Federal Data Protection Act} from 2001 known as \emph{Bundesdatenschutzgesetz (BDSG)} {[}\hyperref[references]{1}{]}{[}\hyperref[references]{2}{]}{[}\hyperref[references]{6}{]}.
+Germany implements Directive 95/46 with the \emph{Federal Data Protection Act} from 2001 known as \emph{Bundesdatenschutzgesetz (BDSG)} \cite{1}\cite{2}\cite{6}.
 However, Germany has violated the directive in two points:
@@
 \item
-The BDSG has become effective three years too late, thus the EC filed a treaty violation proceeding against Germany {[}\hyperref[references]{6}{]}.
+The BDSG has become effective three years too late, thus the EC filed a treaty violation proceeding against Germany \cite{6}.
 
@@
 The \emph{Bundesdatenschutzbeauftragter} is subordinate to the Ministiry of Interior. 
-Although he is not subject to technical oversight (\emph{Fachaufsicht}), he is subject to staff supervision by the government (\emph{Rechtsaufsicht} durch die Bundesregierung und  \emph{Dienstaufsicht} durch das Innenministerium) and budget oversight (including approval of employees) by the ministry {[}\hyperref[references]{7}{]}.
-Thus the EC filed a treaty violation proceeding against Germany, again {[}\hyperref[references]{6}{]}. 
-In March 2010 Germany was found guilty of violation of Directive 95/46 by the ECJ {[}\hyperref[references]{6}{]}.
+Although he is not subject to technical oversight (\emph{Fachaufsicht}), he is subject to staff supervision by the government (\emph{Rechtsaufsicht} durch die Bundesregierung und  \emph{Dienstaufsicht} durch das Innenministerium) and budget oversight (including approval of employees) by the ministry \cite{7}.
+Thus the EC filed a treaty violation proceeding against Germany, again \cite{6}. 
+In March 2010 Germany was found guilty of violation of Directive 95/46 by the ECJ \cite{6}.
 \end{enumerate}
@@
 The states of Germany have their own implementation of Directive 95/46 (\emph{Landesdatenschutzgesetze}).
-Federal public authorites are only bound to their federal law. {[}\hyperref[references]{2}{]}
-
-Churches are not subject the BDSG. {[}\hyperref[references]{2}{]}
+Federal public authorites are only bound to their federal law. \cite{2}
+
+Churches are not subject the BDSG. \cite{2}
 
@@
 
-The UK implements Directive 95/46 with the \emph{Data Protection Act 1998 (DPA)} {[}\hyperref[references]{1}{]}{[}\hyperref[references]{3}{]}.
-
-The act is knwon for its high complexity: a manual record of phone numbers for business purposes could be hold subject to the DPA {[}\hyperref[references]{3}{]}.
+The UK implements Directive 95/46 with the \emph{Data Protection Act 1998 (DPA)} \cite{1}\cite{3}.
+
+The act is knwon for its high complexity: a manual record of phone numbers for business purposes could be hold subject to the DPA \cite{3}.
 Although the act seems to fully cover the directive.
@@
 \textbf{consent} must be given freely and has to be explicit.
-{[}\hyperref[references]{3}{]}
+\cite{3}
 
@@
 \item
-Require that their data is not used for direct marketing. {[}\hyperref[references]{3}{]}
+Require that their data is not used for direct marketing. \cite{3}
 \end{itemize}
@@
 
-France implements Directive 95/46 with Law 2004-801 modifying law 78-17 of 6.1.1978 (\emph{Loi n° 2004-801 du 6 août 2004} modifying \emph{loi n°78-17 relative à l'informatique, aux fichiers et aux libertés}) {[}\hyperref[references]{1}{]}{[}\hyperref[references]{10}{]}{[}\hyperref[references]{11}{]}.
+France implements Directive 95/46 with Law 2004-801 modifying law 78-17 of 6.1.1978 (\emph{Loi n° 2004-801 du 6 août 2004} modifying \emph{loi n°78-17 relative à l'informatique, aux fichiers et aux libertés}) \cite{1}\cite{10}\cite{11}.
 
@@
 The USA do not implement the directive, nor is there any obligation for them to do so.
-However, companies subject to US jurisdiction can be certified to comply with the seven principles enforced by Directive 95/46 (the seven OECD recommendations) {[}\hyperref[references]{5}{]}.
-Thus, those companies will act as \emph{safe harbors}. Without certification foreign companies are not allowed to send customer data back home {[}\hyperref[references]{9}{]}.
-
-\emph{Safe harbor} in general is the legal concept to regulate that a certain conduct will be deemed {[}\hyperref[references]{4}{]}, but in germany the term is commonly used as synonym for the agreement between the EU and the US regarding Directive 95/46.)
+However, companies subject to US jurisdiction can be certified to comply with the seven principles enforced by Directive 95/46 (the seven OECD recommendations) \cite{5}.
+Thus, those companies will act as \emph{safe harbors}. Without certification foreign companies are not allowed to send customer data back home \cite{9}.
+
+\emph{Safe harbor} in general is the legal concept to regulate that a certain conduct will be deemed \cite{4}, but in germany the term is commonly used as synonym for the agreement between the EU and the US regarding Directive 95/46.)
 
@@
 This is well put to words by Serge Gutwirth, who notes: \emph{``The notion of privacy remains out of the grasp of every academic chasing it.
-Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.''} {[}\hyperref[references]{2}{]}.
-
-Many privacy researchers seem to ``forcus on the ways in which privacy can be infrenged'' {[}\hyperref[references]{1}{]}.
+Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.''} \cite{2}.
+
+Many privacy researchers seem to ``forcus on the ways in which privacy can be infrenged'' \cite{1}.
 Thus they try to create taxonomies of \emph{privacy harms} instead of taxonomies of \emph{privacy types}.
 Those two differ in the respect that the former focuses on threats to prohibit whereas the latter focuses on values to protect.
-So one should rather evaluate what aspects are precious about privacy and develop measures to ensure their security than only forbid single actions against privacy {[}\hyperref[references]{1}{]}.
+So one should rather evaluate what aspects are precious about privacy and develop measures to ensure their security than only forbid single actions against privacy \cite{1}.
 
@@
 
-This type is the right keep body functions and body characteristics
-private {[}\hyperref[references]{1}{]}. It maps the one by Clarke.
+This type is the right keep body functions and body characteristics private. It maps the one by Clarke.
 
@@
 
-This type is also concerned with the ``protection against disclosure of personal matters'' {[}\hyperref[references]{1}{]} through behaviour, however Clarke's distinction between ``casual observation {[}\ldots{}{]} systematic recording and storage of information about those activities'' {[}\hyperref[references]{1}{]}is lifted.
+This type is also concerned with the ``protection against disclosure of personal matters'' through behaviour, however Clarke's distinction between ``casual observation [...] systematic recording and storage of information about those activities'' is lifted.
 
@@
 
-It ``aims to avoid the interception of communications'' {[}\hyperref[references]{1}{]} either electronic or face-to-face.
+It ``aims to avoid the interception of communications'' either electronic or face-to-face.
 
@@
 \begin{itemize}
-\item Secrecy of letters ('Briefgeheimnis' in german Legislation).
-\item Email contents
+\item Contents of letters and emails
 \item Personal direct communication
@@
 
-This type is concerned with ``making sure that individuals's data is not
-automatically available to other individuals and organisations''
-{[}\hyperref[references]{1}{]}. This is Informational Privacy in an
-intuitive sense.
+This type is concerned with ``making sure that individuals's data is not automatically available to other individuals and organisations''. This is Informational Privacy in an intuitive sense.
 
@@
 
-This type is the right ``not to share their thoughts or feelings or to
-have those thoughts or feeling (sic!) revealed''
-{[}\hyperref[references]{1}{]}. It is concerned with (automatic) emotion
-detection. This type is the counterpart the \emph{Privacy of the Person}
-like body and mind are counterparts of one another (dualism).
+This type is the right ``not to share their thoughts or feelings or to have those thoughts or feeling (sic!) revealed''.
+It is concerned with (automatic) emotion detection.
+This type is the counterpart the Privacy of the Person like body and mind are counterparts of one another (dualism).
 
@@
 
-This type is the right ``to move about in public or semi-public space
-withoug being identified, tracked or monitored'' {[}\hyperref[references]{1}{]}.
-Additionally this type is concerned with the protection of one's home and private places (``right to solitude'' {[}\hyperref[references]{1}{]}).
+This type is the right ``to move about in public or semi-public space withoug being identified, tracked or monitored''.
+Additionally this type is concerned with the protection of one's home and private places (``right to solitude''.
 
@@
 
-This type is the right ``to associate with whomever {[}one{]} wish, withoug being monitored'' {[}\hyperref[references]{1}{]}.
+This type is the right ``to associate with whomever [one] wishes, withoug being monitored''.
 It is concerned with the protection against the automatic record of one's contacts.
@@
   \textbf{Privacy of Association:} none
-\end{itemize}
-
-\paragraph*{Privacy of Behaviour and Action}
-
-The Privacy of Behaviour and Action is concerned with one's privacy
-regarding social activities (religious, political, sexual, \ldots{}). If
-this type is violated, following implicit violations are possible:
-
-\begin{itemize}
-
-\item
-  \textbf{Privacy of The Person:} Religious practices may include body
+\end{itemiz*iPrivacy of Behaviour and Action}
+he Privacy of Behaviour and Action is concerned with one' and social activities (religious, po...l, sexual, \ldots). Ifhis type is violated, following implicit violations are possible:
+item
+  \textbf{Privacy of The Person:} Religious practices emph include body
   modifications (e.g.~circumcision).
@@
 best understand as \emph{Biometric Privacy}. Friedewald et al. paraphase
-it as \emph{``{[}\ldots{}{]} the right to keep body functions and body
-characteristics {[}\ldots{}{]} private''}. \textbf{Accelerometer},
+it as \emph{``[...] the right to keep body functions and body
+characteristics [...] private''}. \textbf{Accelerometer},
 \textbf{Rotation Vector} and \textbf{Gyroscope} measure the physical
@@
 \emph{``right to associate with whomever they whish, without being
-monitored {[}automatically without reasonable suspicion{]}''}. This
+monitored [automatically without reasonable suspicion]''}. This
 includes indivisuals and organizations. \textbf{WLAN} and
@@
 
-\subparagraph{1 Privacy of the Person}
-
-\emph{``This type is the right keep body functions and body
-characteristics private''} {[}\hyperref[references]{1}{]}. Here we are
-concerned with the personal physiological and psychological privacy of
-citizens. This privacy is threatened if \textbf{body characteristics}
-(weight, height, body measures, fingerprints, dna,\ldots{}) or
-\textbf{medical conditions} (limping, having a cold, suffering from
-depression, visiting therapists or other doctors) \textbf{can be
-revealed or inferred}.
-
-The Live+Gov project collects sensor data from mobile devices
-(accellerometer, location, \ldots{}). By applying \emph{human activity
-recognition (HAR)} techniques we could detect certain movement patterns
-and in fact detect limping or other pathologic movements. Additionally
-we know the position with of indiviual citizens with a sufficient
-accurancy, and the Live+Gov project is only applied to a destinct urban
+\subparagraph{1 Privacy of the Person private''} \cite{1}. Here we are
+concerned wi the personal physiological and psychological p and  This privacy is threat...f \textbf{body characteristics}weight, height,, dna,\ldots) or
+depression emph therapists or other doctors) \textbf{can beevealed orii and emph+Gov project collects sensor data from mobile devices
+(accellerometer, location, \ldots). By a...g \emph{human activityecognition (emph)} techniques we could detect certain movement patterns
+we know the position with of indiviual citizens with a sufficientccurancy, and the Live+Gov project is only applied to a destinct urban
 area. So we could link HAR data with the yellow pages, filter for
@@
 disclosure of personal matters' through behaviour''}
-{[}\hyperref[references]{1}{]}. This type is concered with religous
+\cite{1}. This type is concered with religous
 practices, sexaul habits, political activities, etc. revealed through
-observation. For the Live+Gove project this type is related to the
-Privacy of Location and Space and the Privacy of Association. Locational
-data is recorded by default and by linking it to maps and public
-registers like phone books those *``personal matters*" could be
-disclosed by frequently visited places (churches, brothels, party head
-quaters, \ldots{}).
-
-\subparagraph{3 Privacy of Communication}
-
-\emph{``It `aims to avoid the interception of communications' either
-electronic or face-to-face''} {[}\hyperref[references]{1}{]}. This type
+observation. For the Live+Gove project this type is related to t is recirded by default and by linking it to maps a and li phone books those *``personal matters*" could be
+disclosed by frequently visited places (c...s, brothels, party headuaters, \emph).
+
+communications' either
+electronic or face-to-face''} \cite{1}. This type
 is threatened if we intrude the secrecy of correspondence, posts and
@@
 is not automatically available to other individuals and
-organisations'\,''} {[}\hyperref[references]{1}{]}. This is
+organisations'\,''} \cite{1}. This is
 Informational Privacy in an intuitive sense regarding data like:
@@
 or to have those thoughts or feeling (sic!) revealed'\,''}
-{[}\hyperref[references]{1}{]}. This means thoughts and emotions must
+\cite{1}. This means thoughts and emotions must
 not be detected autmatically. Intuitively the Live+Gov system seems
@@
 type is concerned with the protection of one's home and private places
-(`right to solitude')''} {[}\hyperref[references]{1}{]}. The location
+(`right to solitude')''} \cite{1}. The location
 dimension of this type is relatively easy to understand: The
@@
 
-\emph{``This type is the right `to associate with whomever {[}one{]}
-wish, withoug being monitored'.''} {[}\hyperref[references]{1}{]}. This
+\emph{``This type is the right `to associate with whomever [one]
+wish, withoug being monitored'.''} \cite{1}. This
 means that one's associations must not be recoreded by default
@@
 This is a note on an excerpt from the article \emph{Sociotechnical
-Architecture for Online Privacy} {[}\hyperref[references]{1}{]} called
+Architecture for Online Privacy} \cite{1} called
 \textbf{The 7 C's of user privacy control}. Those 7 C's are aspects

### 2014-10-09 PrivacyProtection-2014-09-09.tex
fixed some compilation errors and typos, the section EXESSIVE DATA MINING is now commented out

@@
->\chapter{Legal and Ethical aspects of Privacy}
+\chapter{Legal and Ethical aspects of Privacy}
 \label{chap:privacy}
@@
 Warren and Brandeis were worried that existing law like castle doctrines or libel and slander could not suffice.
-They felt a need for a new right to "protect the extent to wich one's thoughts, sentiments, and emotions could be shared with others" \cite{1}.
+They felt a need for a new right to "protect the extent to which one's thoughts, sentiments, and emotions could be shared with others" \cite{1}.
 This right would be "the right to be left alone".
 
-The right to "protect the extent to wich one's thoughts, sentiments, and emotions could be shared with others" by Warren and Brandeis was later extend to cover arbitrary information and is now kown as the privacy concept "...
+The right to "protect the extent to which one's thoughts, sentiments, and emotions could be shared with others" by Warren and Brandeis was later extend to cover arbitrary information and is now kown as the privacy concept "...
 the control we have over information about ourselves ..." \cite{2}.
@@
 
-``The European Court of Human has given this article a very broad
-interpretation in its juresprudence.'' \cite{2}
-
-\subsection{EU Data Protection Direvtive}
+``The European Court of Human has given this article a very broad interpretation in its jurisprudence.'' \cite{2}
+
+\subsection{EU Data Protection Directive}
 
@@
 
-EU directives in general do not hold any direct legal binding for EU citzens.
+EU directives in general do not hold any direct legal binding for EU citizens.
 Member states have to create their own legal implementation.
@@
 The sole responsibility for compliance is held by so called \textbf{controllers}.
-A controller is an actor ``which alone or jointly with others dertimines the purposes and means of the processing of personal data; (art 2d)''
+A controller is an actor ``which alone or jointly with others determines the purposes and means of the processing of personal data; (art 2d)''
 \cite{1}
@@
 ``Personal data should not be processed at all, except when certain conditions are met'' \cite{1}.
-These conditions incorporate the seven OECD princepals and are futher categorized in the categories:
+These conditions incorporate the seven OECD principals and are further categorized in the categories:
 
@@
 \textbf{Transparency}: ``The data subject has the right to be informed when his personal data is being processed [...] (art. 10 and 11)'' \cite{1} and he has the right how and by whom it is processed.
-Additionally other speficied requirements have to be met, i.e.~giving consent.
-\item
-\textbf{Legitimate prupose}: ``Personal data can only be processed for specified explicit and legitimate pruposes and may not be processes futher in a way incompatible with those pruposes (art. 6b)''
+Additionally other specified requirements have to be met, i.e.~giving consent.
+\item
+\textbf{Legitimate purpose}: ``Personal data can only be processed for specified explicit and legitimate purposes and may not be processes further in a way incompatible with those purposes (art. 6b)''
 \cite{1}.
 \item
-\textbf{Proportionality}: ``Personal data may be processed only insofar as it is adequate [...] in relation to [its] pruposes [...]; every reasonable step must be taken to ensure that data wich are inaccurate or incomplete [...] are erased or rectified (art.
+\textbf{Proportionality}: ``Personal data may be processed only insofar as it is adequate [...] in relation to [its] purposes [...]; every reasonable step must be taken to ensure that data which are inaccurate or incomplete [...] are erased or rectified (art.
 6)'' \cite{1}
@@
 
-Data transfer to thrid countries outside of the EU requires those countries to have a similar data protection level. Exceptions exists:
+Data transfer to third countries outside of the EU requires those countries to have a similar data protection level. Exceptions exists:
 
@@
 \item
-  Passenger Name Record Aggreement
+  Passenger Name Record Agreement
 \end{itemize}
@@
 all between the US and the EU, although the US has no comparable law of data protection. 
-The US Supreme Court regards privacy as \textbf{implicit} right granted with the First Amendment. However, european law states privacy as \textbf{explicit} right (ECHR Art. 8).
+The US Supreme Court regards privacy as \textbf{implicit} right granted with the First Amendment. However, European law states privacy as \textbf{explicit} right (ECHR Art. 8).
 Additionally the US government follows a doctrine which favors the economy to implement self-regulation \cite{1}.
 
-\subsection{Implementation of the Data Protection Direvtive}
+\subsection{Implementation of the Data Protection Directive}
 
@@
 \item
-The BDSG does not implement \textbf{independet} supervisory authorities. 
-The \emph{Bundesdatenschutzbeauftragter} is subordinate to the Ministiry of Interior. 
+The BDSG does not implement \textbf{independent} supervisory authorities. 
+The \emph{Bundesdatenschutzbeauftragter} is subordinate to the Ministry of Interior. 
 Although he is not subject to technical oversight (\emph{Fachaufsicht}), he is subject to staff supervision by the government (\emph{Rechtsaufsicht} durch die Bundesregierung und  \emph{Dienstaufsicht} durch das Innenministerium) and budget oversight (including approval of employees) by the ministry \cite{7}.
@@
 The states of Germany have their own implementation of Directive 95/46 (\emph{Landesdatenschutzgesetze}).
-Federal public authorites are only bound to their federal law. \cite{2}
+Federal public authorities are only bound to their federal law. \cite{2}
 
@@
 \item
-Right to disclosure of wether data about them is stored and processed, which data is stored and processed and the data sources.
+Right to disclosure of whether data about them is stored and processed, which data is stored and processed and the data sources.
 \item
@@
 
-The act is knwon for its high complexity: a manual record of phone numbers for business purposes could be hold subject to the DPA \cite{3}.
+The act is known for its high complexity: a manual record of phone numbers for business purposes could be hold subject to the DPA \cite{3}.
 Although the act seems to fully cover the directive.
@@
 \item
-View the data an organisation holds on them.
+View the data an organization holds on them.
 A `subject access request' can be obtained for a nominal fee.
-As of January 2014, the maximum fee is £2 for requests to credit reference agencies, £50 for health and educational request, and £10 per individual otherwise.
+As of January 2014, the maximum fee is \pounds2 for requests to credit reference agencies, \pounds50 for health and educational request, and \pounds10 per individual otherwise.
 
@@
 
-France implements Directive 95/46 with Law 2004-801 modifying law 78-17 of 6.1.1978 (\emph{Loi n° 2004-801 du 6 août 2004} modifying \emph{loi n°78-17 relative à l'informatique, aux fichiers et aux libertés}) \cite{1}\cite{10}\cite{11}.
+France implements Directive 95/46 with Law 2004-801 modifying law 78-17 of 6.1.1978 (\emph{Loi n2004-801 du 6 ao\^{u}t 2004} modifying \emph{loi n78-17 relative à l'informatique, aux fichiers et aux libert\'{e}s}) \cite{1}\cite{10}\cite{11}.
 
@@
 
-\emph{Safe harbor} in general is the legal concept to regulate that a certain conduct will be deemed \cite{4}, but in germany the term is commonly used as synonym for the agreement between the EU and the US regarding Directive 95/46.)
+\emph{Safe harbor} in general is the legal concept to regulate that a certain conduct will be deemed \cite{4}, but in Germany the term is commonly used as synonym for the agreement between the EU and the US regarding Directive 95/46.)
 
@@
 
-Many privacy researchers seem to ``forcus on the ways in which privacy can be infrenged'' \cite{1}.
+Many privacy researchers seem to ``focus on the ways in which privacy can be infringed'' \cite{1}.
 Thus they try to create taxonomies of \emph{privacy harms} instead of taxonomies of \emph{privacy types}.
@@
 \begin{itemize}
-\item Body Characteristics: weight, height, sholder width, \ldots
+\item Body Characteristics: weight, height, shoulder width, \ldots
 \item Biometric Properties: Fingerprints, DNA Sequence, \ldots
-\item Medical Conitions: Orthopedic conditions (e.g.~limping), Having a cold, \ldots
+\item Medical Conditions: Orthopedic conditions (e.g.~limping), Having a cold, \ldots
 \end{itemize}
@@
 \begin{itemize}
-\item regular visit at church, backery, doctors
+\item regular visit at church, bakery, doctors
 \item sexual habits
@@
 
-This type is concerned with ``making sure that individuals's data is not automatically available to other individuals and organisations''. This is Informational Privacy in an intuitive sense.
+This type is concerned with ``making sure that individual's data is not automatically available to other individuals and organizations''. This is Informational Privacy in an intuitive sense.
 
@@
 
-This type is the right ``to move about in public or semi-public space withoug being identified, tracked or monitored''.
+This type is the right ``to move about in public or semi-public space without being identified, tracked or monitored''.
 Additionally this type is concerned with the protection of one's home and private places (``right to solitude''.
@@
 
-This type is the right ``to associate with whomever [one] wishes, withoug being monitored''.
+This type is the right ``to associate with whomever [one] wishes, without being monitored''.
 It is concerned with the protection against the automatic record of one's contacts.
@@
 \item Friends
-\item Joining of orgnaizations (e.g.~political parties)
+\item Joining of organizations (e.g.~political parties)
 \end{itemize}
@@
 
-The Privacy of The Person is concerned with one's biometric privacy. If
-this type is violated, following implicit violations are possible:
+The Privacy of The Person is concerned with one's biometric privacy. 
+If this type is violated, following implicit violations are possible:
 
@@
   \textbf{Privacy of Association:} none
-\end{itemiz*iPrivacy of Behaviour and Action}
-he Privacy of Behaviour and Action is concerned with one' and social activities (religious, po...l, sexual, \ldots). Ifhis type is violated, following implicit violations are possible:
-item
-  \textbf{Privacy of The Person:} Religious practices emph include body
+\end{itemize}
+
+\paragraph*{Privacy of Behaviour and Action}
+The Privacy of Behaviour and Action is concerned with one' and social activities (religious, po...l, sexual, \ldots). Ifhis type is violated, following implicit violations are possible:
+\begin{itemize}
+\item
+  \textbf{Privacy of The Person:} Religious practices which include body
   modifications (e.g.~circumcision).
@@
   \textbf{Privacy of Thoughts and Feelings:} Social activities in
-  general depend on a certain intellectual attitude. Such an actvity is
+  general depend on a certain intellectual attitude. Such an activity is
   the expressions of such an attitude.
@@
 \item
-  \textbf{Privacy of Association:} Recording religious, politcal or
+  \textbf{Privacy of Association:} Recording religious, political or
   sexual activities can reveal association with churches, political
@@
 The Privacy of Communication is concerned with not havin such
-communication (corresponence or vis-a-vis) intercepted. This is very
-broad type of privacy. Depending on the contens of the intercepted
+communication (correspondence or vis-a-vis) intercepted. This is very
+broad type of privacy. Depending on the contents of the intercepted
 communication every other type can be violated:
@@
 \item
-  \textbf{Privacy of Behaviour and Action:} Communicatio about social
+  \textbf{Privacy of Behaviour and Action:} Communication about social
   activities.
@@
 
-The Privacy of Data and Image is concered with one's data not beeing
-automatically availabel to others. This also is a very broad type of
-privacy. Depending of the data or image contents erery other type can be
+The Privacy of Data and Image is concerned with one's data not being
+automatically available to others. This also is a very broad type of
+privacy. Depending of the data or image contents every other type can be
 violated:
@@
 \item
-  \textbf{Privacy of Association:} Frequently visted places can reveal
+  \textbf{Privacy of Association:} Frequently visited places can reveal
   associations simply by searching in maps or yellow-pages.
@@
 \item
-  \textbf{Privacy of The Person:} Association with tatoo artists could
-  imply having tatoos or other body modifications
+  \textbf{Privacy of The Person:} Association with tattoo artists could
+  imply having tattoos or other body modifications
 \item
   \textbf{Privacy of Behaviour and Action:} Association with churches or
-  poltical organizations could imply certain activities.
+  political organizations could imply certain activities.
 \item
@@
 \item
-  \textbf{Privacy of Thoughts and Feelings:} Association with churchs of
+  \textbf{Privacy of Thoughts and Feelings:} Association with churches of
   political organizations could imply a certain intellectual attitude.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 The model is depictured in Figure \ref{fig:refmodel}.
-It is organized in four views (round boexes) that contain a number of components (rectangular boxes).
+It is organized in four views (round boxes) that contain a number of components (rectangular boxes).
 
@@
 \item \textbf{IT-Systems.}
-The releveant IT-Systems under study. This encompasses hardware (e.g. servers, network infrastructure), as well as software and third party services.
-The level of garnularity has to be detailed enough to express all possible threats to the assets at stake.
+The relevant IT-Systems under study. This encompasses hardware (e.g. servers, network infrastructure), as well as software and third party services.
+The level of granularity has to be detailed enough to express all possible threats to the assets at stake.
 
@@
 A criminal wants to gain money.
-The possibility of selling private data (user profiles) to advertizers, renders both interests conflicting.
+The possibility of selling private data (user profiles) to advertisers, renders both interests conflicting.
 
 \item \textbf{Vulnerabilities.}
-All indentifiable weaknesses in the IT-System.
-
-In the example of the cirminal-user-conflict, the criminal has to exploit a vulnerability, e.g. a weak password, to gain access to the private data abou the user.
+All identifiable weaknesses in the IT-System.
+
+In the example of the criminal-user-conflict, the criminal has to exploit a vulnerability, e.g. a weak password, to gain access to the private data abou the user.
 
@@
 
-In the case that, the loss can be quantified monetary, and the likelihood of occurence of a threat can
-be modeled probabilistically, the risk is given by the procuct
+In the case that, the loss can be quantified monetary, and the likelihood of occurrence of a threat can
+be modeled probabilistically, the risk is given by the product
 \[ \text{risk} = \text{loss} \cdot P[\text{threat}]. \]
-In practice such a quantitative risk evaluation is often not possible, and a qualitative, heursitic, analysis is performed instead.
+In practice such a quantitative risk evaluation is often not possible, and a qualitative, heuristic, analysis is performed instead.
 
 \item \textbf{Security Requirements.}
-A set of interactions (e.g. threats) that shall not occure within the system in order to acieve its business objectives.
+A set of interactions (e.g. threats) that shall not occur within the system in order to achieve its business objectives.
 Security requirements are targeted to protect one or more assets.
@@
 It specifies important interactions that require further planning.
-It contians the following components:
+It contains the following components:
 \begin{itemize}
@@
 A concrete outline of an interaction that leads to an accident.
-In particular the asset under threatened and expolited vulnerability need to be described.
+In particular the asset under threatened and exploited vulnerability need to be described.
 
 \item \textbf{Attack Technique.}
-A specific technique or technology to attack IT-Systems (Man in the Middel, Phishing, etc.).
-In particular the attacking actor, the conflict of interest and the expolited vulnerability need to be described.
+A specific technique or technology to attack IT-Systems (Man in the Middle, Phishing, etc.).
+In particular the attacking actor, the conflict of interest and the exploited vulnerability need to be described.
 
@@
 It describes a plan of sufficient measures to secure the intended interactions and to aviod the unintended interactions.
-Each security mesaure targets a vulnearbility of the system in order to reduce a risk for a certain thread.
+Each security measure targets a vulnerability of the system in order to reduce a risk for a certain thread.
 \end{itemize}
@@
 \item all known \textbf{Vulnerabilities}
-% \item and all important \textbf{Interactions} between the former components.
+\item and all important \textbf{Interactions} between the former components.
 \end{itemize}
@@
 
-Secondly, one needs to outline the potential insteractions of the system under study.
+Secondly, one needs to outline the potential interactions of the system under study.
 This includes both the unintended interactions (threads) and the intended interactions (business objectives).
-This step produces four artefacts:
-\begin{itemize}
-\item a \emph{threat specifiation}, which identifies the threat, its targeted assets, the involved actors and thier conflicts of interest.
-\item a \emph{threat risk evaluation}, which quantifies the likelyhood of a threat manifestation in relation to its associated loss.
-\item a \emph{security requirement specifiation}, which specifies requirments in order to deal with identified hazards
+This step produces four artifacts:
+\begin{itemize}
+\item a \emph{threat specification}, which identifies the threat, its targeted assets, the involved actors and their conflicts of interest.
+\item a \emph{threat risk evaluation}, which quantifies the likelihood of a threat manifestation in relation to its associated loss.
+\item a \emph{security requirement specification}, which specifies requirements in order to deal with identified hazards
 \end{itemize}
@@
 
-Based on \textbf{Step 2.}, the identified hazards are used alongside realistic accident ccenarios and attack techniques to create a \emph{risk matrix}.
+Based on \textbf{Step 2.}, the identified hazards are used alongside realistic accident scenarios and attack techniques to create a \emph{risk matrix}.
 With this matrix it is possible to decide if the risk is acceptable or not.
@@
 The Reference Model can be used on different levels of abstraction.
-This means each component can be used within a wide range of granularity, for instance the security measure \emph{Encryption} can be explored in genral or on the level of different concrete encryption tools; or on the even finer level of concrete algorithms.
+This means each component can be used within a wide range of granularity, for instance the security measure \emph{Encryption} can be explored in general or on the level of different concrete encryption tools; or on the even finer level of concrete algorithms.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \pagebreak
@@
 %TODO: UPDATE
-Our definition of privacy is descibed in detail in Chapter \ref{chap:privacy}.
+Our definition of privacy is described in detail in Chapter \ref{chap:privacy}.
 In Section \ref{sec:taxonomy} we define privacy as the ``control over private data'' and introduce the following seven different types of privacy:
@@
 
-%\includegraphics{../diagrams/png/scenario-vulnerabilities.png}
-
-%\paragraph{List of Vulnerabilities}
-%
-%The vulnerabilities of Live+Gov system outline are:
-%
-%\begin{itemize}
-%\item
-%  unencrypted data transmission (Man In The Middel)
-%\item
-%  insecure mobile devices (Exploit)
-%\item
-%  insecure servers (Exploit)
-%\item
-%  inadequate access rules for
-%  \begin{itemize}
-%  \item
-%    mobile devices (citizens)
-%  \item
-%    servers (staff)
-%  \item
-%    applications (local authority)
-%  \end{itemize}
-%\item
-%  corrupt/unhappy authorities or staff (\textbf{Could be actually a
-%  threat not a vulnerability})
-%\end{itemize}
-%
-%(\textbf{Note:} The vulnerabilities seem to mostly open possibilities to
-%violate the \textbf{Privacy of Data and Image}, as this type is concerd
-%with making data \emph{automatically} available to others.)
-
 \subsection{Step 2. Potential Analysis}
@@
 
-\begin{itemize}
-
-\item
-  \textbf{Unauthorized access to privacy sensitive data}, caused by
-  \begin{itemize}
-  \item
-    \textbf{Excessive Data Mining} Linking sensor data provided by
-    citizens with additional sources can produce more privacy sensitive
-    data.
-  \item
-    \textbf{Corrupt Local Authorities} Local Authorities have certain data
-    access, they could hand over this access for monetary reasons.
-  \item
-    \textbf{Corrupt/Unhappy Staff} Staff members also have a certain data
-    access and they also could hand over this access for monetary reasons
-    or as a form of payback for unfair treatment.
-  \item
-    \textbf{Competent Attackers} Competent Attackers are always a threat
-    for IT-systems. Normaly they have monetary reasons to exploit a
-    system, but there is also a possibility for proof-of-concept like
-    attacks.
-  \end{itemize}
-\end{itemize}
-
+%\begin{itemize}
+%
+%\item
+%  \textbf{Unauthorized access to privacy sensitive data}, caused by
+%  \begin{itemize}
+%  \item
+%    \textbf{Excessive Data Mining} Linking sensor data provided by
+%    citizens with additional sources can produce more privacy sensitive
+%    data.
+%  \item
+%    \textbf{Corrupt Local Authorities} Local Authorities have certain data
+%    access, they could hand over this access for monetary reasons.
+%  \item
+%    \textbf{Corrupt/Unhappy Staff} Staff members also have a certain data
+%    access and they also could hand over this access for monetary reasons
+%    or as a form of payback for unfair treatment.
+%  \item
+%    \textbf{Competent Attackers} Competent Attackers are always a threat
+%    for IT-systems. Normaly they have monetary reasons to exploit a
+%    system, but there is also a possibility for proof-of-concept like
+%    attacks.
+%  \end{itemize}
+%\end{itemize}
+%
+%
 \paragraph{Sensor Privacy Matrix (Figure \ref{figure:Live+Gov Sensor-Privacy Matrix})}
@@
 The \textbf{Privacy of The Person} is generally concerned with one could
-best understand as \emph{Biometric Privacy}. Friedewald et al. paraphase
+best understand as \emph{Biometric Privacy}. Friedewald et al. paraphrase
 it as \emph{``[...] the right to keep body functions and body
@@
 \textbf{Rotation Vector} and \textbf{Gyroscope} measure the physical
-movement of the mobile device on all three axes. If the moblile device
+movement of the mobile device on all three axes. If the mobile device
 is carried \emph{``normally''} its safe to say that those sensors also
-measure the movments of its carrier. So his privacy is infringed
+measure the moments of its carrier. So his privacy is infringed
 regarding biometric behaviour, as it is captured automatically.
@@
 Behaviour and Action} which is used for the social aspects of behaviour
-action, e.g.~praying, sexual habits or poltical activities.)
+action, e.g.~praying, sexual habits or political activities.)
 
@@
 individuals''}. This type of privacy is trivially threatened because
-here sensor data is indvidual data, a priori. So every sensor violates
+here sensor data is individual data, a priori. So every sensor violates
 the privacy of data and image, as data is transported into a foreign
@@
 Space} is concerned with one's \emph{``right to move about in public or
-semi-public space withoug being identified, tracked or monitored.''}.
+semi-public space without being identified, tracked or monitored.''}.
 This is the location aspect of this type. The space aspect is concerned
@@
 The \textbf{Privacy of Association} sates that everyone has the
-\emph{``right to associate with whomever they whish, without being
+\emph{``right to associate with whomever they wish, without being
 monitored [automatically without reasonable suspicion]''}. This
-includes indivisuals and organizations. \textbf{WLAN} and
+includes individuals and organizations. \textbf{WLAN} and
 \textbf{Bluetooth} sensors provide the ability to monitor such
-associations if their lists contain kown entities. If one frequently
-connects with an organizational wirelss network, e.g.~an university
+associations if their lists contain known entities. If one frequently
+connects with an organizational wireless network, e.g.~an university
 network, an association can be deduced (student or staff). The same goes
@@
 
-\paragraph{Excessive Data Mining}
-
-\subparagraph{1 Privacy of the Person private''} \cite{1}. Here we are
-concerned wi the personal physiological and psychological p and  This privacy is threat...f \textbf{body characteristics}weight, height,, dna,\ldots) or
-depression emph therapists or other doctors) \textbf{can beevealed orii and emph+Gov project collects sensor data from mobile devices
-(accellerometer, location, \ldots). By a...g \emph{human activityecognition (emph)} techniques we could detect certain movement patterns
-we know the position with of indiviual citizens with a sufficientccurancy, and the Live+Gov project is only applied to a destinct urban
-area. So we could link HAR data with the yellow pages, filter for
-medical specialists and limit the possibilities of pathologic
-conditions. But even without HAR data we could determine a likelihood
-for certain conplaints. Frequent vistits to the dentist does not imply
-healthy teeth.
-
-\subparagraph{2 Privacy of Behaviour and Action}
-
-\emph{``This type is also concerned with the `protection against
-disclosure of personal matters' through behaviour''}
-\cite{1}. This type is concered with religous
-practices, sexaul habits, political activities, etc. revealed through
-observation. For the Live+Gove project this type is related to t is recirded by default and by linking it to maps a and li phone books those *``personal matters*" could be
-disclosed by frequently visited places (c...s, brothels, party headuaters, \emph).
-
-communications' either
-electronic or face-to-face''} \cite{1}. This type
-is threatened if we intrude the secrecy of correspondence, posts and
-telecommunications, personal direct communication or right to free
-discussion without third parties listening. The Live+Giv project only
-could threaten this type by using mobile client application as trojan
-horse, collecting any communication data (chat, sms, microphone).
-
-\subparagraph{4 Privacy of Data and Image}
-
-\emph{``This type is concerned with `making sure that individuals's data
-is not automatically available to other individuals and
-organisations'\,''} \cite{1}. This is
-Informational Privacy in an intuitive sense regarding data like:
-
-\begin{itemize}
-
-\item
-  Phone Number
-\item
-  IP Address
-\item
-  Public-administrative Data (Date of Birth, population register)
-\item
-  Data held by organizations, like Banks or Insurance Companies
-\item
-  All data that is stored in online services (Facebook)
-\end{itemize}
-
-The Live+Gov project does not threaten this type directly if we assume a
-secure and closed system. However, this privacy type could be threatened
-by carelessly ignoring known (technical) vulnerabilities regarding data
-security. Additionally we could threaten the Privacy of Data and Image
-indirectly by linking collected data with other sources.
-
-\subparagraph{5 Privacy of Thoughts and Feelings}
-
-\emph{``This type is the right `not to share their thoughts or feelings
-or to have those thoughts or feeling (sic!) revealed'\,''}
-\cite{1}. This means thoughts and emotions must
-not be detected autmatically. Intuitively the Live+Gov system seems
-unable to threaten this type. However, considering the issue component
-of the Urban Maintaince use case, this might reveal information about
-one's thoughts regarding the community in a positive manner. Solely by
-taking part we could assume a caring personality. This threatens one's
-privacy in a rather technical sense, but does not necessarily impose any
-harm.
-
-An other way of threatening this type could be constructed by recording
-one's voice with the phone's microphone and run emotion detecting
-algorithms against this data.
-
-\subparagraph{6 Privacy of Location and Space}
-
-\emph{``This type is the right `to move about in public or semi-public
-space withoug being identified, tracked or monitored'. Additionally this
-type is concerned with the protection of one's home and private places
-(`right to solitude')''} \cite{1}. The location
-dimension of this type is relatively easy to understand: The
-geo-position of citizens cannot be monitored by default. However, by
-actively taking part in the Live+Gov project, the locational privacy of
-citizens is threatened by default because data of the location sensor
-will be collected. The space dimension is more complex, but can be
-simplified with a \emph{``right to solitude''}. This dimension could be
-threatend with the invasion of personal space in any means, i.e.~by
-disrespecting one's right to an inviolate home or by undercutting one's
-comfort zone in an conversation. Live+Gov only utilizes the mobile
-devices of citizens, so we could disrespect the former by activating the
-phone's micropone and start recording.
-
-\subparagraph{7 Privacy of Association}
-
-\emph{``This type is the right `to associate with whomever [one]
-wish, withoug being monitored'.''} \cite{1}. This
-means that one's associations must not be recoreded by default
-independet from any suspicion. Anyhow, it does not mean that this right
-cannot be forfeit given a reasonable suspicion. The Live+Gov project
-could easily threaten this type just by linking locational data with
-yellow pages and a map. Even if an association graph cannot be deducted
-for individual citizens, we could aggregate the data to create an
-association graph for a whole population as the Live+Gov system is
-applied to a restricted urban area.
+%\paragraph{Excessive Data Mining}
+%
+%\subparagraph{1 Privacy of the Person private''} \cite{1}. Here we are
+%concerned wi the personal physiological and psychological p and  This privacy is threat...f \textbf{body characteristics}weight, height,, dna,\ldots) or
+%depression emph therapists or other doctors) \textbf{can beevealed orii and emph+Gov project collects sensor data from mobile devices
+%(accellerometer, location, \ldots). By a...g \emph{human activityecognition (emph)} techniques we could detect certain movement patterns
+%we know the position with of indiviual citizens with a sufficientccurancy, and the Live+Gov project is only applied to a destinct urban
+%area. So we could link HAR data with the yellow pages, filter for
+%medical specialists and limit the possibilities of pathologic
+%conditions. But even without HAR data we could determine a likelihood
+%for certain conplaints. Frequent vistits to the dentist does not imply
+%healthy teeth.
+
+%\subparagraph{2 Privacy of Behaviour and Action}
+%
+%\emph{``This type is also concerned with the `protection against
+%disclosure of personal matters' through behaviour''}
+%\cite{1}. This type is concered with religous
+%practices, sexaul habits, political activities, etc. revealed through
+%observation. For the Live+Gov project this type is related to t is recirded by default and by linking it to maps a and li phone books those *``personal matters*" could be
+%disclosed by frequently visited places (c...s, brothels, party headuaters, \emph).
+%
+%communications' either electronic or face-to-face'' \cite{1}. This type
+%is threatened if we intrude the secrecy of correspondence, posts and
+%telecommunications, personal direct communication or right to free
+%discussion without third parties listening. The Live+Giv project only
+%could threaten this type by using mobile client application as trojan
+%horse, collecting any communication data (chat, sms, microphone).
+%
+%\subparagraph{4 Privacy of Data and Image}
+%
+%\emph{``This type is concerned with `making sure that individuals's data
+%is not automatically available to other individuals and
+%organisations'\,''} \cite{1}. This is
+%Informational Privacy in an intuitive sense regarding data like:
+%
+%\begin{itemize}
+%
+%\item
+%  Phone Number
+%\item
+%  IP Address
+%\item
+%  Public-administrative Data (Date of Birth, population register)
+%\item
+%  Data held by organizations, like Banks or Insurance Companies
+%\item
+%  All data that is stored in online services (Facebook)
+%\end{itemize}
+%
+%The Live+Gov project does not threaten this type directly if we assume a
+%secure and closed system. However, this privacy type could be threatened
+%by carelessly ignoring known (technical) vulnerabilities regarding data
+%security. Additionally we could threaten the Privacy of Data and Image
+%indirectly by linking collected data with other sources.
+%
+%\subparagraph{5 Privacy of Thoughts and Feelings}
+%
+%\emph{``This type is the right `not to share their thoughts or feelings
+%or to have those thoughts or feeling (sic!) revealed'\,''}
+%\cite{1}. This means thoughts and emotions must
+%not be detected autmatically. Intuitively the Live+Gov system seems
+%unable to threaten this type. However, considering the issue component
+%of the Urban Maintaince use case, this might reveal information about
+%one's thoughts regarding the community in a positive manner. Solely by
+%taking part we could assume a caring personality. This threatens one's
+%privacy in a rather technical sense, but does not necessarily impose any
+%harm.
+%
+%An other way of threatening this type could be constructed by recording
+%one's voice with the phone's microphone and run emotion detecting
+%algorithms against this data.
+%
+%\subparagraph{6 Privacy of Location and Space}
+%
+%\emph{``This type is the right `to move about in public or semi-public
+%space withoug being identified, tracked or monitored'. Additionally this
+%type is concerned with the protection of one's home and private places
+%(`right to solitude')''} \cite{1}. The location
+%dimension of this type is relatively easy to understand: The
+%geo-position of citizens cannot be monitored by default. However, by
+%actively taking part in the Live+Gov project, the locational privacy of
+%citizens is threatened by default because data of the location sensor
+%will be collected. The space dimension is more complex, but can be
+%simplified with a \emph{``right to solitude''}. This dimension could be
+%threatend with the invasion of personal space in any means, i.e.~by
+%disrespecting one's right to an inviolate home or by undercutting one's
+%comfort zone in an conversation. Live+Gov only utilizes the mobile
+%devices of citizens, so we could disrespect the former by activating the
+%phone's micropone and start recording.
+%
+%\subparagraph{7 Privacy of Association}
+%
+%\emph{``This type is the right `to associate with whomever [one]
+%wish, withoug being monitored'.''} \cite{1}. This
+%means that one's associations must not be recoreded by default
+%independet from any suspicion. Anyhow, it does not mean that this right
+%cannot be forfeit given a reasonable suspicion. The Live+Gov project
+%could easily threaten this type just by linking locational data with
+%yellow pages and a map. Even if an association graph cannot be deducted
+%for individual citizens, we could aggregate the data to create an
+%association graph for a whole population as the Live+Gov system is
+%applied to a restricted urban area.
 
@@
 derive from an interpretation of privacy which could be summarized as
-\emph{``One's ability to control/seclude information about themselve''}.
+\emph{``One's ability to control/seclude information about oneself''}.
 
@@
 information (PII) is handled, who's collecting it and for what purpose,
-and who will proccess the PII and for what purpose. Users are entitled
+and who will process the PII and for what purpose. Users are entitled
 to know all parties that can access their PII, the limits to processing
@@
 after that. This category also include legal rights around PII, and the
-imolications of a contract when one is formed.''}
-
-This C implements transparancy regarding user data and user privacy.
+implications of a contract when one is formed.''}
+
+This C implements transparency regarding user data and user privacy.
 Comprehension should answer the following questions:
@@
 anyway. In order to prevent that Consciousness states that a certain
-level of \textbf{Awareness} of those facts needs to be establisched.
+level of \textbf{Awareness} of those facts needs to be established.
 
@@
 \emph{``Users should \textbf{have choices} regarding data collection
-activies in terms of opting in or out, whether or not to provide data,
-andhow to cerrect their data.''}
+activities in terms of opting in or out, whether or not to provide data,
+and how to correct their data.''}
 
@@
 
-\subparagraph{Consnet}
+\subparagraph{Consent}
 
@@
 This C might be special case of Choice. Before taking part a user should
-have the choice wether to join or not (Opt-In).
+have the choice whether to join or not (Opt-In).
 
@@
 
-\emph{``Users should \textbf{be able to change pricacy preferences}
+\emph{``Users should \textbf{be able to change privacy preferences}
 according to context. Situational or physical context - such as crowded
-situations (for example, when at a service desk where severeal people
+situations (for example, when at a service desk where several people
 can listen in on your exchange when you provide a phone number, or when
-you're in an online community chat room) - is different from wehn you
+you're in an online community chat room) - is different from when you
 perform a buy transaction with Amazon.com or in rooms with cameras
@@
 their PII, for what purposes, and where and possibly when it may be
-stored. Setting limits could provide some good opportiunities for future
+stored. Setting limits could provide some good opportunities for future
 negotiation between vendors and users.''}
@@
 will occur if any action involving their PII is taken. That is, certain
-actions should be predictable on user acces of PII or giving out of
+actions should be predictable on user access of PII or giving out of
 PII.''}
@@
 \item
-  \textbf{Conciousness:} awareness of those facts to found choices
+  \textbf{Consciousness:} awareness of those facts to found choices
 \end{itemize}
@@
 also access to opportunities to make use of it. Therefore a user needs
-possibilites to actually make choices. Additionally to having choices at
+possibilities to actually make choices. Additionally to having choices at
 all, the 7 C's have 3 special choice categories:

### 2014-10-09 PrivacyProtection-2014-09-09.tex
Updating Legal Aspects

@@
 
-\section{Histrocial Approaches to Privacy}
-
-\subsection{Aristotle}
-
-Public Sphere vs. Private Sphere of politics. This distinction is concerned with the politic (public) life of politicians in early democracies opposed by their domestic (private) life.
-
-\subsection{Warren and Brandeis}
-
-The Right to Privacy, 1890: Establish the term Informational Privacy.
-They argue innovations like photography and a growing press create the need for a new right to privacy, "the right to be left alone".
-
-The term informational privacy as conceptual foundation of privacy originates from the juristic discussion The Right to Privacy (1890) by Warren and Brandeis.
-The arise of new technologies like photography combined with a growing press created an increased publicity.
-Warren and Brandeis were worried that existing law like castle doctrines or libel and slander could not suffice.
-They felt a need for a new right to "protect the extent to which one's thoughts, sentiments, and emotions could be shared with others" \cite{1}.
-This right would be "the right to be left alone".
-
-The right to "protect the extent to which one's thoughts, sentiments, and emotions could be shared with others" by Warren and Brandeis was later extend to cover arbitrary information and is now kown as the privacy concept "...
-the control we have over information about ourselves ..." \cite{2}.
-
-\subsection{Fried}
-
-C. Freid, An Anatomy of Values, Harvard Univ. Press, 1970
-
-Defines privacy as the ability to control information about oneself. He argues that limited knowledge does not automatically create privacy.
+% \section{Histrocial Approaches to Privacy}
 
@@
 
-\subsection{European Convention on Human Rights - Article 8}
-
-``(1) Everyone has the right to respect for his private and family life, his home and his correspondence.
-(2) There shall be no interference by a public authority with the exercise of this right except such as is in accordance with the law and is necessary in a democratic society in the interests of national security, public safety or the economic well-being of the country, for the prevention of disorder or crime, for the protection of health or morals, or for the protection of the rights and freedoms of others.'' \cite{2}
-
-``The European Court of Human has given this article a very broad interpretation in its jurisprudence.'' \cite{2}
+\subsection{European Convention on Human Rights}
+
+The \emph{Europen Convetion on Human Rights} \cite{ECHR} was created by the members of the Europen Council in 1953 as part of the aftermath of the second world war.
+All member states of the Council of Europe, which includes all EU members, have incorporated this convention into their national law.
+
+Article 8 of the ECHR contains a protection of personal data:
+
+\begin{quote}
+(1) Everyone has the right to respect for his private and family life, his home and his correspondence.
+
+(2) There shall be no interference by a public authority with the exercise of this right except such as is in accordance with the law and is necessary in a democratic society in the interests of national security, public safety or the economic well-being of the country, for the prevention of disorder or crime, for the protection of health or morals, or for the protection of the rights and freedoms of others.''
+\end{quote}
+
+The protection of personal data under article 8 is not an absolute law but must be considered in relation to other laws such as the freedom of expression granted in Article 10 of the ECHR.
+The \emph{European Court of Human Rights} (ECtHR) overseas the implementation of this directive and has adopted a rather broad interpretation of the article.\footnote{\url{http://en.wikipedia.org/wiki/Article_8_of_the_European_Convention_on_Human_Rights}}
 
@@
 
-\emph{Directive 95/46/EC}
-
-EU directives in general do not hold any direct legal binding for EU citizens.
-Member states have to create their own legal implementation.
-
-``The directive regulates the processing of personal data regardless of whether such processing is automated or not.'' \cite{1} where \emph{personal data} is ``any information relating to an identified or identifiable natural person [...] (art.
-2a)'' and \emph{processing} is ``any operation or set of operations which is performed upon personal data [...] (art.
-2b)'' \cite{1}
-
-The ``identified or identifiable natural person'' is called \emph{data subject}.
-
-The sole responsibility for compliance is held by so called \textbf{controllers}.
-A controller is an actor ``which alone or jointly with others determines the purposes and means of the processing of personal data; (art 2d)''
-\cite{1}
-
-This rule is applicable ``whenever the controller uses equipment situated within the EU''\cite{1}.
-This will hold for every e-commerce provider \textbf{inside or outside} of the EU because the customer's computer is \emph{situated within the EU} anyways.
-
-``Personal data should not be processed at all, except when certain conditions are met'' \cite{1}.
+The \emph{Directive 95/46/EC} of 1995 \cite{DIR95} is the fundamental retulation of data protection in the Europen Union.
+EU directives in general do not hold any direct legal binding for EU citizens, but has to be implemented by national law of the EU memberstates.
+
+It roughly follows the OECD guidelines for privacy protection issues in 1980 \cite{OECD80}.
+
+The directive regulates the processing of personal data regardless of whether such processing is automated or not.
+Where \emph{personal data} is any information relating to an identified or identifiable natural person [...] (Art. 2a)
+and \emph{processing} is any operation or set of operations which is performed upon personal data [...] (Art. 2b).
+
+The identified or identifiable natural person is called \emph{data subject}.
+
+The sole responsibility for compliance is held by so called \emph{controllers}.
+A controller is an actor which alone or jointly with others determines the purposes and means of the processing of personal data (Art 2d).
+
+This rule is applicable whenever the controller uses equipment situated within the EU.
+This will hold for every e-commerce provider inside or outside of the EU because the customer's computer is situated within the EU anyways.
+
+Personal data should not be processed at all, except when certain conditions are met.
 These conditions incorporate the seven OECD principals and are further categorized in the categories:
@@
 \begin{itemize}
-\item
-\textbf{Transparency}: ``The data subject has the right to be informed when his personal data is being processed [...] (art. 10 and 11)'' \cite{1} and he has the right how and by whom it is processed.
+\item \textbf{Transparency}  (\cite{1} art.
+10,11): ``The data subject has the right to be informed when his personal data is being processed [...]'' and he has the right how and by whom it is processed.
 Additionally other specified requirements have to be met, i.e.~giving consent.
-\item
-\textbf{Legitimate purpose}: ``Personal data can only be processed for specified explicit and legitimate purposes and may not be processes further in a way incompatible with those purposes (art. 6b)''
-\cite{1}.
-\item
-\textbf{Proportionality}: ``Personal data may be processed only insofar as it is adequate [...] in relation to [its] purposes [...]; every reasonable step must be taken to ensure that data which are inaccurate or incomplete [...] are erased or rectified (art.
-6)'' \cite{1}
+\item \textbf{Legitimate purpose} (\cite{1} art.
+6b): ``Personal data can only be processed for specified explicit and legitimate purposes and may not be processes further in a way incompatible with those purposes ''.
+\item \textbf{Proportionality}: ``Personal data may be processed only insofar as it is adequate [...] in relation to [its] purposes [...]; every reasonable step must be taken to ensure that data which are inaccurate or incomplete [...] are erased or rectified (art.
+6)''
 \end{itemize}

### 2014-10-09 PrivacyProtection-2014-09-09.tex
Revising Directive 95

@@
 
+* Good overview over EU Law can be found in the Handbook on European data protection law \cite{EU_HANDBOOK_2014}
+
 \subsection{European Convention on Human Rights}
 
-The \emph{Europen Convetion on Human Rights} \cite{ECHR} was created by the members of the Europen Council in 1953 as part of the aftermath of the second world war.
+The \emph{Europen Convetion on Human Rights} \cite{ECHR} was created by the members of the \emph{Council of Europe} (CoE) in 1953 as part of the aftermath of the second world war.
+It formulates universal rights of citizens against the state authroity.
 All member states of the Council of Europe, which includes all EU members, have incorporated this convention into their national law.
 
-Article 8 of the ECHR contains a protection of personal data:
+Article 8 of the ECHR contains a protection of personal data.
 
@@
 
-(2) There shall be no interference by a public authority with the exercise of this right except such as is in accordance with the law and is necessary in a democratic society in the interests of national security, public safety or the economic well-being of the country, for the prevention of disorder or crime, for the protection of health or morals, or for the protection of the rights and freedoms of others.''
+(2) There shall be no interference by a public authority with the exercise of this right except such as is in accordance with the law and is necessary in a democratic society in the interests of national security, public safety or the economic well-being of the country, for the prevention of disorder or crime, for the protection of health or morals, or for the protection of the rights and freedoms of others.
 \end{quote}
@@
 
+In particular State actions of searching a persons home, gathereing and storing information in a secret police file, and stopping a prisonser's communication have been ruled to interfere with Article 8.
+
+% [CORRECT?] In the context of the Live+Gov project the ECHR is of limited relevance, since the Live+Gov System is offered by private companies and not directly by the State autority.
+
+\subsection{OECD Guidelines and CoE Convention 108}
+
+The emergence of information technologies that allow automated processing of personal data made more detailed rules for safeguarding the privacy of citizens necessary.
+In the light of this developments the OECD  \cite{OECD80} issued Guidelines on Protection of Privacy in 1980 which establish the following basic principles of data protection:
+\begin{enumerate}
+\item Collection Limitation Principle.
+  There should be limits to the collection of personal data and any such data should be obtained [...] with the knowledge or consent of the data subject.
+\item Data Quality Principle.
+  Personal data should be relevant to the purposes for which they are to be used, and,  [...] accurate, complete and kept up-to-date.
+\item Purpose Specification Principle.
+  The purposes for which personal data are collected should be specified [...] and the subsequent use limited to the fulfilment of those purposes [...].
+\item Use Limitation Principle.
+  Personal data should not be disclosed [...] except: a) with the consent of the data subject; or b) by the authority of law.
+\item Security Safeguards Principle.
+  Personal data should be protected by reasonable security safeguards [...].
+\item Openness Principle.
+  {[...]} Means should be readily available of establishing the existence and nature of personal data, and the main purposes of their use, as well as the identity and usual residence of the data controller.
+\item Individual Participation Principle.
+  An individual should have the right
+  \begin{itemize}
+  \item to obtain from a data controller, or otherwise, confirmation of whether or not the data controller has data relating to him;
+  \item to have communicated to him, data relating to him [...]
+  \item {[...]} to have the data erased, rectified, completed or amended.
+  \end{itemize}
+\item Accountability Principle.
+  A data controller should be accountable for complying with measures which give effect to the principles stated above.
+\end{enumerate}
+
+Where the following definitions are undestood.
+\begin{itemize}
+\item \emph{personal data} means any information relating to an identified or identifiable person (\emph{data subject}).
+  An identifiable person is is one who can be identified either directly, by reference to a name or identification number or indirectly,
+  by one or more factors which make it possible to find out who the data subject is by conducting further research.
+\item a \emph{data controller} means any party who is competent to decide about the contents and use of personal data.
+\end{itemize}
+
+The OCED guidelines were adopted by EU law in the Strassbourg Convention 108 of the Council of Europe \cite{CONV108} in 1981.
+
+% UPDATE 2013: http://www.oecd.org/sti/ieconomy/privacy.htm
+
 \subsection{EU Data Protection Directive}
@@
 The \emph{Directive 95/46/EC} of 1995 \cite{DIR95} is the fundamental retulation of data protection in the Europen Union.
-EU directives in general do not hold any direct legal binding for EU citizens, but has to be implemented by national law of the EU memberstates.
-
-It roughly follows the OECD guidelines for privacy protection issues in 1980 \cite{OECD80}.
-
-The directive regulates the processing of personal data regardless of whether such processing is automated or not.
-Where \emph{personal data} is any information relating to an identified or identifiable natural person [...] (Art. 2a)
-and \emph{processing} is any operation or set of operations which is performed upon personal data [...] (Art. 2b).
-
-The identified or identifiable natural person is called \emph{data subject}.
-
-The sole responsibility for compliance is held by so called \emph{controllers}.
-A controller is an actor which alone or jointly with others determines the purposes and means of the processing of personal data (Art 2d).
-
-This rule is applicable whenever the controller uses equipment situated within the EU.
-This will hold for every e-commerce provider inside or outside of the EU because the customer's computer is situated within the EU anyways.
-
-Personal data should not be processed at all, except when certain conditions are met.
-These conditions incorporate the seven OECD principals and are further categorized in the categories:
-
-\begin{itemize}
-\item \textbf{Transparency}  (\cite{1} art.
-10,11): ``The data subject has the right to be informed when his personal data is being processed [...]'' and he has the right how and by whom it is processed.
-Additionally other specified requirements have to be met, i.e.~giving consent.
-\item \textbf{Legitimate purpose} (\cite{1} art.
-6b): ``Personal data can only be processed for specified explicit and legitimate purposes and may not be processes further in a way incompatible with those purposes ''.
-\item \textbf{Proportionality}: ``Personal data may be processed only insofar as it is adequate [...] in relation to [its] purposes [...]; every reasonable step must be taken to ensure that data which are inaccurate or incomplete [...] are erased or rectified (art.
-6)''
-\end{itemize}
-
-(\textbf{Note:} The principles incorporated by the EU directive seem to closely map to \emph{the seven principles for user privacy control}.)
+It was designed to give further substance to the Convetion 108.
+In particular it makes the creation of an independent supervisory authority necessary (Art. 28).
+In Section \ref{sec:ECIMPL} the status of the implementation is discussed in some example cases.
+
+The ambition of the Directive is stated at the very beginning.
+\begin{quote}
+  (Art.
+1.1) In Accordance with this Directive the Memebr States shall protect the fundamental rights and freedoms of natural persons, and in particular their right to privacy with respect to the processing of personal data.
+\end{quote}
+
+The Directive follows the definition of personal data, data subject, and data controller of the OECD Guidelines (cf. Art. 2).
+Furthermore it introduces the role of a \emph{processor} as a person which processes, by wholly or partly automatic means (Art 3.), personal data on behalf of the controller.
+
+Data processing for operations concerning public security, defence, State security, and in areas of
+criminal law are explicitly excluded (Art. 3).
+
+The most important articles are summarized in the following paragraphs.
+\begin{itemize}
+\item Specification of Purpose (Art. 6.1b). Personal data must be collected for specified, explicit and legitimate purposes and not be processed in a way incompatible with this purposes.
+Further processing of data for historical, statistical or scientific purposes shall not be considered incompatible [...].
+
+\item Accuracy (Art. 6.1d). Personal data must be accurate and [...] every reasonable step must be taken to ensure that data which is inaccurate or incomplete [...] are erased or rectified.
+
+\item (Art. 6.1e). Personal data must be kept in a form which permits identification for no longer than is necessary [...].
+
+\item (Art 7). Personal data may be processed only if:
+  \begin{itemize}
+    \item [(a)] the data subject has unambiguously given his consent; or
+    \item [(b)] the processing is necessary for the performance of a contract to which the data subject is party [...]; or
+     \item [(..)] [further exceptions are made for legal obligations of the controller, to protect vital interests of the data subject, the public interest and]
+     \item [(f)] the processing is necessary for purposes of legitimate interests pursued by the controller or by a third party to whom the data are disclosed, except where such interests are overridden by the interests [...] of the data subject which require protection under Art.
+1(1).
+  \end{itemize}
+
+\item (Art.
+8.1).
+Member Stats shall prohibit the processing of personal data revealing racial or ethnic origin, political opinions, religious or philosophical beliefs, trade-union membership, and the processing of data concerning health or sex life.
+[Art.
+8.2 lists several exceptions including explicit consent of the data subject.]
+
+\item (Art. 10) The controller [...] must provide a data subject [...] with the following information:
+  \begin{itemize}
+    \item [(a)] the identity of the controller [...]
+    \item [(b)] the purpose of processing for which the data are intended;
+    \item [(c)] any further information [...] that is necessary to guarantee fair processing in respect of the data subject.
+[This includes in particular the recipients of the data, the existence of the right to of access and the right to rectify the data.]
+  \end{itemize}
+
+\item Right to Access (Art. 12). Every data subject has the right to obtain from the controller
+\begin{itemize}
+  \item (a) [...] confirmation as to whether or not data relating to him are being processed [...], communication to him [...] the data undergoing processing and any available information as to their source, knowledge of the logic involved in any automatic processing [...];
+  \item (b) the rectification, erasure or blocking of data the procession of which does not comply with the provisions of this Directive [...];
+  \item (c) notification to third parties to whom the data have been disclosed [...].
+\end{itemize}
+
+\item Right to Object (Art. 14).
+
+\end{itemize}
 
 Member states have to create independent \textbf{supervisory authorities} for personal data processing where controllers have to register before they start to process any data.
-The record has to be stored in public register \cite{1}.
+The record has to be stored in public register.
 
 Data transfer to third countries outside of the EU requires those countries to have a similar data protection level. Exceptions exists:
-
 \begin{itemize}
@@
 all between the US and the EU, although the US has no comparable law of data protection.
-The US Supreme Court regards privacy as \textbf{implicit} right granted with the First Amendment. However, European law states privacy as \textbf{explicit} right (ECHR Art. 8).
-Additionally the US government follows a doctrine which favors the economy to implement self-regulation \cite{1}.
-
-\subsection{Implementation of the Data Protection Directive}
+The US Supreme Court regards privacy as implicit right granted with the First Amendment.
+However, European law states privacy as explicit right (ECHR Art. 8).
+Additionally the US government follows a doctrine which favors the economy to implement self-regulation.
+
+\subsection{Implementation of the Data Protection Directive}\label{sec:ECIMPL}
 
@@
 
-Germany implements Directive 95/46 with the \emph{Federal Data Protection Act} from 2001 known as \emph{Bundesdatenschutzgesetz (BDSG)} \cite{1}\cite{2}\cite{6}.
+Germany implements Directive 95/46 with the \emph{Federal Data Protection Act} from 2001 known as \emph{Bundesdatenschutzgesetz (BDSG)}.
 However, Germany has violated the directive in two points:
-
 \begin{enumerate}
-\item
-The BDSG has become effective three years too late, thus the EC filed a treaty violation proceeding against Germany \cite{6}.
-
-\item
-The BDSG does not implement \textbf{independent} supervisory authorities. 
-The \emph{Bundesdatenschutzbeauftragter} is subordinate to the Ministry of Interior. 
-Although he is not subject to technical oversight (\emph{Fachaufsicht}), he is subject to staff supervision by the government (\emph{Rechtsaufsicht} durch die Bundesregierung und  \emph{Dienstaufsicht} durch das Innenministerium) and budget oversight (including approval of employees) by the ministry \cite{7}.
-Thus the EC filed a treaty violation proceeding against Germany, again \cite{6}. 
-In March 2010 Germany was found guilty of violation of Directive 95/46 by the ECJ \cite{6}.
+\item The BDSG has become effective three years too late, thus the EC filed a treaty violation proceeding against Germany.
+\item The BDSG does not implement independent supervisory authorities.
+The Bundesdatenschutzbeauftragter is subordinate to the Ministry of Interior.
+Although he is not subject to technical oversight (\emph{Fachaufsicht}), he is subject to staff supervision by the government (Rechtsaufsicht, Dienstaufsicht) and budget oversight by the ministry.
+In March 2010 Germany was found guilty of violation of Directive 95/46 by the ECJ.
 \end{enumerate}
 
-Germany still violates point 2!
-
 The states of Germany have their own implementation of Directive 95/46 (\emph{Landesdatenschutzgesetze}).
-Federal public authorities are only bound to their federal law. \cite{2}
-
-Churches are not subject the BDSG. \cite{2}
-
-\begin{itemize}
-
-\item
-\textbf{Roman Catholic Church:}
-\href{http://de.wikipedia.org/wiki/Anordnung_\%C3\%BCber_den_kirchlichen_Datenschutz}{Anordnung  über den kirchlichen Datenschutz (KDO)}
-
-\item
-\textbf{German Protestant Churches (Synode der Evangelischen Kirche in Deutschland):}
-\href{http://de.wikipedia.org/wiki/Datenschutzgesetz_der_Evangelischen_Kirche_in_Deutschland}{Kirchengesetz  über den Datenschutz der Evangelischen Kirche in Deutschland}
-\end{itemize}
-
-Data subjects have the following rights.
-
-\begin{itemize}
-\item
-Right to disclosure of whether data about them is stored and processed, which data is stored and processed and the data sources.
-\item
-Right to correction of false data.
-
-\item
-Right to file complaints to the supervisory authority.
-
-\item
-Right to have data deleted or blocked, but controllers can prohibit deletion in favour of blocking.
-
-\item
-Right to decline third party access to the data.
-\end{itemize}
+Federal public authorities are only bound to their federal law.
+Churches are not subject the BDSG.
 
@@
 
-The UK implements Directive 95/46 with the \emph{Data Protection Act 1998 (DPA)} \cite{1}\cite{3}.
-
-The act is known for its high complexity: a manual record of phone numbers for business purposes could be hold subject to the DPA \cite{3}.
+The UK implements Directive 95/46 with the \emph{Data Protection Act 1998} (DPA).
+
+The act is known for its high complexity: a manual record of phone numbers for business purposes could be hold subject to the DPA.
 Although the act seems to fully cover the directive.
 Even higher restriction apply for \emph{``sensitive personal data''} (race, ethnicity, politics, religion, trade union status, health, sex life or criminal record), i.e.
-\textbf{consent} must be given freely and has to be explicit.
-\cite{3}
-
-\emph{``The Act's definition of''personal data" covers any data that can be used to identify a living individual.
+consent must be given freely and has to be explicit.
+
+The Act's definition of''personal data" covers any data that can be used to identify a living individual.
 Anonymised or aggregated data is not regulated by the Act, providing the anonymisation or aggregation has not been done in a reversible way.
-Individuals can be identified by various means including their name and address, telephone number or
-Email address.
-The Act applies only to data which is held, or intended to be held, on computers (`equipment operating automatically in response to instructions given for that purpose'), or held in a `relevant filing system'.}
-
-\emph{In some cases even a paper address book can be classified as a `relevant filing system', for example diaries used to support commercial activities such as a salesperson's diary.}
-
-\emph{The Freedom of Information Act 2000 modified the act for public bodies and authorities, and the Durant case modified the interpretation of the act by providing case law and precedent.}
-
-\emph{The Data Protection Act creates rights for those who have their data stored, and responsibilities for those who store, process or transmit such data.
-The person who has their data processed has the right to:}
-
-\begin{itemize}
-
-\item
-View the data an organization holds on them.
+Individuals can be identified by various means including their name and address, telephone number or Email address.
+The Act applies only to data which is held, or intended to be held, on computers (`equipment operating automatically in response to instructions given for that purpose'), or held in a `relevant filing system'.
+
+In some cases even a paper address book can be classified as a `relevant filing system', for example diaries used to support commercial activities such as a salesperson's diary.
+
+The Freedom of Information Act 2000 modified the act for public bodies and authorities, and the Durant case modified the interpretation of the act by providing case law and precedent.
+
+The Data Protection Act creates rights for those who have their data stored, and responsibilities for those who store, process or transmit such data.
+The person who has their data processed has the right to:
+\begin{itemize}
+\item View the data an organization holds on them.
 A `subject access request' can be obtained for a nominal fee.
 As of January 2014, the maximum fee is \pounds2 for requests to credit reference agencies, \pounds50 for health and educational request, and \pounds10 per individual otherwise.
-
-\item
-Request that incorrect information be corrected.
+\item Request that incorrect information be corrected.
 If the company ignores the request, a court can order the data to be corrected or destroyed, and in some cases compensation can be awarded.
-
-\item
-Require that data is not used in any way that may potentially cause damage or distress.
-
-\item
-Require that their data is not used for direct marketing. \cite{3}
+\item Require that data is not used in any way that may potentially cause damage or distress.
+\item Require that their data is not used for direct marketing. \cite{3}
 \end{itemize}
@@
 \emph{Safe harbor} in general is the legal concept to regulate that a certain conduct will be deemed \cite{4}, but in Germany the term is commonly used as synonym for the agreement between the EU and the US regarding Directive 95/46.)
+
+\subsection{EU Charter of Fundamental Rights}
+
+The Treaty of Lisbon of 2007, which layed down the fundamental functioning of the European Union
+includes the \emph{Charter of Fundamental Rights of the Europen Union} \cite{EUFR2010}.
+This Charta summarizes the full range of civil, political and economic rights of EU citizens and contains the following two articles that safeguard the privacy of the citizen.
+\begin{quote}
+  Article 7. Respect for private and family life.\\
+  Everyone has the right to respect for his or her private and family life, home and communications.
+
+  Article 8. Protection of personal data.
+  \begin{enumerate}
+    \item [(1)] Everyone has the right to the protection of personal data concerning him or her.
+    \item [(2)] Such data must be processed fairly for specified purposes and on the basis of the consent of the person concerned or some other legitimate basis laid down by law.
+      Everyone has the right of access to data which has been collected concerning him or her, and the right to have it rectified.
+    \item [(3)] Compliance with these rules shall be subject to control by an independent authority.
+  \end{enumerate}
+\end{quote}
+
+Although the Charter does not extend the pre-exiting Directive 95/46/EC, it amplifies the importance of this privacy protection as a fundamental right.
 

### 2014-10-10 PrivacyProtection-2014-09-09.tex
Updated Directive.

@@
 \label{chap:privacy}
+
+\newcommand{\om}{[...]\xspace} % omission
+%\newcommand{\om}{\ldots} % omission
 
@@
 
-* Good overview over EU Law can be found in the Handbook on European data protection law \cite{EU_HANDBOOK_2014}
-
 \subsection{European Convention on Human Rights}
@@
 The \emph{Europen Convetion on Human Rights} \cite{ECHR} was created by the members of the \emph{Council of Europe} (CoE) in 1953 as part of the aftermath of the second world war.
-It formulates universal rights of citizens against the state authroity.
+It formulates universal rights of citizens against the state authority.
 All member states of the Council of Europe, which includes all EU members, have incorporated this convention into their national law.
@@
 
-The protection of personal data under article 8 is not an absolute law but must be considered in relation to other laws such as the freedom of expression granted in Article 10 of the ECHR.
+The protection of personal data under Article 8 is not an absolute law but must be considered in relation to other laws such as the freedom of expression (ECHR, Art. 10).
 The \emph{European Court of Human Rights} (ECtHR) overseas the implementation of this directive and has adopted a rather broad interpretation of the article.\footnote{\url{http://en.wikipedia.org/wiki/Article_8_of_the_European_Convention_on_Human_Rights}}
 
-In particular State actions of searching a persons home, gathereing and storing information in a secret police file, and stopping a prisonser's communication have been ruled to interfere with Article 8.
-
-% [CORRECT?] In the context of the Live+Gov project the ECHR is of limited relevance, since the Live+Gov System is offered by private companies and not directly by the State autority.
+In particular State actions of searching a persons home, gathering and storing information in a secret police file, and stopping a prisoner's communication have been ruled to interfere with Article 8.
+
+In the context of the Live+Gov project the ECHR is of limited relevance, since the Live+Gov System is offered by private companies and not directly by the State authority.
 
@@
 \item Collection Limitation Principle.
-  There should be limits to the collection of personal data and any such data should be obtained [...] with the knowledge or consent of the data subject.
+  There should be limits to the collection of personal data and any such data should be obtained \om with the knowledge or consent of the data subject.
 \item Data Quality Principle.
-  Personal data should be relevant to the purposes for which they are to be used, and,  [...] accurate, complete and kept up-to-date.
+  Personal data should be relevant to the purposes for which they are to be used, and,  \om accurate, complete and kept up-to-date.
 \item Purpose Specification Principle.
-  The purposes for which personal data are collected should be specified [...] and the subsequent use limited to the fulfilment of those purposes [...].
+  The purposes for which personal data are collected should be specified \om and the subsequent use limited to the fulfillment of those purposes \om.
 \item Use Limitation Principle.
-  Personal data should not be disclosed [...] except: a) with the consent of the data subject; or b) by the authority of law.
+  Personal data should not be disclosed \om except: a) with the consent of the data subject; or b) by the authority of law.
 \item Security Safeguards Principle.
-  Personal data should be protected by reasonable security safeguards [...].
+  Personal data should be protected by reasonable security safeguards \om.
 \item Openness Principle.
-  {[...]} Means should be readily available of establishing the existence and nature of personal data, and the main purposes of their use, as well as the identity and usual residence of the data controller.
+  {\om} Means should be readily available of establishing the existence and nature of personal data, and the main purposes of their use, as well as the identity and usual residence of the data controller.
 \item Individual Participation Principle.
@@
   \item to obtain from a data controller, or otherwise, confirmation of whether or not the data controller has data relating to him;
-  \item to have communicated to him, data relating to him [...]
-  \item {[...]} to have the data erased, rectified, completed or amended.
+  \item to have communicated to him, data relating to him \om
+  \item \om to have the data erased, rectified, completed or amended.
   \end{itemize}
@@
 
-Where the following definitions are undestood.
+Where the following definitions are understood.
 \begin{itemize}
@@
 
-The OCED guidelines were adopted by EU law in the Strassbourg Convention 108 of the Council of Europe \cite{CONV108} in 1981.
+The OECD guidelines were adopted by EU law in the Strassbourg Convention 108 of the Council of Europe \cite{CONV108} in 1981.
+
+Relevant for our investigations is in particular, that sensor data collected from mobile devices is considered personal data if the individual can be identified based on the data by any direct or indirect means.
+Therefore the above principles apply.
 
@@
 
-The \emph{Directive 95/46/EC} of 1995 \cite{DIR95} is the fundamental retulation of data protection in the Europen Union.
-It was designed to give further substance to the Convetion 108.
+The \emph{Directive 95/46/EC} of 1995 \cite{DIR95} is the fundamental regulation of data protection in the European Union.
+It was designed to give further substance to the Convention 108.
 In particular it makes the creation of an independent supervisory authority necessary (Art. 28).
+As this text is the main legal basis for our investigation we review it in detail.
 In Section \ref{sec:ECIMPL} the status of the implementation is discussed in some example cases.
+
+The interpretation and oversight of the Directive lies at the Court of Justice of the European Union (CJEU).
+A summary of the most important rulings can be found in Handbook on European data protection law \cite{EU_HANDBOOK_2014}.
 
@@
   (Art.
-1.1) In Accordance with this Directive the Memebr States shall protect the fundamental rights and freedoms of natural persons, and in particular their right to privacy with respect to the processing of personal data.
+1.1) In Accordance with this Directive the Member States shall protect the fundamental rights and freedoms of natural persons, and in particular their right to privacy with respect to the processing of personal data.
 \end{quote}
@@
 
-Data processing for operations concerning public security, defence, State security, and in areas of
-criminal law are explicitly excluded (Art. 3).
+Data collection of intelligence agencies and the police is not restricted by the Directive (Art. 3).
 
 The most important articles are summarized in the following paragraphs.
-\begin{itemize}
-\item Specification of Purpose (Art. 6.1b). Personal data must be collected for specified, explicit and legitimate purposes and not be processed in a way incompatible with this purposes.
-Further processing of data for historical, statistical or scientific purposes shall not be considered incompatible [...].
-
-\item Accuracy (Art. 6.1d). Personal data must be accurate and [...] every reasonable step must be taken to ensure that data which is inaccurate or incomplete [...] are erased or rectified.
-
-\item (Art. 6.1e). Personal data must be kept in a form which permits identification for no longer than is necessary [...].
-
-\item (Art 7). Personal data may be processed only if:
+In order to keep the text more readable we have simplified the formulation by removing the formulation ``Member States shall provide that ...'' from each paragraph.
+\begin{itemize}
+\item (Art. 6.1.a) Personal data must be processed fairly and lawfully.
+
+[The meaning of the term 'fair processing' is entailed in the following articles.
+In particular transparency of processing, information of the data subject, and right to access the data are implied (cf. \cite{EU_HANDBOOK_2014})]
+
+\item (Art. 6.1.b) \emph{Specification of Purpose.} Personal data must be collected for specified, explicit and legitimate purposes and not be processed in a way incompatible with this purposes.
+
+Further processing of data for historical, statistical or scientific purposes shall not be considered incompatible \om.
+
+\item (Art. 6.1.c) Personal data must be adequate, relevant and non-excessive in relation to the purposes \om.
+
+\item (Art. 6.1.d) \emph{Accuracy.} Personal data must be accurate and \om every reasonable step must be taken to ensure that data which is inaccurate or incomplete \om are erased or rectified.
+
+\item (Art. 6.1e) Personal data must be kept in a form which permits identification for no longer than is necessary \om.
+% \item (Art. 6.2) It shall be for the controller to ensure that (Art. 6.1) is complied with.
+\end{itemize}
+
+In addition to strengthening the OECD Data Quality Principle in Article 6, the Directive requires a strong form of consent of the data subject, before any collection or processing of data can take place.
+
+\begin{itemize}
+\item (Art. 7). Personal data may be processed only if:
   \begin{itemize}
     \item [(a)] the data subject has unambiguously given his consent; or
-    \item [(b)] the processing is necessary for the performance of a contract to which the data subject is party [...]; or
-     \item [(..)] [further exceptions are made for legal obligations of the controller, to protect vital interests of the data subject, the public interest and]
-     \item [(f)] the processing is necessary for purposes of legitimate interests pursued by the controller or by a third party to whom the data are disclosed, except where such interests are overridden by the interests [...] of the data subject which require protection under Art.
+    \item [(..)] [exceptions are made for performance of a contracts with the data subject,  legal obligations of the controller, to protect vital interests of the data subject, the public interest and]
+     \item [(f)] the processing is necessary for purposes of legitimate interests pursued by the controller or by a third party to whom the data are disclosed, except where such interests are overridden by the interests \om of the data subject which require protection under Art.
 1(1).
@@
 \item (Art.
-8.1).
-Member Stats shall prohibit the processing of personal data revealing racial or ethnic origin, political opinions, religious or philosophical beliefs, trade-union membership, and the processing of data concerning health or sex life.
+8.1) \emph{Special Categories.} Member Stats shall prohibit the processing of personal data revealing racial or ethnic origin, political opinions, religious or philosophical beliefs, trade-union membership, and the processing of data concerning health or sex life.
 [Art.
-8.2 lists several exceptions including explicit consent of the data subject.]
-
-\item (Art. 10) The controller [...] must provide a data subject [...] with the following information:
-  \begin{itemize}
-    \item [(a)] the identity of the controller [...]
+8.2 lists several exceptions, including explicitly explicit consent of the data subject.]
+\end{itemize}
+
+The OECD Participation Principle is addressed and extended in the following articles.
+
+\begin{itemize}
+\item (Art. 10, 11) The controller \om must provide a data subject \om with the following information:
+  \begin{itemize}
+    \item [(a)] the identity of the controller \om
     \item [(b)] the purpose of processing for which the data are intended;
-    \item [(c)] any further information [...] that is necessary to guarantee fair processing in respect of the data subject.
+    \item [(c)] any further information \om that is necessary to guarantee fair processing in respect of the data subject.
 [This includes in particular the recipients of the data, the existence of the right to of access and the right to rectify the data.]
@@
 
-\item Right to Access (Art. 12). Every data subject has the right to obtain from the controller
-\begin{itemize}
-  \item (a) [...] confirmation as to whether or not data relating to him are being processed [...], communication to him [...] the data undergoing processing and any available information as to their source, knowledge of the logic involved in any automatic processing [...];
-  \item (b) the rectification, erasure or blocking of data the procession of which does not comply with the provisions of this Directive [...];
-  \item (c) notification to third parties to whom the data have been disclosed [...].
-\end{itemize}
-
-\item Right to Object (Art. 14).
-
-\end{itemize}
-
-Member states have to create independent \textbf{supervisory authorities} for personal data processing where controllers have to register before they start to process any data.
-The record has to be stored in public register.
-
-Data transfer to third countries outside of the EU requires those countries to have a similar data protection level. Exceptions exists:
-\begin{itemize}
-\item
-  Save Harbor
-\item
-  Passenger Name Record Agreement
-\end{itemize}
-
-all between the US and the EU, although the US has no comparable law of data protection.
-The US Supreme Court regards privacy as implicit right granted with the First Amendment.
-However, European law states privacy as explicit right (ECHR Art. 8).
-Additionally the US government follows a doctrine which favors the economy to implement self-regulation.
+\item (Art. 12) \emph{Right to Access.} Every data subject has the right to obtain from the controller
+\begin{itemize}
+  \item (a) \om confirmation as to whether or not data relating to him are being processed \om, communication to him \om the data undergoing processing and any available information as to their source, knowledge of the logic involved in any automatic processing \om;
+  \item (b) the rectification, erasure or blocking of data the procession of which does not comply with the provisions of this Directive \om;
+  \item (c) notification to third parties to whom the data have been disclosed \om.
+\end{itemize}
+
+%\item Confidentiality of Processing (Art. 16).
+%Any person under the authority of the controller or of the processor \om who has access to personal data must not process them except on instructions from the controller, unless he is required to do so by law.
+\end{itemize}
+
+A similar form of the Security Safeguard Principle is contained in Article 17.
+
+\begin{itemize}
+\item (Art. 17) \emph{Security of processing.}
+The controller must implement appropriate technical and organizational measures to protect personal data against accidental loss, alteration, unauthorized disclosure or access \om and against all other forms of unlawful processing.
+
+[The level of security has to be balanced against the risk of processing.]
+
+% Having regard to the state of the art and the cost of their implementation, such measures shall ensure a level of security appropriate to the risk represented by the processing and the nature of the data to be processed.
+\end{itemize}
+
+Furthermore the Directive demands all data processing to be reported to a supervisory authority.
+This obligation can be lifted if an internal data protection official is appointed, who is responsible in particular for keeping a processing register, that has to be made available on request.
+
+\begin{itemize}
+\item Notification (Art. 18. 1).
+The controller \om must notify the supervisory authority referred to in Article 28 before carrying out any wholly or partly automatic processing operation \om.
+
+\item (Art. 18.2).
+Simplification or exemption from notification may be provided only under the following conditions: \om where the controller \om appoints a personal data protection official, responsible in particular:
+\begin{itemize}
+\item for ensuring in an independent manner the internal application of the national provisions taken pursuant to this directive.
+\item for keeping the register of processing operations carried out by the controller, containing the items of information referred to in Article 21 (2)
+\end{itemize}
+thereby ensuring that the rights and freedoms of the data subjects are unlikely to be adversely affected by the processing operations.
+
+\item (Art. 19.1)
+The information to be given in the notification shall include at least:
+\begin{itemize}
+  \item [(a)] the name and address of the controller \om;
+  \item [(b)] the purpose \om of processing;
+  \item [(c)] a description of the category \om of the data \om;
+  \item [(d)] the recipients \om to whom the data might be disclosed;
+  \item [(e)] proposed transfer of data to third countries;
+  \item [(f)] a general description allowing a preliminary assessment to be made of the appropriateness of the measures taken pursuant of Article 17 to ensure the security of processing;
+\end{itemize}
+
+\item (Art. 21.2) \emph{Publication}.
+A register of processing information \om shall be kept at the supervisory authority.
+The register shall contain at least the information listed in Article 19.1 (a) to (e).
+
+The register may be inspected by any person.
+
+\item (Art.
+21.3).
+In relation to processing operations not subject to notification, that controllers \om make available at least the information referred to in Article 19.1 (a) to (e) in an appropriate form to any person on request.
+
+\item (Art. 28) \emph{Supervisory Authority.}
+Each Member State shall provide that one or more public authorities are responsible for monitoring the application \om of this Directive.
+
+These authorities shall act in complete independence in exercising this functions entrusted to them.
+\end{itemize}
+
+Data transfer to third countries outside of the EU requires those countries to have an adequate level of data protection.
+\begin{itemize}
+\item (Art. 25.1) \emph{Transfer to third countries.}
+The transfer to a third country of personal data \om may take place only if \om the third country in question ensures an adequate level of protection.
+\end{itemize}
+The European Commission can decide whether a third country ensures an adequate level of protection (Art 25.6).
+The USA, is not considered to do so.
+Data exchange between the US and EU is possible under the Safe Harbor regulation \cite{SAFE_HABOR}.
+Further Exceptions are provided by the Passenger Name Record Agreement \cite{PNR}.
 
@@
 
-\subsubsection*{Germany}
-
-Germany implements Directive 95/46 with the \emph{Federal Data Protection Act} from 2001 known as \emph{Bundesdatenschutzgesetz (BDSG)}.
-However, Germany has violated the directive in two points:
+\textbf{Germany}.
+Germany implements Directive 95/46 with the \emph{Bundesdatenschutzgesetz (BDSG)}
+of 2001. However, Germany has violated the directive in two points:
 \begin{enumerate}
@@
 
-\subsubsection*{United Kingdom}
-
+\textbf{United Kingdom.}
 The UK implements Directive 95/46 with the \emph{Data Protection Act 1998} (DPA).
@@
 
-This type is also concerned with the ``protection against disclosure of personal matters'' through behaviour, however Clarke's distinction between ``casual observation [...] systematic recording and storage of information about those activities'' is lifted.
+This type is also concerned with the ``protection against disclosure of personal matters'' through behaviour, however Clarke's distinction between ``casual observation \om systematic recording and storage of information about those activities'' is lifted.
 
@@
 best understand as \emph{Biometric Privacy}. Friedewald et al. paraphrase
-it as \emph{``[...] the right to keep body functions and body
-characteristics [...] private''}. \textbf{Accelerometer},
+it as \emph{``\om the right to keep body functions and body
+characteristics \om private''}. \textbf{Accelerometer},
 \textbf{Rotation Vector} and \textbf{Gyroscope} measure the physical

### 2014-10-13 PrivacyProtection-2014-09-09.tex
Deleted "Excessive data mining" (Commented Out)

@@
 
-The Act's definition of''personal data" covers any data that can be used to identify a living individual.
+The Act's definition of personal data covers any data that can be used to identify a living individual.
 Anonymised or aggregated data is not regulated by the Act, providing the anonymisation or aggregation has not been done in a reversible way.
-Individuals can be identified by various means including their name and address, telephone number or Email address.
-The Act applies only to data which is held, or intended to be held, on computers (`equipment operating automatically in response to instructions given for that purpose'), or held in a `relevant filing system'.
-
-In some cases even a paper address book can be classified as a `relevant filing system', for example diaries used to support commercial activities such as a salesperson's diary.
 
@@
 
-%\paragraph{Excessive Data Mining}
-%
-%\subparagraph{1 Privacy of the Person private''} \cite{1}. Here we are
-%concerned wi the personal physiological and psychological p and  This privacy is threat...f \textbf{body characteristics}weight, height,, dna,\ldots) or
-%depression emph therapists or other doctors) \textbf{can beevealed orii and emph+Gov project collects sensor data from mobile devices
-%(accellerometer, location, \ldots). By a...g \emph{human activityecognition (emph)} techniques we could detect certain movement patterns
-%we know the position with of indiviual citizens with a sufficientccurancy, and the Live+Gov project is only applied to a destinct urban
-%area. So we could link HAR data with the yellow pages, filter for
-%medical specialists and limit the possibilities of pathologic
-%conditions. But even without HAR data we could determine a likelihood
-%for certain conplaints. Frequent vistits to the dentist does not imply
-%healthy teeth.
-
-%\subparagraph{2 Privacy of Behaviour and Action}
-%
-%\emph{``This type is also concerned with the `protection against
-%disclosure of personal matters' through behaviour''}
-%\cite{1}. This type is concered with religous
-%practices, sexaul habits, political activities, etc. revealed through
-%observation. For the Live+Gov project this type is related to t is recirded by default and by linking it to maps a and li phone books those *``personal matters*" could be
-%disclosed by frequently visited places (c...s, brothels, party headuaters, \emph).
-%
-%communications' either electronic or face-to-face'' \cite{1}. This type
-%is threatened if we intrude the secrecy of correspondence, posts and
-%telecommunications, personal direct communication or right to free
-%discussion without third parties listening. The Live+Giv project only
-%could threaten this type by using mobile client application as trojan
-%horse, collecting any communication data (chat, sms, microphone).
-%
-%\subparagraph{4 Privacy of Data and Image}
-%
-%\emph{``This type is concerned with `making sure that individuals's data
-%is not automatically available to other individuals and
-%organisations'\,''} \cite{1}. This is
-%Informational Privacy in an intuitive sense regarding data like:
-%
-%\begin{itemize}
-%
-%\item
-%  Phone Number
-%\item
-%  IP Address
-%\item
-%  Public-administrative Data (Date of Birth, population register)
-%\item
-%  Data held by organizations, like Banks or Insurance Companies
-%\item
-%  All data that is stored in online services (Facebook)
-%\end{itemize}
-%
-%The Live+Gov project does not threaten this type directly if we assume a
-%secure and closed system. However, this privacy type could be threatened
-%by carelessly ignoring known (technical) vulnerabilities regarding data
-%security. Additionally we could threaten the Privacy of Data and Image
-%indirectly by linking collected data with other sources.
-%
-%\subparagraph{5 Privacy of Thoughts and Feelings}
-%
-%\emph{``This type is the right `not to share their thoughts or feelings
-%or to have those thoughts or feeling (sic!) revealed'\,''}
-%\cite{1}. This means thoughts and emotions must
-%not be detected autmatically. Intuitively the Live+Gov system seems
-%unable to threaten this type. However, considering the issue component
-%of the Urban Maintaince use case, this might reveal information about
-%one's thoughts regarding the community in a positive manner. Solely by
-%taking part we could assume a caring personality. This threatens one's
-%privacy in a rather technical sense, but does not necessarily impose any
-%harm.
-%
-%An other way of threatening this type could be constructed by recording
-%one's voice with the phone's microphone and run emotion detecting
-%algorithms against this data.
-%
-%\subparagraph{6 Privacy of Location and Space}
-%
-%\emph{``This type is the right `to move about in public or semi-public
-%space withoug being identified, tracked or monitored'. Additionally this
-%type is concerned with the protection of one's home and private places
-%(`right to solitude')''} \cite{1}. The location
-%dimension of this type is relatively easy to understand: The
-%geo-position of citizens cannot be monitored by default. However, by
-%actively taking part in the Live+Gov project, the locational privacy of
-%citizens is threatened by default because data of the location sensor
-%will be collected. The space dimension is more complex, but can be
-%simplified with a \emph{``right to solitude''}. This dimension could be
-%threatend with the invasion of personal space in any means, i.e.~by
-%disrespecting one's right to an inviolate home or by undercutting one's
-%comfort zone in an conversation. Live+Gov only utilizes the mobile
-%devices of citizens, so we could disrespect the former by activating the
-%phone's micropone and start recording.
-%
-%\subparagraph{7 Privacy of Association}
-%
-%\emph{``This type is the right `to associate with whomever [one]
-%wish, withoug being monitored'.''} \cite{1}. This
-%means that one's associations must not be recoreded by default
-%independet from any suspicion. Anyhow, it does not mean that this right
-%cannot be forfeit given a reasonable suspicion. The Live+Gov project
-%could easily threaten this type just by linking locational data with
-%yellow pages and a map. Even if an association graph cannot be deducted
-%for individual citizens, we could aggregate the data to create an
-%association graph for a whole population as the Live+Gov system is
-%applied to a restricted urban area.
-
 \subsection{Step 3. Plan Development}

### 2014-10-13 PrivacyProtection-2014-09-09.tex
Updated 7 Types of Privacy.

@@
 As background information we include an overview about historical treatments of privacy as well as legal regulation of privacy in the European Union.
-Based on this we propose a definition of privacy as {\em control over personal data}, and introduce a taxonomy of privacy attributes that give specify the term {\em personal data} in the context of mobile sensor data collection.
+Based on this we propose a definition of privacy as {\em control over personal data}, and introduce seven privacy types that give specify the term {\em personal data} in the context of mobile sensor data collection.
 
@@
 
-* Present legal view on privacy from European angle: ``Data Protection Directive`` and Implementation.
-
-* Some US Law
+In the following sections we outline the most relevant legislature regarding privacy protection,
+from on European angle.
+Of particular importance is the EU Directive 95/46/EC that is discussed in section \ref{EUDIR}.
+
+Based on the legal background presented here, we will derive the legal privacy requirements for our systems in \ref{sec:LEGAL_REQ}.
 
@@
 \item \emph{personal data} means any information relating to an identified or identifiable person (\emph{data subject}).
-  An identifiable person is is one who can be identified either directly, by reference to a name or identification number or indirectly,
-  by one or more factors which make it possible to find out who the data subject is by conducting further research.
+  An identifiable person is one who can be identified either directly, by reference to a name or identification number or indirectly, by one or more factors which make it possible to find out who the data subject is by conducting further research.
 \item a \emph{data controller} means any party who is competent to decide about the contents and use of personal data.
@@
 
-\subsection{EU Data Protection Directive}
+\subsection{EU Data Protection Directive}\label{EUDIR}
 
@@
 
-\textbf{United Kingdom.}
+\textbf{United Kingdom.} %TODO: HOW MUCH IS COPY PASTE FROM WIKIPEDIA?
 The UK implements Directive 95/46 with the \emph{Data Protection Act 1998} (DPA).
@@
 
-\subsubsection*{France}
-
-France implements Directive 95/46 with Law 2004-801 modifying law 78-17 of 6.1.1978 (\emph{Loi n2004-801 du 6 ao\^{u}t 2004} modifying \emph{loi n78-17 relative à l'informatique, aux fichiers et aux libert\'{e}s}) \cite{1}\cite{10}\cite{11}.
-
-\subsubsection*{United States of America}
-
+\textbf{United States of America}
 The USA do not implement the directive, nor is there any obligation for them to do so.
-However, companies subject to US jurisdiction can be certified to comply with the seven principles enforced by Directive 95/46 (the seven OECD recommendations) \cite{5}.
-Thus, those companies will act as \emph{safe harbors}. Without certification foreign companies are not allowed to send customer data back home \cite{9}.
-
-\emph{Safe harbor} in general is the legal concept to regulate that a certain conduct will be deemed \cite{4}, but in Germany the term is commonly used as synonym for the agreement between the EU and the US regarding Directive 95/46.)
+However, companies subject to US jurisdiction can be certified to comply with the seven principles enforced by Directive 95/46.
+Thus, those companies will act as \emph{safe harbors}.
+Without certification foreign companies are not allowed to store and process customer data in their country.
 
 \subsection{EU Charter of Fundamental Rights}
-
-The Treaty of Lisbon of 2007, which layed down the fundamental functioning of the European Union
+The Treaty of Lisbon of 2007, which introduced the fundamental functioning of the European Union
 includes the \emph{Charter of Fundamental Rights of the Europen Union} \cite{EUFR2010}.
-This Charta summarizes the full range of civil, political and economic rights of EU citizens and contains the following two articles that safeguard the privacy of the citizen.
+This Charter summarizes the full range of civil, political and economic rights of EU citizens and contains the following two articles that safeguard the privacy of the citizen.
+
 \begin{quote}
@@
 
-Defining privacy is a challenge which seems impossible.
-This is well put to words by Serge Gutwirth, who notes: \emph{``The notion of privacy remains out of the grasp of every academic chasing it.
-Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.''} \cite{2}.
+% TODO: Expand this section!!
+
+Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes in \cite{Gutwirth}:
+
+\begin{quote}
+The notion of privacy remains out of the grasp of every academic chasing it. Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.
+\end{quote}
 
@@
 
-This type is the right keep body functions and body characteristics private. It maps the one by Clarke.
+This type is the right keep body functions and body characteristics private.
 
@@
 
-This type is concerned with ``making sure that individual's data is not automatically available to other individuals and organizations''. This is Informational Privacy in an intuitive sense.
+This type is concerned with ``making sure that individual's data is not automatically available to other individuals and organizations''.
+This is Informational Privacy in an intuitive sense.
 
@@
 
-\subsection{Implicit Privacy Type Violation}
-
-Cf. Figure \ref{figure:Implicit Privacy Violation Matrix}
-
-%\includegraphics{../diagrams/png/implicit-privacy-violation-matrix.png}
-\input{./figures/Implicit-Privacy-Violation-Matrix}
-
-\paragraph*{Privacy of The Person}
+\subsection{Privacy Type Inferences}
+
+The above types of privacy types are not disjoint to each other.
+For example if the location of a citizen is known it is possible to infer information about political associations (e.g. his visits to a party meeting).
+In this section we explore possibilities how personal data can be inferred from each other.
+Our findings are summarized in Figure \ref{figure:Implicit Privacy Violation Matrix}.
+
+\paragraph*{1. Privacy of The Person}
 
@@
 \begin{itemize}
-
-\item
-  \textbf{Privacy of The Person:} reflexive violation
-\item
-  \textbf{Privacy of Behaviour and Action:} none
-\item
-  \textbf{Privacy of Communication:} none
-\item
-  \textbf{Privacy of Data and Image:} none
-\item
-  \textbf{Privacy of Thoughts and Feelings:} Some psychological diseases
+\item [(1-5)]
+  Privacy of Thoughts and Feelings. Some psychological diseases
   (e.g.~depression) have physiological impact. Such physiological
   patterns could be detected.
-\item
-  \textbf{Privacy of Location and Space:} none
-\item
-  \textbf{Privacy of Association:} none
-\end{itemize}
-
-\paragraph*{Privacy of Behaviour and Action}
-The Privacy of Behaviour and Action is concerned with one' and social activities (religious, po...l, sexual, \ldots). Ifhis type is violated, following implicit violations are possible:
-\begin{itemize}
-\item
-  \textbf{Privacy of The Person:} Religious practices which include body
-  modifications (e.g.~circumcision).
-\item
-  \textbf{Privacy of Behaviour and Action:} reflexive violation
-\item
-  \textbf{Privacy of Communication:} none
-\item
-  \textbf{Privacy of Data and Image:} none
-\item
-  \textbf{Privacy of Thoughts and Feelings:} Social activities in
-  general depend on a certain intellectual attitude. Such an activity is
-  the expressions of such an attitude.
-\item
-  \textbf{Privacy of Location and Space:} none
-\item
-  \textbf{Privacy of Association:} Recording religious, political or
-  sexual activities can reveal association with churches, political
-  parties or sexual partners.
-\end{itemize}
-
-\paragraph*{Privacy of Communication}
-
-The Privacy of Communication is concerned with not havin such
-communication (correspondence or vis-a-vis) intercepted. This is very
-broad type of privacy. Depending on the contents of the intercepted
-communication every other type can be violated:
-
-\begin{itemize}
-
-\item
-  \textbf{Privacy of The Person:} Communication about body
-  characteristics.
-\item
-  \textbf{Privacy of Behaviour and Action:} Communication about social
-  activities.
-\item
-  \textbf{Privacy of Communication:} reflexive violation
-\item
-  \textbf{Privacy of Data and Image:} Communication containing one's
-  passwords or other sensitive data.
-\item
-  \textbf{Privacy of Thoughts and Feelings:} Communication of thoughts
-  and feelings, e.g.~wiretapping a flirt or a catholic confession
+\end{itemize}
+
+\paragraph*{2. Privacy of Behaviour and Action}
+
+The Privacy of Behaviour and Action is concerned with one's social, political, religious, sexual, etc. activities.
+If his type is violated, following implicit violations are possible:
+
+\begin{itemize}
+\item [(2-1)] Privacy of The Person.
+  Religious practices which include body modifications (e.g.~circumcision).
+\item [(2-5)] Privacy of Thoughts and Feelings.
+  Social activities in general depend on a certain intellectual attitude.
+  Such an activity is the expressions of such an attitude.
+\item [(2-7)] Privacy of Association.
+  Recording religious, political or sexual activities can reveal association with churches, political parties or sexual partners.
+\end{itemize}
+
+\paragraph*{3. Privacy of Communication}
+
+The Privacy of Communication is concerned with not having such communication (correspondence or vis-a-vis) intercepted.
+This is very broad type of privacy.
+Depending on the contents of the intercepted communication every other type can be violated:
+
+\begin{itemize}
+\item [(3-1)] Privacy of The Person. Communication about body characteristics.
+\item [(3-2)] Privacy of Behaviour and Action. Communication about social activities.
+\item [(3-4)] Privacy of Data and Image.
+  Communication containing one's passwords or other sensitive data.
+\item [(3-5)] Privacy of Thoughts and Feelings.
+  Communication of thoughts and feelings, e.g.~wiretapping a flirt or a catholic confession
   ritual.
-\item
-  \textbf{Privacy of Location and Space:} Interception of face-to-face
-  communication is only possible if one's location and space is violated
-  (wiretapping).
-\item
-  \textbf{Privacy of Association:} Communication about one's
-  associations (family members, churches, etc.).
-\end{itemize}
-
-\paragraph*{Privacy of Data and Image}
-
-The Privacy of Data and Image is concerned with one's data not being
-automatically available to others. This also is a very broad type of
-privacy. Depending of the data or image contents every other type can be
-violated:
-
-\begin{itemize}
-
-\item
-  \textbf{Privacy of The Person:} Images or stored biometric information
-  reveal one's physical characteristics.
-\item
-  \textbf{Privacy of Behaviour and Action:} Images or diaries can reveal
-  one's social activities.
-\item
-  \textbf{Privacy of Communication:} Modern communication systems
-  usually contain some sort of archive function, e.g.~E-mail clients do
-  not automatically delete messages. Such messages are data and reveal
-  one's communication.
-\item
-  \textbf{Privacy of Data and Image:} reflexive violation
-\item
-  \textbf{Privacy of Thoughts and Feelings:} Images can show one's
-  emotional state.
-\item
-  \textbf{Privacy of Location and Space:} Images can reveal one's
-  location, e.g.~making a picture in front of the Eifel Tower.
-\item
-  \textbf{Privacy of Association:} E-mail data can also reveal
+\item [(3-6)] Privacy of Location and Space.
+  Interception of face-to-face communication is only possible if one's location and space is violated (wiretapping).
+\item [(3-7)] Privacy of Association.
+  Communication about one's associations (family members, churches, etc.).
+\end{itemize}
+
+\paragraph*{4. Privacy of Data and Image}
+
+The Privacy of Data and Image is concerned with one's data not being automatically available to others.
+This also is a very broad type of privacy.
+Depending of the data or image contents every other type can be violated:
+
+\begin{itemize}
+
+\item [(4-1)] Privacy of The Person.
+  Images or stored biometric information reveal one's physical characteristics.
+\item [(4-2)]Privacy of Behaviour and Action.
+  Images or diaries can reveal one's social activities.
+\item [(4-3)]Privacy of Communication.
+  Modern communication systems usually contain some sort of archive function, e.g.~E-mail clients do not automatically delete messages.
+  Such messages are data and reveal one's communication.
+\item [(4-4)]Privacy of Thoughts and Feelings.
+  Images can show one's emotional state.
+\item [(4-5)]Privacy of Location and Space.
+  Images can reveal one's location, e.g.~making a picture in front of the Eifel Tower.
+\item [(4-6)]Privacy of Association. E-mail data can also reveal
   association.
@@
 
-\paragraph*{Privacy of Thoughts and Feelings}
+\paragraph*{5. Privacy of Thoughts and Feelings}
 
@@
 \begin{itemize}
-
-\item
-  \textbf{Privacy of The Person:} Thoughts and feelings can reveal
-  medical conditions.
-\item
-  \textbf{Privacy of Behaviour and Action:} Thoughts and feelings can
-  reveal a certain attitudes which create a foundation for certain
-  social activities.
-\item
-  \textbf{Privacy of Communication:} none
-\item
-  \textbf{Privacy of Data and Image:} none
-\item
-  \textbf{Privacy of Thoughts and Feelings:} reflexive violation
-\item
-  \textbf{Privacy of Location and Space:} none
-\item
-  \textbf{Privacy of Association:} Thoughts and feelings can reveal
-  individual association, e.g amorous feelings for a certain person.
-\end{itemize}
-
-\paragraph*{Privacy of Location and Space}
-
-The Privacy of Location and Space is concerned with one's right to move
-freely without being tracked and one's right to private places. If this
-type is violated, following implicit violations are possible:
-
-\begin{itemize}
-
-\item
-  \textbf{Privacy of The Person:} Frequently visited doctors can reveal
-  certain medical conditions, if such doctors are known specialsts. In
-  general it could imply ill-being.
-\item
-  \textbf{Privacy of Behaviour and Action:} Frequently visited places in
-  general can reveal association and hence implies social activities.
-\item
-  \textbf{Privacy of Communication:} If one's location is knwon, it is
-  possible to intercept (wiretap) one's communication. This also may
-  violate the right to private spaces.
-\item
-  \textbf{Privacy of Data and Image:} If one's location is known, it is
-  possible shoot pictures. This violates the right to one's image
-  (\emph{``Recht am eigenen Bild''}).
-\item
-  \textbf{Privacy of Thoughts and Feelings:} Frequently visited persons
-  may imply certain thoughts and feelings, e.g.~having a mistress.
-\item
-  \textbf{Privacy of Location and Space:} reflexive violation
-\item
-  \textbf{Privacy of Association:} Frequently visited places can reveal
-  associations simply by searching in maps or yellow-pages.
-\end{itemize}
-
-\paragraph*{Privacy of Association}
-
-The Privacy of Association is concerned with one's right to associate
-with whomever one wants, without that association having recorded.If
-this type is violated, following implicit violations are possible:
-
-\begin{itemize}
-
-\item
-  \textbf{Privacy of The Person:} Association with tattoo artists could
-  imply having tattoos or other body modifications
-\item
-  \textbf{Privacy of Behaviour and Action:} Association with churches or
-  political organizations could imply certain activities.
-\item
-  \textbf{Privacy of Communication:} none
-\item
-  \textbf{Privacy of Data and Image:} none
-\item
-  \textbf{Privacy of Thoughts and Feelings:} Association with churches of
-  political organizations could imply a certain intellectual attitude.
-\item
-  \textbf{Privacy of Location and Space:} none
-\item
-  \textbf{Privacy of Association:} reflexive violation
-\end{itemize}
+\item [(5-1)] Privacy of The Person.
+  Thoughts and feelings can reveal medical conditions.
+\item [(5-2)] Privacy of Behaviour and Action.
+  Thoughts and feelings can reveal a certain attitudes which create a foundation for certain   social activities.
+\item [(5-7)] Privacy of Association.
+  Thoughts and feelings can reveal individual association, e.g amorous feelings for a certain person.
+\end{itemize}
+
+\paragraph*{6. Privacy of Location and Space}
+
+The Privacy of Location and Space is concerned with one's right to move freely without being tracked and one's right to private places.
+If this type is violated, following implicit violations are possible:
+
+\begin{itemize}
+\item [(6-1)] Privacy of The Person.
+  Frequently visited doctors can reveal certain medical conditions, if such doctors are known specialsts.
+  In general it could imply illness.
+\item [(6-2)] Privacy of Behaviour and Action.
+  Frequently visited places in general can reveal association and hence implies social activities.
+\item [(6-3)] Privacy of Communication.
+  If one's location is knwon, it is possible to intercept (wiretap) one's communication.
+  This also may violate the right to private spaces.
+\item [(6-4)] Privacy of Data and Image.
+  If one's location is known, it is possible shoot pictures.
+\item [(6-5)] Privacy of Thoughts and Feelings.
+  Frequently visited persons may imply certain thoughts and feelings, e.g.~having a mistress.
+\item [(6-6)] Privacy of Association.
+  Frequently visited places can reveal associations simply by searching in maps or yellow-pages.
+\end{itemize}
+
+\paragraph*{7. Privacy of Association}
+
+The Privacy of Association is concerned with one's right to associate with whomever one wants, without that association having recorded.
+If this type is violated, following implicit violations are possible:
+
+\begin{itemize}
+
+\item [(7-1)] Privacy of The Person.
+  Association with tattoo artists could imply having tattoos or other body modifications.
+\item [(7-2)] Privacy of Behaviour and Action.
+  Association with churches or political organizations could imply certain activities.
+\item [(7-5)] Privacy of Thoughts and Feelings.
+  Association with churches of political organizations could imply a certain intellectual attitude.
+\end{itemize}
+
+%\includegraphics{../diagrams/png/implicit-privacy-violation-matrix.png}
+\input{./figures/Implicit-Privacy-Violation-Matrix}
 
@@
 
+\subsection{Legal Privacy Requirements} \ref{sec:LEGAL_REQ}
+
 \subsubsection{Security Measures}

### 2014-10-14 PrivacyProtection-2014-09-09.tex
added bibtex for privacy protection references, still unclear on citation of legal documents

--- 2014-10-10 PrivacyProtection-2014-09-09.tex
@@
+% test citing
+\cite{CFried:Privacy}
+\cite{WarrenBrandeis:RightToPrivacy}
+\cite{sep-privacy}
+\cite{7ToP}
+\cite{Grimm:ItSecRefModel}
+\cite{ItSecGlossary}
+\cite{SociotechnicalArchitectureForOnlinePrivacy}
+\cite{EUDataProtectionDirective}
+\cite{EUConventionOnHumanRights}
+\cite{Bundesdatenschutzgesetz}
+\cite{Gutwirth:DefiningPrivacy}
+\cite{UKDataProtectionAct}
+
 \chapter{Legal and Ethical aspects of Privacy}
 \label{chap:privacy}
-
-\newcommand{\om}{[...]\xspace} % omission
-%\newcommand{\om}{\ldots} % omission
 
@@
 
-% \section{Histrocial Approaches to Privacy}
+\section{Histrocial Approaches to Privacy}
+
+\subsection{Aristotle}
+
+Public Sphere vs. Private Sphere of politics. This distinction is concerned with the politic (public) life of politicians in early democracies opposed by their domestic (private) life.
+
+\subsection{Warren and Brandeis}
+
+The Right to Privacy, 1890: Establish the term Informational Privacy.
+They argue innovations like photography and a growing press create the need for a new right to privacy, "the right to be left alone".
+
+The term informational privacy as conceptual foundation of privacy originates from the juristic discussion The Right to Privacy (1890) by Warren and Brandeis.
+The arise of new technologies like photography combined with a growing press created an increased publicity.
+Warren and Brandeis were worried that existing law like castle doctrines or libel and slander could not suffice.
+They felt a need for a new right to "protect the extent to which one's thoughts, sentiments, and emotions could be shared with others" \cite{1}.
+This right would be "the right to be left alone".
+
+The right to "protect the extent to which one's thoughts, sentiments, and emotions could be shared with others" by Warren and Brandeis was later extend to cover arbitrary information and is now kown as the privacy concept "...
+the control we have over information about ourselves ..." \cite{2}.
+
+\subsection{Fried}
+
+C. Freid, An Anatomy of Values, Harvard Univ. Press, 1970
+
+Defines privacy as the ability to control information about oneself. He argues that limited knowledge does not automatically create privacy.
 
@@
 
-\subsection{European Convention on Human Rights}
-
-The \emph{Europen Convetion on Human Rights} \cite{ECHR} was created by the members of the \emph{Council of Europe} (CoE) in 1953 as part of the aftermath of the second world war.
-It formulates universal rights of citizens against the state authority.
-All member states of the Council of Europe, which includes all EU members, have incorporated this convention into their national law.
-
-Article 8 of the ECHR contains a protection of personal data.
-
-\begin{quote}
-(1) Everyone has the right to respect for his private and family life, his home and his correspondence.
-
-(2) There shall be no interference by a public authority with the exercise of this right except such as is in accordance with the law and is necessary in a democratic society in the interests of national security, public safety or the economic well-being of the country, for the prevention of disorder or crime, for the protection of health or morals, or for the protection of the rights and freedoms of others.
-\end{quote}
-
-The protection of personal data under Article 8 is not an absolute law but must be considered in relation to other laws such as the freedom of expression (ECHR, Art. 10).
-The \emph{European Court of Human Rights} (ECtHR) overseas the implementation of this directive and has adopted a rather broad interpretation of the article.\footnote{\url{http://en.wikipedia.org/wiki/Article_8_of_the_European_Convention_on_Human_Rights}}
-
-In particular State actions of searching a persons home, gathering and storing information in a secret police file, and stopping a prisoner's communication have been ruled to interfere with Article 8.
-
-In the context of the Live+Gov project the ECHR is of limited relevance, since the Live+Gov System is offered by private companies and not directly by the State authority.
-
-\subsection{OECD Guidelines and CoE Convention 108}
-
-The emergence of information technologies that allow automated processing of personal data made more detailed rules for safeguarding the privacy of citizens necessary.
-In the light of this developments the OECD  \cite{OECD80} issued Guidelines on Protection of Privacy in 1980 which establish the following basic principles of data protection:
+\subsection{European Convention on Human Rights - Article 8}
+
+``(1) Everyone has the right to respect for his private and family life, his home and his correspondence.
+(2) There shall be no interference by a public authority with the exercise of this right except such as is in accordance with the law and is necessary in a democratic society in the interests of national security, public safety or the economic well-being of the country, for the prevention of disorder or crime, for the protection of health or morals, or for the protection of the rights and freedoms of others.'' \cite{2}
+
+``The European Court of Human has given this article a very broad interpretation in its jurisprudence.'' \cite{2}
+
+\subsection{EU Data Protection Directive}
+
+\emph{Directive 95/46/EC}
+
+EU directives in general do not hold any direct legal binding for EU citizens.
+Member states have to create their own legal implementation.
+
+``The directive regulates the processing of personal data regardless of whether such processing is automated or not.'' \cite{1} where \emph{personal data} is ``any information relating to an identified or identifiable natural person [...] (art.
+2a)'' and \emph{processing} is ``any operation or set of operations which is performed upon personal data [...] (art.
+2b)'' \cite{1}
+
+The ``identified or identifiable natural person'' is called \emph{data subject}.
+
+The sole responsibility for compliance is held by so called \textbf{controllers}.
+A controller is an actor ``which alone or jointly with others determines the purposes and means of the processing of personal data; (art 2d)''
+\cite{1}
+
+This rule is applicable ``whenever the controller uses equipment situated within the EU''\cite{1}.
+This will hold for every e-commerce provider \textbf{inside or outside} of the EU because the customer's computer is \emph{situated within the EU} anyways.
+
+``Personal data should not be processed at all, except when certain conditions are met'' \cite{1}.
+These conditions incorporate the seven OECD principals and are further categorized in the categories:
+
+\begin{itemize}
+\item
+\textbf{Transparency}: ``The data subject has the right to be informed when his personal data is being processed [...] (art. 10 and 11)'' \cite{1} and he has the right how and by whom it is processed.
+Additionally other specified requirements have to be met, i.e.~giving consent.
+\item
+\textbf{Legitimate purpose}: ``Personal data can only be processed for specified explicit and legitimate purposes and may not be processes further in a way incompatible with those purposes (art. 6b)''
+\cite{1}.
+\item
+\textbf{Proportionality}: ``Personal data may be processed only insofar as it is adequate [...] in relation to [its] purposes [...]; every reasonable step must be taken to ensure that data which are inaccurate or incomplete [...] are erased or rectified (art.
+6)'' \cite{1}
+\end{itemize}
+
+(\textbf{Note:} The principles incorporated by the EU directive seem to closely map to \emph{the seven principles for user privacy control}.)
+
+Member states have to create independent \textbf{supervisory authorities} for personal data processing where controllers have to register before they start to process any data.
+The record has to be stored in public register \cite{1}.
+
+Data transfer to third countries outside of the EU requires those countries to have a similar data protection level. Exceptions exists:
+
+\begin{itemize}
+\item
+  Save Harbor
+\item
+  Passenger Name Record Agreement
+\end{itemize}
+
+all between the US and the EU, although the US has no comparable law of data protection. 
+The US Supreme Court regards privacy as \textbf{implicit} right granted with the First Amendment. However, European law states privacy as \textbf{explicit} right (ECHR Art. 8).
+Additionally the US government follows a doctrine which favors the economy to implement self-regulation \cite{1}.
+
+\subsection{Implementation of the Data Protection Directive}
+
+\subsubsection*{Germany}
+
+Germany implements Directive 95/46 with the \emph{Federal Data Protection Act} from 2001 known as \emph{Bundesdatenschutzgesetz (BDSG)} \cite{1}\cite{2}\cite{6}.
+However, Germany has violated the directive in two points:
+
 \begin{enumerate}
-\item Collection Limitation Principle.
-  There should be limits to the collection of personal data and any such data should be obtained \om with the knowledge or consent of the data subject.
-\item Data Quality Principle.
-  Personal data should be relevant to the purposes for which they are to be used, and,  \om accurate, complete and kept up-to-date.
-\item Purpose Specification Principle.
-  The purposes for which personal data are collected should be specified \om and the subsequent use limited to the fulfillment of those purposes \om.
-\item Use Limitation Principle.
-  Personal data should not be disclosed \om except: a) with the consent of the data subject; or b) by the authority of law.
-\item Security Safeguards Principle.
-  Personal data should be protected by reasonable security safeguards \om.
-\item Openness Principle.
-  {\om} Means should be readily available of establishing the existence and nature of personal data, and the main purposes of their use, as well as the identity and usual residence of the data controller.
-\item Individual Participation Principle.
-  An individual should have the right
-  \begin{itemize}
-  \item to obtain from a data controller, or otherwise, confirmation of whether or not the data controller has data relating to him;
-  \item to have communicated to him, data relating to him \om
-  \item \om to have the data erased, rectified, completed or amended.
-  \end{itemize}
-\item Accountability Principle.
-  A data controller should be accountable for complying with measures which give effect to the principles stated above.
+\item
+The BDSG has become effective three years too late, thus the EC filed a treaty violation proceeding against Germany \cite{6}.
+
+\item
+The BDSG does not implement \textbf{independent} supervisory authorities. 
+The \emph{Bundesdatenschutzbeauftragter} is subordinate to the Ministry of Interior. 
+Although he is not subject to technical oversight (\emph{Fachaufsicht}), he is subject to staff supervision by the government (\emph{Rechtsaufsicht} durch die Bundesregierung und  \emph{Dienstaufsicht} durch das Innenministerium) and budget oversight (including approval of employees) by the ministry \cite{7}.
+Thus the EC filed a treaty violation proceeding against Germany, again \cite{6}. 
+In March 2010 Germany was found guilty of violation of Directive 95/46 by the ECJ \cite{6}.
 \end{enumerate}
 
-Where the following definitions are understood.
-\begin{itemize}
-\item \emph{personal data} means any information relating to an identified or identifiable person (\emph{data subject}).
-  An identifiable person is is one who can be identified either directly, by reference to a name or identification number or indirectly,
-  by one or more factors which make it possible to find out who the data subject is by conducting further research.
-\item a \emph{data controller} means any party who is competent to decide about the contents and use of personal data.
-\end{itemize}
-
-The OECD guidelines were adopted by EU law in the Strassbourg Convention 108 of the Council of Europe \cite{CONV108} in 1981.
-
-Relevant for our investigations is in particular, that sensor data collected from mobile devices is considered personal data if the individual can be identified based on the data by any direct or indirect means.
-Therefore the above principles apply.
-
-% UPDATE 2013: http://www.oecd.org/sti/ieconomy/privacy.htm
-
-\subsection{EU Data Protection Directive}
-
-The \emph{Directive 95/46/EC} of 1995 \cite{DIR95} is the fundamental regulation of data protection in the European Union.
-It was designed to give further substance to the Convention 108.
-In particular it makes the creation of an independent supervisory authority necessary (Art. 28).
-As this text is the main legal basis for our investigation we review it in detail.
-In Section \ref{sec:ECIMPL} the status of the implementation is discussed in some example cases.
-
-The interpretation and oversight of the Directive lies at the Court of Justice of the European Union (CJEU).
-A summary of the most important rulings can be found in Handbook on European data protection law \cite{EU_HANDBOOK_2014}.
-
-The ambition of the Directive is stated at the very beginning.
-\begin{quote}
-  (Art.
-1.1) In Accordance with this Directive the Member States shall protect the fundamental rights and freedoms of natural persons, and in particular their right to privacy with respect to the processing of personal data.
-\end{quote}
-
-The Directive follows the definition of personal data, data subject, and data controller of the OECD Guidelines (cf. Art. 2).
-Furthermore it introduces the role of a \emph{processor} as a person which processes, by wholly or partly automatic means (Art 3.), personal data on behalf of the controller.
-
-Data collection of intelligence agencies and the police is not restricted by the Directive (Art. 3).
-
-The most important articles are summarized in the following paragraphs.
-In order to keep the text more readable we have simplified the formulation by removing the formulation ``Member States shall provide that ...'' from each paragraph.
-\begin{itemize}
-\item (Art. 6.1.a) Personal data must be processed fairly and lawfully.
-
-[The meaning of the term 'fair processing' is entailed in the following articles.
-In particular transparency of processing, information of the data subject, and right to access the data are implied (cf. \cite{EU_HANDBOOK_2014})]
-
-\item (Art. 6.1.b) \emph{Specification of Purpose.} Personal data must be collected for specified, explicit and legitimate purposes and not be processed in a way incompatible with this purposes.
-
-Further processing of data for historical, statistical or scientific purposes shall not be considered incompatible \om.
-
-\item (Art. 6.1.c) Personal data must be adequate, relevant and non-excessive in relation to the purposes \om.
-
-\item (Art. 6.1.d) \emph{Accuracy.} Personal data must be accurate and \om every reasonable step must be taken to ensure that data which is inaccurate or incomplete \om are erased or rectified.
-
-\item (Art. 6.1e) Personal data must be kept in a form which permits identification for no longer than is necessary \om.
-% \item (Art. 6.2) It shall be for the controller to ensure that (Art. 6.1) is complied with.
-\end{itemize}
-
-In addition to strengthening the OECD Data Quality Principle in Article 6, the Directive requires a strong form of consent of the data subject, before any collection or processing of data can take place.
-
-\begin{itemize}
-\item (Art. 7). Personal data may be processed only if:
-  \begin{itemize}
-    \item [(a)] the data subject has unambiguously given his consent; or
-    \item [(..)] [exceptions are made for performance of a contracts with the data subject,  legal obligations of the controller, to protect vital interests of the data subject, the public interest and]
-     \item [(f)] the processing is necessary for purposes of legitimate interests pursued by the controller or by a third party to whom the data are disclosed, except where such interests are overridden by the interests \om of the data subject which require protection under Art.
-1(1).
-  \end{itemize}
-
-\item (Art.
-8.1) \emph{Special Categories.} Member Stats shall prohibit the processing of personal data revealing racial or ethnic origin, political opinions, religious or philosophical beliefs, trade-union membership, and the processing of data concerning health or sex life.
-[Art.
-8.2 lists several exceptions, including explicitly explicit consent of the data subject.]
-\end{itemize}
-
-The OECD Participation Principle is addressed and extended in the following articles.
-
-\begin{itemize}
-\item (Art. 10, 11) The controller \om must provide a data subject \om with the following information:
-  \begin{itemize}
-    \item [(a)] the identity of the controller \om
-    \item [(b)] the purpose of processing for which the data are intended;
-    \item [(c)] any further information \om that is necessary to guarantee fair processing in respect of the data subject.
-[This includes in particular the recipients of the data, the existence of the right to of access and the right to rectify the data.]
-  \end{itemize}
-
-\item (Art. 12) \emph{Right to Access.} Every data subject has the right to obtain from the controller
-\begin{itemize}
-  \item (a) \om confirmation as to whether or not data relating to him are being processed \om, communication to him \om the data undergoing processing and any available information as to their source, knowledge of the logic involved in any automatic processing \om;
-  \item (b) the rectification, erasure or blocking of data the procession of which does not comply with the provisions of this Directive \om;
-  \item (c) notification to third parties to whom the data have been disclosed \om.
-\end{itemize}
-
-%\item Confidentiality of Processing (Art. 16).
-%Any person under the authority of the controller or of the processor \om who has access to personal data must not process them except on instructions from the controller, unless he is required to do so by law.
-\end{itemize}
-
-A similar form of the Security Safeguard Principle is contained in Article 17.
-
-\begin{itemize}
-\item (Art. 17) \emph{Security of processing.}
-The controller must implement appropriate technical and organizational measures to protect personal data against accidental loss, alteration, unauthorized disclosure or access \om and against all other forms of unlawful processing.
-
-[The level of security has to be balanced against the risk of processing.]
-
-% Having regard to the state of the art and the cost of their implementation, such measures shall ensure a level of security appropriate to the risk represented by the processing and the nature of the data to be processed.
-\end{itemize}
-
-Furthermore the Directive demands all data processing to be reported to a supervisory authority.
-This obligation can be lifted if an internal data protection official is appointed, who is responsible in particular for keeping a processing register, that has to be made available on request.
-
-\begin{itemize}
-\item Notification (Art. 18. 1).
-The controller \om must notify the supervisory authority referred to in Article 28 before carrying out any wholly or partly automatic processing operation \om.
-
-\item (Art. 18.2).
-Simplification or exemption from notification may be provided only under the following conditions: \om where the controller \om appoints a personal data protection official, responsible in particular:
-\begin{itemize}
-\item for ensuring in an independent manner the internal application of the national provisions taken pursuant to this directive.
-\item for keeping the register of processing operations carried out by the controller, containing the items of information referred to in Article 21 (2)
-\end{itemize}
-thereby ensuring that the rights and freedoms of the data subjects are unlikely to be adversely affected by the processing operations.
-
-\item (Art. 19.1)
-The information to be given in the notification shall include at least:
-\begin{itemize}
-  \item [(a)] the name and address of the controller \om;
-  \item [(b)] the purpose \om of processing;
-  \item [(c)] a description of the category \om of the data \om;
-  \item [(d)] the recipients \om to whom the data might be disclosed;
-  \item [(e)] proposed transfer of data to third countries;
-  \item [(f)] a general description allowing a preliminary assessment to be made of the appropriateness of the measures taken pursuant of Article 17 to ensure the security of processing;
-\end{itemize}
-
-\item (Art. 21.2) \emph{Publication}.
-A register of processing information \om shall be kept at the supervisory authority.
-The register shall contain at least the information listed in Article 19.1 (a) to (e).
-
-The register may be inspected by any person.
-
-\item (Art.
-21.3).
-In relation to processing operations not subject to notification, that controllers \om make available at least the information referred to in Article 19.1 (a) to (e) in an appropriate form to any person on request.
-
-\item (Art. 28) \emph{Supervisory Authority.}
-Each Member State shall provide that one or more public authorities are responsible for monitoring the application \om of this Directive.
-
-These authorities shall act in complete independence in exercising this functions entrusted to them.
-\end{itemize}
-
-Data transfer to third countries outside of the EU requires those countries to have an adequate level of data protection.
-\begin{itemize}
-\item (Art. 25.1) \emph{Transfer to third countries.}
-The transfer to a third country of personal data \om may take place only if \om the third country in question ensures an adequate level of protection.
-\end{itemize}
-The European Commission can decide whether a third country ensures an adequate level of protection (Art 25.6).
-The USA, is not considered to do so.
-Data exchange between the US and EU is possible under the Safe Harbor regulation \cite{SAFE_HABOR}.
-Further Exceptions are provided by the Passenger Name Record Agreement \cite{PNR}.
-
-\subsection{Implementation of the Data Protection Directive}\label{sec:ECIMPL}
-
-\textbf{Germany}.
-Germany implements Directive 95/46 with the \emph{Bundesdatenschutzgesetz (BDSG)}
-of 2001. However, Germany has violated the directive in two points:
-\begin{enumerate}
-\item The BDSG has become effective three years too late, thus the EC filed a treaty violation proceeding against Germany.
-\item The BDSG does not implement independent supervisory authorities.
-The Bundesdatenschutzbeauftragter is subordinate to the Ministry of Interior.
-Although he is not subject to technical oversight (\emph{Fachaufsicht}), he is subject to staff supervision by the government (Rechtsaufsicht, Dienstaufsicht) and budget oversight by the ministry.
-In March 2010 Germany was found guilty of violation of Directive 95/46 by the ECJ.
-\end{enumerate}
+Germany still violates point 2!
 
 The states of Germany have their own implementation of Directive 95/46 (\emph{Landesdatenschutzgesetze}).
-Federal public authorities are only bound to their federal law.
-Churches are not subject the BDSG.
-
-\textbf{United Kingdom.}
-The UK implements Directive 95/46 with the \emph{Data Protection Act 1998} (DPA).
-
-The act is known for its high complexity: a manual record of phone numbers for business purposes could be hold subject to the DPA.
+Federal public authorities are only bound to their federal law. \cite{2}
+
+Churches are not subject the BDSG. \cite{2}
+
+\begin{itemize}
+
+\item
+\textbf{Roman Catholic Church:}
+\href{http://de.wikipedia.org/wiki/Anordnung_\%C3\%BCber_den_kirchlichen_Datenschutz}{Anordnung  über den kirchlichen Datenschutz (KDO)}
+
+\item
+\textbf{German Protestant Churches (Synode der Evangelischen Kirche in Deutschland):}
+\href{http://de.wikipedia.org/wiki/Datenschutzgesetz_der_Evangelischen_Kirche_in_Deutschland}{Kirchengesetz  über den Datenschutz der Evangelischen Kirche in Deutschland}
+\end{itemize}
+
+Data subjects have the following rights.
+
+\begin{itemize}
+\item
+Right to disclosure of whether data about them is stored and processed, which data is stored and processed and the data sources.
+\item
+Right to correction of false data.
+
+\item
+Right to file complaints to the supervisory authority.
+
+\item
+Right to have data deleted or blocked, but controllers can prohibit deletion in favour of blocking.
+
+\item
+Right to decline third party access to the data.
+\end{itemize}
+
+\subsubsection*{United Kingdom}
+
+The UK implements Directive 95/46 with the \emph{Data Protection Act 1998 (DPA)} \cite{1}\cite{3}.
+
+The act is known for its high complexity: a manual record of phone numbers for business purposes could be hold subject to the DPA \cite{3}.
 Although the act seems to fully cover the directive.
 Even higher restriction apply for \emph{``sensitive personal data''} (race, ethnicity, politics, religion, trade union status, health, sex life or criminal record), i.e.
-consent must be given freely and has to be explicit.
-
-The Act's definition of''personal data" covers any data that can be used to identify a living individual.
+\textbf{consent} must be given freely and has to be explicit.
+\cite{3}
+
+\emph{``The Act's definition of''personal data" covers any data that can be used to identify a living individual.
 Anonymised or aggregated data is not regulated by the Act, providing the anonymisation or aggregation has not been done in a reversible way.
-Individuals can be identified by various means including their name and address, telephone number or Email address.
-The Act applies only to data which is held, or intended to be held, on computers (`equipment operating automatically in response to instructions given for that purpose'), or held in a `relevant filing system'.
-
-In some cases even a paper address book can be classified as a `relevant filing system', for example diaries used to support commercial activities such as a salesperson's diary.
-
-The Freedom of Information Act 2000 modified the act for public bodies and authorities, and the Durant case modified the interpretation of the act by providing case law and precedent.
-
-The Data Protection Act creates rights for those who have their data stored, and responsibilities for those who store, process or transmit such data.
-The person who has their data processed has the right to:
-\begin{itemize}
-\item View the data an organization holds on them.
+Individuals can be identified by various means including their name and address, telephone number or
+Email address.
+The Act applies only to data which is held, or intended to be held, on computers (`equipment operating automatically in response to instructions given for that purpose'), or held in a `relevant filing system'.}
+
+\emph{In some cases even a paper address book can be classified as a `relevant filing system', for example diaries used to support commercial activities such as a salesperson's diary.}
+
+\emph{The Freedom of Information Act 2000 modified the act for public bodies and authorities, and the Durant case modified the interpretation of the act by providing case law and precedent.}
+
+\emph{The Data Protection Act creates rights for those who have their data stored, and responsibilities for those who store, process or transmit such data.
+The person who has their data processed has the right to:}
+
+\begin{itemize}
+
+\item
+View the data an organization holds on them.
 A `subject access request' can be obtained for a nominal fee.
 As of January 2014, the maximum fee is \pounds2 for requests to credit reference agencies, \pounds50 for health and educational request, and \pounds10 per individual otherwise.
-\item Request that incorrect information be corrected.
+
+\item
+Request that incorrect information be corrected.
 If the company ignores the request, a court can order the data to be corrected or destroyed, and in some cases compensation can be awarded.
-\item Require that data is not used in any way that may potentially cause damage or distress.
-\item Require that their data is not used for direct marketing. \cite{3}
+
+\item
+Require that data is not used in any way that may potentially cause damage or distress.
+
+\item
+Require that their data is not used for direct marketing. \cite{3}
 \end{itemize}
@@
 \emph{Safe harbor} in general is the legal concept to regulate that a certain conduct will be deemed \cite{4}, but in Germany the term is commonly used as synonym for the agreement between the EU and the US regarding Directive 95/46.)
-
-\subsection{EU Charter of Fundamental Rights}
-
-The Treaty of Lisbon of 2007, which layed down the fundamental functioning of the European Union
-includes the \emph{Charter of Fundamental Rights of the Europen Union} \cite{EUFR2010}.
-This Charta summarizes the full range of civil, political and economic rights of EU citizens and contains the following two articles that safeguard the privacy of the citizen.
-\begin{quote}
-  Article 7. Respect for private and family life.\\
-  Everyone has the right to respect for his or her private and family life, home and communications.
-
-  Article 8. Protection of personal data.
-  \begin{enumerate}
-    \item [(1)] Everyone has the right to the protection of personal data concerning him or her.
-    \item [(2)] Such data must be processed fairly for specified purposes and on the basis of the consent of the person concerned or some other legitimate basis laid down by law.
-      Everyone has the right of access to data which has been collected concerning him or her, and the right to have it rectified.
-    \item [(3)] Compliance with these rules shall be subject to control by an independent authority.
-  \end{enumerate}
-\end{quote}
-
-Although the Charter does not extend the pre-exiting Directive 95/46/EC, it amplifies the importance of this privacy protection as a fundamental right.
 
@@
 
-This type is also concerned with the ``protection against disclosure of personal matters'' through behaviour, however Clarke's distinction between ``casual observation \om systematic recording and storage of information about those activities'' is lifted.
+This type is also concerned with the ``protection against disclosure of personal matters'' through behaviour, however Clarke's distinction between ``casual observation [...] systematic recording and storage of information about those activities'' is lifted.
 
@@
 best understand as \emph{Biometric Privacy}. Friedewald et al. paraphrase
-it as \emph{``\om the right to keep body functions and body
-characteristics \om private''}. \textbf{Accelerometer},
+it as \emph{``[...] the right to keep body functions and body
+characteristics [...] private''}. \textbf{Accelerometer},
 \textbf{Rotation Vector} and \textbf{Gyroscope} measure the physical

### 2014-10-14 PrivacyProtection-2014-09-09.tex
added threat example diagrams for man-in-the-middle and intransparency
did some refactoring on the sensor privacy matrix section

--- 2014-10-13 PrivacyProtection-2014-09-09.tex
@@
+% test citing
+%\cite{CFried:Privacy}
+%\cite{WarrenBrandeis:RightToPrivacy}
+%\cite{sep-privacy}
+%\cite{7ToP}
+%\cite{Grimm:ItSecRefModel}
+%\cite{ItSecGlossary}
+%\cite{SociotechnicalArchitectureForOnlinePrivacy}
+%\cite{EUDataProtectionDirective}
+%\cite{EUConventionOnHumanRights}
+%\cite{Bundesdatenschutzgesetz}
+%\cite{Gutwirth:DefiningPrivacy}
+%\cite{UKDataProtectionAct}
+
 \chapter{Legal and Ethical aspects of Privacy}
@@
 
-\subsubsection{Threats}
+\subsubsection{Threat Specification}
+
+\textbf{Man In The Middle:}
+Lorem ipsum dolor suit amet, ...
+
+\textbf{Exploit:}
+Lorem ipsum dolor suit amet, ...
+
+\textbf{Corruption:}
+Lorem ipsum dolor suit amet, ...
+
+\textbf{Intrasparency:}
+Lorem ipsum dolor suit amet, ...
+
 \input{./figures/LG-Threats}
+\input{./figures/Threat-Example-1}
+\input{./figures/Threat-Example-2}
 \input{./figures/ThreatTable}
 
-%\begin{itemize}
-%
-%\item
-%  \textbf{Unauthorized access to privacy sensitive data}, caused by
-%  \begin{itemize}
-%  \item
-%    \textbf{Excessive Data Mining} Linking sensor data provided by
-%    citizens with additional sources can produce more privacy sensitive
-%    data.
-%  \item
-%    \textbf{Corrupt Local Authorities} Local Authorities have certain data
-%    access, they could hand over this access for monetary reasons.
-%  \item
-%    \textbf{Corrupt/Unhappy Staff} Staff members also have a certain data
-%    access and they also could hand over this access for monetary reasons
-%    or as a form of payback for unfair treatment.
-%  \item
-%    \textbf{Competent Attackers} Competent Attackers are always a threat
-%    for IT-systems. Normaly they have monetary reasons to exploit a
-%    system, but there is also a possibility for proof-of-concept like
-%    attacks.
-%  \end{itemize}
-%\end{itemize}
-%
-%
-\paragraph{Sensor Privacy Matrix (Figure \ref{figure:Live+Gov Sensor-Privacy Matrix})}
+\subsubsection{Threat Risk Evaluation}
+
+% Todo: use sensor privacy matrix and implicit sensor privacy matrix for threat risk evaluation
+
+\textbf{Sensor Privacy Matrix}
 
@@
 
-\subparagraph{Privacy of The Person}
+\begin{itemize}
+
+\item
+\textbf{Privacy of The Person}
 
@@
 
-\subparagraph{Privacy of Data and Image}
+\item
+\textbf{Privacy of Data and Image}
 
@@
 
-\subparagraph{Privacy of Location and Space}
+\item
+\textbf{Privacy of Location and Space}
 
@@
 
-\subparagraph{Privacy of Associtation}
+\item
+\textbf{Privacy of Associtation}
 
@@
 
+\end{itemize}
+
+\subsubsection{Security Requirement Specification}
+
 \subsection{Step 3. Plan Development}

### 2014-10-15 PrivacyProtection-2014-09-09.tex
Added %%% separators

@@
 In order to ensure their privacy, System Provider would have to implement additional mechanisms, which allow Citizens to exercise control of their data.
-But this will eventually raise the complexity of the monitoring and mining system.
+This will not only raise the complexity of the monitoring and mining system, System Providers also have to layout the complexity in a comprehensible manner.
+To effectively enable Citizens to preserve their privacy, they need to know what happens with their data.
 
@@
 \subsubsection{Vulnerabilities}
+\label{subsubsection:Vulnerabilities}
 This section outlines the vulnerabilities (Figure \ref{figure:Live+Gov Vulnerabilities}) of the proposed monitoring and mining system.
@@
 
+\subsection{Step 2. Potential Analysis}
+
+\subsubsection{Threats}
+This section outlines the potential threats (Figure \ref{figure:Live+Gov Threats}) for Citizen privacy in the proposed monitoring and mining System.
+We derive threats from known conflicts of interests \ref{subsubsection:Conflicts of Interests} and vulnerabilities \ref{subsubsection:Vulnerabilities}
+=======
 \subsection{Step 2. Potential Analysis}

### 2014-10-15 PrivacyProtection-2014-09-09.tex
Updated IT Systems description

--- 2014-10-14 PrivacyProtection-2014-09-09.tex
@@
 
-In the context of the Live+Gov project the ECHR is of limited relevance, since the Live+Gov System is offered by private companies and not directly by the State authority.
+The Live+Gov systems can violate this fundamental human right if it is used to gather information about the private life without the knowledge and consent of the citizen.
+
+% In the context of the Live+Gov project the ECHR is of limited relevance, since the Live+Gov System is offered by private companies and not directly by the State authority.
 
@@
 
-The act is known for its high complexity: a manual record of phone numbers for business purposes could be hold subject to the DPA.
-Although the act seems to fully cover the directive.
-Even higher restriction apply for \emph{``sensitive personal data''} (race, ethnicity, politics, religion, trade union status, health, sex life or criminal record), i.e.
+%TODO: Reformulate Wikipedia Text
+
+%% The act is known for its high complexity: a manual record of phone numbers for business purposes could be hold subject to the DPA.
+%% Although the act seems to fully cover the directive.
+%% Even higher restriction apply for \emph{``sensitive personal data''} (race, ethnicity, politics, religion, trade union status, health, sex life or criminal record), i.e.
 consent must be given freely and has to be explicit.
 
-The Act's definition of personal data covers any data that can be used to identify a living individual.
-Anonymised or aggregated data is not regulated by the Act, providing the anonymisation or aggregation has not been done in a reversible way.
-
-The Freedom of Information Act 2000 modified the act for public bodies and authorities, and the Durant case modified the interpretation of the act by providing case law and precedent.
-
-The Data Protection Act creates rights for those who have their data stored, and responsibilities for those who store, process or transmit such data.
-The person who has their data processed has the right to:
-\begin{itemize}
-\item View the data an organization holds on them.
-A `subject access request' can be obtained for a nominal fee.
-As of January 2014, the maximum fee is \pounds2 for requests to credit reference agencies, \pounds50 for health and educational request, and \pounds10 per individual otherwise.
-\item Request that incorrect information be corrected.
-If the company ignores the request, a court can order the data to be corrected or destroyed, and in some cases compensation can be awarded.
-\item Require that data is not used in any way that may potentially cause damage or distress.
-\item Require that their data is not used for direct marketing. \cite{3}
-\end{itemize}
+%% The Act's definition of personal data covers any data that can be used to identify a living individual.
+%% Anonymised or aggregated data is not regulated by the Act, providing the anonymisation or aggregation has not been done in a reversible way.
+
+%% The Freedom of Information Act 2000 modified the act for public bodies and authorities, and the Durant case modified the interpretation of the act by providing case law and precedent.
+
+%% The Data Protection Act creates rights for those who have their data stored, and responsibilities for those who store, process or transmit such data.
+%% The person who has their data processed has the right to:
+%% \begin{itemize}
+%% \item View the data an organization holds on them.
+%% A `subject access request' can be obtained for a nominal fee.
+%% As of January 2014, the maximum fee is \pounds2 for requests to credit reference agencies, \pounds50 for health and educational request, and \pounds10 per individual otherwise.
+%% \item Request that incorrect information be corrected.
+%% If the company ignores the request, a court can order the data to be corrected or destroyed, and in some cases compensation can be awarded.
+%% \item Require that data is not used in any way that may potentially cause damage or distress.
+%% \item Require that their data is not used for direct marketing. \cite{3}
+%% \end{itemize}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%
+
 The \emph{world view} contains all components describing the current state.
@@
 \end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%
+
 The \emph{concept view} is a realization of the the potential view of the system.
@@
 
-A citizen (a) carries a mobile device running the \emph{sensor collector} (d) application.
-The sensor collector application collects sensor data (accelerometer, GPS, GSM, ...) and sends it to the \emph{data center} (c).
+A \emph{citizen} (a) carries a mobile device running the \emph{sensor collector} application (d).
+The sensor collector application is able to collect various kinds of sensor data (accelerometer, GPS, GSM, ...) and can sent the raw data back to the \emph{data center} (c).
 The sensor collector can also be able to perform certain data mining operations.
+
 Examples of such data mining operations include human activity recognition,the detection of service lines, detection of characters and faces from images.
 
+In the Live+Gov Project mobile devices are used in particular for the following activities:
+\begin{enumerate}
+\item collection of GPS samples,
+\item mining of Human activities (HAR) based on accelerometer samples,
+\item collection of reports consisting of an image, free text and selected categories.
+\end{enumerate}
+
 The \emph{data center} stores and processes sensor data collected with the sensor collector application.
-Again the amount
-
-Vice versa, it sends beneficial information (traffic jam reports, bus schedule) for \textit{Citizens} to the mobile device.
-The \textit{System Provider (b)} provides and operates technical infrastructure like the \textit{Data Center} and the \textit{Report Tool (f)}.
-The \textit{Report Tool} queries the \textit{Data Center} for aggregated data to visualize in form of charts and other means suitable to help understanding of monitored \textit{Citizens}. 
-\textit{Local Authorities (c)} use the \textit{Report Tool} to get information in order to understand citizen movement and improve public services.
-
-In such systems, the most valuable information is not the raw sensor data, the beneficial information for \textit{Local Authorities} is produced via \textit{Data Mining} (Human Activity Recognition: detect running, walking, standing, ...).
-Usually, \textit{Data Mining} is heavily conducted within the \textit{Data Center}, but simpler mining processes can also be conducted on mobile devices.
+It can also take into account data obtained from third parties, like the current positions of trains.
+The data center can sent mining end products (traffic jam reports, bus schedule) back to the mobile device of the citizen.
+
+In the Live+Gov Project data centers, in particular, perform the following activities:
+\begin{enumerate}
+\item storage of login credentials, name, and email address for each citizen,
+\item storage of collected GPS samples (user id, timestamp, GPS location),
+\item storage of HAR results (user id, timestamp, HAR result),
+\item detection of service lines based received GPS samples (SLD),
+\item storage of SLD results (userid, timestamp, SLD result),
+\item sent back SLD results to mobile device,
+\item storage of received reports (user id, timestamp, report),
+\item detection of inherent patterns of the received reports.
+\end{enumerate}
+
+The \emph{System Provider} (b) provides and operates technical infrastructure like the Data Center and the \emph{Reporting Tool} (f).
+The Reporting Tool queries the Data Center for aggregated data to visualize in form of charts and other means suitable to help understanding of monitored citizens.
+
+\textit{Local Authorities} (c) use the reporting tool to get information in order to understand citizen movement and improve public services.
+In such systems, the most valuable information for local authorities is not the raw sensor data, but aggregated views on the mining end products.
+
+In the Live+Gov Project the reporting tools allows, in particular the following queries:
+\begin{enumerate}
+\item show aggregate information about which routes citizens take starting from a given location,
+\item show the average waiting time of a citizen for each bus stop,
+\item show routes where citizens where running to catch a bus,
+\item show locations of all reports in a given time window,
+\item visualize detected patterns in the report data,
+\item view individual reports.
+\end{enumerate}
 

### 2014-10-15 PrivacyProtection-2014-09-09.tex
Updated Actors and CoI sections.

@@
 
-This section outlines the human actors previously introduced within the IT system architecture (\ref{subsubsection:it-systems}).
-It is described which individual interests they have invested in monitoring systems comparable to the Live+Gov system. 
-Also the additional actor \textit{Criminal} is introduced as an possible obstacle.
-
-\textbf{Citizen:} 
-Citizens are persons who take part in the monitoring system as the monitored subjects.
-They offer their privacy sensitive data, because by participating they expect a higher level of comfort in their lives through the offered services in return.
-E.g. they will have access to real time bus schedules and traffic jam reports, localized social networking or they generally benefit from public improvement by Local Authorities.
-
-However, they have a legal right to privacy and are interested in having that privacy respected by others.
-For example, all humans have a natural interest in physical wellbeing and personal health. 
-If someone monitors their immediate location (GPS) or frequently used routes (stalking), people with access to that information could attack and harm citizens.
-
-In general, citizens are not interested in having information about them exploited by others. 
+This section describes the human actors previously introduced within the IT system architecture (\ref{subsubsection:it-systems}).
+Also the additional actor \textit{Criminal} is introduced as an external person with harmful intentions.
+
+\textbf{Citizen.}
+Citizens are persons who use the mobile device as users of the provided software.
+Their main motivation for using the software is to gain a higher level of convenience in their daily activities.
+For example they may have access to real time bus schedules or reports about traffic jams, that help them to avoid long waiting times.
+Also they generally benefit from improvements of public infrastructure by local authorities, which is triggered by issue reports.
+
+By using the application citizens are sharing personal information like name and address, as well as data gathered from mobile sensor with the service provider.
+This data can be exploited in ways that are harmful to the citizen.
+This need to protect this protection is manifest in several laws and constitutions as the right to privacy and data protection.
+The citizen has a vital interest in having his legal rights protected and enforced.
+
+If the right to privacy is violated, there is a magnitude of potential harms that interfere with other interests of the citizen.
 This includes just annoying spam, where their technical identity is used to send unwanted commercials.
-But it also includes more severe phishing attacks, which try to rob citizens of their TAN numbers and disrespect their financial interests.
-
-Such financial interests of monitored citizens, the interest in not losing / gaining money, could be disrespected on an even larger scale.
-Collected data could be used for scoring and analysis, which may affect the pricing of policies or credits, if sold to insurances or banks.
-
-So citizens have an inherent interest in transparency and disclosure of who uses their data for which purposes, as it may affect their physical and financial wellbeing.
-
-In short, Citizens are interested in ...
-\begin{itemize}
-\item ... confidential use of privacy sensitive data
-\item ... legitimate use of privacy sensitive data
-\item ... proportional use of privacy sensitive data
-\item ... successful use of privacy sensitive data
-\end{itemize}
-
-\textbf{Criminal:}
-Criminals are persons who do not have legal access to the IT systems of a proposed monitoring system.
-Usually, monetary interests lead their activities.
+More severe phishing attacks can exploit personal information, and try to manipulate citizens into disclosing credentials like TAN numbers and disrespect their financial interests.
+Information about the current location (GPS) or frequently used routes (stalking) can be used to attack and directly harm the health of the citizen.
+Information about medical conditions inferred by sensor data (e.g.
+fintness trackers) are of interest to insurance companies, which may affect the pricing of policies and can interfere with financial interests of the citizen.
+
+Also it is well established (cf. \cite{GuardienMassSurveillance}) that the very act of being monitored can have impact on mental health and performance, promotes distrust and breeds conformity.
+
+In short, citizens are interested in
+\begin{itemize}
+\item convenience
+\item legitimate use of personal data
+\item physical wellbeing and health
+\item financial profit
+\item not being monitored.
+\end{itemize}
+
+\textbf{Criminal.}
+Criminals are persons who do not have privileged access to the IT systems,
+and are willing to break laws and norms in order to promote their interests.
+
 For example they want to obtain access to critical systems to steal sensitive data or to get the system under their control.
@@
 
-Because Local Authorities are involved in the general outline of a public monitoring system, the possibility for politically motivated attacks cannot be ignored. 
-Criminals could want to harm or destroy the systems in order to damage the reputation of Local Authorities (politicians or other officials) or to make a political statement of their own, e.g.
-"Stop Surveillance!".
+Because local authorities are involved in the general outline of the Live+Gov system, the possibility for politically motivated attacks is given.
+
+Criminals could want to harm or destroy the systems in order to damage the reputation of local authorities (politicians or other officials) or to make a political statement of their own.
 
@@
 
-In short, Criminals are interested in ...
-\begin{itemize}
-\item ... financial profit
-\item ... high social standing
-\item ... political power
-\end{itemize}
-
-\textbf{System Provider:}
-System Providers provide and operate the technical infrastructure (hardware and software) of the proposed monitoring system.
-They are normal companies which offer to provide a service, so different roles of employees are involved:
- an administrator, who maintains and operates the running system;
- a programmer/developer, who develops the system;
- a support manager, who handles customer relations;
- etc.
+In short, Criminals are interested in
+\begin{itemize}
+\item financial profit
+\item political activism
+\item social standing.
+\end{itemize}
+
+\textbf{System Provider.}
+System Providers operate the technical infrastructure (hardware and software) of the IT System.
+They are private companies and legal persons in their own right, but also employ a number of people with diverging interests, including:
+administrators, who maintain and operate the running system;
+programmer/developer, who develop the system;
+a support manager, who handles customer relations.
 
@@
 Moreover, unhappy employees can produce higher costs through sick days.
-Ill employees stay on payroll, but stress the project schedule by splitting the workload on fewer heads.
 The worst case scenario could be, that a discontent employee gets angry and steals data or harms the running systems.
 
-At last, the financial success of System Providers depends on the task complexity.
-The complexity of a task has to be in reasonable bounds, so that System Providers can complete it within time, with a satisfying quality.
-If a task has a higher complexity than expected, financial loss is almost determined.
+At last, the financial success of system providers depends on the task complexity of the maintained infrastructure.
+The complexity of a task has to be in reasonable bounds, so that system providers can complete it within time, with a satisfying quality.
+If a task has a higher complexity than expected, financial loss is almost certain.
 Either System Providers need to hire additional competence to meet schedule and requirements.
-Or System Providers they stress the time-line, which also results in a higher man-hour salary ratio and additionally endangers customer satisfaction.
+Or system providers they stress the time-line, which also results in a higher man-hour salary ratio and additionally endangers customer satisfaction.
 Ultimately, high task complexities can affect employee happiness, if employees cannot complete it within their claim to professional excellence.
 
-In short, System Providers are interested in ...
-\begin{itemize}
-\item ... financial profit
-\item ... good working conditions
-\item ... professional excellence
-\item ... reasonable complexity
-\end{itemize}
-
-\textbf{Local Authority:} 
-Local Authorities are public offices (ministry, agency, department, ...) or other external public service providers which act as direct customers of Service Providers.
-They order a proposed monitoring and mining system specialized for their needs. 
-For example a low level department, like departments for public mobility, orders a monitoring and mining system to better understand and improve urban traffic flow.
-
-Such systems are heavy investments, so naturally Local Authorities are interested in a profitable return.
+In short, System Providers are interested in
+\begin{itemize}
+\item financial profit
+\item good working conditions
+\item professional excellence
+\item manageable complexity.
+\end{itemize}
+
+\textbf{Local Authority.}
+Local authorities are public offices (ministry, agency, department, ...) or other external public entities which act as direct customers of service providers.
+They purchase a system specialized for their needs.
+For example a department for urban mobility, orders a system to better understand usage patterns and make improvement to the urban traffic flow.
+
+Such systems are investments, and so naturally local authorities are interested in a profitable return, like increased ticket sales.
 However, the return of investment is not directly of financial nature.
-Like Service Providers their financial gain depends on customer satisfaction. 
-Customers for Local Authorities are either citizens, who use their services, or politicians, who order their services.
-The happiness of both sides is interdependent. 
-Citizens are happy customers if the services, e.g. public mobility, work well. 
-If citizens are happy, it is more likely that politicians gain reputation, as they organize the public services through Local Authorities.
-
-One way of achieving these business objectives is by utilizing business intelligence. 
-A proposed monitoring and mining system can provide such intelligence.
-For public mobility services it can provide traffic jam detection to maximize traffic flow.
-Or it can provide user profiling to maximize (targeted) publicity.
-
-Additionally, since Local Authorities act like corporations comparable to Service Providers, they are also interested in good working conditions.
-As discontent employees may harm the system by disclosure of business intelligence.
-
-In short, Local Authorities are interested in ...
-\begin{itemize}
-\item ... business intelligence
-\item ... political reputation
-\item ... good working conditions
+Like service providers their financial gain depends on customer satisfaction.
+Customers for local authorities are either citizens, who use their services, or politicians, who order their services.
+The satisfaction of both sides is interdependent.
+
+Citizens are satisfied customers if the services, e.g. public mobility, work well.
+If citizens are happy, it is more likely that politicians gain reputation, as they organize the public services through local authorities.
+
+The first important step of improving the public services is by obtaining business intelligence.
+For the urban mobility scenario the Live+Gov systems provide insight in form of traffic jam detection and usage pattern mining, which allow local authorities to focus their efforts to the most important sites.
+
+Additionally, since local authorities act like corporations comparable to service providers, they are also interested in good working conditions.
+Discontent employees may harm the system by e.g. disclosure of privacy sensitive data.
+
+In short, local authorities are interested in
+\begin{itemize}
+\item financial profit
+\item political reputation
+\item business intelligence
+\item good working conditions.
 \end{itemize}
@@
 This section outlines the Conflicts of Interests (Figure \ref{figure:Live+Gov Conflicts of Interests}) between the actors of the proposed IT system architecture.
+
 The individual interests of all actors is already described in the previous section and are not elaborated any further.
@@
 
-\textbf{System Complexity vs Privacy:}
-System Providers offer a service to Local Authorities, which is to provide and maintain a monitoring and mining system, e.g. for public mobility.
-This system shall produce business intelligence, so that Local Authorities can improve their public services.
-This task in it self has a high technical complexity and is the sole asset with financial return for System Providers.
-However, this task operates on privacy sensitive data provided by monitored Citizens.
-In order to ensure their privacy, System Provider would have to implement additional mechanisms, which allow Citizens to exercise control of their data.
-But this will eventually raise the complexity of the monitoring and mining system.
-
-\textbf{Business Intelligence vs Privacy:}
-Local Authorities order a monitoring and mining system from System Providers, which allows them to produce business intelligence for public services.
-The system is a heavy investment for Local Authorities, so they are interested in as much intelligence as possible to achieve a profitable return.
-But the intelligence is the result of data mining conducted on privacy sensitive data of monitored Citizens.
+\textbf{System Complexity vs Privacy of the Citizen.}
+System Providers offer a service to local authorities, which is to provide and maintain the IT system.
+This system shall produce business intelligence, so that local authorities can improve their public services.
+This task in it self has a high technical complexity and is the sole asset with financial return for system provider.
+
+However, this task operates on privacy sensitive data provided by monitored citizens.
+In order to ensure their privacy, system provider would have to implement additional mechanisms, which allow citizens to exercise control of their data.
+This rises the complexity of the IT system considerably and thereby also raises the costs for its production and maintenance.
+
+\textbf{Business Intelligence vs Privacy.}
+Local authorities order a monitoring and mining system from system providers, which allows them to produce business intelligence for public services.
+The system is an investment for local authorities, so they are interested in as much intelligence as possible to achieve a profitable return.
+
+The gained intelligence is the result of data mining conducted on privacy sensitive data of participating citizens.
 They are interested in the successful usage of their data, in a sense that they are also benefactors, e.g. improvement of public mobility.
-However, the main interest of Citizens lies in maintaining control over their data.
-In order do that, they need disclosure of the exact purposes their data is used for, so they can select which data to share.
-These purposes cannot be (secretly) exceeded.
-
-\textbf{Money vs Privacy:}
+The main interest of citizens lies in maintaining control over their data and protecting their rights to privacy.
+In order do that, they need full disclosure of the processing steps and the purposes their data is used for, and to be given a choice whether such processing should be allowed for their own data.
+
+\textbf{Financial Profit of Criminal vs Privacy.}
 Criminals can gain financial profit from stealing privacy sensitive data.
-For example by selling raw contact information to advertisers or by selling mined data to scoring services.
-In such cases, Citizens lose complete control over their data.
-
-\textbf{Money vs Professional Reputation:}
+For example by selling raw contact information to advertisers or by selling mined data to insurance companies, or intermediaries like scoring companies.
+In such cases, citizens lose complete control over their data.
+
+\textbf{Financial Profit of Criminal vs Reputation of System Providers.}
 Criminals have various business models as optional foundation for attacks on System Providers.
@@
 Gathered information can be sold, zombie servers can be leased.
-A successful attack proves the technical competence of System Providers wrong and subsequently harms their professional reputation.
+
+A successful attack proves the technical competence of system providers wrong and subsequently harms their professional reputation.
 This can lead to a loss of future customers or a decrease of stock price for registered companies.
-Eventually the financial interests of System Providers are endangered.
-
-\textbf{Political Power vs Political Reputation:}
+Eventually also the financial interests of system providers are endangered.
+
+\textbf{Political Activism vs Reputation of Local Authority.}
 Besides monetary reasons, Criminals can be motivated by political reasons to attack the monitoring and mining system.
-Criminals can break the system to make a political statement of their own, e.g. "Stop Surveillance!".
-Or they can steal user data to prove the system insecure.
-Both would harm the political reputation of Local Authorities, who endangered the privacy of monitored Citizens.
-Criminals would gain political power, because their point of view would seem to be valid.
+Criminals can break the system to make a political statement of their own,
+or they can steal user data to prove the system insecure.
+Both would harm the reputation of local authorities, who endangered the privacy of the citizens.
 

### 2014-10-15 PrivacyProtection-2014-09-09.tex
started summary fried

--- 2014-10-15 PrivacyProtection-2014-09-09.tex
@@
 
-Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes in \cite{Gutwirth}:
+Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes:
 
 \begin{quote}
-The notion of privacy remains out of the grasp of every academic chasing it. Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.
+The notion of privacy remains out of the grasp of every academic chasing it.
+Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.
+\cite{Gutwirth}
 \end{quote}
 
-Many privacy researchers seem to ``focus on the ways in which privacy can be infringed'' \cite{1}.
-Thus they try to create taxonomies of \emph{privacy harms} instead of taxonomies of \emph{privacy types}.
-Those two differ in the respect that the former focuses on threats to prohibit whereas the latter focuses on values to protect.
-So one should rather evaluate what aspects are precious about privacy and develop measures to ensure their security than only forbid single actions against privacy \cite{1}.
+%Many researchers seem to only ``focus on the ways in which privacy can be infringed'' \cite{7ToP}.
+%Thus they try to create taxonomies of \emph{privacy harms} instead of taxonomies of \emph{privacy types}.
+%Those two differ in the respect that the former focuses on threats to prohibit whereas the latter focuses on values to protect.
+%So one should rather evaluate what aspects are precious about privacy and develop measures to ensure their security than only forbid single actions against it \cite{7ToP}.
+
+Many researchers seem to only ``focus on the ways in which privacy can be infringed'' \cite{7ToP}.
+Thus they invest a great amount of work in defining threats to prohibit instead of describing why privacy is so valuable to us.
+This way around one could derive measures to ensure and secure its value \cite{7ToP}.
+One particular scholar who does this in the context of digital monitoring is Charles Fried.
+He questioned, why we are so sensitive to violations of privacy.
+But he did not assert privacy as an intrinsic value by itself, he rather sated:
+
+\begin{quote}
+Privacy is not simply an absence of information about us in the minds of others;
+rather it is \textbf{the control we have over information about ourselves}. \cite{CFried:Privacy}
+\end{quote}
+
+According to Fried privacy is a \textit{rational context}.
+Which means that one is aware of such context's existence during a rational action.
+If we are to share private information with others, we are most likely aware that this action is privacy related.
+In that case we are able to selectively disclose information along the two dimensions of quantity and quality.
+This leads Fried to his thesis, that privacy is one's ability to create and modulate his or hers social relationships, namely: friendship, love and trust. \cite{CFried:Privacy}
+
+Depending on the conversation partner, we change the degree of intimate information we share if it is a total stranger, colleague, close friend or lover.
+With close friends or lovers we share information of great intimacy we do not share with anyone else.
+Moreover, we trust those friends or lovers to not reveal information about us to others by respecting their privacy.
+Trust needs the possibility of unknown failure.
+If we would constantly monitor our partners, they cannot fail unnoticed nor can they willingly share that information with us.
+Thus they could not trust us anymore. \cite{CFried:Privacy}
+
+So privacy or the its possibility, according to Fried, is the foundation of our core relations: friendship, love and trust.
+And thus it is valuable, because those relations are essential to human society \cite{CFried:Privacy}. 
+Using his anatomy of privacy as foundation of our analysis is suitable because of two points.
+At first, Fried's study on the understanding of privacy provided a great contribution to the research on the same term in philosophy \cite{sep-privacy} and computer science \cite{SociotechnicalArchitectureForOnlinePrivacy}.
+Secondly, despite the fact his text was published in 1970, he already included technologies to its viewpoint that did not only monitor location, but also record biometric data.
 

### 2014-10-15 PrivacyProtection-2014-09-09.tex

--- 2014-10-15 PrivacyProtection-2014-09-09.tex
@@
 
-However, this task operates on privacy sensitive data provided by monitored citizens.
-In order to ensure their privacy, system provider would have to implement additional mechanisms, which allow citizens to exercise control of their data.
-This rises the complexity of the IT system considerably and thereby also raises the costs for its production and maintenance.
+\textbf{System Complexity vs Privacy:}
+System Providers offer a service to Local Authorities, which is to provide and maintain a monitoring and mining system, e.g. for public mobility.
+This system shall produce business intelligence, so that Local Authorities can improve their public services.
+This task in it self has a high technical complexity and is the sole asset with financial return for System Providers.
+However, this task operates on privacy sensitive data provided by monitored Citizens.
+In order to ensure their privacy, System Provider would have to implement additional mechanisms, which allow Citizens to exercise control of their data.
+This will not only raise the complexity of the monitoring and mining system, System Providers also have to layout the complexity in a comprehensible manner.
+To effectively enable Citizens to preserve their privacy, they need to know what happens with their data.
 
@@
 Subjects who are not aware of these things cannot effectively preserve control and thus lose their privacy.
+
+\subsection{Step 2. Potential Analysis}
+
+\subsubsection{Threats}
+This section outlines the potential threats (Figure \ref{figure:Live+Gov Threats}) for Citizen privacy in the proposed monitoring and mining System.
+We derive threats from known conflicts of interests \ref{subsubsection:Conflicts of Interests} and vulnerabilities \ref{subsubsection:Vulnerabilities}
 

### 2014-10-20 PrivacyProtection-2014-09-09.tex
local changes

--- 2014-10-15 PrivacyProtection-2014-09-09.tex
@@
 
-Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes:
+Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes in \cite{Gutwirth}:
 
 \begin{quote}
-The notion of privacy remains out of the grasp of every academic chasing it.
-Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.
-\cite{Gutwirth}
+The notion of privacy remains out of the grasp of every academic chasing it. Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.
 \end{quote}
 
-%Many researchers seem to only ``focus on the ways in which privacy can be infringed'' \cite{7ToP}.
-%Thus they try to create taxonomies of \emph{privacy harms} instead of taxonomies of \emph{privacy types}.
-%Those two differ in the respect that the former focuses on threats to prohibit whereas the latter focuses on values to protect.
-%So one should rather evaluate what aspects are precious about privacy and develop measures to ensure their security than only forbid single actions against it \cite{7ToP}.
-
-Many researchers seem to only ``focus on the ways in which privacy can be infringed'' \cite{7ToP}.
-Thus they invest a great amount of work in defining threats to prohibit instead of describing why privacy is so valuable to us.
-This way around one could derive measures to ensure and secure its value \cite{7ToP}.
-One particular scholar who does this in the context of digital monitoring is Charles Fried.
-He questioned, why we are so sensitive to violations of privacy.
-But he did not assert privacy as an intrinsic value by itself, he rather sated:
-
-\begin{quote}
-Privacy is not simply an absence of information about us in the minds of others;
-rather it is \textbf{the control we have over information about ourselves}. \cite{CFried:Privacy}
-\end{quote}
-
-According to Fried privacy is a \textit{rational context}.
-Which means that one is aware of such context's existence during a rational action.
-If we are to share private information with others, we are most likely aware that this action is privacy related.
-In that case we are able to selectively disclose information along the two dimensions of quantity and quality.
-This leads Fried to his thesis, that privacy is one's ability to create and modulate his or hers social relationships, namely: friendship, love and trust. \cite{CFried:Privacy}
-
-Depending on the conversation partner, we change the degree of intimate information we share if it is a total stranger, colleague, close friend or lover.
-With close friends or lovers we share information of great intimacy we do not share with anyone else.
-Moreover, we trust those friends or lovers to not reveal information about us to others by respecting their privacy.
-Trust needs the possibility of unknown failure.
-If we would constantly monitor our partners, they cannot fail unnoticed nor can they willingly share that information with us.
-Thus they could not trust us anymore. \cite{CFried:Privacy}
-
-So privacy or the its possibility, according to Fried, is the foundation of our core relations: friendship, love and trust.
-And thus it is valuable, because those relations are essential to human society \cite{CFried:Privacy}. 
-Using his anatomy of privacy as foundation of our analysis is suitable because of two points.
-At first, Fried's study on the understanding of privacy provided a great contribution to the research on the same term in philosophy \cite{sep-privacy} and computer science \cite{SociotechnicalArchitectureForOnlinePrivacy}.
-Secondly, despite the fact his text was published in 1970, he already included technologies to its viewpoint that did not only monitor location, but also record biometric data.
+Many privacy researchers seem to ``focus on the ways in which privacy can be infringed'' \cite{1}.
+Thus they try to create taxonomies of \emph{privacy harms} instead of taxonomies of \emph{privacy types}.
+Those two differ in the respect that the former focuses on threats to prohibit whereas the latter focuses on values to protect.
+So one should rather evaluate what aspects are precious about privacy and develop measures to ensure their security than only forbid single actions against privacy \cite{1}.
 
@@
 In order to ensure their privacy, System Provider would have to implement additional mechanisms, which allow Citizens to exercise control of their data.
-This will not only raise the complexity of the monitoring and mining system, System Providers also have to layout the complexity in a comprehensible manner.
-To effectively enable Citizens to preserve their privacy, they need to know what happens with their data.
+But this will eventually raise the complexity of the monitoring and mining system.
 
@@
 \subsubsection{Vulnerabilities}
-\label{subsubsection:Vulnerabilities}
 This section outlines the vulnerabilities (Figure \ref{figure:Live+Gov Vulnerabilities}) of the proposed monitoring and mining system.
@@
 
-\subsection{Step 2. Potential Analysis}
-
-\subsubsection{Threats}
-This section outlines the potential threats (Figure \ref{figure:Live+Gov Threats}) for Citizen privacy in the proposed monitoring and mining System.
-We derive threats from known conflicts of interests \ref{subsubsection:Conflicts of Interests} and vulnerabilities \ref{subsubsection:Vulnerabilities}
-=======
 \subsection{Step 2. Potential Analysis}
@@
 
-\paragraph{The 7 C's of user privacy control (Figure \ref{figure:The Seven Cs of User Privacy Control})}
+\textbf{The 7 C's of user privacy control (Figure \ref{figure:The Seven Cs of User Privacy Control})}
 
@@
 
-\subparagraph{Comprehension}
+\begin{itemize}
+
+\item
+\textbf{Comprehension}
 
@@
 
-\subparagraph{Consciousness \textbf{(critical!)}}
+\item
+\textbf{Consciousness \textbf{(critical!)}}
 
@@
 
-\subparagraph{Choice}
+\item
+\textbf{Choice}
 
@@
 
-\subparagraph{Consent}
+\item
+\textbf{Consent}
 
@@
 
-\subparagraph{Context}
+\item
+\textbf{Context}
 
@@
 
-\subparagraph{Confinement}
+\item
+\textbf{Confinement}
 
@@
 
-\subparagraph{Consistency}
+\item
+\textbf{Consistency}
 
@@
 
-\paragraph{The 2 Steps of the 7 C's (Figure \ref{figure:The 2 Steps of the 7 Cs of User Privacy Control})}
+\end{itemize}
+
+\textbf{The 2 Steps of the 7 C's (Figure \ref{figure:The 2 Steps of the 7 Cs of User Privacy Control})}
 
@@
 
-\subparagraph{(I) Enable \emph{Adequate} Control}
+\begin{enumerate}
+
+\item
+\textbf{Enable \emph{Adequate} Control}
 
@@
 
-\subparagraph{(II) Enable \emph{Actual} Control}
+\item
+\textbf{Enable \emph{Actual} Control}
 
@@
 
-% \section{References}
+\end{enumerate}

### 2014-10-21 PrivacyProtection-2014-09-09.tex
refined introduction to 7ToP

@@
 One particular scholar who does this in the context of digital monitoring is Charles Fried.
-He questioned, why we are so sensitive to violations of privacy.
+He questioned, why we are intuitively sensitive to violations of privacy.
 But he did not assert privacy as an intrinsic value by itself, he rather sated:
@@
 With close friends or lovers we share information of great intimacy we do not share with anyone else.
-Moreover, we trust those friends or lovers to not reveal information about us to others by respecting their privacy.
+Moreover, we trust those persons to not reveal information about us to others by respecting their privacy.
 Trust needs the possibility of unknown failure.
@@
 
-* We follow Fried in defining privacy as the ability to control data about personal information.
-
-* What is the relation to legal aspects?
-
-* What should be personal information?
-
-* Use Seven Privacy Types be Friedewald, Finn and Wrigtht.
+%* We follow Fried in defining privacy as the ability to control data about personal information.
+%
+%* What is the relation to legal aspects?
+%
+%* What should be personal information?
+%
+%* Use Seven Privacy Types be Friedewald, Finn and Wrigtht.
 
@@
 
-Seven Types of Privacy based on four types of privacy by Clarke.
-Clarke's four types are outdated by contemporary technologies and no longer adequate.
-In order to fix this Friedewald, Finn and Wright extend the former four to the now introduced seven types privacy as follows:
+Although we use Fried's definition of privacy as one's control over information about oneself, we do not feel safe relying only on that to conduct a privacy protection analysis.
+We need additional means to analyze in more detail which aspects of our lives need protection.
+Friedewald, Finn and Wright \cite{7ToP} introduced a privacy taxonomy called \textit{Seven Types of Privacy} that can be of help.
+The seven types of privacy are an extension to the four types of privacy by Roger Clarke \cite{RClarke:4ToP}:
+\begin{itemize}
+\item Privacy of the Person
+\item Privacy of Personal Behaviour
+\item Privacy of Personal Communication
+\item Privacy of Personal Data
+\end{itemize}
+However, Friedewald et al. felt that Clarke's taxonomy are technologically outdated and no longer adequate.
+In order to fix this they extend the former four to the now introduced seven types privacy as follows:
 
@@
 
-\item \textbf{Privacy of the Person.}
-
-This type is the right keep body functions and body characteristics private.
-
-Examples include:
-\begin{itemize}
-\item Body Characteristics: weight, height, shoulder width, \ldots
-\item Biometric Properties: Fingerprints, DNA Sequence, \ldots
-\item Medical Conditions: Orthopedic conditions (e.g.~limping), Having a cold, \ldots
-\end{itemize}
-
-\item \textbf{Privacy of Behaviour and Action}
-
-This type is also concerned with the ``protection against disclosure of personal matters'' through behaviour, however Clarke's distinction between ``casual observation \om systematic recording and storage of information about those activities'' is lifted.
-
-Examples include:
-\begin{itemize}
-\item regular visit at church, bakery, doctors
-\item sexual habits
-\item political activities
-\end{itemize}
-
-\item \textbf{Privacy of Communication}
-
-It ``aims to avoid the interception of communications'' either electronic or face-to-face.
-
-Examples include:
-\begin{itemize}
-\item Contents of letters and emails
-\item Personal direct communication
-\item Right to free discussion, i.e.~without third parties listening
-\end{itemize}
-
-\item \textbf{Privacy of Data and Image}
-
-This type is concerned with ``making sure that individual's data is not automatically available to other individuals and organizations''.
-This is Informational Privacy in an intuitive sense.
-
-Examples include:
-\begin{itemize}
-\item Phone number
-\item IP address
-\item Public-administrative Data (Date of Birth, Melderegister)
-\item Data held by organizations, like Banks or Insurance Companies
-\item All data that is stored in online services (facebook)
-\end{itemize}
-
-\item \textbf{Privacy of Thoughts and Feelings}
-
-This type is the right ``not to share their thoughts or feelings or to have those thoughts or feeling (sic!) revealed''.
-It is concerned with (automatic) emotion detection.
-This type is the counterpart the Privacy of the Person like body and mind are counterparts of one another (dualism).
-
-Examples include:
-\begin{itemize}
-\item Current feeling: Depression, tiredness, stressed, awake
-\item Thoughts in general.
-\end{itemize}
-
-\item \textbf{Privacy of Location and Space}
-
-This type is the right ``to move about in public or semi-public space without being identified, tracked or monitored''.
-Additionally this type is concerned with the protection of one's home and private places (``right to solitude''.
-
-Examples include:
-\begin{itemize}
-\item GPS position tracking
-\item Location of home address
-\end{itemize}
-
-\item \textbf{Privacy of Association}
-
-This type is the right ``to associate with whomever [one] wishes, without being monitored''.
-It is concerned with the protection against the automatic record of one's contacts.
-It does not imply that one is monitored because of the associations.
-
-Examples include:
-\begin{itemize}
-\item Friends
-\item Joining of organizations (e.g.~political parties)
-\end{itemize}
+\item 
+\textbf{Privacy of the Person}
+
+This privacy type is generally concerned with one could best understand as biometric privacy. 
+Friedewald et al. paraphrase it as \emph{``\om the right to keep body functions and body characteristics \om private''}. 
+This includes but is not limited to measures like weight, height or shoulder width;
+biometric identifiers like fingerprints and DNA sequences;
+or medical conditions such as limping or having a cold. \cite{7ToP}
+
+\item 
+\textbf{Privacy of Behaviour and Action}
+
+This privacy type is concerned with one's activities in public as well as in private spaces. 
+It includes but is not limited to religious practices, political activities and sexual preferences or habits. \cite{7ToP}
+
+\item 
+\textbf{Privacy of Communication}
+
+This privacy type is concerned with one's communication in a broad sense.
+It includes written correspondence, but also conversations conducted either vis-a-vis or via electronic devices.
+Friedewald et al. put it as the right to free discussion without unknown interception by third parties. \cite{7ToP}
+
+\item 
+\textbf{Privacy of Data and Image}
+
+This privacy type is concerned with the secrecy of one's personal data, especially its automatic disclosure to other individuals and organizations.
+It includes data such as paychecks, insurance information or records of public administration.
+However, it also refers to pictures taken without consent and digital identifiers like IP addresses or social security numbers. \cite{7ToP}
+
+\item 
+\textbf{Privacy of Thoughts and Feelings}
+
+This privacy type is the counterpart to Privacy of the Person like body and mind are counterparts of one another.
+Comparable to the Privacy of Data and Image, Friedewald et al. state that one's thoughts and feelings must not be automatically revealed to others.
+This could simply happen by the disclosure of one's diary or by technologies which allow emotion detection through biometric means.
+One's body temperature or iris reflexes might infer stress or excitation. \cite{7ToP}
+
+\item 
+\textbf{Privacy of Location and Space}
+
+This privacy type is concerned with one's movements in public spaces and the inviolateness of one's private spaces.
+Friedewald et al. qualify the first dimension as one's right to move without being identified, tracked or monitored.
+The second dimension is qualified as one's general right to solitude, especially the right to the inviolability of the home. \cite{7ToP}
+
+\item 
+\textbf{Privacy of Association}
+
+This privacy type is also put as group privacy.
+Friedewald et al. state that one must have the possibility with whomever without being recorded.
+Associations like friends or organizations such as political parties must not automatically  be recorded because one associates with them, and vice versa.
+
 \end{enumerate}
@@
 Our findings are summarized in Figure \ref{figure:Implicit Privacy Violation Matrix}.
+
+%\includegraphics{../diagrams/png/implicit-privacy-violation-matrix.png}
+\input{./figures/Implicit-Privacy-Violation-Matrix}
 
@@
 
-%\includegraphics{../diagrams/png/implicit-privacy-violation-matrix.png}
-\input{./figures/Implicit-Privacy-Violation-Matrix}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The \textbf{Privacy of The Person} is generally concerned with one could
-best understand as \emph{Biometric Privacy}. Friedewald et al. paraphrase
-it as \emph{``\om the right to keep body functions and body
-characteristics \om private''}. \textbf{Accelerometer},
-\textbf{Rotation Vector} and \textbf{Gyroscope} measure the physical
-movement of the mobile device on all three axes. If the mobile device
-is carried \emph{``normally''} its safe to say that those sensors also
-measure the moments of its carrier. So his privacy is infringed
-regarding biometric behaviour, as it is captured automatically.
-(\textbf{Note:} This is not to be confused with the \textbf{Privacy of
-Behaviour and Action} which is used for the social aspects of behaviour
-action, e.g.~praying, sexual habits or political activities.)
+The \textbf{Privacy of The Person} is generally concerned with one could best understand as \emph{Biometric Privacy}. 
+Friedewald et al. paraphrase it as \emph{``\om the right to keep body functions and body characteristics \om private''}. 
+\textbf{Accelerometer}, \textbf{Rotation Vector} and \textbf{Gyroscope} measure the physical movement of the mobile device on all three axes. 
+If the mobile device is carried \emph{``normally''} its safe to say that those sensors also measure the moments of its carrier. 
+So his privacy is infringed regarding biometric behaviour, as it is captured automatically.
+(\textbf{Note:} This is not to be confused with the \textbf{Privacy of Behaviour and Action} which is used for the social aspects of behaviour action, e.g.~praying, sexual habits or political activities.)
 
@@
 
-The \textbf{Privacy of Data and Image} demands, that
-\emph{``individual's data is not automatically available to other
-individuals''}. This type of privacy is trivially threatened because
-here sensor data is individual data, a priori. So every sensor violates
-the privacy of data and image, as data is transported into a foreign
-system where operators have access to it.
+The \textbf{Privacy of Data and Image} demands, that \emph{``individual's data is not automatically available to other individuals''}. 
+This type of privacy is trivially threatened because here sensor data is individual data, a priori. 
+So every sensor violates the privacy of data and image, as data is transported into a foreign system where operators have access to it.
 
@@
 
-According to Friedewald et al., the \textbf{Privacy of Location and
-Space} is concerned with one's \emph{``right to move about in public or
-semi-public space without being identified, tracked or monitored.''}.
-This is the location aspect of this type. The space aspect is concerned
-with one's \emph{``right to solitude''}, which generally includes one's
-right to an inviolate home and other private spaces. Obviously the
-\textbf{GPS} and \textbf{GSM} sensors violate such right about not being
-tracked, because they reveal the position of the mobile device and its
-carrier. The \textbf{GSM} sensor gives the exact cell, the mobile device
-has registered with at the current moment. The \textbf{GPS} sensor gives
-the current longitude and latitude, the current global position of the
-mobile device and its carrier, although there is some artificial
-inaccuracy within civil use.
-
-The \textbf{WLAN} and \textbf{Bluetooth} sensors record lists of the
-currently available local wireless networks and bluetooth clients. If
-such are known stationary entities, those sensors are considered as
-dangerous as the \textbf{GSM} sensor for the carriers locational
-privacy.
-
-The \textbf{Magnetic field} sensor is not regarded very dangerous to the
-carrier's privacy, because it does not allow very precise localization.
-But it can limit the possibilities for the global position of the mobile
-device. Here, it is just named for completeness sake.
+According to Friedewald et al., the \textbf{Privacy of Location and Space} is concerned with one's \emph{``right to move about in public or semi-public space without being identified, tracked or monitored.''}.
+This is the location aspect of this type. The space aspect is concerned with one's \emph{``right to solitude''}, which generally includes one's right to an inviolate home and other private spaces. 
+Obviously the \textbf{GPS} and \textbf{GSM} sensors violate such right about not being tracked, because they reveal the position of the mobile device and its carrier. 
+The \textbf{GSM} sensor gives the exact cell, the mobile device has registered with at the current moment. 
+The \textbf{GPS} sensor gives the current longitude and latitude, the current global position of the mobile device and its carrier, although there is some artificial inaccuracy within civil use.
+
+The \textbf{WLAN} and \textbf{Bluetooth} sensors record lists of the currently available local wireless networks and bluetooth clients. 
+If such are known stationary entities, those sensors are considered as dangerous as the \textbf{GSM} sensor for the carriers locational privacy.
+
+The \textbf{Magnetic field} sensor is not regarded very dangerous to the carrier's privacy, because it does not allow very precise localization.
+But it can limit the possibilities for the global position of the mobile device. 
+Here, it is just named for completeness sake.
 
@@
 
-The \textbf{Privacy of Association} sates that everyone has the
-\emph{``right to associate with whomever they wish, without being
-monitored [automatically without reasonable suspicion]''}. This
-includes individuals and organizations. \textbf{WLAN} and
-\textbf{Bluetooth} sensors provide the ability to monitor such
-associations if their lists contain known entities. If one frequently
-connects with an organizational wireless network, e.g.~an university
-network, an association can be deduced (student or staff). The same goes
-for the \textbf{Bluetooth} sensor, if it is stationary. Additionally, if
-the recorded bluetooth clients are mobile, it is more or less possible
-to deduce association with the technical identity of (yet) anonymous
-individuals.
-
-Additionally, the \textbf{GSM} sensor could provide the association with
-the GSM operator (\textbf{NEEDS TO BE VERIFIED!}).
+The \textbf{Privacy of Association} sates that everyone has the \emph{``right to associate with whomever they wish, without being monitored [automatically without reasonable suspicion]''}. 
+This includes individuals and organizations. 
+\textbf{WLAN} and \textbf{Bluetooth} sensors provide the ability to monitor such associations if their lists contain known entities. 
+If one frequently connects with an organizational wireless network, e.g.~an university network, an association can be deduced (student or staff).
+The same goes for the \textbf{Bluetooth} sensor, if it is stationary.
+Additionally, if the recorded bluetooth clients are mobile, it is more or less possible to deduce association with the technical identity of (yet) anonymous individuals.
+
+Additionally, the \textbf{GSM} sensor could provide the association with the GSM operator (\textbf{NEEDS TO BE VERIFIED!}).
 

### 2014-10-21 PrivacyProtection-2014-09-09.tex
added figure for implicit sensor privacy matrix, fixed some errors

@@
 
-Although we use Fried's definition of privacy as one's control over information about oneself, we do not feel safe relying only on that to conduct a privacy protection analysis.
+Although we use Fried's definition of privacy as one's control over information about oneself, we do not feel safe conducting a privacy protection analysis relying only him.
 We need additional means to analyze in more detail which aspects of our lives need protection.
@@
 \end{itemize}
-However, Friedewald et al. felt that Clarke's taxonomy are technologically outdated and no longer adequate.
+However, Friedewald et al. felt that Clarke's taxonomy is technologically outdated and no longer adequate.
 In order to fix this they extend the former four to the now introduced seven types privacy as follows:
@@
 \input{./figures/LG-Sensor-Privacy-Matrix}
+\input{./figures/LG-Implicit-Sensor-Privacy-Matrix}
 

### 2014-10-21 PrivacyProtection-2014-09-09.tex
Added some TODOs. Renamed Crimial -> External.

@@
 
-\textbf{United Kingdom.} %TODO: HOW MUCH IS COPY PASTE FROM WIKIPEDIA?
+\textbf{United Kingdom.}
 The UK implements Directive 95/46 with the \emph{Data Protection Act 1998} (DPA).
@@
 
-%% TODO %%
-
-%* We follow Fried in defining privacy as the ability to control data about personal information.
-%
-%* What is the relation to legal aspects?
-%
-%* What should be personal information?
-%
-%* Use Seven Privacy Types be Friedewald, Finn and Wrigtht.
+%TODO: Iterate Section. Exmplain Triangular Relationship:
+- Subject
+- Trusted Person
+- Third Party which shall NOT obtain sensitive information.
 
@@
 
-\item 
-\textbf{Privacy of the Person}
+\item \textbf{Privacy of the Person}
 
@@
 
-\item 
-\textbf{Privacy of Behaviour and Action}
+\item \textbf{Privacy of Behaviour and Action}
 
@@
 
-\item 
-\textbf{Privacy of Communication}
+\item \textbf{Privacy of Communication}
 
@@
 
-\item 
-\textbf{Privacy of Data and Image}
+\item \textbf{Privacy of Data and Image}
 
@@
 
-\item 
-\textbf{Privacy of Thoughts and Feelings}
+\item \textbf{Privacy of Thoughts and Feelings}
 
@@
 
-\item 
-\textbf{Privacy of Location and Space}
+\item \textbf{Privacy of Location and Space}
 
@@
 
-\item 
-\textbf{Privacy of Association}
+\item \textbf{Privacy of Association}
 
@@
 The complexity of our systems and the variety of threads make a great number of counter measures plausible.
-We approach this complexity with the aid of a general security analysis model developed in \cite{Grimm}.
+We approach this complexity with the aid of a general security analysis model developed in \cite{Grimm:ItSecRefModel}.
 We give a brief introduction to this model and perform a IT Security Analysis with respect to the privacy asset for our system.
@@
 
-We follow the Reference Model for IT Security Analyis as described in \cite{Grimm}.
+We follow the Reference Model for IT Security Analyis as described in \cite{Grimm:ItSecRefModel}.
 It supersedes earlier efforts by \cite{Avizienis}.
@@
 All identifiable stakeholders of the system under study.
-Typical actors include, users, developers, clients and criminals.
+Typical actors include, users, developers, clients and externals.
 
@@
 
-In the example of the criminal-user-conflict, the criminal has to exploit a vulnerability, e.g. a weak password, to gain access to the private data abou the user.
+In the example of the criminal-user-conflict, the criminal has to exploit a vulnerability, e.g. a weak password, to gain access to the private data about the user.
 
@@
 
-An example of a security requirement is that a given communication channel shall not be infringed by criminals.
+An example of a security requirement is that a given communication channel shall not be infringed by externals.
 \end{itemize}
@@
 
-The \emph{concept view} is a realization of the the potential view of the system.
+The \emph{concept view} is a realization of the potential view of the system.
 It specifies important interactions that require further planning.
@@
 \item \textbf{Security Measures.}
-It describes a plan of sufficient measures to secure the intended interactions and to aviod the unintended interactions.
+It describes a plan of sufficient measures to secure the intended interactions and to avoid the unintended interactions.
 Each security measure targets a vulnerability of the system in order to reduce a risk for a certain thread.
@@
 This section describes the human actors previously introduced within the IT system architecture (\ref{subsubsection:it-systems}).
-Also the additional actor \textit{Criminal} is introduced as an external person with harmful intentions.
+Also the additional actor \textit{External} is introduced as a person with no special access rights.
 
@@
 
-\textbf{Criminal.}
-Criminals are persons who do not have privileged access to the IT systems,
-and are willing to break laws and norms in order to promote their interests.
-
-For example they want to obtain access to critical systems to steal sensitive data or to get the system under their control.
+%TODO:
+- Nondisclosure of sensitive information to peers of the citizen.
+
+\textbf{External.}
+Externals are persons who do not have privileged access to the IT systems, and are willing to break laws, security constrains and norms in order to promote their interests.
+
+If the external is in some kind of relationship to the citizen, like a friendship or business partnership, the external can have a direct interest in gaining information about the citizen in order to increase their power.
+
+Another common interest of an external is financial profit. For example they want to obtain access to critical systems to steal sensitive data or to get the system under their control.
 Controlled systems could be leased as part of a bot net.
@@
 
-Criminals could want to harm or destroy the systems in order to damage the reputation of local authorities (politicians or other officials) or to make a political statement of their own.
-
-Another possible motivation for criminal activities could be social appreciation.
+Externals could want to harm or destroy the systems in order to damage the reputation of local authorities (politicians or other officials) or to make a political statement of their own.
+
+Another possible motivation for external  activities could be social appreciation.
 A hacker could attack critical infrastructure just to prove his skills.
 
-In short, Criminals are interested in
-\begin{itemize}
+In short, externals are interested in
+\begin{itemize}
+\item increase power over citizen
 \item financial profit
@@
 \input{./figures/LG-Conflicts-Of-Interests}
-
-\textbf{System Complexity vs Privacy of the Citizen.}
-System Providers offer a service to local authorities, which is to provide and maintain the IT system.
-This system shall produce business intelligence, so that local authorities can improve their public services.
-This task in it self has a high technical complexity and is the sole asset with financial return for system provider.
-
-\textbf{System Complexity vs Privacy:}
+%TODO: Adapt Figure
+
+\textbf{System Complexity vs Privacy.}
 System Providers offer a service to Local Authorities, which is to provide and maintain a monitoring and mining system, e.g. for public mobility.
@@
 
-\textbf{Financial Profit of Criminal vs Privacy.}
-Criminals can gain financial profit from stealing privacy sensitive data.
+\textbf{Power of External vs. Privacy.}
+Externals which are in a social relation to the citizen can have an interest in obtaining further information in order to gain power.
+In the most simplistic example this could be a man wanting monitor the activities of his spouse.
+Another example is a Government spying on it's citizens in order to suppress opposition.
+An additional twist in the last example is, that there can be legal regulations that require the service provider to support the Governments invasion of the citizens privacy.
+
+\textbf{Financial Profit of External vs Privacy.}
+Externals can gain financial profit from stealing privacy sensitive data.
 For example by selling raw contact information to advertisers or by selling mined data to insurance companies, or intermediaries like scoring companies.
@@
 
-\textbf{Financial Profit of Criminal vs Reputation of System Providers.}
-Criminals have various business models as optional foundation for attacks on System Providers.
+\textbf{Financial Profit of External vs Reputation of System Providers.}
+Externals have various business models as optional foundation for attacks on System Providers.
 For instance, they can try to invade the infrastructure for e-espionage reasons, to get control over servers to create a bot-net or to steal user data.
@@
 \textbf{Political Activism vs Reputation of Local Authority.}
-Besides monetary reasons, Criminals can be motivated by political reasons to attack the monitoring and mining system.
-Criminals can break the system to make a political statement of their own,
+Besides monetary reasons, externals can be motivated by political reasons to attack the monitoring and mining system.
+Externals can break the system to make a political statement of their own,
 or they can steal user data to prove the system insecure.
@@
 
-\textbf{Unencrypted Data Transmission:}
-The proposed monitoring and mining system uses HTTP to exchange data between the Sensor Collector, the Data Center and the Report Tool.
-Per default, HTTP is a clear text protocol.
-This means, one can intercept the connection and read all sensitive information, which is send between the components.
-That is: passwords, raw sensor data and data mining results
-
-\textbf{Inadequate Access Rules:}
-The proposed IT system infrastructure has various accesses to privacy sensitive data.
-System Provider staff has access to Data Center hardware and software like databases, web-servers and other inspection tools.
-Local Authority staff has access to the Report Tool.
-This all enables staff members to have potential access to privacy sensitive information.
-Those accesses have to be secured against unauthorized third parties, e.g. Criminals.
-Moreover, we need to ensure that no single person has to many access rights.
-For example, a system administrator should not be able to secretly download the whole database on a flash-drive.
-
-\textbf{Insufficiently Secured Infrastructure:}
+\textbf{Insecure Infrastructure.}
 The proposed monitoring system consists of many hardware and software components, each with its own concrete weaknesses.
@@
 
-\textbf{Unaware Monitoring Subjects:}
+\textbf{Insecure Data Transmission.}
+The proposed monitoring and mining system uses HTTP to exchange data between the Sensor Collector, the Data Center and the Report Tool.
+Per default, HTTP is a clear text protocol.
+This means, one can intercept the connection and read all sensitive information, which is send between the components.
+That is: passwords, raw sensor data and data mining results
+
+\textbf{Unhappy Employees.} Susceptible for social engineering.
+%TODO: Add description
+
+\textbf{Inadequate Access Rules.}
+The proposed IT system infrastructure has various accesses to privacy sensitive data.
+System Provider staff has access to Data Center hardware and software like databases, web-servers and other inspection tools.
+Local Authority staff has access to the Report Tool.
+This all enables staff members to have potential access to privacy sensitive information.
+Those accesses have to be secured against unauthorized third parties.
+Moreover, we need to ensure that no single person has to many access rights.
+For example, a system administrator should not be able to secretly download the whole database on a flash-drive.
+
+\textbf{Unaware Monitoring Subjects.}
 We define privacy as one's ability to control information about oneself. 
@@
 Subjects who are not aware of these things cannot effectively preserve control and thus lose their privacy.
+
+%TODO
+- No Privacy Policy
+- No Information about Processing, Access rules, 3rd party disclosure
 
@@
 \input{./figures/LG-Sensor-Privacy-Matrix}
-\input{./figures/LG-Implicit-Sensor-Privacy-Matrix}
+%\input{./figures/LG-Implicit-Sensor-Privacy-Matrix}
 

### 2014-10-21 PrivacyProtection-2014-09-09.tex
Added a first list of threats.

@@
 
+%TODO: Explain relation to DP Directive
+- Control over information in law
+
 \subsection{The Seven Types of Privacy}
 
-Although we use Fried's definition of privacy as one's control over information about oneself, we do not feel safe conducting a privacy protection analysis relying only him.
-We need additional means to analyze in more detail which aspects of our lives need protection.
-Friedewald, Finn and Wright \cite{7ToP} introduced a privacy taxonomy called \textit{Seven Types of Privacy} that can be of help.
-The seven types of privacy are an extension to the four types of privacy by Roger Clarke \cite{RClarke:4ToP}:
+In Fried's definition of privacy as control over "information about onself'', the specification of what constitues such information remains open.
+There is a vast amount of information that relates to a person and we need to get a better understanding in order to perform a thorough analysis.
+To this end we use the the categorization by Friedewald, Finn and Wright \cite{7ToP} called the \textit{Seven Types of Privacy}.
+The seven types of privacy are an extension to the four types of privacy by Roger Clarke \cite{RClarke:4ToP}, which are:
 \begin{itemize}
@@
 \end{itemize}
-However, Friedewald et al. felt that Clarke's taxonomy is technologically outdated and no longer adequate.
-In order to fix this they extend the former four to the now introduced seven types privacy as follows:
+It is important to note, that this categories do not form no taxonomy, since the categories are not mutually exclusive.
+For instance a written email is considered personal communication as well as personal data stored on a computer.
+
+Moreover, Friedewald et al. argue that Clarke's taxonomy is outdated and no longer adequate in order to describe the privacy aspect of our modern, technology driven, world.
+In order to address this shortcoming they extend the former four to the now introduced seven types privacy as follows:
 
@@
 \item \textbf{Privacy of the Person}
-
 This privacy type is generally concerned with one could best understand as biometric privacy.
@@
 biometric identifiers like fingerprints and DNA sequences;
-or medical conditions such as limping or having a cold. \cite{7ToP}
+or medical conditions such as limping or having a cold.
 
 \item \textbf{Privacy of Behaviour and Action}
-
 This privacy type is concerned with one's activities in public as well as in private spaces.
-It includes but is not limited to religious practices, political activities and sexual preferences or habits. \cite{7ToP}
+It includes but is not limited to religious practices, political activities and sexual preferences or habits.
 
 \item \textbf{Privacy of Communication}
-
 This privacy type is concerned with one's communication in a broad sense.
 It includes written correspondence, but also conversations conducted either vis-a-vis or via electronic devices.
-Friedewald et al. put it as the right to free discussion without unknown interception by third parties. \cite{7ToP}
+Friedewald et al. put it as the right to free discussion without unknown interception by third parties.
 
 \item \textbf{Privacy of Data and Image}
-
-This privacy type is concerned with the secrecy of one's personal data, especially its automatic disclosure to other individuals and organizations.
+This privacy type is concerned with the secrecy of personal data, especially its automatic disclosure to other individuals and organizations.
 It includes data such as paychecks, insurance information or records of public administration.
-However, it also refers to pictures taken without consent and digital identifiers like IP addresses or social security numbers. \cite{7ToP}
+However, it also refers to pictures taken without consent and digital identifiers like IP addresses or social security numbers.
 
 \item \textbf{Privacy of Thoughts and Feelings}
-
 This privacy type is the counterpart to Privacy of the Person like body and mind are counterparts of one another.
@@
 This could simply happen by the disclosure of one's diary or by technologies which allow emotion detection through biometric means.
-One's body temperature or iris reflexes might infer stress or excitation. \cite{7ToP}
+One's body temperature or iris reflexes might infer stress or excitation.
 
 \item \textbf{Privacy of Location and Space}
-
-This privacy type is concerned with one's movements in public spaces and the inviolateness of one's private spaces.
+This privacy type is concerned with one's movements in public spaces and the protection of private spaces.
 Friedewald et al. qualify the first dimension as one's right to move without being identified, tracked or monitored.
-The second dimension is qualified as one's general right to solitude, especially the right to the inviolability of the home. \cite{7ToP}
+The second dimension is qualified as one's general right to solitude, especially the right to the inviolability of the home.
 
 \item \textbf{Privacy of Association}
-
 This privacy type is also put as group privacy.
 Friedewald et al. state that one must have the possibility with whomever without being recorded.
-Associations like friends or organizations such as political parties must not automatically  be recorded because one associates with them, and vice versa.
-
+Associations like friends or organizations such as political parties must not automatically be recorded because one associates with them, and vice-versa.
 \end{enumerate}
@@
 
+%TODO: Update Threat Table
+
+0. Bad Architecture
+-------------------
+Description: The Service Provider does not offer tools for the citizen to control his data.
+COI:            Privacy vs. System Complexity
+Vulneratbility: Bad system design
+Asset:          Control over Data
+
+1. Surveillance
+---------------
+Description:   External infiltrates infrastructure in order to spy on citizen
+COI:           Privacy vs. External Power
+Vulnerability: Insecure Infrastructure, Insecure Communication
+Asset:         Data transferred and stored in data center.
+
+2. Data Theft
+-------------
+Description: External infiltrates infrastructure in order to steal personal data and sell it on the black market.
+Also the external might be motivated politically and wants to harm the reputation of the Service Provider or the Local Authority.
+
+COI:         Privacy vs. Financial Profit
+             Reputation of Service Provider vs. Political Activism
+             Reputation of Local Authority vs. Political Activism
+Vuln.:       Insecure Infrastructure, Insecure Communication
+Asset:       Data transmitted or stored in data center.
+
+3. Information Leak
+-------------------
+Description: Like an external person in Scenario (2) an employee of the service provider has selfish interests to gain money, make political statements or harm his employer.
+In order to pursue this interest he can steal personal data and sell it or release it to the public.
+
+COI:         Privacy vs. Financial Profit
+             Reputation of Service Provider vs. Political Activism
+             Reputation of Local Authority vs. Political Activism
+Vuln.:       Unhappy Employee, Lax Access Rules.
+Asset:       Data stored in data center.
+
+4. Social Engineering
+---------------------
+Description: An External can manipulate Local Authorities or Service Providers to give them personal information that he can then sell or disclose.
+
+COI:         Privacy vs. Financial Profit
+             Reputation of Service Provider vs. Political Activism
+             Reputation of Local Authority vs. Political Activism
+Vuln.:       Unhappy Employee, Lax Access Rules.
+Asset:       Data stored in data center.
+
 \subsection{Step 2. Potential Analysis}

### 2014-10-23 Implementations.tex
Added Privacy Policy Section

\chapter{Implementation of Privacy Protection Measures}
\subsection{Privacy Policy}

In complience legal requirements of the Data Protection Directive (cf. Section \ref{EUDIR}) we include the following
Privacy Policy on our Mobile Application and Website.

\begin{quote}
\tt \small
\textbf{Privacy Policy}\\
        ==============

\textbf{What Personal Information do we Collect?}\\
The Live+Gov Service Center stores the following personal information:\\
- Name\\
- Email Address\\

The Mobile Sensor Collection Application collects the following data from your mobile phone:\\
- Sensor Data from Accelerometer (50Hz)\\
- Sensor Data from GPS sensors (~every 30 sec)\\
- Daily Human Activities (walking, standing, sitting, etc.) (every 30sec)\\
- Used Service Lines (bus, train, tram) (every 2 minutes)\\

All collected data is transferred to our data center.

\textbf{Which processing is applied to the data?}\\
We process Accelerometer samples on the Mobile device in order to extract human activities.
We process GPS coordinates on the Live+Gov server to extract the currently used service lines.

The stored information is used to produce aggregate reports.
These reports are based on anonymized data that do not allow the reconstruction of the routes of individual persons at a given point in time.

\textbf{When is the data deleted?}\\
The full data-set will be stored until January 2015.

Afterwards we will delete Names and Email addresses and anonymize
the sensor data and processing results.  The anonymized data will
be used by the the Consortium Partners for research purposes.

\textbf{For which purposes is the data collected?}\\
The collected data will be used for personalization of user experience of the mobile application.
Moreover, we will analyize travel patterns in the Helsinki Urban region for general research and for improvement and optimization of the
infrastructure provided by HSL.

\textbf{Identity of Data Controller}\\
The data is controlled by Heinrich Hartmann <hartmann@uni-koblenz.de>

The following organizations have access to the collected data:
\begin{enumerate}
\item University of Koblenz-Landau
\url{http://www.uni-koblenz-landau.de/}\\
Universitätsstraße 1\\
56070 Koblenz

\item Mattersoft
\url{http://www.mattersoft.fi/}
Hämeenkatu 13, 33100 Tampere, Finland\\
+358 10 3225000

\item Centre for Research and Technology Hellas (CERTH)
\url{http://mklab.iti.gr/contact}\\
6th km Charilaou-Thermi Road\\
P.O. Box 60361,\\
57001 Thermi-Thessaloniki, Greece
\end{enumerate}

\textbf{Access to Stored Data}\\
All personal data can be accessed at our Privacy Dashboard at
\url{http://liveandgov.uni-koblenz.de/trial/dashboard}

There you have the possibility to:\\
- View all recordings made\\
- View the raw data collected for each recording\\
- Download the data for each recording\\
- Delete records

\textbf{Collaboration with Third Parties}\\
The raw data is not shared with any parties external to the Live+Gov Consortium.

Aggregated report based on anonymized data are disclosed to officials at HSL \url{http://www.hsl.fi/} and might be made
public in the form of a blog post or research article.  These aggregated reports show distributions of all routes that
have been collected in the system and do not allow to infere the location of a single user at a given point in time.

\textbf{Consent}\\
I have read and understood the above privacy policy and agree that my data is collected and processed accordingly.

\end{verbatim}
\end{quote}

\subsection{Processing Journal}
* name of the controller
* purpose of processing
* description of the data categories
* recipients of the data if disclosed
* transfer to third countries
* description of security of processing

The journal is available for anybody to access via Email.

\subsection{Privacy Dashboard}
\subsection{Transfer Encryption (HTTPS/SSL)}
\subsection{Database Access Policies}

\subsection{Anonymization of recorded trips}
- remove precise date information, only day of week.

\subsection{Randomization of GPS data}

### 2014-10-24 PrivacyProtection-2014-09-09.tex
Added Privacy Policy Section
Revised Threats.

@@
 This section outlines the potential threats (Figure \ref{figure:Live+Gov Threats}) for Citizen privacy in the proposed monitoring and mining System.
+
 We derive threats from known conflicts of interests \ref{subsubsection:Conflicts of Interests} and vulnerabilities \ref{subsubsection:Vulnerabilities}
 
+Abstract Description of Threats.
+
 %TODO: Update Threat Table
+
+Loss Calibration:
+* High:   Leak of information to Peers, e.g. spouse gets GPS data and finds out cheating.
+          Subjects is added to No-Fly list due to Government spying.
+          Public leak of private information;
+* Medium: Undisclosed processing of personal data;
+          Potentially disclosure to third parties that are unrelated to subjct.
+* Low:    Loss of control over data.
+* None: -
 
@@
 Vulneratbility: Bad system design
-Asset:          Control over Data
-
-1. Surveillance
+Asset:          All collected data
+
+Likeliness: Always
+Loss: Low (Control is lost, no direct harm done)
+Risk:      5
+
+1. Excessive Data Mining
+------------------------
+Description: The Service Provider and/or the Local Authority extract more private information from the collected data, than the citizen agreed to.
+E.g. the Service Provider wants to test a new product, and uses old data collected for a different purpose.
+E.g. the Local Authority wants to analyze collected data from fare evasions.
+
+COI: Privacy vs. Buisness Intelligence of Local Authority, Privacy vs. Financial Profit of Serivce Provider.
+Vuln.: Lax data handling policies; Unaware citizen; Weak law enforcement
+Affected Data: Potential Information extracted from raw data.
+
+Likeliness: High
+Loss:      Medium (Control is lost, no direct harm done)
+Risk:      5
+
+2. Data Theft
+-------------
+Description: External infiltrates infrastructure in order to steal personal data and sell it on the black market.
+Also the external might be motivated politically and wants to harm the reputation of the Service Provider or the Local Authority.
+
+COI:         Privacy vs. Financial Profit
+             Reputation of Service Provider vs. Political Activism
+             Reputation of Local Authority vs. Political Activism
+Vuln.:       Insecure Infrastructure, Insecure Communication
+Affected Data:  Data transmitted or stored in data center: GPS trips, HAR-records.
+
+Likeliness: Low-Medium
+Loss: High
+Risk: 4-5
+
+3. Surveillance
 ---------------
@@
 Vulnerability: Insecure Infrastructure, Insecure Communication
-Asset:         Data transferred and stored in data center.
-
-2. Data Theft
--------------
-Description: External infiltrates infrastructure in order to steal personal data and sell it on the black market.
-Also the external might be motivated politically and wants to harm the reputation of the Service Provider or the Local Authority.
-
-COI:         Privacy vs. Financial Profit
-             Reputation of Service Provider vs. Political Activism
-             Reputation of Local Authority vs. Political Activism
-Vuln.:       Insecure Infrastructure, Insecure Communication
-Asset:       Data transmitted or stored in data center.
-
-3. Information Leak
+Affected Data:         Data transferred and stored in data center.
+
+Likeliness: Medium (Government*), Low (Third parties)
+Loss: High (Control is lost, potential harm by due to Government action)
+Risk:      6/4
+
+* Government, has also other tools at it's disposal
+
+4. Information Leak
 -------------------
@@
 Vuln.:       Unhappy Employee, Lax Access Rules.
-Asset:       Data stored in data center.
-
-4. Social Engineering
+Affected Data:       Data stored in data center.
+
+Likeliness: Low
+Loss: High (public leak)
+Risk: 4
+
+5. Social Engineering
 ---------------------
@@
 Vuln.:       Unhappy Employee, Lax Access Rules.
-Asset:       Data stored in data center.
+Affected Data:       Data stored in data center.
+
+Likeliness: Low
+Loss: High
+Risk: 4
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %\includegraphics{../diagrams/png/sensor-privacy-matrix.png}
-\input{./figures/LG-Sensor-Privacy-Matrix}
-%\input{./figures/LG-Implicit-Sensor-Privacy-Matrix}
+%\input{./figures/LG-Sensor-Privacy-Matrix}
+\input{./figures/LG-Implicit-Sensor-Privacy-Matrix}
 
@@
 
-\item
-\textbf{Privacy of Data and Image}
+\item \textbf{Privacy of Data and Image}
 
@@
 
-\item
-\textbf{Privacy of Location and Space}
+\item \textbf{Privacy of Location and Space}
 
@@
 
-\item
-\textbf{Privacy of Associtation}
+\item \textbf{Privacy of Associtation}
 

### 2014-10-24 PrivacyProtection-2014-09-09.tex
Updated Threat Intro.

@@
 
-\section{IT Security Analysis according to Grimm et. al}
+\section{IT Security Analysis according to Grimm et. al}\label{sec:GrimmModel}
 
@@
 \subsubsection{Threats}
-This section outlines the potential threats (Figure \ref{figure:Live+Gov Threats}) for Citizen privacy in the proposed monitoring and mining System.
-
-We derive threats from known conflicts of interests \ref{subsubsection:Conflicts of Interests} and vulnerabilities \ref{subsubsection:Vulnerabilities}
-
-Abstract Description of Threats.
+
+This section outlines the potential threats for the citizen privacy in the Live+Gov system.
+Recall from the security model description in section \ref{sec:GrimmModel}, that a threat is
+a potential interaction of the components that harms an asset.
+
+We restrict ourselves to the case of attacks, and the asset of privacy.
+An attack is an interaction that is executed by an actor in response to a conflict of interest by exploiting a vulnerability of the system.
+The alternative interaction of accidents are less relevant for us, since the violation of privacy always requires an actor that takes advantage of the personal data.
+
+In Figure \ref{fig:ThreatExamples} we illustrate the structure of threats at the example of ``Data Theft".
+Here the attack is executed by an external person, that harms the privacy of the citizen.
+He is motivated to do so by financial profit gained by selling personal information, which
+is in conflict with the citizens interest in his privacy.
+To get hold of the data the external exploits an insecured HTTP transfer of recorded data.
 

### 2014-10-24 PrivacyProtection-2014-09-09.tex
refactored figure threat example

--- 2014-10-24 PrivacyProtection-2014-09-09.tex
@@
 
-\subsubsection{Threats}
+\subsubsection{Threat Specification}
+
 This section outlines the potential threats (Figure \ref{figure:Live+Gov Threats}) for Citizen privacy in the proposed monitoring and mining System.
@@
 
-\subsection{Step 2. Potential Analysis}
-
-\subsubsection{Threat Specification}
-
 \textbf{Man In The Middle:}
@@
 \input{./figures/LG-Threats}
-\input{./figures/Threat-Example-1}
-\input{./figures/Threat-Example-2}
+\input{./figures/Threat-Example}
 \input{./figures/ThreatTable}

### 2014-10-24 PrivacyProtection-2014-09-09.tex
WIP

--- 2014-10-24 PrivacyProtection-2014-09-09.tex
@@
 
-This section outlines the potential threats for the citizen privacy in the Live+Gov system.
 Recall from the security model description in section \ref{sec:GrimmModel}, that a threat is
@@
 
-In Figure \ref{fig:ThreatExamples} we illustrate the structure of threats at the example of ``Data Theft".
+\input{./figures/Threat-Example-1}
+
+In Figure \ref{figure:ThreatExample} we illustrate the structure of threats at the example of ``Data Theft".
 Here the attack is executed by an external person, that harms the privacy of the citizen.
@@
 
+In this section we describe the most important threats for the citizen privacy in the Live+Gov system.
+
 %TODO: Update Threat Table
+\input{./figures/ThreatTable}
+
+0. Bad Architecture
+-------------------
+Description:    The Service Provider does not offer tools for the citizen to control his data.
+COI:            Privacy vs. System Complexity
+Vulneratbility: Bad system design
+Asset:          All collected data
+
+Likeliness: Always
+Loss: Low (Control is lost, no direct harm done)
+Risk:      5
+
+1. Excessive Data Mining
+------------------------
+Description: The Service Provider and/or the Local Authority extract more private information from the collected data, than the citizen agreed to.
+E.g. the Service Provider wants to test a new product, and uses old data collected for a different purpose.
+E.g. the Local Authority wants to analyze collected data from fare evasions.
+
+COI: Privacy vs. Buisness Intelligence of Local Authority, Privacy vs. Financial Profit of Serivce Provider.
+Vuln.: Lax data handling policies; Unaware citizen; Weak law enforcement
+Affected Data: Potential Information extracted from raw data.
+
+Likeliness: High
+Loss:      Medium (Control is lost, no direct harm done)
+Risk:      5
+
+2. Data Theft
+-------------
+Description: External infiltrates infrastructure in order to steal personal data and sell it on the black market.
+Also the external might be motivated politically and wants to harm the reputation of the Service Provider or the Local Authority.
+
+COI:         Privacy vs. Financial Profit
+             Reputation of Service Provider vs. Political Activism
+             Reputation of Local Authority vs. Political Activism
+Vuln.:       Insecure Infrastructure, Insecure Communication
+Affected Data:  Data transmitted or stored in data center: GPS trips, HAR-records.
+
+Likeliness: Low-Medium
+Loss: High
+Risk: 4-5
+
+3. Surveillance
+---------------
+Description:   External infiltrates infrastructure in order to spy on citizen
+COI:           Privacy vs. External Power
+Vulnerability: Insecure Infrastructure, Insecure Communication
+Affected Data:         Data transferred and stored in data center.
+
+Likeliness: Medium (Government*), Low (Third parties)
+Loss: High (Control is lost, potential harm by due to Government action)
+Risk:      6/4
+
+* Government, has also other tools at it's disposal
+
+4. Information Leak
+-------------------
+Description: Like an external person in Scenario (2) an employee of the service provider has selfish interests to gain money, make political statements or harm his employer.
+In order to pursue this interest he can steal personal data and sell it or release it to the public.
+
+COI:         Privacy vs. Financial Profit
+             Reputation of Service Provider vs. Political Activism
+             Reputation of Local Authority vs. Political Activism
+Vuln.:       Unhappy Employee, Lax Access Rules.
+Affected Data:       Data stored in data center.
+
+Likeliness: Low
+Loss: High (public leak)
+Risk: 4
+
+5. Social Engineering
+---------------------
+Description: An External can manipulate Local Authorities or Service Providers to give them personal information that he can then sell or disclose.
+
+COI:         Privacy vs. Financial Profit
+             Reputation of Service Provider vs. Political Activism
+             Reputation of Local Authority vs. Political Activism
+Vuln.:       Unhappy Employee, Lax Access Rules.
+Affected Data:       Data stored in data center.
+
+Likeliness: Low
+Loss: High
+Risk: 4
+
+\subsubsection{Threat Risk Evaluation}
 
@@
 
-0. Bad Architecture
--------------------
-Description:    The Service Provider does not offer tools for the citizen to control his data.
-COI:            Privacy vs. System Complexity
-Vulneratbility: Bad system design
-Asset:          All collected data
-
-Likeliness: Always
-Loss: Low (Control is lost, no direct harm done)
-Risk:      5
-
-1. Excessive Data Mining
-------------------------
-Description: The Service Provider and/or the Local Authority extract more private information from the collected data, than the citizen agreed to.
-E.g. the Service Provider wants to test a new product, and uses old data collected for a different purpose.
-E.g. the Local Authority wants to analyze collected data from fare evasions.
-
-COI: Privacy vs. Buisness Intelligence of Local Authority, Privacy vs. Financial Profit of Serivce Provider.
-Vuln.: Lax data handling policies; Unaware citizen; Weak law enforcement
-Affected Data: Potential Information extracted from raw data.
-
-Likeliness: High
-Loss:      Medium (Control is lost, no direct harm done)
-Risk:      5
-
-2. Data Theft
--------------
-Description: External infiltrates infrastructure in order to steal personal data and sell it on the black market.
-Also the external might be motivated politically and wants to harm the reputation of the Service Provider or the Local Authority.
-
-COI:         Privacy vs. Financial Profit
-             Reputation of Service Provider vs. Political Activism
-             Reputation of Local Authority vs. Political Activism
-Vuln.:       Insecure Infrastructure, Insecure Communication
-Affected Data:  Data transmitted or stored in data center: GPS trips, HAR-records.
-
-Likeliness: Low-Medium
-Loss: High
-Risk: 4-5
-
-3. Surveillance
----------------
-Description:   External infiltrates infrastructure in order to spy on citizen
-COI:           Privacy vs. External Power
-Vulnerability: Insecure Infrastructure, Insecure Communication
-Affected Data:         Data transferred and stored in data center.
-
-Likeliness: Medium (Government*), Low (Third parties)
-Loss: High (Control is lost, potential harm by due to Government action)
-Risk:      6/4
-
-* Government, has also other tools at it's disposal
-
-4. Information Leak
--------------------
-Description: Like an external person in Scenario (2) an employee of the service provider has selfish interests to gain money, make political statements or harm his employer.
-In order to pursue this interest he can steal personal data and sell it or release it to the public.
-
-COI:         Privacy vs. Financial Profit
-             Reputation of Service Provider vs. Political Activism
-             Reputation of Local Authority vs. Political Activism
-Vuln.:       Unhappy Employee, Lax Access Rules.
-Affected Data:       Data stored in data center.
-
-Likeliness: Low
-Loss: High (public leak)
-Risk: 4
-
-5. Social Engineering
----------------------
-Description: An External can manipulate Local Authorities or Service Providers to give them personal information that he can then sell or disclose.
-
-COI:         Privacy vs. Financial Profit
-             Reputation of Service Provider vs. Political Activism
-             Reputation of Local Authority vs. Political Activism
-Vuln.:       Unhappy Employee, Lax Access Rules.
-Affected Data:       Data stored in data center.
-
-Likeliness: Low
-Loss: High
-Risk: 4
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Step 2. Potential Analysis}
-
-\subsubsection{Threat Specification}
-
-\textbf{Man In The Middle:}
-Lorem ipsum dolor suit amet, ...
-
-\textbf{Exploit:}
-Lorem ipsum dolor suit amet, ...
-
-\textbf{Corruption:}
-Lorem ipsum dolor suit amet, ...
-
-\textbf{Intrasparency:}
-Lorem ipsum dolor suit amet, ...
-
-\input{./figures/LG-Threats}
-\input{./figures/Threat-Example-1}
-\input{./figures/Threat-Example-2}
-\input{./figures/ThreatTable}
-
-\subsubsection{Threat Risk Evaluation}
-
-% Todo: use sensor privacy matrix and implicit sensor privacy matrix for threat risk evaluation
-
 \textbf{Sensor Privacy Matrix}
-
-%\includegraphics{../diagrams/png/sensor-privacy-matrix.png}
-%\input{./figures/LG-Sensor-Privacy-Matrix}
 \input{./figures/LG-Implicit-Sensor-Privacy-Matrix}
@@
 \begin{itemize}
-
-\item
-\textbf{Privacy of The Person}
+\item \textbf{Privacy of The Person}
 
@@
 \item \textbf{Privacy of Data and Image}
-
-The \textbf{Privacy of Data and Image} demands, that \emph{``individual's data is not automatically available to other individuals''}.
+The Privacy of Data and Image demands, that \emph{``individual's data is not automatically available to other individuals''}.
 This type of privacy is trivially threatened because here sensor data is individual data, a priori.
@@
 \item \textbf{Privacy of Location and Space}
-
-According to Friedewald et al., the \textbf{Privacy of Location and Space} is concerned with one's \emph{``right to move about in public or semi-public space without being identified, tracked or monitored.''}.
+According to Friedewald et al., the Privacy of Location and Space is concerned with one's \emph{``right to move about in public or semi-public space without being identified, tracked or monitored.''}.
 This is the location aspect of this type. The space aspect is concerned with one's \emph{``right to solitude''}, which generally includes one's right to an inviolate home and other private spaces.

### 2014-10-24 PrivacyProtection-2014-09-09.tex
Updated Threat Intro. Fixed Figure References.

@@
 
-\subsubsection{Threats}
-
-Recall from the security model description in section \ref{sec:GrimmModel}, that a threat is
-a potential interaction of the components that harms an asset.
-
+\subsubsection{Threat Specification}
+
+\input{./figures/Threat-Example}
+
+Recall from the security model description in Section \ref{sec:GrimmModel}, that a threat is a potential interaction of the components that harms an asset.
 We restrict ourselves to the case of attacks, and the asset of privacy.
@@
 
-\input{./figures/Threat-Example-1}
-
-In Figure \ref{figure:ThreatExample} we illustrate the structure of threats at the example of ``Data Theft".
+In Figure \ref{figure:Threat-Example} we illustrate the structure of threats at the example of ``Data Theft".
 Here the attack is executed by an external person, that harms the privacy of the citizen.
@@
 
-In this section we describe the most important threats for the citizen privacy in the Live+Gov system.
+In this section we describe, in a similar fashin, threats for the citizen privacy in the Live+Gov system.
+This list can necessarily not be complete, but we make a best effort to cover the most relevant cases.
 

### 2014-10-24 PrivacyProtection-2014-09-09.tex
added new threat-table figure

--- 2014-10-24 PrivacyProtection-2014-09-09.tex
@@
 
-We derive threats from known conflicts of interests \ref{subsubsection:Conflicts of Interests} and vulnerabilities \ref{subsubsection:Vulnerabilities}
-
-Abstract Description of Threats.
+This section outlines the potential threats for the citizen privacy in the Live+Gov system.
+Recall from the security model description in section \ref{sec:GrimmModel}, that a threat is
+a potential interaction of the components that harms an asset.
+
+We restrict ourselves to the case of attacks, and the asset of privacy.
+An attack is an interaction that is executed by an actor in response to a conflict of interest by exploiting a vulnerability of the system.
+The alternative interaction of accidents are less relevant for us, since the violation of privacy always requires an actor that takes advantage of the personal data.
+
+In Figure \ref{fig:ThreatExamples} we illustrate the structure of threats at the example of ``Data Theft".
+Here the attack is executed by an external person, that harms the privacy of the citizen.
+He is motivated to do so by financial profit gained by selling personal information, which
+is in conflict with the citizens interest in his privacy.
+To get hold of the data the external exploits an insecured HTTP transfer of recorded data.
 
@@
 \input{./figures/Threat-Example}
-\input{./figures/ThreatTable}
+\input{./figures/Threats}
 

### 2014-10-24 PrivacyProtection-2014-09-09.tex
Revised Threat Risk Evaluation Section.

--- 2014-10-24 PrivacyProtection-2014-09-09.tex
@@
 
+\subsection{Sensor Data Privacy Impact}\label{sec:SensorPrivacyImpact}
+\input{./figures/LG-Implicit-Sensor-Privacy-Matrix}
+
+In this section we analyze the impact of the disclosure sensor data and certain processing results to the citizens privacy.
+This analyisis builds upon the preceeding analyis, but is more focused on the concrete type of data vailable.
+For example, the disclosure of Service Line Detection results, does violate the Privacy of Location and Space but not the Privacy of Association that is violated by general GPS tracking.
+
+%TODO: Reformulate the section.
+% - Too much overlap with preceeding sections.
+% - Structure should be per Sensor Type
+% - Add HAR and SLD!
+
+The \textbf{Privacy of The Person} is generally concerned with one could best understand as \emph{Biometric Privacy}.
+Friedewald et al. paraphrase it as \emph{``\om the right to keep body functions and body characteristics \om private''}.
+Accelerometer, Rotation Vector and Gyroscope measure the physical movement of the mobile device on all three axes.
+If the mobile device is carried ``normally'' its safe to say that those sensors also measure the moments of its carrier.
+So his privacy is infringed regarding biometric behaviour, as it is captured automatically.
+
+\textbf{Privacy of Data and Image} demands, that \emph{``individual's data is not automatically available to other individuals''}.
+This type of privacy is trivially threatened because here sensor data is individual data, a priori.
+So every sensor violates the privacy of data and image, as data is transported into a foreign system where operators have access to it.
+
+According to Friedewald et al., the \textbf{Privacy of Location and Space} is concerned with one's \emph{``right to move about in public or semi-public space without being identified, tracked or monitored.''}.
+This is the location aspect of this type. The space aspect is concerned with one's \emph{``right to solitude''}, which generally includes one's right to an inviolate home and other private spaces.
+Obviously the GPS and GSM sensors violate such right about not being tracked, because they reveal the position of the mobile device and its carrier.
+The GSM sensor gives the exact cell, the mobile device has registered with at the current moment.
+The GPS sensor gives the current longitude and latitude, the current global position of the mobile device and its carrier, although there is some artificial inaccuracy within civil use.
+
+The WLAN and Bluetooth sensors record lists of the currently available local wireless networks and bluetooth clients.
+If such are known stationary entities, those sensors are considered as dangerous as the GSM sensor for the carriers locational privacy.
+
+The Magnetic field sensor is not regarded very dangerous to the carrier's privacy, because it does not allow very precise localization.
+But it can limit the possibilities for the global position of the mobile device.
+Here, it is just named for completeness sake.
+
+The \textbf{Privacy of Association} sates that everyone has the \emph{``right to associate with whomever they wish, without being monitored [automatically without reasonable suspicion]''}.
+This includes individuals and organizations.
+WLAN and Bluetooth sensors provide the ability to monitor such associations if their lists contain known entities.
+If one frequently connects with an organizational wireless network, e.g.~an university network, an association can be deduced (student or staff).
+The same goes for the Bluetooth sensor, if it is stationary.
+Additionally, if the recorded bluetooth clients are mobile, it is more or less possible to deduce association with the technical identity of (yet) anonymous individuals.
+
+Additionally, the GSM sensor could provide the association with the GSM operator.
+%TODO: NEEDS TO BE VERIFIED!
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Here the attack is executed by an external person, that harms the privacy of the citizen.
-He is motivated to do so by financial profit gained by selling personal information, which
-is in conflict with the citizens interest in his privacy.
+He is motivated to do so by financial profit gained by selling personal information, which is in conflict with the citizens interest in his privacy.
 To get hold of the data the external exploits an insecured HTTP transfer of recorded data.
@@
 
-Loss Calibration:
-* High:   Leak of information to Peers, e.g. spouse gets GPS data and finds out cheating.
+In this section we will associate to every identified threat a corresponding risk.
+Recall from \ref{sec:GrimmModel} that a risk is the expected loss that is associated to the threat.
+Therefore, we have to quantify the likeliness of the threat to occure and the harm or loss done in this case.
+The quantification of likeliness will be solely based on rough judgment of the authors.
+The quantification of loss, will be made in a two step process.
+For each threat listed in table \ref{firgure:Threats}, we have analyzed the affected personal data of the citizen.
+For each possible data type (e.g. GPS) we analyze the impact on the seven different types of privacy in Section \ref{sec:SensorPrivacyImpact}.
+In combination we can quantify roughly the impact of each threat on the citizens privacy.
+
+Both evaluations are necessarily fraught with a high level of uncertainty.
+
+Loss Calibration
+\begin{itemize}
+\item 3: High. Leak of information to Peers which impacts citizen.
+  E.g. spouse gets GPS data and finds out cheating.
   Subjects is added to No-Fly list due to Government spying.
-          Public leak of private information;
-* Medium: Undisclosed processing of personal data;
-          Potentially disclosure to third parties that are unrelated to subjct.
-* Low:    Loss of control over data.
-* None: -
-
-\textbf{Sensor Privacy Matrix}
-\input{./figures/LG-Implicit-Sensor-Privacy-Matrix}
-
-\begin{itemize}
-\item \textbf{Privacy of The Person}
-
-The \textbf{Privacy of The Person} is generally concerned with one could best understand as \emph{Biometric Privacy}.
-Friedewald et al. paraphrase it as \emph{``\om the right to keep body functions and body characteristics \om private''}.
-\textbf{Accelerometer}, \textbf{Rotation Vector} and \textbf{Gyroscope} measure the physical movement of the mobile device on all three axes.
-If the mobile device is carried \emph{``normally''} its safe to say that those sensors also measure the moments of its carrier.
-So his privacy is infringed regarding biometric behaviour, as it is captured automatically.
-(\textbf{Note:} This is not to be confused with the \textbf{Privacy of Behaviour and Action} which is used for the social aspects of behaviour action, e.g.~praying, sexual habits or political activities.)
-
-\item \textbf{Privacy of Data and Image}
-The Privacy of Data and Image demands, that \emph{``individual's data is not automatically available to other individuals''}.
-This type of privacy is trivially threatened because here sensor data is individual data, a priori.
-So every sensor violates the privacy of data and image, as data is transported into a foreign system where operators have access to it.
-
-\item \textbf{Privacy of Location and Space}
-According to Friedewald et al., the Privacy of Location and Space is concerned with one's \emph{``right to move about in public or semi-public space without being identified, tracked or monitored.''}.
-This is the location aspect of this type. The space aspect is concerned with one's \emph{``right to solitude''}, which generally includes one's right to an inviolate home and other private spaces.
-Obviously the \textbf{GPS} and \textbf{GSM} sensors violate such right about not being tracked, because they reveal the position of the mobile device and its carrier.
-The \textbf{GSM} sensor gives the exact cell, the mobile device has registered with at the current moment.
-The \textbf{GPS} sensor gives the current longitude and latitude, the current global position of the mobile device and its carrier, although there is some artificial inaccuracy within civil use.
-
-The \textbf{WLAN} and \textbf{Bluetooth} sensors record lists of the currently available local wireless networks and bluetooth clients.
-If such are known stationary entities, those sensors are considered as dangerous as the \textbf{GSM} sensor for the carriers locational privacy.
-
-The \textbf{Magnetic field} sensor is not regarded very dangerous to the carrier's privacy, because it does not allow very precise localization.
-But it can limit the possibilities for the global position of the mobile device.
-Here, it is just named for completeness sake.
-
-\item \textbf{Privacy of Associtation}
-
-The \textbf{Privacy of Association} sates that everyone has the \emph{``right to associate with whomever they wish, without being monitored [automatically without reasonable suspicion]''}.
-This includes individuals and organizations.
-\textbf{WLAN} and \textbf{Bluetooth} sensors provide the ability to monitor such associations if their lists contain known entities.
-If one frequently connects with an organizational wireless network, e.g.~an university network, an association can be deduced (student or staff).
-The same goes for the \textbf{Bluetooth} sensor, if it is stationary.
-Additionally, if the recorded bluetooth clients are mobile, it is more or less possible to deduce association with the technical identity of (yet) anonymous individuals.
-
-Additionally, the \textbf{GSM} sensor could provide the association with the GSM operator (\textbf{NEEDS TO BE VERIFIED!}).
-
-\end{itemize}
+  Public leak of private information.
+\item 2: Medium. Undisclosed processing of personal data or disclosure to third parties that are unrelated to subject.
+\item 1: Low. Loss of control over data.
+\item 0: None
+\end{itemize}
+
+Likeliness is measured on a the following scale
+\begin{itemize}
+\item 0 (impossible)
+\item 1: maybe occure once in 1 million cases.
+\item 2: maybe occure once in 100
+\item 3: maybe occure once in 10
+\item 4: always the case.
+\end{itemize}
+The risk scale is defined as addition of the loss and likeliness scales.\footnote{
+Note, that loss and likeliness scales have a logarithmic character, so that that addition of those scales corresponds to multiplication of the usual scales.
+}
+
+%TODO:
+Insert Table with Likeliness / Loss / Risk Assesment.
 
 \subsubsection{Security Requirement Specification}
+%TODO SKIP?
 

### 2014-10-24 PrivacyProtection-2014-09-09.tex
finished spec of insufficient control features

--- 2014-10-24 PrivacyProtection-2014-09-09.tex
@@
 
+% Extend to general missing privacy wareness of citizen, system provider and local authority
 \textbf{Unaware Monitoring Subjects.}
@@
 
-Recall from the security model description in Section \ref{sec:GrimmModel}, that a threat is a potential interaction of the components that harms an asset.
+Recall from the security model description in Section \ref{sec:GrimmModel}, that a threat is a potential interaction of the components that targets an asset.
 We restrict ourselves to the case of attacks, and the asset of privacy.
 An attack is an interaction that is executed by an actor in response to a conflict of interest by exploiting a vulnerability of the system.
-The alternative interaction of accidents are less relevant for us, since the violation of privacy always requires an actor that takes advantage of the personal data.
+The alternative interaction of accidents are less relevant for us, since the violation of privacy always requires an actor that takes advantage of personal data.
 
@@
 
-In this section we describe, in a similar fashin, threats for the citizen privacy in the Live+Gov system.
+In this section we describe, in a similar fashion, threats for the citizen privacy in the Live+Gov system.
 This list can necessarily not be complete, but we make a best effort to cover the most relevant cases.
@@
 %TODO: Update Threat Table
-\input{./figures/ThreatTable}
-
-0. Bad Architecture
--------------------
-Description:    The Service Provider does not offer tools for the citizen to control his data.
-COI:            Privacy vs. System Complexity
-Vulneratbility: Bad system design
-Asset:          All collected data
-
-Likeliness: Always
-Loss: Low (Control is lost, no direct harm done)
-Risk:      5
-
-1. Excessive Data Mining
-------------------------
-Description: The Service Provider and/or the Local Authority extract more private information from the collected data, than the citizen agreed to.
-E.g. the Service Provider wants to test a new product, and uses old data collected for a different purpose.
-E.g. the Local Authority wants to analyze collected data from fare evasions.
-
-COI: Privacy vs. Buisness Intelligence of Local Authority, Privacy vs. Financial Profit of Serivce Provider.
-Vuln.: Lax data handling policies; Unaware citizen; Weak law enforcement
-Affected Data: Potential Information extracted from raw data.
-
-Likeliness: High
-Loss:      Medium (Control is lost, no direct harm done)
-Risk:      5
-
-2. Data Theft
--------------
-Description: External infiltrates infrastructure in order to steal personal data and sell it on the black market.
+\input{./figures/Threats}
+
+%0. Insufficient Control Features
+%-------------------
+%Description:    The Service Provider does not offer tools for the citizen to control his data.
+%COI:            Privacy vs. System Complexity
+%Vulneratbility: missing privacy wareness
+%Asset:          All collected data
+%
+%Likeliness: Always
+%Loss: Low (Control is lost, no direct harm done)
+%Risk:      5
+%
+\textbf{Insufficient Control Features.}
+The System Provider does not offer tools for the Citizen to control his data.
+However, besides the missing features he provides a safe and secure system.
+This threat contradicts the definition of privacy as control over one's information about oneself.
+As soon as collected data of Citizens is stored on System Provider servers, all control over that data is lost.
+This is not necessarily due to bad intention, System Providers may simply have forgotten to include such features during the development process.
+Although, control capabilities for Citizens add to the system complexity, which could motivate to omit those. 
+Therefore interests of Citizens and System Providers are in conflict, namely it is the Citizen's \textit{Privacy vs. System Complexity} for System Providers.
+This is an abstract threat provoked by a general \textit{Missing Privacy Awareness} in the minds of all actors in the Live+Gov context.
+
+%1. Excessive Data Mining
+%------------------------
+%Description: The Service Provider and/or the Local Authority extract more private information from the collected data, than the citizen agreed to.
+%E.g. the Service Provider wants to test a new product, and uses old data collected for a different purpose.
+%E.g. the Local Authority wants to analyze collected data from fare evasions.
+%
+%COI: Privacy vs. Buisness Intelligence of Local Authority, Privacy vs. Financial Profit of Serivce Provider.
+%Vuln.: Lax data handling policies; Unaware citizen; Weak law enforcement
+%Affected Data: Potential Information extracted from raw data.
+%
+%Likeliness: High
+%Loss:      Medium (Control is lost, no direct harm done)
+%Risk:      5
+%
+\textbf{Excessive Data Mining}
+Lorem Ipsum
+
+%2. Data Theft
+%-------------
+%Description: External infiltrates infrastructure in order to steal personal data and sell it on the black market.
 Also the external might be motivated politically and wants to harm the reputation of the Service Provider or the Local Authority.
-
-COI:         Privacy vs. Financial Profit
-             Reputation of Service Provider vs. Political Activism
-             Reputation of Local Authority vs. Political Activism
-Vuln.:       Insecure Infrastructure, Insecure Communication
-Affected Data:  Data transmitted or stored in data center: GPS trips, HAR-records.
-
-Likeliness: Low-Medium
-Loss: High
-Risk: 4-5
-
-3. Surveillance
----------------
-Description:   External infiltrates infrastructure in order to spy on citizen
-COI:           Privacy vs. External Power
-Vulnerability: Insecure Infrastructure, Insecure Communication
-Affected Data:         Data transferred and stored in data center.
-
-Likeliness: Medium (Government*), Low (Third parties)
-Loss: High (Control is lost, potential harm by due to Government action)
-Risk:      6/4
-
-* Government, has also other tools at it's disposal
-
-4. Information Leak
--------------------
-Description: Like an external person in Scenario (2) an employee of the service provider has selfish interests to gain money, make political statements or harm his employer.
-In order to pursue this interest he can steal personal data and sell it or release it to the public.
-
-COI:         Privacy vs. Financial Profit
-             Reputation of Service Provider vs. Political Activism
-             Reputation of Local Authority vs. Political Activism
-Vuln.:       Unhappy Employee, Lax Access Rules.
-Affected Data:       Data stored in data center.
-
-Likeliness: Low
-Loss: High (public leak)
-Risk: 4
-
-5. Social Engineering
----------------------
-Description: An External can manipulate Local Authorities or Service Providers to give them personal information that he can then sell or disclose.
-
-COI:         Privacy vs. Financial Profit
-             Reputation of Service Provider vs. Political Activism
-             Reputation of Local Authority vs. Political Activism
-Vuln.:       Unhappy Employee, Lax Access Rules.
-Affected Data:       Data stored in data center.
-
-Likeliness: Low
-Loss: High
-Risk: 4
+%
+%COI:         Privacy vs. Financial Profit
+%             Reputation of Service Provider vs. Political Activism
+%             Reputation of Local Authority vs. Political Activism
+%Vuln.:       Insecure Infrastructure, Insecure Communication
+%Affected Data:  Data transmitted or stored in data center: GPS trips, HAR-records.
+%
+%Likeliness: Low-Medium
+%Loss: High
+%Risk: 4-5
+%
+\textbf{Data Theft}
+Lorem Ipsum
+
+%3. Surveillance
+%---------------
+%Description:   External infiltrates infrastructure in order to spy on citizen
+%COI:           Privacy vs. External Power
+%Vulnerability: Insecure Infrastructure, Insecure Communication
+%Affected Data:         Data transferred and stored in data center.
+%
+%Likeliness: Medium (Government*), Low (Third parties)
+%Loss: High (Control is lost, potential harm by due to Government action)
+%Risk:      6/4
+%
+%* Government, has also other tools at it's disposal
+%
+%
+\textbf{Surveillance}
+Lorem Ipsum
+
+%4. Information Leak
+%-------------------
+%Description: Like an external person in Scenario (2) an employee of the service provider has selfish interests to gain money, make political statements or harm his employer.
+%In order to pursue this interest he can steal personal data and sell it or release it to the public.
+%
+%COI:         Privacy vs. Financial Profit
+%             Reputation of Service Provider vs. Political Activism
+%             Reputation of Local Authority vs. Political Activism
+%Vuln.:       Unhappy Employee, Lax Access Rules.
+%Affected Data:       Data stored in data center.
+%
+%Likeliness: Low
+%Loss: High (public leak)
+%Risk: 4
+%
+\textbf{Information Leak}
+Lorem Ipsum
+
+%5. Social Engineering
+%---------------------
+%Description: An External can manipulate Local Authorities or Service Providers to give them personal information that he can then sell or disclose.
+%
+%COI:         Privacy vs. Financial Profit
+%             Reputation of Service Provider vs. Political Activism
+%             Reputation of Local Authority vs. Political Activism
+%Vuln.:       Unhappy Employee, Lax Access Rules.
+%Affected Data:       Data stored in data center.
+%
+%Likeliness: Low
+%Loss: High
+%Risk: 4
+%
+\textbf{Social Engineering}
+Lorem Ipsum
 

### 2014-10-24 Implementations.tex
Added Screenshots
Revised Threats.
Added short description of Implementation.

@@
 I have read and understood the above privacy policy and agree that my data is collected and processed accordingly.
-
-\end{verbatim}
 \end{quote}
@@
 
+Reporting to Data Protection Authority (Bundesamt für Datenschutz)?
+
 \subsection{Privacy Dashboard}
+TODO: Add description of dashboard. With relation to 7C's of Privacy.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{screenshots/login.png}
+\caption{Live+Gov Privacy Dashboard Login}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{screenshots/recordings.png}
+\caption{Live+Gov Privacy Dashboard Recording Overview}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{screenshots/raw.png}
+\caption{Live+Gov Privacy Dashboard Raw Data View}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{screenshots/processed.png}
+\caption{Live+Gov Privacy Dashboard Processed Data View}
+\end{figure}
+
 \subsection{Transfer Encryption (HTTPS/SSL)}
+\begin{figure}
+\includegraphics[width=\textwidth]{screenshots/HTTPS.png}
+\caption{HTTPS connection to Privacy Dashboard}
+\end{figure}
+
 \subsection{Database Access Policies}
+Access rules for data controller and processor.
+Data Processing Journal.
+Python script for deletion of data when it is older than a given date.
 
-\subsection{Anonymization of recorded trips}
-- remove precise date information, only day of week.
+\subsection{Anonymization of trips}
+Privacy by Design: We do never collect name and email address in our DB.
+Python script that deletes user information from trip.
+% Python script that reomves precise date information, only day of week.
+
+\subsection{K-Anonymity of GPS routes}
+Python script that deletes all parts of journeys that have are not near to at least k other journeys.
+
+TODO: Insert Literature Reference.
+TODO: Add screenshots.
 
 \subsection{Randomization of GPS data}
+Python Script that adds random noise to GPS data.

### 2014-10-24 PrivacyProtection-2014-09-09.tex
Added short description of Implementation.
finsihed spec of excessive data mining

@@
 %
-\textbf{Excessive Data Mining}
-Lorem Ipsum
+\textbf{Excessive Data Mining.}
+The System Provider and/or the Local Authority secretly extract more private information from the collected data, than the Citizen agreed to.
+But results of the mining process create no disadvantages for Citizens, because there is no disclosure to third parties.
+This could be the case for a System Provider, who wants to test a new product and uses the pre-existing data collection.
+Or for a Local Authority, who wants to analyze the data collection regarding fare evasion.
+However, the Citizen has not agreed to such data processing nor could he, since it is conducted secretly.
+This disables a Citizen to control his data adequately.
+Thus there are two possible conflicts: \textit{Privacy vs. Financial Profit of Serivce Provider} and  \textit{Privacy vs. Buisness Intelligence of Local Authority}
+The threat can be provoked by either lax data handling policies of both System Providers and Local Authorities, or a weak law enforcement of existing supervision.
+But the main issue, which can lead to such threats, is again a general \textit{Missing Privacy Awareness}.
 

### 2014-10-24 PrivacyProtection-2014-09-09.tex
finished spec data theft

@@
 
-We follow the Reference Model for IT Security Analyis as described in \cite{Grimm:ItSecRefModel}.
+We follow the Reference Model for IT Security Analysis as described in \cite{Grimm:ItSecRefModel}.
 It supersedes earlier efforts by \cite{Avizienis}.
@@
 
-The model is depictured in Figure \ref{fig:refmodel}.
+The model is depicted in Figure \ref{fig:refmodel}.
 It is organized in four views (round boxes) that contain a number of components (rectangular boxes).
@@
 \textbf{Data Theft}
-Lorem Ipsum
+An External infiltrates infrastructure in order to steal personal data and sell it on the black market. 
+Also the External might be motivated politically and wants to harm the reputation of the System Provider or the Local Authority.
+Anyways, this threat would be manifested through a technical attack on either hardware (Packet Capture) or software (buffer overflow, SQL injection).
+Such a successful attack could harm the reputation of both System Provider and Local Authority.
+The technical competence of System Providers would be proven wrong, therefore they would lose professional reputation.
+Local Authorities would lose their political reputation, as it would seem like they had endangered data of Citizens, which lose complete control.
+Thus this threat is defined by three conflicts: \textit{Privacy vs.
+Financial Profit}, \textit{ Reputation of Service Provider vs.
+Political Activism} and \textit{Reputation of Local Authority vs.
+Political Activism}.
+Also this threat describes the classical scenario, where attacks are provoked by \textit{Insecure Infrastructure} (SQL injection) and \textit{Insecure Communication} (Packet Capture).
 
@@
 choices.
-
 \end{itemize}

### 2014-10-24 Implementations.tex
WIP

@@
 
-\subsection{Processing Journal}
+\section{Processing Journal}
 * name of the controller
@@
 
-\subsection{Privacy Dashboard}
+\section{Privacy Dashboard}
 TODO: Add description of dashboard. With relation to 7C's of Privacy.
@@
 
-\subsection{Transfer Encryption (HTTPS/SSL)}
+\section{Transfer Encryption (HTTPS/SSL)}
 \begin{figure}
@@
 
-\subsection{Database Access Policies}
+\section{Database Access Policies}
 Access rules for data controller and processor.
@@
 
-\subsection{Anonymization of trips}
+\section{Anonymization of trips}
 Privacy by Design: We do never collect name and email address in our DB.
@@
 
-\subsection{K-Anonymity of GPS routes}
+\section{K-Anonymity of GPS routes}
 Python script that deletes all parts of journeys that have are not near to at least k other journeys.
@@
 
-\subsection{Randomization of GPS data}
+\section{Randomization of GPS data}
 Python Script that adds random noise to GPS data.

### 2014-10-24 PrivacyProtection-2014-09-09.tex
WIP

--- 2014-10-24 PrivacyProtection-2014-09-09.tex
@@
 
-In this section we analyze the impact of the disclosure sensor data and certain processing results to the citizens privacy.
-This analyisis builds upon the preceeding analyis, but is more focused on the concrete type of data vailable.
-For example, the disclosure of Service Line Detection results, does violate the Privacy of Location and Space but not the Privacy of Association that is violated by general GPS tracking.
-
 %TODO: Reformulate the section.
@@
 
+In this section we analyze the impact of the disclosure sensor data and certain processing results to the citizens privacy.
+This analyisis builds upon the preceeding analyis, but is more focused on the concrete type of data vailable.
+For example, the disclosure of Service Line Detection results, does violate the Privacy of Location and Space but not the Privacy of Association that is violated by general GPS tracking.
+
+\textbf{GPS Data}.
+The collection of GPS data does directly infringe the citizens privacy of Location and Space, and the privacy of Data an Image.
+By inference we also get implicit violations of the Privacy of the Person, Privacy of Behavior and Action and Privacy of Association.
+
+\textbf{Motion Sensors.}
+Accelerometer, Rotation Vector, Gyroscope and Magnetic field sensor measure the physical movement of the mobile device on all three axes.
+If the mobile device is carried ``normally'' its safe to say that those sensors also measure the moments of its carrier.
+So his privacy is infringed regarding biometric behaviour, as it is captured automatically
+Privacy of Data and Image is trivially threatened because here sensor data is individual data, a priori.
+
+By inference we also get implicit violations of the Privacy of Behavior and Action.
+
+\textbf{Network Sensors}.
+The GSM and WLAN sensors reveal the position of the mobile device and its carrier, when used in connection with external databases.
+The GSM sensor gives the exact cell, the mobile device has registered with at the current moment.
+The GPS sensor gives the current longitude and latitude, the current global position of the mobile device and its carrier, although there is some artificial inaccuracy within civil use.
+
+The WLAN and Bluetooth sensors record lists of the currently available local wireless networks and bluetooth clients.
+If such are known stationary entities, those sensors are considered as dangerous as the GSM sensor for the carriers locational privacy.
+
 The \textbf{Privacy of The Person} is generally concerned with one could best understand as \emph{Biometric Privacy}.
 Friedewald et al. paraphrase it as \emph{``\om the right to keep body functions and body characteristics \om private''}.
-Accelerometer, Rotation Vector and Gyroscope measure the physical movement of the mobile device on all three axes.
-If the mobile device is carried ``normally'' its safe to say that those sensors also measure the moments of its carrier.
-So his privacy is infringed regarding biometric behaviour, as it is captured automatically.
 
 \textbf{Privacy of Data and Image} demands, that \emph{``individual's data is not automatically available to other individuals''}.
-This type of privacy is trivially threatened because here sensor data is individual data, a priori.
 So every sensor violates the privacy of data and image, as data is transported into a foreign system where operators have access to it.
@@
 This is the location aspect of this type. The space aspect is concerned with one's \emph{``right to solitude''}, which generally includes one's right to an inviolate home and other private spaces.
-Obviously the GPS and GSM sensors violate such right about not being tracked, because they reveal the position of the mobile device and its carrier.
-The GSM sensor gives the exact cell, the mobile device has registered with at the current moment.
-The GPS sensor gives the current longitude and latitude, the current global position of the mobile device and its carrier, although there is some artificial inaccuracy within civil use.
-
-The WLAN and Bluetooth sensors record lists of the currently available local wireless networks and bluetooth clients.
-If such are known stationary entities, those sensors are considered as dangerous as the GSM sensor for the carriers locational privacy.
-
-The Magnetic field sensor is not regarded very dangerous to the carrier's privacy, because it does not allow very precise localization.
-But it can limit the possibilities for the global position of the mobile device.
-Here, it is just named for completeness sake.
 
@@
 
-\subsubsection{Security Requirement Specification}
+%\subsubsection{Security Requirement Specification}
 %TODO SKIP?
@@
 
-\subsection{Legal Privacy Requirements} \ref{sec:LEGAL_REQ}
+% \subsection{Legal Privacy Requirements} \ref{sec:LEGAL_REQ}
 
@@
 
-\textbf{The 7 C's of user privacy control (Figure \ref{figure:The Seven Cs of User Privacy Control})}
-
-This is a note on an excerpt from the article \emph{Sociotechnical
-Architecture for Online Privacy} \cite{1} called
-\textbf{The 7 C's of user privacy control}. Those 7 C's are aspects
-which should be covered by measures for implementing user privacy. They
-derive from an interpretation of privacy which could be summarized as
-\emph{``One's ability to control/seclude information about oneself''}.
-
-%\includegraphics{../diagrams/png/The7Cs.png}
-\input{./figures/The-7-Cs-Of-User-Privacy-Control}
-
-\begin{itemize}
-
-\item
-\textbf{Comprehension}
-
-\emph{``Users should \textbf{understand} ho personal identifiable
-information (PII) is handled, who's collecting it and for what purpose,
-and who will process the PII and for what purpose. Users are entitled
-to know all parties that can access their PII, the limits to processing
-transparency, why the PII data is being requested, when the data will
-expire (Either from a collection or database), and what happens to it
-after that. This category also include legal rights around PII, and the
-implications of a contract when one is formed.''}
-
-This C implements transparency regarding user data and user privacy.
-Comprehension should answer the following questions:
-
-\begin{itemize}
-
-\item
-  \emph{WHO} collects data?
-\item
-  \emph{WHAT} data will be collected?
-\item
-  \emph{WHY} will data be collected and processed?
-\item
-  \emph{HOW} will data be collected and processed?
-\item
-  \emph{WHEN} will data expire?
-\item
-  What is allowed?
-\item
-  What choices ar possible?
-\end{itemize}
-
-All in all information of what's happening and why has to made
-accessible for users.
-
-\item
-\textbf{Consciousness \textbf{(critical!)}}
-
-\emph{``Users should \textbf{be aware} of when data collections occurs,
-when a contract is being formed between a user and data collector when
-their PII is set to expire, who's collecting the data, with whom the
-data will be shared, how to subsequently access the PII, and the
-purposes for which the data is being collected.''}
-
-This C seems to be critical for privacy protection. Consciousness
-complements Comprehension in respect that the latter just states that
-hard facts need to be delivered. However, those facts might get hidden
-in a terms and conditions section which nobody reads but still accepts
-anyway. In order to prevent that Consciousness states that a certain
-level of \textbf{Awareness} of those facts needs to be established.
-
-\item
-\textbf{Choice}
-
-\emph{``Users should \textbf{have choices} regarding data collection
-activities in terms of opting in or out, whether or not to provide data,
-and how to correct their data.''}
-
-Self explaining. This is the actual control enabled by the 7 C's.
-
-\item
-\textbf{Consent}
-
-\emph{``Users must first \textbf{consent} (meaning informed, explicit,
-unambiguous agreement) to data collection, use, and storage proposals
-for any PII. Privacy consent mechanisms should explicitly incorporate
-the mechanisms of comprehension, consciousness, limitations, and
-choice.''}
-
-This C might be special case of Choice. Before taking part a user should
-have the choice whether to join or not (Opt-In).
-
-\item
-\textbf{Context}
-
-\emph{``Users should \textbf{be able to change privacy preferences}
-according to context. Situational or physical context - such as crowded
-situations (for example, when at a service desk where several people
-can listen in on your exchange when you provide a phone number, or when
-you're in an online community chat room) - is different from when you
-perform a buy transaction with Amazon.com or in rooms with cameras
-(where digitization makes the information permanent and unmistakably
-you) and data context (such as the sensitivity of data, for example
-health data could dictate different actions on the same PII in different
-contexts.''}
-
-Self explaining. Refines Choice in context sensitive manner.
-
-\item
-\textbf{Confinement}
-
-\emph{``Users should \textbf{be able to set limits} on who may access
-their PII, for what purposes, and where and possibly when it may be
-stored. Setting limits could provide some good opportunities for future
-negotiation between vendors and users.''}
-
-Self explaining. Refines Choice regarding data collection and
-processing.
-
-\item
-\textbf{Consistency}
-
-\emph{``Users should \textbf{anticipate} with reasonable certainty what
-will occur if any action involving their PII is taken. That is, certain
-actions should be predictable on user access of PII or giving out of
-PII.''}
-
-Information given by Comprehension needs to be reliable to found
-choices.
-
-\end{itemize}
-
-\textbf{The 2 Steps of the 7 C's (Figure \ref{figure:The 2 Steps of the 7 Cs of User Privacy Control})}
-
-If we look closer at the 7 C's and how they try to enable control, we
-see that a 2 step approach is taken:
-
-%\includegraphics{../diagrams/png/7Cs2Steps.png}
-\input{./figures/The-2-Stepos-Of-The-7-Cs-Of-User-Privacy-Control}
-
-\begin{enumerate}
-
-\item
-\textbf{Enable \emph{Adequate} Control}
-
-The 7 C's try to enable control through choices. A user should be able
-to choose which data can be collected and processed depending on context
-and who will have access to the data. But cannot be random. In order to
-make substantiated choices and enable \emph{adequate} control a user
-needs have
-
-\begin{itemize}
-
-\item
-  \textbf{Comprehension:} access to hard facts
-\item
-  \textbf{Consistency:} trust that those facts are reliable
-\item
-  \textbf{Consciousness:} awareness of those facts to found choices
-\end{itemize}
-
-\item
-\textbf{Enable \emph{Actual} Control}
-
-Naturally after creating a certain amount of knowledge, a user needs
-also access to opportunities to make use of it. Therefore a user needs
-possibilities to actually make choices. Additionally to having choices at
-all, the 7 C's have 3 special choice categories:
-
-\begin{itemize}
-
-\item
-  \textbf{Choice}
-  \begin{itemize}
-  \item
-    \textbf{Consent:} the choice to opt-in
-  \item
-    \textbf{Confinement:} the choice to set limits regarding user data
-  \item
-    \textbf{Context:} the choice to set limits regarding user data
-    depending on certain context
-  \end{itemize}
-
-\end{itemize}
-
-\end{enumerate}
+% Seven Cs are now in separate Document

### 2014-10-24 PrivacyProtection-2014-09-09.tex
Updated Sensor Data Privacy Impact.

@@
 \textbf{GPS Data}.
-The collection of GPS data does directly infringe the citizens privacy of Location and Space, and the privacy of Data an Image.
+The GPS sensor gives the current longitude and latitude, the current global position of the mobile device and its carrier, although there is some artificial inaccuracy within civil use.
+Threefore, the collection of GPS data violates directly the citizens privacy of Location and Space, and the privacy of Data an Image.
+
 By inference we also get implicit violations of the Privacy of the Person, Privacy of Behavior and Action and Privacy of Association.
@@
 
-\textbf{Network Sensors}.
+\textbf{Network Sensors.}
 The GSM and WLAN sensors reveal the position of the mobile device and its carrier, when used in connection with external databases.
 The GSM sensor gives the exact cell, the mobile device has registered with at the current moment.
-The GPS sensor gives the current longitude and latitude, the current global position of the mobile device and its carrier, although there is some artificial inaccuracy within civil use.
-
-The WLAN and Bluetooth sensors record lists of the currently available local wireless networks and bluetooth clients.
-If such are known stationary entities, those sensors are considered as dangerous as the GSM sensor for the carriers locational privacy.
-
-The \textbf{Privacy of The Person} is generally concerned with one could best understand as \emph{Biometric Privacy}.
-Friedewald et al. paraphrase it as \emph{``\om the right to keep body functions and body characteristics \om private''}.
-
-\textbf{Privacy of Data and Image} demands, that \emph{``individual's data is not automatically available to other individuals''}.
-So every sensor violates the privacy of data and image, as data is transported into a foreign system where operators have access to it.
-
-According to Friedewald et al., the \textbf{Privacy of Location and Space} is concerned with one's \emph{``right to move about in public or semi-public space without being identified, tracked or monitored.''}.
-This is the location aspect of this type. The space aspect is concerned with one's \emph{``right to solitude''}, which generally includes one's right to an inviolate home and other private spaces.
-
-The \textbf{Privacy of Association} sates that everyone has the \emph{``right to associate with whomever they wish, without being monitored [automatically without reasonable suspicion]''}.
-This includes individuals and organizations.
-WLAN and Bluetooth sensors provide the ability to monitor such associations if their lists contain known entities.
-If one frequently connects with an organizational wireless network, e.g.~an university network, an association can be deduced (student or staff).
-The same goes for the Bluetooth sensor, if it is stationary.
-Additionally, if the recorded bluetooth clients are mobile, it is more or less possible to deduce association with the technical identity of (yet) anonymous individuals.
-
-Additionally, the GSM sensor could provide the association with the GSM operator.
-%TODO: NEEDS TO BE VERIFIED!
+
+The Bluetooth sensors record lists of the bluetooth clients in the direct neighbourhood.
+Since those clients are usually moving, inference of the position is usually not possible.
+Instead, bluetooth clients carried by a third person may infringe the Privacy of Association.
+
+A similar argumentat applies to WLAN snsors. If one frequently connects with an organizational wireless network, e.g.~an university network, an association can be deduced (student or staff).
+
+\textbf{Human Activity Recognition.}
+The detection and collection of human activities like walking, standing and running, can interfer with the Privacy of the Person, e.g.~since these movement patterns can be indicators for a person's health.
+Also, trivially, Privacy of Data and Image is violated.
+
+\textbf{Service Line Detection.}
+The detection and collection of the service line the user currently uses in the public transportation system allows inference of the location of the user, at least when entering and leaving those service lines at e.g.~bus stops.
+Implicit privacy violations apply accordingly.
+
+%TODO: Update Graphic: Violation of Privacy of Association is implicit for GSM and WLAN.
+%TODO: Add HAR and SLD to Table
 
@@
 
-%\subsubsection{Security Requirement Specification}
+\subsubsection{Security Requirement Specification}
 %TODO SKIP?
-
-\subsection{Step 3. Plan Development}
+TODO: Formulate requirements that address the high risk scenarios.
+
+%\subsection{Step 3. Plan Development}
 
@@
 
-\subsubsection{Security Measures}
+%\subsubsection{Security Measures}
 

### 2014-11-03 Implementations.tex
Updated Sensor Data Privacy Impact.
Added dashboard implementation description

@@
 \chapter{Implementation of Privacy Protection Measures}
-\subsection{Privacy Policy}
+\section{Privacy Policy}
 
@@
 \textbf{What Personal Information do we Collect?}\\
-The Live+Gov Service Center stores the following personal information:\\
-- Name\\
-- Email Address\\
-
 The Mobile Sensor Collection Application collects the following data from your mobile phone:\\
@@
 All collected data is transferred to our data center.
+
+The Live+Gov Services does NOT store or collect names and email addresses.
 
@@
 
-\section{Processing Journal}
-* name of the controller
-* purpose of processing
-* description of the data categories
-* recipients of the data if disclosed
-* transfer to third countries
-* description of security of processing
+\section{Privacy Dashboard}
+The Live+Gov Privacy Dashboard is a software component that allows the
+citizens to view the data about themselves that is stored in the
+Sensor Data Storage server.  The foundation of the Privacy Dashboard
+component is the inspection tool that was introduced in D1.1 as a
+means to perform basic quality controlling of the collected data.
 
-The journal is available for anybody to access via Email.
+The Privacy Dashboard is currently endowed with the following views:
+\begin{itemize}
+  \item \textbf{Login Screen} (cf. Figure \ref{fig:PDLogin}).  Before
+    access to the dashboard is granted, the user is prompted for
+    credentials.  The credentials are generated by the sensor
+    collection application on the mobile device.  The username can be
+    freely chosen by the user. The password is then generated by the
+    mobile application and transferred to the server alongside with
+    the recording.
+ \item \textbf{Recording Overview.} (Cf. Figure \ref{fig:PDOverview}).
+   This view shows all recordings that the user submitted to the Data
+   Storage Service including basic meta information like start data of
+   recording, duration. The user can click on each recording in order
+   to view the recorded data in the Raw Data View.
 
-Reporting to Data Protection Authority (Bundesamt für Datenschutz)?
+   An important feature of the recording overview table is the delete
+   button (cross on the very left.) That allows the user to
+   selectively delete recordings, that shall not be analyzed by the
+   Live+Gov system. Internally the click on the delete button only
+   flags the corresponding trip for deletion. The actual deletion is
+   carried out as a batch job on a regular bases. This addition step
+   prevents accidental loss of data.
 
-\section{Privacy Dashboard}
-TODO: Add description of dashboard. With relation to 7C's of Privacy.
+ \item \textbf{Raw Data View.} (cf. Figure \ref{fig:PDRawData}) The
+   raw data view shows all recorded raw data to the user.  GPS data is
+   visualized on a map. Motion sensor data is displayed as plots. Wifi
+   and GSM data is displayed in lists.
+   The raw data from each sensor can be downloaded individually as CSV file.
+
+ \item \textbf{Processing View.} (cf. Figure
+   \ref{fig:PDProcessedData}).  The processing view shows the data
+   mining end-products for each processing step.  The Figure shows the
+   processing view for the Activity Recognition. On the top a
+   time-line is show that visualizes the different recognized
+   activities via color-coding. A map below displays the corresponding
+   GPS track with the same color coding.
+
+   A similar view for Service Line Detection results is currently
+   under development.
+\end{itemize}
 
@@
 \caption{Live+Gov Privacy Dashboard Login}
+\label{fig:PDLogin}
 \end{figure}
@@
 \caption{Live+Gov Privacy Dashboard Recording Overview}
+\label{fig:PDOverview}
 \end{figure}
@@
 \caption{Live+Gov Privacy Dashboard Raw Data View}
+\label{fig:PDRawData}
 \end{figure}
@@
 \caption{Live+Gov Privacy Dashboard Processed Data View}
+\label{fig:PDProcessedData}
 \end{figure}
@@
 \section{Transfer Encryption (HTTPS/SSL)}
-\begin{figure}
+The data transfer from the mobile device to the Data Storage component
+as well as the transfer to the Privacy Dashboard relay on HTTP. In
+order to avoid eavesdropping by externals that have access to the
+communication channels, we secured all HTTP connections with SSL
+encryption.
+
+The implementation of this measure was rather straight forward. A SSL
+certificate was issued by the University of Koblenz, and installed to
+our webservers. Requests to this servers can now be done using HTTPS.
+Figure \ref{fig:PDHTTPS} shows a screenshot of such a secured
+connection to our Privacy Dashboard. The necessary adjustments to the
+Sensor Collection Component were minimal, since Android comes with
+native support for HTTPS connections.
+
+\begin{figure}[h]
 \includegraphics[width=\textwidth]{screenshots/HTTPS.png}
 \caption{HTTPS connection to Privacy Dashboard}
+\label{fig:PDHTTPS}
 \end{figure}
 
-\section{Database Access Policies}
-Access rules for data controller and processor.
-Data Processing Journal.
-Python script for deletion of data when it is older than a given date.
+\section{Data Expiring Dates}
+Trips stored in the Sensor Data Storage Service have been endowed with
+a data expiration date. A python script is provided that performs the
+of expired data. In our installation we have this script running as a
+cron job every night. The resulting schema of the trip table listed in
+Figure \ref{fig:TripScheme}.
+
+By default the expiration data is set to one year after the recording
+was uploaded. We are working on including a feature for setting the
+expiration date for each recording inside the privacy dashboard.
+
+\begin{figure}
+\centering
+\begin{verbatim}
+  Column  |          Type
+----------+------------------------
+ trip_id  | integer
+ user_id  | character varying(36)
+ start_ts | bigint
+ stop_ts  | bigint
+ name     | character varying(255)
+ deleted  | boolean
+ expires  | bigint
+\end{verbatim}
+\caption{Database Schema of the Trip Table}
+\label{fig:TripScheme}
+\end{figure}
 
 \section{Anonymization of trips}
-Privacy by Design: We do never collect name and email address in our DB.
-Python script that deletes user information from trip.
-% Python script that reomves precise date information, only day of week.
+The Sensor Data Storage Service follows a privacy by design guideline
+in that it does not store personal information like names and email
+address at all.
+
+In the case that such information should be linked to the user-id
+stored in the trip table we provide a python script that randomizes
+the user-ids and thereby removes all links that might have existed
+before.
+
+\begin{figure}[fontsize=\tiny]
+\centering
+\begin{verbatim}
+Usage: privacy.py [-d database name] [-u database user]
+                  [-p database password]
+                  [-h database host] [-o operation]
+
+Valid Operations:
+
+cleanup: Flags all expired rows as deleted
+    No Options
+
+anonymization: Sets all user_ids to 0
+    Option:
+        SQL Select Query: Every row found by this query will
+        be set to user_id = 0.
+
+haircut: Cuts trips where they split and could be traceable
+    Options:
+        First - the radius in which we look
+        Second - the number of users that needs to be in the
+                 radius
+
+blur: Offsets every GPS point by a random amount
+    Options:
+        The maximum distance a point can have to its origin
+
+k_anonymity: Creates a centroid for every GPS point
+    Options:
+        First - Number of maximum GPS points to use. If 0 then
+                we grab every point inside the radius
+        Second - Maximum distance a point can have to be used
+                for the creation of the new point
+\end{verbatim}
+\caption{Usage options for the privacy.py script.}
+\end{figure}
 
 \section{K-Anonymity of GPS routes}
-Python script that deletes all parts of journeys that have are not near to at least k other journeys.
+Python script that deletes all parts of journeys that have are not
+near to at least k other journeys.
 

### 2014-11-05 PrivacyProtection-2014-09-09.tex
Added dashboard implementation description
started table

--- 2014-10-24 PrivacyProtection-2014-09-09.tex
@@
 
+In this section we analyze the impact of the disclosure sensor data and certain processing results to the citizens privacy.
+This analyisis builds upon the preceeding analyis, but is more focused on the concrete type of data vailable.
+For example, the disclosure of Service Line Detection results, does violate the Privacy of Location and Space but not the Privacy of Association that is violated by general GPS tracking.
+
 %TODO: Reformulate the section.
@@
 
-In this section we analyze the impact of the disclosure sensor data and certain processing results to the citizens privacy.
-This analyisis builds upon the preceeding analyis, but is more focused on the concrete type of data vailable.
-For example, the disclosure of Service Line Detection results, does violate the Privacy of Location and Space but not the Privacy of Association that is violated by general GPS tracking.
-
-\textbf{GPS Data}.
-The collection of GPS data does directly infringe the citizens privacy of Location and Space, and the privacy of Data an Image.
-By inference we also get implicit violations of the Privacy of the Person, Privacy of Behavior and Action and Privacy of Association.
-
-\textbf{Motion Sensors.}
-Accelerometer, Rotation Vector, Gyroscope and Magnetic field sensor measure the physical movement of the mobile device on all three axes.
+The \textbf{Privacy of The Person} is generally concerned with one could best understand as \emph{Biometric Privacy}.
+Friedewald et al. paraphrase it as \emph{``\om the right to keep body functions and body characteristics \om private''}.
+Accelerometer, Rotation Vector and Gyroscope measure the physical movement of the mobile device on all three axes.
 If the mobile device is carried ``normally'' its safe to say that those sensors also measure the moments of its carrier.
-So his privacy is infringed regarding biometric behaviour, as it is captured automatically
-Privacy of Data and Image is trivially threatened because here sensor data is individual data, a priori.
-
-By inference we also get implicit violations of the Privacy of Behavior and Action.
-
-\textbf{Network Sensors}.
-The GSM and WLAN sensors reveal the position of the mobile device and its carrier, when used in connection with external databases.
+So his privacy is infringed regarding biometric behaviour, as it is captured automatically.
+
+\textbf{Privacy of Data and Image} demands, that \emph{``individual's data is not automatically available to other individuals''}.
+This type of privacy is trivially threatened because here sensor data is individual data, a priori.
+So every sensor violates the privacy of data and image, as data is transported into a foreign system where operators have access to it.
+
+According to Friedewald et al., the \textbf{Privacy of Location and Space} is concerned with one's \emph{``right to move about in public or semi-public space without being identified, tracked or monitored.''}.
+This is the location aspect of this type. The space aspect is concerned with one's \emph{``right to solitude''}, which generally includes one's right to an inviolate home and other private spaces.
+Obviously the GPS and GSM sensors violate such right about not being tracked, because they reveal the position of the mobile device and its carrier.
 The GSM sensor gives the exact cell, the mobile device has registered with at the current moment.
@@
 
-The \textbf{Privacy of The Person} is generally concerned with one could best understand as \emph{Biometric Privacy}.
-Friedewald et al. paraphrase it as \emph{``\om the right to keep body functions and body characteristics \om private''}.
-
-\textbf{Privacy of Data and Image} demands, that \emph{``individual's data is not automatically available to other individuals''}.
-So every sensor violates the privacy of data and image, as data is transported into a foreign system where operators have access to it.
-
-According to Friedewald et al., the \textbf{Privacy of Location and Space} is concerned with one's \emph{``right to move about in public or semi-public space without being identified, tracked or monitored.''}.
-This is the location aspect of this type. The space aspect is concerned with one's \emph{``right to solitude''}, which generally includes one's right to an inviolate home and other private spaces.
+The Magnetic field sensor is not regarded very dangerous to the carrier's privacy, because it does not allow very precise localization.
+But it can limit the possibilities for the global position of the mobile device.
+Here, it is just named for completeness sake.
 
@@
 
-%\subsubsection{Security Requirement Specification}
+\begin{tabular}{|c|c|c|c|}
+
+\hline
+Threat & Likeliness & Loss & Risk
+\\\hline
+
+Threat & Likeliness & Loss & Risk
+\\\hline
+\hline
+
+\end{tabular}
+
+\subsubsection{Security Requirement Specification}
 %TODO SKIP?
@@
 
-% \subsection{Legal Privacy Requirements} \ref{sec:LEGAL_REQ}
+\subsection{Legal Privacy Requirements} \ref{sec:LEGAL_REQ}
 
@@
 
-% Seven Cs are now in separate Document
+\textbf{The 7 C's of user privacy control (Figure \ref{figure:The Seven Cs of User Privacy Control})}
+
+This is a note on an excerpt from the article \emph{Sociotechnical
+Architecture for Online Privacy} \cite{1} called
+\textbf{The 7 C's of user privacy control}. Those 7 C's are aspects
+which should be covered by measures for implementing user privacy. They
+derive from an interpretation of privacy which could be summarized as
+\emph{``One's ability to control/seclude information about oneself''}.
+
+%\includegraphics{../diagrams/png/The7Cs.png}
+\input{./figures/The-7-Cs-Of-User-Privacy-Control}
+
+\begin{itemize}
+
+\item
+\textbf{Comprehension}
+
+\emph{``Users should \textbf{understand} ho personal identifiable
+information (PII) is handled, who's collecting it and for what purpose,
+and who will process the PII and for what purpose. Users are entitled
+to know all parties that can access their PII, the limits to processing
+transparency, why the PII data is being requested, when the data will
+expire (Either from a collection or database), and what happens to it
+after that. This category also include legal rights around PII, and the
+implications of a contract when one is formed.''}
+
+This C implements transparency regarding user data and user privacy.
+Comprehension should answer the following questions:
+
+\begin{itemize}
+
+\item
+  \emph{WHO} collects data?
+\item
+  \emph{WHAT} data will be collected?
+\item
+  \emph{WHY} will data be collected and processed?
+\item
+  \emph{HOW} will data be collected and processed?
+\item
+  \emph{WHEN} will data expire?
+\item
+  What is allowed?
+\item
+  What choices ar possible?
+\end{itemize}
+
+All in all information of what's happening and why has to made
+accessible for users.
+
+\item
+\textbf{Consciousness \textbf{(critical!)}}
+
+\emph{``Users should \textbf{be aware} of when data collections occurs,
+when a contract is being formed between a user and data collector when
+their PII is set to expire, who's collecting the data, with whom the
+data will be shared, how to subsequently access the PII, and the
+purposes for which the data is being collected.''}
+
+This C seems to be critical for privacy protection. Consciousness
+complements Comprehension in respect that the latter just states that
+hard facts need to be delivered. However, those facts might get hidden
+in a terms and conditions section which nobody reads but still accepts
+anyway. In order to prevent that Consciousness states that a certain
+level of \textbf{Awareness} of those facts needs to be established.
+
+\item
+\textbf{Choice}
+
+\emph{``Users should \textbf{have choices} regarding data collection
+activities in terms of opting in or out, whether or not to provide data,
+and how to correct their data.''}
+
+Self explaining. This is the actual control enabled by the 7 C's.
+
+\item
+\textbf{Consent}
+
+\emph{``Users must first \textbf{consent} (meaning informed, explicit,
+unambiguous agreement) to data collection, use, and storage proposals
+for any PII. Privacy consent mechanisms should explicitly incorporate
+the mechanisms of comprehension, consciousness, limitations, and
+choice.''}
+
+This C might be special case of Choice. Before taking part a user should
+have the choice whether to join or not (Opt-In).
+
+\item
+\textbf{Context}
+
+\emph{``Users should \textbf{be able to change privacy preferences}
+according to context. Situational or physical context - such as crowded
+situations (for example, when at a service desk where several people
+can listen in on your exchange when you provide a phone number, or when
+you're in an online community chat room) - is different from when you
+perform a buy transaction with Amazon.com or in rooms with cameras
+(where digitization makes the information permanent and unmistakably
+you) and data context (such as the sensitivity of data, for example
+health data could dictate different actions on the same PII in different
+contexts.''}
+
+Self explaining. Refines Choice in context sensitive manner.
+
+\item
+\textbf{Confinement}
+
+\emph{``Users should \textbf{be able to set limits} on who may access
+their PII, for what purposes, and where and possibly when it may be
+stored. Setting limits could provide some good opportunities for future
+negotiation between vendors and users.''}
+
+Self explaining. Refines Choice regarding data collection and
+processing.
+
+\item
+\textbf{Consistency}
+
+\emph{``Users should \textbf{anticipate} with reasonable certainty what
+will occur if any action involving their PII is taken. That is, certain
+actions should be predictable on user access of PII or giving out of
+PII.''}
+
+Information given by Comprehension needs to be reliable to found
+choices.
+\end{itemize}
+
+\textbf{The 2 Steps of the 7 C's (Figure \ref{figure:The 2 Steps of the 7 Cs of User Privacy Control})}
+
+If we look closer at the 7 C's and how they try to enable control, we
+see that a 2 step approach is taken:
+
+%\includegraphics{../diagrams/png/7Cs2Steps.png}
+\input{./figures/The-2-Stepos-Of-The-7-Cs-Of-User-Privacy-Control}
+
+\begin{enumerate}
+
+\item
+\textbf{Enable \emph{Adequate} Control}
+
+The 7 C's try to enable control through choices. A user should be able
+to choose which data can be collected and processed depending on context
+and who will have access to the data. But cannot be random. In order to
+make substantiated choices and enable \emph{adequate} control a user
+needs have
+
+\begin{itemize}
+
+\item
+  \textbf{Comprehension:} access to hard facts
+\item
+  \textbf{Consistency:} trust that those facts are reliable
+\item
+  \textbf{Consciousness:} awareness of those facts to found choices
+\end{itemize}
+
+\item
+\textbf{Enable \emph{Actual} Control}
+
+Naturally after creating a certain amount of knowledge, a user needs
+also access to opportunities to make use of it. Therefore a user needs
+possibilities to actually make choices. Additionally to having choices at
+all, the 7 C's have 3 special choice categories:
+
+\begin{itemize}
+
+\item
+  \textbf{Choice}
+  \begin{itemize}
+  \item
+    \textbf{Consent:} the choice to opt-in
+  \item
+    \textbf{Confinement:} the choice to set limits regarding user data
+  \item
+    \textbf{Context:} the choice to set limits regarding user data
+    depending on certain context
+  \end{itemize}
+
+\end{itemize}
+
+\end{enumerate}

### 2014-11-05 D1-3.tex
Added Privacy Policy Section
WIP

@@
 \documentclass[external]{20120615_deliverable_template_ukob}
-
 \usepackage{xspace}
@@
 \usepackage{color, colortbl}
+\usepackage{listings}
+\usepackage{subfigure}
 %\usepackage{showkeys}
@@
 \LGaddauthorNoPhone{UKob}{Maximilan Meffert}%
-{}
+{\small maxmeffert@uni-koblenz.de}
 
@@
 
-\chapter{Introduction}
-\label{chap:Introduction}
+\chapter{Introduction}\label{chap:Introduction}
+In recent days the importance of privacy protection has been amplified
+by the reports about the mass surveillance of ordinary citizens on a
+global scale by the NSA and other intelligence agencies around the
+world.
+
+While aiming at the noble cause of enhancing eParticipation using
+mobile technologies, Live+Gov systems do process a large variety data
+that is potentially infringing the citizens privacy. The captured data
+includes personal information like name, phone numbers and email
+addresses and sensor data from GPS and accelerometer sensors. Also
+with some applications it is possible to gather images and textual
+input from the citizen.
+
+While the collection of this data is necessary for providing the
+advanced services that Live+Gov aims to deliver, at he same time, the
+available raw data can be used to draw a very detailed picture of the
+private life of the citizen. For instance can GPS location tracking be
+used to reveal shopping habits (e.g. when a car seller is visited) and
+associations to political groups (when a meeting is
+attended). Accelerometer data can be used to infer medical conditions
+like walking disabilities.
+
+Images can contain faces of nearby persons to with whom the citizen is
+associated. All this data is highly sensitive to the citizens privacy
+and can be used against the citizen if it falls in the wrong hands.
+
+The great importance of protecting the citizens privacy should be
+apparent from these examples. The European Union, as well as many
+other countries in the past, has set out a number of directives that
+regulate the collection, processing and use of privacy sensitive
+data. The most important legislation on the European Level is the Data
+Protection Directive 95/46/EC [1], which is also summarized in the
+European Charta of Fundamental Rights [2] of EU citizens:
+
+\begin{quote}
+Article 7. Respect for private and family life.
+
+Everyone has the right to respect for his or her private and family
+life, home and communications.
+
+Article 8. Protection of personal data.
+
+(1) Everyone has the right to the protection of personal data
+concerning him or her.
+
+(2) Such data must be processed fairly for specified purposes and on
+the basis of the consent of the person concerned or some other
+legitimate basis laid down by law.  Everyone has the right of access
+to data which has been collected concerning him or her, and the right
+to have it rectified.
+
+(3) Compliance with these rules shall be subject to control by an
+independent authority.  The Live+Gov fully consortium complies with
+this legislation and is moreover committed to convey and enact the
+ethical values inherent in these texts.
+\end{quote}
+
+Privacy as Control over Private Data
+
+The ethical aspects of privacy have been the subject of study of many
+social scientists and philosophers [3]. Many researchers do focus on
+the ways in which privacy can be infringed. Thus they invest a great
+amount of work in exploring threats, instead of describing why privacy
+is so valuable to us. One scholar which follows an alternative aproach
+in the context of digital monitoring is Charles Fried [4]. He
+investiages, why we are intuitively so sensitive to violations of our
+privacy. For him privacy is not asserted as an intrinsic value by
+itself, he rather sated:
+
+\begin{quote}
+Privacy is not simply an absence of information about us in the minds of others;
+rather it is the control we have over information about ourselves.
+\end{quote}
+
+Depending on the conversation partner, we change the degree of
+information we share if it is a total stranger, a colleague, close
+friend or a doctor. With doctors we share information of great
+intimacy we do not share with anyone else.
+
+Moreover, we trust those friends to not reveal information about us to
+others by respecting their privacy. Trust needs the possibility of
+unknown failure. If we would constantly monitor our partners, they
+cannot fail unnoticed nor can they willingly share that information
+with us.  So privacy, according to Fried, is the foundation of our
+core relations. And thus it is valuable, because those relations are
+essential to human society.
+
+Fried’s study on the understanding of privacy provided a great
+contribution to the research on the same term in philosophy and
+computer science [5] and despite the fact his text was published in
+1970, he already included technologies to its viewpoint (like location
+monitoring) that are particularly relevant to our Context.
+
+Live+Gov Privacy Dashboard gives citizens back the Control
+
+In order to meet the the legal requirements and offer the citiens to
+keep a maximum of contol about their personal data, Live+Gov systems
+implements a number of privacy protection measure. One of the measures
+that is under current development is a Privacy Dashboard (Figure
+1). Using this dashboard the citizen will be able to take the
+following actions: View and export all data from the citizen that is
+currently stored in the Live+Gov system Selectively delete parts of
+the stored data Get information about processing applied to this data
+Consent or decline selectively to processing steps
+
+View the end poducts of data mining applied to his data in a
+comprehensive, graphical way These measures provide the citizen with a
+maximal respect for their privacy while at the same time allowing
+advanced data processing. Although, abusive exploitation of the data
+is still possible, the level of transparency offered by the dashboard
+builds trust between the citizen and the service provider.
 
@@
 
-\chapter{Implementation of Privacy Protection Measures}
-\subsection{Privacy Policy}
-\subsection{Privacy Dashboard}
-\subsection{Transfer Encryption (HTTPS/SSL)}
-\subsection{Database Access Policies}
-\subsection{Anonymization of recorded trips}
-\subsection{Randomization of GPS data}
+\input{Implementations}
+
+\chapter{Summary}\label{chap:Summary}
 

### 2014-11-05 Implementations.tex

@@
 \chapter{Implementation of Privacy Protection Measures}
-\section{Privacy Policy}
+\section{Privacy Policy} \label{sec:PP}
 
@@
 The Live+Gov Privacy Dashboard is a software component that allows the
-citizens to view the data about themselves that is stored in the
-Sensor Data Storage server.  The foundation of the Privacy Dashboard
-component is the inspection tool that was introduced in D1.1 as a
-means to perform basic quality controlling of the collected data.
+citizens to view and control the data about themselves that is stored
+in the Sensor Data Storage server.  The foundation of the Privacy
+Dashboard component is the inspection tool that was introduced in D1.1
+as a means to perform basic quality controlling of the collected data.
 
@@
     the recording.
- \item \textbf{Recording Overview.} (Cf. Figure \ref{fig:PDOverview}).
+
+ \item \textbf{Recording Overview} (Cf. Figure \ref{fig:PDOverview}).
    This view shows all recordings that the user submitted to the Data
@@
 
- \item \textbf{Raw Data View.} (cf. Figure \ref{fig:PDRawData}) The
-   raw data view shows all recorded raw data to the user.  GPS data is
-   visualized on a map. Motion sensor data is displayed as plots. Wifi
-   and GSM data is displayed in lists.
-   The raw data from each sensor can be downloaded individually as CSV file.
+ \item \textbf{Raw Data View} (cf. Figure \ref{fig:PDRawData}).
+   The raw data view shows all recorded raw data to the user.  GPS
+   data is visualized on a map. Motion sensor data is displayed as
+   plots. Wifi and GSM data is displayed in lists.  The raw data from
+   each sensor can be downloaded individually as CSV file.
 
- \item \textbf{Processing View.} (cf. Figure
-   \ref{fig:PDProcessedData}).  The processing view shows the data
-   mining end-products for each processing step.  The Figure shows the
-   processing view for the Activity Recognition. On the top a
-   time-line is show that visualizes the different recognized
-   activities via color-coding. A map below displays the corresponding
-   GPS track with the same color coding.
+ \item \textbf{Processing View} (cf. Figure \ref{fig:PDProcessedData}).
+   The processing view shows the data mining end-products for each
+   processing step.  The Figure shows the processing view for the
+   Activity Recognition. On the top a time-line is show that
+   visualizes the different recognized activities via color-coding. A
+   map below displays the corresponding GPS track with the same color
+   coding.
 
@@
    under development.
+ \item \textbf{Privacy Settings.}
+   This view contains the privacy policy shown in Section \ref{sec:PP}.
+
+   We intend to include settings also settings concerning the
+   expiration date of the recordings, agreements to certain processing
+   steps, view logs of performed processing.
 \end{itemize}
+
+The Privacy Dashboard is implemented on a
+nodes.js\footnote{\url{http://nodejs.org}} web server using
+Express.js\footnote{\url{http://expressjs.com/}} as a web-framework
+and d3.js\footnote{\url{http://d3js.org/}} as plotting library. The
+data is pulled directly from the PostgreSQL database of the Data
+Storage Service.
 
@@
 \end{figure}
+
+\pagebreak
 
@@
 cron job every night. The resulting schema of the trip table listed in
-Figure \ref{fig:TripScheme}.
+Figure \ref{fig:TripScheme}. The expiring date is encoded as a 64bit
+integer holding the unix timestamp in seconds of the expiration date.
 
@@
 \begin{figure}
-\centering
-\begin{verbatim}
+\begin{center}
+\begin{alltt}
   Column  |          Type
@@
  deleted  | boolean
- expires  | bigint
-\end{verbatim}
+ \textbf{expires}  | \textbf{bigint}
+\end{alltt}
+\end{center}
 \caption{Database Schema of the Trip Table}
@@
 
-anonymization: Sets all user_ids to 0
+anonymization: Randomizes user_ids
     Option:
         SQL Select Query: Every row found by this query will
-        be set to user_id = 0.
+        be randomized.
 
@@
     Options:
-        The maximum distance a point can have to its origin
+        The expected distance a point has to its origin
 
@@
 \caption{Usage options for the privacy.py script.}
+\label{fig:privacypy}
 \end{figure}
 
-\section{K-Anonymity of GPS routes}
-Python script that deletes all parts of journeys that have are not
-near to at least k other journeys.
+\section{GPS Anonymization: Gaussian Noise}
 
-TODO: Insert Literature Reference.
-TODO: Add screenshots.
+The Gaussian Noise anonymization method adds random noise to the
+stored GPS points. The \texttt{privacy.py} script (cf. Figure
+\ref{fig:privacypy}) implements this feature as "blur'' and allows the
+specification of the displacement variance.  Figure \ref{fig:noise}
+shows an example of a GPS track with and without added noise.
 
-\section{Randomization of GPS data}
-Python Script that adds random noise to GPS data.
+The Gaussian Noise anonymization is a basic method is usually mixed
+well with other methods.
+
+\begin{figure}
+  \centering
+  \subfigure[Original GPS samples]{
+    \includegraphics[width=0.45\textwidth]{screenshots/GPS_sample.png}
+    }
+  \quad
+  \subfigure[GPS samples with noise]{
+    \includegraphics[width=0.45\textwidth]{screenshots/AfterBlur.png}
+  }
+  \caption{GPS Anonymization with Gaussian Noise}
+  \label{fig:noise}
+\end{figure}
+
+\section{GPS Anonymization: K-Anonymity}
+This method of GPS anonymization is a variant of Sweeney's K-Anonymity
+\cite{sweeney2002k} applied to GPS data along the lines of Grutser and
+Grunwald \cite{Gruteser2003}.  The basic idea is that the resulting
+data from a single person should not distinguishable form data from
+K-1 other persons. In order to achieve such kind of anonymization we
+calculate for each GPS point $P$ by a user $U$ the $K-1$ nearest $GPS$
+points $Q_1, \dots, Q_{K-1}$ which are recorded by $K-1$ different
+users.  The anonymized point $R$ is now the centroid of the $K$ points
+$(P,Q_1,\dots,Q_{K-1})$. This operations applied to all data points
+$P$ in the dataset. The resulting anonymized points $R$ are stored in
+a new table, so that these points do not take part in further the
+centroid calculations.
+
+Figure \ref{fig:k-ano} shows the results of application of K-anonymity
+to the original GPS trip displayed in Figure \ref{fig:noise}.
+$K$-Anonymity implementation script (cf. Figure \ref{fig:privacypy})
+supports a second parameter that specifies the maximal distance that a
+point can get displaced by this method.
+
+\begin{figure}
+  \centering
+  \subfigure[K=7 anonymization]{
+    \includegraphics[width=0.4\textwidth]{screenshots/k_anon_7_nearest_points.png}
+  }
+  \quad
+  \subfigure[K=15 Anonymization]{
+    \includegraphics[width=0.4\textwidth]{screenshots/k_ano_15_nearest_points.png}
+  }
+  \caption{GPS Anonymization with K-Anonimity}
+  \label{fig:k-ano}
+\end{figure}
+
+\section{GPS Anonymization: Haircut}
+
+The \emph{haircut} anonymization method was developed by UKob to
+address the problem of GPS tracks that show locations, like the home
+address, that are unique to a single user. Looking on our data we see
+many trips that originate at a unique location and later on meet other
+tracks at locations of public interest like bus stops or larger
+streets. The haircut methods cuts away parts of tracks that
+"stick-out'' the whole set of recorded trajectories. For a given GPS
+location $P$ the number of users is computed that have GPS samples
+near to $P$. If this number is lower than a given threshold (e.g. $3$)
+the point $P$ is dropped from the dataset.
+
+The implementation script (cf. Figure \ref{fig:privacypy}) takes two
+supports two parameters that specify the radius to look for other GPS
+samples and the number of different users that are required to have
+samples inside this radius. Figure \ref{fig:haircut} shows an example
+a two-trips before and after the haircut.
+
+\begin{figure}
+  \centering
+  \subfigure[Original GPS tracks]{
+    \includegraphics[width=0.4\textwidth]{screenshots/haircut_pre.png}
+  }
+  \quad
+  \subfigure[GPS tracks with haircut]{
+    \includegraphics[width=0.4\textwidth]{screenshots/haircut_post.png}
+  }
+  \caption{GPS Anonymization with Haircut Method}
+  \label{fig:haircut}
+\end{figure}

### 2014-11-05 SevenCs.tex

\textbf{The 7 C's of user privacy control (Figure \ref{figure:The Seven Cs of User Privacy Control})}

This is a note on an excerpt from the article \emph{Sociotechnical
Architecture for Online Privacy} \cite{1} called
\textbf{The 7 C's of user privacy control}. Those 7 C's are aspects
which should be covered by measures for implementing user privacy. They
derive from an interpretation of privacy which could be summarized as
\emph{``One's ability to control/seclude information about oneself''}.

%\includegraphics{../diagrams/png/The7Cs.png}
\input{./figures/The-7-Cs-Of-User-Privacy-Control}

\begin{itemize}

\item
\textbf{Comprehension}

\emph{``Users should \textbf{understand} ho personal identifiable
information (PII) is handled, who's collecting it and for what purpose,
and who will process the PII and for what purpose. Users are entitled
to know all parties that can access their PII, the limits to processing
transparency, why the PII data is being requested, when the data will
expire (Either from a collection or database), and what happens to it
after that. This category also include legal rights around PII, and the
implications of a contract when one is formed.''}

This C implements transparency regarding user data and user privacy.
Comprehension should answer the following questions:

\begin{itemize}

\item
  \emph{WHO} collects data?
\item
  \emph{WHAT} data will be collected?
\item
  \emph{WHY} will data be collected and processed?
\item
  \emph{HOW} will data be collected and processed?
\item
  \emph{WHEN} will data expire?
\item
  What is allowed?
\item
  What choices ar possible?
\end{itemize}

All in all information of what's happening and why has to made
accessible for users.

\item
\textbf{Consciousness \textbf{(critical!)}}

\emph{``Users should \textbf{be aware} of when data collections occurs,
when a contract is being formed between a user and data collector when
their PII is set to expire, who's collecting the data, with whom the
data will be shared, how to subsequently access the PII, and the
purposes for which the data is being collected.''}

This C seems to be critical for privacy protection. Consciousness
complements Comprehension in respect that the latter just states that
hard facts need to be delivered. However, those facts might get hidden
in a terms and conditions section which nobody reads but still accepts
anyway. In order to prevent that Consciousness states that a certain
level of \textbf{Awareness} of those facts needs to be established.

\item
\textbf{Choice}

\emph{``Users should \textbf{have choices} regarding data collection
activities in terms of opting in or out, whether or not to provide data,
and how to correct their data.''}

Self explaining. This is the actual control enabled by the 7 C's.

\item
\textbf{Consent}

\emph{``Users must first \textbf{consent} (meaning informed, explicit,
unambiguous agreement) to data collection, use, and storage proposals
for any PII. Privacy consent mechanisms should explicitly incorporate
the mechanisms of comprehension, consciousness, limitations, and
choice.''}

This C might be special case of Choice. Before taking part a user should
have the choice whether to join or not (Opt-In).

\item
\textbf{Context}

\emph{``Users should \textbf{be able to change privacy preferences}
according to context. Situational or physical context - such as crowded
situations (for example, when at a service desk where several people
can listen in on your exchange when you provide a phone number, or when
you're in an online community chat room) - is different from when you
perform a buy transaction with Amazon.com or in rooms with cameras
(where digitization makes the information permanent and unmistakably
you) and data context (such as the sensitivity of data, for example
health data could dictate different actions on the same PII in different
contexts.''}

Self explaining. Refines Choice in context sensitive manner.

\item
\textbf{Confinement}

\emph{``Users should \textbf{be able to set limits} on who may access
their PII, for what purposes, and where and possibly when it may be
stored. Setting limits could provide some good opportunities for future
negotiation between vendors and users.''}

Self explaining. Refines Choice regarding data collection and
processing.

\item
\textbf{Consistency}

\emph{``Users should \textbf{anticipate} with reasonable certainty what
will occur if any action involving their PII is taken. That is, certain
actions should be predictable on user access of PII or giving out of
PII.''}

Information given by Comprehension needs to be reliable to found
choices.
\end{itemize}

\textbf{The 2 Steps of the 7 C's (Figure \ref{figure:The 2 Steps of the 7 Cs of User Privacy Control})}

If we look closer at the 7 C's and how they try to enable control, we
see that a 2 step approach is taken:

%\includegraphics{../diagrams/png/7Cs2Steps.png}
\input{./figures/The-2-Stepos-Of-The-7-Cs-Of-User-Privacy-Control}

\begin{enumerate}

\item
\textbf{Enable \emph{Adequate} Control}

The 7 C's try to enable control through choices. A user should be able
to choose which data can be collected and processed depending on context
and who will have access to the data. But cannot be random. In order to
make substantiated choices and enable \emph{adequate} control a user
needs have

\begin{itemize}

\item
  \textbf{Comprehension:} access to hard facts
\item
  \textbf{Consistency:} trust that those facts are reliable
\item
  \textbf{Consciousness:} awareness of those facts to found choices
\end{itemize}

\item
\textbf{Enable \emph{Actual} Control}

Naturally after creating a certain amount of knowledge, a user needs
also access to opportunities to make use of it. Therefore a user needs
possibilities to actually make choices. Additionally to having choices at
all, the 7 C's have 3 special choice categories:

\begin{itemize}

\item
  \textbf{Choice}
  \begin{itemize}
  \item
    \textbf{Consent:} the choice to opt-in
  \item
    \textbf{Confinement:} the choice to set limits regarding user data
  \item
    \textbf{Context:} the choice to set limits regarding user data
    depending on certain context
  \end{itemize}

\end{itemize}

\end{enumerate}

### 2014-11-05 D1-3.tex
Extended Threat Scenarios

@@
 
-Live+Gov Privacy Dashboard gives citizens back the Control
+\input{PrivacyProtection-2014-09-09}
+
+\clearpage
+
+\input{Implementations}
+
+\chapter{Summary}\label{chap:Summary}
+* Discussed Legal and ethical aspects of privacy
+* Made Privacy Security Analysis of Live+Gov System
+* Presented several Privacy Protection Measures that have been implemented.
+* Most importantly the Privacy Dashboard, which gives the citizen control over their data.
 
@@
 
-\input{PrivacyProtection-2014-09-09}
-
-\clearpage
-
-\input{Implementations}
-
-\chapter{Summary}\label{chap:Summary}
-
 \bibliography{D1-3}

### 2014-11-05 Implementations.tex
Extended Threat Scenarios

@@
 The Mobile Sensor Collection Application collects the following data from your mobile phone:\\
-- Sensor Data from Accelerometer (50Hz)\\
-- Sensor Data from GPS sensors (~every 30 sec)\\
-- Daily Human Activities (walking, standing, sitting, etc.) (every 30sec)\\
-- Used Service Lines (bus, train, tram) (every 2 minutes)\\
+- Sensor Data from Accelerometer\\
+- Sensor Data from GPS sensors\\
+- Daily Human Activities (walking, standing, sitting, etc.)\\
+- Used Service Lines (bus, train, tram)
 
-All collected data is transferred to our data center.
-
-The Live+Gov Services does NOT store or collect names and email addresses.
+All collected data is transferred to our data center. The Live+Gov
+Services does NOT store or collect names and email addresses.
 
@@
 
+\textbf{Access to Stored Data}\\
+All personal data can be accessed at our Privacy Dashboard at
+\url{http://liveandgov.uni-koblenz.de/trial/dashboard}
+
+There you have the possibility to:\\
+- View all recordings made\\
+- Delete individual recordings\\
+- View the raw data collected for each recording\\
+- Download the data for each recording\\
+- View the results of the applied data processing
+
 \textbf{When is the data deleted?}\\
-The full data-set will be stored until January 2015.
+The full data-set will be stored until January 2016.
 
-Afterwards we will delete Names and Email addresses and anonymize
-the sensor data and processing results.  The anonymized data will
-be used by the the Consortium Partners for research purposes.
+Anonymized artifacts of this data are stored longer than that and will
+be used by the Consortium Partners for research purposes.
 
 \textbf{For which purposes is the data collected?}\\
-The collected data will be used for personalization of user experience of the mobile application.
-Moreover, we will analyize travel patterns in the Helsinki Urban region for general research and for improvement and optimization of the
-infrastructure provided by HSL.
+The collected data will be used for personalization of user experience
+of the mobile application.  Moreover, we will analyze travel patterns
+in the Helsinki Urban region for general research and for improvement
+and optimization of the infrastructure provided by HSL.
 
-\textbf{Identity of Data Controller}\\
-The data is controlled by Heinrich Hartmann <hartmann@uni-koblenz.de>
+\textbf{Collaboration with Third Parties}\\
+The raw data is not shared with any parties external to the Live+Gov
+Consortium.
 
-The following organizations have access to the collected data:
+\textbf{Who has access to the data?}
+The following organizations inside the Live+Gov Consortium have access
+to the collected data:
 \begin{enumerate}
@@
 
-\item Mattersoft
-\url{http://www.mattersoft.fi/}
+\item Mattersoft \\
+\url{http://www.mattersoft.fi/} \\
 Hämeenkatu 13, 33100 Tampere, Finland\\
@@
 
-\item Centre for Research and Technology Hellas (CERTH)
+\item Centre for Research and Technology Hellas (CERTH)\\
 \url{http://mklab.iti.gr/contact}\\
@@
 
-\textbf{Access to Stored Data}\\
-All personal data can be accessed at our Privacy Dashboard at
-\url{http://liveandgov.uni-koblenz.de/trial/dashboard}
+Aggregated report based on anonymized data are disclosed to officials
+at HSL \url{http://www.hsl.fi/} and might be made public in the form
+of a blog post or research article.  These aggregated reports show
+distributions of all routes that have been collected in the system and
+do not allow to infere the location of a single user at a given point
+in time.
 
-There you have the possibility to:\\
-- View all recordings made\\
-- View the raw data collected for each recording\\
-- Download the data for each recording\\
-- Delete records
-
-\textbf{Collaboration with Third Parties}\\
-The raw data is not shared with any parties external to the Live+Gov Consortium.
-
-Aggregated report based on anonymized data are disclosed to officials at HSL \url{http://www.hsl.fi/} and might be made
-public in the form of a blog post or research article.  These aggregated reports show distributions of all routes that
-have been collected in the system and do not allow to infere the location of a single user at a given point in time.
+\textbf{Identity of Data Controller}\\
+The data is controlled by Dr. H.  Hartmann <hartmann@uni-koblenz.de>.
 
@@
 In the case that such information should be linked to the user-id
-stored in the trip table we provide a python script that randomizes
-the user-ids and thereby removes all links that might have existed
-before.
+stored in the trip table we provide a python script (cf. Figure
+\ref{fig:privacypy}) that randomizes the user-ids and thereby removes
+all links that might have existed before. The scripts allows as a
+parameter a general SQL query that selects a sub-set of the trips that
+shall be anonymized.
 
@@
 
-The Gaussian Noise anonymization is a basic method is usually mixed
-well with other methods.
+The Gaussian Noise anonymization, even with a low noise radius of 5m,
+disables the possibility to correlate GPS data with data from cameras
+to identify a single person. It is typically used in conjunction with
+other method to provide a maximum of anonymity.
 

### 2014-11-05 PrivacyProtection-2014-09-09.tex
finished threat eval table
Extended Threat Scenarios

--- 2014-10-24 PrivacyProtection-2014-09-09.tex
@@
 
-\section{Privacy Definition and Taxonomy}
+\section{Definition and Types of Privacy}
 \label{sec:taxonomy}
@@
 
-Please refer to Section \ref{sec:taxonomy} for more details.
+We refer to Section \ref{sec:taxonomy} for more details.
 
@@
 
-\textbf{Unhappy Employees.} Susceptible for social engineering.
-%TODO: Add description
+\textbf{Unhappy Employees.}
+
+An Employee that is frustrated with his situation for a long time
+period constitutes a security vulnerability. On the one hand he might
+want to harm his employer directly, on the other he is increasingly
+susceptible for social engineering.
 
@@
 
-% Extend to general missing privacy wareness of citizen, system provider and local authority
+%TODO: Extend to general missing privacy wareness of citizen, system provider and local authority
 \textbf{Unaware Monitoring Subjects.}
-We define privacy as one's ability to control information about oneself.
-In order to do that, monitored subjects need to know, that they are monitored, who monitors them, what information is recorded and for what purposes.
-Subjects who are not aware of these things cannot effectively preserve control and thus lose their privacy.
-
-%TODO
-- No Privacy Policy
-- No Information about Processing, Access rules, 3rd party disclosure
+We define privacy as one's ability to control information about
+oneself.  In order to do that, monitored subjects need to know, that
+they are monitored, who monitors them, what information is recorded
+and for what purposes.  Subjects who are not aware of these things
+cannot effectively preserve control and thus lose their privacy.  This
+vulnerability expresses itself as the lack of information material
+like a Privacy Policy including concise information about applied
+processing steps, access rules and disclosure to 3rd parties.
 
@@
 %---------------
-%Description:   External infiltrates infrastructure in order to spy on citizen
 %COI:           Privacy vs. External Power
@@
 %
-\textbf{Surveillance}
-Lorem Ipsum
+\textbf{Surveillance} External infiltrates infrastructure in order to
+obtain direct about the citizen. In this scenario the external is
+supposed to have some direct relationship to the citizen which
+motivates his interest to obtain personal information. Examples could
+be a public institution that wants to gain information about planned
+activities of the citizens (e.g. Nixon's Watergate scandal or the
+recent prosecution of Guardian journalists by GHCQ). Another example
+is an insurance company that seeks to get information about the
+citizens life-style in relation to the insured risk, like car
+accidents or health hazards.
+
+In this threat the privacy interest of the citizen is in conflict with
+the aspirations for power over the citizen by the externals.
 
@@
 %-------------------
-%Description: Like an external person in Scenario (2) an employee of the service provider has selfish interests to gain money, make political statements or harm his employer.
-%In order to pursue this interest he can steal personal data and sell it or release it to the public.
-%
 %COI:         Privacy vs. Financial Profit
@@
 %
+
 \textbf{Information Leak}
-Lorem Ipsum
+Like an external person in Scenario (2) an employee of the service provider has selfish interests to gain money, make political statements or harm his employer.
+In order to pursue this interest he can steal personal data and sell it or release it to the public.
 
@@
 %---------------------
-%Description: An External can manipulate Local Authorities or Service Providers to give them personal information that he can then sell or disclose.
-%
+%Description: 
 %COI:         Privacy vs. Financial Profit
@@
 \textbf{Social Engineering}
-Lorem Ipsum
+An External can manipulate Local Authorities or Service Providers to
+give them personal information that he can then sell or disclose.
 
@@
 For each possible data type (e.g. GPS) we analyze the impact on the seven different types of privacy in Section \ref{sec:SensorPrivacyImpact}.
-In combination we can quantify roughly the impact of each threat on the citizens privacy.
-
-Both evaluations are necessarily fraught with a high level of uncertainty.
-
-Loss Calibration
+In combination we can quantify roughly the impact of each threat on the citizens privacy. Both evaluations are necessarily fraught with a high level of uncertainty.
+
+For the quantification of the loss in case of a threat scenario we use the following rough calibration:
 \begin{itemize}
@@
 
-Likeliness is measured on a the following scale
-\begin{itemize}
+For the quantification of likeliness the following scale is used:
+\begin{itemize}
+\item 4: always the case.
+\item 3: maybe occure once in 10
+\item 2: maybe occure once in 100
+\item 1: maybe occure once in 1 million cases.
 \item 0 (impossible)
-\item 1: maybe occure once in 1 million cases.
-\item 2: maybe occure once in 100
-\item 3: maybe occure once in 10
-\item 4: always the case.
-\end{itemize}
-The risk scale is defined as addition of the loss and likeliness scales.\footnote{
-Note, that loss and likeliness scales have a logarithmic character, so that that addition of those scales corresponds to multiplication of the usual scales.
-}
+\end{itemize}
+
+The quantification of the risk, we add the values for loss and
+likeliness of the corresponding threats. Note, that loss and
+likeliness scales have a logarithmic character, so that that addition
+of those scales corresponds to multiplication of the usual scales.
 
@@
 
-\subsubsection{Security Requirement Specification}
-%TODO SKIP?
-TODO: Formulate requirements that address the high risk scenarios.
-
-%\subsection{Step 3. Plan Development}
-
-% \subsection{Legal Privacy Requirements} \ref{sec:LEGAL_REQ}
-
-%\subsubsection{Security Measures}
-
-% Seven Cs are now in separate Document
+\subsubsection{Security Recommendations and Requirements}
+\textbf{Legal Requirements}
+
+Privacy Policy
+
+Processing Journal
+\begin{itemize}
+\item name of the controller
+\item purpose of processing
+\item description of the data categories
+\item recipients of the data if disclosed
+\item transfer to third countries
+\item description of security of processing
+\end{itemize}
+The journal is available for anybody to access via Email.
+Reporting to Data Protection Authority (Bundesamt für Datenschutz).

### 2014-11-06 D1-3.tex
Added link between Recommendations and Implementations.

--- 2014-11-05 D1-3.tex
@@
 attended). Accelerometer data can be used to infer medical conditions
-like walking disabilities.
-
-Images can contain faces of nearby persons to with whom the citizen is
-associated. All this data is highly sensitive to the citizens privacy
-and can be used against the citizen if it falls in the wrong hands.
+like walking disabilities. Images can contain faces of nearby persons
+to with whom the citizen is associated. All this data is highly
+sensitive to the citizens privacy and can be used against the citizen
+if it falls in the wrong hands.
 
@@
 data. The most important legislation on the European Level is the Data
-Protection Directive 95/46/EC [1], which is also summarized in the
-European Charta of Fundamental Rights [2] of EU citizens:
+Protection Directive 95/46/EC (cf. Section \ref{sec:EUDIR}), which is
+also summarized in the European Charta of Fundamental Rights
+(cf. Section \ref{sec:FR}) of EU citizens:
 
@@
 
-Live+Gov Privacy Dashboard gives citizens back the Control
+\input{PrivacyProtection-2014-09-09}
+
+\clearpage
+
+\input{Implementations}
+
+\chapter{Summary}\label{chap:Summary}
+
+* Discussed Legal and ethical aspects of privacy
+* Made Privacy Security Analysis of Live+Gov System
+* Presented several Privacy Protection Measures that have been implemented.
+* Most importantly the Privacy Dashboard, which gives the citizen control over their data.
 

### 2014-11-06 Implementations.tex
Added link between Recommendations and Implementations.

@@
 \chapter{Implementation of Privacy Protection Measures}
+\label{chap:impl}
+
+In this chapter we explain the measures we have taken in order to
+protect the citizens privacy. We have used the recommendations derived
+from the Security Analysis performed in Chapter \ref{chap:analysis} as
+a guideline.
+
+\section{Privacy Dashboard}
+The Live+Gov Privacy Dashboard is a software component that allows the
+citizens to view and control the data about themselves that is stored
+in the Sensor Data Storage server. It implements privacy
+recommendation R1 of Section \ref{sec:prec}.  The foundation of the
+Privacy Dashboard component is the inspection tool that was introduced
+in D1.1 as a means to perform basic quality controlling of the
+collected data.
+
+The Privacy Dashboard is currently endowed with the following views:
+\begin{itemize}
+  \item \textbf{Login Screen} (cf. Figure \ref{fig:PDLogin}).  Before
+    access to the dashboard is granted, the user is prompted for
+    credentials.  The credentials are generated by the sensor
+    collection application on the mobile device.  The username can be
+    freely chosen by the user. The password is then generated by the
+    mobile application and transferred to the server alongside with
+    the recording.
+
+ \item \textbf{Recording Overview} (Cf. Figure \ref{fig:PDOverview}).
+   This view shows all recordings that the user submitted to the Data
+   Storage Service including basic meta information like start data of
+   recording, duration. The user can click on each recording in order
+   to view the recorded data in the Raw Data View.
+
+   An important feature of the recording overview table is the delete
+   button (cross on the very left.) That allows the user to
+   selectively delete recordings, that shall not be analyzed by the
+   Live+Gov system. Internally the click on the delete button only
+   flags the corresponding trip for deletion. The actual deletion is
+   carried out as a batch job on a regular bases. This addition step
+   prevents accidental loss of data.
+
+ \item \textbf{Raw Data View} (cf. Figure \ref{fig:PDRawData}).
+   The raw data view shows all recorded raw data to the user.  GPS
+   data is visualized on a map. Motion sensor data is displayed as
+   plots. Wifi and GSM data is displayed in lists.  The raw data from
+   each sensor can be downloaded individually as CSV file.
+
+ \item \textbf{Processing View} (cf. Figure \ref{fig:PDProcessedData}).
+   The processing view shows the data mining end-products for each
+   processing step.  The Figure shows the processing view for the
+   Activity Recognition. On the top a time-line is show that
+   visualizes the different recognized activities via color-coding. A
+   map below displays the corresponding GPS track with the same color
+   coding.
+
+   A similar view for Service Line Detection results is currently
+   under development.
+ \item \textbf{Privacy Settings.}
+   This view contains the privacy policy shown in Section \ref{sec:PP}.
+
+   We intend to include settings also settings concerning the
+   expiration date of the recordings, agreements to certain processing
+   steps, view logs of performed processing.
+\end{itemize}
+
+The Privacy Dashboard is implemented on a
+nodes.js\footnote{\url{http://nodejs.org}} web server using
+Express.js\footnote{\url{http://expressjs.com/}} as a web-framework
+and d3.js\footnote{\url{http://d3js.org/}} as plotting library. The
+data is pulled directly from the PostgreSQL database of the Data
+Storage Service.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{screenshots/login.png}
+\caption{Live+Gov Privacy Dashboard Login}
+\label{fig:PDLogin}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{screenshots/recordings.png}
+\caption{Live+Gov Privacy Dashboard Recording Overview}
+\label{fig:PDOverview}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{screenshots/raw.png}
+\caption{Live+Gov Privacy Dashboard Raw Data View}
+\label{fig:PDRawData}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{screenshots/processed.png}
+\caption{Live+Gov Privacy Dashboard Processed Data View}
+\label{fig:PDProcessedData}
+\end{figure}
+
 \section{Privacy Policy} \label{sec:PP}
-
-In complience legal requirements of the Data Protection Directive (cf. Section \ref{EUDIR}) we include the following
-Privacy Policy on our Mobile Application and Website.
+Implementing Privacy Recommendation R1 of Section \ref{sec:prec} and
+following legal requirements of the Data Protection Directive
+(cf. Section \ref{EUDIR}) we include the following Privacy Policy on
+our Mobile Application and Website.
 
@@
 
-\section{Privacy Dashboard}
-The Live+Gov Privacy Dashboard is a software component that allows the
-citizens to view and control the data about themselves that is stored
-in the Sensor Data Storage server.  The foundation of the Privacy
-Dashboard component is the inspection tool that was introduced in D1.1
-as a means to perform basic quality controlling of the collected data.
-
-The Privacy Dashboard is currently endowed with the following views:
-\begin{itemize}
-  \item \textbf{Login Screen} (cf. Figure \ref{fig:PDLogin}).  Before
-    access to the dashboard is granted, the user is prompted for
-    credentials.  The credentials are generated by the sensor
-    collection application on the mobile device.  The username can be
-    freely chosen by the user. The password is then generated by the
-    mobile application and transferred to the server alongside with
-    the recording.
-
- \item \textbf{Recording Overview} (Cf. Figure \ref{fig:PDOverview}).
-   This view shows all recordings that the user submitted to the Data
-   Storage Service including basic meta information like start data of
-   recording, duration. The user can click on each recording in order
-   to view the recorded data in the Raw Data View.
-
-   An important feature of the recording overview table is the delete
-   button (cross on the very left.) That allows the user to
-   selectively delete recordings, that shall not be analyzed by the
-   Live+Gov system. Internally the click on the delete button only
-   flags the corresponding trip for deletion. The actual deletion is
-   carried out as a batch job on a regular bases. This addition step
-   prevents accidental loss of data.
-
- \item \textbf{Raw Data View} (cf. Figure \ref{fig:PDRawData}).
-   The raw data view shows all recorded raw data to the user.  GPS
-   data is visualized on a map. Motion sensor data is displayed as
-   plots. Wifi and GSM data is displayed in lists.  The raw data from
-   each sensor can be downloaded individually as CSV file.
-
- \item \textbf{Processing View} (cf. Figure \ref{fig:PDProcessedData}).
-   The processing view shows the data mining end-products for each
-   processing step.  The Figure shows the processing view for the
-   Activity Recognition. On the top a time-line is show that
-   visualizes the different recognized activities via color-coding. A
-   map below displays the corresponding GPS track with the same color
-   coding.
-
-   A similar view for Service Line Detection results is currently
-   under development.
- \item \textbf{Privacy Settings.}
-   This view contains the privacy policy shown in Section \ref{sec:PP}.
-
-   We intend to include settings also settings concerning the
-   expiration date of the recordings, agreements to certain processing
-   steps, view logs of performed processing.
-\end{itemize}
-
-The Privacy Dashboard is implemented on a
-nodes.js\footnote{\url{http://nodejs.org}} web server using
-Express.js\footnote{\url{http://expressjs.com/}} as a web-framework
-and d3.js\footnote{\url{http://d3js.org/}} as plotting library. The
-data is pulled directly from the PostgreSQL database of the Data
-Storage Service.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{screenshots/login.png}
-\caption{Live+Gov Privacy Dashboard Login}
-\label{fig:PDLogin}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{screenshots/recordings.png}
-\caption{Live+Gov Privacy Dashboard Recording Overview}
-\label{fig:PDOverview}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{screenshots/raw.png}
-\caption{Live+Gov Privacy Dashboard Raw Data View}
-\label{fig:PDRawData}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{screenshots/processed.png}
-\caption{Live+Gov Privacy Dashboard Processed Data View}
-\label{fig:PDProcessedData}
-\end{figure}
-
-\pagebreak
-
 \section{Transfer Encryption (HTTPS/SSL)}
@@
 communication channels, we secured all HTTP connections with SSL
-encryption.
+encryption. This measure implements the security recommendation R7
+Securing Communiction Channels of Section \ref{sec:prec}.
 
@@
 \end{figure}
+
+\section{Security of Infrastructure}
+The servers hosting the Sensor Data Storage component have been
+secured using the following security measures:
+\begin{itemize}
+  \item Firewall. A Firewall restricts access to the minimal necessary
+    ports allowing upload and inspection of the data as well, database
+    access by our consortium partners and maintenance over SSH.
+  \item Security Upgrades.  The operating system (Ubuntu Linux 12.04)
+    is configured to install security updates of all installed
+    packages on a daily basis.
+  \item Access Rules. Access to the System (via SSH) and the database
+    is restricted to a minimal amount of persons. Those persons have
+    been advised to use strong passwords.
+\end{itemize}
+These measures are a basic implementation of recommendation R6.
 
@@
 
-\section{Anonymization of trips}
+\section{Pseudonymization of trips.}
 The Sensor Data Storage Service follows a privacy by design guideline
 in that it does not store personal information like names and email
-address at all.
+address at all, only pseudonyms in the form of user-ids.
 
@@
 shall be anonymized.
+
+This measure implements part of the security recommendation R5
+Anonymize personal data of Section \ref{sec:prec}.
 
@@
 
+This measure implements part of the security recommendation R5
+Anonymize personal data of Section \ref{sec:prec}.
+
 \begin{figure}

### 2014-11-06 PrivacyProtection-2014-09-09.tex
Added link between Recommendations and Implementations.

@@
 
-\subsection{EU Charter of Fundamental Rights}
+\subsection{EU Charter of Fundamental Rights}\label{sec:FR}
 The Treaty of Lisbon of 2007, which introduced the fundamental functioning of the European Union
@@
 \chapter{Privacy Analysis of Live+Gov Systems}
+\label{chap:analysis}
 
@@
 
-%TODO: Update Threat Table
+% \label{figure:Threats}
 \input{./figures/Threats}
 
-%0. Insufficient Control Features
-%-------------------
-%Description:    The Service Provider does not offer tools for the citizen to control his data.
-%COI:            Privacy vs. System Complexity
-%Vulneratbility: missing privacy wareness
-%Asset:          All collected data
-%
-%Likeliness: Always
-%Loss: Low (Control is lost, no direct harm done)
-%Risk:      5
-%
-\textbf{Insufficient Control Features.}
+\textbf{T1. Insufficient Control Features.}
 The System Provider does not offer tools for the Citizen to control his data.
@@
 
-%1. Excessive Data Mining
-%------------------------
-%Description: The Service Provider and/or the Local Authority extract more private information from the collected data, than the citizen agreed to.
-%E.g. the Service Provider wants to test a new product, and uses old data collected for a different purpose.
-%E.g. the Local Authority wants to analyze collected data from fare evasions.
-%
-%COI: Privacy vs. Buisness Intelligence of Local Authority, Privacy vs. Financial Profit of Serivce Provider.
-%Vuln.: Lax data handling policies; Unaware citizen; Weak law enforcement
-%Affected Data: Potential Information extracted from raw data.
-%
-%Likeliness: High
-%Loss:      Medium (Control is lost, no direct harm done)
-%Risk:      5
-%
-\textbf{Excessive Data Mining.}
+\textbf{T2. Excessive Data Mining.}
 The System Provider and/or the Local Authority secretly extract more private information from the collected data, than the Citizen agreed to.
@@
 
-%2. Data Theft
-%-------------
-%Description: External infiltrates infrastructure in order to steal personal data and sell it on the black market.
-Also the external might be motivated politically and wants to harm the reputation of the Service Provider or the Local Authority.
-%
-%COI:         Privacy vs. Financial Profit
-%             Reputation of Service Provider vs. Political Activism
-%             Reputation of Local Authority vs. Political Activism
-%Vuln.:       Insecure Infrastructure, Insecure Communication
-%Affected Data:  Data transmitted or stored in data center: GPS trips, HAR-records.
-%
-%Likeliness: Low-Medium
-%Loss: High
-%Risk: 4-5
-%
-\textbf{Data Theft}
+\textbf{T3. Data Theft}
 An External infiltrates infrastructure in order to steal personal data and sell it on the black market.
@@
 Thus this threat is defined by three conflicts: \textit{Privacy vs.
-Financial Profit}, \textit{ Reputation of Service Provider vs.
+Financial Profit}, \textit{Reputation of Service Provider vs.
 Political Activism} and \textit{Reputation of Local Authority vs.
@@
 
-%3. Surveillance
-%---------------
-%COI:           Privacy vs. External Power
-%Vulnerability: Insecure Infrastructure, Insecure Communication
-%Affected Data:         Data transferred and stored in data center.
-%
-%Likeliness: Medium (Government*), Low (Third parties)
-%Loss: High (Control is lost, potential harm by due to Government action)
-%Risk:      6/4
-%
-%* Government, has also other tools at it's disposal
-%
-%
-\textbf{Surveillance} External infiltrates infrastructure in order to
-obtain direct about the citizen. In this scenario the external is
-supposed to have some direct relationship to the citizen which
-motivates his interest to obtain personal information. Examples could
-be a public institution that wants to gain information about planned
-activities of the citizens (e.g. Nixon's Watergate scandal or the
-recent prosecution of Guardian journalists by GHCQ). Another example
-is an insurance company that seeks to get information about the
-citizens life-style in relation to the insured risk, like car
+\textbf{T4. Surveillance}
+
+An External infiltrates infrastructure in order to obtain information
+about the citizen and exploit it directly.  In this scenario the
+external is supposed to have some direct relationship to the citizen
+which motivates his interest to obtain personal information.  Examples
+could be a public institution that wants to gain information about
+planned activities of the citizens (e.g. Nixon's Watergate scandal or
+the recent prosecution of Guardian journalists by GHCQ).  Another
+example is an insurance company that seeks to get information about
+the citizens life-style in relation to the insured risk, like car
 accidents or health hazards.
@@
 
-%4. Information Leak
-%-------------------
-%COI:         Privacy vs. Financial Profit
-%             Reputation of Service Provider vs. Political Activism
-%             Reputation of Local Authority vs. Political Activism
-%Vuln.:       Unhappy Employee, Lax Access Rules.
-%Affected Data:       Data stored in data center.
-%
-%Likeliness: Low
-%Loss: High (public leak)
-%Risk: 4
-%
-
-\textbf{Information Leak}
-Like an external person in Scenario (2) an employee of the service provider has selfish interests to gain money, make political statements or harm his employer.
-In order to pursue this interest he can steal personal data and sell it or release it to the public.
-
-%5. Social Engineering
-%---------------------
-%Description: 
-%COI:         Privacy vs. Financial Profit
-%             Reputation of Service Provider vs. Political Activism
-%             Reputation of Local Authority vs. Political Activism
-%Vuln.:       Unhappy Employee, Lax Access Rules.
-%Affected Data:       Data stored in data center.
-%
-%Likeliness: Low
-%Loss: High
-%Risk: 4
-%
-\textbf{Social Engineering}
-An External can manipulate Local Authorities or Service Providers to
-give them personal information that he can then sell or disclose.
+\textbf{T5. Information Leak}
+
+Like an external person the Data Theft Scenario an employee of the
+service provider or the Local Authority has selfish interests to gain
+money, make political statements or harm his employer.  In order to
+pursue this interest he can steal personal data and sell it or release
+it to the public.  The corresponding conflicts of interests are:
+\textit{Privacy vs. Financial Profit} of the Employee, \textit{Reputation of Service
+  Provider vs. Political Activism} of the Employee and \textit{Reputation of Local
+  Authority vs. Political Activism} of the Employee.  The vulnerability constitutes of
+the existence of \textit{Unhappy employees} itself and possibly \textit{lax access
+rules} that enable the employee to obtain large amounts of data
+unnoticed.
+
+\textbf{T6. Social Engineering}
+
+This scenario an external manipulates an employee of a Service
+Provider or the Local Authority to leak information to the external
+person.  It is thus combination of the Data Theft and Information Leak
+scenario.
+The conflicts of interest are \textit{Privacy vs. Financial Profit} of
+the External, \textit{Reputation of Service Provider vs. Political
+  Activism} of the external and \textit{Reputation of Local Authority
+  vs. Political Activism} of the external.  The exploited
+vulnerabilities are, again, the existence of \textit{Unhappy
+  employees} and possibly \textit{lax access rules} that enable the
+employee to obtain large amounts of data unnoticed.
 
@@
 \begin{itemize}
-\item 3: High. Leak of information to Peers which impacts citizen.
-  E.g. spouse gets GPS data and finds out cheating.
-  Subjects is added to No-Fly list due to Government spying.
-  Public leak of private information.
+\item 3: High. Leak of information to peers (e.g. public) which impacts citizen.
 \item 2: Medium. Undisclosed processing of personal data or disclosure to third parties that are unrelated to subject.
@@
 \begin{itemize}
-\item 4: always the case.
-\item 3: maybe occure once in 10
-\item 2: maybe occure once in 100
-\item 1: maybe occure once in 1 million cases.
-\item 0 (impossible)
+\item 4: Always.
+\item 3: High. Occurs once in 10 cases
+\item 2: Medium. Occurs once in 100 cases
+\item 1: Low. Occurs once in 1 million cases
+\item 0: Impossible
 \end{itemize}
@@
 
-%TODO:
-Insert Table with Likeliness / Loss / Risk Assesment.
-
-\subsubsection{Security Recommendations and Requirements}
-\textbf{Legal Requirements}
-
-Privacy Policy
-
-Processing Journal
-\begin{itemize}
-\item name of the controller
-\item purpose of processing
-\item description of the data categories
-\item recipients of the data if disclosed
-\item transfer to third countries
-\item description of security of processing
-\end{itemize}
-The journal is available for anybody to access via Email.
-Reporting to Data Protection Authority (Bundesamt für Datenschutz).
+The likeliness, loss and the resulting risks assigned to the threats
+are discussed in the following paragraphs and summarized in Figure
+\ref{fig:risks}.
+
+\textbf{T1. Insufficient Control Features.}  The occurrence of this
+threat is dependent on the design on the system and given in our case,
+since we do not give the citizen control over his data once it is
+recorded. Therefore the Likeliness is evaluated as $4$ (Always).  The
+associated, risk is $1$ Low on our scale, since no direct harm is done to
+the citizen by exploiting the data.
+
+Hence the resulting risk is calculated as $4+1 = 5$.
+
+\textbf{T2. Excessive Data Mining.}
+We assess the likeliness of excessive data mining to be $3$ High, since
+these kind of analysis can be performed within the walls of the service
+provider, without somebody else noticing, and the service provider himself
+has an interest in this activity.
+
+The associated loss, on the other hand can be substantial (Medium
+$2$).  For example when GPS data is linked to data from telephone
+books the identity of the citizen can be revealed and personal details
+like visits to doctors. In the threat scenario, this information is
+not leaked to third parties, (which would justify an even higher loss
+assessment), but the very existence of this information violates the
+citizens privacy.
+
+Hence the resulting risk is calculated as $3+2 = 5$.
+
+\textbf{T3. Data Theft.}
+The likeliness of a targeted attack by a third party is dependent on
+the popularity of the offered service and financial value of the
+captured information. Moreover, the amount of manual work required to
+infiltrate a custom build system is significantly higher that that of
+compromising a standard software solution. In the scenario we assume a
+moderate popularity in a single metropolitan area, with around 10.000
+users and storage of data of only limited financial value (no
+addresses, no payment information). Therefore the likeliness
+assessment is $1-2$ (Low-Medium).
+
+The harm of leaked information to a criminal party is $3$ High.
+Hence the resulting risk is calculated as $4-5$.
+
+\textbf{T4. Surveillance.}  In the surveillance scenario an party
+related to the citizen, like a company where he is customer of, or a
+government agency, seeks to obtain sensitive information from our
+service.
+
+The likeliness of such an intrusion is hard to asses, and depends
+again on the popularity of the service. If a high popularity is
+reached we have recently learned that spying by government agencies is
+very likely to occur. The barrier for companies that do not operate
+the infrastructure used to transmit the data a surveillance attack is
+however very hard to perform. Therefore we assess the likeliness of
+the threat with $1-2$ (Low-Medium).
+
+The harm of leaked information to a related party is $3$ High.
+Hence the resulting risk is calculated as $4-5$.
+
+\textbf{T5/6. Information Leak and Social Engineering.}
+
+In our scenario we assume that the culture and ethics inside the
+service provider company and local authority are very high, so that
+the information leak scenario has a likeliness of $1$ (Low).
+
+The harm of such an information leaked is $3$ High, so that the
+resulting risk is calculated as $4$.
+
+\begin{figure}
+\centering
+\begin{tabular}{|l|l|l|l|l|}
+\hline
+\textbf{Threat}                   & \textbf{Likeliness} & \textbf{Loss} & \textbf{Risk} & \textbf{Recommendation}
+\\\hline
+T1. Insufficient Control Features & 4                   & 1             & 5             & R1, R2
+\\\hline
+T2. Excessive Data Mining         & 3                   & 2             & 5             & R3, R4, R5
+\\\hline
+T3. Data Theft                    & 1 - 2               & 3             & 4 - 5         & R6
+\\\hline
+T4. Surveillance                  & 1 - 2               & 3             & 4 - 5         & R7
+\\\hline
+T5. Information Leak              & 1                   & 3             & 4             & R8
+\\\hline
+T6. Social Engineering            & 1                   & 3             & 4             & R8
+\\\hline
+\end{tabular}
+\caption{Live+Gov Risk Evaluation and Recommendations}
+\label{fig:risks}
+\end{figure}
+
+\subsubsection{Privacy Recommendations}
+\label{sec:prec}
+
+In the preceding section we have identified the main risks for the
+users privacy.  In this section we derive recommendations or
+requirements for a system that addresses these risks.  Some of these
+requirements are implemented as security measures in our systems and
+discussed in the following chapter \ref{chap:impl}.
+
+In order to address the threat with the highers risk, Insufficient
+Control (T1) of the citizen, we need to give the citizen back the
+control over its data inside the system. The most direct way to do
+this is to provide a web-based \emph{Privacy Dashboard (R1)} which allows
+the citizen to view, edit and delete all information about his person
+that is stored inside the system. Also control applied processing and
+disclosure of the data to third parties should be given to the user,
+at least in the form of an opt-out or veto option.
+
+A necessary pre-requirement for effective control of the citizen over
+his data is information and comprehension of the intended data
+capturing and processing steps. Therefore a \emph{Privacy Policy (R2)} that
+is easily readable and contains all important information is
+essential.  Moreover, the existence of a Privacy Policy is a legal
+requirement (cf. Section \ref{EUDIR}).
+
+The threat with the second largest risk is (T2) Excessive Data Mining.
+Contrary to common belief, it is neither legal nor ethical to process
+personal data for by new methods or for new purposes that were not
+stated and explained to the citizen at the time of data
+collection. Also the common practice of obtaining far-reaching
+permissions from the citizens inside the privacy policy is neither an
+ethical or legal solution to the problem (cf. Art. 6 in Section
+\ref{EUDIR}).
+
+To address this threat awareness about the limitations of data
+processors inside the company is a key element. As one mean to
+establish such a culture of privacy respect, we recommend to prepare
+an document called \emph{Data Handling Guidelines (R3)} intended for
+internal use that explains the concrete processing steps and purposes
+that are permitted by the citizens. This guideline should also be
+structured in a way that it covers the legal notification requirement
+from the EU Data Protection Directive (cf. Section \ref{EUDIR}). In
+particular the following information should be provided for each
+processing task: The name of the controller, purpose of processing,
+description of the data categories, recipients of the data if
+disclosed, transfer to third countries and a description of security
+of processing.
+
+If further processing should be performed, it is necessary to seek
+additional permissions from the citizen. A simple email explaining the
+planned processing steps, and \emph{asking for permission (R4)} would be
+enough for this purpose. The permission can be given via an embedded
+link that shall be followed in order to signal agreement.
+
+An alternative measure to address the risk of excessive data mining is
+the \emph{anonymization (R5)} of data. When all direct- or indirect
+links to the identity of the person are removed, no violation of the
+citizens privacy caused by arbitrary processing. However removing all
+such links is a challenging tasks, and full anonymity is often not
+achieved, cf. \cite{krumm2009}.
+
+The protection from threat scenario (T3) Data Theft is a case of
+classical \emph{IT infrastructure security (R6)}. The storage and processing infrastructure has
+to be secured using firewalls, up-to data software versions and proper
+authentication mechanisms.
+
+The protection from threat scenario (T4) Surveillance focuses on the
+communication channels. They are target of wiretapping attacks by
+intermediaries or externals with access to the communication
+infrastructure. Strong \emph{encryption (R7)} should be used to make
+it harder for externals to read the content of the transmitted data.
+
+Threat scenarios (T5) Information Leak and (T6) Social Engineering
+target the vulnerability of unhappy employees. Therefore a trustful,
+\emph{healthy company culture (R8)} should be maintained.
+
+In summary we recommend the following measures to secure the citizens
+privacy:
+\begin{enumerate}
+\item[R1] Privacy Dashboard. A tool which allows the citizen to view, edit
+  and delete all data personal data that is stored in the system.
+\item[R2] Privacy Policy. A document, that informs the citizen about the
+  collection and processing of personal information. It should at
+  least contain the legally required information.
+\item[R3] Issue Data Handling Guidelines that explain the permitted
+  processing methods and purposes.
+\item[R4] Ask the citizens for permissions before applying further
+  processing via Email.
+\item[R5] Anonymize personal data before processing.
+\item[R6] Securing of Storage and Processing infrastructure using e.g. firewalls.
+\item[R7] Securing communication channels using encryption.
+\item[R8] Maintain a healthy, trustful relationship with your employees.
+\end{enumerate}
+The mapping of these recommendations to the threats is summarized in Figure \ref{fig:risks}.

### 2014-11-07 D1-3.tex
Updated intro and summary.

--- 2014-11-05 D1-3.tex
@@
 
+This deliverable presents the privacy aware, server based reality
+sensing infrastructure for Live+Gov applications.  Privacy protection
+is a crucial point for public acceptance of advanced data mining
+techniques in eParticipation scenarios.  In order to provide such
+protection we perform an extensive analysis of all threats to privacy
+and implement a number of measures that address these threats in an
+adequate way.
+
+The term "privacy'' itself is highly ambiguous and there is no common
+understanding what privacy really is in the literature.  In
+Chapter \ref{chap:privacy} we discuss different legal and
+philosophical texts and about the subject.  The main relevant
+legislation is the EU Data Protection Directive (cf. Section
+\ref{EUDIR}). The main ethical treatment is provided by Charles Fried
+(cf. Section \ref{sec:ethics}), who defines privacy as the
+\emph{control over personal information}. Furthermore, seven concrete
+privacy types are introduced (cf. Section \ref{sec:privacytypes}) and
+the relation to collected sensor data from mobile phones is made.
+
+On the basis of these foundations perform a thorough analysis of all
+risks for the privacy of a citizen. The main difficulty we are
+presented with is the selection of sensible measures. There are a lot
+of anonymization techniques in the literature and it is easy to jump
+ahead and implement a few of them while leaving out massive privacy
+gaps at other ends. Therefore a structured approach to this problem
+was taken.
+
+We have chosen to use the IT Security Reference Model
+\cite{Grimm:ItSecRefModel} (cf. Section \ref{sec:GrimmModel}), that
+was developed by our colleague from Prof. Grimm from University of
+Koblenz, as a framework to conduct a systematic analysis of all risks
+that are posed to the privacy of the citizens. The analysis is carried
+out in Section \ref{sec:PPA}. Prof. Grimm also contributed with his
+expertise and guidance to the execution of this analysis.
+
+The analysis starts by describing an abstract version of the IT
+Systems and the involved actors with their interests. We then describe
+vulnerabilities of the IT Systems and conflicts between the interests
+of the stakeholders. All threats that are posed to the privacy of the
+citizen are caused by a conflict of interest between two actors and
+exploit a vulnerability of the system.
+
+We identity six main threats that include ``T1. Insufficent Control
+Features'', ``T2. Excessive Data Minig'' and ``T4. Surveillance'' and
+asses their risk of occurrence. The biggest threat to the citizens
+privacy is posed by Insufficient Control Features of the system. As
+soon as collected data of Citizens is stored on the servers, all
+control over that data is lost if the service provider does not offer
+tools for the citizen to control his data.
+
+These risks are addressed by eight recommendations, that when
+implemented provide an adequate protection of the citizens privacy.
+One surprising outcome, is that the main privacy threats can be
+eliminated by enhancing communication with the citizen and
+establishing policies and awareness for privacy aspects inside the
+company. Thus, a central point in the protection of the citizens privacy is
+a ``Privacy Dashboard'' that gives the citizen a maximum of control
+over his data. Using this dashboard he is able to view and export all
+data from the citizen that is currently stored in the Live+Gov system,
+selectively delete parts of the stored data and get information about
+processing applied and view the processing results themselves.
+
+In addition, we have implemented four different anonymization methods
+for GPS samples, several security mechanisms for infrastructure and
+communication channels, moreover we wrote a concise and comprehensive
+privacy policy that informs the citizen about all processing steps.
+
 The deliverable is accompanied with source code of the components in
-  Java, API documentation (javadoc), and pre-compiled packages for
-  direct installation and testing on mobile devices.
+Java, Python and JavaScript, documentation, and pre-compiled packages
+of the executable parts.
 
@@
 regulate the collection, processing and use of privacy sensitive
-data. The most important legislation on the European Level is the Data
-Protection Directive 95/46/EC [1], which is also summarized in the
-European Charta of Fundamental Rights [2] of EU citizens:
-
-\begin{quote}
-Article 7. Respect for private and family life.
-
-Everyone has the right to respect for his or her private and family
-life, home and communications.
-
-Article 8. Protection of personal data.
-
-(1) Everyone has the right to the protection of personal data
-concerning him or her.
-
-(2) Such data must be processed fairly for specified purposes and on
-the basis of the consent of the person concerned or some other
-legitimate basis laid down by law.  Everyone has the right of access
-to data which has been collected concerning him or her, and the right
-to have it rectified.
-
-(3) Compliance with these rules shall be subject to control by an
-independent authority.  The Live+Gov fully consortium complies with
-this legislation and is moreover committed to convey and enact the
-ethical values inherent in these texts.
-\end{quote}
-
-Privacy as Control over Private Data
+data. We explain the most relevant legislation in Section
+\ref{sec:legal}.
 
 The ethical aspects of privacy have been the subject of study of many
-social scientists and philosophers [3]. Many researchers do focus on
-the ways in which privacy can be infringed. Thus they invest a great
-amount of work in exploring threats, instead of describing why privacy
-is so valuable to us. One scholar which follows an alternative aproach
-in the context of digital monitoring is Charles Fried [4]. He
-investiages, why we are intuitively so sensitive to violations of our
-privacy. For him privacy is not asserted as an intrinsic value by
-itself, he rather sated:
+social scientists and philosophers. One scholar which is particularly
+relevant in our context is Charles Fried. He investiages, why we are
+intuitively so sensitive to violations of our privacy. For him privacy
+is not asserted as an intrinsic value by itself, he rather sated:
 
@@
 
-Depending on the conversation partner, we change the degree of
-information we share if it is a total stranger, a colleague, close
-friend or a doctor. With doctors we share information of great
-intimacy we do not share with anyone else.
+Fried`s study on the understanding of privacy provided a great
+contribution to the research on the same term in philosophy and
+computer science and despite the fact his text was published in 1970,
+he already included technologies to its viewpoint (like location
+monitoring) that are particularly relevant to our Context.  We explain
+his theory in \ref{sec:ethics} and follow his definition of privacy in
+this document.
 
-Moreover, we trust those friends to not reveal information about us to
-others by respecting their privacy. Trust needs the possibility of
-unknown failure. If we would constantly monitor our partners, they
-cannot fail unnoticed nor can they willingly share that information
-with us.  So privacy, according to Fried, is the foundation of our
-core relations. And thus it is valuable, because those relations are
-essential to human society.
+In this document we perform a thorough analysis of how to protect the
+privacy of a citizen and present several implementations that form the
+core of our privacy aware Sensor Data Storage and Mining Infrastructure.
 
-Fried’s study on the understanding of privacy provided a great
-contribution to the research on the same term in philosophy and
-computer science [5] and despite the fact his text was published in
-1970, he already included technologies to its viewpoint (like location
-monitoring) that are particularly relevant to our Context.
-
-Live+Gov Privacy Dashboard gives citizens back the Control
-
-In order to meet the the legal requirements and offer the citiens to
-keep a maximum of contol about their personal data, Live+Gov systems
-implements a number of privacy protection measure. One of the measures
-that is under current development is a Privacy Dashboard (Figure
-1). Using this dashboard the citizen will be able to take the
-following actions: View and export all data from the citizen that is
-currently stored in the Live+Gov system Selectively delete parts of
-the stored data Get information about processing applied to this data
-Consent or decline selectively to processing steps
-
-View the end poducts of data mining applied to his data in a
-comprehensive, graphical way These measures provide the citizen with a
-maximal respect for their privacy while at the same time allowing
-advanced data processing. Although, abusive exploitation of the data
-is still possible, the level of transparency offered by the dashboard
-builds trust between the citizen and the service provider.
+We identify six main threats for the citizens privacy, and derive
+eight recommendations that should be followed in order to reduce the
+associated risks for hazards.  This analysis form the guideline for
+the selection of 9 measures that were implemented in our system and
+documented in Chapter \ref{chap:impl}.
 
@@
 
-\chapter{Summary}\label{chap:Summary}
-
 \bibliography{D1-3}

### 2014-11-07 D1-3.tex
Added abstract. Updated Graphics.

@@
 
-\LGversion{12}
+\LGversion{9}
 %\LGstaffmonths{Resources spent on the deliverable}
@@
 
-% \TODO{Add abstract}
+This deliverable presents the privacy aware, server based reality
+sensing infrastructure for Live+Gov applications.
 
-%  This deliverable describes the mobile sensor mining component.
-%  Along with this document we supply the source code of the
-%  components, documentation (javadoc), and pre-compiled packages for
-%  direct installation and testing on mobile devices.
+We review the legal and ethical aspects of privacy and perform a
+thorough security analysis of the Live+Gov applications. We identify
+six main threats to privacy and make eight recommendations for
+addressing these threats. These recommendations form the guidelines
+for several privacy enhancements to our infrastructure which we
+describe in this document.
 
+Along with this document we supply the source code of the components,
+documentation, and pre-compiled packages for direct installation.
 }
@@
 \begin{LGhistory}
-\LGaddhistory{01}{2014-07-17}{Outline}{Heinrich Hartmann}
-
+\LGaddhistory{01}{2014-07-17}{Added Outline of Deliverable}{Heinrich Hartmann}
+\LGaddhistory{02}{2014-09-10}{Started IT Security Analysis}{Heinrich Hartmann}
+\LGaddhistory{03}{2014-09-13}{Added Threat Table with Conflicts and Vulnearbilities}{Maximilian Meffert}
+\LGaddhistory{03}{2014-09-15}{Description of IT Security Analysis}{Heinrich Hartmann}
+\LGaddhistory{04}{2014-09-17}{Finished Description of Actors and Interests}{Maximilian Meffert}
+\LGaddhistory{05}{2014-10-08}{Added chapter on Legal Aspects of Privacy}{Heinrich Hartmann}
+\LGaddhistory{06}{2014-10-13}{Updated 7 Types of Privacy}{Heinrich Hartmann}
+\LGaddhistory{07}{2014-10-15}{Updated IT Systems Description}{Heinrich Hartmann}
+\LGaddhistory{08}{2014-10-21}{Added Implicit Privacy Type Violations}{Maximilian Meffert}
+\LGaddhistory{09}{2014-10-25}{Added Description of Implementations}{Heinrich Hartmann}
+\LGaddhistory{09}{2014-11-05}{Extended Threat Scenarios}{Heinrich Hartmann}
+\LGaddhistory{09}{2014-11-07}{Added Introduction and Summary}{Heinrich Hartmann}
 \end{LGhistory}
@@
 
-\LGaddauthor{UKob}{Christoph Schaefer}{+49 261 287 2786 }%
-{+49 261 287 100 2786}{\small chrisschaefer@uni-koblenz.de}
+%\LGaddauthor{UKob}{Christoph Schaefer}{+49 261 287 2786 }%
+%{+49 261 287 100 2786}{\small chrisschaefer@uni-koblenz.de}
 
@@
 
-\LGaddauthorNoPhone{BIZ}{Maite de Arana Agiretxe}%
-{\small komunikazioa.tekniko1@biscaytik.eu}
+%\LGaddauthorNoPhone{BIZ}{Maite de Arana Agiretxe}%
+%{\small komunikazioa.tekniko1@biscaytik.eu}
 
@@
 
-\LGaddabbreviation{\textbf{ALC}}
-{Attribute Language with Complement}
-
 \LGaddabbreviation{\textbf{API}}
 {Application Programming Interface}
-
-\LGaddabbreviation{\textbf{GeSA}}
-{Geographical Semantic Analysis}
 
@@
 {Global System for Mobile Communications}
+
+\LGaddabbreviation{\textbf{HAR}}
+{Human Activity Recognition}
 
@@
 
-\LGaddabbreviation{\textbf{IEEE 802.11}}
-{see WIFI, WLAN}
-
 \LGaddabbreviation{\textbf{JSON}}
@@
 
-\LGaddabbreviation{\textbf{RF-ID}}
-{Radio-Frequency Identification}
-
-\LGaddabbreviation{\textbf{SDCF}}
-{Sensor Data Collection Framework}
+\LGaddabbreviation{\textbf{SLD}}
+{Service Line Detection}
 
@@
 {Structured Query Language}
-
-\LGaddabbreviation{\textbf{SVM}}
-{Support Vector Machine}
-
-\LGaddabbreviation{\textbf{TBox}}
-{Terminological Box}
 

### 2014-11-07 PrivacyProtection-2014-09-09.tex
Updated intro and summary.
Added abstract. Updated Graphics.

@@
 
-\section{Legal Aspects of Privacy}
+\section{Legal Aspects of Privacy} \label{sec:legal}
 
@@
 Of particular importance is the EU Directive 95/46/EC that is discussed in section \ref{EUDIR}.
-
-Based on the legal background presented here, we will derive the legal privacy requirements for our systems in \ref{sec:LEGAL_REQ}.
 
@@
 
-\subsection{Defining Privacy}
+In this section we introduce our definition of privacy which as \emph{control over personal information} following C. Fried.
+As this description is still on a very high level we go on to specify seven different privacy types in the next section.
+These privacy types are then analyzed for their implications in the presence of different kinds of sensor data.
+
+\subsection{Defining Privacy}\label{sec:ethics}
 
@@
 
-\subsection{The Seven Types of Privacy}
+\subsection{The Seven Types of Privacy}\label{sec:privacytypes}
 
@@
 We follow the Reference Model for IT Security Analysis as described in \cite{Grimm:ItSecRefModel}.
-It supersedes earlier efforts by \cite{Avizienis}.
-
+It supersedes earlier efforts by e.g. \cite{Avizienis}.
 The reference model consists of a \emph{model} and a \emph{procedure}.
@@
 
-\section{Live+Gov Privacy Protection Analysis}
+\section{Live+Gov Privacy Protection Analysis}\label{sec:PPA}
 
@@
 \begin{itemize}
+\item physical wellbeing and health
+\item financial profit
 \item convenience
 \item legitimate use of personal data
-\item physical wellbeing and health
-\item financial profit
+\item non-disclosure of personal data to peers of the citizen
 \item not being monitored.
 \end{itemize}
-
-%TODO:
-- Nondisclosure of sensitive information to peers of the citizen.
 

### 2014-12-18 figures/LG-Implicit-Sensor-Privacy-Matrix.tex

\begin{figure}
\centering
\includegraphics[width=\textwidth]{diagrams/png/implicit-sensor-privacy-matrix.png}

%\begin{flushleft}
%\scriptsize
%\textbf{Legend:}
%\begin{itemize}
%\itemsep1pt\parskip0pt\parsep0pt
%\item
%  The x-axis lists the 7 Types of Privacy according to Friedewald et al.
%\item
%  The y-axis lists all mobile sensors currently used by the Live+Gov
%  procject.
%\item
%  The big black bullet point dentotes that there is a \textbf{direct}
%  violoation of a privacy type from a sensor
%\end{itemize}
%\end{flushleft}

\caption{Live+Gov Implicit Sensor-Privacy Matrix}
\label{figure:Live+Gov Implicit Sensor-Privacy Matrix}
\end{figure}

### 2014-12-18 figures/LG-Operation-Scenario-02-Mobile-Mining.tex
Added Privacy Protection tex from Max.
Signed-off-by: Heinrich Hartmann <hartmann@uni-koblenz.de>

--- 2014-09-11 figures/LG-Operation-Scenario-01-Server-Side-Mining.tex
@@
-\begin{figure}[h]
+\begin{figure}
 \centering
-\includegraphics{diagrams/png/scenario01-ServerSideMining.png}
+\includegraphics[width=\textwidth]{diagrams/png/scenario02-MobileMining.png}
 
@@
 \textbf{Legend:}
+
 \begin{itemize}
@@
 
-\caption{Live+Gov Operation Scenario 01 - Server Side Mining}
-\label{figure:Live+Gov Operation Scenario 01 - Server Side Mining}
+\caption{Live+Gov Operation Scenario 02 - Mobile Mining}
+\label{figure:Live+Gov Operation Scenario 02 - Mobile Mining}
 \end{figure}

### 2014-12-18 llncs/paper.tex
Added template

\documentclass[runningheads,a4paper]{llncs}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{listings, color}
\usepackage[bookmarks,bookmarksopen,bookmarksdepth=2]{hyperref}

\graphicspath{ {./img/} }

\begin{document}

%============================================================

\title{Privacy}
\author{-}
\institute{University of Koblenz-Landau}
\maketitle

%============================================================

\begin{abstract}

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.

\keywords{
privacy
}

\end{abstract}

%============================================================

\section{Introduction}
Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.

\end{document}

### 2014-12-18 llncs/paper.tex
added basic structure to paper.tex

@@
 
+\newcommand{\TODO}[1]{\textcolor{magenta}{TODO: #1}}
+
 \graphicspath{ {./img/} }
@@
 
-\title{Privacy}
-\author{-}
+\title{Privacy Analysis of a Mobile Sensor Application}
+\author{Heinrich Hartmann \and Tim Wambach \and Maximilian Meffert \and R\"udiger Grimm}
 \institute{University of Koblenz-Landau}
@@
 
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
-At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
-At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
+\TODO{later}
 
 \keywords{
-privacy
+privacy protection, IT security analysis, sensor data, mobile phones, traffic survey
 }
@@
 \section{Introduction}
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
-At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
-Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua.
-At vero eos et accusam et justo duo dolores et ea rebum.
-Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.
+
+\TODO{RG}
+
+%============================================================
+
+\section{Scenario Description: Mobile Traffic Survey}
+
+\TODO{HH}
+
+%============================================================
+
+\section{Definition of Privacy}
+
+\TODO{MM}
+
+%============================================================
+
+\section{Privacy Analysis}
+
+\TODO{TW}
+
+%============================================================
+
+\section{Conclusion}
+
+\thebibliography
 

### 2014-12-18 llncs/paper.tex
Filled in section from LG deliverable.

@@
 
-\newcommand{\TODO}[1]{\textcolor{magenta}{TODO: #1}}
+\graphicspath{ {./img/} }
 
-\graphicspath{ {./img/} }
+%% CUSTOM BINDINGS %%
+
+\usepackage{amsmath}
+
+\usepackage{alltt}
+\newcommand{\TODO}[1]{\begin{alltt}\textcolor{magenta}{TODO: #1}\end{alltt}}
+
+\newenvironment{LGContent}
+{ \par\color{blue} \it \small }
+{ \par }
+
+\usepackage{xspace}
+\newcommand{\om}{[...]\xspace} % omission
+
+%% END CUSTOM BINDINGS %%
 
@@
 
-\TODO{later}
+\TODO{(later)}
 
@@
 
-\TODO{RG}
+\TODO{RG\\
+* Context: Reference model IT Security analysis\\
+* Application: Travel Survey\\
+* Different definitions of privacy in the literature.\\
+* Study outcomes depend on privacy definition
+}
+
+\begin{LGContent}
+In recent days the importance of privacy protection has been amplified
+by the reports about the mass surveillance of ordinary citizens on a
+global scale by the NSA and other intelligence agencies around the
+world.
+
+While aiming at the noble cause of enhancing eParticipation using
+mobile technologies, Live+Gov systems do process a large variety data
+that is potentially infringing the citizens privacy. The captured data
+includes personal information like name, phone numbers and email
+addresses and sensor data from GPS and accelerometer sensors. Also
+with some applications it is possible to gather images and textual
+input from the citizen.
+
+While the collection of this data is necessary for providing the
+advanced services that Live+Gov aims to deliver, at he same time, the
+available raw data can be used to draw a very detailed picture of the
+private life of the citizen. For instance can GPS location tracking be
+used to reveal shopping habits (e.g. when a car seller is visited) and
+associations to political groups (when a meeting is
+attended). Accelerometer data can be used to infer medical conditions
+like walking disabilities. Images can contain faces of nearby persons
+to with whom the citizen is associated. All this data is highly
+sensitive to the citizens privacy and can be used against the citizen
+if it falls in the wrong hands.
+
+The great importance of protecting the citizens privacy should be
+apparent from these examples. The European Union, as well as many
+other countries in the past, has set out a number of directives that
+regulate the collection, processing and use of privacy sensitive
+data. We explain the most relevant legislation in Section
+\ref{sec:legal}.
+
+The ethical aspects of privacy have been the subject of study of many
+social scientists and philosophers. One scholar which is particularly
+relevant in our context is Charles Fried. He investiages, why we are
+intuitively so sensitive to violations of our privacy. For him privacy
+is not asserted as an intrinsic value by itself, he rather sated:
+
+\begin{quote}
+Privacy is not simply an absence of information about us in the minds of others;
+rather it is the control we have over information about ourselves.
+\end{quote}
+
+Fried`s study on the understanding of privacy provided a great
+contribution to the research on the same term in philosophy and
+computer science and despite the fact his text was published in 1970,
+he already included technologies to its viewpoint (like location
+monitoring) that are particularly relevant to our Context.  We explain
+his theory in \ref{sec:ethics} and follow his definition of privacy in
+this document.
+
+In this document we perform a thorough analysis of how to protect the
+privacy of a citizen and present several implementations that form the
+core of our privacy aware Sensor Data Storage and Mining Infrastructure.
+
+We identify six main threats for the citizens privacy, and derive
+eight recommendations that should be followed in order to reduce the
+associated risks for hazards.  This analysis form the guideline for
+the selection of 9 measures that were implemented in our system and
+documented in Chapter \ref{chap:impl}.
+\end{LGContent}
 
@@
 
-\TODO{HH}
+\TODO{HH:\\
+* System description: (Raw version copied from Privacy Analysis)\\
+* Information needs of transport agency\\
+}
+
+\subsubsection{IT-Systems \& Interactions}
+\label{subsubsection:it-systems}
+
+\begin{LGContent}
+This section outlines the general architecture (Figure \ref{figure:IT Systems}) of IT systems for public monitoring comparable to the Live+Gov project.
+This includes a description of it technical infrastructure and the interactions between its components.
+
+The IT infrastructure of the Live+Gov system, i.e.
+the Live+Gov toolkit and the customization of the software components are described in detail in various project deliverables: D4.1, D4.3, D1.1, D5.1.
+In this section we give an abstraction of those systems from the perspective of WP1.
+
+\input{./figures/IT-Systems}
+
+A \emph{citizen} (a) carries a mobile device running the \emph{sensor collector} application (d).
+The sensor collector application is able to collect various kinds of sensor data (accelerometer, GPS, GSM, ...) and can sent the raw data back to the \emph{data center} (c).
+The sensor collector can also be able to perform certain data mining operations.
+
+Examples of such data mining operations include human activity recognition,the detection of service lines, detection of characters and faces from images.
+
+In the Live+Gov Project mobile devices are used in particular for the following activities:
+\begin{enumerate}
+\item collection of GPS samples,
+\item mining of Human activities (HAR) based on accelerometer samples,
+\item collection of reports consisting of an image, free text and selected categories.
+\end{enumerate}
+
+The \emph{data center} stores and processes sensor data collected with the sensor collector application.
+It can also take into account data obtained from third parties, like the current positions of trains.
+The data center can sent mining end products (traffic jam reports, bus schedule) back to the mobile device of the citizen.
+
+In the Live+Gov Project data centers, in particular, perform the following activities:
+\begin{enumerate}
+\item storage of login credentials, name, and email address for each citizen,
+\item storage of collected GPS samples (user id, timestamp, GPS location),
+\item storage of HAR results (user id, timestamp, HAR result),
+\item detection of service lines based received GPS samples (SLD),
+\item storage of SLD results (userid, timestamp, SLD result),
+\item sent back SLD results to mobile device,
+\item storage of received reports (user id, timestamp, report),
+\item detection of inherent patterns of the received reports.
+\end{enumerate}
+
+The \emph{System Provider} (b) provides and operates technical infrastructure like the Data Center and the \emph{Reporting Tool} (f).
+The Reporting Tool queries the Data Center for aggregated data to visualize in form of charts and other means suitable to help understanding of monitored citizens.
+
+\textit{Local Authorities} (c) use the reporting tool to get information in order to understand citizen movement and improve public services.
+In such systems, the most valuable information for local authorities is not the raw sensor data, but aggregated views on the mining end products.
+
+In the Live+Gov Project the reporting tools allows, in particular the following queries:
+\begin{enumerate}
+\item show aggregate information about which routes citizens take starting from a given location,
+\item show the average waiting time of a citizen for each bus stop,
+\item show routes where citizens where running to catch a bus,
+\item show locations of all reports in a given time window,
+\item visualize detected patterns in the report data,
+\item view individual reports.
+\end{enumerate}
+\end{LGContent}
 
@@
 
-\TODO{MM}
+\TODO{MM:\\
+- Shorten\\
+- Remove Implicit Type Violations\\
+- Update figure\\
+- RG: Discuss alternative definitions of privacy\\
+}
+
+\begin{LGContent}
+Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes:
+
+\begin{quote}
+The notion of privacy remains out of the grasp of every academic chasing it.
+Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.
+\cite{Gutwirth}
+\end{quote}
+
+%Many researchers seem to only ``focus on the ways in which privacy can be infringed'' \cite{7ToP}.
+%Thus they try to create taxonomies of \emph{privacy harms} instead of taxonomies of \emph{privacy types}.
+%Those two differ in the respect that the former focuses on threats to prohibit whereas the latter focuses on values to protect.
+%So one should rather evaluate what aspects are precious about privacy and develop measures to ensure their security than only forbid single actions against it \cite{7ToP}.
+
+Many researchers seem to only ``focus on the ways in which privacy can be infringed'' \cite{7ToP}.
+Thus they invest a great amount of work in defining threats to prohibit instead of describing why privacy is so valuable to us.
+This way around one could derive measures to ensure and secure its value \cite{7ToP}.
+One particular scholar who does this in the context of digital monitoring is Charles Fried.
+He questioned, why we are intuitively sensitive to violations of privacy.
+But he did not assert privacy as an intrinsic value by itself, he rather sated:
+
+\begin{quote}
+Privacy is not simply an absence of information about us in the minds of others;
+rather it is \textbf{the control we have over information about ourselves}. \cite{CFried:Privacy}
+\end{quote}
+
+According to Fried privacy is a \textit{rational context}.
+Which means that one is aware of such context's existence during a rational action.
+If we are to share private information with others, we are most likely aware that this action is privacy related.
+In that case we are able to selectively disclose information along the two dimensions of quantity and quality.
+This leads Fried to his thesis, that privacy is one's ability to create and modulate his or hers social relationships, namely: friendship, love and trust. \cite{CFried:Privacy}
+
+Depending on the conversation partner, we change the degree of intimate information we share if it is a total stranger, colleague, close friend or lover.
+With close friends or lovers we share information of great intimacy we do not share with anyone else.
+Moreover, we trust those persons to not reveal information about us to others by respecting their privacy.
+Trust needs the possibility of unknown failure.
+If we would constantly monitor our partners, they cannot fail unnoticed nor can they willingly share that information with us.
+Thus they could not trust us anymore. \cite{CFried:Privacy}
+
+So privacy or the its possibility, according to Fried, is the foundation of our core relations: friendship, love and trust.
+And thus it is valuable, because those relations are essential to human society \cite{CFried:Privacy}.
+Using his anatomy of privacy as foundation of our analysis is suitable because of two points.
+At first, Fried's study on the understanding of privacy provided a great contribution to the research on the same term in philosophy \cite{sep-privacy} and computer science \cite{SociotechnicalArchitectureForOnlinePrivacy}.
+Secondly, despite the fact his text was published in 1970, he already included technologies to its viewpoint that did not only monitor location, but also record biometric data.
+
+\subsection{The Seven Types of Privacy}\label{sec:privacytypes}
+
+In Fried's definition of privacy as control over "information about onself'', the specification of what constitues such information remains open.
+There is a vast amount of information that relates to a person and we need to get a better understanding in order to perform a thorough analysis.
+To this end we use the the categorization by Friedewald, Finn and Wright \cite{7ToP} called the \textit{Seven Types of Privacy}.
+The seven types of privacy are an extension to the four types of privacy by Roger Clarke \cite{RClarke:4ToP}, which are:
+\begin{itemize}
+\item Privacy of the Person
+\item Privacy of Personal Behaviour
+\item Privacy of Personal Communication
+\item Privacy of Personal Data
+\end{itemize}
+It is important to note, that this categories do not form no taxonomy, since the categories are not mutually exclusive.
+For instance a written email is considered personal communication as well as personal data stored on a computer.
+
+Moreover, Friedewald et al. argue that Clarke's taxonomy is outdated and no longer adequate in order to describe the privacy aspect of our modern, technology driven, world.
+In order to address this shortcoming they extend the former four to the now introduced seven types privacy as follows:
+
+\begin{enumerate}
+
+\item \textbf{Privacy of the Person}
+This privacy type is generally concerned with one could best understand as biometric privacy.
+Friedewald et al. paraphrase it as \emph{``\om the right to keep body functions and body characteristics \om private''}.
+This includes but is not limited to measures like weight, height or shoulder width;
+biometric identifiers like fingerprints and DNA sequences;
+or medical conditions such as limping or having a cold.
+
+\item \textbf{Privacy of Behaviour and Action}
+This privacy type is concerned with one's activities in public as well as in private spaces.
+It includes but is not limited to religious practices, political activities and sexual preferences or habits.
+
+\item \textbf{Privacy of Communication}
+This privacy type is concerned with one's communication in a broad sense.
+It includes written correspondence, but also conversations conducted either vis-a-vis or via electronic devices.
+Friedewald et al. put it as the right to free discussion without unknown interception by third parties.
+
+\item \textbf{Privacy of Data and Image}
+This privacy type is concerned with the secrecy of personal data, especially its automatic disclosure to other individuals and organizations.
+It includes data such as paychecks, insurance information or records of public administration.
+However, it also refers to pictures taken without consent and digital identifiers like IP addresses or social security numbers.
+
+\item \textbf{Privacy of Thoughts and Feelings}
+This privacy type is the counterpart to Privacy of the Person like body and mind are counterparts of one another.
+Comparable to the Privacy of Data and Image, Friedewald et al. state that one's thoughts and feelings must not be automatically revealed to others.
+This could simply happen by the disclosure of one's diary or by technologies which allow emotion detection through biometric means.
+One's body temperature or iris reflexes might infer stress or excitation.
+
+\item \textbf{Privacy of Location and Space}
+This privacy type is concerned with one's movements in public spaces and the protection of private spaces.
+Friedewald et al. qualify the first dimension as one's right to move without being identified, tracked or monitored.
+The second dimension is qualified as one's general right to solitude, especially the right to the inviolability of the home.
+
+\item \textbf{Privacy of Association}
+This privacy type is also put as group privacy.
+Friedewald et al. state that one must have the possibility with whomever without being recorded.
+Associations like friends or organizations such as political parties must not automatically be recorded because one associates with them, and vice-versa.
+\end{enumerate}
+
+\subsection{Privacy Type Inferences}
+
+The above types of privacy types are not disjoint to each other.
+For example if the location of a citizen is known it is possible to infer information about political associations (e.g. his visits to a party meeting).
+In this section we explore possibilities how personal data can be inferred from each other.
+Our findings are summarized in Figure \ref{figure:Implicit Privacy Violation Matrix}.
+
+%\includegraphics{../diagrams/png/implicit-privacy-violation-matrix.png}
+\input{./figures/Implicit-Privacy-Violation-Matrix}
+
+\paragraph*{1. Privacy of The Person}
+
+The Privacy of The Person is concerned with one's biometric privacy.
+If this type is violated, following implicit violations are possible:
+
+\begin{itemize}
+\item [(1-5)]
+  Privacy of Thoughts and Feelings. Some psychological diseases
+  (e.g.~depression) have physiological impact. Such physiological
+  patterns could be detected.
+\end{itemize}
+
+\paragraph*{2. Privacy of Behaviour and Action}
+
+The Privacy of Behaviour and Action is concerned with one's social, political, religious, sexual, etc. activities.
+If his type is violated, following implicit violations are possible:
+
+\begin{itemize}
+\item [(2-1)] Privacy of The Person.
+  Religious practices which include body modifications (e.g.~circumcision).
+\item [(2-5)] Privacy of Thoughts and Feelings.
+  Social activities in general depend on a certain intellectual attitude.
+  Such an activity is the expressions of such an attitude.
+\item [(2-7)] Privacy of Association.
+  Recording religious, political or sexual activities can reveal association with churches, political parties or sexual partners.
+\end{itemize}
+
+\paragraph*{3. Privacy of Communication}
+
+The Privacy of Communication is concerned with not having such communication (correspondence or vis-a-vis) intercepted.
+This is very broad type of privacy.
+Depending on the contents of the intercepted communication every other type can be violated:
+
+\begin{itemize}
+\item [(3-1)] Privacy of The Person. Communication about body characteristics.
+\item [(3-2)] Privacy of Behaviour and Action. Communication about social activities.
+\item [(3-4)] Privacy of Data and Image.
+  Communication containing one's passwords or other sensitive data.
+\item [(3-5)] Privacy of Thoughts and Feelings.
+  Communication of thoughts and feelings, e.g.~wiretapping a flirt or a catholic confession
+  ritual.
+\item [(3-6)] Privacy of Location and Space.
+  Interception of face-to-face communication is only possible if one's location and space is violated (wiretapping).
+\item [(3-7)] Privacy of Association.
+  Communication about one's associations (family members, churches, etc.).
+\end{itemize}
+
+\paragraph*{4. Privacy of Data and Image}
+
+The Privacy of Data and Image is concerned with one's data not being automatically available to others.
+This also is a very broad type of privacy.
+Depending of the data or image contents every other type can be violated:
+
+\begin{itemize}
+
+\item [(4-1)] Privacy of The Person.
+  Images or stored biometric information reveal one's physical characteristics.
+\item [(4-2)]Privacy of Behaviour and Action.
+  Images or diaries can reveal one's social activities.
+\item [(4-3)]Privacy of Communication.
+  Modern communication systems usually contain some sort of archive function, e.g.~E-mail clients do not automatically delete messages.
+  Such messages are data and reveal one's communication.
+\item [(4-4)]Privacy of Thoughts and Feelings.
+  Images can show one's emotional state.
+\item [(4-5)]Privacy of Location and Space.
+  Images can reveal one's location, e.g.~making a picture in front of the Eifel Tower.
+\item [(4-6)]Privacy of Association. E-mail data can also reveal
+  association.
+\end{itemize}
+
+\paragraph*{5. Privacy of Thoughts and Feelings}
+
+The Privacy of Thoughts and Feelings is concerned with keeping such
+thoughts and feelings secret. If this type is violated, following
+implicit violations are possible:
+
+\begin{itemize}
+\item [(5-1)] Privacy of The Person.
+  Thoughts and feelings can reveal medical conditions.
+\item [(5-2)] Privacy of Behaviour and Action.
+  Thoughts and feelings can reveal a certain attitudes which create a foundation for certain   social activities.
+\item [(5-7)] Privacy of Association.
+  Thoughts and feelings can reveal individual association, e.g amorous feelings for a certain person.
+\end{itemize}
+
+\paragraph*{6. Privacy of Location and Space}
+
+The Privacy of Location and Space is concerned with one's right to move freely without being tracked and one's right to private places.
+If this type is violated, following implicit violations are possible:
+
+\begin{itemize}
+\item [(6-1)] Privacy of The Person.
+  Frequently visited doctors can reveal certain medical conditions, if such doctors are known specialsts.
+  In general it could imply illness.
+\item [(6-2)] Privacy of Behaviour and Action.
+  Frequently visited places in general can reveal association and hence implies social activities.
+\item [(6-3)] Privacy of Communication.
+  If one's location is knwon, it is possible to intercept (wiretap) one's communication.
+  This also may violate the right to private spaces.
+\item [(6-4)] Privacy of Data and Image.
+  If one's location is known, it is possible shoot pictures.
+\item [(6-5)] Privacy of Thoughts and Feelings.
+  Frequently visited persons may imply certain thoughts and feelings, e.g.~having a mistress.
+\item [(6-6)] Privacy of Association.
+  Frequently visited places can reveal associations simply by searching in maps or yellow-pages.
+\end{itemize}
+
+\paragraph*{7. Privacy of Association}
+
+The Privacy of Association is concerned with one's right to associate with whomever one wants, without that association having recorded.
+If this type is violated, following implicit violations are possible:
+
+\begin{itemize}
+
+\item [(7-1)] Privacy of The Person.
+  Association with tattoo artists could imply having tattoos or other body modifications.
+\item [(7-2)] Privacy of Behaviour and Action.
+  Association with churches or political organizations could imply certain activities.
+\item [(7-5)] Privacy of Thoughts and Feelings.
+  Association with churches of political organizations could imply a certain intellectual attitude.
+\end{itemize}
+
+\end{LGContent}
+
+\subsection{Sensor Data Privacy Impact}\label{sec:SensorPrivacyImpact}
+
+\begin{LGContent}
+
+\input{./figures/LG-Implicit-Sensor-Privacy-Matrix}
+
+%TODO: Reformulate the section.
+% - Too much overlap with preceeding sections.
+% - Structure should be per Sensor Type
+% - Add HAR and SLD!
+
+In this section we analyze the impact of the disclosure sensor data and certain processing results to the citizens privacy.
+This analyisis builds upon the preceeding analyis, but is more focused on the concrete type of data vailable.
+For example, the disclosure of Service Line Detection results, does violate the Privacy of Location and Space but not the Privacy of Association that is violated by general GPS tracking.
+
+\textbf{GPS Data}.
+The GPS sensor gives the current longitude and latitude, the current global position of the mobile device and its carrier, although there is some artificial inaccuracy within civil use.
+Threefore, the collection of GPS data violates directly the citizens privacy of Location and Space, and the privacy of Data an Image.
+
+By inference we also get implicit violations of the Privacy of the Person, Privacy of Behavior and Action and Privacy of Association.
+
+\textbf{Motion Sensors.}
+Accelerometer, Rotation Vector, Gyroscope and Magnetic field sensor measure the physical movement of the mobile device on all three axes.
+If the mobile device is carried ``normally'' its safe to say that those sensors also measure the moments of its carrier.
+So his privacy is infringed regarding biometric behaviour, as it is captured automatically
+Privacy of Data and Image is trivially threatened because here sensor data is individual data, a priori.
+
+By inference we also get implicit violations of the Privacy of Behavior and Action.
+
+\textbf{Network Sensors.}
+The GSM and WLAN sensors reveal the position of the mobile device and its carrier, when used in connection with external databases.
+The GSM sensor gives the exact cell, the mobile device has registered with at the current moment.
+
+The Bluetooth sensors record lists of the bluetooth clients in the direct neighbourhood.
+Since those clients are usually moving, inference of the position is usually not possible.
+Instead, bluetooth clients carried by a third person may infringe the Privacy of Association.
+
+A similar argumentat applies to WLAN snsors. If one frequently connects with an organizational wireless network, e.g.~an university network, an association can be deduced (student or staff).
+
+\textbf{Human Activity Recognition.}
+The detection and collection of human activities like walking, standing and running, can interfer with the Privacy of the Person, e.g.~since these movement patterns can be indicators for a person's health.
+Also, trivially, Privacy of Data and Image is violated.
+
+\textbf{Service Line Detection.}
+The detection and collection of the service line the user currently uses in the public transportation system allows inference of the location of the user, at least when entering and leaving those service lines at e.g.~bus stops.
+Implicit privacy violations apply accordingly.
+
+\end{LGContent}
 
 %============================================================
+
+\section{IT Security Reference Model}
+
+\TODO{TW}
+
+\begin{LGContent}
+
+The goal of this chapter to analyze and identify the threads to
+personal privacy that are posed by collecting, storing and processing
+sensor data from mobile phones.  We derive concrete privacy protection
+measures that address the main risks involved with handling such data.
+
+The complexity of our systems and the variety of threads make a great
+number of counter measures plausible.  We approach this complexity
+with the aid of a general security analysis model developed in
+\cite{Grimm:ItSecRefModel}.  We give a brief introduction to this
+model and perform a IT Security Analysis with respect to the privacy
+asset for our system.
+
+We follow the Reference Model for IT Security Analysis as described in
+\cite{Grimm:ItSecRefModel}.  It supersedes earlier efforts by
+e.g. \cite{Avizienis}.  The reference model consists of a \emph{model}
+and a \emph{procedure}.  The model aims to organize common security
+terminology in a reasonable and practical way.  The procedure
+describes a method to analysis the IT system based on that model.  In
+this section we give a brief overview over the reference model.
+
+\subsection{Model}
+
+%\includegraphics{../diagrams/png/itsec-ref-model-grimm.png}
+\input{./figures/The-IT-Security-Reference-Model-Grimm}
+
+The model is depicted in Figure \ref{fig:refmodel}.  It is organized
+in four views (round boxes) that contain a number of components
+(rectangular boxes).
+
+%%%%%%%%%%%%%%%%%%%%
+
+The \emph{world view} contains all components describing the current state.
+It consists of the following components:
+\begin{itemize}
+\item \textbf{Actors.}
+All identifiable stakeholders of the system under study.
+Typical actors include, users, developers, clients and externals.
+
+\item
+\textbf{Assets.} Things of value to one or more stakeholders.
+The value can be hard (money, data, etc.) or soft (trust, privacy,etc.).
+In our case the only asset we are concerned with is the privacy of the citizen.
+
+\item \textbf{IT-Systems.}
+The relevant IT-Systems under study. This encompasses hardware (e.g. servers, network infrastructure), as well as software and third party services.
+The level of granularity has to be detailed enough to express all possible threats to the assets at stake.
+
+\item \textbf{Conflicts of Interests.}
+Different actors have different interests which can be in conflict which each other.
+These conflicts of interest are the origin of all attacks to the system.
+
+A typical conflict is the \emph{Criminal-User-Conflict}:
+A user wants to keep control over their private data.
+A criminal wants to gain money.
+The possibility of selling private data (user profiles) to advertisers, renders both interests conflicting.
+
+\item \textbf{Vulnerabilities.}
+All identifiable weaknesses in the IT-System.
+
+In the example of the criminal-user-conflict, the criminal has to exploit a vulnerability, e.g. a weak password, to gain access to the private data about the user.
+
+\item \textbf{Interactions.}
+This point captures all possible interactions between assets, IT-Systems, humans and vulnerabilites.
+It is described in more detail in the next view.
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%
+
+The \emph{potential view} displays the intended and unintended interactions of the components in the world view.
+The intended interactions support the underlying business objectives.
+Unintended interactions lead to threats.
+The potential view consists of the following components.
+\begin{itemize}
+\item \textbf{Business Objectives.}
+Interaction of IT-system and actors that realize a business goals of the system owner.
+
+\item \textbf{Threats.}
+A threat is a potential interaction that destroys or harms assets of the system.
+Concrete realizations of threats can be \emph{attacks} or \emph{accidents}.
+Attacks are executed by an actor in response to a conflict of interest.
+Accidents are harmful interactions that are not willfully caused by an actor.
+
+\item \textbf{Chances/Risk.}
+Evaluation of chances and risks associated to the business objectives and threats.
+The risk associated to a threat is its expected loss.
+A chance associated to a intended interaction is its expected gain.
+
+In the case that, the loss can be quantified monetary, and the likelihood of occurrence of a threat can
+be modeled probabilistically, the risk is given by the product
+\[ \text{risk} = \text{loss} \cdot P[\text{threat}]. \]
+In practice such a quantitative risk evaluation is often not possible, and a qualitative, heuristic, analysis is performed instead.
+
+\item \textbf{Security Requirements.}
+A set of interactions (e.g. threats) that shall not occur within the system in order to achieve its business objectives.
+Security requirements are targeted to protect one or more assets.
+
+An example of a security requirement is that a given communication channel shall not be infringed by externals.
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%
+
+The \emph{concept view} is a realization of the potential view of the system.
+It specifies important interactions that require further planning.
+It contains the following components:
+\begin{itemize}
+\item \textbf{Business Model.}
+The plan to achieve business objectives.
+
+\item \textbf{Accident Scenario.}
+A concrete outline of an interaction that leads to an accident.
+In particular the asset under threatened and exploited vulnerability need to be described.
+
+\item \textbf{Attack Technique.}
+A specific technique or technology to attack IT-Systems (Man in the Middle, Phishing, etc.).
+In particular the attacking actor, the conflict of interest and the exploited vulnerability need to be described.
+
+\item \textbf{Security Measures.}
+It describes a plan of sufficient measures to secure the intended interactions and to avoid the unintended interactions.
+Each security measure targets a vulnerability of the system in order to reduce a risk for a certain thread.
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%
+
+The \emph{event view} contains all actual events through out the lifetime of the system.
+The event view instantiates the concept view of the system.
+It contains the following components:
+\begin{itemize}
+\item \textbf{Business Process.}
+The actual, running instance of the business model.
+
+\item \textbf{Accidents.}
+All actually happened accidents.
+
+\item \textbf{Attacks.}
+All actually happened attacks.
+
+\item \textbf{Security Operations.}
+Instances of security measures.
+\end{itemize}
+
+\subsection{Procedure}
+
+The analysis procedure is an incremental and iterative process following the four views of the previously described model.
+
+%\includegraphics{../diagrams/png/itsec-ref-model-grimm-procedure.png}
+\input{./figures/The-IT-Security-Reference-Analysis-Grimm}
+
+\paragraph*{Step 1. World Analysis}
+
+At first, one has to outline the current state of the system under study. This includes description of:
+\begin{itemize}
+\item all \textbf{Assets} which must be protected
+\item all relevant \textbf{IT-Systems}
+\item all involved \textbf{Humans} and their \textbf{Conflicts of Interests}
+\item all known \textbf{Vulnerabilities}
+\item and all important \textbf{Interactions} between the former components.
+\end{itemize}
+
+\paragraph*{Step 2. Potential Analysis}
+
+Secondly, one needs to outline the potential interactions of the system under study.
+This includes both the unintended interactions (threads) and the intended interactions (business objectives).
+This step produces four artifacts:
+\begin{itemize}
+\item a \emph{threat specification}, which identifies the threat, its targeted assets, the involved actors and their conflicts of interest.
+\item a \emph{threat risk evaluation}, which quantifies the likelihood of a threat manifestation in relation to its associated loss.
+\item a \emph{security requirement specification}, which specifies requirements in order to deal with identified hazards
+\end{itemize}
+
+\paragraph*{Step 3. Concept Development}
+
+Based on \textbf{Step 2.}, the identified hazards are used alongside realistic accident scenarios and attack techniques to create a \emph{risk matrix}.
+With this matrix it is possible to decide if the risk is acceptable or not.
+Together with the previously specified security requirements, the matrix is used to define adequate security measures.
+Like a business model is an abstract concept to achieve business objectives, this step creates an concept to improve the system's security.
+
+\paragraph*{Step 4. Concept Deployment}
+
+Finally, the security measures have to be implemented.
+Additionally, all business operations, accidents, attacks and executed security operations will be recorded in the following time.
+
+The implementation of security measures changes the world view (e.g. IT-systems, actors) and renders the conducted analysis outdated.
+So this analysis procedure needs to be conducted again.
+
+\subsection{Abstraction Levels of the Reference Model}
+
+The Reference Model can be used on different levels of abstraction.
+This means each component can be used within a wide range of granularity, for instance the security measure \emph{Encryption} can be explored in general or on the level of different concrete encryption tools; or on the even finer level of concrete algorithms.
+
+The utilized abstraction level is not important for the analysis procedure, it depends on the intended audience for the analysis.
+However, it is important to use one abstraction level consistently through out the analysis.
+\end{LGContent}
 
@@
 
+\begin{LGContent}
+\subsection{Step 1. World Analysis}
+
+\subsubsection{Assets: Privacy}
+
+In this document we focus our attention to only one asset: The privacy of the citizen.
+
+%TODO: UPDATE
+Our definition of privacy is described in detail in Chapter \ref{chap:privacy}.
+In Section \ref{sec:taxonomy} we define privacy as the ``control over private data'' and introduce the following seven different types of privacy:
+\begin{enumerate}
+\item Privacy of the Person
+\item Privacy of Behaviour and Action
+\item Privacy of Communication
+\item Privacy of Data and Image
+\item Privacy of Thoughts and Feelings
+\item Privacy of Location and Space
+\item Privacy os Association
+\end{enumerate}
+
+We refer to Section \ref{sec:taxonomy} for more details.
+
+%
+%
+% SECTION: IT SYSTEM DESCRIPTION MOVED ABOVE
+%
+%
+
+\subsubsection{Actors}
+\label{subsubsection:humans}
+
+This section describes the human actors previously introduced within the IT system architecture (\ref{subsubsection:it-systems}).
+Also the additional actor \textit{External} is introduced as a person with no special access rights.
+
+\textbf{Citizen.}
+Citizens are persons who use the mobile device as users of the provided software.
+Their main motivation for using the software is to gain a higher level of convenience in their daily activities.
+For example they may have access to real time bus schedules or reports about traffic jams, that help them to avoid long waiting times.
+Also they generally benefit from improvements of public infrastructure by local authorities, which is triggered by issue reports.
+
+By using the application citizens are sharing personal information like name and address, as well as data gathered from mobile sensor with the service provider.
+This data can be exploited in ways that are harmful to the citizen.
+This need to protect this protection is manifest in several laws and constitutions as the right to privacy and data protection.
+The citizen has a vital interest in having his legal rights protected and enforced.
+
+If the right to privacy is violated, there is a magnitude of potential harms that interfere with other interests of the citizen.
+This includes just annoying spam, where their technical identity is used to send unwanted commercials.
+More severe phishing attacks can exploit personal information, and try to manipulate citizens into disclosing credentials like TAN numbers and disrespect their financial interests.
+Information about the current location (GPS) or frequently used routes (stalking) can be used to attack and directly harm the health of the citizen.
+Information about medical conditions inferred by sensor data (e.g.
+fintness trackers) are of interest to insurance companies, which may affect the pricing of policies and can interfere with financial interests of the citizen.
+
+Also it is well established (cf. \cite{GuardienMassSurveillance}) that the very act of being monitored can have impact on mental health and performance, promotes distrust and breeds conformity.
+
+In short, citizens are interested in
+\begin{itemize}
+\item physical wellbeing and health
+\item financial profit
+\item convenience
+\item legitimate use of personal data
+\item non-disclosure of personal data to peers of the citizen
+\item not being monitored.
+\end{itemize}
+
+\textbf{External.}
+Externals are persons who do not have privileged access to the IT systems, and are willing to break laws, security constrains and norms in order to promote their interests.
+
+If the external is in some kind of relationship to the citizen, like a friendship or business partnership, the external can have a direct interest in gaining information about the citizen in order to increase their power.
+
+Another common interest of an external is financial profit. For example they want to obtain access to critical systems to steal sensitive data or to get the system under their control.
+Controlled systems could be leased as part of a bot net.
+Stolen data could simply be sold as is or used for illegitimate purposes, e.g. spam or phishing attacks - or excessive data mining.
+
+Because local authorities are involved in the general outline of the Live+Gov system, the possibility for politically motivated attacks is given.
+
+Externals could want to harm or destroy the systems in order to damage the reputation of local authorities (politicians or other officials) or to make a political statement of their own.
+
+Another possible motivation for external  activities could be social appreciation.
+A hacker could attack critical infrastructure just to prove his skills.
+
+In short, externals are interested in
+\begin{itemize}
+\item increase power over citizen
+\item financial profit
+\item political activism
+\item social standing.
+\end{itemize}
+
+\textbf{System Provider.}
+System Providers operate the technical infrastructure (hardware and software) of the IT System.
+They are private companies and legal persons in their own right, but also employ a number of people with diverging interests, including:
+administrators, who maintain and operate the running system;
+programmer/developer, who develop the system;
+a support manager, who handles customer relations.
+
+As companies, they are interested in gaining financial profit.
+Among other things, this depends on customer satisfaction, employee happiness and task complexity.
+Unsatisfied customers may not want to pay for the service or do not continue the business relation.
+Moreover, unsatisfied customers can create a bad reputation, which affects the market for future customers.
+
+Customer satisfaction is connected with the quality of the offered product or service.
+This quality depends on the happiness of employees.
+Employees have a claim to professional excellence.
+They want to deliver a good job within their means.
+If employees cannot satisfy their demand for professional excellence, they might get discontent and deliver poor work.
+Moreover, unhappy employees can produce higher costs through sick days.
+The worst case scenario could be, that a discontent employee gets angry and steals data or harms the running systems.
+
+At last, the financial success of system providers depends on the task complexity of the maintained infrastructure.
+The complexity of a task has to be in reasonable bounds, so that system providers can complete it within time, with a satisfying quality.
+If a task has a higher complexity than expected, financial loss is almost certain.
+Either System Providers need to hire additional competence to meet schedule and requirements.
+Or system providers they stress the time-line, which also results in a higher man-hour salary ratio and additionally endangers customer satisfaction.
+Ultimately, high task complexities can affect employee happiness, if employees cannot complete it within their claim to professional excellence.
+
+In short, System Providers are interested in
+\begin{itemize}
+\item financial profit
+\item good working conditions
+\item professional excellence
+\item manageable complexity.
+\end{itemize}
+
+\textbf{Local Authority.}
+Local authorities are public offices (ministry, agency, department, ...) or other external public entities which act as direct customers of service providers.
+They purchase a system specialized for their needs.
+For example a department for urban mobility, orders a system to better understand usage patterns and make improvement to the urban traffic flow.
+
+Such systems are investments, and so naturally local authorities are interested in a profitable return, like increased ticket sales.
+However, the return of investment is not directly of financial nature.
+Like service providers their financial gain depends on customer satisfaction.
+Customers for local authorities are either citizens, who use their services, or politicians, who order their services.
+The satisfaction of both sides is interdependent.
+
+Citizens are satisfied customers if the services, e.g. public mobility, work well.
+If citizens are happy, it is more likely that politicians gain reputation, as they organize the public services through local authorities.
+
+The first important step of improving the public services is by obtaining business intelligence.
+For the urban mobility scenario the Live+Gov systems provide insight in form of traffic jam detection and usage pattern mining, which allow local authorities to focus their efforts to the most important sites.
+
+Additionally, since local authorities act like corporations comparable to service providers, they are also interested in good working conditions.
+Discontent employees may harm the system by e.g. disclosure of privacy sensitive data.
+
+In short, local authorities are interested in
+\begin{itemize}
+\item financial profit
+\item political reputation
+\item business intelligence
+\item good working conditions.
+\end{itemize}
+
+\subsubsection{Conflicts of Interests}
+\label{subsubsection:Conflicts of Interests}
+This section outlines the Conflicts of Interests (Figure \ref{figure:Live+Gov Conflicts of Interests}) between the actors of the proposed IT system architecture.
+
+The individual interests of all actors is already described in the previous section and are not elaborated any further.
+The emphasis here is put on prominent existing conflicts, because they provide a foundation for vulnerabilities and subsequent threats.
+
+\input{./figures/LG-Conflicts-Of-Interests}
+%TODO: Adapt Figure
+
+\textbf{System Complexity vs Privacy.}
+System Providers offer a service to Local Authorities, which is to provide and maintain a monitoring and mining system, e.g. for public mobility.
+This system shall produce business intelligence, so that Local Authorities can improve their public services.
+This task in it self has a high technical complexity and is the sole asset with financial return for System Providers.
+However, this task operates on privacy sensitive data provided by monitored Citizens.
+In order to ensure their privacy, System Provider would have to implement additional mechanisms, which allow Citizens to exercise control of their data.
+This will not only raise the complexity of the monitoring and mining system, System Providers also have to layout the complexity in a comprehensible manner.
+To effectively enable Citizens to preserve their privacy, they need to know what happens with their data.
+
+\textbf{Business Intelligence vs Privacy.}
+Local authorities order a monitoring and mining system from system providers, which allows them to produce business intelligence for public services.
+The system is an investment for local authorities, so they are interested in as much intelligence as possible to achieve a profitable return.
+
+The gained intelligence is the result of data mining conducted on privacy sensitive data of participating citizens.
+They are interested in the successful usage of their data, in a sense that they are also benefactors, e.g. improvement of public mobility.
+The main interest of citizens lies in maintaining control over their data and protecting their rights to privacy.
+In order do that, they need full disclosure of the processing steps and the purposes their data is used for, and to be given a choice whether such processing should be allowed for their own data.
+
+\textbf{Power of External vs. Privacy.}
+Externals which are in a social relation to the citizen can have an interest in obtaining further information in order to gain power.
+In the most simplistic example this could be a man wanting monitor the activities of his spouse.
+Another example is a Government spying on it's citizens in order to suppress opposition.
+An additional twist in the last example is, that there can be legal regulations that require the service provider to support the Governments invasion of the citizens privacy.
+
+\textbf{Financial Profit of External vs Privacy.}
+Externals can gain financial profit from stealing privacy sensitive data.
+For example by selling raw contact information to advertisers or by selling mined data to insurance companies, or intermediaries like scoring companies.
+In such cases, citizens lose complete control over their data.
+
+\textbf{Financial Profit of External vs Reputation of System Providers.}
+Externals have various business models as optional foundation for attacks on System Providers.
+For instance, they can try to invade the infrastructure for e-espionage reasons, to get control over servers to create a bot-net or to steal user data.
+All these approaches are motivated by financial interests.
+Gathered information can be sold, zombie servers can be leased.
+
+A successful attack proves the technical competence of system providers wrong and subsequently harms their professional reputation.
+This can lead to a loss of future customers or a decrease of stock price for registered companies.
+Eventually also the financial interests of system providers are endangered.
+
+\textbf{Political Activism vs Reputation of Local Authority.}
+Besides monetary reasons, externals can be motivated by political reasons to attack the monitoring and mining system.
+Externals can break the system to make a political statement of their own,
+or they can steal user data to prove the system insecure.
+Both would harm the reputation of local authorities, who endangered the privacy of the citizens.
+
+\subsubsection{Vulnerabilities}
+\label{subsubsection:Vulnerabilities}
+This section outlines the vulnerabilities (Figure \ref{figure:Live+Gov Vulnerabilities}) of the proposed monitoring and mining system.
+Note that vulnerabilities are not necessarily of technical nature.
+The weaknesses of IT systems are often created due to misuse or misconfiguration of the various components by one or more actors.
+
+\input{./figures/LG-Vulnerabilities}
+
+\textbf{Insecure Infrastructure.}
+The proposed monitoring system consists of many hardware and software components, each with its own concrete weaknesses.
+For instance, operating systems can be outdated or not subject to frequent updates or virus scans.
+Web-applications can be carelessly implemented and not protected against SQL-Injections or Cross-Site-Scripting attacks.
+Databases can be ill-configured, so that access from outside the system is possible.
+All those weak points can be subject to various known exploit techniques.
+
+\textbf{Insecure Data Transmission.}
+The proposed monitoring and mining system uses HTTP to exchange data between the Sensor Collector, the Data Center and the Report Tool.
+Per default, HTTP is a clear text protocol.
+This means, one can intercept the connection and read all sensitive information, which is send between the components.
+That is: passwords, raw sensor data and data mining results
+
+\textbf{Unhappy Employees.}
+
+An Employee that is frustrated with his situation for a long time
+period constitutes a security vulnerability. On the one hand he might
+want to harm his employer directly, on the other he is increasingly
+susceptible for social engineering.
+
+\textbf{Inadequate Access Rules.}
+The proposed IT system infrastructure has various accesses to privacy sensitive data.
+System Provider staff has access to Data Center hardware and software like databases, web-servers and other inspection tools.
+Local Authority staff has access to the Report Tool.
+This all enables staff members to have potential access to privacy sensitive information.
+Those accesses have to be secured against unauthorized third parties.
+Moreover, we need to ensure that no single person has to many access rights.
+For example, a system administrator should not be able to secretly download the whole database on a flash-drive.
+
+%TODO: Extend to general missing privacy wareness of citizen, system provider and local authority
+\textbf{Unaware Monitoring Subjects.}
+We define privacy as one's ability to control information about
+oneself.  In order to do that, monitored subjects need to know, that
+they are monitored, who monitors them, what information is recorded
+and for what purposes.  Subjects who are not aware of these things
+cannot effectively preserve control and thus lose their privacy.  This
+vulnerability expresses itself as the lack of information material
+like a Privacy Policy including concise information about applied
+processing steps, access rules and disclosure to 3rd parties.
+
+\subsection{Step 2. Potential Analysis}
+
+\subsubsection{Threat Specification}
+
+\input{./figures/Threat-Example}
+
+Recall from the security model description in Section \ref{sec:GrimmModel}, that a threat is a potential interaction of the components that targets an asset.
+We restrict ourselves to the case of attacks, and the asset of privacy.
+An attack is an interaction that is executed by an actor in response to a conflict of interest by exploiting a vulnerability of the system.
+The alternative interaction of accidents are less relevant for us, since the violation of privacy always requires an actor that takes advantage of personal data.
+
+In Figure \ref{figure:Threat-Example} we illustrate the structure of threats at the example of ``Data Theft".
+Here the attack is executed by an external person, that harms the privacy of the citizen.
+He is motivated to do so by financial profit gained by selling personal information, which is in conflict with the citizens interest in his privacy.
+To get hold of the data the external exploits an insecured HTTP transfer of recorded data.
+
+In this section we describe, in a similar fashion, threats for the citizen privacy in the Live+Gov system.
+This list can necessarily not be complete, but we make a best effort to cover the most relevant cases.
+
+% \label{figure:Threats}
+\input{./figures/Threats}
+
+\textbf{T1. Insufficient Control Features.}
+The System Provider does not offer tools for the Citizen to control his data.
+However, besides the missing features he provides a safe and secure system.
+This threat contradicts the definition of privacy as control over one's information about oneself.
+As soon as collected data of Citizens is stored on System Provider servers, all control over that data is lost.
+This is not necessarily due to bad intention, System Providers may simply have forgotten to include such features during the development process.
+Although, control capabilities for Citizens add to the system complexity, which could motivate to omit those. 
+Therefore interests of Citizens and System Providers are in conflict, namely it is the Citizen's \textit{Privacy vs. System Complexity} for System Providers.
+This is an abstract threat provoked by a general \textit{Missing Privacy Awareness} in the minds of all actors in the Live+Gov context.
+
+\textbf{T2. Excessive Data Mining.}
+The System Provider and/or the Local Authority secretly extract more private information from the collected data, than the Citizen agreed to.
+But results of the mining process create no disadvantages for Citizens, because there is no disclosure to third parties.
+This could be the case for a System Provider, who wants to test a new product and uses the pre-existing data collection.
+Or for a Local Authority, who wants to analyze the data collection regarding fare evasion.
+However, the Citizen has not agreed to such data processing nor could he, since it is conducted secretly.
+This disables a Citizen to control his data adequately.
+Thus there are two possible conflicts: \textit{Privacy vs. Financial Profit of Serivce Provider} and  \textit{Privacy vs. Buisness Intelligence of Local Authority}
+The threat can be provoked by either lax data handling policies of both System Providers and Local Authorities, or a weak law enforcement of existing supervision.
+But the main issue, which can lead to such threats, is again a general \textit{Missing Privacy Awareness}.
+
+\textbf{T3. Data Theft}
+An External infiltrates infrastructure in order to steal personal data and sell it on the black market.
+Also the External might be motivated politically and wants to harm the reputation of the System Provider or the Local Authority.
+Anyways, this threat would be manifested through a technical attack on either hardware (Packet Capture) or software (buffer overflow, SQL injection).
+Such a successful attack could harm the reputation of both System Provider and Local Authority.
+The technical competence of System Providers would be proven wrong, therefore they would lose professional reputation.
+Local Authorities would lose their political reputation, as it would seem like they had endangered data of Citizens, which lose complete control.
+Thus this threat is defined by three conflicts: \textit{Privacy vs.
+Financial Profit}, \textit{Reputation of Service Provider vs.
+Political Activism} and \textit{Reputation of Local Authority vs.
+Political Activism}.
+Also this threat describes the classical scenario, where attacks are provoked by \textit{Insecure Infrastructure} (SQL injection) and \textit{Insecure Communication} (Packet Capture).
+
+\textbf{T4. Surveillance}
+
+An External infiltrates infrastructure in order to obtain information
+about the citizen and exploit it directly.  In this scenario the
+external is supposed to have some direct relationship to the citizen
+which motivates his interest to obtain personal information.  Examples
+could be a public institution that wants to gain information about
+planned activities of the citizens (e.g. Nixon's Watergate scandal or
+the recent prosecution of Guardian journalists by GHCQ).  Another
+example is an insurance company that seeks to get information about
+the citizens life-style in relation to the insured risk, like car
+accidents or health hazards.
+
+In this threat the privacy interest of the citizen is in conflict with
+the aspirations for power over the citizen by the externals.
+
+\textbf{T5. Information Leak}
+
+Like an external person the Data Theft Scenario an employee of the
+service provider or the Local Authority has selfish interests to gain
+money, make political statements or harm his employer.  In order to
+pursue this interest he can steal personal data and sell it or release
+it to the public.  The corresponding conflicts of interests are:
+\textit{Privacy vs. Financial Profit} of the Employee, \textit{Reputation of Service
+  Provider vs. Political Activism} of the Employee and \textit{Reputation of Local
+  Authority vs. Political Activism} of the Employee.  The vulnerability constitutes of
+the existence of \textit{Unhappy employees} itself and possibly \textit{lax access
+rules} that enable the employee to obtain large amounts of data
+unnoticed.
+
+\textbf{T6. Social Engineering}
+
+This scenario an external manipulates an employee of a Service
+Provider or the Local Authority to leak information to the external
+person.  It is thus combination of the Data Theft and Information Leak
+scenario.
+The conflicts of interest are \textit{Privacy vs. Financial Profit} of
+the External, \textit{Reputation of Service Provider vs. Political
+  Activism} of the external and \textit{Reputation of Local Authority
+  vs. Political Activism} of the external.  The exploited
+vulnerabilities are, again, the existence of \textit{Unhappy
+  employees} and possibly \textit{lax access rules} that enable the
+employee to obtain large amounts of data unnoticed.
+
+\subsubsection{Threat Risk Evaluation}
+
+In this section we will associate to every identified threat a corresponding risk.
+Recall from \ref{sec:GrimmModel} that a risk is the expected loss that is associated to the threat.
+Therefore, we have to quantify the likeliness of the threat to occure and the harm or loss done in this case.
+The quantification of likeliness will be solely based on rough judgment of the authors.
+The quantification of loss, will be made in a two step process.
+For each threat listed in table \ref{firgure:Threats}, we have analyzed the affected personal data of the citizen.
+For each possible data type (e.g. GPS) we analyze the impact on the seven different types of privacy in Section \ref{sec:SensorPrivacyImpact}.
+In combination we can quantify roughly the impact of each threat on the citizens privacy. Both evaluations are necessarily fraught with a high level of uncertainty.
+
+For the quantification of the loss in case of a threat scenario we use the following rough calibration:
+\begin{itemize}
+\item 3: High. Leak of information to peers (e.g. public) which impacts citizen.
+\item 2: Medium. Undisclosed processing of personal data or disclosure to third parties that are unrelated to subject.
+\item 1: Low. Loss of control over data.
+\item 0: None
+\end{itemize}
+
+For the quantification of likeliness the following scale is used:
+\begin{itemize}
+\item 4: Always.
+\item 3: High. Occurs once in 10 cases
+\item 2: Medium. Occurs once in 100 cases
+\item 1: Low. Occurs once in 1 million cases
+\item 0: Impossible
+\end{itemize}
+
+The quantification of the risk, we add the values for loss and
+likeliness of the corresponding threats. Note, that loss and
+likeliness scales have a logarithmic character, so that that addition
+of those scales corresponds to multiplication of the usual scales.
+
+The likeliness, loss and the resulting risks assigned to the threats
+are discussed in the following paragraphs and summarized in Figure
+\ref{fig:risks}.
+
+\textbf{T1. Insufficient Control Features.}  The occurrence of this
+threat is dependent on the design on the system and given in our case,
+since we do not give the citizen control over his data once it is
+recorded. Therefore the Likeliness is evaluated as $4$ (Always).  The
+associated, risk is $1$ Low on our scale, since no direct harm is done to
+the citizen by exploiting the data.
+
+Hence the resulting risk is calculated as $4+1 = 5$.
+
+\textbf{T2. Excessive Data Mining.}
+We assess the likeliness of excessive data mining to be $3$ High, since
+these kind of analysis can be performed within the walls of the service
+provider, without somebody else noticing, and the service provider himself
+has an interest in this activity.
+
+The associated loss, on the other hand can be substantial (Medium
+$2$).  For example when GPS data is linked to data from telephone
+books the identity of the citizen can be revealed and personal details
+like visits to doctors. In the threat scenario, this information is
+not leaked to third parties, (which would justify an even higher loss
+assessment), but the very existence of this information violates the
+citizens privacy.
+
+Hence the resulting risk is calculated as $3+2 = 5$.
+
+\textbf{T3. Data Theft.}
+The likeliness of a targeted attack by a third party is dependent on
+the popularity of the offered service and financial value of the
+captured information. Moreover, the amount of manual work required to
+infiltrate a custom build system is significantly higher that that of
+compromising a standard software solution. In the scenario we assume a
+moderate popularity in a single metropolitan area, with around 10.000
+users and storage of data of only limited financial value (no
+addresses, no payment information). Therefore the likeliness
+assessment is $1-2$ (Low-Medium).
+
+The harm of leaked information to a criminal party is $3$ High.
+Hence the resulting risk is calculated as $4-5$.
+
+\textbf{T4. Surveillance.}  In the surveillance scenario an party
+related to the citizen, like a company where he is customer of, or a
+government agency, seeks to obtain sensitive information from our
+service.
+
+The likeliness of such an intrusion is hard to asses, and depends
+again on the popularity of the service. If a high popularity is
+reached we have recently learned that spying by government agencies is
+very likely to occur. The barrier for companies that do not operate
+the infrastructure used to transmit the data a surveillance attack is
+however very hard to perform. Therefore we assess the likeliness of
+the threat with $1-2$ (Low-Medium).
+
+The harm of leaked information to a related party is $3$ High.
+Hence the resulting risk is calculated as $4-5$.
+
+\textbf{T5/6. Information Leak and Social Engineering.}
+
+In our scenario we assume that the culture and ethics inside the
+service provider company and local authority are very high, so that
+the information leak scenario has a likeliness of $1$ (Low).
+
+The harm of such an information leaked is $3$ High, so that the
+resulting risk is calculated as $4$.
+
+\begin{figure}
+\centering
+\begin{tabular}{|l|l|l|l|l|}
+\hline
+\textbf{Threat}                   & \textbf{Likeliness} & \textbf{Loss} & \textbf{Risk} & \textbf{Recommendation}
+\\\hline
+T1. Insufficient Control Features & 4                   & 1             & 5             & R1, R2
+\\\hline
+T2. Excessive Data Mining         & 3                   & 2             & 5             & R3, R4, R5
+\\\hline
+T3. Data Theft                    & 1 - 2               & 3             & 4 - 5         & R6
+\\\hline
+T4. Surveillance                  & 1 - 2               & 3             & 4 - 5         & R7
+\\\hline
+T5. Information Leak              & 1                   & 3             & 4             & R8
+\\\hline
+T6. Social Engineering            & 1                   & 3             & 4             & R8
+\\\hline
+\end{tabular}
+\caption{Live+Gov Risk Evaluation and Recommendations}
+\label{fig:risks}
+\end{figure}
+
+\subsubsection{Privacy Recommendations}
+\label{sec:prec}
+
+In the preceding section we have identified the main risks for the
+users privacy.  In this section we derive recommendations or
+requirements for a system that addresses these risks.  Some of these
+requirements are implemented as security measures in our systems and
+discussed in the following chapter \ref{chap:impl}.
+
+In order to address the threat with the highers risk, Insufficient
+Control (T1) of the citizen, we need to give the citizen back the
+control over its data inside the system. The most direct way to do
+this is to provide a web-based \emph{Privacy Dashboard (R1)} which allows
+the citizen to view, edit and delete all information about his person
+that is stored inside the system. Also control applied processing and
+disclosure of the data to third parties should be given to the user,
+at least in the form of an opt-out or veto option.
+
+A necessary pre-requirement for effective control of the citizen over
+his data is information and comprehension of the intended data
+capturing and processing steps. Therefore a \emph{Privacy Policy (R2)} that
+is easily readable and contains all important information is
+essential.  Moreover, the existence of a Privacy Policy is a legal
+requirement (cf. Section \ref{EUDIR}).
+
+The threat with the second largest risk is (T2) Excessive Data Mining.
+Contrary to common belief, it is neither legal nor ethical to process
+personal data for by new methods or for new purposes that were not
+stated and explained to the citizen at the time of data
+collection. Also the common practice of obtaining far-reaching
+permissions from the citizens inside the privacy policy is neither an
+ethical or legal solution to the problem (cf. Art. 6 in Section
+\ref{EUDIR}).
+
+To address this threat awareness about the limitations of data
+processors inside the company is a key element. As one mean to
+establish such a culture of privacy respect, we recommend to prepare
+an document called \emph{Data Handling Guidelines (R3)} intended for
+internal use that explains the concrete processing steps and purposes
+that are permitted by the citizens. This guideline should also be
+structured in a way that it covers the legal notification requirement
+from the EU Data Protection Directive (cf. Section \ref{EUDIR}). In
+particular the following information should be provided for each
+processing task: The name of the controller, purpose of processing,
+description of the data categories, recipients of the data if
+disclosed, transfer to third countries and a description of security
+of processing.
+
+If further processing should be performed, it is necessary to seek
+additional permissions from the citizen. A simple email explaining the
+planned processing steps, and \emph{asking for permission (R4)} would be
+enough for this purpose. The permission can be given via an embedded
+link that shall be followed in order to signal agreement.
+
+An alternative measure to address the risk of excessive data mining is
+the \emph{anonymization (R5)} of data. When all direct- or indirect
+links to the identity of the person are removed, no violation of the
+citizens privacy caused by arbitrary processing. However removing all
+such links is a challenging tasks, and full anonymity is often not
+achieved, cf. \cite{krumm2009}.
+
+The protection from threat scenario (T3) Data Theft is a case of
+classical \emph{IT infrastructure security (R6)}. The storage and processing infrastructure has
+to be secured using firewalls, up-to data software versions and proper
+authentication mechanisms.
+
+The protection from threat scenario (T4) Surveillance focuses on the
+communication channels. They are target of wiretapping attacks by
+intermediaries or externals with access to the communication
+infrastructure. Strong \emph{encryption (R7)} should be used to make
+it harder for externals to read the content of the transmitted data.
+
+Threat scenarios (T5) Information Leak and (T6) Social Engineering
+target the vulnerability of unhappy employees. Therefore a trustful,
+\emph{healthy company culture (R8)} should be maintained.
+
+In summary we recommend the following measures to secure the citizens
+privacy:
+\begin{enumerate}
+\item[R1] Privacy Dashboard. A tool which allows the citizen to view, edit
+  and delete all data personal data that is stored in the system.
+\item[R2] Privacy Policy. A document, that informs the citizen about the
+  collection and processing of personal information. It should at
+  least contain the legally required information.
+\item[R3] Issue Data Handling Guidelines that explain the permitted
+  processing methods and purposes.
+\item[R4] Ask the citizens for permissions before applying further
+  processing via Email.
+\item[R5] Anonymize personal data before processing.
+\item[R6] Securing of Storage and Processing infrastructure using e.g. firewalls.
+\item[R7] Securing communication channels using encryption.
+\item[R8] Maintain a healthy, trustful relationship with your employees.
+\end{enumerate}
+The mapping of these recommendations to the threats is summarized in Figure \ref{fig:risks}.
+\end{LGContent}
+
 %============================================================
@@
 
-\thebibliography
+\TODO{later.\\
+* Key aspect control, can be given back to citizens via ``Privacy Dashboard''
+}
+
+\bibliographystyle{splncs}
+\bibliography{paper}{}
 

### 2014-12-28 paper.tex
removed privacy type inferences (implicit privacy violations)

--- 2014-12-18 llncs/paper.tex
@@
 - Shorten\\
-- Remove Implicit Type Violations\\
+- DONE - Remove Implicit Type Violations\\
 - Update figure\\
@@
 \end{enumerate}
-
-\subsection{Privacy Type Inferences}
-
-The above types of privacy types are not disjoint to each other.
-For example if the location of a citizen is known it is possible to infer information about political associations (e.g. his visits to a party meeting).
-In this section we explore possibilities how personal data can be inferred from each other.
-Our findings are summarized in Figure \ref{figure:Implicit Privacy Violation Matrix}.
-
-%\includegraphics{../diagrams/png/implicit-privacy-violation-matrix.png}
-\input{./figures/Implicit-Privacy-Violation-Matrix}
-
-\paragraph*{1. Privacy of The Person}
-
-The Privacy of The Person is concerned with one's biometric privacy.
-If this type is violated, following implicit violations are possible:
-
-\begin{itemize}
-\item [(1-5)]
-  Privacy of Thoughts and Feelings. Some psychological diseases
-  (e.g.~depression) have physiological impact. Such physiological
-  patterns could be detected.
-\end{itemize}
-
-\paragraph*{2. Privacy of Behaviour and Action}
-
-The Privacy of Behaviour and Action is concerned with one's social, political, religious, sexual, etc. activities.
-If his type is violated, following implicit violations are possible:
-
-\begin{itemize}
-\item [(2-1)] Privacy of The Person.
-  Religious practices which include body modifications (e.g.~circumcision).
-\item [(2-5)] Privacy of Thoughts and Feelings.
-  Social activities in general depend on a certain intellectual attitude.
-  Such an activity is the expressions of such an attitude.
-\item [(2-7)] Privacy of Association.
-  Recording religious, political or sexual activities can reveal association with churches, political parties or sexual partners.
-\end{itemize}
-
-\paragraph*{3. Privacy of Communication}
-
-The Privacy of Communication is concerned with not having such communication (correspondence or vis-a-vis) intercepted.
-This is very broad type of privacy.
-Depending on the contents of the intercepted communication every other type can be violated:
-
-\begin{itemize}
-\item [(3-1)] Privacy of The Person. Communication about body characteristics.
-\item [(3-2)] Privacy of Behaviour and Action. Communication about social activities.
-\item [(3-4)] Privacy of Data and Image.
-  Communication containing one's passwords or other sensitive data.
-\item [(3-5)] Privacy of Thoughts and Feelings.
-  Communication of thoughts and feelings, e.g.~wiretapping a flirt or a catholic confession
-  ritual.
-\item [(3-6)] Privacy of Location and Space.
-  Interception of face-to-face communication is only possible if one's location and space is violated (wiretapping).
-\item [(3-7)] Privacy of Association.
-  Communication about one's associations (family members, churches, etc.).
-\end{itemize}
-
-\paragraph*{4. Privacy of Data and Image}
-
-The Privacy of Data and Image is concerned with one's data not being automatically available to others.
-This also is a very broad type of privacy.
-Depending of the data or image contents every other type can be violated:
-
-\begin{itemize}
-
-\item [(4-1)] Privacy of The Person.
-  Images or stored biometric information reveal one's physical characteristics.
-\item [(4-2)]Privacy of Behaviour and Action.
-  Images or diaries can reveal one's social activities.
-\item [(4-3)]Privacy of Communication.
-  Modern communication systems usually contain some sort of archive function, e.g.~E-mail clients do not automatically delete messages.
-  Such messages are data and reveal one's communication.
-\item [(4-4)]Privacy of Thoughts and Feelings.
-  Images can show one's emotional state.
-\item [(4-5)]Privacy of Location and Space.
-  Images can reveal one's location, e.g.~making a picture in front of the Eifel Tower.
-\item [(4-6)]Privacy of Association. E-mail data can also reveal
-  association.
-\end{itemize}
-
-\paragraph*{5. Privacy of Thoughts and Feelings}
-
-The Privacy of Thoughts and Feelings is concerned with keeping such
-thoughts and feelings secret. If this type is violated, following
-implicit violations are possible:
-
-\begin{itemize}
-\item [(5-1)] Privacy of The Person.
-  Thoughts and feelings can reveal medical conditions.
-\item [(5-2)] Privacy of Behaviour and Action.
-  Thoughts and feelings can reveal a certain attitudes which create a foundation for certain   social activities.
-\item [(5-7)] Privacy of Association.
-  Thoughts and feelings can reveal individual association, e.g amorous feelings for a certain person.
-\end{itemize}
-
-\paragraph*{6. Privacy of Location and Space}
-
-The Privacy of Location and Space is concerned with one's right to move freely without being tracked and one's right to private places.
-If this type is violated, following implicit violations are possible:
-
-\begin{itemize}
-\item [(6-1)] Privacy of The Person.
-  Frequently visited doctors can reveal certain medical conditions, if such doctors are known specialsts.
-  In general it could imply illness.
-\item [(6-2)] Privacy of Behaviour and Action.
-  Frequently visited places in general can reveal association and hence implies social activities.
-\item [(6-3)] Privacy of Communication.
-  If one's location is knwon, it is possible to intercept (wiretap) one's communication.
-  This also may violate the right to private spaces.
-\item [(6-4)] Privacy of Data and Image.
-  If one's location is known, it is possible shoot pictures.
-\item [(6-5)] Privacy of Thoughts and Feelings.
-  Frequently visited persons may imply certain thoughts and feelings, e.g.~having a mistress.
-\item [(6-6)] Privacy of Association.
-  Frequently visited places can reveal associations simply by searching in maps or yellow-pages.
-\end{itemize}
-
-\paragraph*{7. Privacy of Association}
-
-The Privacy of Association is concerned with one's right to associate with whomever one wants, without that association having recorded.
-If this type is violated, following implicit violations are possible:
-
-\begin{itemize}
-
-\item [(7-1)] Privacy of The Person.
-  Association with tattoo artists could imply having tattoos or other body modifications.
-\item [(7-2)] Privacy of Behaviour and Action.
-  Association with churches or political organizations could imply certain activities.
-\item [(7-5)] Privacy of Thoughts and Feelings.
-  Association with churches of political organizations could imply a certain intellectual attitude.
-\end{itemize}
 

### 2014-12-29 paper.tex
started reworking privacy definition (fried)

@@
 { \par }
+
+\usepackage{verbatim}
+% Use this to hide text in TODO commands and LGContent environments.
+%\renewcommand{\TODO}[1]{}
+%\renewenvironment{LGContent}{ \comment  }{ }
+\newenvironment{LGContent-Hidden}{ \comment  }{ }
 
@@
 
-\TODO{MM:\\
-- Shorten\\
-- DONE - Remove Implicit Type Violations\\
-- Update figure\\
-- RG: Discuss alternative definitions of privacy\\
-}
-
-\begin{LGContent}
+%\TODO{MM:\\
+%- Shorten\\
+%- DONE - Remove Implicit Type Violations\\
+%- Update figure\\
+%- RG: Discuss alternative definitions of privacy\\
+%}
+
+\noindent
+Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes:
+\begin{quote}
+The notion of privacy remains out of the grasp of every academic chasing it.
+Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.
+\cite{Gutwirth}
+\end{quote}
+In this section we introduce the concept of privacy as control over information.
+We specify kinds of information we have to deal with in the context of privacy.
+And finally we outline the impact mobile sensors can have on one's privacy.
+
+\subsection{Privacy as Control over Information}
+%We may need to remove the following paragraph depinding on RG's alternative definition of privacy
+%Many definitions seem to focus on the vulnerabilities of privacy.
+%Thus a great amount of work is invested in finding threats to prohibit instead of describing why privacy is so valuable to us.
+%This way around one could derive measures to ensure and secure its value. \cite{7ToP}
+
+%One particular scholar, who did not follow this way of thinking is Charles Fried.
+One possible definition of privacy is given by Charles Fried in his work \textit{An Anatomy of Values}\cite{CFried:Privacy}.
+He questioned, why we are intuitively sensitive to privacy violations.
+But he did not assert privacy as an intrinsic value by itself, he rather stated:
+\begin{quote}
+\textbf{Privacy is} not simply an absence of information about us in the minds of others;
+rather it is \textbf{the control we have over information about ourselves}. \cite{CFried:Privacy}
+\end{quote}
+According to Fried privacy is our ability to create and modulate the relationships to other humans.
+Which makes us intuitively sensitive for its violation, because this affects what defines our identity.
+Also, the ability to build and maintain relationships is essential to society.
+This is what makes privacy valuable and deserving of protection.
+
+\subsubsection{Relationship Creation.}
+
+% Privacy + Trust ~> Relationship
+% ===============================
+% - A shares intimate information with B
+% - A trusts B not to reveal that information by not constantly monitoring B, i.e. respecting B's privacy
+% - B has now the possibility to share intimate information with A
+
+We share information of great intimacy with persons we deem trustworthy we do not share with anyone else. 
+Moreover, we trust those persons to not reveal information about us to others by not constantly monitoring them, i.e. respecting their privacy.
+Trust needs the possibility of unknown failure.
+If we would constantly monitor our partners, they cannot fail unnoticed nor can they willingly share intimate information with us.
+By trusting them as described before, we create the possibility for them to share intimate information with us.
+
+\subsubsection{Relationship Modulation.}
+Fried argues, depending on the person we are interacting with, we are changing the degree of intimacy we share.
+If we are to share private information with others, we are most likely aware that this action is privacy related.
+In that case we are able to selectively disclose information along the two dimensions of quantity and quality.
+
+\subsubsection{Alternative Definitions.}
+So privacy and its possibility, according to Fried, is the ability to build and maintain relationships between humans.
+Thus it is essential to society, \cite{CFried:Privacy}.
+Using his anatomy of privacy as foundation of our analysis is suitable because of two points.
+At first, Fried's study on the understanding of privacy provided a great contribution to the research on the same term in philosophy \cite{sep-privacy} and computer science \cite{SociotechnicalArchitectureForOnlinePrivacy}.
+Secondly, despite the fact his text was published in 1970, he already included technologies to its viewpoint that did not only monitor location, but also record biometric data.
+
+\TODO{RG: Discuss alternative definitions of privacy}
+
+%=======================================================================================
+
+\begin{LGContent-Hidden}
 Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes:
@@
 Secondly, despite the fact his text was published in 1970, he already included technologies to its viewpoint that did not only monitor location, but also record biometric data.
-
-\subsection{The Seven Types of Privacy}\label{sec:privacytypes}
-
+\end{LGContent-Hidden}
+
+\subsection{The Seven Types of Privacy}
+\label{sec:privacytypes}
+
+\begin{LGContent}
 In Fried's definition of privacy as control over "information about onself'', the specification of what constitues such information remains open.
@@
 
-\subsection{Sensor Data Privacy Impact}\label{sec:SensorPrivacyImpact}
-
-\begin{LGContent}
-
-\input{./figures/LG-Implicit-Sensor-Privacy-Matrix}
+\subsection{Sensor Data Privacy Impact}
+\label{sec:SensorPrivacyImpact}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{diagrams/png/implicit-sensor-privacy-matrix.png}
+
+\caption{Live+Gov Implicit Sensor-Privacy Matrix}
+\label{figure:Live+Gov Implicit Sensor-Privacy Matrix}
+\end{figure}
+
+In this section we analyze the impact of the disclosure sensor data and certain processing results to the citizens privacy.
+This analyisis builds upon the preceeding analyis, but is more focused on the concrete type of data vailable.
+For example, the disclosure of Service Line Detection results, does violate the Privacy of Location and Space but not the Privacy of Association that is violated by general GPS tracking.
+
+\subsubsection{GPS Data.}
+The GPS sensor gives the current longitude and latitude, the current global position of the mobile device and its carrier, although there is some artificial inaccuracy within civil use.
+Threefore, the collection of GPS data violates directly the citizens privacy of Location and Space, and the privacy of Data an Image.
+
+By inference we also get implicit violations of the Privacy of the Person, Privacy of Behavior and Action and Privacy of Association.
+
+\subsubsection{Motion Sensors.}
+Accelerometer, Rotation Vector, Gyroscope and Magnetic field sensor measure the physical movement of the mobile device on all three axes.
+If the mobile device is carried ``normally'' its safe to say that those sensors also measure the moments of its carrier.
+So his privacy is infringed regarding biometric behaviour, as it is captured automatically
+Privacy of Data and Image is trivially threatened because here sensor data is individual data, a priori.
+
+By inference we also get implicit violations of the Privacy of Behavior and Action.
+
+\subsubsection{Network Sensors.}
+The GSM and WLAN sensors reveal the position of the mobile device and its carrier, when used in connection with external databases.
+The GSM sensor gives the exact cell, the mobile device has registered with at the current moment.
+
+The Bluetooth sensors record lists of the bluetooth clients in the direct neighbourhood.
+Since those clients are usually moving, inference of the position is usually not possible.
+Instead, bluetooth clients carried by a third person may infringe the Privacy of Association.
+
+A similar argumentat applies to WLAN snsors. If one frequently connects with an organizational wireless network, e.g.~an university network, an association can be deduced (student or staff).
+
+\subsubsection{Human Activity Recognition.}
+The detection and collection of human activities like walking, standing and running, can interfer with the Privacy of the Person, e.g.~since these movement patterns can be indicators for a person's health.
+Also, trivially, Privacy of Data and Image is violated.
+
+\subsubsection{Service Line Detection.}
+The detection and collection of the service line the user currently uses in the public transportation system allows inference of the location of the user, at least when entering and leaving those service lines at e.g.~bus stops.
+Implicit privacy violations apply accordingly.
+
+\begin{LGContent-Hidden}
+
+%\input{./figures/LG-Implicit-Sensor-Privacy-Matrix}
 
@@
 
-\end{LGContent}
+\end{LGContent-Hidden}
 

### 2014-12-30 paper.tex
started subsubsection for alternative definition of privacy

@@
 In this section we introduce the concept of privacy as control over information.
-We specify kinds of information we have to deal with in the context of privacy.
+We specify types of information we have to deal with in the context of privacy.
 And finally we outline the impact mobile sensors can have on one's privacy.
@@
 This is what makes privacy valuable and deserving of protection. 
+\cite{sep-privacy}\cite{CFried:Privacy}
 
@@
 
-We share information of great intimacy with persons we deem trustworthy we do not share with anyone else. 
+We share information of great intimacy only with persons we deem trustworthy. 
 Moreover, we trust those persons to not reveal information about us to others by not constantly monitoring them, i.e. respecting their privacy.
@@
 If we would constantly monitor our partners, they cannot fail unnoticed nor can they willingly share intimate information with us.
-By trusting them as described before, we create the possibility for them to share intimate information with us.
+By trusting them, we create the possibility for them to share intimate information with us.
+\cite{CFried:Privacy}
 
@@
 In that case we are able to selectively disclose information along the two dimensions of quantity and quality.
+\cite{CFried:Privacy}
 
 \subsubsection{Alternative Definitions.}
-So privacy and its possibility, according to Fried, is the ability to build and maintain relationships between humans.
-Thus it is essential to society, \cite{CFried:Privacy}.
-Using his anatomy of privacy as foundation of our analysis is suitable because of two points.
-At first, Fried's study on the understanding of privacy provided a great contribution to the research on the same term in philosophy \cite{sep-privacy} and computer science \cite{SociotechnicalArchitectureForOnlinePrivacy}.
-Secondly, despite the fact his text was published in 1970, he already included technologies to its viewpoint that did not only monitor location, but also record biometric data.
+Using Fried's anatomy of privacy for this analysis is suitable because of two points.
+Despite the fact his text was published in 1970, he already included technologies to its viewpoint that did not only monitor location, but also record biometric data \cite{CFried:Privacy}. 
+Secondly, the concept of control over information does not enforce data minimization.
+Fried's definition complies with what we call \textit{self data protection}, where persons communicate their data self-determined.
+This is the opposite of \textit{system data protection}, where the collection of certain data is prohibited.
 
@@
 
-\begin{LGContent}
+%\begin{LGContent}
 In Fried's definition of privacy as control over "information about onself'', the specification of what constitues such information remains open.
@@
 The seven types of privacy are an extension to the four types of privacy by Roger Clarke \cite{RClarke:4ToP}, which are:
-\begin{itemize}
-\item Privacy of the Person
-\item Privacy of Personal Behaviour
-\item Privacy of Personal Communication
-\item Privacy of Personal Data
-\end{itemize}
+\textit{Privacy of the Person},
+\textit{Privacy of Personal Behaviour},
+\textit{Privacy of Personal Communication},
+\textit{Privacy of Personal Data}.
 It is important to note, that this categories do not form no taxonomy, since the categories are not mutually exclusive.
@@
 
-\end{LGContent}
+%\end{LGContent}
 
@@
 \label{sec:SensorPrivacyImpact}
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{diagrams/png/implicit-sensor-privacy-matrix.png}
-
-\caption{Live+Gov Implicit Sensor-Privacy Matrix}
-\label{figure:Live+Gov Implicit Sensor-Privacy Matrix}
-\end{figure}
 
@@
 For example, the disclosure of Service Line Detection results, does violate the Privacy of Location and Space but not the Privacy of Association that is violated by general GPS tracking.
+
+\begin{figure}
+\newcommand{\tcell}[1]{\parbox[c]{4cm}{\vspace{3mm}#1\vspace{3mm}}}
+\newcommand{\thead}[1]{\tcell{\Large\centering\textbf{#1}}}
+\newcommand{\tbody}[1]{\tcell{\Large\centering #1}}
+\centering
+\resizebox{\textwidth}{!}{
+\begin{tabular}{|c||c|c|c|c|c|c|c|}
+\hline
+& \thead{Pricacy of the Person}
+& \thead{Pricacy of Behaviour and Action}
+& \thead{Pricacy of Communication}
+& \thead{Pricacy of Data and Image}
+& \thead{Pricacy of Thoughts and Feelings}
+& \thead{Pricacy of Location and Space}
+& \thead{Pricacy of Association}
+\\\hline\hline
+\thead{Geo Position}
+& \tbody{0.5} % Pricacy of the Person
+& \tbody{0.5} % Pricacy of Behaviour and Action
+& \tbody{0} % Pricacy of Communication
+& \tbody{1} % Pricacy of Data and Image
+& \tbody{0} % Pricacy of Thoughts and Feelings
+& \tbody{1} % Pricacy of Location and Space
+& \tbody{0.5} % Pricacy of Association
+\\\hline
+\thead{Motion}
+& \tbody{1} % Pricacy of the Person
+& \tbody{0.5} % Pricacy of Behaviour and Action
+& \tbody{0} % Pricacy of Communication
+& \tbody{1} % Pricacy of Data and Image
+& \tbody{0} % Pricacy of Thoughts and Feelings
+& \tbody{0} % Pricacy of Location and Space
+& \tbody{0} % Pricacy of Association
+\\\hline
+\thead{Networking}
+& \tbody{0.5} % Pricacy of the Person
+& \tbody{0.5} % Pricacy of Behaviour and Action
+& \tbody{0} % Pricacy of Communication
+& \tbody{1} % Pricacy of Data and Image
+& \tbody{0} % Pricacy of Thoughts and Feelings
+& \tbody{1} % Pricacy of Location and Space
+& \tbody{1} % Pricacy of Association
+\\\hline
+%\thead{HAR}
+%& \tbody{-} % Pricacy of the Person
+%& \tbody{-} % Pricacy of Behaviour and Action
+%& \tbody{-} % Pricacy of Communication
+%& \tbody{-} % Pricacy of Data and Image
+%& \tbody{-} % Pricacy of Thoughts and Feelings
+%& \tbody{-} % Pricacy of Location and Space
+%& \tbody{-} % Pricacy of Association
+%\\\hline
+%\thead{SLD}
+%& \tbody{-} % Pricacy of the Person
+%& \tbody{-} % Pricacy of Behaviour and Action
+%& \tbody{-} % Pricacy of Communication
+%& \tbody{-} % Pricacy of Data and Image
+%& \tbody{-} % Pricacy of Thoughts and Feelings
+%& \tbody{-} % Pricacy of Location and Space
+%& \tbody{-} % Pricacy of Association
+%\\\hline
+\end{tabular}
+}
+\newline
+\newline
+{\scriptsize 
+0: No Impact,
+0.5: Light Impact,
+1: Hard Impact}
+\caption{Sensor Data Privacy Impact Matrix}
+\label{figure:Sensor Data Privacy Impact Matrix}
+\end{figure}
 
@@
 
-\subsubsection{Human Activity Recognition.}
-The detection and collection of human activities like walking, standing and running, can interfer with the Privacy of the Person, e.g.~since these movement patterns can be indicators for a person's health.
+%\subsubsection{Human Activity Recognition.}
+%The detection and collection of human activities like walking, standing and running, can interfer with the Privacy of the Person, e.g.~since these movement patterns can be indicators for a person's health.
 Also, trivially, Privacy of Data and Image is violated.
-
-\subsubsection{Service Line Detection.}
-The detection and collection of the service line the user currently uses in the public transportation system allows inference of the location of the user, at least when entering and leaving those service lines at e.g.~bus stops.
-Implicit privacy violations apply accordingly.
+%
+%\subsubsection{Service Line Detection.}
+%The detection and collection of the service line the user currently uses in the public transportation system allows inference of the location of the user, at least when entering and leaving those service lines at e.g.~bus stops.
+%Implicit privacy violations apply accordingly.
 

### 2014-12-31 paper.tex
finished shortening of sensor privacy impact ouline

@@
 
-%\begin{LGContent}
+In Fried's definition of privacy as control over information, the specification of what constitues such information remains open.
+There is a vast amount of information that relates to a person and we need to get a better understanding in order to perform a thorough analysis.
+To this end we use the the categorization by Friedewald, Finn and Wright \cite{7ToP} called the \textit{Seven Types of Privacy}.
+The seven types of privacy are an extension to the four types of privacy by Roger Clarke \cite{RClarke:4ToP}, which are:
+\textit{Privacy of the Person},
+\textit{Privacy of Personal Behaviour},
+\textit{Privacy of Personal Communication},
+\textit{Privacy of Personal Data}.
+It is important to note, that these categories do not form a taxonomy, since the categories are not mutually exclusive.
+For instance a written email is considered personal communication as well as personal data stored on a computer.
+
+Moreover, Friedewald et al. argue that Clarke's types are outdated and no longer adequate in order to describe the privacy aspect of our modern, technology driven, world.
+In order to address this shortcoming they extend the former four to the now introduced seven types privacy as follows:
+
+\begin{enumerate}
+
+\item \textbf{Privacy of the Person}
+This privacy type is generally concerned with one could best understand as biometric privacy.
+Friedewald et al. paraphrase it as \emph{``\om the right to keep body functions and body characteristics \om private''}.
+This includes but is not limited to measures like weight, height or shoulder width;
+biometric identifiers like fingerprints and DNA sequences;
+or medical conditions such as limping or having a cold.
+
+\item \textbf{Privacy of Behaviour and Action}
+This privacy type is concerned with one's activities in public as well as in private spaces.
+It includes but is not limited to religious practices, political activities and sexual preferences or habits.
+
+\item \textbf{Privacy of Communication}
+This privacy type is concerned with one's communication in a broad sense.
+It includes written correspondence, but also conversations conducted either vis-a-vis or via electronic devices.
+Friedewald et al. put it as the right to free discussion without unknown interception by third parties.
+
+\item \textbf{Privacy of Data and Image}
+This privacy type is concerned with the secrecy of personal data, especially its automatic disclosure to other individuals and organizations.
+It includes data such as paychecks, insurance information or records of public administration.
+However, it also refers to pictures taken without consent and digital identifiers like IP addresses or social security numbers.
+
+\item \textbf{Privacy of Thoughts and Feelings}
+This privacy type is the counterpart to Privacy of the Person like body and mind are counterparts of one another.
+Comparable to the Privacy of Data and Image, Friedewald et al. state that one's thoughts and feelings must not be automatically revealed to others.
+This could simply happen by the disclosure of one's diary or by technologies which allow emotion detection through biometric means.
+One's body temperature or iris reflexes might infer stress or excitation.
+
+\item \textbf{Privacy of Location and Space}
+This privacy type is concerned with one's movements in public spaces and the protection of private spaces.
+Friedewald et al. qualify the first dimension as one's right to move without being identified, tracked or monitored.
+The second dimension is qualified as one's general right to solitude, especially the right to the inviolability of the home.
+
+\item \textbf{Privacy of Association}
+This privacy type is also put as group privacy.
+Friedewald et al. state that one must have the possibility with whomever without being recorded.
+Associations like friends or organizations such as political parties must not automatically be recorded because one associates with them, and vice-versa.
+\end{enumerate}
+
+\begin{LGContent-Hidden}
 In Fried's definition of privacy as control over "information about onself'', the specification of what constitues such information remains open.
@@
 
-%\end{LGContent}
+\end{LGContent-Hidden}
 
@@
 
-In this section we analyze the impact of the disclosure sensor data and certain processing results to the citizens privacy.
-This analyisis builds upon the preceeding analyis, but is more focused on the concrete type of data vailable.
-For example, the disclosure of Service Line Detection results, does violate the Privacy of Location and Space but not the Privacy of Association that is violated by general GPS tracking.
+Modern mobile devices have a broad collection of sensors.
+Disclosure or processing of sensor data can impact one's privacy.
+In this section we identify groups of sensors and their potential impact on a certain type of privacy.
+Figure \ref{figure:Sensor Data Privacy Impact Matrix} qualifies that impact on a simple scale.
+Privacy of Data and Image is trivially threatened because here sensor data is individual data, a priori.
+Indirect impact is caused by combining sensor data with additional knowledge.
+For instance, comparing a contemporary map with locational data can imply behaviour if the position matches a church.
 
 \begin{figure}
-\newcommand{\tcell}[1]{\parbox[c]{4cm}{\vspace{3mm}#1\vspace{3mm}}}
+\newcommand{\tcell}[1]{\parbox[c][2cm]{4cm}{\vspace{3mm}#1\vspace{3mm}}}
 \newcommand{\thead}[1]{\tcell{\Large\centering\textbf{#1}}}
@@
 \\\hline\hline
-\thead{Geo Position}
+\thead{GPS Sensor}
 & \tbody{0.5} % Pricacy of the Person
@@
 \\\hline
-\thead{Motion}
+\thead{Motion Sensors}
 & \tbody{1} % Pricacy of the Person
@@
 \\\hline
-\thead{Networking}
+\thead{Networking Sensors}
 & \tbody{0.5} % Pricacy of the Person
@@
 0: No Impact,
-0.5: Light Impact,
-1: Hard Impact}
+0.5: Indirect Impact,
+1: Direct Impact}
 \caption{Sensor Data Privacy Impact Matrix}
@@
 
-\subsubsection{GPS Data.}
+\subsubsection{GPS Sensor.}
 The GPS sensor gives the current longitude and latitude, the current global position of the mobile device and its carrier, although there is some artificial inaccuracy within civil use.
-Threefore, the collection of GPS data violates directly the citizens privacy of Location and Space, and the privacy of Data an Image.
-
-By inference we also get implicit violations of the Privacy of the Person, Privacy of Behavior and Action and Privacy of Association.
+Therefore, the collection of GPS data violates directly the citizens privacy of Location and Space.
+
+%By inference we also get implicit violations of the Privacy of the Person, Privacy of Behavior and Action and Privacy of Association.
 
@@
 If the mobile device is carried ``normally'' its safe to say that those sensors also measure the moments of its carrier.
-So his privacy is infringed regarding biometric behaviour, as it is captured automatically
-Privacy of Data and Image is trivially threatened because here sensor data is individual data, a priori.
-
-By inference we also get implicit violations of the Privacy of Behavior and Action.
+So his privacy is infringed regarding biometric behaviour, i.e. the Privacy of the Person.
+
+%By inference we also get implicit violations of the Privacy of Behavior and Action.
 
@@
 The GSM and WLAN sensors reveal the position of the mobile device and its carrier, when used in connection with external databases.
-The GSM sensor gives the exact cell, the mobile device has registered with at the current moment.
+The both sensors give the exact cell or network, the mobile device has registered with at the current moment.
+Frequent connection to one particular network also reveals association, e.g. university networks.
 
@@
 
-A similar argumentat applies to WLAN snsors. If one frequently connects with an organizational wireless network, e.g.~an university network, an association can be deduced (student or staff).
+%A similar argumentat applies to WLAN snsors. 
+%If one frequently connects with an organizational wireless network, e.g.~an university network, an association can be deduced (student or staff).
 

### 2015-01-07 paper.tex
resized figures

@@
 
-\input{./figures/IT-Systems}
+%input{./figures/IT-Systems}
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.7\textwidth]{diagrams/png/it-systems.png}
+
+\begin{flushleft}
+\scriptsize
+\textbf{Legend:}
+\begin{itemize}
+\itemsep1pt\parskip0pt\parsep0pt
+
+\item
+\textbf{(a) Citizen:} 
+User of the L+G client application whose privacy is at stake.
+
+\item
+\textbf{(b) Service Provider:} Provides technical infrastructure.
+
+\item
+\textbf{(c) Local Authority:} Provider of the L+G system.
+
+\item
+\textbf{(d) Mobile Device:}
+Runs the L+G client application, produces and stores data sensitive to the users privacy.
+
+\item
+\textbf{(e) Data Center:} Runs the L+G services, processes and stores user data.
+
+\item
+\textbf{(f) Report Tool:} Interface to aggregated user data.
+
+\end{itemize}
+\end{flushleft}
+
+\caption{IT Systems}
+\label{figure:IT Systems}
+\end{figure}
 
@@
 
-%\input{./figures/LG-Implicit-Sensor-Privacy-Matrix}
+%input{./figures/LG-Implicit-Sensor-Privacy-Matrix}
 
@@
 %\includegraphics{../diagrams/png/itsec-ref-model-grimm.png}
-\input{./figures/The-IT-Security-Reference-Model-Grimm}
+%input{./figures/The-IT-Security-Reference-Model-Grimm}
+\begin{figure}
+\centering
+\includegraphics[width=0.6\textwidth]{diagrams/png/itsec-ref-model-grimm.png}
+\caption{The IT Security Reference Model (Grimm)}
+\label{fig:refmodel}
+\end{figure}
 
@@
 
-%\includegraphics{../diagrams/png/itsec-ref-model-grimm-procedure.png}
-\input{./figures/The-IT-Security-Reference-Analysis-Grimm}
+%includegraphics{../diagrams/png/itsec-ref-model-grimm-procedure.png}
+%input{./figures/The-IT-Security-Reference-Analysis-Grimm}
+\begin{figure}
+\centering
+\includegraphics[width=0.6\textwidth]{diagrams/png/itsec-ref-model-grimm-procedure.png}
+\caption{The IT Security Reference Analysis (Grimm)}
+\end{figure}
 
@@
 
-\input{./figures/LG-Conflicts-Of-Interests}
+%input{./figures/LG-Conflicts-Of-Interests}
+\begin{figure}
+\centering
+\includegraphics[width=0.6\textwidth]{diagrams/png/conflict-of-interests.png}
+\caption{Live+Gov Conflicts of Interests}
+\label{figure:Live+Gov Conflicts of Interests}
+\end{figure}
 %TODO: Adapt Figure
@@
 
-\input{./figures/LG-Vulnerabilities}
+%input{./figures/LG-Vulnerabilities}
+\begin{figure}
+\centering
+\includegraphics[width=0.6\textwidth]{diagrams/png/vulnerabilities.png}
+
+%\begin{flushleft}
+%\scriptsize
+%\textbf{Legend:}
+%
+%\begin{itemize}
+%\itemsep1pt\parskip0pt\parsep0pt
+%\item
+%  \textbf{(a) Citizen:} User of the L+G client application whose privacy
+%  is at stake.
+%\item
+%  \textbf{(b) Mobile Device:} Runs the L+G client application, produces
+%  and stores data sensitive to the users privacy.
+%\item
+%  \textbf{(c) L+G Data Center:} Runs the L+G services, proccesses and
+%  stores user data.
+%\item
+%  \textbf{(d) L+G Maintenance Crew:} Technical staff with access to
+%  critical infrastructure.
+%\item
+%  \textbf{(e) Aggregation Monitor:} Interface to aggregated user data.
+%\item
+%  \textbf{(f) Local Authority:} Provider of the L+G system.
+%\item
+%  \textbf{(g) Criminal:} Threatens the L+G system and Citizen privacy.
+%\end{itemize}
+%\end{flushleft}
+
+\caption{Live+Gov Vulnerabilities}
+\label{figure:Live+Gov Vulnerabilities}
+\end{figure}
 
@@
 
-\input{./figures/Threat-Example}
+%input{./figures/Threat-Example}
+\begin{figure}
+\centering
+\includegraphics[width=0.6\textwidth]{diagrams/png/threat-example.png}
+\caption{Threat Example}
+\label{figure:Threat-Example}
+\end{figure}
 
@@
 % \label{figure:Threats}
-\input{./figures/Threats}
+%input{./figures/Threats}
+\begin{figure}
+\centering
+%\includegraphics[width=\textwidth]{diagrams/png/threat-table.png}
+\newcommand{\tcell}[1]{\parbox[c]{6cm}{#1\vspace{3mm}}}
+\newcommand{\thead}[1]{\tcell{\vspace{3mm}\centering\textbf{#1}}}
+\newcommand{\tbody}[1]{\tcell{\footnotesize\flushleft #1}}
+\resizebox{\textwidth}{!}{
+\begin{tabular}{|c|c|c|c|}
+
+\hline
+\thead{Threat}
+& \thead{Conflict of Interest}
+& \thead{Vulnerability}
+& \thead{Affected Data}
+\\\hline\hline
+\tbody{Insufficient Control Features}
+& \tbody{Privacy vs. System Complexity}
+& \tbody{Missing Privacy Awareness}
+& \tbody{All collected data}
+\\\hline
+
+\tbody{Exessive Data Mining}
+& \tbody{Privacy vs. Business Intelligence (LA),
+\\Privacy vs. Financial Profit (SP)}
+& \tbody{Lax data handling policies
+\\Missing Privacy Awareness
+\\Weak law enforcement}
+& \tbody{Mined Information}
+\\\hline
+
+\tbody{Data Theft}
+& \tbody{Privacy vs. Financial Profit
+\\Reputation (SP) vs. Political Activism
+\\Reputation (LA) vs. Political Activism}
+& \tbody{Insecure Infrastructure,
+\\Insecure Communication}
+& \tbody{Transmitted data 
+\\Data stored in data center}
+\\\hline
+
+\tbody{Surveillance}
+& \tbody{Privacy vs. External Power}
+& \tbody{Insecure Infrastructure
+\\Insecure Communication}
+& \tbody{Transmitted data 
+\\Data stored in data center}
+\\\hline
+
+\tbody{Information Leak}
+& \tbody{Privacy vs. Financial Profit
+\\Reputation (SP) vs. Political Activism
+\\Reputation (LA) vs. Political Activism}
+& \tbody{Unhappy Employee 
+\\Lax Access Rules}
+& \tbody{Data stored in data center}
+\\\hline
+
+\tbody{Social Engineering}
+& \tbody{Privacy vs. Financial Profit
+\\Reputation (SP) vs. Political Activism
+\\Reputation (LA) vs. Political Activism}
+& \tbody{Unhappy Employee
+\\Lax Access Rules}
+& \tbody{Data stored in data center}
+\\\hline
+\end{tabular}
+}
+\caption{Threats}
+\label{figure:Threats}
+\end{figure}
 

### 2015-01-15 paper.tex
- merged section 4+5 to 4

@@
 \newcommand{\om}{[...]\xspace} % omission
+
+\authorrunning{H. Hartmann, T. Wambach, M. Meffert, R. Grimm}
 
@@
 \section{Definition of Privacy}
+\label{sec:privacydef}
 
@@
 
-\section{IT Security Reference Model}
-
-\TODO{TW}
-
-\begin{LGContent}
-
-The goal of this chapter to analyze and identify the threads to
+\section{Privacy Analysis}
+
+The goal of this chapter is to analyze and identify the threads to
 personal privacy that are posed by collecting, storing and processing
@@
 
-The complexity of our systems and the variety of threads make a great
-number of counter measures plausible.  We approach this complexity
-with the aid of a general security analysis model developed in
-\cite{Grimm:ItSecRefModel}.  We give a brief introduction to this
-model and perform a IT Security Analysis with respect to the privacy
-asset for our system.
-
-We follow the Reference Model for IT Security Analysis as described in
-\cite{Grimm:ItSecRefModel}.  It supersedes earlier efforts by
+In section \ref{subsec:world-analysis} and \ref{subsec:pot-analysis} we apply the first two steps of the Reference Model for IT Security Analysis as described in \cite{Grimm:ItSecRefModel}.
+It supersedes earlier efforts by
 e.g. \cite{Avizienis}.  The reference model consists of a \emph{model}
-and a \emph{procedure}.  The model aims to organize common security
-terminology in a reasonable and practical way.  The procedure
-describes a method to analysis the IT system based on that model.  In
-this section we give a brief overview over the reference model.
-
-\subsection{Model}
-
-%\includegraphics{../diagrams/png/itsec-ref-model-grimm.png}
-%input{./figures/The-IT-Security-Reference-Model-Grimm}
-\begin{figure}
-\centering
-\includegraphics[width=0.6\textwidth]{diagrams/png/itsec-ref-model-grimm.png}
-\caption{The IT Security Reference Model (Grimm)}
-\label{fig:refmodel}
-\end{figure}
-
-The model is depicted in Figure \ref{fig:refmodel}.  It is organized
-in four views (round boxes) that contain a number of components
-(rectangular boxes).
-
-%%%%%%%%%%%%%%%%%%%%
+and a \emph{procedure}: The model aims to organize common security terminology in a reasonable and practical way, the procedure describes a method to analysis the IT system based on that model.
+In section \ref{subsec:priv-rec} we provide specific privacy recommendations as a result of the previous analysis that the operated system must comply.
+
+\subsection{Step 1. World Analysis}
+
+\label{subsec:world-analysis}
 
@@
 \begin{itemize}
+
+	\item
+	\textbf{Assets.} Things of value to one or more stakeholders.
+
+	\item \textbf{IT-Systems.}
+	The relevant IT-Systems under study. Already described in Section \ref{subsubsection:it-systems}.
+
 	\item \textbf{Actors.}
 	All identifiable stakeholders of the system under study.
-Typical actors include, users, developers, clients and externals.
-
-\item
-\textbf{Assets.} Things of value to one or more stakeholders.
-The value can be hard (money, data, etc.) or soft (trust, privacy,etc.).
-In our case the only asset we are concerned with is the privacy of the citizen.
-
-\item \textbf{IT-Systems.}
-The relevant IT-Systems under study. This encompasses hardware (e.g. servers, network infrastructure), as well as software and third party services.
-The level of granularity has to be detailed enough to express all possible threats to the assets at stake.
 
@@
 	Different actors have different interests which can be in conflict which each other.
-These conflicts of interest are the origin of all attacks to the system.
-
-A typical conflict is the \emph{Criminal-User-Conflict}:
-A user wants to keep control over their private data.
-A criminal wants to gain money.
-The possibility of selling private data (user profiles) to advertisers, renders both interests conflicting.
 
 	\item \textbf{Vulnerabilities.}
-All identifiable weaknesses in the IT-System.
-
-In the example of the criminal-user-conflict, the criminal has to exploit a vulnerability, e.g. a weak password, to gain access to the private data about the user.
+	All identifiable weaknesses in the IT-System e.g. a weak passwords.
 
 	\item \textbf{Interactions.}
-This point captures all possible interactions between assets, IT-Systems, humans and vulnerabilites.
-It is described in more detail in the next view.
+	This point captures all possible interactions between assets, IT-Systems, humans, and vulnerabilites as partly described in Section \ref{subsubsection:it-systems} and will be further analyzed in \ref{subsec:pot-analysis}.
+
 \end{itemize}
 
-%%%%%%%%%%%%%%%%%%%%
-
-The \emph{potential view} displays the intended and unintended interactions of the components in the world view.
+\subsubsection{Assets}
+
+In this scenario we focus our attention to only one asset: The privacy of the citizen.
+Our definition of privacy is described in detail in Section \ref{sec:privacydef} and introduced in Section \ref{sec:privacytypes} the seven different types of privacy.
+
+\subsubsection{Actors}
+\label{subsubsection:humans}
+
+This section describes the human actors previously introduced within the IT system architecture (\ref{subsubsection:it-systems}).
+Also the additional actor \textit{External} is introduced as a person with no special access rights.
+Although the text is only a short description, a list of the most important interests is provided for each actor.
+
+\textbf{Citizen.}
+Citizens are persons who use the mobile device as users of the provided software.
+Their main motivation for using the software is to gain a higher level of convenience in their daily activities.
+For example they may have access to real time bus schedules or reports about traffic jams, that help them to avoid long waiting times.
+Also they generally benefit from improvements of public infrastructure by local authorities, which is triggered by issue reports.
+
+By using the application, citizens are sharing personal information like name and address, as well as data gathered from mobile sensor with the service provider.
+This data can be exploited in ways that are harmful to the citizen (cf.
+\cite{GuardienMassSurveillance}).
+
+Citizens are interested in: physical wellbeing and health, financial profit, convenience, legitimate use of personal data, non-disclosure of personal data to peers of the citizen, and not being monitored.
+
+\textbf{External.}
+Externals are persons who do not have privileged access to the IT systems, and are willing to break laws, security constrains and norms in order to promote their interests.
+
+A common interest of an external is financial profit.
+For example they want to obtain access to critical systems to steal sensitive data or to get the system under their control.
+Stolen data could simply be sold as is or used for illegitimate purposes, e.g.
+spam or phishing attacks - or excessive data mining.
+
+In short, externals could be interested in: increase power over citizen, financial profit, political activism, and their social standing.
+
+\textbf{System Provider.}
+System Providers operate the technical infrastructure (hardware and software) of the IT System.
+They are private companies and legal persons in their own right, but also employ a number of people with diverging interests including administrators, programmer/developer, and a support manager.
+
+As companies, they are interested in gaining financial profit.
+The financial success of system providers depends on the task complexity of the maintained infrastructure.
+The complexity of a task has to be in reasonable bounds, so that system providers can complete it within time, with a satisfying quality.
+
+System Providers are interested in: financial profit, manageable complexity, professional excellence, and good working conditions.
+
+\textbf{Local Authority.}
+Local authorities are public offices (ministry, agency, department, ...) or other external public entities which act as direct customers of service providers.
+They purchase a system specialized for their needs.
+For example a department for urban mobility, orders a system to better understand usage patterns and make improvement to the urban traffic flow.
+Such systems are investments, and so naturally local authorities are interested in a profitable return, like increased ticket sales.
+
+The first important step of improving the public services is by obtaining business intelligence.
+For the urban mobility scenario the Live+Gov systems provide insight in form of traffic jam detection and usage pattern mining, which allow local authorities to focus their efforts to the most important sites.
+
+Local authorities are interested in: financial profit, political reputation, business intelligence, and good working conditions.
+
+\subsubsection{Conflicts of Interests}
+\label{subsubsection:Conflicts of Interests}
+Different actors have different interests which can be in conflict each other. This section outlines the Conflicts of Interests between the actors of the proposed IT system architecture.
+
+The individual interests of all actors is already described in the previous section and are not elaborated any further.
+The emphasis here is put on prominent existing conflicts, because they provide a foundation for vulnerabilities and subsequent threats.
+
+\textbf{System Complexity vs Privacy.}
+System Providers offer a service to Local Authorities, so that Local Authorities can improve their public services.
+This task in it self has a high technical complexity and operates on privacy sensitive data provided by monitored Citizens.
+
+\textbf{Business Intelligence vs Privacy.}
+Local authorities order a monitoring and mining system from system providers as much as possible, which allows them to produce business intelligence for public services.
+They are interested in the successful usage of their data, although the interest of citizens lies in maintaining control over their data and protecting their rights to privacy.
+
+\textbf{Power of External vs. Privacy.}
+Externals which are in a social relation to the citizen can have an interest in obtaining further information in order to gain power.
+In the most simplistic example this could be a man wanting monitor the activities of his spouse.
+
+\textbf{Financial Profit of External vs Privacy.}
+Externals can gain financial profit from stealing privacy sensitive data.
+For example by selling raw contact information to advertisers or by selling mined data to insurance companies, or intermediaries like scoring companies.
+In such cases, citizens lose complete control over their data.
+
+\textbf{Financial Profit of External vs Reputation of System Providers.}
+Externals have various business models as optional foundation for attacks on System Providers.
+For instance, they can try to invade the infrastructure for e-espionage reasons, to get control over servers to create a bot network or to steal user data.
+A successful attack proves the technical competence of system providers wrong and subsequently harms their professional reputation.
+
+\textbf{Political Activism vs Reputation of Local Authority.}
+Besides monetary reasons, externals can be motivated by political reasons to attack the monitoring and mining system.
+Externals can break the system to make a political statement of their own,
+or they can steal user data to prove the system insecure.
+Both would harm the reputation of local authorities, who endangered the privacy of the citizens.
+
+\subsubsection{Vulnerabilities}
+\label{subsubsection:Vulnerabilities}
+This section outlines the vulnerabilities (Figure \ref{figure:Live+Gov Vulnerabilities}) of the proposed monitoring and mining system.
+Note that vulnerabilities are not necessarily of technical nature.
+The weaknesses of IT systems are often created due to misuse or misconfiguration of the various components by one or more actors.
+
+%input{./figures/LG-Vulnerabilities}
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{diagrams/png/vulnerabilities.png}
+
+\caption{Overview of vulnerabilities of the Live+Gov System; legend see \ref{figure:IT Systems} }
+\label{figure:Live+Gov Vulnerabilities}
+\end{figure}
+
+\textbf{Insecure Infrastructure.}
+The proposed monitoring system consists of many hardware and software components, each with its own concrete weaknesses.
+For instance, operating systems can be outdated or not subject to frequent updates or virus scans.
+
+\textbf{Insecure Data Transmission.}
+The proposed monitoring and mining system uses HTTP to exchange data between the Sensor Collector, the Data Center and the Report Tool.
+Data can be intercept and read all sensitive information, which is send between the components e.g. passwords, raw sensor data and data mining results.
+
+\textbf{Unhappy Employees.}
+An Employee that is frustrated with his situation for a long time
+period constitutes a security vulnerability. On the one hand he might
+want to harm his employer directly, on the other he is increasingly
+susceptible for social engineering.
+
+\textbf{Inadequate Access Rules.}
+The proposed IT system infrastructure has various accesses to privacy sensitive data.
+System Provider staff has access to Data Center hardware and software like databases, web-servers and other inspection tools.
+Local Authority staff has access to the Report Tool.
+This all enables staff members to have potential access to privacy sensitive information.
+
+\textbf{Unaware Monitoring Subjects.}
+We define privacy as one's ability to control information about
+oneself.  In order to do that, monitored subjects need to know, that
+they are monitored, who monitors them, what information is recorded
+and for what purposes.  Subjects who are not aware of these things
+cannot effectively preserve control and thus lose their privacy.
+
+\subsection{Step 2. Potential Analysis}
+\label{subsec:pot-analysis}
+
+The \emph{Potential Analysis} (or \emph{Potential View}) displays the intended and unintended interactions of the components in the world view.
 The intended interactions support the underlying business objectives.
-Unintended interactions lead to threats.
-The potential view consists of the following components.
+Unintended interactions lead to threats. The potential view consists of the following components:
+
 \begin{itemize}
+
 	\item \textbf{Business Objectives.}
-Interaction of IT-system and actors that realize a business goals of the system owner.
+	Interaction of IT-system and actors that realize a business goals of the system owner were already described in Section \ref{subsubsection:it-systems} and \ref{subsec:world-analysis}.
 
@@
 	\item \textbf{Chances/Risk.}
-Evaluation of chances and risks associated to the business objectives and threats.
-The risk associated to a threat is its expected loss.
-A chance associated to a intended interaction is its expected gain.
-
-In the case that, the loss can be quantified monetary, and the likelihood of occurrence of a threat can
-be modeled probabilistically, the risk is given by the product
-\[ \text{risk} = \text{loss} \cdot P[\text{threat}]. \]
-In practice such a quantitative risk evaluation is often not possible, and a qualitative, heuristic, analysis is performed instead.
+	Evaluation of chances and risks associated to the business objectives and threats. The risk associated to a threat is its expected loss.
 
 	\item \textbf{Security Requirements.}
-A set of interactions (e.g. threats) that shall not occur within the system in order to achieve its business objectives.
-Security requirements are targeted to protect one or more assets.
-
-An example of a security requirement is that a given communication channel shall not be infringed by externals.
+	Interactions (e.g. threats) that shall not occur within the system in order to achieve its business objectives.
+
 \end{itemize}
 
-%%%%%%%%%%%%%%%%%%%%
-
-The \emph{concept view} is a realization of the potential view of the system.
-It specifies important interactions that require further planning.
-It contains the following components:
-\begin{itemize}
-\item \textbf{Business Model.}
-The plan to achieve business objectives.
-
-\item \textbf{Accident Scenario.}
-A concrete outline of an interaction that leads to an accident.
-In particular the asset under threatened and exploited vulnerability need to be described.
-
-\item \textbf{Attack Technique.}
-A specific technique or technology to attack IT-Systems (Man in the Middle, Phishing, etc.).
-In particular the attacking actor, the conflict of interest and the exploited vulnerability need to be described.
-
-\item \textbf{Security Measures.}
-It describes a plan of sufficient measures to secure the intended interactions and to avoid the unintended interactions.
-Each security measure targets a vulnerability of the system in order to reduce a risk for a certain thread.
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%
-
-The \emph{event view} contains all actual events through out the lifetime of the system.
-The event view instantiates the concept view of the system.
-It contains the following components:
-\begin{itemize}
-\item \textbf{Business Process.}
-The actual, running instance of the business model.
-
-\item \textbf{Accidents.}
-All actually happened accidents.
-
-\item \textbf{Attacks.}
-All actually happened attacks.
-
-\item \textbf{Security Operations.}
-Instances of security measures.
-\end{itemize}
-
-\subsection{Procedure}
-
-The analysis procedure is an incremental and iterative process following the four views of the previously described model.
-
-%includegraphics{../diagrams/png/itsec-ref-model-grimm-procedure.png}
-%input{./figures/The-IT-Security-Reference-Analysis-Grimm}
-\begin{figure}
-\centering
-\includegraphics[width=0.6\textwidth]{diagrams/png/itsec-ref-model-grimm-procedure.png}
-\caption{The IT Security Reference Analysis (Grimm)}
-\end{figure}
-
-\paragraph*{Step 1. World Analysis}
-
-At first, one has to outline the current state of the system under study. This includes description of:
-\begin{itemize}
-\item all \textbf{Assets} which must be protected
-\item all relevant \textbf{IT-Systems}
-\item all involved \textbf{Humans} and their \textbf{Conflicts of Interests}
-\item all known \textbf{Vulnerabilities}
-\item and all important \textbf{Interactions} between the former components.
-\end{itemize}
-
-\paragraph*{Step 2. Potential Analysis}
-
-Secondly, one needs to outline the potential interactions of the system under study.
-This includes both the unintended interactions (threads) and the intended interactions (business objectives).
-This step produces four artifacts:
-\begin{itemize}
-\item a \emph{threat specification}, which identifies the threat, its targeted assets, the involved actors and their conflicts of interest.
-\item a \emph{threat risk evaluation}, which quantifies the likelihood of a threat manifestation in relation to its associated loss.
-\item a \emph{security requirement specification}, which specifies requirements in order to deal with identified hazards
-\end{itemize}
-
-\paragraph*{Step 3. Concept Development}
-
-Based on \textbf{Step 2.}, the identified hazards are used alongside realistic accident scenarios and attack techniques to create a \emph{risk matrix}.
-With this matrix it is possible to decide if the risk is acceptable or not.
-Together with the previously specified security requirements, the matrix is used to define adequate security measures.
-Like a business model is an abstract concept to achieve business objectives, this step creates an concept to improve the system's security.
-
-\paragraph*{Step 4. Concept Deployment}
-
-Finally, the security measures have to be implemented.
-Additionally, all business operations, accidents, attacks and executed security operations will be recorded in the following time.
-
-The implementation of security measures changes the world view (e.g. IT-systems, actors) and renders the conducted analysis outdated.
-So this analysis procedure needs to be conducted again.
-
-\subsection{Abstraction Levels of the Reference Model}
-
-The Reference Model can be used on different levels of abstraction.
-This means each component can be used within a wide range of granularity, for instance the security measure \emph{Encryption} can be explored in general or on the level of different concrete encryption tools; or on the even finer level of concrete algorithms.
-
-The utilized abstraction level is not important for the analysis procedure, it depends on the intended audience for the analysis.
-However, it is important to use one abstraction level consistently through out the analysis.
-\end{LGContent}
-
-\section{Privacy Analysis}
-
-\TODO{TW}
-
-\begin{LGContent}
-\subsection{Step 1. World Analysis}
-
-\subsubsection{Assets: Privacy}
-
-In this document we focus our attention to only one asset: The privacy of the citizen.
-
-%TODO: UPDATE
-Our definition of privacy is described in detail in Chapter \ref{chap:privacy}.
-In Section \ref{sec:taxonomy} we define privacy as the ``control over private data'' and introduce the following seven different types of privacy:
-\begin{enumerate}
-\item Privacy of the Person
-\item Privacy of Behaviour and Action
-\item Privacy of Communication
-\item Privacy of Data and Image
-\item Privacy of Thoughts and Feelings
-\item Privacy of Location and Space
-\item Privacy os Association
-\end{enumerate}
-
-We refer to Section \ref{sec:taxonomy} for more details.
-
-%
-%
-% SECTION: IT SYSTEM DESCRIPTION MOVED ABOVE
-%
-%
-
-\subsubsection{Actors}
-\label{subsubsection:humans}
-
-This section describes the human actors previously introduced within the IT system architecture (\ref{subsubsection:it-systems}).
-Also the additional actor \textit{External} is introduced as a person with no special access rights.
-
-\textbf{Citizen.}
-Citizens are persons who use the mobile device as users of the provided software.
-Their main motivation for using the software is to gain a higher level of convenience in their daily activities.
-For example they may have access to real time bus schedules or reports about traffic jams, that help them to avoid long waiting times.
-Also they generally benefit from improvements of public infrastructure by local authorities, which is triggered by issue reports.
-
-By using the application citizens are sharing personal information like name and address, as well as data gathered from mobile sensor with the service provider.
-This data can be exploited in ways that are harmful to the citizen.
-This need to protect this protection is manifest in several laws and constitutions as the right to privacy and data protection.
-The citizen has a vital interest in having his legal rights protected and enforced.
-
-If the right to privacy is violated, there is a magnitude of potential harms that interfere with other interests of the citizen.
-This includes just annoying spam, where their technical identity is used to send unwanted commercials.
-More severe phishing attacks can exploit personal information, and try to manipulate citizens into disclosing credentials like TAN numbers and disrespect their financial interests.
-Information about the current location (GPS) or frequently used routes (stalking) can be used to attack and directly harm the health of the citizen.
-Information about medical conditions inferred by sensor data (e.g.
-fintness trackers) are of interest to insurance companies, which may affect the pricing of policies and can interfere with financial interests of the citizen.
-
-Also it is well established (cf. \cite{GuardienMassSurveillance}) that the very act of being monitored can have impact on mental health and performance, promotes distrust and breeds conformity.
-
-In short, citizens are interested in
-\begin{itemize}
-\item physical wellbeing and health
-\item financial profit
-\item convenience
-\item legitimate use of personal data
-\item non-disclosure of personal data to peers of the citizen
-\item not being monitored.
-\end{itemize}
-
-\textbf{External.}
-Externals are persons who do not have privileged access to the IT systems, and are willing to break laws, security constrains and norms in order to promote their interests.
-
-If the external is in some kind of relationship to the citizen, like a friendship or business partnership, the external can have a direct interest in gaining information about the citizen in order to increase their power.
-
-Another common interest of an external is financial profit. For example they want to obtain access to critical systems to steal sensitive data or to get the system under their control.
-Controlled systems could be leased as part of a bot net.
-Stolen data could simply be sold as is or used for illegitimate purposes, e.g. spam or phishing attacks - or excessive data mining.
-
-Because local authorities are involved in the general outline of the Live+Gov system, the possibility for politically motivated attacks is given.
-
-Externals could want to harm or destroy the systems in order to damage the reputation of local authorities (politicians or other officials) or to make a political statement of their own.
-
-Another possible motivation for external  activities could be social appreciation.
-A hacker could attack critical infrastructure just to prove his skills.
-
-In short, externals are interested in
-\begin{itemize}
-\item increase power over citizen
-\item financial profit
-\item political activism
-\item social standing.
-\end{itemize}
-
-\textbf{System Provider.}
-System Providers operate the technical infrastructure (hardware and software) of the IT System.
-They are private companies and legal persons in their own right, but also employ a number of people with diverging interests, including:
-administrators, who maintain and operate the running system;
-programmer/developer, who develop the system;
-a support manager, who handles customer relations.
-
-As companies, they are interested in gaining financial profit.
-Among other things, this depends on customer satisfaction, employee happiness and task complexity.
-Unsatisfied customers may not want to pay for the service or do not continue the business relation.
-Moreover, unsatisfied customers can create a bad reputation, which affects the market for future customers.
-
-Customer satisfaction is connected with the quality of the offered product or service.
-This quality depends on the happiness of employees.
-Employees have a claim to professional excellence.
-They want to deliver a good job within their means.
-If employees cannot satisfy their demand for professional excellence, they might get discontent and deliver poor work.
-Moreover, unhappy employees can produce higher costs through sick days.
-The worst case scenario could be, that a discontent employee gets angry and steals data or harms the running systems.
-
-At last, the financial success of system providers depends on the task complexity of the maintained infrastructure.
-The complexity of a task has to be in reasonable bounds, so that system providers can complete it within time, with a satisfying quality.
-If a task has a higher complexity than expected, financial loss is almost certain.
-Either System Providers need to hire additional competence to meet schedule and requirements.
-Or system providers they stress the time-line, which also results in a higher man-hour salary ratio and additionally endangers customer satisfaction.
-Ultimately, high task complexities can affect employee happiness, if employees cannot complete it within their claim to professional excellence.
-
-In short, System Providers are interested in
-\begin{itemize}
-\item financial profit
-\item good working conditions
-\item professional excellence
-\item manageable complexity.
-\end{itemize}
-
-\textbf{Local Authority.}
-Local authorities are public offices (ministry, agency, department, ...) or other external public entities which act as direct customers of service providers.
-They purchase a system specialized for their needs.
-For example a department for urban mobility, orders a system to better understand usage patterns and make improvement to the urban traffic flow.
-
-Such systems are investments, and so naturally local authorities are interested in a profitable return, like increased ticket sales.
-However, the return of investment is not directly of financial nature.
-Like service providers their financial gain depends on customer satisfaction.
-Customers for local authorities are either citizens, who use their services, or politicians, who order their services.
-The satisfaction of both sides is interdependent.
-
-Citizens are satisfied customers if the services, e.g. public mobility, work well.
-If citizens are happy, it is more likely that politicians gain reputation, as they organize the public services through local authorities.
-
-The first important step of improving the public services is by obtaining business intelligence.
-For the urban mobility scenario the Live+Gov systems provide insight in form of traffic jam detection and usage pattern mining, which allow local authorities to focus their efforts to the most important sites.
-
-Additionally, since local authorities act like corporations comparable to service providers, they are also interested in good working conditions.
-Discontent employees may harm the system by e.g. disclosure of privacy sensitive data.
-
-In short, local authorities are interested in
-\begin{itemize}
-\item financial profit
-\item political reputation
-\item business intelligence
-\item good working conditions.
-\end{itemize}
-
-\subsubsection{Conflicts of Interests}
-\label{subsubsection:Conflicts of Interests}
-This section outlines the Conflicts of Interests (Figure \ref{figure:Live+Gov Conflicts of Interests}) between the actors of the proposed IT system architecture.
-
-The individual interests of all actors is already described in the previous section and are not elaborated any further.
-The emphasis here is put on prominent existing conflicts, because they provide a foundation for vulnerabilities and subsequent threats.
-
-%input{./figures/LG-Conflicts-Of-Interests}
-\begin{figure}
-\centering
-\includegraphics[width=0.6\textwidth]{diagrams/png/conflict-of-interests.png}
-\caption{Live+Gov Conflicts of Interests}
-\label{figure:Live+Gov Conflicts of Interests}
-\end{figure}
-%TODO: Adapt Figure
-
-\textbf{System Complexity vs Privacy.}
-System Providers offer a service to Local Authorities, which is to provide and maintain a monitoring and mining system, e.g. for public mobility.
-This system shall produce business intelligence, so that Local Authorities can improve their public services.
-This task in it self has a high technical complexity and is the sole asset with financial return for System Providers.
-However, this task operates on privacy sensitive data provided by monitored Citizens.
-In order to ensure their privacy, System Provider would have to implement additional mechanisms, which allow Citizens to exercise control of their data.
-This will not only raise the complexity of the monitoring and mining system, System Providers also have to layout the complexity in a comprehensible manner.
-To effectively enable Citizens to preserve their privacy, they need to know what happens with their data.
-
-\textbf{Business Intelligence vs Privacy.}
-Local authorities order a monitoring and mining system from system providers, which allows them to produce business intelligence for public services.
-The system is an investment for local authorities, so they are interested in as much intelligence as possible to achieve a profitable return.
-
-The gained intelligence is the result of data mining conducted on privacy sensitive data of participating citizens.
-They are interested in the successful usage of their data, in a sense that they are also benefactors, e.g. improvement of public mobility.
-The main interest of citizens lies in maintaining control over their data and protecting their rights to privacy.
-In order do that, they need full disclosure of the processing steps and the purposes their data is used for, and to be given a choice whether such processing should be allowed for their own data.
-
-\textbf{Power of External vs. Privacy.}
-Externals which are in a social relation to the citizen can have an interest in obtaining further information in order to gain power.
-In the most simplistic example this could be a man wanting monitor the activities of his spouse.
-Another example is a Government spying on it's citizens in order to suppress opposition.
-An additional twist in the last example is, that there can be legal regulations that require the service provider to support the Governments invasion of the citizens privacy.
-
-\textbf{Financial Profit of External vs Privacy.}
-Externals can gain financial profit from stealing privacy sensitive data.
-For example by selling raw contact information to advertisers or by selling mined data to insurance companies, or intermediaries like scoring companies.
-In such cases, citizens lose complete control over their data.
-
-\textbf{Financial Profit of External vs Reputation of System Providers.}
-Externals have various business models as optional foundation for attacks on System Providers.
-For instance, they can try to invade the infrastructure for e-espionage reasons, to get control over servers to create a bot-net or to steal user data.
-All these approaches are motivated by financial interests.
-Gathered information can be sold, zombie servers can be leased.
-
-A successful attack proves the technical competence of system providers wrong and subsequently harms their professional reputation.
-This can lead to a loss of future customers or a decrease of stock price for registered companies.
-Eventually also the financial interests of system providers are endangered.
-
-\textbf{Political Activism vs Reputation of Local Authority.}
-Besides monetary reasons, externals can be motivated by political reasons to attack the monitoring and mining system.
-Externals can break the system to make a political statement of their own,
-or they can steal user data to prove the system insecure.
-Both would harm the reputation of local authorities, who endangered the privacy of the citizens.
-
-\subsubsection{Vulnerabilities}
-\label{subsubsection:Vulnerabilities}
-This section outlines the vulnerabilities (Figure \ref{figure:Live+Gov Vulnerabilities}) of the proposed monitoring and mining system.
-Note that vulnerabilities are not necessarily of technical nature.
-The weaknesses of IT systems are often created due to misuse or misconfiguration of the various components by one or more actors.
-
-%input{./figures/LG-Vulnerabilities}
-\begin{figure}
-\centering
-\includegraphics[width=0.6\textwidth]{diagrams/png/vulnerabilities.png}
-
-%\begin{flushleft}
-%\scriptsize
-%\textbf{Legend:}
-%
-%\begin{itemize}
-%\itemsep1pt\parskip0pt\parsep0pt
-%\item
-%  \textbf{(a) Citizen:} User of the L+G client application whose privacy
-%  is at stake.
-%\item
-%  \textbf{(b) Mobile Device:} Runs the L+G client application, produces
-%  and stores data sensitive to the users privacy.
-%\item
-%  \textbf{(c) L+G Data Center:} Runs the L+G services, proccesses and
-%  stores user data.
-%\item
-%  \textbf{(d) L+G Maintenance Crew:} Technical staff with access to
-%  critical infrastructure.
-%\item
-%  \textbf{(e) Aggregation Monitor:} Interface to aggregated user data.
-%\item
-%  \textbf{(f) Local Authority:} Provider of the L+G system.
-%\item
-%  \textbf{(g) Criminal:} Threatens the L+G system and Citizen privacy.
-%\end{itemize}
-%\end{flushleft}
-
-\caption{Live+Gov Vulnerabilities}
-\label{figure:Live+Gov Vulnerabilities}
-\end{figure}
-
-\textbf{Insecure Infrastructure.}
-The proposed monitoring system consists of many hardware and software components, each with its own concrete weaknesses.
-For instance, operating systems can be outdated or not subject to frequent updates or virus scans.
-Web-applications can be carelessly implemented and not protected against SQL-Injections or Cross-Site-Scripting attacks.
-Databases can be ill-configured, so that access from outside the system is possible.
-All those weak points can be subject to various known exploit techniques.
-
-\textbf{Insecure Data Transmission.}
-The proposed monitoring and mining system uses HTTP to exchange data between the Sensor Collector, the Data Center and the Report Tool.
-Per default, HTTP is a clear text protocol.
-This means, one can intercept the connection and read all sensitive information, which is send between the components.
-That is: passwords, raw sensor data and data mining results
-
-\textbf{Unhappy Employees.}
-
-An Employee that is frustrated with his situation for a long time
-period constitutes a security vulnerability. On the one hand he might
-want to harm his employer directly, on the other he is increasingly
-susceptible for social engineering.
-
-\textbf{Inadequate Access Rules.}
-The proposed IT system infrastructure has various accesses to privacy sensitive data.
-System Provider staff has access to Data Center hardware and software like databases, web-servers and other inspection tools.
-Local Authority staff has access to the Report Tool.
-This all enables staff members to have potential access to privacy sensitive information.
-Those accesses have to be secured against unauthorized third parties.
-Moreover, we need to ensure that no single person has to many access rights.
-For example, a system administrator should not be able to secretly download the whole database on a flash-drive.
-
-%TODO: Extend to general missing privacy wareness of citizen, system provider and local authority
-\textbf{Unaware Monitoring Subjects.}
-We define privacy as one's ability to control information about
-oneself.  In order to do that, monitored subjects need to know, that
-they are monitored, who monitors them, what information is recorded
-and for what purposes.  Subjects who are not aware of these things
-cannot effectively preserve control and thus lose their privacy.  This
-vulnerability expresses itself as the lack of information material
-like a Privacy Policy including concise information about applied
-processing steps, access rules and disclosure to 3rd parties.
-
-\subsection{Step 2. Potential Analysis}
-
 \subsubsection{Threat Specification}
 
-%input{./figures/Threat-Example}
-\begin{figure}
-\centering
-\includegraphics[width=0.6\textwidth]{diagrams/png/threat-example.png}
-\caption{Threat Example}
-\label{figure:Threat-Example}
-\end{figure}
-
-Recall from the security model description in Section \ref{sec:GrimmModel}, that a threat is a potential interaction of the components that targets an asset.
+A threat is a potential interaction of the components that targets an asset.
 We restrict ourselves to the case of attacks, and the asset of privacy.
@@
 
-In Figure \ref{figure:Threat-Example} we illustrate the structure of threats at the example of ``Data Theft".
-Here the attack is executed by an external person, that harms the privacy of the citizen.
-He is motivated to do so by financial profit gained by selling personal information, which is in conflict with the citizens interest in his privacy.
-To get hold of the data the external exploits an insecured HTTP transfer of recorded data.
-
-In this section we describe, in a similar fashion, threats for the citizen privacy in the Live+Gov system.
+We describe threats for the citizen privacy in the Live+Gov system.
 This list can necessarily not be complete, but we make a best effort to cover the most relevant cases.
@@
 \textbf{T1. Insufficient Control Features.}
-The System Provider does not offer tools for the Citizen to control his data.
-However, besides the missing features he provides a safe and secure system.
-This threat contradicts the definition of privacy as control over one's information about oneself.
 As soon as collected data of Citizens is stored on System Provider servers, all control over that data is lost.
-This is not necessarily due to bad intention, System Providers may simply have forgotten to include such features during the development process.
 Although, control capabilities for Citizens add to the system complexity, which could motivate to omit those.
-Therefore interests of Citizens and System Providers are in conflict, namely it is the Citizen's \textit{Privacy vs. System Complexity} for System Providers.
+Therefore interests of Citizens and System Providers are in conflict, namely it is the Citizen's \textit{Privacy vs.
+System Complexity} for System Providers.
 This is an abstract threat provoked by a general \textit{Missing Privacy Awareness} in the minds of all actors in the Live+Gov context.
@@
 \textbf{T4. Surveillance}
-
 An External infiltrates infrastructure in order to obtain information
@@
 \textbf{T6. Social Engineering}
-
 This scenario an external manipulates an employee of a Service
@@
 
-In this section we will associate to every identified threat a corresponding risk.
-Recall from \ref{sec:GrimmModel} that a risk is the expected loss that is associated to the threat.
+In this section we will associate to every identified threat a corresponding risk. A risk is the expected loss that is associated to the threat.
 Therefore, we have to quantify the likeliness of the threat to occure and the harm or loss done in this case.
@@
 The quantification of loss, will be made in a two step process.
-For each threat listed in table \ref{firgure:Threats}, we have analyzed the affected personal data of the citizen.
+For each threat listed in table \ref{figure:Threats}, we have analyzed the affected personal data of the citizen.
 For each possible data type (e.g. GPS) we analyze the impact on the seven different types of privacy in Section \ref{sec:SensorPrivacyImpact}.
@@
 
-For the quantification of the loss in case of a threat scenario we use the following rough calibration:
-\begin{itemize}
-\item 3: High. Leak of information to peers (e.g. public) which impacts citizen.
-\item 2: Medium. Undisclosed processing of personal data or disclosure to third parties that are unrelated to subject.
-\item 1: Low. Loss of control over data.
-\item 0: None
-\end{itemize}
-
-For the quantification of likeliness the following scale is used:
-\begin{itemize}
-\item 4: Always.
-\item 3: High. Occurs once in 10 cases
-\item 2: Medium. Occurs once in 100 cases
-\item 1: Low. Occurs once in 1 million cases
-\item 0: Impossible
-\end{itemize}
-
-The quantification of the risk, we add the values for loss and
-likeliness of the corresponding threats. Note, that loss and
-likeliness scales have a logarithmic character, so that that addition
+For the quantification of the loss in case of a threat scenario we use the following rough calibration between 3 (high) and 0 (none).
+For the quantification of likeliness the following scale between 4 (always) and 0 (impossible) is used.
+The quantification of the risk, we add the values for loss and likeliness of the corresponding threats.
+Note, that loss and
+likeliness scales have a logarithmic character, so that addition
 of those scales corresponds to multiplication of the usual scales.
@@
 
-\subsubsection{Privacy Recommendations}
-\label{sec:prec}
+\subsubsection{Security Requirements}
+
+A set of interactions (e.g. threats) that shall not occur within the system in order to achieve its business objectives. Security requirements are targeted to protect one or more assets.
+
+An example of a security requirement is that a given communication channel shall not be infringed by externals.
+We included these security requirements into our analysis in Section \ref{subsec:priv-rec}.
+
+\subsection{Privacy Recommendations}
+\label{subsec:priv-rec}
 
@@
 users privacy.  In this section we derive recommendations or
-requirements for a system that addresses these risks.  Some of these
-requirements are implemented as security measures in our systems and
-discussed in the following chapter \ref{chap:impl}.
+requirements for a system that addresses these risks..
 
@@
 is easily readable and contains all important information is
-essential.  Moreover, the existence of a Privacy Policy is a legal
-requirement (cf. Section \ref{EUDIR}).
+essential and a legal requirement.
 
@@
 permissions from the citizens inside the privacy policy is neither an
-ethical or legal solution to the problem (cf. Art. 6 in Section
-\ref{EUDIR}).
-
-To address this threat awareness about the limitations of data
+ethical or legal solution to the problem. To address this threat awareness about the limitations of data
 processors inside the company is a key element. As one mean to
@@
 internal use that explains the concrete processing steps and purposes
-that are permitted by the citizens. This guideline should also be
-structured in a way that it covers the legal notification requirement
-from the EU Data Protection Directive (cf. Section \ref{EUDIR}). In
-particular the following information should be provided for each
-processing task: The name of the controller, purpose of processing,
+that are permitted by the citizens. In particular the following information should be provided for each processing task: The name of the controller, purpose of processing,
 description of the data categories, recipients of the data if
@@
 \begin{enumerate}
-\item[R1] Privacy Dashboard. A tool which allows the citizen to view, edit
+\item[R1] \textbf{Privacy Dashboard}. A tool which allows the citizen to view, edit
   and delete all data personal data that is stored in the system.
-\item[R2] Privacy Policy. A document, that informs the citizen about the
+\item[R2] \textbf{Privacy Policy}. A document, that informs the citizen about the
   collection and processing of personal information. It should at
   least contain the legally required information.
-\item[R3] Issue Data Handling Guidelines that explain the permitted
+\item[R3] Issue \textbf{Data Handling Guidelines} that explain the permitted
   processing methods and purposes.
-\item[R4] Ask the citizens for permissions before applying further
+\item[R4] Ask the citizens for \textbf{permissions} before applying further
   processing via Email.
-\item[R5] Anonymize personal data before processing.
-\item[R6] Securing of Storage and Processing infrastructure using e.g. firewalls.
-\item[R7] Securing communication channels using encryption.
-\item[R8] Maintain a healthy, trustful relationship with your employees.
+\item[R5] \textbf{Anonymize} personal data before processing.
+\item[R6] \textbf{Securing of Storage} and Processing infrastructure using e.g. firewalls.
+\item[R7] \textbf{Securing communication} channels using encryption.
+\item[R8] Maintain a healthy, trustful \textbf{relationship} with your employees.
 \end{enumerate}
 The mapping of these recommendations to the threats is summarized in Figure \ref{fig:risks}.
-\end{LGContent}
+%\end{LGContent}
 

### 2015-01-15 paper.tex
Further reductions in Section 4

@@
 
-	\item
-	\textbf{Assets.} Things of value to one or more stakeholders.
+	\item \textbf{Assets.} 
+	Things of value to one or more stakeholders.
 
@@
 Such systems are investments, and so naturally local authorities are interested in a profitable return, like increased ticket sales.
-
-The first important step of improving the public services is by obtaining business intelligence.
-For the urban mobility scenario the Live+Gov systems provide insight in form of traffic jam detection and usage pattern mining, which allow local authorities to focus their efforts to the most important sites.
 
@@
 	\item \textbf{Security Requirements.}
-	Interactions (e.g. threats) that shall not occur within the system in order to achieve its business objectives.
+	Interactions (e.g.
+threats) that shall not occur within the system in order to achieve its business objectives.
+We included these security requirements into our analysis in Section \ref{subsec:priv-rec}.
 
@@
 Also the External might be motivated politically and wants to harm the reputation of the System Provider or the Local Authority.
-Anyways, this threat would be manifested through a technical attack on either hardware (Packet Capture) or software (buffer overflow, SQL injection).
 Such a successful attack could harm the reputation of both System Provider and Local Authority.
-The technical competence of System Providers would be proven wrong, therefore they would lose professional reputation.
-Local Authorities would lose their political reputation, as it would seem like they had endangered data of Citizens, which lose complete control.
 Thus this threat is defined by three conflicts: \textit{Privacy vs.
@@
 planned activities of the citizens (e.g. Nixon's Watergate scandal or
-the recent prosecution of Guardian journalists by GHCQ).  Another
-example is an insurance company that seeks to get information about
-the citizens life-style in relation to the insured risk, like car
-accidents or health hazards.
-
+the recent prosecution of Guardian journalists by GHCQ). 
 In this threat the privacy interest of the citizen is in conflict with
@@
 provider, without somebody else noticing, and the service provider himself
-has an interest in this activity.
-
-The associated loss, on the other hand can be substantial (Medium
-$2$).  For example when GPS data is linked to data from telephone
-books the identity of the citizen can be revealed and personal details
-like visits to doctors. In the threat scenario, this information is
-not leaked to third parties, (which would justify an even higher loss
-assessment), but the very existence of this information violates the
-citizens privacy.
+has an interest in this activity. The associated loss, on the other hand can be substantial (Medium $2$).
 
@@
 \end{figure}
-
-\subsubsection{Security Requirements}
-
-A set of interactions (e.g. threats) that shall not occur within the system in order to achieve its business objectives. Security requirements are targeted to protect one or more assets.
-
-An example of a security requirement is that a given communication channel shall not be infringed by externals.
-We included these security requirements into our analysis in Section \ref{subsec:priv-rec}.
 

### 2015-01-19 paper.tex
minor changes

@@
 e.g. \cite{Avizienis}.  The reference model consists of a \emph{model}
-and a \emph{procedure}: The model aims to organize common security terminology in a reasonable and practical way, the procedure describes a method to analysis the IT system based on that model.
-In section \ref{subsec:priv-rec} we provide specific privacy recommendations as a result of the previous analysis that the operated system must comply.
+and a \emph{procedure}: The model aims to organize common security terminology in a reasonable and practical way.
+The procedure describes a method to analysis the IT system based on that model.
+In these sections we give a brief overview over the reference model by applying it to our scenario.
+
+In section \ref{subsec:priv-rec} we provide specific privacy recommendations and requirements as a result of the previous analysis that the system must comply with.
 
@@
 
-The \emph{world view} contains all components describing the current state.
-It consists of the following components:
+The first step is the \emph{world view} where all components are described in their current state. It consists of the following components:
 \begin{itemize}
@@
 This list can necessarily not be complete, but we make a best effort to cover the most relevant cases.
-
-% \label{figure:Threats}
-%input{./figures/Threats}
-\begin{figure}
-\centering
-%\includegraphics[width=\textwidth]{diagrams/png/threat-table.png}
-\newcommand{\tcell}[1]{\parbox[c]{6cm}{#1\vspace{3mm}}}
-\newcommand{\thead}[1]{\tcell{\vspace{3mm}\centering\textbf{#1}}}
-\newcommand{\tbody}[1]{\tcell{\footnotesize\flushleft #1}}
-\resizebox{\textwidth}{!}{
-\begin{tabular}{|c|c|c|c|}
-
-\hline
-\thead{Threat}
-& \thead{Conflict of Interest}
-& \thead{Vulnerability}
-& \thead{Affected Data}
-\\\hline\hline
-\tbody{Insufficient Control Features}
-& \tbody{Privacy vs. System Complexity}
-& \tbody{Missing Privacy Awareness}
-& \tbody{All collected data}
-\\\hline
-
-\tbody{Exessive Data Mining}
-& \tbody{Privacy vs. Business Intelligence (LA),
-\\Privacy vs. Financial Profit (SP)}
-& \tbody{Lax data handling policies
-\\Missing Privacy Awareness
-\\Weak law enforcement}
-& \tbody{Mined Information}
-\\\hline
-
-\tbody{Data Theft}
-& \tbody{Privacy vs. Financial Profit
-\\Reputation (SP) vs. Political Activism
-\\Reputation (LA) vs. Political Activism}
-& \tbody{Insecure Infrastructure,
-\\Insecure Communication}
-& \tbody{Transmitted data 
-\\Data stored in data center}
-\\\hline
-
-\tbody{Surveillance}
-& \tbody{Privacy vs. External Power}
-& \tbody{Insecure Infrastructure
-\\Insecure Communication}
-& \tbody{Transmitted data 
-\\Data stored in data center}
-\\\hline
-
-\tbody{Information Leak}
-& \tbody{Privacy vs. Financial Profit
-\\Reputation (SP) vs. Political Activism
-\\Reputation (LA) vs. Political Activism}
-& \tbody{Unhappy Employee 
-\\Lax Access Rules}
-& \tbody{Data stored in data center}
-\\\hline
-
-\tbody{Social Engineering}
-& \tbody{Privacy vs. Financial Profit
-\\Reputation (SP) vs. Political Activism
-\\Reputation (LA) vs. Political Activism}
-& \tbody{Unhappy Employee
-\\Lax Access Rules}
-& \tbody{Data stored in data center}
-\\\hline
-\end{tabular}
-}
-\caption{Threats}
-\label{figure:Threats}
-\end{figure}
 
@@
 The quantification of loss, will be made in a two step process.
-For each threat listed in table \ref{figure:Threats}, we have analyzed the affected personal data of the citizen.
+For each threat listed in the previous section, we have analyzed the affected personal data of the citizen.
 For each possible data type (e.g. GPS) we analyze the impact on the seven different types of privacy in Section \ref{sec:SensorPrivacyImpact}.

### 2015-01-19 paper.tex
removed legend from fig1, scaled fig1 and fig4 down to 60% textwidth

@@
 \centering
-\includegraphics[width=0.7\textwidth]{diagrams/png/it-systems.png}
-
-\begin{flushleft}
-\scriptsize
-\textbf{Legend:}
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-
-\item
-\textbf{(a) Citizen:} 
-User of the L+G client application whose privacy is at stake.
-
-\item
-\textbf{(b) Service Provider:} Provides technical infrastructure.
-
-\item
-\textbf{(c) Local Authority:} Provider of the L+G system.
-
-\item
-\textbf{(d) Mobile Device:}
-Runs the L+G client application, produces and stores data sensitive to the users privacy.
-
-\item
-\textbf{(e) Data Center:} Runs the L+G services, processes and stores user data.
-
-\item
-\textbf{(f) Report Tool:} Interface to aggregated user data.
-
-\end{itemize}
-\end{flushleft}
+\includegraphics[width=0.6\textwidth]{diagrams/png/it-systems.png}
+
+%\begin{flushleft}
+%\scriptsize
+%\textbf{Legend:}
+%\begin{itemize}
+%\itemsep1pt\parskip0pt\parsep0pt
+%
+%\item
+%\textbf{(a) Citizen:} 
+%User of the L+G client application whose privacy is at stake.
+%
+%\item
+%\textbf{(b) Service Provider:} Provides technical infrastructure.
+%
+%\item
+%\textbf{(c) Local Authority:} Provider of the L+G system.
+%
+%\item
+%\textbf{(d) Mobile Device:}
+%Runs the L+G client application, produces and stores data sensitive to the users privacy.
+%
+%\item
+%\textbf{(e) Data Center:} Runs the L+G services, processes and stores user data.
+%
+%\item
+%\textbf{(f) Report Tool:} Interface to aggregated user data.
+%
+%\end{itemize}
+%\end{flushleft}
 
@@
 \centering
-\includegraphics[width=0.8\textwidth]{diagrams/png/vulnerabilities.png}
+\includegraphics[width=0.6\textwidth]{diagrams/png/vulnerabilities.png}
 

### 2015-01-21 paper.tex
added 2.1 changed introduction of 4

--- 2015-01-19 paper.tex
@@
 
-\subsubsection{IT-Systems \& Interactions}
+\subsection{A Privacy Respecting Mobile Traffic Survey}
+
+The scenario of our privacy analysis is a mobile traffic survey that supports local travel services by collecting and evaluating mobile sensor data of their customers.
+The survey is executed by a trustworthy third-party service.
+The travel services need the personal mobile data for two reasons: firstly, they want to serve their customers individually, for example by guiding them to proximity stations, to recommend adequate travel connections, to inform about delays, or to sell electronic tickets.
+For this purpose, the personal data for evaluation are related to the individual customers.
+Secondly, the travel services want to enhance their system, in that they evaluate the demand of stations, connections, changes, ways and time spent by their customers.
+For this purpose, personal data are anonymized before evaluation.
+
+\subsection{IT-Systems \& Interactions}
 \label{subsubsection:it-systems}
@@
 
-The goal of this chapter is to analyze and identify the threads to
-personal privacy that are posed by collecting, storing and processing
-sensor data from mobile phones.  We derive concrete privacy protection
-measures that address the main risks involved with handling such data.
-
-In section \ref{subsec:world-analysis} and \ref{subsec:pot-analysis} we apply the first two steps of the Reference Model for IT Security Analysis as described in \cite{Grimm:ItSecRefModel}.
-It supersedes earlier efforts by
-e.g. \cite{Avizienis}.  The reference model consists of a \emph{model}
-and a \emph{procedure}: The model aims to organize common security terminology in a reasonable and practical way.
-The procedure describes a method to analysis the IT system based on that model.
-In these sections we give a brief overview over the reference model by applying it to our scenario.
-
-In section \ref{subsec:priv-rec} we provide specific privacy recommendations and requirements as a result of the previous analysis that the system must comply with.
+The goal of this chapter is to analyze and identify the threats to personal privacy that are posed by collecting, storing and processing sensor data from mobile phones.
+We derive concrete privacy protection measures that address the main risks involved with handling such data.
+
+In our analysis we follow the “reference model for IT security analysis” as described in \cite{Grimm:ItSecRefModel}.
+It supersedes earlier efforts by e.g.
+\cite{Avizienis}.
+The reference model consists of a model and a procedure.
+The model organizes a common security terminology in a reasonable and practical way.
+The procedure describes a method for analyzing the IT system based on that model.
+The reference model provides four views: (1) the real world of persons and their assets, (2) the potential world of requirements and threats, (3) measures and plans specified by programs, business models and attack strategies, and, finally, (4) events of running programs, data accesses and performed attacks as well as their defense.
+In the following sections we apply the proposed procedures of the reference model to our scenario of a travel service that collects personal data from mobile users in order to serve the users and to enhance the service.
+The service intends to respect the privacy of its users.
+
+In sections \ref{subsec:world-analysis} and \ref{subsec:pot-analysis} we apply the first two steps of the reference model \cite{Grimm:ItSecRefModel}, which are related to the views on the real world and on the measures and plans.
+In section \ref{subsec:priv-rec} we apply the third step of the reference model by providing specific privacy recommendations and requirements as a result of the previous analysis that the system must comply with.
 
@@
 The \emph{Potential Analysis} (or \emph{Potential View}) displays the intended and unintended interactions of the components in the world view.
+
 The intended interactions support the underlying business objectives.
-Unintended interactions lead to threats. The potential view consists of the following components:
+Unintended interactions can lead to threats.
+Obviously this provides a conflict of interest with the victim's business model, to keep the asset away from unauthorized access.
+From the point of view of the victim, an attack would be an unintended interaction.
+
+The potential view consists of the following components:
 

### 2015-01-22 paper.tex
section 2 and bib-file updated

@@
 
-%\TODO{MM:\\
-%- Shorten\\
-%- DONE - Remove Implicit Type Violations\\
-%- Update figure\\
-%- RG: Discuss alternative definitions of privacy\\
-%}
-
-\noindent
-Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes:
+More than one hundred years ago, 1890, the American politician Louis Brandis had specified privacy as the right to be left alone \cite{WarrenBrandeis:RightToPrivacy}.
+From the data processing point of view, this right is best expressed by the absence of information about a person in the mind of others.
+Indeed, this principle of “data minimization” is still fundamental to modern data protection legislation.
+The modern principles of data protection are codified by many legal systems in different countries, especially in Europe \cite{EU-directive2}, and by the US save harbor principles.
+However, modern data protection is more than only the absence of personal data.
+It is based on the personal right on self-determination over information and communication \cite{BVerfG:census}.
+This is, for example, very well expressed by Fried’s definition of 1970:
 \begin{quote}
-The notion of privacy remains out of the grasp of every academic chasing it.
-Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.
-\cite{Gutwirth}
+	Privacy is not simply an absence of information about us in the minds of others; rather it is the control we have over information about ourselves. \cite{CFried:Privacy}
 \end{quote}
-In this section we introduce the concept of privacy as control over information.
-We specify types of information we have to deal with in the context of privacy.
-And finally we outline the impact mobile sensors can have on one's privacy.
-
-\subsection{Privacy as Control over Information}
-%We may need to remove the following paragraph depinding on RG's alternative definition of privacy
-%Many definitions seem to focus on the vulnerabilities of privacy.
-%Thus a great amount of work is invested in finding threats to prohibit instead of describing why privacy is so valuable to us.
-%This way around one could derive measures to ensure and secure its value. \cite{7ToP}
-
-%One particular scholar, who did not follow this way of thinking is Charles Fried.
-One possible definition of privacy is given by Charles Fried in his work \textit{An Anatomy of Values}\cite{CFried:Privacy}.
-He questioned, why we are intuitively sensitive to privacy violations.
-But he did not assert privacy as an intrinsic value by itself, he rather stated:
-\begin{quote}
-\textbf{Privacy is} not simply an absence of information about us in the minds of others;
-rather it is \textbf{the control we have over information about ourselves}. \cite{CFried:Privacy}
-\end{quote}
-According to Fried privacy is our ability to create and modulate the relationships to other humans.
-Which makes us intuitively sensitive for its violation, because this affects what defines our identity.
-Also, the ability to build and maintain relationships is essential to society.
-This is what makes privacy valuable and deserving of protection. 
-\cite{sep-privacy}\cite{CFried:Privacy}
-
-\subsubsection{Relationship Creation.}
-
-% Privacy + Trust ~> Relationship
-% ===============================
-% - A shares intimate information with B
-% - A trusts B not to reveal that information by not constantly monitoring B, i.e. respecting B's privacy
-% - B has now the possibility to share intimate information with A
-
-We share information of great intimacy only with persons we deem trustworthy. 
-Moreover, we trust those persons to not reveal information about us to others by not constantly monitoring them, i.e. respecting their privacy.
-Trust needs the possibility of unknown failure.
-If we would constantly monitor our partners, they cannot fail unnoticed nor can they willingly share intimate information with us.
-By trusting them, we create the possibility for them to share intimate information with us.
-\cite{CFried:Privacy}
-
-\subsubsection{Relationship Modulation.}
-Fried argues, depending on the person we are interacting with, we are changing the degree of intimacy we share.
-If we are to share private information with others, we are most likely aware that this action is privacy related.
-In that case we are able to selectively disclose information along the two dimensions of quantity and quality.
-\cite{CFried:Privacy}
-
-\subsubsection{Alternative Definitions.}
-Using Fried's anatomy of privacy for this analysis is suitable because of two points.
-Despite the fact his text was published in 1970, he already included technologies to its viewpoint that did not only monitor location, but also record biometric data \cite{CFried:Privacy}. 
-Secondly, the concept of control over information does not enforce data minimization.
-Fried's definition complies with what we call \textit{self data protection}, where persons communicate their data self-determined.
-This is the opposite of \textit{system data protection}, where the collection of certain data is prohibited.
-
-\TODO{RG: Discuss alternative definitions of privacy}
-
-%=======================================================================================
-
-\begin{LGContent-Hidden}
-Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes:
-
-\begin{quote}
-The notion of privacy remains out of the grasp of every academic chasing it.
-Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.
-\cite{Gutwirth}
-\end{quote}
-
-%Many researchers seem to only ``focus on the ways in which privacy can be infringed'' \cite{7ToP}.
-%Thus they try to create taxonomies of \emph{privacy harms} instead of taxonomies of \emph{privacy types}.
-%Those two differ in the respect that the former focuses on threats to prohibit whereas the latter focuses on values to protect.
-%So one should rather evaluate what aspects are precious about privacy and develop measures to ensure their security than only forbid single actions against it \cite{7ToP}.
-
-Many researchers seem to only ``focus on the ways in which privacy can be infringed'' \cite{7ToP}.
-Thus they invest a great amount of work in defining threats to prohibit instead of describing why privacy is so valuable to us.
-This way around one could derive measures to ensure and secure its value \cite{7ToP}.
-One particular scholar who does this in the context of digital monitoring is Charles Fried.
-He questioned, why we are intuitively sensitive to violations of privacy.
-But he did not assert privacy as an intrinsic value by itself, he rather sated:
-
-\begin{quote}
-Privacy is not simply an absence of information about us in the minds of others;
-rather it is \textbf{the control we have over information about ourselves}. \cite{CFried:Privacy}
-\end{quote}
-
-According to Fried privacy is a \textit{rational context}.
-Which means that one is aware of such context's existence during a rational action.
-If we are to share private information with others, we are most likely aware that this action is privacy related.
-In that case we are able to selectively disclose information along the two dimensions of quantity and quality.
-This leads Fried to his thesis, that privacy is one's ability to create and modulate his or hers social relationships, namely: friendship, love and trust. \cite{CFried:Privacy}
-
-Depending on the conversation partner, we change the degree of intimate information we share if it is a total stranger, colleague, close friend or lover.
-With close friends or lovers we share information of great intimacy we do not share with anyone else.
-Moreover, we trust those persons to not reveal information about us to others by respecting their privacy.
-Trust needs the possibility of unknown failure.
-If we would constantly monitor our partners, they cannot fail unnoticed nor can they willingly share that information with us.
-Thus they could not trust us anymore. \cite{CFried:Privacy}
-
-So privacy or the its possibility, according to Fried, is the foundation of our core relations: friendship, love and trust.
-And thus it is valuable, because those relations are essential to human society \cite{CFried:Privacy}.
-Using his anatomy of privacy as foundation of our analysis is suitable because of two points.
-At first, Fried's study on the understanding of privacy provided a great contribution to the research on the same term in philosophy \cite{sep-privacy} and computer science \cite{SociotechnicalArchitectureForOnlinePrivacy}.
-Secondly, despite the fact his text was published in 1970, he already included technologies to its viewpoint that did not only monitor location, but also record biometric data.
-\end{LGContent-Hidden}
+In order to realize this requirement in the modern IT world users are provided with user control functions, for example to decide about the access on their data for future use, and to view, modify, and delete unwanted data that had been collected.
+In its fundamental decision of May 13, 2014, the European Court has convicted Google to accept user demands to delete links to incorrect or irrelevant personal data in the Internet (C-131/12 – Google Spain SL).
+Such functions would constitute a so-called self-data-protection.
+It requires users to be highly aware of their rights and how to use them.
+Complementary, system-data-protection puts the load of data protection enforcement on the data collectors, ideally even without bothering the users.
+Typical system data protection functions are the deletion of personal data that are not bound to service purpose, the anonymization of personal data for research, and the abstinence from forwarding personal data to third parties.
+Together, system-data-protection and self-data-protection are supposed to provide a fair balance between the interests of service providers and their users.
+In our research work we focus on self-data-protection according to Fried’s user control requirement \cite{CFried:Privacy}.
+A system data protection point of view would lead to other results, namely to a set of obligations for service providers.
+This is subject to further research.
 
@@
 
-In Fried's definition of privacy as control over information, the specification of what constitues such information remains open.
+In Fried's definition of privacy as control over information, the specification of what constitutes such information remains open.
 There is a vast amount of information that relates to a person and we need to get a better understanding in order to perform a thorough analysis.
-To this end we use the the categorization by Friedewald, Finn and Wright \cite{7ToP} called the \textit{Seven Types of Privacy}.
-The seven types of privacy are an extension to the four types of privacy by Roger Clarke \cite{RClarke:4ToP}, which are:
-\textit{Privacy of the Person},
-\textit{Privacy of Personal Behaviour},
-\textit{Privacy of Personal Communication},
-\textit{Privacy of Personal Data}.
-It is important to note, that these categories do not form a taxonomy, since the categories are not mutually exclusive.
+To this end we use the categorization by Friedewald, Finn and Wright of 2013 called the Seven Types of Privacy \cite{7ToP}.
+These are an extension to the four types of privacy by Roger Clarke of 1997 \cite{RClarke:4ToP}.
+It is important to note, that these categories are not mutually exclusive.
 For instance a written email is considered personal communication as well as personal data stored on a computer.
-
-Moreover, Friedewald et al. argue that Clarke's types are outdated and no longer adequate in order to describe the privacy aspect of our modern, technology driven, world.
-In order to address this shortcoming they extend the former four to the now introduced seven types privacy as follows:
-
+However, the categories are very helpful for a privacy analysis with a focus on self-data-protection.
+The seven types of privacy are as follows:
 \begin{enumerate}
-
-\item \textbf{Privacy of the Person}
-This privacy type is generally concerned with one could best understand as biometric privacy.
-Friedewald et al. paraphrase it as \emph{``\om the right to keep body functions and body characteristics \om private''}.
-This includes but is not limited to measures like weight, height or shoulder width;
-biometric identifiers like fingerprints and DNA sequences;
-or medical conditions such as limping or having a cold.
-
-\item \textbf{Privacy of Behaviour and Action}
-This privacy type is concerned with one's activities in public as well as in private spaces.
-It includes but is not limited to religious practices, political activities and sexual preferences or habits.
-
-\item \textbf{Privacy of Communication}
-This privacy type is concerned with one's communication in a broad sense.
-It includes written correspondence, but also conversations conducted either vis-a-vis or via electronic devices.
-Friedewald et al. put it as the right to free discussion without unknown interception by third parties.
-
-\item \textbf{Privacy of Data and Image}
-This privacy type is concerned with the secrecy of personal data, especially its automatic disclosure to other individuals and organizations.
-It includes data such as paychecks, insurance information or records of public administration.
-However, it also refers to pictures taken without consent and digital identifiers like IP addresses or social security numbers.
-
-\item \textbf{Privacy of Thoughts and Feelings}
-This privacy type is the counterpart to Privacy of the Person like body and mind are counterparts of one another.
-Comparable to the Privacy of Data and Image, Friedewald et al. state that one's thoughts and feelings must not be automatically revealed to others.
-This could simply happen by the disclosure of one's diary or by technologies which allow emotion detection through biometric means.
-One's body temperature or iris reflexes might infer stress or excitation.
-
-\item \textbf{Privacy of Location and Space}
-This privacy type is concerned with one's movements in public spaces and the protection of private spaces.
-Friedewald et al. qualify the first dimension as one's right to move without being identified, tracked or monitored.
-The second dimension is qualified as one's general right to solitude, especially the right to the inviolability of the home.
-
-\item \textbf{Privacy of Association}
-This privacy type is also put as group privacy.
-Friedewald et al. state that one must have the possibility with whomever without being recorded.
-Associations like friends or organizations such as political parties must not automatically be recorded because one associates with them, and vice-versa.
+	\item Privacy of the Person (with respect to body functions and body characteristics)
+	\item Privacy of Behavior and Action
+	\item Privacy of Communication
+	\item Privacy of Data and Image
+	\item Privacy of Thoughts and Feelings
+	\item Privacy of Location and Space 
+	\item Privacy of Association (with persons, e.g. friends, and organizations, e.g. political parties)
 \end{enumerate}
-
-\begin{LGContent-Hidden}
-In Fried's definition of privacy as control over "information about onself'', the specification of what constitues such information remains open.
-There is a vast amount of information that relates to a person and we need to get a better understanding in order to perform a thorough analysis.
-To this end we use the the categorization by Friedewald, Finn and Wright \cite{7ToP} called the \textit{Seven Types of Privacy}.
-The seven types of privacy are an extension to the four types of privacy by Roger Clarke \cite{RClarke:4ToP}, which are:
-\textit{Privacy of the Person},
-\textit{Privacy of Personal Behaviour},
-\textit{Privacy of Personal Communication},
-\textit{Privacy of Personal Data}.
-It is important to note, that this categories do not form no taxonomy, since the categories are not mutually exclusive.
-For instance a written email is considered personal communication as well as personal data stored on a computer.
-
-Moreover, Friedewald et al. argue that Clarke's taxonomy is outdated and no longer adequate in order to describe the privacy aspect of our modern, technology driven, world.
-In order to address this shortcoming they extend the former four to the now introduced seven types privacy as follows:
-
-\begin{enumerate}
-
-\item \textbf{Privacy of the Person}
-This privacy type is generally concerned with one could best understand as biometric privacy.
-Friedewald et al. paraphrase it as \emph{``\om the right to keep body functions and body characteristics \om private''}.
-This includes but is not limited to measures like weight, height or shoulder width;
-biometric identifiers like fingerprints and DNA sequences;
-or medical conditions such as limping or having a cold.
-
-\item \textbf{Privacy of Behaviour and Action}
-This privacy type is concerned with one's activities in public as well as in private spaces.
-It includes but is not limited to religious practices, political activities and sexual preferences or habits.
-
-\item \textbf{Privacy of Communication}
-This privacy type is concerned with one's communication in a broad sense.
-It includes written correspondence, but also conversations conducted either vis-a-vis or via electronic devices.
-Friedewald et al. put it as the right to free discussion without unknown interception by third parties.
-
-\item \textbf{Privacy of Data and Image}
-This privacy type is concerned with the secrecy of personal data, especially its automatic disclosure to other individuals and organizations.
-It includes data such as paychecks, insurance information or records of public administration.
-However, it also refers to pictures taken without consent and digital identifiers like IP addresses or social security numbers.
-
-\item \textbf{Privacy of Thoughts and Feelings}
-This privacy type is the counterpart to Privacy of the Person like body and mind are counterparts of one another.
-Comparable to the Privacy of Data and Image, Friedewald et al. state that one's thoughts and feelings must not be automatically revealed to others.
-This could simply happen by the disclosure of one's diary or by technologies which allow emotion detection through biometric means.
-One's body temperature or iris reflexes might infer stress or excitation.
-
-\item \textbf{Privacy of Location and Space}
-This privacy type is concerned with one's movements in public spaces and the protection of private spaces.
-Friedewald et al. qualify the first dimension as one's right to move without being identified, tracked or monitored.
-The second dimension is qualified as one's general right to solitude, especially the right to the inviolability of the home.
-
-\item \textbf{Privacy of Association}
-This privacy type is also put as group privacy.
-Friedewald et al. state that one must have the possibility with whomever without being recorded.
-Associations like friends or organizations such as political parties must not automatically be recorded because one associates with them, and vice-versa.
-\end{enumerate}
-
-\end{LGContent-Hidden}
 
@@
 
-The first step is the \emph{world view} where all components are described in their current state. It consists of the following components:
-\begin{itemize}
-
-	\item \textbf{Assets.} 
-	Things of value to one or more stakeholders.
-
-	\item \textbf{IT-Systems.}
-	The relevant IT-Systems under study. Already described in Section \ref{subsubsection:it-systems}.
-
-	\item \textbf{Actors.}
-	All identifiable stakeholders of the system under study.
-
-	\item \textbf{Conflicts of Interests.}
-	Different actors have different interests which can be in conflict which each other.
-
-	\item \textbf{Vulnerabilities.}
-	All identifiable weaknesses in the IT-System e.g. a weak passwords.
-
-	\item \textbf{Interactions.}
-	This point captures all possible interactions between assets, IT-Systems, humans, and vulnerabilites as partly described in Section \ref{subsubsection:it-systems} and will be further analyzed in \ref{subsec:pot-analysis}.
-
-\end{itemize}
+The first step is the \emph{world view} where all components are described in their current state.
+It consists of the following components: \texttt{Assets}, \texttt{IT-Systems}, \texttt{Actors}, \texttt{Conflicts of Interests}, \texttt{Vulnerabilities}, and \texttt{Interactions}.
+
+The relevant \texttt{IT-Systems} were already described in Section \ref{subsubsection:it-systems}.
+\texttt{Interactions} between assets, IT-Systems, humans, and vulnerabilites as partly described in Section \ref{subsubsection:it-systems} and will be further analyzed in Section \ref{subsec:pot-analysis}.
 
@@
 
-The \emph{Potential Analysis} (or \emph{Potential View}) displays the intended and unintended interactions of the components in the world view.
+The \emph{Potential Analysis} displays the intended and unintended interactions of the components in the world view.
 

### 2015-01-22 paper.tex

@@
 \centering
-\includegraphics[width=0.7\textwidth]{diagrams/png/it-systems.png}
+\includegraphics[width=0.6\textwidth]{diagrams/png/it-systems.png}
 
-\begin{flushleft}
-\scriptsize
-\textbf{Legend:}
-\begin{itemize}
-\itemsep1pt\parskip0pt\parsep0pt
-
-\item
-\textbf{(a) Citizen:} 
-User of the L+G client application whose privacy is at stake.
-
-\item
-\textbf{(b) Service Provider:} Provides technical infrastructure.
-
-\item
-\textbf{(c) Local Authority:} Provider of the L+G system.
-
-\item
-\textbf{(d) Mobile Device:}
-Runs the L+G client application, produces and stores data sensitive to the users privacy.
-
-\item
-\textbf{(e) Data Center:} Runs the L+G services, processes and stores user data.
-
-\item
-\textbf{(f) Report Tool:} Interface to aggregated user data.
-
-\end{itemize}
-\end{flushleft}
+%\begin{flushleft}
+%\scriptsize
+%\textbf{Legend:}
+%\begin{itemize}
+%\itemsep1pt\parskip0pt\parsep0pt
+%
+%\item
+%\textbf{(a) Citizen:} 
+%User of the L+G client application whose privacy is at stake.
+%
+%\item
+%\textbf{(b) Service Provider:} Provides technical infrastructure.
+%
+%\item
+%\textbf{(c) Local Authority:} Provider of the L+G system.
+%
+%\item
+%\textbf{(d) Mobile Device:}
+%Runs the L+G client application, produces and stores data sensitive to the users privacy.
+%
+%\item
+%\textbf{(e) Data Center:} Runs the L+G services, processes and stores user data.
+%
+%\item
+%\textbf{(f) Report Tool:} Interface to aggregated user data.
+%
+%\end{itemize}
+%\end{flushleft}
 

### 2015-01-22 paper.tex
futher changings in section 2

@@
 
+\subsection{Privacy and Self Data Protection}
 More than one hundred years ago, 1890, the American politician Louis Brandis had specified privacy as the right to be left alone \cite{WarrenBrandeis:RightToPrivacy}.

### 2015-01-22 paper.tex
further reductions in section 4

@@
 The first step is the \emph{world view} where all components are described in their current state.
-It consists of the following components: \texttt{Assets}, \texttt{IT-Systems}, \texttt{Actors}, \texttt{Conflicts of Interests}, \texttt{Vulnerabilities}, and \texttt{Interactions}.
+It consists of the following components: Assets, IT-Systems, Actors, Conflicts of Interests, Vulnerabilities, and Interactions.
 
-The relevant \texttt{IT-Systems} were already described in Section \ref{subsubsection:it-systems}.
-\texttt{Interactions} between assets, IT-Systems, humans, and vulnerabilites as partly described in Section \ref{subsubsection:it-systems} and will be further analyzed in Section \ref{subsec:pot-analysis}.
+The relevant \textbf{IT-Systems} were already described in Section \ref{subsubsection:it-systems}.
+\textbf{Interactions} between assets, IT-Systems, humans, and vulnerabilites as partly described in Section \ref{subsubsection:it-systems} and will be further analyzed in Section \ref{subsec:pot-analysis}.
 
@@
 
-In this scenario we focus our attention to only one asset: The privacy of the citizen.
-Our definition of privacy is described in detail in Section \ref{sec:privacydef} and introduced in Section \ref{sec:privacytypes} the seven different types of privacy.
+In this scenario we focus our attention to only one asset: The privacy of the citizen. Our definition of privacy is described in detail in Section \ref{sec:privacydef},
 
@@
 This section describes the human actors previously introduced within the IT system architecture (\ref{subsubsection:it-systems}).
-Also the additional actor \textit{External} is introduced as a person with no special access rights.
 Although the text is only a short description, a list of the most important interests is provided for each actor.
@@
 Local authorities are public offices (ministry, agency, department, ...) or other external public entities which act as direct customers of service providers.
-They purchase a system specialized for their needs.
 For example a department for urban mobility, orders a system to better understand usage patterns and make improvement to the urban traffic flow.
@@
 This section outlines the vulnerabilities (Figure \ref{figure:Live+Gov Vulnerabilities}) of the proposed monitoring and mining system.
-Note that vulnerabilities are not necessarily of technical nature.
-The weaknesses of IT systems are often created due to misuse or misconfiguration of the various components by one or more actors.
 
@@
 
-The potential view consists of the following components:
+The potential world view consists of the following components: Business Objectives, Threats, hances/Risk, and Security Requirements. 
+\textbf{Business Objectives} of the system owner were already described in Section \ref{subsubsection:it-systems} and \ref{subsec:world-analysis}.
+\textbf{Security Requirement}s are included to our analysis in Section \ref{subsec:priv-rec}.
 
-\begin{itemize}
-
-	\item \textbf{Business Objectives.}
-	Interaction of IT-system and actors that realize a business goals of the system owner were already described in Section \ref{subsubsection:it-systems} and \ref{subsec:world-analysis}.
-
-	\item \textbf{Threats.}
-	A threat is a potential interaction that destroys or harms assets of the system.
-	Concrete realizations of threats can be \emph{attacks} or \emph{accidents}.
-	Attacks are executed by an actor in response to a conflict of interest.
-	Accidents are harmful interactions that are not willfully caused by an actor.
-
-	\item \textbf{Chances/Risk.}
-	Evaluation of chances and risks associated to the business objectives and threats. The risk associated to a threat is its expected loss.
-
-	\item \textbf{Security Requirements.}
-	Interactions (e.g.
-threats) that shall not occur within the system in order to achieve its business objectives.
-We included these security requirements into our analysis in Section \ref{subsec:priv-rec}.
-
-\end{itemize}
-
-\subsubsection{Threat Specification}
+\subsubsection{Threats}
 
@@
 We restrict ourselves to the case of attacks, and the asset of privacy.
-An attack is an interaction that is executed by an actor in response to a conflict of interest by exploiting a vulnerability of the system.
-The alternative interaction of accidents are less relevant for us, since the violation of privacy always requires an actor that takes advantage of personal data.
-
-We describe threats for the citizen privacy in the Live+Gov system.
+An attack is an interaction that is executed by an actor in response to a conflict of interest by exploiting a vulnerability of the system. We describe threats for the citizen privacy in the system.
 This list can necessarily not be complete, but we make a best effort to cover the most relevant cases.
@@
 System Complexity} for System Providers.
-This is an abstract threat provoked by a general \textit{Missing Privacy Awareness} in the minds of all actors in the Live+Gov context.
 
@@
 This could be the case for a System Provider, who wants to test a new product and uses the pre-existing data collection.
-Or for a Local Authority, who wants to analyze the data collection regarding fare evasion.
-However, the Citizen has not agreed to such data processing nor could he, since it is conducted secretly.
-This disables a Citizen to control his data adequately.
+
 Thus there are two possible conflicts: \textit{Privacy vs. Financial Profit of Serivce Provider} and  \textit{Privacy vs. Buisness Intelligence of Local Authority}
@@
 
-\subsubsection{Threat Risk Evaluation}
+\subsubsection{Chances/Risk}
 

### 2015-01-26 paper.tex
references changed paper title changed changed title of section 4.2 abstract added

@@
 
-\title{Privacy Analysis of a Mobile Sensor Application}
+\title{A Privacy Aware Mobile Sensor Application}
 \author{Heinrich Hartmann \and Tim Wambach \and Maximilian Meffert \and R\"udiger Grimm}
@@
 
-\TODO{(later)}
+In our research work we have implemented a privacy aware mobile sensor application [diesen ersten Satz ggf.
+wieder streichen].
+In this paper we outline a scenario of a mobile traffic survey and analyse its privacy risks according to a reference model for IT security analysis.
+This leads to concrete recommendations for privacy protecting measures that are implemented in the system, for example a Web-based dashboard which allows the users to view, modify and delete all personal data that is stored in the system.
 
@@
 
-In this scenario we focus our attention to only one asset: The privacy of the citizen. Our definition of privacy is described in detail in Section \ref{sec:privacydef},
+In this scenario we focus our attention to only one asset: The privacy of the citizen. Our definition of privacy is described in detail in Section \ref{sec:privacydef}.
 
@@
 
-\subsection{Step 2. Potential Analysis}
+\subsection{Step 2. Potential World Analysis}
 \label{subsec:pot-analysis}
 
-The \emph{Potential Analysis} displays the intended and unintended interactions of the components in the world view.
+The \emph{Potential World Analysis} displays the intended and unintended interactions of the components in the world view.
 

### 2015-01-30 paper.tex
conclusion added

@@
 
-\TODO{later.\\
-* Key aspect control, can be given back to citizens via ``Privacy Dashboard''
-}
+Our goal was to find a balance between the demand of the Local Authority for location-based data to improve their public services on one side, and to keep the privacy of citizens as untouched as possible on the other side.
+In a structured manner we identified all actors and what kind of conflict of interests can occour.
+Doing this, we found potential risks and provided finally specific recommendations how to protect the citizen’s privacy.
+One central idea of these recommendations is to give the citizen as much control over their data back as possible.
+This can be e.g.
+done via a Privacy Dashboard that gives a full overview about all personal data that is stored on the system.
+By implementing these privacy recommendations we found a way to satisfy the above-mentioned demand of both sides and comply with legal data protection requirements in addition.
 

### 2015-01-31 paper.tex
Rewrote IT System description.

--- 2015-01-22 paper.tex
@@
 
-\TODO{(later)}
+* The system has been built in Live+Gov EU project
+* Analysis of Privacy aspects following a structured approach
+* Derive eight privacy recommendations
+* Some of them have been implemented in the service
 
@@
 
-\TODO{RG\\
-* Context: Reference model IT Security analysis\\
-* Application: Travel Survey\\
-* Different definitions of privacy in the literature.\\
-* Study outcomes depend on privacy definition
-}
-
-\begin{LGContent}
-In recent days the importance of privacy protection has been amplified
-by the reports about the mass surveillance of ordinary citizens on a
-global scale by the NSA and other intelligence agencies around the
-world.
-
-While aiming at the noble cause of enhancing eParticipation using
-mobile technologies, Live+Gov systems do process a large variety data
-that is potentially infringing the citizens privacy. The captured data
-includes personal information like name, phone numbers and email
-addresses and sensor data from GPS and accelerometer sensors. Also
-with some applications it is possible to gather images and textual
-input from the citizen.
-
-While the collection of this data is necessary for providing the
-advanced services that Live+Gov aims to deliver, at he same time, the
-available raw data can be used to draw a very detailed picture of the
-private life of the citizen. For instance can GPS location tracking be
-used to reveal shopping habits (e.g. when a car seller is visited) and
-associations to political groups (when a meeting is
-attended). Accelerometer data can be used to infer medical conditions
-like walking disabilities. Images can contain faces of nearby persons
-to with whom the citizen is associated. All this data is highly
-sensitive to the citizens privacy and can be used against the citizen
-if it falls in the wrong hands.
-
-The great importance of protecting the citizens privacy should be
-apparent from these examples. The European Union, as well as many
-other countries in the past, has set out a number of directives that
-regulate the collection, processing and use of privacy sensitive
-data. We explain the most relevant legislation in Section
-\ref{sec:legal}.
-
-The ethical aspects of privacy have been the subject of study of many
-social scientists and philosophers. One scholar which is particularly
-relevant in our context is Charles Fried. He investiages, why we are
-intuitively so sensitive to violations of our privacy. For him privacy
-is not asserted as an intrinsic value by itself, he rather sated:
-
-\begin{quote}
-Privacy is not simply an absence of information about us in the minds of others;
-rather it is the control we have over information about ourselves.
-\end{quote}
-
-Fried`s study on the understanding of privacy provided a great
-contribution to the research on the same term in philosophy and
-computer science and despite the fact his text was published in 1970,
-he already included technologies to its viewpoint (like location
-monitoring) that are particularly relevant to our Context.  We explain
-his theory in \ref{sec:ethics} and follow his definition of privacy in
-this document.
-
-In this document we perform a thorough analysis of how to protect the
-privacy of a citizen and present several implementations that form the
-core of our privacy aware Sensor Data Storage and Mining Infrastructure.
-
-We identify six main threats for the citizens privacy, and derive
-eight recommendations that should be followed in order to reduce the
-associated risks for hazards.  This analysis form the guideline for
-the selection of 9 measures that were implemented in our system and
-documented in Chapter \ref{chap:impl}.
-\end{LGContent}
+* Privacy Protection very important topic in current time Fundamental
+* Fundamental trade-off between enhanced functionality and convenience
+  of the service (service demands) and personal data revealed by the
+  use (privacy demands).
+* Challenge for service provider. Find the right balance.
+* We study concrete use case of a travel-survey application.
 
@@
 
-\subsection{A Privacy Respecting Mobile Traffic Survey}
-
-The scenario of our privacy analysis is a mobile traffic survey that supports local travel services by collecting and evaluating mobile sensor data of their customers.
-The survey is executed by a trustworthy third-party service.
-The travel services need the personal mobile data for two reasons: firstly, they want to serve their customers individually, for example by guiding them to proximity stations, to recommend adequate travel connections, to inform about delays, or to sell electronic tickets.
-For this purpose, the personal data for evaluation are related to the individual customers.
-Secondly, the travel services want to enhance their system, in that they evaluate the demand of stations, connections, changes, ways and time spent by their customers.
-For this purpose, personal data are anonymized before evaluation.
-
-\subsection{IT-Systems \& Interactions}
+\subsubsection{IT-Systems \& Interactions}
 \label{subsubsection:it-systems}
 
-\begin{LGContent}
-This section outlines the general architecture (Figure \ref{figure:IT Systems}) of IT systems for public monitoring comparable to the Live+Gov project.
-This includes a description of it technical infrastructure and the interactions between its components.
+Travel surveys are regularly conducted by providers of public
+transportation infrastructure in order to asses the current usage of
+their services and inform future planning of e.g. bus or tram lines.
+In this events a large number of citizens (e.g. 5000) is asked to keep
+track of their travel patterns for a constraint period of time
+(e.g. one month). Traditionally the travel patterns have been recorded
+manually in the form of travel diaries. Modern smart-phones are
+equipped with a wide range of sensors (in particular GPS) that allow
+the recording of travel-patterns in a fully automated way.
 
-The IT infrastructure of the Live+Gov system, i.e.
-the Live+Gov toolkit and the customization of the software components are described in detail in various project deliverables: D4.1, D4.3, D1.1, D5.1.
-In this section we give an abstraction of those systems from the perspective of WP1.
+Within the Live+Gov EU project, we have established a first
+prototypical implementation of such an automated travel-survey
+service. In this section we give a high level description of a generic
+travel-survey service. This description along with the introduced
+terminology will serve as a basis for our privacy analysis.
 
-%input{./figures/IT-Systems}
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.6\textwidth]{diagrams/png/it-systems.png}
+There are three stakeholders involved with the sysem: The
+\emph{citizen}, the \emph{service provider} and the \emph{travel
+  agency}. The service provider is the operator of all IT systems and
+sells the travel-survey service to the travel agency.
 
-%\begin{flushleft}
-%\scriptsize
-%\textbf{Legend:}
-%\begin{itemize}
-%\itemsep1pt\parskip0pt\parsep0pt
-%
-%\item
-%\textbf{(a) Citizen:} 
-%User of the L+G client application whose privacy is at stake.
-%
-%\item
-%\textbf{(b) Service Provider:} Provides technical infrastructure.
-%
-%\item
-%\textbf{(c) Local Authority:} Provider of the L+G system.
-%
-%\item
-%\textbf{(d) Mobile Device:}
-%Runs the L+G client application, produces and stores data sensitive to the users privacy.
-%
-%\item
-%\textbf{(e) Data Center:} Runs the L+G services, processes and stores user data.
-%
-%\item
-%\textbf{(f) Report Tool:} Interface to aggregated user data.
-%
-%\end{itemize}
-%\end{flushleft}
+The citizen is a volunteer who is willing to share his personal travel
+patterns within the travel-survey event.  He carries a mobile
+smart-phone which runs the \emph{sensor collector application}. The
+application collects data from various sensors available on modern
+smart-phones. In particular accelerometer- and GPS-samples are
+collected. Based on the accelerometer data, the application extracts
+human activities (e.g. running, walking, standing). Also, the
+application can send raw-sensor data back to a central data center.
 
-\caption{IT Systems}
-\label{figure:IT Systems}
-\end{figure}
+The \emph{data center} stores and processes sensor data collected with
+the sensor collector application. It can also take into account data
+obtained from third parties, like the current positions of trains. In
+particular, the data center determines if the citizen is using public
+transportation and if so which lines of service (bus, train, tram) are
+used. The data center can also sent mining end products and messages
+(e.g. traffic jam reports, bus schedule) back to the mobile device of
+the citizen.
 
-A \emph{citizen} (a) carries a mobile device running the \emph{sensor collector} application (d).
-The sensor collector application is able to collect various kinds of sensor data (accelerometer, GPS, GSM, ...) and can sent the raw data back to the \emph{data center} (c).
-The sensor collector can also be able to perform certain data mining operations.
-
-Examples of such data mining operations include human activity recognition,the detection of service lines, detection of characters and faces from images.
-
-In the Live+Gov Project mobile devices are used in particular for the following activities:
-\begin{enumerate}
-\item collection of GPS samples,
-\item mining of Human activities (HAR) based on accelerometer samples,
-\item collection of reports consisting of an image, free text and selected categories.
-\end{enumerate}
-
-The \emph{data center} stores and processes sensor data collected with the sensor collector application.
-It can also take into account data obtained from third parties, like the current positions of trains.
-The data center can sent mining end products (traffic jam reports, bus schedule) back to the mobile device of the citizen.
-
-In the Live+Gov Project data centers, in particular, perform the following activities:
-\begin{enumerate}
-\item storage of login credentials, name, and email address for each citizen,
-\item storage of collected GPS samples (user id, timestamp, GPS location),
-\item storage of HAR results (user id, timestamp, HAR result),
-\item detection of service lines based received GPS samples (SLD),
-\item storage of SLD results (userid, timestamp, SLD result),
-\item sent back SLD results to mobile device,
-\item storage of received reports (user id, timestamp, report),
-\item detection of inherent patterns of the received reports.
-\end{enumerate}
-
-The \emph{System Provider} (b) provides and operates technical infrastructure like the Data Center and the \emph{Reporting Tool} (f).
-The Reporting Tool queries the Data Center for aggregated data to visualize in form of charts and other means suitable to help understanding of monitored citizens.
-
-\textit{Local Authorities} (c) use the reporting tool to get information in order to understand citizen movement and improve public services.
-In such systems, the most valuable information for local authorities is not the raw sensor data, but aggregated views on the mining end products.
-
-In the Live+Gov Project the reporting tools allows, in particular the following queries:
-\begin{enumerate}
-\item show aggregate information about which routes citizens take starting from a given location,
-\item show the average waiting time of a citizen for each bus stop,
-\item show routes where citizens where running to catch a bus,
-\item show locations of all reports in a given time window,
-\item visualize detected patterns in the report data,
-\item view individual reports.
-\end{enumerate}
-\end{LGContent}
+The \emph{service provider} provides and operates all technical
+infrastructure like the data center and the sensor application. He
+also operates a web-based \emph{reporting tool} that aggregates
+information about travel patterns of citizens within the
+travel-survey. The reporting tool is used by the travel agency to gain
+insights for planning and accounting of the public transportation
+infrastructure.
 
@@
 
-\subsection{Privacy and Self Data Protection}
-More than one hundred years ago, 1890, the American politician Louis Brandis had specified privacy as the right to be left alone \cite{WarrenBrandeis:RightToPrivacy}.
-From the data processing point of view, this right is best expressed by the absence of information about a person in the mind of others.
-Indeed, this principle of “data minimization” is still fundamental to modern data protection legislation.
-The modern principles of data protection are codified by many legal systems in different countries, especially in Europe \cite{EU-directive2}, and by the US save harbor principles.
-However, modern data protection is more than only the absence of personal data.
-It is based on the personal right on self-determination over information and communication \cite{BVerfG:census}.
-This is, for example, very well expressed by Fried’s definition of 1970:
+%\TODO{MM:\\
+%- Shorten\\
+%- DONE - Remove Implicit Type Violations\\
+%- Update figure\\
+%- RG: Discuss alternative definitions of privacy\\
+%}
+
+\noindent
+Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes:
 \begin{quote}
-	Privacy is not simply an absence of information about us in the minds of others; rather it is the control we have over information about ourselves. \cite{CFried:Privacy}
+The notion of privacy remains out of the grasp of every academic chasing it.
+Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.
+\cite{Gutwirth}
 \end{quote}
-In order to realize this requirement in the modern IT world users are provided with user control functions, for example to decide about the access on their data for future use, and to view, modify, and delete unwanted data that had been collected.
-In its fundamental decision of May 13, 2014, the European Court has convicted Google to accept user demands to delete links to incorrect or irrelevant personal data in the Internet (C-131/12 – Google Spain SL).
-Such functions would constitute a so-called self-data-protection.
-It requires users to be highly aware of their rights and how to use them.
-Complementary, system-data-protection puts the load of data protection enforcement on the data collectors, ideally even without bothering the users.
-Typical system data protection functions are the deletion of personal data that are not bound to service purpose, the anonymization of personal data for research, and the abstinence from forwarding personal data to third parties.
-Together, system-data-protection and self-data-protection are supposed to provide a fair balance between the interests of service providers and their users.
-In our research work we focus on self-data-protection according to Fried’s user control requirement \cite{CFried:Privacy}.
-A system data protection point of view would lead to other results, namely to a set of obligations for service providers.
-This is subject to further research.
+In this section we introduce the concept of privacy as control over information.
+We specify types of information we have to deal with in the context of privacy.
+And finally we outline the impact mobile sensors can have on one's privacy.
+
+\subsection{Privacy as Control over Information}
+%We may need to remove the following paragraph depinding on RG's alternative definition of privacy
+%Many definitions seem to focus on the vulnerabilities of privacy.
+%Thus a great amount of work is invested in finding threats to prohibit instead of describing why privacy is so valuable to us.
+%This way around one could derive measures to ensure and secure its value. \cite{7ToP}
+
+%One particular scholar, who did not follow this way of thinking is Charles Fried.
+One possible definition of privacy is given by Charles Fried in his work \textit{An Anatomy of Values}\cite{CFried:Privacy}.
+He questioned, why we are intuitively sensitive to privacy violations.
+But he did not assert privacy as an intrinsic value by itself, he rather stated:
+\begin{quote}
+\textbf{Privacy is} not simply an absence of information about us in the minds of others;
+rather it is \textbf{the control we have over information about ourselves}. \cite{CFried:Privacy}
+\end{quote}
+According to Fried privacy is our ability to create and modulate the relationships to other humans.
+Which makes us intuitively sensitive for its violation, because this affects what defines our identity.
+Also, the ability to build and maintain relationships is essential to society.
+This is what makes privacy valuable and deserving of protection. 
+\cite{sep-privacy}\cite{CFried:Privacy}
+
+\subsubsection{Relationship Creation.}
+
+% Privacy + Trust ~> Relationship
+% ===============================
+% - A shares intimate information with B
+% - A trusts B not to reveal that information by not constantly monitoring B, i.e. respecting B's privacy
+% - B has now the possibility to share intimate information with A
+
+We share information of great intimacy only with persons we deem trustworthy. 
+Moreover, we trust those persons to not reveal information about us to others by not constantly monitoring them, i.e. respecting their privacy.
+Trust needs the possibility of unknown failure.
+If we would constantly monitor our partners, they cannot fail unnoticed nor can they willingly share intimate information with us.
+By trusting them, we create the possibility for them to share intimate information with us.
+\cite{CFried:Privacy}
+
+\subsubsection{Relationship Modulation.}
+Fried argues, depending on the person we are interacting with, we are changing the degree of intimacy we share.
+If we are to share private information with others, we are most likely aware that this action is privacy related.
+In that case we are able to selectively disclose information along the two dimensions of quantity and quality.
+\cite{CFried:Privacy}
+
+\subsubsection{Alternative Definitions.}
+Using Fried's anatomy of privacy for this analysis is suitable because of two points.
+Despite the fact his text was published in 1970, he already included technologies to its viewpoint that did not only monitor location, but also record biometric data \cite{CFried:Privacy}. 
+Secondly, the concept of control over information does not enforce data minimization.
+Fried's definition complies with what we call \textit{self data protection}, where persons communicate their data self-determined.
+This is the opposite of \textit{system data protection}, where the collection of certain data is prohibited.
+
+\TODO{RG: Discuss alternative definitions of privacy}
+
+%=======================================================================================
+
+\begin{LGContent-Hidden}
+Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes:
+
+\begin{quote}
+The notion of privacy remains out of the grasp of every academic chasing it.
+Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.
+\cite{Gutwirth}
+\end{quote}
+
+%Many researchers seem to only ``focus on the ways in which privacy can be infringed'' \cite{7ToP}.
+%Thus they try to create taxonomies of \emph{privacy harms} instead of taxonomies of \emph{privacy types}.
+%Those two differ in the respect that the former focuses on threats to prohibit whereas the latter focuses on values to protect.
+%So one should rather evaluate what aspects are precious about privacy and develop measures to ensure their security than only forbid single actions against it \cite{7ToP}.
+
+Many researchers seem to only ``focus on the ways in which privacy can be infringed'' \cite{7ToP}.
+Thus they invest a great amount of work in defining threats to prohibit instead of describing why privacy is so valuable to us.
+This way around one could derive measures to ensure and secure its value \cite{7ToP}.
+One particular scholar who does this in the context of digital monitoring is Charles Fried.
+He questioned, why we are intuitively sensitive to violations of privacy.
+But he did not assert privacy as an intrinsic value by itself, he rather sated:
+
+\begin{quote}
+Privacy is not simply an absence of information about us in the minds of others;
+rather it is \textbf{the control we have over information about ourselves}. \cite{CFried:Privacy}
+\end{quote}
+
+According to Fried privacy is a \textit{rational context}.
+Which means that one is aware of such context's existence during a rational action.
+If we are to share private information with others, we are most likely aware that this action is privacy related.
+In that case we are able to selectively disclose information along the two dimensions of quantity and quality.
+This leads Fried to his thesis, that privacy is one's ability to create and modulate his or hers social relationships, namely: friendship, love and trust. \cite{CFried:Privacy}
+
+Depending on the conversation partner, we change the degree of intimate information we share if it is a total stranger, colleague, close friend or lover.
+With close friends or lovers we share information of great intimacy we do not share with anyone else.
+Moreover, we trust those persons to not reveal information about us to others by respecting their privacy.
+Trust needs the possibility of unknown failure.
+If we would constantly monitor our partners, they cannot fail unnoticed nor can they willingly share that information with us.
+Thus they could not trust us anymore. \cite{CFried:Privacy}
+
+So privacy or the its possibility, according to Fried, is the foundation of our core relations: friendship, love and trust.
+And thus it is valuable, because those relations are essential to human society \cite{CFried:Privacy}.
+Using his anatomy of privacy as foundation of our analysis is suitable because of two points.
+At first, Fried's study on the understanding of privacy provided a great contribution to the research on the same term in philosophy \cite{sep-privacy} and computer science \cite{SociotechnicalArchitectureForOnlinePrivacy}.
+Secondly, despite the fact his text was published in 1970, he already included technologies to its viewpoint that did not only monitor location, but also record biometric data.
+\end{LGContent-Hidden}
 
@@
 
-In Fried's definition of privacy as control over information, the specification of what constitutes such information remains open.
+In Fried's definition of privacy as control over information, the specification of what constitues such information remains open.
 There is a vast amount of information that relates to a person and we need to get a better understanding in order to perform a thorough analysis.
-To this end we use the categorization by Friedewald, Finn and Wright of 2013 called the Seven Types of Privacy \cite{7ToP}.
-These are an extension to the four types of privacy by Roger Clarke of 1997 \cite{RClarke:4ToP}.
-It is important to note, that these categories are not mutually exclusive.
+To this end we use the the categorization by Friedewald, Finn and Wright \cite{7ToP} called the \textit{Seven Types of Privacy}.
+The seven types of privacy are an extension to the four types of privacy by Roger Clarke \cite{RClarke:4ToP}, which are:
+\textit{Privacy of the Person},
+\textit{Privacy of Personal Behaviour},
+\textit{Privacy of Personal Communication},
+\textit{Privacy of Personal Data}.
+It is important to note, that these categories do not form a taxonomy, since the categories are not mutually exclusive.
 For instance a written email is considered personal communication as well as personal data stored on a computer.
-However, the categories are very helpful for a privacy analysis with a focus on self-data-protection.
-The seven types of privacy are as follows:
+
+Moreover, Friedewald et al. argue that Clarke's types are outdated and no longer adequate in order to describe the privacy aspect of our modern, technology driven, world.
+In order to address this shortcoming they extend the former four to the now introduced seven types privacy as follows:
+
 \begin{enumerate}
-	\item Privacy of the Person (with respect to body functions and body characteristics)
-	\item Privacy of Behavior and Action
-	\item Privacy of Communication
-	\item Privacy of Data and Image
-	\item Privacy of Thoughts and Feelings
-	\item Privacy of Location and Space 
-	\item Privacy of Association (with persons, e.g. friends, and organizations, e.g. political parties)
+
+\item \textbf{Privacy of the Person}
+This privacy type is generally concerned with one could best understand as biometric privacy.
+Friedewald et al. paraphrase it as \emph{``\om the right to keep body functions and body characteristics \om private''}.
+This includes but is not limited to measures like weight, height or shoulder width;
+biometric identifiers like fingerprints and DNA sequences;
+or medical conditions such as limping or having a cold.
+
+\item \textbf{Privacy of Behaviour and Action}
+This privacy type is concerned with one's activities in public as well as in private spaces.
+It includes but is not limited to religious practices, political activities and sexual preferences or habits.
+
+\item \textbf{Privacy of Communication}
+This privacy type is concerned with one's communication in a broad sense.
+It includes written correspondence, but also conversations conducted either vis-a-vis or via electronic devices.
+Friedewald et al. put it as the right to free discussion without unknown interception by third parties.
+
+\item \textbf{Privacy of Data and Image}
+This privacy type is concerned with the secrecy of personal data, especially its automatic disclosure to other individuals and organizations.
+It includes data such as paychecks, insurance information or records of public administration.
+However, it also refers to pictures taken without consent and digital identifiers like IP addresses or social security numbers.
+
+\item \textbf{Privacy of Thoughts and Feelings}
+This privacy type is the counterpart to Privacy of the Person like body and mind are counterparts of one another.
+Comparable to the Privacy of Data and Image, Friedewald et al. state that one's thoughts and feelings must not be automatically revealed to others.
+This could simply happen by the disclosure of one's diary or by technologies which allow emotion detection through biometric means.
+One's body temperature or iris reflexes might infer stress or excitation.
+
+\item \textbf{Privacy of Location and Space}
+This privacy type is concerned with one's movements in public spaces and the protection of private spaces.
+Friedewald et al. qualify the first dimension as one's right to move without being identified, tracked or monitored.
+The second dimension is qualified as one's general right to solitude, especially the right to the inviolability of the home.
+
+\item \textbf{Privacy of Association}
+This privacy type is also put as group privacy.
+Friedewald et al. state that one must have the possibility with whomever without being recorded.
+Associations like friends or organizations such as political parties must not automatically be recorded because one associates with them, and vice-versa.
 \end{enumerate}
+
+\begin{LGContent-Hidden}
+In Fried's definition of privacy as control over "information about onself'', the specification of what constitues such information remains open.
+There is a vast amount of information that relates to a person and we need to get a better understanding in order to perform a thorough analysis.
+To this end we use the the categorization by Friedewald, Finn and Wright \cite{7ToP} called the \textit{Seven Types of Privacy}.
+The seven types of privacy are an extension to the four types of privacy by Roger Clarke \cite{RClarke:4ToP}, which are:
+\textit{Privacy of the Person},
+\textit{Privacy of Personal Behaviour},
+\textit{Privacy of Personal Communication},
+\textit{Privacy of Personal Data}.
+It is important to note, that this categories do not form no taxonomy, since the categories are not mutually exclusive.
+For instance a written email is considered personal communication as well as personal data stored on a computer.
+
+Moreover, Friedewald et al. argue that Clarke's taxonomy is outdated and no longer adequate in order to describe the privacy aspect of our modern, technology driven, world.
+In order to address this shortcoming they extend the former four to the now introduced seven types privacy as follows:
+
+\begin{enumerate}
+
+\item \textbf{Privacy of the Person}
+This privacy type is generally concerned with one could best understand as biometric privacy.
+Friedewald et al. paraphrase it as \emph{``\om the right to keep body functions and body characteristics \om private''}.
+This includes but is not limited to measures like weight, height or shoulder width;
+biometric identifiers like fingerprints and DNA sequences;
+or medical conditions such as limping or having a cold.
+
+\item \textbf{Privacy of Behaviour and Action}
+This privacy type is concerned with one's activities in public as well as in private spaces.
+It includes but is not limited to religious practices, political activities and sexual preferences or habits.
+
+\item \textbf{Privacy of Communication}
+This privacy type is concerned with one's communication in a broad sense.
+It includes written correspondence, but also conversations conducted either vis-a-vis or via electronic devices.
+Friedewald et al. put it as the right to free discussion without unknown interception by third parties.
+
+\item \textbf{Privacy of Data and Image}
+This privacy type is concerned with the secrecy of personal data, especially its automatic disclosure to other individuals and organizations.
+It includes data such as paychecks, insurance information or records of public administration.
+However, it also refers to pictures taken without consent and digital identifiers like IP addresses or social security numbers.
+
+\item \textbf{Privacy of Thoughts and Feelings}
+This privacy type is the counterpart to Privacy of the Person like body and mind are counterparts of one another.
+Comparable to the Privacy of Data and Image, Friedewald et al. state that one's thoughts and feelings must not be automatically revealed to others.
+This could simply happen by the disclosure of one's diary or by technologies which allow emotion detection through biometric means.
+One's body temperature or iris reflexes might infer stress or excitation.
+
+\item \textbf{Privacy of Location and Space}
+This privacy type is concerned with one's movements in public spaces and the protection of private spaces.
+Friedewald et al. qualify the first dimension as one's right to move without being identified, tracked or monitored.
+The second dimension is qualified as one's general right to solitude, especially the right to the inviolability of the home.
+
+\item \textbf{Privacy of Association}
+This privacy type is also put as group privacy.
+Friedewald et al. state that one must have the possibility with whomever without being recorded.
+Associations like friends or organizations such as political parties must not automatically be recorded because one associates with them, and vice-versa.
+\end{enumerate}
+
+\end{LGContent-Hidden}
 
@@
 
-The goal of this chapter is to analyze and identify the threats to personal privacy that are posed by collecting, storing and processing sensor data from mobile phones.
-We derive concrete privacy protection measures that address the main risks involved with handling such data.
+The goal of this chapter is to analyze and identify the threads to
+personal privacy that are posed by collecting, storing and processing
+sensor data from mobile phones.  We derive concrete privacy protection
+measures that address the main risks involved with handling such data.
 
-In our analysis we follow the “reference model for IT security analysis” as described in \cite{Grimm:ItSecRefModel}.
-It supersedes earlier efforts by e.g.
-\cite{Avizienis}.
-The reference model consists of a model and a procedure.
-The model organizes a common security terminology in a reasonable and practical way.
-The procedure describes a method for analyzing the IT system based on that model.
-The reference model provides four views: (1) the real world of persons and their assets, (2) the potential world of requirements and threats, (3) measures and plans specified by programs, business models and attack strategies, and, finally, (4) events of running programs, data accesses and performed attacks as well as their defense.
-In the following sections we apply the proposed procedures of the reference model to our scenario of a travel service that collects personal data from mobile users in order to serve the users and to enhance the service.
-The service intends to respect the privacy of its users.
+In section \ref{subsec:world-analysis} and \ref{subsec:pot-analysis} we apply the first two steps of the Reference Model for IT Security Analysis as described in \cite{Grimm:ItSecRefModel}.
+It supersedes earlier efforts by
+e.g. \cite{Avizienis}.  The reference model consists of a \emph{model}
+and a \emph{procedure}: The model aims to organize common security terminology in a reasonable and practical way.
+The procedure describes a method to analysis the IT system based on that model.
+In these sections we give a brief overview over the reference model by applying it to our scenario.
 
-In sections \ref{subsec:world-analysis} and \ref{subsec:pot-analysis} we apply the first two steps of the reference model \cite{Grimm:ItSecRefModel}, which are related to the views on the real world and on the measures and plans.
-In section \ref{subsec:priv-rec} we apply the third step of the reference model by providing specific privacy recommendations and requirements as a result of the previous analysis that the system must comply with.
+In section \ref{subsec:priv-rec} we provide specific privacy recommendations and requirements as a result of the previous analysis that the system must comply with.
 
@@
 
-The first step is the \emph{world view} where all components are described in their current state.
-It consists of the following components: Assets, IT-Systems, Actors, Conflicts of Interests, Vulnerabilities, and Interactions.
+The first step is the \emph{world view} where all components are described in their current state. It consists of the following components:
+\begin{itemize}
 
-The relevant \textbf{IT-Systems} were already described in Section \ref{subsubsection:it-systems}.
-\textbf{Interactions} between assets, IT-Systems, humans, and vulnerabilites as partly described in Section \ref{subsubsection:it-systems} and will be further analyzed in Section \ref{subsec:pot-analysis}.
+	\item \textbf{Assets.} 
+	Things of value to one or more stakeholders.
+
+	\item \textbf{IT-Systems.}
+	The relevant IT-Systems under study. Already described in Section \ref{subsubsection:it-systems}.
+
+	\item \textbf{Actors.}
+	All identifiable stakeholders of the system under study.
+
+	\item \textbf{Conflicts of Interests.}
+	Different actors have different interests which can be in conflict which each other.
+
+	\item \textbf{Vulnerabilities.}
+	All identifiable weaknesses in the IT-System e.g. a weak passwords.
+
+	\item \textbf{Interactions.}
+	This point captures all possible interactions between assets, IT-Systems, humans, and vulnerabilites as partly described in Section \ref{subsubsection:it-systems} and will be further analyzed in \ref{subsec:pot-analysis}.
+
+\end{itemize}
 
@@
 
-In this scenario we focus our attention to only one asset: The privacy of the citizen. Our definition of privacy is described in detail in Section \ref{sec:privacydef},
+In this scenario we focus our attention to only one asset: The privacy of the citizen.
+Our definition of privacy is described in detail in Section \ref{sec:privacydef} and introduced in Section \ref{sec:privacytypes} the seven different types of privacy.
 
@@
 This section describes the human actors previously introduced within the IT system architecture (\ref{subsubsection:it-systems}).
+Also the additional actor \textit{External} is introduced as a person with no special access rights.
 Although the text is only a short description, a list of the most important interests is provided for each actor.
@@
 Local authorities are public offices (ministry, agency, department, ...) or other external public entities which act as direct customers of service providers.
+They purchase a system specialized for their needs.
 For example a department for urban mobility, orders a system to better understand usage patterns and make improvement to the urban traffic flow.
@@
 This section outlines the vulnerabilities (Figure \ref{figure:Live+Gov Vulnerabilities}) of the proposed monitoring and mining system.
+Note that vulnerabilities are not necessarily of technical nature.
+The weaknesses of IT systems are often created due to misuse or misconfiguration of the various components by one or more actors.
 
@@
 
-The \emph{Potential Analysis} displays the intended and unintended interactions of the components in the world view.
+The \emph{Potential Analysis} (or \emph{Potential View}) displays the intended and unintended interactions of the components in the world view.
+The intended interactions support the underlying business objectives.
+Unintended interactions lead to threats. The potential view consists of the following components:
 
-The intended interactions support the underlying business objectives.
-Unintended interactions can lead to threats.
-Obviously this provides a conflict of interest with the victim's business model, to keep the asset away from unauthorized access.
-From the point of view of the victim, an attack would be an unintended interaction.
+\begin{itemize}
 
-The potential world view consists of the following components: Business Objectives, Threats, hances/Risk, and Security Requirements. 
-\textbf{Business Objectives} of the system owner were already described in Section \ref{subsubsection:it-systems} and \ref{subsec:world-analysis}.
-\textbf{Security Requirement}s are included to our analysis in Section \ref{subsec:priv-rec}.
+	\item \textbf{Business Objectives.}
+	Interaction of IT-system and actors that realize a business goals of the system owner were already described in Section \ref{subsubsection:it-systems} and \ref{subsec:world-analysis}.
 
-\subsubsection{Threats}
+	\item \textbf{Threats.}
+	A threat is a potential interaction that destroys or harms assets of the system.
+	Concrete realizations of threats can be \emph{attacks} or \emph{accidents}.
+	Attacks are executed by an actor in response to a conflict of interest.
+	Accidents are harmful interactions that are not willfully caused by an actor.
+
+	\item \textbf{Chances/Risk.}
+	Evaluation of chances and risks associated to the business objectives and threats. The risk associated to a threat is its expected loss.
+
+	\item \textbf{Security Requirements.}
+	Interactions (e.g.
+threats) that shall not occur within the system in order to achieve its business objectives.
+We included these security requirements into our analysis in Section \ref{subsec:priv-rec}.
+
+\end{itemize}
+
+\subsubsection{Threat Specification}
 
@@
 We restrict ourselves to the case of attacks, and the asset of privacy.
-An attack is an interaction that is executed by an actor in response to a conflict of interest by exploiting a vulnerability of the system. We describe threats for the citizen privacy in the system.
+An attack is an interaction that is executed by an actor in response to a conflict of interest by exploiting a vulnerability of the system.
+The alternative interaction of accidents are less relevant for us, since the violation of privacy always requires an actor that takes advantage of personal data.
+
+We describe threats for the citizen privacy in the Live+Gov system.
 This list can necessarily not be complete, but we make a best effort to cover the most relevant cases.
@@
 System Complexity} for System Providers.
+This is an abstract threat provoked by a general \textit{Missing Privacy Awareness} in the minds of all actors in the Live+Gov context.
 
@@
 This could be the case for a System Provider, who wants to test a new product and uses the pre-existing data collection.
-
+Or for a Local Authority, who wants to analyze the data collection regarding fare evasion.
+However, the Citizen has not agreed to such data processing nor could he, since it is conducted secretly.
+This disables a Citizen to control his data adequately.
 Thus there are two possible conflicts: \textit{Privacy vs. Financial Profit of Serivce Provider} and  \textit{Privacy vs. Buisness Intelligence of Local Authority}
@@
 
-\subsubsection{Chances/Risk}
+\subsubsection{Threat Risk Evaluation}
 

### 2015-01-31 paper.tex
Rewrote Introduction.

@@
 
-\keywords{
-privacy protection, IT security analysis, sensor data, mobile phones, traffic survey
-}
-
 \end{abstract}
@@
 
-* Privacy Protection very important topic in current time Fundamental
-* Fundamental trade-off between enhanced functionality and convenience
-  of the service (service demands) and personal data revealed by the
-  use (privacy demands).
-* Challenge for service provider. Find the right balance.
-* We study concrete use case of a travel-survey application.
+The rise of mobile smart-phones equipped with a wide range of sensors
+and the abundance mobile internet connectivity has paved the way to a
+whole new generation of mobile services which ease the daily life of
+citizens of the western developed world. Apps like Google
+Maps\footnote{\url{http://google.com/maps}} allows the user to
+navigate effectively in unknown cities. Others track sports activities
+in order to optimize training plans or engage in virtual competitions
+(e.g. Strava\footnote{Citation Needed}).
+
+Besides the immediate benefits, citizens using these services, reveal
+a wealth of which can be used to draw a very detailed picture of the
+private life of the citizen. For instance, GPS location tracking can
+be used to infer shopping habits or associations with political groups
+(when meetings are attended). Accelerometer data can be used to detect
+medical conditions like walking disabilities. All this data is highly
+sensitive to the citizens privacy and can be used against the citizen
+if it falls into the wrong hands.
+
+Providers of these new services are faced with a fundamental trade-off
+between features and convenience of the service and the protection of
+the citizen's privacy. The service has to collect enough data to
+support the basic service-promises. If too much data is collected the
+privacy of the citizen is put at risk, with immediate implications for
+the customer trust relation and the acceptance of the service. Also
+leaks of private data can jeopardize the whole business.
+
+In this article we study the privacy risks of a travel-survey service
+for urban mobility.
 
@@
 \section{Scenario Description: Mobile Traffic Survey}
-
-\TODO{HH:\\
-* System description: (Raw version copied from Privacy Analysis)\\
-* Information needs of transport agency\\
-}
-
-\subsubsection{IT-Systems \& Interactions}
-\label{subsubsection:it-systems}
+\label{sec:ScenarioDescription}
 
@@
   agency}. The service provider is the operator of all IT systems and
-sells the travel-survey service to the travel agency.
+sells the travel-survey service to the travel agency. Figure
+\ref{figure:Live+Gov Vulnerabilities} contains a schematic
+visualization of the relation between these stakeholders.
 

### 2015-01-31 paper.tex
Added abstract draft.

@@
 
-\title{Privacy Analysis of a Mobile Sensor Application}
+\title{Privacy Aware Mobile Sensor Applications}
 \author{Heinrich Hartmann \and Tim Wambach \and Maximilian Meffert \and R\"udiger Grimm}
@@
 
-* The system has been built in Live+Gov EU project
-* Analysis of Privacy aspects following a structured approach
-* Derive eight privacy recommendations
-* Some of them have been implemented in the service
+In this article we analyze the privacy aspects of a mobile sensor
+application used for recording urban travel patterns as part of a
+travel-survey service. This service has been developed and
+field-tested within the Live+Gov EU Project. The privacy analysis
+follows a structured approach established in
+\cite{Grimm:ItSecRefModel}. Eight privacy recommendations are derived,
+and have already led to corresponding enhancements of the
+travel-survey service.
 
@@
 Maps\footnote{\url{http://google.com/maps}} allows the user to
-navigate effectively in unknown cities. Others track sports activities
-in order to optimize training plans or engage in virtual competitions
-(e.g. Strava\footnote{Citation Needed}).
-
-Besides the immediate benefits, citizens using these services, reveal
+navigate effectively in unknown cities. Others
+(e.g. Strava\footnote{Citation Needed}) track sports activities in
+order to optimize training plans or engage in virtual competitions .
+
+Besides the immediate benefits citizens using these services reveal
 a wealth of which can be used to draw a very detailed picture of the
@@
 
-Travel surveys are regularly conducted by providers of public
-transportation infrastructure in order to asses the current usage of
-their services and inform future planning of e.g. bus or tram lines.
-In this events a large number of citizens (e.g. 5000) is asked to keep
+Travel surveys are regularly conducted travel agencies which provide
+public transportation infrastructure. With these surveys the current
+usage of the public transportation system is assessed and the insights
+are used for future planning of bus, tram and subway lines, etc..  For
+each survey a large number of citizens (e.g. 5000) is asked to keep
 track of their travel patterns for a constraint period of time
@@
 prototypical implementation of such an automated travel-survey
-service. In this section we give a high level description of a generic
-travel-survey service. This description along with the introduced
-terminology will serve as a basis for our privacy analysis.
-
-There are three stakeholders involved with the sysem: The
-\emph{citizen}, the \emph{service provider} and the \emph{travel
+service. In this section we give a high level description of the
+generic service architecture. This description along with the
+introduced terminology will serve as a basis for our privacy analysis.
+
+There are three stakeholders involved in the travel-survey service:
+The \emph{citizen}, the \emph{service provider} and the \emph{travel
   agency}. The service provider is the operator of all IT systems and
-sells the travel-survey service to the travel agency. Figure
+offers the travel-survey service to the travel agency. Figure
 \ref{figure:Live+Gov Vulnerabilities} contains a schematic

### 2015-02-01 paper.tex

@@
 
-\title{Privacy Aware Mobile Sensor Applications}
+\title{A Privacy Aware Mobile Sensor Application}
+
 \author{Heinrich Hartmann \and Tim Wambach \and Maximilian Meffert \and R\"udiger Grimm}
@@
 
-In this article we analyze the privacy aspects of a mobile sensor
+In our research work we have implemented a privacy aware mobile sensor application [diesen ersten Satz ggf. wieder streichen]. 
+In this paper we outline a scenario of a mobile traffic survey and analyse its privacy risks according to a reference model for IT security analysis.
+This leads to concrete recommendations for privacy protecting measures that are implemented in the system, for example a Web-based dashboard which allows the users to view, modify and delete all personal data that is stored in the system.
+
+Vorschlag HH: In this article we analyze the privacy aspects of a mobile sensor
 application used for recording urban travel patterns as part of a
@@
 travel-survey service.
+
+\keywords{
+privacy protection, IT security analysis, sensor data, mobile phones, traffic survey
+}
 
@@
 whole new generation of mobile services which ease the daily life of
-citizens of the western developed world. Apps like Google
-Maps\footnote{\url{http://google.com/maps}} allows the user to
-navigate effectively in unknown cities. Others
-(e.g. Strava\footnote{Citation Needed}) track sports activities in
-order to optimize training plans or engage in virtual competitions .
+citizens using them. Apps like Google
+Maps\footnote{\url{http://www.google.com/mobile/maps/}} allows the user to
+navigate effectively in unknown cities; others track sports activities
+in order to optimize training plans or engage in virtual competitions.
 
-Besides the immediate benefits citizens using these services reveal
-a wealth of which can be used to draw a very detailed picture of the
-private life of the citizen. For instance, GPS location tracking can
-be used to infer shopping habits or associations with political groups
-(when meetings are attended). Accelerometer data can be used to detect
+Besides the immediate benefits these services reveal a wealth of
+private information about the citizen that capture a very detailed
+picture of his life. For instance, GPS location tracking can be used
+to infer shopping habits or associations with political groups (when
+meetings are attended) and accelerometer data can be used to detect
 medical conditions like walking disabilities. All this data is highly
@@
 the citizen's privacy. The service has to collect enough data to
-support the basic service-promises. If too much data is collected the
+support the basic service promises. If too much data is collected the
 privacy of the citizen is put at risk, with immediate implications for
@@
 
-%\TODO{MM:\\
-%- Shorten\\
-%- DONE - Remove Implicit Type Violations\\
-%- Update figure\\
-%- RG: Discuss alternative definitions of privacy\\
-%}
-
-\noindent
-Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes:
+\subsection{Privacy and Self Data Protection}
+More than one hundred years ago, 1890, the American politician Louis Brandis had specified privacy as the right to be left alone \cite{WarrenBrandeis:RightToPrivacy}.
+From the data processing point of view, this right is best expressed by the absence of information about a person in the mind of others.
+Indeed, this principle of “data minimization” is still fundamental to modern data protection legislation.
+The modern principles of data protection are codified by many legal systems in different countries, especially in Europe \cite{EU-directive2}, and by the US save harbor principles.
+However, modern data protection is more than only the absence of personal data.
+It is based on the personal right on self-determination over information and communication \cite{BVerfG:census}.
+This is, for example, very well expressed by Fried’s definition of 1970:
 \begin{quote}
-The notion of privacy remains out of the grasp of every academic chasing it.
-Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.
-\cite{Gutwirth}
+	Privacy is not simply an absence of information about us in the minds of others; rather it is the control we have over information about ourselves. \cite{CFried:Privacy}
 \end{quote}
-In this section we introduce the concept of privacy as control over information.
-We specify types of information we have to deal with in the context of privacy.
-And finally we outline the impact mobile sensors can have on one's privacy.
-
-\subsection{Privacy as Control over Information}
-%We may need to remove the following paragraph depinding on RG's alternative definition of privacy
-%Many definitions seem to focus on the vulnerabilities of privacy.
-%Thus a great amount of work is invested in finding threats to prohibit instead of describing why privacy is so valuable to us.
-%This way around one could derive measures to ensure and secure its value. \cite{7ToP}
-
-%One particular scholar, who did not follow this way of thinking is Charles Fried.
-One possible definition of privacy is given by Charles Fried in his work \textit{An Anatomy of Values}\cite{CFried:Privacy}.
-He questioned, why we are intuitively sensitive to privacy violations.
-But he did not assert privacy as an intrinsic value by itself, he rather stated:
-\begin{quote}
-\textbf{Privacy is} not simply an absence of information about us in the minds of others;
-rather it is \textbf{the control we have over information about ourselves}. \cite{CFried:Privacy}
-\end{quote}
-According to Fried privacy is our ability to create and modulate the relationships to other humans.
-Which makes us intuitively sensitive for its violation, because this affects what defines our identity.
-Also, the ability to build and maintain relationships is essential to society.
-This is what makes privacy valuable and deserving of protection. 
-\cite{sep-privacy}\cite{CFried:Privacy}
-
-\subsubsection{Relationship Creation.}
-
-% Privacy + Trust ~> Relationship
-% ===============================
-% - A shares intimate information with B
-% - A trusts B not to reveal that information by not constantly monitoring B, i.e. respecting B's privacy
-% - B has now the possibility to share intimate information with A
-
-We share information of great intimacy only with persons we deem trustworthy. 
-Moreover, we trust those persons to not reveal information about us to others by not constantly monitoring them, i.e. respecting their privacy.
-Trust needs the possibility of unknown failure.
-If we would constantly monitor our partners, they cannot fail unnoticed nor can they willingly share intimate information with us.
-By trusting them, we create the possibility for them to share intimate information with us.
-\cite{CFried:Privacy}
-
-\subsubsection{Relationship Modulation.}
-Fried argues, depending on the person we are interacting with, we are changing the degree of intimacy we share.
-If we are to share private information with others, we are most likely aware that this action is privacy related.
-In that case we are able to selectively disclose information along the two dimensions of quantity and quality.
-\cite{CFried:Privacy}
-
-\subsubsection{Alternative Definitions.}
-Using Fried's anatomy of privacy for this analysis is suitable because of two points.
-Despite the fact his text was published in 1970, he already included technologies to its viewpoint that did not only monitor location, but also record biometric data \cite{CFried:Privacy}. 
-Secondly, the concept of control over information does not enforce data minimization.
-Fried's definition complies with what we call \textit{self data protection}, where persons communicate their data self-determined.
-This is the opposite of \textit{system data protection}, where the collection of certain data is prohibited.
-
-\TODO{RG: Discuss alternative definitions of privacy}
-
-%=======================================================================================
-
-\begin{LGContent-Hidden}
-Defining privacy is a challenge which seems impossible. This is well put to words by Serge Gutwirth, who notes:
-
-\begin{quote}
-The notion of privacy remains out of the grasp of every academic chasing it.
-Even when it is cornered by such additional modifiers as `our' privacy, it still finds a way to remain elusive.
-\cite{Gutwirth}
-\end{quote}
-
-%Many researchers seem to only ``focus on the ways in which privacy can be infringed'' \cite{7ToP}.
-%Thus they try to create taxonomies of \emph{privacy harms} instead of taxonomies of \emph{privacy types}.
-%Those two differ in the respect that the former focuses on threats to prohibit whereas the latter focuses on values to protect.
-%So one should rather evaluate what aspects are precious about privacy and develop measures to ensure their security than only forbid single actions against it \cite{7ToP}.
-
-Many researchers seem to only ``focus on the ways in which privacy can be infringed'' \cite{7ToP}.
-Thus they invest a great amount of work in defining threats to prohibit instead of describing why privacy is so valuable to us.
-This way around one could derive measures to ensure and secure its value \cite{7ToP}.
-One particular scholar who does this in the context of digital monitoring is Charles Fried.
-He questioned, why we are intuitively sensitive to violations of privacy.
-But he did not assert privacy as an intrinsic value by itself, he rather sated:
-
-\begin{quote}
-Privacy is not simply an absence of information about us in the minds of others;
-rather it is \textbf{the control we have over information about ourselves}. \cite{CFried:Privacy}
-\end{quote}
-
-According to Fried privacy is a \textit{rational context}.
-Which means that one is aware of such context's existence during a rational action.
-If we are to share private information with others, we are most likely aware that this action is privacy related.
-In that case we are able to selectively disclose information along the two dimensions of quantity and quality.
-This leads Fried to his thesis, that privacy is one's ability to create and modulate his or hers social relationships, namely: friendship, love and trust. \cite{CFried:Privacy}
-
-Depending on the conversation partner, we change the degree of intimate information we share if it is a total stranger, colleague, close friend or lover.
-With close friends or lovers we share information of great intimacy we do not share with anyone else.
-Moreover, we trust those persons to not reveal information about us to others by respecting their privacy.
-Trust needs the possibility of unknown failure.
-If we would constantly monitor our partners, they cannot fail unnoticed nor can they willingly share that information with us.
-Thus they could not trust us anymore. \cite{CFried:Privacy}
-
-So privacy or the its possibility, according to Fried, is the foundation of our core relations: friendship, love and trust.
-And thus it is valuable, because those relations are essential to human society \cite{CFried:Privacy}.
-Using his anatomy of privacy as foundation of our analysis is suitable because of two points.
-At first, Fried's study on the understanding of privacy provided a great contribution to the research on the same term in philosophy \cite{sep-privacy} and computer science \cite{SociotechnicalArchitectureForOnlinePrivacy}.
-Secondly, despite the fact his text was published in 1970, he already included technologies to its viewpoint that did not only monitor location, but also record biometric data.
-\end{LGContent-Hidden}
+In order to realize this requirement in the modern IT world users are provided with user control functions, for example to decide about the access on their data for future use, and to view, modify, and delete unwanted data that had been collected.
+In its fundamental decision of May 13, 2014, the European Court has convicted Google to accept user demands to delete links to incorrect or irrelevant personal data in the Internet (C-131/12 – Google Spain SL).
+Such functions would constitute a so-called self-data-protection.
+It requires users to be highly aware of their rights and how to use them.
+Complementary, system-data-protection puts the load of data protection enforcement on the data collectors, ideally even without bothering the users.
+Typical system data protection functions are the deletion of personal data that are not bound to service purpose, the anonymization of personal data for research, and the abstinence from forwarding personal data to third parties.
+Together, system-data-protection and self-data-protection are supposed to provide a fair balance between the interests of service providers and their users.
+In our research work we focus on self-data-protection according to Fried’s user control requirement \cite{CFried:Privacy}.
+A system data protection point of view would lead to other results, namely to a set of obligations for service providers.
+This is subject to further research.
 
@@
 
-The goal of this chapter is to analyze and identify the threads to
-personal privacy that are posed by collecting, storing and processing
-sensor data from mobile phones.  We derive concrete privacy protection
-measures that address the main risks involved with handling such data.
+The goal of this chapter is to analyze and identify the threats to personal privacy that are posed by collecting, storing and processing sensor data from mobile phones.
+We derive concrete privacy protection measures that address the main risks involved with handling such data.
 
-In section \ref{subsec:world-analysis} and \ref{subsec:pot-analysis} we apply the first two steps of the Reference Model for IT Security Analysis as described in \cite{Grimm:ItSecRefModel}.
-It supersedes earlier efforts by
-e.g. \cite{Avizienis}.  The reference model consists of a \emph{model}
-and a \emph{procedure}: The model aims to organize common security terminology in a reasonable and practical way.
-The procedure describes a method to analysis the IT system based on that model.
-In these sections we give a brief overview over the reference model by applying it to our scenario.
+In our analysis we follow the “reference model for IT security analysis” as described in \cite{Grimm:ItSecRefModel}.
+It supersedes earlier efforts by e.g.
+\cite{Avizienis}.
+The reference model consists of a model and a procedure.
+The model organizes a common security terminology in a reasonable and practical way.
+The procedure describes a method for analyzing the IT system based on that model.
+The reference model provides four views: (1) the real world of persons and their assets, (2) the potential world of requirements and threats, (3) measures and plans specified by programs, business models and attack strategies, and, finally, (4) events of running programs, data accesses and performed attacks as well as their defense.
+In the following sections we apply the proposed procedures of the reference model to our scenario of a travel service that collects personal data from mobile users in order to serve the users and to enhance the service.
+The service intends to respect the privacy of its users.
 
-In section \ref{subsec:priv-rec} we provide specific privacy recommendations and requirements as a result of the previous analysis that the system must comply with.
+In sections \ref{subsec:world-analysis} and \ref{subsec:pot-analysis} we apply the first two steps of the reference model \cite{Grimm:ItSecRefModel}, which are related to the views on the real world and on the measures and plans.
+In section \ref{subsec:priv-rec} we apply the third step of the reference model by providing specific privacy recommendations and requirements as a result of the previous analysis that the system must comply with.
 
@@
 
-The first step is the \emph{world view} where all components are described in their current state. It consists of the following components:
-\begin{itemize}
+The first step is the \emph{world view} where all components are described in their current state.
+It consists of the following components: Assets, IT-Systems, Actors, Conflicts of Interests, Vulnerabilities, and Interactions.
 
-	\item \textbf{Assets.} 
-	Things of value to one or more stakeholders.
-
-	\item \textbf{IT-Systems.}
-	The relevant IT-Systems under study. Already described in Section \ref{subsubsection:it-systems}.
-
-	\item \textbf{Actors.}
-	All identifiable stakeholders of the system under study.
-
-	\item \textbf{Conflicts of Interests.}
-	Different actors have different interests which can be in conflict which each other.
-
-	\item \textbf{Vulnerabilities.}
-	All identifiable weaknesses in the IT-System e.g. a weak passwords.
-
-	\item \textbf{Interactions.}
-	This point captures all possible interactions between assets, IT-Systems, humans, and vulnerabilites as partly described in Section \ref{subsubsection:it-systems} and will be further analyzed in \ref{subsec:pot-analysis}.
-
-\end{itemize}
+The relevant \textbf{IT-Systems} were already described in Section \ref{subsubsection:it-systems}.
+\textbf{Interactions} between assets, IT-Systems, humans, and vulnerabilites as partly described in Section \ref{subsubsection:it-systems} and will be further analyzed in Section \ref{subsec:pot-analysis}.
 
@@
 
-In this scenario we focus our attention to only one asset: The privacy of the citizen.
-Our definition of privacy is described in detail in Section \ref{sec:privacydef} and introduced in Section \ref{sec:privacytypes} the seven different types of privacy.
+In this scenario we focus our attention to only one asset: The privacy of the citizen. Our definition of privacy is described in detail in Section \ref{sec:privacydef}.
 
@@
 
-\subsection{Step 2. Potential Analysis}
+\subsection{Step 2. Potential World Analysis}
 \label{subsec:pot-analysis}
 
-The \emph{Potential Analysis} (or \emph{Potential View}) displays the intended and unintended interactions of the components in the world view.
+The \emph{Potential World Analysis} displays the intended and unintended interactions of the components in the world view.
+
 The intended interactions support the underlying business objectives.
-Unintended interactions lead to threats. The potential view consists of the following components:
+Unintended interactions can lead to threats.
+Obviously this provides a conflict of interest with the victim's business model, to keep the asset away from unauthorized access.
+From the point of view of the victim, an attack would be an unintended interaction.
 
-\begin{itemize}
+The potential world view consists of the following components: Business Objectives, Threats, hances/Risk, and Security Requirements. 
+\textbf{Business Objectives} of the system owner were already described in Section \ref{subsubsection:it-systems} and \ref{subsec:world-analysis}.
+\textbf{Security Requirement}s are included to our analysis in Section \ref{subsec:priv-rec}.
 
-	\item \textbf{Business Objectives.}
-	Interaction of IT-system and actors that realize a business goals of the system owner were already described in Section \ref{subsubsection:it-systems} and \ref{subsec:world-analysis}.
-
-	\item \textbf{Threats.}
-	A threat is a potential interaction that destroys or harms assets of the system.
-	Concrete realizations of threats can be \emph{attacks} or \emph{accidents}.
-	Attacks are executed by an actor in response to a conflict of interest.
-	Accidents are harmful interactions that are not willfully caused by an actor.
-
-	\item \textbf{Chances/Risk.}
-	Evaluation of chances and risks associated to the business objectives and threats. The risk associated to a threat is its expected loss.
-
-	\item \textbf{Security Requirements.}
-	Interactions (e.g.
-threats) that shall not occur within the system in order to achieve its business objectives.
-We included these security requirements into our analysis in Section \ref{subsec:priv-rec}.
-
-\end{itemize}
-
-\subsubsection{Threat Specification}
+\subsubsection{Threats}
 

### 2015-02-02 paper.tex
further changings from RG changed from local authorities to travel agencies

@@
 \section{Introduction}
+\label{sec:intro}
 
@@
 However, the categories are very helpful for a privacy analysis with a focus on self-data-protection.
-The seven types of privacy are as follows:
-\begin{enumerate}
-	\item Privacy of the Person (with respect to body functions and body characteristics)
-	\item Privacy of Behavior and Action
-	\item Privacy of Communication
-	\item Privacy of Data and Image
-	\item Privacy of Thoughts and Feelings
-	\item Privacy of Location and Space 
-	\item Privacy of Association (with persons, e.g. friends, and organizations, e.g. political parties)
-\end{enumerate}
+The seven types of privacy are as follows: Privacy of the Person (with respect to body functions and body characteristics), Privacy of Behavior and Action, Privacy of Communication, Privacy of Data and Image, Privacy of Thoughts and Feelings, Privacy of Location and Space, Privacy of Association (with persons, e.g.
+friends, and organizations, e.g.
+political parties).
 
@@
 
-The relevant \textbf{IT-Systems} were already described in Section \ref{subsubsection:it-systems}.
-\textbf{Interactions} between assets, IT-Systems, humans, and vulnerabilites as partly described in Section \ref{subsubsection:it-systems} and will be further analyzed in Section \ref{subsec:pot-analysis}.
+The relevant \textbf{IT-Systems} were already described in Section \ref{sec:intro}.
+\textbf{Interactions} between assets, IT-Systems, humans, and vulnerabilites as partly described in Section \ref{sec:intro} and will be further analyzed in Section \ref{subsec:pot-analysis}.
 
@@
 For example they may have access to real time bus schedules or reports about traffic jams, that help them to avoid long waiting times.
-Also they generally benefit from improvements of public infrastructure by local authorities, which is triggered by issue reports.
+Also they generally benefit from improvements of public infrastructure by travel agencies, which is triggered by issue reports.
 
@@
 For example a department for urban mobility, orders a system to better understand usage patterns and make improvement to the urban traffic flow.
-Such systems are investments, and so naturally local authorities are interested in a profitable return, like increased ticket sales.
+Such systems are investments, and so naturally travel agencies are interested in a profitable return, like increased ticket sales.
 
@@
 or they can steal user data to prove the system insecure.
-Both would harm the reputation of local authorities, who endangered the privacy of the citizens.
+Both would harm the reputation of travel agencies, who endangered the privacy of the citizens.
 
@@
 The potential world view consists of the following components: Business Objectives, Threats, hances/Risk, and Security Requirements. 
-\textbf{Business Objectives} of the system owner were already described in Section \ref{subsubsection:it-systems} and \ref{subsec:world-analysis}.
+\textbf{Business Objectives} of the system owner were already described in Section \ref{sec:intro} and \ref{subsec:world-analysis}.
 \textbf{Security Requirement}s are included to our analysis in Section \ref{subsec:priv-rec}.
@@
 
-\subsubsection{Chances/Risk}
+\subsubsection{Chances/Risks}
 
@@
 
-\textbf{T1. Insufficient Control Features.}  The occurrence of this
+\textbf{Risk of T1. Insufficient Control Features.}  The occurrence of this
 threat is dependent on the design on the system and given in our case,
@@
 
-\textbf{T2. Excessive Data Mining.}
+\textbf{Risk of T2. Excessive Data Mining.}
 We assess the likeliness of excessive data mining to be $3$ High, since
@@
 
-\textbf{T3. Data Theft.}
+\textbf{Risk of T3. Data Theft.}
 The likeliness of a targeted attack by a third party is dependent on
@@
 
-\textbf{T4. Surveillance.}  In the surveillance scenario an party
+\textbf{Risk of T4. Surveillance.}  In the surveillance scenario an party
 related to the citizen, like a company where he is customer of, or a
@@
 
-\textbf{T5/6. Information Leak and Social Engineering.}
-
-In our scenario we assume that the culture and ethics inside the
-service provider company and local authority are very high, so that
+\textbf{Risk of T5/6. Information Leak and Social Engineering.} In our scenario we assume that the culture and ethics inside the
+service provider company and travel agencies are very high, so that
 the information leak scenario has a likeliness of $1$ (Low).
@@
 control over its data inside the system. The most direct way to do
-this is to provide a web-based \emph{Privacy Dashboard (R1)} which allows
+this is to provide a web-based \textbf{Privacy Dashboard} (R1) which allows
 the citizen to view, edit and delete all information about his person
-that is stored inside the system. Also control applied processing and
-disclosure of the data to third parties should be given to the user,
-at least in the form of an opt-out or veto option.
+that is stored inside the system. Also, control over processing and onward transfer of the data to third parties should be given to the user, at least in the form of an opt-out or veto option.
 
@@
 his data is information and comprehension of the intended data
-capturing and processing steps. Therefore a \emph{Privacy Policy (R2)} that
+capturing and processing steps. Therefore a \textbf{Privacy Policy} (R2) that
 is easily readable and contains all important information is
@@
 establish such a culture of privacy respect, we recommend to prepare
-an document called \emph{Data Handling Guidelines (R3)} intended for
+an document called \textbf{Data Handling Guidelines} (R3) intended for
 internal use that explains the concrete processing steps and purposes
@@
 additional permissions from the citizen. A simple email explaining the
-planned processing steps, and \emph{asking for permission (R4)} would be
+planned processing steps, and asking for \textbf{permission} (R4) would be
 enough for this purpose. The permission can be given via an embedded
@@
 An alternative measure to address the risk of excessive data mining is
-the \emph{anonymization (R5)} of data. When all direct- or indirect
+the \textbf{anonymization} (R5) of data. When all direct- or indirect
 links to the identity of the person are removed, no violation of the
@@
 The protection from threat scenario (T3) Data Theft is a case of
-classical \emph{IT infrastructure security (R6)}. The storage and processing infrastructure has
+classical \textbf{IT infrastructure security} (R6). The storage and processing infrastructure has
 to be secured using firewalls, up-to data software versions and proper
@@
 intermediaries or externals with access to the communication
-infrastructure. Strong \emph{encryption (R7)} should be used to make
+infrastructure. Strong \textbf{encryption} (R7) should be used to make
 it harder for externals to read the content of the transmitted data.
@@
 target the vulnerability of unhappy employees. Therefore a trustful,
-\emph{healthy company culture (R8)} should be maintained.
+\textbf{healthy company culture} (R8) should be maintained.
 
-In summary we recommend the following measures to secure the citizens
-privacy:
-\begin{enumerate}
-\item[R1] \textbf{Privacy Dashboard}. A tool which allows the citizen to view, edit
-  and delete all data personal data that is stored in the system.
-\item[R2] \textbf{Privacy Policy}. A document, that informs the citizen about the
-  collection and processing of personal information. It should at
-  least contain the legally required information.
-\item[R3] Issue \textbf{Data Handling Guidelines} that explain the permitted
-  processing methods and purposes.
-\item[R4] Ask the citizens for \textbf{permissions} before applying further
-  processing via Email.
-\item[R5] \textbf{Anonymize} personal data before processing.
-\item[R6] \textbf{Securing of Storage} and Processing infrastructure using e.g. firewalls.
-\item[R7] \textbf{Securing communication} channels using encryption.
-\item[R8] Maintain a healthy, trustful \textbf{relationship} with your employees.
-\end{enumerate}
 The mapping of these recommendations to the threats is summarized in Figure \ref{fig:risks}.
@@
 
-Our goal was to find a balance between the demand of the Local Authority for location-based data to improve their public services on one side, and to keep the privacy of citizens as untouched as possible on the other side.
-In a structured manner we identified all actors and what kind of conflict of interests can occour.
-Doing this, we found potential risks and provided finally specific recommendations how to protect the citizen’s privacy.
-One central idea of these recommendations is to give the citizen as much control over their data back as possible.
-This can be e.g.
-done via a Privacy Dashboard that gives a full overview about all personal data that is stored on the system.
-By implementing these privacy recommendations we found a way to satisfy the above-mentioned demand of both sides and comply with legal data protection requirements in addition.
+The travel agency has a demand for location-based personal data in order to improve its public services.
+Our goal was to find a balance between this demand of the travel agency and the privacy of the citizens.
+In a structured manner we identified all actors and their conflicts of interest.
+We found potential risks and derived recommendations how to protect the citizen's privacy.
+A central idea of these recommendations is to give the citizen as much control over their data back as possible.
+For example, this can be done by a privacy dashboard in the hand of every user that shows all his or her personal data stored on the system.
+The implementation of these privacy recommendations satisfies the above-mentioned demand of both sides in compliance with legal data protection requirements.
 

### 2015-02-05 paper.tex
abstract set added 'organized' section into introduction section 2 modified small error corrections

@@
 \begin{abstract}
-
-In our research work we have implemented a privacy aware mobile sensor application [diesen ersten Satz ggf. wieder streichen]. 
-In this paper we outline a scenario of a mobile traffic survey and analyse its privacy risks according to a reference model for IT security analysis.
-This leads to concrete recommendations for privacy protecting measures that are implemented in the system, for example a Web-based dashboard which allows the users to view, modify and delete all personal data that is stored in the system.
-
-Vorschlag HH: In this article we analyze the privacy aspects of a mobile sensor
+In this article we analyze the privacy aspects of a mobile sensor
 application used for recording urban travel patterns as part of a
@@
 In this article we study the privacy risks of a travel-survey service
-for urban mobility.
+for urban mobility that is further specified in section \ref{sec:ScenarioDescription}.
+In section \ref{sec:privacydef} below, we will explain our view on privacy.
+In order to find the right user control functions we will perform a systematic privacy analysis of our sensor data storage and mining infrastructure following the reference model of IT security analysis \cite{Grimm:ItSecRefModel} in section \ref{sec:privacyanalysis} below.
 
@@
 
-Travel surveys are regularly conducted travel agencies which provide
+Travel surveys are regularly conducted by transport authorities which provide
 public transportation infrastructure. With these surveys the current
@@
 
+The travel services need the personal mobile data for two reasons: firstly, they want to serve their customers individually, for example by guiding them to proximity stations, to recommend adequate travel connections, to inform about delays, or to sell electronic tickets.
+For this purpose, the personal data for evaluation are related to the individual customers.
+Secondly, the travel services want to enhance their system, in that they evaluate the demand of stations, connections, changes, ways and time spent by their customers.
+For this purpose, personal data are anonymized before evaluation.
+
 Within the Live+Gov EU project, we have established a first
@@
 There are three stakeholders involved in the travel-survey service:
-The \emph{citizen}, the \emph{service provider} and the \emph{travel
-  agency}. The service provider is the operator of all IT systems and
-offers the travel-survey service to the travel agency. Figure
+The \emph{citizen}, the \emph{service provider} and the \emph{transport
+  authority}. The service provider is the operator of all IT systems and
+offers the travel-survey service to the transport authority. Figure
 \ref{figure:Live+Gov Vulnerabilities} contains a schematic
@@
 
-The citizen is a volunteer who is willing to share his personal travel
+The \emph{citizen} is a volunteer who is willing to share his personal travel
 patterns within the travel-survey event.  He carries a mobile
@@
 information about travel patterns of citizens within the
-travel-survey. The reporting tool is used by the travel agency to gain
+travel-survey. The reporting tool is used by the transport authority to gain
 insights for planning and accounting of the public transportation
@@
 \subsection{Privacy and Self Data Protection}
-More than one hundred years ago, 1890, the American politician Louis Brandis had specified privacy as the right to be left alone \cite{WarrenBrandeis:RightToPrivacy}.
+1890, the American politician Louis Brandis had specified privacy as the right to be left alone \cite{WarrenBrandeis:RightToPrivacy}.
 From the data processing point of view, this right is best expressed by the absence of information about a person in the mind of others.
@@
 However, the categories are very helpful for a privacy analysis with a focus on self-data-protection.
-The seven types of privacy are as follows: Privacy of the Person (with respect to body functions and body characteristics), Privacy of Behavior and Action, Privacy of Communication, Privacy of Data and Image, Privacy of Thoughts and Feelings, Privacy of Location and Space, Privacy of Association (with persons, e.g.
+The seven types of privacy are as follows: Privacy of the Person (with respect to body functions), Privacy of Behavior and Action, Privacy of Communication, Privacy of Data and Image, Privacy of Thoughts and Feelings, Privacy of Location and Space, Privacy of Association (with persons, e.g.
 friends, and organizations, e.g.
@@
 The both sensors give the exact cell or network, the mobile device has registered with at the current moment.
-Frequent connection to one particular network also reveals association, e.g. university networks.
-
-The Bluetooth sensors record lists of the bluetooth clients in the direct neighbourhood.
+Frequent connection to one particular network also reveals association, e.g. university networks. The Bluetooth sensors record lists of the bluetooth clients in the direct neighbourhood.
 Since those clients are usually moving, inference of the position is usually not possible.
@@
 \section{Privacy Analysis}
+\label{sec:privacyanalysis}
 
@@
 
-This section describes the human actors previously introduced within the IT system architecture (\ref{subsubsection:it-systems}).
+This section describes the human actors previously introduced in section \ref{sec:ScenarioDescription}.
 Although the text is only a short description, a list of the most important interests is provided for each actor.
@@
 Also they generally benefit from improvements of public infrastructure by travel agencies, which is triggered by issue reports.
-
 By using the application, citizens are sharing personal information like name and address, as well as data gathered from mobile sensor with the service provider.
@@
 \cite{GuardienMassSurveillance}).
-
 Citizens are interested in: physical wellbeing and health, financial profit, convenience, legitimate use of personal data, non-disclosure of personal data to peers of the citizen, and not being monitored.
@@
 Externals are persons who do not have privileged access to the IT systems, and are willing to break laws, security constrains and norms in order to promote their interests.
-
 A common interest of an external is financial profit.
@@
 spam or phishing attacks - or excessive data mining.
-
 In short, externals could be interested in: increase power over citizen, financial profit, political activism, and their social standing.
@@
 They are private companies and legal persons in their own right, but also employ a number of people with diverging interests including administrators, programmer/developer, and a support manager.
-
 As companies, they are interested in gaining financial profit.
@@
 The complexity of a task has to be in reasonable bounds, so that system providers can complete it within time, with a satisfying quality.
-
 System Providers are interested in: financial profit, manageable complexity, professional excellence, and good working conditions.
 
-\textbf{Local Authority.}
-Local authorities are public offices (ministry, agency, department, ...) or other external public entities which act as direct customers of service providers.
+\textbf{Transport Authority.}
+ Transport authorities are public offices (ministry, agency, department, ...) or other external public entities which act as direct customers of service providers.
 For example a department for urban mobility, orders a system to better understand usage patterns and make improvement to the urban traffic flow.
 Such systems are investments, and so naturally travel agencies are interested in a profitable return, like increased ticket sales.
-
-Local authorities are interested in: financial profit, political reputation, business intelligence, and good working conditions.
+Transport authorities are interested in: financial profit, political reputation, business intelligence, and good working conditions.
 
@@
 \label{subsubsection:Conflicts of Interests}
-Different actors have different interests which can be in conflict each other. This section outlines the Conflicts of Interests between the actors of the proposed IT system architecture.
-
+Different actors have different interests which can be in conflict each other.
+This section outlines the Conflicts of Interests between the actors of the proposed IT system architecture.
 The individual interests of all actors is already described in the previous section and are not elaborated any further.
@@
 \textbf{System Complexity vs Privacy.}
-System Providers offer a service to Local Authorities, so that Local Authorities can improve their public services.
+System Providers offer a service to Transport Authorities, so that they are able to improve their public services.
 This task in it self has a high technical complexity and operates on privacy sensitive data provided by monitored Citizens.
@@
 \textbf{Business Intelligence vs Privacy.}
-Local authorities order a monitoring and mining system from system providers as much as possible, which allows them to produce business intelligence for public services.
+Transport authorities order a monitoring and mining system from system providers as much as possible, which allows them to produce business intelligence for public services.
 They are interested in the successful usage of their data, although the interest of citizens lies in maintaining control over their data and protecting their rights to privacy.
@@
 
-\textbf{Political Activism vs Reputation of Local Authority.}
+\textbf{Political Activism vs Reputation of Transport Authority.}
 Besides monetary reasons, externals can be motivated by political reasons to attack the monitoring and mining system.
@@
 
-\caption{Overview of vulnerabilities of the Live+Gov System; legend see \ref{figure:IT Systems} }
+\caption{Overview of vulnerabilities of the Live+Gov System }
 \label{figure:Live+Gov Vulnerabilities}
@@
 System Provider staff has access to Data Center hardware and software like databases, web-servers and other inspection tools.
-Local Authority staff has access to the Report Tool.
+Transport authority staff has access to the Report Tool.
 This all enables staff members to have potential access to privacy sensitive information.
@@
 The \emph{Potential World Analysis} displays the intended and unintended interactions of the components in the world view.
-
 The intended interactions support the underlying business objectives.
@@
 From the point of view of the victim, an attack would be an unintended interaction.
-
 The potential world view consists of the following components: Business Objectives, Threats, hances/Risk, and Security Requirements.
 \textbf{Business Objectives} of the system owner were already described in Section \ref{sec:intro} and \ref{subsec:world-analysis}.
-\textbf{Security Requirement}s are included to our analysis in Section \ref{subsec:priv-rec}.
+\textbf{Security Requirements} are included to our analysis in Section \ref{subsec:priv-rec}.
 
@@
 We restrict ourselves to the case of attacks, and the asset of privacy.
-An attack is an interaction that is executed by an actor in response to a conflict of interest by exploiting a vulnerability of the system. We describe threats for the citizen privacy in the system.
+An attack is an interaction that is executed by an actor in response to a conflict of interest by exploiting a vulnerability of the system.
+We describe threats for the citizen privacy in the system.
 This list can necessarily not be complete, but we make a best effort to cover the most relevant cases.
@@
 \textbf{T2. Excessive Data Mining.}
-The System Provider and/or the Local Authority secretly extract more private information from the collected data, than the Citizen agreed to.
+The System Provider and/or the transport authority secretly extract more private information from the collected data, than the Citizen agreed to.
 But results of the mining process create no disadvantages for Citizens, because there is no disclosure to third parties.
 This could be the case for a System Provider, who wants to test a new product and uses the pre-existing data collection.
-
-Thus there are two possible conflicts: \textit{Privacy vs. Financial Profit of Serivce Provider} and  \textit{Privacy vs. Buisness Intelligence of Local Authority}
-The threat can be provoked by either lax data handling policies of both System Providers and Local Authorities, or a weak law enforcement of existing supervision.
+Thus there are two possible conflicts: \textit{Privacy vs.
+Financial Profit of Serivce Provider} and  \textit{Privacy vs.
+Buisness Intelligence of Transport Authority} The threat can be provoked by either lax data handling policies of both System Providers and Transport Authorities, or a weak law enforcement of existing supervision.
 But the main issue, which can lead to such threats, is again a general \textit{Missing Privacy Awareness}.
@@
 An External infiltrates infrastructure in order to steal personal data and sell it on the black market.
-Also the External might be motivated politically and wants to harm the reputation of the System Provider or the Local Authority.
-Such a successful attack could harm the reputation of both System Provider and Local Authority.
+Also the External might be motivated politically and wants to harm the reputation of the System Provider or the Transport Authority.
+Such a successful attack could harm the reputation of both System Provider and Transport Authority.
 Thus this threat is defined by three conflicts: \textit{Privacy vs.
 Financial Profit}, \textit{Reputation of Service Provider vs.
-Political Activism} and \textit{Reputation of Local Authority vs.
+Political Activism} and \textit{Reputation of Transport Authority vs.
 Political Activism}.
@@
 \textbf{T5. Information Leak} 
-
 Like an external person the Data Theft Scenario an employee of the
-service provider or the Local Authority has selfish interests to gain
+service provider or the transport authority has selfish interests to gain
 money, make political statements or harm his employer.  In order to
@@
 \textit{Privacy vs. Financial Profit} of the Employee, \textit{Reputation of Service
-  Provider vs. Political Activism} of the Employee and \textit{Reputation of Local
+  Provider vs. Political Activism} of the Employee and \textit{Reputation of Transport
   Authority vs. Political Activism} of the Employee.  The vulnerability constitutes of
@@
 This scenario an external manipulates an employee of a Service
-Provider or the Local Authority to leak information to the external
+Provider or the Transport Authority to leak information to the external
 person.  It is thus combination of the Data Theft and Information Leak
@@
 the External, \textit{Reputation of Service Provider vs. Political
-  Activism} of the external and \textit{Reputation of Local Authority
+  Activism} of the external and \textit{Reputation of Transport Authority
   vs. Political Activism} of the external.  The exploited
@@
 likeliness scales have a logarithmic character, so that addition
-of those scales corresponds to multiplication of the usual scales.
-
-The likeliness, loss and the resulting risks assigned to the threats
+of those scales corresponds to multiplication of the usual scales. The likeliness, loss and the resulting risks assigned to the threats
 are discussed in the following paragraphs and summarized in Figure
@@
 associated, risk is $1$ Low on our scale, since no direct harm is done to
-the citizen by exploiting the data.
-
-Hence the resulting risk is calculated as $4+1 = 5$.
+the citizen by exploiting the data. Hence the resulting risk is calculated as $4+1 = 5$.
 
@@
 provider, without somebody else noticing, and the service provider himself
-has an interest in this activity. The associated loss, on the other hand can be substantial (Medium $2$).
-
-Hence the resulting risk is calculated as $3+2 = 5$.
+has an interest in this activity. The associated loss, on the other hand can be substantial (Medium $2$). Hence the resulting risk is calculated as $3+2 = 5$.
 
@@
 addresses, no payment information). Therefore the likeliness
-assessment is $1-2$ (Low-Medium).
-
-The harm of leaked information to a criminal party is $3$ High.
-Hence the resulting risk is calculated as $4-5$.
+assessment is $1-2$ (Low-Medium). The harm of leaked information to a criminal party is $3$ High. Hence the resulting risk is calculated as $4-5$.
 
@@
 government agency, seeks to obtain sensitive information from our
-service.
-
-The likeliness of such an intrusion is hard to asses, and depends
+service. The likeliness of such an intrusion is hard to asses, and depends
 again on the popularity of the service. If a high popularity is
@@
 however very hard to perform. Therefore we assess the likeliness of
-the threat with $1-2$ (Low-Medium).
-
-The harm of leaked information to a related party is $3$ High.
-Hence the resulting risk is calculated as $4-5$.
+the threat with $1-2$ (Low-Medium). The harm of leaked information to a related party is $3$ High. Hence the resulting risk is calculated as $4-5$.
 
@@
 service provider company and travel agencies are very high, so that
-the information leak scenario has a likeliness of $1$ (Low).
-
-The harm of such an information leaked is $3$ High, so that the
+the information leak scenario has a likeliness of $1$ (Low). The harm of such an information leaked is $3$ High, so that the
 resulting risk is calculated as $4$.
@@
 
-In the preceding section we have identified the main risks for the
-users privacy.  In this section we derive recommendations or
-requirements for a system that addresses these risks..
+In this section we derive recommendations or
+requirements for a system that addresses these risks. The mapping of these recommendations to the threats is summarized in Figure \ref{fig:risks}.
 
-In order to address the threat with the highers risk, Insufficient
-Control (T1) of the citizen, we need to give the citizen back the
+\textbf{Privacy Dashboard (R1)}. In order to address the threat with the highers risk, Insufficient Control (T1) of the citizen, we need to give the citizen back the
 control over its data inside the system. The most direct way to do
-this is to provide a web-based \textbf{Privacy Dashboard} (R1) which allows
-the citizen to view, edit and delete all information about his person
-that is stored inside the system. Also, control over processing and onward transfer of the data to third parties should be given to the user, at least in the form of an opt-out or veto option.
+this is to provide a web-based overview which allows the citizen to view, edit and delete all information about his person
+that is stored inside the system.
 
-A necessary pre-requirement for effective control of the citizen over
-his data is information and comprehension of the intended data
-capturing and processing steps. Therefore a \textbf{Privacy Policy} (R2) that
-is easily readable and contains all important information is
-essential and a legal requirement.
+\textbf{Privacy Policy (R2)}.
+A necessary pre-requirement for effective control of the citizen over his data is information and comprehension of the intended data capturing and processing steps.
+Therefore a policy that
+is easily readable and contains all important information is essential and a legal requirement.
 
-The threat with the second largest risk is (T2) Excessive Data Mining.
+\textbf{Data Handling Guidelines (R3)} The threat with the second largest risk is (T2) Excessive Data Mining.
 Contrary to common belief, it is neither legal nor ethical to process
@@
 establish such a culture of privacy respect, we recommend to prepare
-an document called \textbf{Data Handling Guidelines} (R3) intended for
-internal use that explains the concrete processing steps and purposes
-that are permitted by the citizens. In particular the following information should be provided for each processing task: The name of the controller, purpose of processing,
-description of the data categories, recipients of the data if
-disclosed, transfer to third countries and a description of security
-of processing.
+an guideline intended for internal use that explains the concrete processing steps and purposes that are permitted by the citizens.
 
-If further processing should be performed, it is necessary to seek
-additional permissions from the citizen. A simple email explaining the
-planned processing steps, and asking for \textbf{permission} (R4) would be
-enough for this purpose. The permission can be given via an embedded
-link that shall be followed in order to signal agreement.
+\textbf{User Permissions (R4)}.
+If further processing should be performed, it is necessary to seek additional permissions from the citizen.
+A simple email explaining the planned processing steps, and asking for permissions would be enough for this purpose.
+The permission can be given via an embedded link that shall be followed in order to signal agreement.
 
-An alternative measure to address the risk of excessive data mining is
-the \textbf{anonymization} (R5) of data. When all direct- or indirect
+\textbf{Anonymization (R5)} An alternative measure to address the risk of excessive data mining is the anonymization of data. When all direct- or indirect
 links to the identity of the person are removed, no violation of the
@@
 
-The protection from threat scenario (T3) Data Theft is a case of
-classical \textbf{IT infrastructure security} (R6). The storage and processing infrastructure has
+\textbf{IT Infrastructure Security (R6)}. The protection from threat scenario (T3) Data Theft is a case of
+classical IT infrastructure security. The storage and processing infrastructure has
 to be secured using firewalls, up-to data software versions and proper
@@
 
-The protection from threat scenario (T4) Surveillance focuses on the
+\textbf{Encryption (R7)} The protection from threat scenario (T4) Surveillance focuses on the
 communication channels. They are target of wiretapping attacks by
 intermediaries or externals with access to the communication
-infrastructure. Strong \textbf{encryption} (R7) should be used to make
+infrastructure. Strong encryption should be used to make
 it harder for externals to read the content of the transmitted data.
 
-Threat scenarios (T5) Information Leak and (T6) Social Engineering
+\textbf{Relationship (R8)}. Threat scenarios (T5) Information Leak and (T6) Social Engineering
 target the vulnerability of unhappy employees. Therefore a trustful,
-\textbf{healthy company culture} (R8) should be maintained.
+healthy company culture should be maintained.
 
-The mapping of these recommendations to the threats is summarized in Figure \ref{fig:risks}.
 %\end{LGContent}
@@
 
-The travel agency has a demand for location-based personal data in order to improve its public services.
-Our goal was to find a balance between this demand of the travel agency and the privacy of the citizens.
+The transport authority has a demand for location-based personal data in order to improve its public services.
+Our goal was to find a balance between this demand of the transport authority and the privacy of the citizens.
 In a structured manner we identified all actors and their conflicts of interest.

### 2015-02-05 paper.tex
privacy recommendation changed

@@
 
-\textbf{Privacy Dashboard (R1)}. In order to address the threat with the highers risk, Insufficient Control (T1) of the citizen, we need to give the citizen back the
+\textbf{R1. Privacy Dashboard}. In order to address the threat with the highers risk, Insufficient Control (T1) of the citizen, we need to give the citizen back the
 control over its data inside the system. The most direct way to do
@@
 
-\textbf{Privacy Policy (R2)}.
+\textbf{R2.
+Privacy Policy}.
 A necessary pre-requirement for effective control of the citizen over his data is information and comprehension of the intended data capturing and processing steps.
@@
 
-\textbf{Data Handling Guidelines (R3)} The threat with the second largest risk is (T2) Excessive Data Mining.
+\textbf{R3. Data Handling Guidelines}. The threat with the second largest risk is (T2) Excessive Data Mining.
 Contrary to common belief, it is neither legal nor ethical to process
@@
 
-\textbf{User Permissions (R4)}.
+\textbf{R4.
+User Permissions}.
 If further processing should be performed, it is necessary to seek additional permissions from the citizen.
@@
 
-\textbf{Anonymization (R5)} An alternative measure to address the risk of excessive data mining is the anonymization of data. When all direct- or indirect
+\textbf{R5. Anonymization}. An alternative measure to address the risk of excessive data mining is the anonymization of data. When all direct- or indirect
 links to the identity of the person are removed, no violation of the
@@
 
-\textbf{IT Infrastructure Security (R6)}. The protection from threat scenario (T3) Data Theft is a case of
+\textbf{R6. IT Infrastructure Security}. The protection from threat scenario (T3) Data Theft is a case of
 classical IT infrastructure security. The storage and processing infrastructure has
@@
 
-\textbf{Encryption (R7)} The protection from threat scenario (T4) Surveillance focuses on the
+\textbf{R7. Encryption}. The protection from threat scenario (T4) Surveillance focuses on the
 communication channels. They are target of wiretapping attacks by
@@
 
-\textbf{Relationship (R8)}. Threat scenarios (T5) Information Leak and (T6) Social Engineering
+\textbf{R8. Relationship}. Threat scenarios (T5) Information Leak and (T6) Social Engineering
 target the vulnerability of unhappy employees. Therefore a trustful,

### 2015-02-17 paper.tex
further changings from RG

@@
 
-\usepackage{amsmath}
+%\usepackage{amsmath}
 
@@
 travel-survey service. This service has been developed and
-field-tested within the Live+Gov EU Project. The privacy analysis
+field-tested within the Live+Gov EU Project \cite{LiveGov:Prog}. The privacy analysis
 follows a structured approach established in
@@
 The rise of mobile smart-phones equipped with a wide range of sensors
-and the abundance mobile internet connectivity has paved the way to a
+and the abundance mobile Internet connectivity has paved the way to a
 whole new generation of mobile services which ease the daily life of
@@
 medical conditions like walking disabilities. All this data is highly
-sensitive to the citizens privacy and can be used against the citizen
+sensitive to the citizen's privacy and can be used against the citizen 
 if it falls into the wrong hands.
@@
 privacy of the citizen is put at risk, with immediate implications for
-the customer trust relation and the acceptance of the service. Also
+the customer trust relation and the acceptance of the service. Also,
 leaks of private data can jeopardize the whole business.
@@
 Travel surveys are regularly conducted by transport authorities which provide
-public transportation infrastructure. With these surveys the current
+a public transportation infrastructure. With these surveys the current
 usage of the public transportation system is assessed and the insights
-are used for future planning of bus, tram and subway lines, etc..  For
+are used for future planning of bus, tram and subway lines, etc.  For
 each survey a large number of citizens (e.g. 5000) is asked to keep
@@
 
-Within the Live+Gov EU project, we have established a first
+Within the Live+Gov EU project \cite{LiveGov:Prog}, we have established a first
 prototypical implementation of such an automated travel-survey
@@
 offers the travel-survey service to the transport authority. Figure
-\ref{figure:Live+Gov Vulnerabilities} contains a schematic
+\ref{figure:Live+Gov Vulnerabilities} in section \ref{subsubsection:Conflicts of Interests} below contains a schematic
 visualization of the relation between these stakeholders.
@@
 
-The \emph{data center} stores and processes sensor data collected with
+The \emph{data center} is under the control of the service provider. It 
+stores and processes sensor data collected with
 the sensor collector application. It can also take into account data
@@
 transportation and if so which lines of service (bus, train, tram) are
-used. The data center can also sent mining end products and messages
+used. The data center can also send mining end products and messages
 (e.g. traffic jam reports, bus schedule) back to the mobile device of
@@
 information about travel patterns of citizens within the
-travel-survey. The reporting tool is used by the transport authority to gain
+travel-survey. The reporting tool is used by the \emph{transport authority} to gain
 insights for planning and accounting of the public transportation
@@
 Indirect impact is caused by combining sensor data with additional knowledge.
-For instance, comparing a contemporary map with locational data can imply behaviour if the position matches a church.
+For instance, comparing a contemporary map with locational data can reveal behaviour and thoughts if the position matches a church.
 
@@
 If the mobile device is carried ``normally'' its safe to say that those sensors also measure the moments of its carrier.
-So his privacy is infringed regarding biometric behaviour, i.e. the Privacy of the Person.
+So privacy is infringed regarding biometric behaviour, i.e. the Privacy of the Person.
 
@@
 \subsubsection{Network Sensors.}
-The GSM and WLAN sensors reveal the position of the mobile device and its carrier, when used in connection with external databases.
-The both sensors give the exact cell or network, the mobile device has registered with at the current moment.
-Frequent connection to one particular network also reveals association, e.g. university networks. The Bluetooth sensors record lists of the bluetooth clients in the direct neighbourhood.
+The GSM and WLAN sensors reveal the position of the mobile device and its carrier, when used in connection with external databases (location).
+Both sensors give the exact cell or network, the mobile device has registered with at the current moment.
+Frequent connection to one particular network also reveals Association, e.g. university networks. The Bluetooth sensors record lists of the bluetooth clients in the direct neighbourhood.
 Since those clients are usually moving, inference of the position is usually not possible.
@@
 
-In sections \ref{subsec:world-analysis} and \ref{subsec:pot-analysis} we apply the first two steps of the reference model \cite{Grimm:ItSecRefModel}, which are related to the views on the real world and on the measures and plans.
-In section \ref{subsec:priv-rec} we apply the third step of the reference model by providing specific privacy recommendations and requirements as a result of the previous analysis that the system must comply with.
+In sections \ref{subsec:world-analysis} and \ref{subsec:pot-analysis} we apply the first two steps of the reference model \cite{Grimm:ItSecRefModel}, which are related to the views on the real world and its privacy risks.
+In section \ref{subsec:priv-rec} we apply the third step of the reference model by providing specific recommendations and for privacy measures as a result of the previous analysis that the system must comply with.
 
@@
 
-The relevant \textbf{IT-Systems} were already described in Section \ref{sec:intro}.
-\textbf{Interactions} between assets, IT-Systems, humans, and vulnerabilites as partly described in Section \ref{sec:intro} and will be further analyzed in Section \ref{subsec:pot-analysis}.
+The relevant \textbf{IT-Systems} were already described in section \ref{sec:intro}.
+\textbf{Interactions} between assets, IT-Systems, humans, and vulnerabilites is partly described in section \ref{sec:intro} and will be further analyzed in section \ref{subsec:pot-analysis}.
 
@@
 
-In this scenario we focus our attention to only one asset: The privacy of the citizen. Our definition of privacy is described in detail in Section \ref{sec:privacydef}.
+In this scenario we focus our attention to only one asset: The privacy of the citizen. Our definition of privacy is described in detail in section \ref{sec:privacydef}.
 
@@
 
-This section describes the human actors previously introduced in section \ref{sec:ScenarioDescription}.
-Although the text is only a short description, a list of the most important interests is provided for each actor.
+This section describes the human actors previously introduced in section \ref{sec:ScenarioDescription}. A short description for each actor includes a list of the most important interests.
 
@@
 Their main motivation for using the software is to gain a higher level of convenience in their daily activities.
-For example they may have access to real time bus schedules or reports about traffic jams, that help them to avoid long waiting times.
-Also they generally benefit from improvements of public infrastructure by travel agencies, which is triggered by issue reports.
-By using the application, citizens are sharing personal information like name and address, as well as data gathered from mobile sensor with the service provider.
+For example they may have access to real time bus schedules or reports about traffic jams that help them to avoid long waiting times.
+Also they generally benefit from improvements of public infrastructure by transport authorities, which is triggered by issue reports.
+By using the application, citizens are sharing personal information like name and address, as well as data gathered from mobile sensors with the service provider.
 This data can be exploited in ways that are harmful to the citizen (cf.
@@
 
-\textbf{External.}
-Externals are persons who do not have privileged access to the IT systems, and are willing to break laws, security constrains and norms in order to promote their interests.
+\textbf{Malicious External.}
+Malicious externals are persons who do not have privileged access to the IT systems, and are willing to break laws, security constrains and norms in order to promote their interests.
 A common interest of an external is financial profit.
@@
 The complexity of a task has to be in reasonable bounds, so that system providers can complete it within time, with a satisfying quality.
-System Providers are interested in: financial profit, manageable complexity, professional excellence, and good working conditions.
+System Providers are interested in: financial profit, manageable complexity, professional excellence, and good working conditions, and a good reputation.
 
@@
  Transport authorities are public offices (ministry, agency, department, ...) or other external public entities which act as direct customers of service providers.
-For example a department for urban mobility, orders a system to better understand usage patterns and make improvement to the urban traffic flow.
-Such systems are investments, and so naturally travel agencies are interested in a profitable return, like increased ticket sales.
+For example a department for urban mobility orders a system to better understand usage patterns and make improvement to the urban traffic flow.
+Such systems are investments, and so naturally transport authorities are interested in a profitable return, like increased ticket sales.
 Transport authorities are interested in: financial profit, political reputation, business intelligence, and good working conditions.
@@
 \label{subsubsection:Conflicts of Interests}
-Different actors have different interests which can be in conflict each other.
+Different actors have different interests which can be in conflict with each other.
 This section outlines the Conflicts of Interests between the actors of the proposed IT system architecture.
-The individual interests of all actors is already described in the previous section and are not elaborated any further.
-The emphasis here is put on prominent existing conflicts, because they provide a foundation for vulnerabilities and subsequent threats.
+The emphasis here is to put on prominent existing conflicts, because they provide a foundation for vulnerabilities and subsequent threats.
 
@@
 \textbf{Power of External vs. Privacy.}
-Externals which are in a social relation to the citizen can have an interest in obtaining further information in order to gain power.
-In the most simplistic example this could be a man wanting monitor the activities of his spouse.
+Malicious externals which are in a social relation to the citizen can have an interest in obtaining further information in order to gain power.
+In the most simplistic example this could be a man wanting to monitor the activities of his spouse.
 
 \textbf{Financial Profit of External vs Privacy.}
-Externals can gain financial profit from stealing privacy sensitive data.
+Malicious Externals can gain financial profit from stealing privacy sensitive data.
 For example by selling raw contact information to advertisers or by selling mined data to insurance companies, or intermediaries like scoring companies.
@@
 \label{subsubsection:Vulnerabilities}
-This section outlines the vulnerabilities (Figure \ref{figure:Live+Gov Vulnerabilities}) of the proposed monitoring and mining system.
+This section outlines the vulnerabilities of the proposed monitoring and mining system. This is illustrated by figure \ref{figure:Live+Gov Vulnerabilities} below.
 
@@
 \textbf{Unhappy Employees.}
-An Employee that is frustrated with his situation for a long time
-period constitutes a security vulnerability. On the one hand he might
+As an insider, an Employee who is frustrated with his situation constitutes a security vulnerability. On the one hand he might
 want to harm his employer directly, on the other he is increasingly
@@
 Transport authority staff has access to the Report Tool.
-This all enables staff members to have potential access to privacy sensitive information.
+Moreover, access control systems are security bottlenecks,
 
@@
 From the point of view of the victim, an attack would be an unintended interaction.
-The potential world view consists of the following components: Business Objectives, Threats, hances/Risk, and Security Requirements.
-\textbf{Business Objectives} of the system owner were already described in Section \ref{sec:intro} and \ref{subsec:world-analysis}.
-\textbf{Security Requirements} are included to our analysis in Section \ref{subsec:priv-rec}.
+The potential world view consists of the following components: Business Objectives, Threats, Chances/Risks, and Security Requirements.
+\textbf{Business Objectives} of the system owner were already described in section \ref{sec:intro} and \ref{subsec:world-analysis}.
+\textbf{Security Requirements} are included to our analysis in section \ref{subsec:priv-rec}.
 
@@
 An attack is an interaction that is executed by an actor in response to a conflict of interest by exploiting a vulnerability of the system.
-We describe threats for the citizen privacy in the system.
-This list can necessarily not be complete, but we make a best effort to cover the most relevant cases.
+We describe threats for the citizen's privacy in the system.
+As a matter of course, this list cannot be complete, but we make a best effort to cover the most relevant cases.
 
@@
 Financial Profit of Serivce Provider} and  \textit{Privacy vs.
-Buisness Intelligence of Transport Authority} The threat can be provoked by either lax data handling policies of both System Providers and Transport Authorities, or a weak law enforcement of existing supervision.
+Buisness Intelligence of Transport Authority}.
+The threat can be provoked by either careless data handling policies of both System Providers and Transport Authorities, or a weak law enforcement of existing supervision.
 But the main issue, which can lead to such threats, is again a general \textit{Missing Privacy Awareness}.
@@
 \textbf{T3. Data Theft}
-An External infiltrates infrastructure in order to steal personal data and sell it on the black market.
+An malicious external infiltrates the infrastructure in order to steal personal data and sell it on the black market.
 Also the External might be motivated politically and wants to harm the reputation of the System Provider or the Transport Authority.
 Such a successful attack could harm the reputation of both System Provider and Transport Authority.
-Thus this threat is defined by three conflicts: \textit{Privacy vs.
+This threat is defined by three conflicts: \textit{Privacy vs.
 Financial Profit}, \textit{Reputation of Service Provider vs.
@@
 An External infiltrates infrastructure in order to obtain information
-about the citizen and exploit it directly.  In this scenario the
+about citizens and exploit it directly.  In this scenario the
 external is supposed to have some direct relationship to the citizen
@@
 In this threat the privacy interest of the citizen is in conflict with
-the aspirations for power over the citizen by the externals.
+the aspirations for power over citizens by the externals.
 
 \textbf{T5. Information Leak} 
-Like an external person the Data Theft Scenario an employee of the
+Like a malicious external person in the Data Theft scenario an employee of the
 service provider or the transport authority has selfish interests to gain
@@
   Authority vs. Political Activism} of the Employee.  The vulnerability constitutes of
-the existence of \textit{Unhappy employees} itself and possibly \textit{lax access
-rules} that enable the employee to obtain large amounts of data
+the existence of \textit{Unhappy employees} and possibly \textit{careless access
+rules}, which enable the employee to obtain large amounts of data
 unnoticed.
@@
 \textbf{T6. Social Engineering}
-This scenario an external manipulates an employee of a Service
+In this scenario an external manipulates an employee of a Service
 Provider or the Transport Authority to leak information to the external
-person.  It is thus combination of the Data Theft and Information Leak
+person.  It is thus a combination of the Data Theft and Information Leak
 scenario.
 The conflicts of interest are \textit{Privacy vs. Financial Profit} of
-the External, \textit{Reputation of Service Provider vs. Political
+the external, \textit{Reputation of Service Provider vs. Political
   Activism} of the external and \textit{Reputation of Transport Authority
@@
 vulnerabilities are, again, the existence of \textit{Unhappy
-  employees} and possibly \textit{lax access rules} that enable the
+  employees} and possibly \textit{careless access rules} that enable the
 employee to obtain large amounts of data unnoticed.
@@
 
-In this section we will associate to every identified threat a corresponding risk. A risk is the expected loss that is associated to the threat.
+In this section we will associate to every identified threat with a corresponding risk. A risk is the expected loss that is associated to the threat.
 Therefore, we have to quantify the likeliness of the threat to occure and the harm or loss done in this case.
 The quantification of likeliness will be solely based on rough judgment of the authors.
-The quantification of loss, will be made in a two step process.
+The quantification of loss will be made in a two step process.
 For each threat listed in the previous section, we have analyzed the affected personal data of the citizen.
-For each possible data type (e.g. GPS) we analyze the impact on the seven different types of privacy in Section \ref{sec:SensorPrivacyImpact}.
-In combination we can quantify roughly the impact of each threat on the citizens privacy. Both evaluations are necessarily fraught with a high level of uncertainty.
+For each possible data type (e.g. GPS) we analyze the impact on the seven different types of privacy in section \ref{sec:SensorPrivacyImpact}.
+In combination we can quantify roughly the impact of each threat on the citizens privacy.
 
@@
 For the quantification of likeliness the following scale between 4 (always) and 0 (impossible) is used.
-The quantification of the risk, we add the values for loss and likeliness of the corresponding threats.
+For the quantification of the risk, we add the values for loss and likeliness of the corresponding threats.
 Note, that loss and
@@
 of those scales corresponds to multiplication of the usual scales. The likeliness, loss and the resulting risks assigned to the threats
-are discussed in the following paragraphs and summarized in Figure
+are discussed in the following paragraphs and summarized in figure
 \ref{fig:risks}.
 
-\textbf{Risk of T1. Insufficient Control Features.}  The occurrence of this
-threat is dependent on the design on the system and given in our case,
+\textbf{Risk of T1. Insufficient Control Features.}  This threat depends on the design of the system. In our case it is always there,
 since we do not give the citizen control over his data once it is
@@
 We assess the likeliness of excessive data mining to be $3$ High, since
-these kind of analysis can be performed within the walls of the service
+this kind of analysis can be performed within the walls of the service
 provider, without somebody else noticing, and the service provider himself
@@
 \textbf{Risk of T3. Data Theft.}
-The likeliness of a targeted attack by a third party is dependent on
-the popularity of the offered service and financial value of the
+The likeliness of a targeted attack by a third party dependens on
+the popularity of the offered service and on the financial value of the
 captured information. Moreover, the amount of manual work required to
@@
 
-\textbf{Risk of T4. Surveillance.}  In the surveillance scenario an party
-related to the citizen, like a company where he is customer of, or a
+\textbf{Risk of T4. Surveillance.}  In the surveillance scenario a party
+related to the citizen, like a company that he is customer of, or a
 government agency, seeks to obtain sensitive information from our
@@
 very likely to occur. The barrier for companies that do not operate
-the infrastructure used to transmit the data a surveillance attack is
-however very hard to perform. Therefore we assess the likeliness of
+the infrastructure used to transmit the data a surveillance attack is very 
+hard to break. Therefore we assess the likeliness of
 the threat with $1-2$ (Low-Medium). The harm of leaked information to a related party is $3$ High. Hence the resulting risk is calculated as $4-5$.
@@
 In this section we derive recommendations or
-requirements for a system that addresses these risks. The mapping of these recommendations to the threats is summarized in Figure \ref{fig:risks}.
+requirements for a system that addresses these risks. The mapping of these recommendations to the threats is summarized in figure \ref{fig:risks}.
 
-\textbf{R1. Privacy Dashboard}. In order to address the threat with the highers risk, Insufficient Control (T1) of the citizen, we need to give the citizen back the
-control over its data inside the system. The most direct way to do
+\textbf{R1. Privacy Dashboard}. In order to address the threat with the highest risk, Insufficient Control (T1) of the citizen, we need to give the citizen back the
+control over his data inside the system. The most direct way to do
 this is to provide a web-based overview which allows the citizen to view, edit and delete all information about his person
@@
 Contrary to common belief, it is neither legal nor ethical to process
-personal data for by new methods or for new purposes that were not
+personal data by new methods or for new purposes that were not
 stated and explained to the citizen at the time of data
@@
 establish such a culture of privacy respect, we recommend to prepare
-an guideline intended for internal use that explains the concrete processing steps and purposes that are permitted by the citizens.
+an guideline intended for internal use that explains the concrete processing steps and purposes that have been permitted by the citizens.
 
@@
 User Permissions}.
-If further processing should be performed, it is necessary to seek additional permissions from the citizen.
+If further processing should be performed, it is necessary to seek additional permissions from the citizens.
 A simple email explaining the planned processing steps, and asking for permissions would be enough for this purpose.
@@
 links to the identity of the person are removed, no violation of the
-citizens privacy caused by arbitrary processing. However removing all
+citizen's privacy is caused by arbitrary processing. However, removing all
 such links is a challenging tasks, and full anonymity is often not
@@
 We found potential risks and derived recommendations how to protect the citizen's privacy.
-A central idea of these recommendations is to give the citizen as much control over their data back as possible.
+A central idea of these recommendations is to give the citizens as much control over their data back as possible.
 For example, this can be done by a privacy dashboard in the hand of every user that shows all his or her personal data stored on the system.

### 2015-02-21 paper.tex
Minor changes.

@@
 Citizens are persons who use the mobile device as users of the provided software.
-Their main motivation for using the software is to gain a higher level of convenience in their daily activities.
-For example they may have access to real time bus schedules or reports about traffic jams that help them to avoid long waiting times.
-Also they generally benefit from improvements of public infrastructure by transport authorities, which is triggered by issue reports.
-By using the application, citizens are sharing personal information like name and address, as well as data gathered from mobile sensors with the service provider.
-This data can be exploited in ways that are harmful to the citizen (cf.
-\cite{GuardienMassSurveillance}).
-Citizens are interested in: physical wellbeing and health, financial profit, convenience, legitimate use of personal data, non-disclosure of personal data to peers of the citizen, and not being monitored.
+Their main motivation gain convenience in their daily transport activities.
+Also they generally benefit from improvements of public infrastructure enabled by data collection.
+By using the application, citizens are sharing personal information with the service provider which put their privacy at risk.
+Other interests of the citizens include: physical wellbeing and health, financial profit, legitimate use of personal data, non-disclosure of personal data to peers of the citizen, and not being monitored.
 
@@
 \textbf{System Provider.}
-System Providers operate the technical infrastructure (hardware and software) of the IT System.
-They are private companies and legal persons in their own right, but also employ a number of people with diverging interests including administrators, programmer/developer, and a support manager.
+System providers operate the technical infrastructure (hardware and software) of the IT System.
+They are private companies and legal persons in their own right, but also employ a number of people with diverging interests including administrators, developers, and support managers.
 As companies, they are interested in gaining financial profit.
-The financial success of system providers depends on the task complexity of the maintained infrastructure.
-The complexity of a task has to be in reasonable bounds, so that system providers can complete it within time, with a satisfying quality.
-System Providers are interested in: financial profit, manageable complexity, professional excellence, and good working conditions, and a good reputation.
+% The financial success of system providers depends on the task complexity of the maintained infrastructure. 
+% The complexity of a task has to be in reasonable bounds, so that system providers can complete it within time, with a satisfying quality.
+System providers are interested in: financial profit, limited infrastructure complexity, professional excellence, good working conditions for their employees and good public reputation.
 
 \textbf{Transport Authority.}
- Transport authorities are public offices (ministry, agency, department, ...) or other external public entities which act as direct customers of service providers.
+Transport authorities are public offices (ministry, agency, department, \dots) or other external public entities which act as direct customers of service providers.
 For example a department for urban mobility orders a system to better understand usage patterns and make improvement to the urban traffic flow.
@@
 \textbf{System Complexity vs Privacy.}
-System Providers offer a service to Transport Authorities, so that they are able to improve their public services.
-This task in it self has a high technical complexity and operates on privacy sensitive data provided by monitored Citizens.
+System providers offer a service to transport authorities, so that they are able to improve their public services.
+This task in it self has a high technical complexity and operates on privacy sensitive data provided by monitored citizens.
 
 \textbf{Business Intelligence vs Privacy.}
-Transport authorities order a monitoring and mining system from system providers as much as possible, which allows them to produce business intelligence for public services.
-They are interested in the successful usage of their data, although the interest of citizens lies in maintaining control over their data and protecting their rights to privacy.
+Transport authorities are purchasing monitoring and mining services from system providers.
+They are naturally interested in gaining as much business intelligence from those services as possible.
+This interest is in conflict with the citizens' interest of maintaining control over his data and protecting his privacy.
 
@@
 \textbf{Financial Profit of External vs Reputation of System Providers.}
-Externals have various business models as optional foundation for attacks on System Providers.
+Externals have various business models as optional foundation for attacks on system providers.
 For instance, they can try to invade the infrastructure for e-espionage reasons, to get control over servers to create a bot network or to steal user data.
@@
 \textbf{Insecure Infrastructure.}
-The proposed monitoring system consists of many hardware and software components, each with its own concrete weaknesses.
-For instance, operating systems can be outdated or not subject to frequent updates or virus scans.
+The proposed monitoring system consists of many hardware and software components, each with its own concrete weaknesses. For instance, outdated operating systems are frequently vulnerable to attacks.
 
 \textbf{Insecure Data Transmission.}
-The proposed monitoring and mining system uses HTTP to exchange data between the Sensor Collector, the Data Center and the Report Tool.
-Data can be intercept and read all sensitive information, which is send between the components e.g. passwords, raw sensor data and data mining results.
+The proposed monitoring and mining system uses HTTP to exchange data between the sensor collector, the data center and the reporting tool.
+Data can be intercept and read all sensitive information, which is send between the components e.g. sensor data and data mining results.
 
 \textbf{Unhappy Employees.}
-As an insider, an Employee who is frustrated with his situation constitutes a security vulnerability. On the one hand he might
+As an insider, an employee who is frustrated with his situation constitutes a security vulnerability. On the one hand he might
 want to harm his employer directly, on the other he is increasingly
@@
 \textbf{Inadequate Access Rules.}
-The proposed IT system infrastructure has various accesses to privacy sensitive data.
-System Provider staff has access to Data Center hardware and software like databases, web-servers and other inspection tools.
-Transport authority staff has access to the Report Tool.
-Moreover, access control systems are security bottlenecks,
+The proposed IT system infrastructure has various accesses rules for privacy sensitive data.
+Administrators of the system provider need full access to data center hardware and software.
+Developers and sales managers only have limited access.
+Transport authority staff has access to the reporting tool but not to the raw data.
+Ill configured or ineffective access systems expose data to non-authorized personal and create a vulnerability.
 
 \textbf{Unaware Monitoring Subjects.}
-We define privacy as one's ability to control information about
-oneself.  In order to do that, monitored subjects need to know, that
+We define privacy as the ability to control information about oneself. 
+In order to do that, monitored subjects need to know, that
 they are monitored, who monitors them, what information is recorded
@@
 \textbf{T1. Insufficient Control Features.}
-As soon as collected data of Citizens is stored on System Provider servers, all control over that data is lost.
-Although, control capabilities for Citizens add to the system complexity, which could motivate to omit those.
-Therefore interests of Citizens and System Providers are in conflict, namely it is the Citizen's \textit{Privacy vs.
-System Complexity} for System Providers.
+As soon as collected data of citizens is stored on system provider servers all control over that data is lost,
+unless features are built into the system that enable citizens to control their data explicitly.
+These control features add to the system complexity and require extra effort by the system provider.
+Therefore interests of citizens and system providers are in conflict, namely it is the citizen's \textit{Privacy vs. System Complexity} for system providers.
 
 \textbf{T2. Excessive Data Mining.}
-The System Provider and/or the transport authority secretly extract more private information from the collected data, than the Citizen agreed to.
-But results of the mining process create no disadvantages for Citizens, because there is no disclosure to third parties.
-This could be the case for a System Provider, who wants to test a new product and uses the pre-existing data collection.
-Thus there are two possible conflicts: \textit{Privacy vs.
-Financial Profit of Serivce Provider} and  \textit{Privacy vs.
-Buisness Intelligence of Transport Authority}.
-The threat can be provoked by either careless data handling policies of both System Providers and Transport Authorities, or a weak law enforcement of existing supervision.
+The system provider and/or the transport authority secretly extract more private information from the collected data, than the citizen agreed to.
+This could be the case for a system provider, who wants to test a new product and uses the pre-existing data collection.
+Although this mining process do not need to create a direct harm for the citizen (e.g.
+when there is no disclosure to third parties) the control over his data is lost and thus his privacy is violated.
+There are two conflicts of interest involved: \textit{Privacy vs. Financial Profit of Serivce Provider} and  \textit{Privacy vs. Buisness Intelligence of Transport Authority}.
+The threat can be provoked by either careless data handling policies of both system providers and transport authorities, or a weak rule enforcement of existing supervision.
 But the main issue, which can lead to such threats, is again a general \textit{Missing Privacy Awareness}.
@@
 An malicious external infiltrates the infrastructure in order to steal personal data and sell it on the black market.
-Also the External might be motivated politically and wants to harm the reputation of the System Provider or the Transport Authority.
-Such a successful attack could harm the reputation of both System Provider and Transport Authority.
+Also the External might be motivated politically and wants to harm the reputation of the system provider or the transport authority.
+Such a successful attack could harm the reputation of both system provider and transport authority.
 This threat is defined by three conflicts: \textit{Privacy vs.
@@
 Political Activism}.
-Also this threat describes the classical scenario, where attacks are provoked by \textit{Insecure Infrastructure} (SQL injection) and \textit{Insecure Communication} (Packet Capture).
+Also this threat describes the classical scenario, where attacks are provoked by \textit{Insecure Infrastructure} (SQL injection) and \textit{Insecure Communication}.
 
@@
 \textbf{T5. Information Leak} 
-Like a malicious external person in the Data Theft scenario an employee of the
+Like a malicious external person in the data theft scenario an employee of the
 service provider or the transport authority has selfish interests to gain
@@
 it to the public.  The corresponding conflicts of interests are:
-\textit{Privacy vs. Financial Profit} of the Employee, \textit{Reputation of Service
-  Provider vs. Political Activism} of the Employee and \textit{Reputation of Transport
-  Authority vs. Political Activism} of the Employee.  The vulnerability constitutes of
+\textit{Privacy vs. Financial Profit} of the employee, \textit{Reputation of Service
+  Provider vs. Political Activism} of the employee and \textit{Reputation of Transport
+  Authority vs. Political Activism} of the employee.  The vulnerability constitutes of
 the existence of \textit{Unhappy employees} and possibly \textit{careless access
@@
 \textbf{T6. Social Engineering}
-In this scenario an external manipulates an employee of a Service
-Provider or the Transport Authority to leak information to the external
-person.  It is thus a combination of the Data Theft and Information Leak
+In this scenario an external manipulates an employee of a service
+provider or the transport authority to leak information to the external
+person.  It is thus a combination of the data theft and Information Leak
 scenario.
@@
 In this section we will associate to every identified threat with a corresponding risk. A risk is the expected loss that is associated to the threat.
-Therefore, we have to quantify the likeliness of the threat to occure and the harm or loss done in this case.
+Therefore, we have to quantify the likeliness of the threat to occur and the harm or loss done in this case.
 The quantification of likeliness will be solely based on rough judgment of the authors.
@@
 We found potential risks and derived recommendations how to protect the citizen's privacy.
-A central idea of these recommendations is to give the citizens as much control over their data back as possible.
+A central idea of these recommendations is to give the citizens back the control over their data.
 For example, this can be done by a privacy dashboard in the hand of every user that shows all his or her personal data stored on the system.

### 2015-06-11 paper.tex
rescaled vul. graphic
'save harbor' -> 'safe harbor' 'Pricacy' -> 'Privacy'

@@
 Indeed, this principle of “data minimization” is still fundamental to modern data protection legislation.
-The modern principles of data protection are codified by many legal systems in different countries, especially in Europe \cite{EU-directive2}, and by the US save harbor principles.
+The modern principles of data protection are codified by many legal systems in different countries, especially in Europe \cite{EU-directive2}, and by the US safe harbor principles.
 However, modern data protection is more than only the absence of personal data.
@@
 \hline
-& \thead{Pricacy of the Person}
-& \thead{Pricacy of Behaviour and Action}
-& \thead{Pricacy of Communication}
-& \thead{Pricacy of Data and Image}
-& \thead{Pricacy of Thoughts and Feelings}
-& \thead{Pricacy of Location and Space}
-& \thead{Pricacy of Association}
+& \thead{Privacy of the Person}
+& \thead{Privacy of Behaviour and Action}
+& \thead{Privacy of Communication}
+& \thead{Privacy of Data and Image}
+& \thead{Privacy of Thoughts and Feelings}
+& \thead{Privacy of Location and Space}
+& \thead{Privacy of Association}
 \\\hline\hline
 \thead{GPS Sensor}
-& \tbody{0.5} % Pricacy of the Person
-& \tbody{0.5} % Pricacy of Behaviour and Action
-& \tbody{0} % Pricacy of Communication
-& \tbody{1} % Pricacy of Data and Image
-& \tbody{0} % Pricacy of Thoughts and Feelings
-& \tbody{1} % Pricacy of Location and Space
-& \tbody{0.5} % Pricacy of Association
+& \tbody{0.5} % Privacy of the Person
+& \tbody{0.5} % Privacy of Behaviour and Action
+& \tbody{0} % Privacy of Communication
+& \tbody{1} % Privacy of Data and Image
+& \tbody{0} % Privacy of Thoughts and Feelings
+& \tbody{1} % Privacy of Location and Space
+& \tbody{0.5} % Privacy of Association
 \\\hline
 \thead{Motion Sensors}
-& \tbody{1} % Pricacy of the Person
-& \tbody{0.5} % Pricacy of Behaviour and Action
-& \tbody{0} % Pricacy of Communication
-& \tbody{1} % Pricacy of Data and Image
-& \tbody{0} % Pricacy of Thoughts and Feelings
-& \tbody{0} % Pricacy of Location and Space
-& \tbody{0} % Pricacy of Association
+& \tbody{1} % Privacy of the Person
+& \tbody{0.5} % Privacy of Behaviour and Action
+& \tbody{0} % Privacy of Communication
+& \tbody{1} % Privacy of Data and Image
+& \tbody{0} % Privacy of Thoughts and Feelings
+& \tbody{0} % Privacy of Location and Space
+& \tbody{0} % Privacy of Association
 \\\hline
@@
 \centering
-\includegraphics[width=0.6\textwidth]{diagrams/png/vulnerabilities.png}
+\includegraphics[width=\textwidth]{diagrams/png/vulnerabilities.png}
 

### 2015-06-11 paper.tex
- minor changins

@@
 & \tbody{1} % Privacy of Data and Image
-& \tbody{0} % Privacy of Thoughts and Feelings
+& \tbody{0.5} % Privacy of Thoughts and Feelings
 & \tbody{1} % Privacy of Location and Space
@@
 & \tbody{1} % Pricacy of Data and Image
-& \tbody{0} % Pricacy of Thoughts and Feelings
+& \tbody{0.5} % Pricacy of Thoughts and Feelings
 & \tbody{1} % Pricacy of Location and Space
@@
 The emphasis here is to put on prominent existing conflicts, because they provide a foundation for vulnerabilities and subsequent threats.
+A profit-driven conflict between the Service Provider and the privacy of citizens has been omitted for the sake of clarity.
 
@@
 
-\textbf{Risk of T5/6. Information Leak and Social Engineering.} In our scenario we assume that the culture and ethics inside the
+\textbf{Risk of T5/6. Information Leak and Social Engineering.} In our scenario we assume that the culture and ethics of the employees inside the
 service provider company and travel agencies are very high, so that
